### 2018-04-30 VFTS682.tex
started a tex file

\documentclass{aa}
\usepackage{graphicx}
\usepackage[varg]{txfonts}
\usepackage[%draft, 
colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
\usepackage{amsmath}
\usepackage[usenames]{xcolor}
\usepackage{comment}
\usepackage{multirow}
\usepackage{chngpage}
\usepackage{lscape}
\usepackage{url}
\newcommand{\todo}[1]{{\large $\blacksquare$~\textbf{\color{red}[#1]}}~$\blacksquare$}
\newcommand{\udef}{\stackrel{\mathrm{def}}{=}}
% for tree diagram
\usepackage{forest}
\usepackage{tikz-qtree}
\usetikzlibrary{shadows,trees}

%Selma's comments
\definecolor{Wildstrawberry}{rgb}{1.0, 0.26, 0.64}
\newcommand{\SdM}[1]{{\color{Wildstrawberry}\bf{#1}}}
\newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}

\renewcommand{\labelitemii}{$\bullet$}
\newcommand{\kms}{{\mathrm{km\ s^{-1}}}}
\newcommand{\Msun}{{\mathrm{M}_\odot}}

\DeclareRobustCommand{\Eqref}[1]{Eq.~\ref{#1}}
\DeclareRobustCommand{\Figref}[1]{Fig.~\ref{#1}}
\DeclareRobustCommand{\Tabref}[1]{Table~\ref{#1}}
\DeclareRobustCommand{\Secref}[1]{Sec.~\ref{#1}}

\defcitealias{tauris:98}{TT98}
\defcitealias{ramirez-agudelo:15}{R-A15}

\interfootnotelinepenalty=10000    % brute-forces the footnote not to break over two pages

\begin{document}

\title{Massive Runaway and Walkaway Stars}
\subtitle{A study of the kinematical imprints of the physical processes \\ governing the evolution and explosion of their binary progenitors}

\author{M.~Renzo\inst{1} \and E.~Zapartas\inst{1} \and S.~E.~de~Mink\inst{1} \and
 Y.~G\"otberg\inst{1} \and S.~Justham\inst{2,3} \and
 R.~J.~Farmer\inst{1} \and R.~G.~Izzard\inst{4}
 \and S.~Toonen\inst{1}
 \and H.~Sana\inst{5}.}

\institute{{Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands}
  \and{School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}
  \and{National Astronomical Observatories, Chinese Academy of
    Sciences, Beijing 100012, China}
  \and{Astrophysics Research Group, Faculty of Engineering and Physical Sciences, University of Surrey, Guildford, Surrey, GU2 7XH, United Kingdom}
  \and{Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}}

\offprints{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\abstract{Anticipating the kinematic constraints from
  the Gaia mission, we perform an extensive numerical study of
  the evolution of massive binary systems to predict the peculiar
  velocities that stars obtain when their companion collapses and disrupts the system. Our aim is to (1) identify which predictions are robust against model uncertainties and assess their implications, (2) investigate which physical processes leave a clear imprint and may therefore be constrained observationally and (3) provide a suite of publicly available model predictions.  
We find that $22_{-8}^{+26}$\% of all massive binary systems merge
prior to the first core collapse in the system. Of the remainder,
$86_{-9}^{+11}\%$ become unbound because of the
core-collapse. Remarkably, this rarely produce runaway stars (i.e.,
stars with velocities above $30\,\kms$). These are outnumbered by more
than an order of magnitude by slower unbound companions, or ``walkaway
stars’’. This is a robust outcome of our simulations and is due to the
reversal of the mass ratio prior to the explosion and widening of the
orbit, as we show analytically and numerically. We estimate a
$10^{+5}_{-8}$\% of massive stars to be walkaways and only $0.5^{+1.0}_{-0.4}$\% to be runaways, nearly all of which have
accreted mass from their companion. Our findings are consistent with
earlier studies, however the low runaway fraction we find is in
tension with observed fractions $\sim$10\%. If Gaia confirms these high fractions of
massive runaway stars resulting from binaries, it would imply that we
are currently missing physics in the binary models. Finally, we show
that high end of the mass distributions of runaway stars is very
sensitive to the assumed black hole natal kicks and propose this as a
potentially stringent test for the explosion mechanism. We also
discuss companions remaining bound which can evolve into X-ray
 and gravitational wave sources.}

\keywords{stars: kinematics, binaries, massive, supernovae }
\maketitle{}

\section{Introduction}
\label{sec:intro}

Stars with initial mass larger than
about $7.5\,M_\odot$ are the progenitors of black holes (BH) and
neutron stars (NS). These stars play an important role in shaping galaxies
through their radiative, chemical, mechanical feedback
(e.g., \citealt{larson:74, Heger+2003, ceverino:09, lamers:13}, see
also \citealt{langer:12, nomoto:13} for reviews).
Most young, unevolved, massive stars have a nearby companion with
which they form a close binary system
\citep[e.g.,][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. 
Binary systems that remain bound throughout the entire evolution of both
stars can give rise to many exotic phenomena, including X-ray binaries
\citep[e.g., ][]{gott:71,Bolton1972, Webster+1972,
  van-den-Heuvel+1972}, binary neutron stars
\citep[e.g.,][]{hulse:75,wijers:92}, gamma-ray bursts
\citep[e.g.,][]{izzard:04b,becerra:16, LVC:17:grb}, and gravitational wave events
\citep{LVC:16b, BNSmerger:17}.  
However, only a small fraction of massive stars that are born as
binary systems are expected to stay together their entire lives. The
majority of systems is disrupted by the first core collapse event,
which can separate the newly formed compact object from its former
companion star
\citep[][]{zwicky:57,blaauw:61, tauris:98, belczynski:99,
  belczynski:08, eldridge:11}.

Single NSs can be detected as pulsars or magnetars. Many of
them are observed to have large proper motions (e.g.,
\citealt{gunn:70,gott:70,lyne:94,kaspi:96,hobbs:05,chatterjee:05}, but see also \citealt{verbunt:17}).
Conversely, single BHs are only detectable under special
circumstances, for example through lensing events when passing in
front of a background star \citep[e.g.,][]{wyrzykowski:16}, or if they
accrete gas from the ambient medium
\citep[e.g.,][]{fender:13, gaggero:17}. Therefore, to probe the
population of stellar-mass black holes we are limited to X-ray or
gravitational wave observations that can only target the rare
cases that remain bound to their companion and are close enough for
the BH to accrete or merge. To learn more about the black holes that
form in less special cases, it is worth to investigate the imprints
they may leave on their former companion star.

The primary focus of this study is the population of unbound main sequence companions,
which can be identified observationally because of their peculiar
spatial velocities compared to the surrounding population. Their velocities can either be detected as proper
motions (i.e.,\ their motion in the plane of the sky measured directly
from the displacement of the star) or as radial velocities (i.e.,\ their
motion perpendicular to the plane of the sky, measured from the
Doppler shift of the spectral lines). Large spatial velocities have
been inferred for a significant sub-population of young massive stars
\citep[e.g.][]{blaauw:61, Cruz-Gonzalez+1974, gies:86, gies:87,
  hoogerwerf:00, hoogerwerf:01, tetzlaff:11, boubert:18}.

\citet[][]{blaauw:61} introduced the term ``runaway stars'' for those %stars
in the fast tail of the velocity distribution for a given spectral type. The typical threshold
adopted to define the tail of this distribution for O and B-type stars is $v \gtrsim 30\,\mathrm{km \ s^{-1}}$
\citep[e.g.,][]{blaauw:56, gies:86, dedonder:97, hoogerwerf:00, hoogerwerf:01, dray:05, eldridge:11}, although sometimes other values
have been considered \citep[e.g., $40\,\mathrm{km\ s^{-1}}$
in][]{blaauw:61, dewit:05, boubert:18}. However, as we will argue based on simulations presented in this work, the majority of unbound companions is expected to exhibit velocities well below these thresholds.  
We will refer to these slow unbound former companions resulting from
disrupted binary systems as ``walkaway stars'' \citep[to our knowledge first coined
by][]{de-Mink+2012} to distinguish them from the faster counterparts.

A proposed explanation for the large spatial velocity of runaways is that
they originate from disrupted binary systems \citep[]{zwicky:57,
  blaauw:61, boersma:61},
which naturally explains the lower number of companions they have,
compared to typical massive stars \citep[e.g.,][]{blaauw:61, gies:86, sana:14}. \Figref{fig:cartoon} sketches the
typical evolution of such a binary system. Most binaries are disrupted
at the time of the first core-collapse.

An alternative mechanism to produce stars with peculiar spatial
velocities is dynamical ejection from a star cluster \citep[e.g.,][]{poveda:67, leonard:91, Evans+2010a,
  fujii:11, allison:12, oh:16}. Dynamical interactions with
a supermassive BH can also disrupt a binary, but the ejection
velocities achieved in this scenario are typically much higher
\citep[$\gtrsim$$10^3\,\mathrm{km\ s^{-1}}$,][]{hills:88}.

Both mechanisms, i.e. the ``binary
disruption scenario'' and the ``dynamical ejection scenario'', are
expected to act in nature, but their relative importance is not not well
constrained \citep[e.g.,][]{hoogerwerf:00,hoogerwerf:01,guseinov:05}.
\citet{hoogerwerf:01} analyzed the properties of a sample of 56 nearby
runaway stars and 9 radio pulsars, and traced back the runaways to
their most likely parent stellar group. From the sub-sample for which a
clear identification of the parent group was possible, they estimated that the disruption of binaries is
responsible for roughly two thirds of observed runaways \citep[see
also][]{gies:86, gies:87, stone:91, hoogerwerf:00, dincel:15,
  boubert:17b}. However, this claim could not be confirmed by
\cite{jilinski:10}.

Unbound stars resulting from the disruption of a binary are of
potential interest for several topical questions in
astrophysics. Their kinematics and stellar properties bear imprints of the uncertain physical processes that govern the evolution of their
binary progenitor systems, including the expected phase of mass
transfer between the two stars (phase B. in
\Figref{fig:cartoon}). Detailed knowledge of the spatial velocity of
runaway stars can improve the accuracy of wind mass-loss rate
determinations relying on their bow shocks \citep[e.g.][]{gull:79,
  kobulnicky:18}. Of particular interest is the
question whether they can provide any unique constraints on the physics of
core-collapse, in particular the natal kick on the compact objects
that they produce. Such kicks are expected either from asymmetries in
the explosion and/or the neutrino emission
\citep[][]{shklovskii:70,kusenko:96,fryer:04,socrates:05,wongwathanarat:13,
  janka:13,janka:17}, and they determine which systems disrupt and
eject the companion star and which systems remain bound and thus have a
potential as future X-ray and gravitational wave sources.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.5\textwidth]{./figures/cartoon}
  \caption{Schematic depiction of the evolution of a close massive
    binary through stable mass transfer. The
    evolution forks at the first core-collapse event in the system: the vast
  majority of the systems is disrupted and produces a runaway or
  walkaway star which can travel far from its birth location. The
  systems remaining bound are possible progenitors of X-ray binaries
  and gravitational wave sources. The fraction of binaries disrupted
  $\mathcal{D}$ comes from the simulations presented in this study.
}
  \label{fig:cartoon}
\end{figure}

These unbound stars have also been considered for their potential
importance as non-canonical sources of stellar feedback
\citep[e.g.,][]{larson:74, ceverino:09}. 
They can travel long distances and end their lives tens to hundreds of
parsecs away from their birth location. Because of their motion, their
ionizing photons are
less likely to be absorbed by their birth clouds. Therefore, in the context of the
re-ionization of the early universe, the ionizing radiation of stars
ejected from a binary are more likely to
escape and become available for ionization of hydrogen in the
intergalactic medium \citep[e.g.][]{conroy:12}. This has been
suggested to be of importance in the 
context of the epoch of reionization
\citep[e.g.,][]{conroy:12,kimm:14,Ma+2016}. Furthermore they are
expected to explode in lower density regions, which may enhance the
impact they have as sources of turbulence in the ISM \citep[e.g.][]{gatto:15}.

The present and upcoming second Gaia data releases provide an important
observational motivation for this study \citep[e.g.,][]{perryman:01}. The determination of
distances, proper motions and radial velocities by the Gaia satellite
is expected to drastically increase the available sample of massive
stars with precisely known velocities
\citep{maoileidigh:05,  cacciari:09, Jeffries+2014,
  Gaia-Collaboration+2016, Gaia-Collaboration+2016a}.

In this study, we present a systematic theoretical study of the
kinematical signatures of stars ejected from binary systems and how
these depend on the uncertain physical processes governing binary
stellar evolution. For this purpose we employ a rapid binary
population synthesis code, which we have updated to account for the main relevant processes that affect the disruption of the binary systems (see \Secref{sec:methods}).

We first provide insight in our simulations by presenting the results
for an individual system in \Secref{sec:example}.
We present in \Secref{sec:analytics} an analytic estimate of the
typical and maximum velocity of ejected companions expected from the
disruption of binaries. Then, in \Secref{sec:fiducial_res} we describe the results of a fiducial simulation of a
full population of binary stars, for which we adopt realistic
input assumptions provided by detailed simulations and/or observations
when available.  In addition, we present in
\Secref{sec:param_variation} an extensive grid of models where we vary the most relevant uncertain assumptions. This allows us to study (i) which predictions are robust against the model uncertainties and (ii) which uncertain physical processes leave a clear imprint on the observables that can be used to test and constrain them.

Various earlier studies have discussed the evolution and
interaction of populations of massive binary systems. Examples
include, but are not limited to, \citet[][]{vanbeveren:82, dedonder:97, Fryer+1997, Fryer+2001, fryer:12,
  Belczynski+2012b, repetto:12, Fragos+2013a, Grudzinska+2015,
  bray:18, boubert:18}. Most
of these studies focus on the minor fraction of systems
that remain bound and are the progenitors of X-ray binaries and/or
gravitational wave sources. A few studies focus specifically on
disrupted systems and unbound stars. Examples are
\cite{vanrensbergen:96,dedonder:97, dray:05, eldridge:11, bray:16, boubert:17b,
  zapartas:17b, boubert:18}. We expand on these studies by using updated physical
assumptions, and by focusing on which physical processes could be
constrained using the Galactic population of unbound stars. A
comparison to previously published results is given in \Secref{sec:literature}.

Our first main result is that the majority of disrupted binaries
ejects a slow-moving walkaway star ($v < 30\,\mathrm{km\ s^{-1}}$), rather than a faster runaway. This result is robust against
variations of uncertain parameters in the model, as we discuss in \Secref{sec:param_variation}.
This would imply that the more easily detected runaway stars only
reveal a small subset of the population of unbound stars that are
former companions of disrupted binary systems.  We discuss the implications in \Secref{sec:discussion}.

Our second main result is that the kinematic properties and absolute
number of unbound companions depend sensitively on assumptions
concerning the  natal kicks of BHs. We show the imprints this leaves on
the mass distribution and discuss whether future observations can be
used to constrain these processes.

Section~\ref{sec:bound} briefly describes the population of systems remaining bound after the first
core-collapse.

We also find that the fraction of runaways among massive stars
predicted by our simulations is much lower than the observed value in
all our parameter variations. This finding agrees with the results of
\cite{eldridge:11}, but is in potential contrast
with the observational results of \cite{hoogerwerf:01}, and needs further
investigation.

Finally, we discuss how our results  may be useful to
include the effect of unbound binaries in models for stellar
feedback. We will provide our numerical results upon publication. %provide our numerical results at \todo{where?}\todo{zenodo/CDS?}.

\section{Binary population synthesis calculations}
\label{sec:methods}

We carry out population synthesis calculations of isolated binaries with the rapid binary evolution code \texttt{binary\_c}
\citep[][]{izzard:04,izzard:06,izzard:09,demink:13,schneider:15,izzard:18}. This
code is based on the algorithms by \cite{tout:97,hurley:00,hurley:02}, which rely on the analytic fits to the
single stellar evolution models from \cite{pols:98}.

First, we compute a fiducial population using observationally
validated assumptions for the free parameters that describe the initial
conditions and physical assumptions.
Then, we check the robustness of our results (or
equivalently, to which assumptions they are most sensitive) by varying
the free parameters that are most relevant for the
velocity distribution of disrupted binaries. To limit
computational costs, we explore variations in each parameter
one-by-one while keeping the other parameters fixed to our
fiducial choices, following the approach of
\cite{Fragos+2013a,demink:13,demink:14,zapartas:17,zapartas:17b,
  belczynski:17}. Effectively, we treat
each parameter in our models as independent from the others, therefore this approach does not account for possible
correlations between either initial distributions \citep[e.g., the one between initial period and mass ratio
suggested by][]{moe:17} or uncertain
physical processes (e.g., core spin and natal kick amplitude). Varying
two (or more) parameters simultaneously, or changing the algorithmic
representation of the uncertain physical processes, might possibly result in
larger variations than those we present in
\Secref{sec:param_variation}.

\subsection{Initial distributions and parameters}

Each binary system in our calculations is characterized by a zero age
main sequence (ZAMS)
mass for the primary\footnote{Throughout this study, we define the primary star
  to be the initially more massive star, even
  if it becomes the less massive star in the binary during the evolution.}
$M_1^\mathrm{ZAMS}$, initial mass ratio $q\udef M_2/M_1$, and initial orbital
period $P^\mathrm{ZAMS}$. We build a grid in this parameter space and
weigh each system in our grid according to the initial distributions
described below.

We select $N_{M_1}$ primary stars with ZAMS mass $M_1^\mathrm{ZAMS}$
at logarithmically spaced intervals in the range $7.5\,M_\odot\leq M_1^\mathrm{ZAMS}
\leq 100\,M_\odot$. We weigh each primary star with an initial mass function (IMF) with
slope $\alpha=-2.3$ 
\citep[][]{kroupa:01}. In our model
variations, we explore values of $\alpha$ of -1.9
\citep[][]{schneider:18} and -3.

For each primary star of mass $M_1^\mathrm{ZAMS}$, we select $N_q$ secondaries with mass
$M_2^\mathrm{ZAMS} = qM_1^\mathrm{ZAMS}$, taken at regular intervals
in mass ratio $q$ between 0.1 and 1 assuming a
flat distribution \citep[slope $\kappa=0$, e.g.,][]{kouwenhoven:05,sana:12}. We consider in
\Secref{sec:param_variation} also variations with $\kappa=\pm1$.

Finally, for each pair of masses$(M_1^\mathrm{ZAMS},M_2^\mathrm{ZAMS})$ we choose $N_p$ different periods
$P^\mathrm{ZAMS}$ equally spaced in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$
between 0.15 and 5.5. We weigh the birth probability of each binary system with a
mass-dependent distribution for the initial orbital period: if $M_1<15\,M_\odot$ we assume
a flat distribution in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$
\citep[][]{opik:24,kobulnicky:07}; while for $M_1\geq15\,M_\odot$ we assume a
powerlaw distribution in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$ with exponent
$\pi=-0.55$, \citep[][]{sana:12}. We explore values of $\pi=0\,$and\,$\pi=-1$
(for all values of $M_1^\mathrm{ZAMS}$)
in our model variations. We chose our upper-limit on the initial period, $10^{5.5}$\,days, to include
wide systems in which the stars effectively evolve as single stars \citep[][]{demink:15}.

To limit the dimensions of the parameter space, we assume
all orbits to be initially circular (i.e., eccentricity
$e=0$). Because we focus on the population of stars resulting from the
disruption of post-interaction binaries, this assumption is not
critical \citep[][]{demink:15}: orbits are expected to
circularize
because of tides before or during mass transfer
(\citealt{belczynski:99,hurley:02}, but see also \citealt{eldridge:09}
for arguments against circularization of the orbit in post-main
sequence mass transfer).

In our fiducial model, we assume a canonical metallicity (i.e., mass fraction of
elements heavier than helium) $Z=0.02$. This value is slightly above
the most recent determination for solar neighborhood
\cite{asplund:09}, but it should generally describe young population
of massive stars in the Milky-Way. The models from \cite{pols:98}, which serve as
input for our computations, adopt the
isotopic mixture of \cite{anders:89}. In
\Secref{sec:param_variation}, we also consider $Z=0.0002$, $0.0047$, $0.008$,
and $0.03$.

We initialize the stellar rotation rate with a mass-dependent
equatorial velocity according to \cite{hurley:00}, assuming
alignment between the stellar spins and the orbital angular momentum. In one of our
parameter variations, we draw the initial spin velocity randomly from
the distribution given by \citealt{ramirez-agudelo:15} (R15, see last
line in \Tabref{tab:parameters}), but maintain the assumption of
alignment. However, changing the initial rotation rate has almost no
impact on our results because binary interaction processes overwrite it \citep[][]{demink:13}.

We build a grid in the initial parameter space, and initialize a
binary system in each cell of such grid. 
Each system is assigned a ``probability''
which can then be multiplied by the
amount of time spent by the system in that cell to obtain a quantity comparable to
observed number counts. See Appendix~\ref{app:data} for more details.

For each set of parameters, the total size of our model grid is
$N_{M_1}\times N_q \times N_p = 50 \times 50 \times 100 = 250\,000$ 
binary systems\footnote{This number is then multiplied by the number
  of natal kicks that we draw for each core-collapse, see \Secref{sec:SN_kick}.}.  Increasing this resolution to
$432\,000$ introduced no
significant variation in our results.

\subsection{Physical assumptions}

We follow the evolution of each binary system until the
first core-collapse (CC) event or until they merge. 
Stellar winds can have a dramatic effect on the evolution of single
stars \citep[e.g.,][]{renzo:17} and impact significantly the orbital
evolution in a binary. We include wind mass loss in our single star models as implemented by
\cite{demink:13}, i.e.\ we use a combination of mass loss rates from
\cite{vink:00,vink:01} for hot hydrogen-rich stars,
\cite{nieuwenhuijzen:90} for cool stars, % (or for luminosities larger than 4\,000$L_\odot$ if the mass loss rate
% is larger than what suggested by \cite{vink:00,vink:01})
 including a ``luminous blue variable'' enhancement as in
 \citealt{hurley:00}, and from \cite{hamann:98} reduced by a factor of 10 (for
 Wolf-Rayet stars. We also include the wind mass loss
 enhancement for fast rotating stars following \cite{maeder:00}.

We include the effects
of tides on the spins and orbital angular momentum using the
algorithm from \cite{hurley:02} based on the calculations of \cite{zahn:77,hut:81}.

When the radius of the most massive star exceeds the Roche radius,
(calculated using \citealt{eggleton:83} fitting formula)
we use the algorithm of \cite{claeys:14} to determine the
mass transfer rate (see their Eq.~10).

We limit the accretion rate of the
accretor to 10 times its Kelvin-Helmholtz timescale, i.e., we assume a
mass transfer efficiency $\beta_\mathrm{RLOF}\equiv |
\dot{M}_\mathrm{acc}|/|\dot{M}_\mathrm{don}|\equiv
\beta_\mathrm{th}$, with $\dot{M}_\mathrm{don}$ and
$\dot{M}_\mathrm{acc}$ temporal variation of the donor and accretor mass, respectively. Larger rates are
likely to drive the accretor out of thermal equilibrium\footnote{We do
not model the internal structure of the stars, therefore, we do not
follow its possible bloating during mass transfer, which could
potentially enhance its mass loss and related spin down.} and lead to
unstable mass transfer \citep[][]{neo:77,hurley:02}. However, for a
large portion of the parameter space, $\beta_\mathrm{th}$ results in a
rather conservative mass transfer \citep[][]{schneider:15}. Since the
efficiency of mass transfer is a major uncertainty in binary evolution
\citep[][]{demink:07}, we also consider the extreme parameter
variations with $\beta=0,\,0.5,$ and 1, respectively.

Matter that is not accreted is assumed to leave the system with
the specific angular momentum of the orbit of the accretor
\citep[e.g.,][]{soberman:97,vandenheuvel:17}, $h = \gamma_\mathrm{RLOF}
J_\mathrm{orb}/(M_\mathrm{don}+M_\mathrm{acc})$,
with $\gamma_\mathrm{RLOF}=M_\mathrm{don}/M_\mathrm{acc}$. In our
model variations, we also explore a scenario where the mass that is not
accreted leaves the binary from the outer Lagrangian point L2 and is assumed to form a circumbinary disk \citep[i.e.,
$\gamma_\mathrm{RLOF}=\sqrt{2}(M_\mathrm{don}+M_\mathrm{acc})^2/(M_\mathrm{acc}M_\mathrm{don})\equiv
\gamma_\mathrm{disk}$,][]{artymowicz:94}. We further consider the
assumptions that the mass that is not accreted leaves the system with the
orbital specific angular momentum, i.e.~$\gamma=1$, which is the standard model of \cite{belczynski:08, dominik:12, dominik:13}.

%qcrits
If, by the time the first star fills its Roche lobe, the mass
ratio of the accretor to the donor is smaller than a certain threshold
($M_\mathrm{acc}/M_\mathrm{don}<q_\mathrm{crit}$), we assume that the
system enters a common envelope phase. The threshold value
$q_\mathrm{crit}$ is uncertain and
depends on the evolutionary stage of the donor. In our fiducial simulation $q_\mathrm{crit,A}=0.65$ for a
main sequence (MS) donor \citep[][]{demink:07}, and $q_\mathrm{crit,B}=0.4$ for a Hertzsprung gap
donor \citep[][]{hurley:02}.
We also consider values of $q_\mathrm{crit,A}=0.25,\,0.8$ and $q_\mathrm{crit,B}=1,\,0.5,\,0$
in our model variations.

We treat common envelope evolution using the $\alpha_{\rm CE}
\lambda$-formalism \citep[][]{webbink:84, livio:88, dekool:90, hurley:02}. In all
variations, we use an analytic fit to the $\lambda_{g}$ values of
\cite{dewi:00} for the binding energy parameter $\lambda$. Those values do not include the energy stored in thermal
motions and the ionization state of the material within the
envelope. It is unknown what fraction of that internal energy is
useful in unbinding the envelope \citep[see, e.g.,][and references
therein]{dewi:00, ivanova:13}.  We assume $\alpha_{\rm CE} = 1$ for
our fiducial simulation, i.e., perfectly efficient use of the
liberated gravitational potential energy from the orbit but without
additional energy sources.  We explore variations with  $\alpha_{\rm
  CE} = 0.1$, i.e., inefficient use of the orbital energy to eject the
common envelope, and $\alpha_{\rm CE} = 10$, intended as indicative of
a fairly extreme case of additional energy input. For
binaries that interact while neither stars have a well defined core-envelope structure, such as
MS stars and helium-MS stars, we assume that the common envelope
phase leads to a merger as described in \cite{demink:13, schneider:15}.

\subsubsection{Natal kick}
\label{sec:SN_kick}

When a CC event happens in a binary system three different physical
ingredients contribute to possibly disrupting the system and ejecting the
companion star.

\begin{enumerate}
\item[i.] The orbit is modified by the sudden change in gravitational potential, because
of the mass lost through ejecta \citep[the so-called ``Blaauw
kick'',][]{zwicky:57,blaauw:61, boersma:61}. However, this is not the
dominant effect, because the typical evolution of a massive close binary
system involves mass transfer through Roche lobe overflow (RLOF) before the first CC (cf.~\Figref{fig:cartoon}). RLOF
removes the envelope of the initially more massive star, limiting
severely the amount of mass that can be ejected by the SN. The
``Blaauw kick'' alone rarely unbinds the system \citep[][]{huang:63,tutukov:73,leonard:94}.
\item[ii.] The SN shock can interact with the companion star.
  \citep[e.g.,][]{wheeler:75,liu:15, rimoldi:16, hirai:18}. The shock can dynamically
  remove mass from the companion (stripping), heat the envelope of the
  secondary, thus enhancing its own mass
  loss (ablation), and deposit mass and momentum on the secondary. The injection of energy can inflate the companion and make
  it look redder for a duration comparable to its thermal timescale. Our
  treatment of accretion only changes the mass of the companion, without
  checking for structural readjustments of the star. However, the
  interaction between the ejecta and the secondary is typically a small effect \citep[e.g.,][]{liu:15}.
\item[iii.]  The natal kick of the compact object changes %instantly
  its kinetic energy and momentum. This kick is caused by asymmetries
  in the SN ejecta and/or neutrino flux at the explosion, possibly
  seeded by the late core and shell burning phases of stellar
  evolution \citep[see e.g.,][]{shklovskii:70,kusenko:96,fryer:04,socrates:05,wongwathanarat:13,
    janka:13,janka:17,holland-ashford:17,grefenstette:17,katsuda:18}. It
is typically parametrized using a kick velocity $\mathbf{v}_k$, drawn
from a distribution of amplitude $v_k\equiv |\mathbf{v}_k|$ and directions (see below). The natal
kick is the dominant reason for the disruption of binaries.
\end{enumerate}

We have updated the treatment of binary disruptions by a CC event in
\texttt{binary\_c}, following \citealt{tauris:98} (TT98 hereafter). To re-calculate the post-CC
orbital parameters, their algorithm
assumes instantaneous
loss of the ejecta, because the ejecta velocity $v_\mathrm{ej}$ is much larger than the
orbital velocity $v_\mathrm{orb}$ ($v_\mathrm{orb} \lesssim 10^2\,\mathrm{km \
  s^{-1}} \ll v_\mathrm{ej}\sim10^4\,\mathrm{km \
  s^{-1}}$). This algorithm considers the motion of the compact
object within the pre-CC orbit, and the interaction between
the ejecta and the companion star. 
We set the
relative change of mass 
of the secondary due to stripping, ablation, and
accretion of SN ejecta following a fit to the three-dimensional
hydrodynamical simulations
of \cite{liu:15} for the impact of the ejecta on a
$M_2=3.5\,M_\odot$ star, which is the most massive companion they
considered. We also assume an efficiency of momentum
transfer\footnote{$\eta=0.5$ in the algorithm from \citetalias{tauris:98}.} from the
ejecta to the companion star of 0.5 (\citetalias{tauris:98};\citealt{hirai:18}). The contribution of 
the interactions between the SN ejecta and the secondary star to the total natal kick is
typically small, $\lesssim10\,\mathrm{km \ s^{-1}}$
(\citetalias{tauris:98}; \citealt{liu:15}; \citealt{hirai:18}).

In our fiducial run, the kick direction is assumed to be isotropically
distributed in the frame of the collapsing star \citep[][]{wongwathanarat:13,
  bear:17}, although  \cite{johnston:05} and \cite{kaplan:08} suggest that there
is weak evidence for kick-spin alignment in well observed pulsars
such as the Crab or Vela. \cite{johnston:05} also underlined that the
direction of the pulsar spin does not necessarily match the direction
of the pre-CC spin of the stellar core, if significant torques
develop during the CC. In our model
variations, we also consider kicks constrained in a cone with an
opening angle of
$\alpha=10\,\mathrm{degrees}$ oriented along the spin of the exploding
star (which we assume to be perpendicular to the orbital plane), and
kicks constrained at angles from the orbital plane $90 - \alpha
\leq 45\,\mathrm{degrees}$. The kick amplitude $|\mathbf{v}_k| \equiv
v_k$ is drawn from a Maxwellian distribution with one-dimensional root
mean squared dispersion $\sigma_\mathrm{kick}=265\,\mathrm{km \  s^{-1}}$ \citep[][]{hobbs:05}. Such
distribution is motivated by the observation of the line of sight velocities of pulsars in
\cite{hobbs:05} (see also \citealt{lyne:94}). 
We also compute
populations with $\sigma_\mathrm{kick}=0,\,300,\,1000\,\mathrm{km\
  s^{-1}}$ in our model variations.

The value of $v_k$ drawn is then reduced to consider the amount of matter that 
falls back after the successful launch of the SN shock,
i.e.~$v_k \rightarrow v_k(1-f_b)$, were $f_b$ is the fallback
fraction taken from\footnote{We correct for a missing parenthesis in
  their equations, which can be found by dimensional analysis \citep[][]{belczynski:17}.} Eq.~16 in \cite{fryer:12}, corresponding to their
``rapid SN engine'' which reproduces the NS-BH mass gap between
$\sim2-5\,M_\odot$ (e.g., \citealt{farr:11}, however see also \citealt{wyrzykowski:16}
regarding the existence of this mass gap). 
The inclusion of fallback
also determines the mass of the compact remnant obtained after each
CC event, and we set the mass boundary between NSs and BHs to
$2.5\,M_\odot$.

The fallback fraction is highly uncertain in BH formation. The algorithm we use
here prescribes a large amount of fallback, implying close to zero
BH kick amplitudes, sometimes referred to as a ``BH momentum kick'' \citep[][]{belczynski:08,stevenson:17}. Whether
this is realistic is subject of debate in the literature. Evidence for
non-zero BH natal kicks comes from the observed
Galactic latitude of BH X-ray binaries
(e.g., \citealt{fragos:09,repetto:12,repetto:15,repetto:17}, although
see also \citealt{mandel:16}), the
possibility of retrograde BH spin \citep[e.g.,][]{morningstar:14}, eccentric orbits in these systems
\citep[e.g.][]{remillard:06}, the
small number of WR+BH binaries \citep[e.g.,][]{dray:05}, and most
recently, the gravitational wave constraints on
BH spins \citep[e.g.,][]{oshaughnessy:17,wysocki:18}. Recently,
multi-dimensional calculations of fallback in CC resulting in the
formation of a BH showed that large natal kicks might be possible \citep[][]{chan:18}. Therefore, we
also consider a model without the rescaling of the kick amplitude, but still
including fallback for the remnant mass calculation. This variation is
effectively equivalent to a so-called ``velocity kick'' for the
BHs. To test an ``intermediate'' BH kick, we run a simulation using
$\sigma_\mathrm{kick}=100\,\mathrm{km\ s^{-1}}$ and no fallback
down scaling of the kick amplitude for
the CC resulting in BH formation, and out fiducial kick
($\sigma_\mathrm{kick}=265\,\mathrm{km\ s^{-1}}$, including fallback scaling)
for CC resulting in NS formation.

Another debated issue is whether the CC of low-mass iron
(or oxygen-neon-magnesium) cores produces small natal
kicks, possibly resulting in a bi-modal kick distribution
\citep[][]{katz:75, arzoumanian:02, pfahl:02,
  podsiadlowski:04, martin:09, knigge:11, wongwathanarat:13,
  beniamini:16, tauris:17, verbunt:17, verbunt:17b}. To
explore the possibility of low kicks for low mass collapsing cores, we
run one model variation reducing the natal kick for NSs less massive
than $1.35\,M_\odot$ \citep[][]{schwab:10,knigge:11} drawing
the natal kicks for these NSs from a Maxwellian with
$\sigma_\mathrm{kick}=30\,\mathrm{km\ s^{-1}}$, effectively creating a
double peaked distribution. For more massive NSs and BHs the kick is drawn from a
Maxwellian with $\sigma_\mathrm{kick}=265\,\mathrm{km\ s^{-1}}$.

In all our simulations, we draw $N_\mathrm{kick}=20$ intrinsic natal
kick directions and amplitudes for each CC in our population, effectively increasing the size of our sample by this
factor. Convergence tests showed no significant variations
of our results increasing $N_\mathrm{kick}$ to 50.

\subsection{Caveats}
\label{sec:caveats}
In addition to the limitations discussed  earlier in this section there are a few more caveats that should be kept in mind when comparing our simulations with observations (aside from the possible effects of biases).

We only consider runaways and walkaways resulting from the disruption of binary systems at the first stellar collapse. Observational samples of runaway stars also include another sub-population, coming from cluster ejection \citep[e.g.,][]{poveda:67, leonard:91, fujii:11,  allison:12, perets:12, banerjee:12, oh:16}.  Similarly, walkaway stars produced by the disruption of binaries will add to slow moving stars generated by the dissolution of clusters which did not go through binary interactions, see e.g.~\cite{allison:12}. In the Discussion we consider how spectroscopic measurements may allow to disentangle binary products from the outcome of purely dynamical processes.

All the calculations presented here implicitly assume a constant star formation history (SFH) for a duration longer than the longest stellar lifetime of interest.  However, a recent changes in the the SFH of a region will affect the ratio of runaways and walkaway stars over the number of normal stars.  For example, a recent increase in the SFH leads to an increase of systems that did not yet have time to interact. For a binary to be disrupted, a CC event is needed, and it can only happen after the lifetime of the primary star is over. This  introduces a delay corresponding to at least the lifetime of the most massive stars \citep[$\sim$3\,Myr,][]{zapartas:17}, between a SFH event and the first binary disruption.  A recent increase in the SFH can therefore decrease the fraction of runaway and walkaway stars in a population.

Finally we warn for the possible effects of stochasticity \citep[][]{eldridge:12}.  We have aimed to simulate sufficient binary systems such that are results are converged, but it should  be kept in mind that the fastest runaways come from relatively rare channels. In a relatively low mass stellar population will not fully sample these rare events and stochastic effects can be large.

\section{Example of the evolution of a binary system}
\label{sec:example}

\begin{figure*}[!htbp]
  \centering  
\includegraphics[width=\textwidth]{./figures/evol_example}  
  \caption{Evolution of an example binary system with 
    $M_1^\mathrm{ZAMS}=20\,M_\odot$, $M_2^\mathrm{ZAMS}=15\,M_\odot$
    binary with $P^\mathrm{ZAMS}=100\,\mathrm{days}$, described in
    \Secref{sec:example} (see also the cartoon in \Figref{fig:cartoon}). The top panel shows the
    primary, secondary, and total mass. The bottom panel shows the
    orbital velocity of the secondary (left
    y-axis) and period of the binary (right y-axis). The labels
    indicate the following phases: A. the MS of the primary, B. case B
    RLOF, emphasized by the vertical dashed line, C. the binary is
    made of a He star (possibly looking like a WR) and a
    rejuvenated MS secondary; D. (and the vertical dot-dashed line)
    marks the explosion of the primary in a SNIb/Ic which unbinds the
    NS and its companion. During the phase E1. the secondary is a slow
    moving walkaway star, traveling for a duration of about half the
    main sequence lifetime of the primary.}
  \label{fig:example}
\end{figure*}

We first describe the typical evolution of a binary system that
produces an unbound companion as a result of the disruption of the
system by the first CC event. The aim is to provide the reader with
some insight in our simulations. As a representative example, we consider
a  $M_1^\mathrm{ZAMS}=20\,M_\odot$ primary star,
a $M_2^\mathrm{ZAMS}=15\,M_\odot$ secondary with
$P^\mathrm{ZAMS}=100$\,days (corresponding to a separation
$a_\mathrm{ZAMS}\simeq300\,R_\odot$). \Figref{fig:example} shows the
evolution of the masses, orbital period, and orbital (and ejection,
after the CC of the primary) velocity of the secondary
as a function of time. The labels on the top axis correspond to the phases depicted in the cartoon shown in \Figref{fig:cartoon}.

\subsection{Binary evolution until first CC}
\label{sec:example_evol}

Our example system remains detached during the MS evolution
of the primary (phase A. in \Figref{fig:example}). Because of wind mass loss, the period increases by
about $7\%$.

After $\sim$8.8\,Myr, the primary leaves
the MS, and starts expanding on a thermal timescale as it burns hydrogen
in a shell. Shortly after, at the vertical dashed line B. in \Figref{fig:example}, the primary fills its Roche lobe
initiating case B
RLOF \citep[][]{kippenhahn:67}.
For a system which has initial mass ratio close to one
($q_\mathrm{ZAMS}=0.75$ in this example),
RLOF is almost conservative in our calculations, i.e.~nearly all the
mass lost by the primary is accreted by the companion (cf.~\citealt{schneider:15}). The
secondary star accretes  $\Delta M_2\simeq14\,M_\odot$, and becomes the
more massive star, now with a mass of about $28\,M_\odot$ (red line in the top
panel of \Figref{fig:example}). It is also spun up to
near to breakup rotation \citep[][]{packet:81,demink:09}, and it is rejuvenated because of the increased
mass and consequent growth of its convective core
\citep[][]{hellings:83,schneider:16}. The primary  becomes
a $\sim$$5.3\,M_\odot$ helium star \citep[][]{gotberg:17, gotberg:18}, and
the orbits widens during mass transfer, reaching a period
longer than 600\,days (phase C. in \Figref{fig:example}).

After about $10$\,Myr, the primary reaches CC (dot-dashed vertical line
D. in \Figref{fig:example}). Just before the core-collapse
of the primary, the orbit is circular with a period of
 about 700\,days. The secondary has a mass of about $28\,M_\odot$ and a pre-SN orbital velocity of
about $v_2 = (M_1/(M_1+M_2))v_\mathrm{orb} \simeq 12.5\,\mathrm{km\ s^{-1}}$.

If the system is disrupted at the first CC event, the final spatial velocity of the
secondary is nearly equal to its pre-explosion orbital velocity
\citep[e.g.,][]{blaauw:61,eldridge:11}. For comparison, similar
initial masses but a shorter initial period of $P^\mathrm{ZAMS}=7$\,days would experience a very similar
evolutionary path, but it would only widen to a pre-CC period of
$\sim$$45$\,days, resulting in a pre-CC orbital velocity of the
secondary slightly larger than $30\,\mathrm{km\  s^{-1}}$.

Because of the rejuvenating effect of
mass accretion,
the secondary remains on the MS for another $\sim$$3.6\,\mathrm{Myr}$ after
the CC of the primary.

\subsection{Effects of the natal kick on the post-CC orbit}
\label{sec:example_kick}

In our example system, the CC of
the primary results in a stripped envelope SN of type\footnote{We do not attempt to distinguish between stripped SN types (IIb, Ib or Ic depending on
the presence or lack of He lines in the spectrum). For a discussion of
how this can tentatively be done in population synthesis calculations,
and the related uncertainties, see \cite{zapartas:17b}.} Ib/Ic 
\citep[][]{filippenko:97}, which forms a NS of about
$M_\mathrm{NS}=1.6\,M_\odot$, with a fallback fraction $f_b=0.13$. The SN ejecta mass is $\Delta M_\mathrm{SN} = M_1^\mathrm{pre-CC}-M_\mathrm{NS}
\simeq 3.7\,M_\odot$, where $M_1^\mathrm{pre-CC}$ is the primary mass just
before the CC, and $M_\mathrm{NS}$ is the mass of the resulting
NS. Because the ejecta mass is less than half of the total pre-CC mass
of the system, $M_1^\mathrm{pre-CC}+M_2^\mathrm{pre-CC}\simeq33\,M_\odot$,
the loss of the SN ejecta alone is insufficient to
unbind the system \citep[][]{blaauw:61}. Whether the system is
disrupted depends on the amplitude and direction of the natal kick
imparted to the newly formed NS.

We distinguish three cases: (i) a large kick with $v_\mathrm{k}\gg v_\mathrm{orb}$ results in disruption of the binary regardless of the
direction, (ii) an intermediate kick $v_\mathrm{k} \simeq v_\mathrm{orb}$ disrupts the system only if the kick
orientation is favorable, while (iii) a small kick, $v_\mathrm{k}
\ll v_\mathrm{orb}$, never unbinds the binary, although it
can still modify its orbit (e.g., \citealt{brandt:95, kalogera:96}, \citetalias{tauris:98}).

To illustrate each case, we consider here three different fixed natal kick
amplitudes $v_k\simeq 870, 230$, and $43\,\mathrm{km\ s^{-1}}$ (respectively
corresponding to $1000,265$, and $50\,\mathrm{km \ s^{-1}}$ before the
rescaling of the kick amplitude with the fallback fraction). For comparison, the pre-CC orbital
velocity in our example system is $v_\mathrm{orb}^\mathrm{pre-CC}\simeq
100\,\mathrm{km\ s^{-1}}$ (corresponding to
$v_2^\mathrm{pre-CC}\simeq12.5\,\mathrm{km\ s^{-1}}$). For each kick amplitude $v_k$, we draw 10\,000 different directions,
isotropically distributed in the frame of the exploding star \citep[e.g.,][]{wongwathanarat:13}.

For large natal kicks, our example system is disrupted in 99.9\% of
all the directions drawn. In the remaining 0.1\%, the
compact object is shot in the envelope of the secondary (see
\Secref{sec:bound} for a brief discussion).
In the case of disruption, the ejected star acquires a velocity
$v_\mathrm{dis}\simeq v_2^\mathrm{pre-SN}=12.5\,\mathrm{km\
  s^{-1}}$. With $v_\mathrm{dis} < 30\,\mathrm{km\
  s^{-1}}$, it is below the threshold to
be classified as runaway as defined by \cite{blaauw:56,
  hoogerwerf:00, hoogerwerf:01}, and thus it is
an example of what we refer to as a walkaway star.

In the case of intermediate kick amplitudes (here we use
$230\,\mathrm{km\ s^{-1}}$), the angle $\theta$ between the kick direction and the
pre-CC orbital velocity of the primary, given by $\cos\theta = \mathbf{v_k}\cdot\mathbf{v}_1^\mathrm{pre-SN}/(v_k
\ v_1^\mathrm{pre-SN})$, determines whether the kick disrupts the
system. Kicks that are roughly aligned with the pre-CC orbital
velocity, i.e.\ $0 \lesssim \theta\lesssim\pi/4$, successfully disrupt the
binary. Conversely, kicks that are roughly oriented opposite to the
pre-CC orbital velocity of the exploding star just tend to slow the
orbital motion: the system will remain bound on an eccentric
orbit. In our example, we find that about $84\%$ of the kick
directions result in disruptions. All cases produce a walkaway
star with $v_\mathrm{dis}\simeq v_2^\mathrm{pre-SN}$.

With small kick amplitudes (here $v_k\lesssim50\,\mathrm{km \ s^{-1}}$),
our example system is never disrupted. The post-SN eccentricity, separation, and systemic velocity of the
binary vary with the kick direction \citep[e.g.,][]{brandt:95,
  kalogera:96, tauris:98}. We discuss the population of binaries surviving the
first CC (and hosting a compact object) in \Secref{sec:bound}.

In wide pre-CC systems, small kicks have a larger relative impact on the
velocity of the unbound secondary. The reason is that in wide pre-CC
orbits, smaller kicks result in a longer time for the compact object
to exit the orbit and exert a gravitational pull on the other
star. Conversely, for short pre-SN period systems (unlike the example
considered here) the effects of large kicks 
are more important. This is because the time for the compact object to
exit the pre-CC orbit is short, but larger kicks are more
likely to result in the disruption of the system than small kicks.

\section{Analytical  estimates}
\label{sec:analytics}

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.33\textwidth]{figures/specificAM}
  \includegraphics[width=0.33\textwidth]{figures/widening}
  \includegraphics[width=0.33\textwidth]{figures/speedup}
  \caption{Analytic estimates of the final-to-ZAMS ratio of the
     specific orbital angular momentum $h$ (left panel), the orbital
     separation $a$ (center panel), and orbital velocity of the
     secondary $v_2$ (right panel), assuming $q^\mathrm{ZAMS}=0.75$ and
     $\mu_\mathrm{env}=0.75$, which are representative for the example
     system of \Secref{sec:example}. Blue colors correspond to a
     decrease, red to an increase. The mass transfer efficiency
     $\beta_\mathrm{RLOF}$ and the angular momentum parameter
     $\gamma_\mathrm{RLOF}$ are assumed constant throughout the
     evolution. The horizontal dashed lines show the fraction of the specific
     orbital angular momentum removed for circumbinary disk mass loss
     ($\gamma_\mathrm{disk}$, yellow) and for isotropic re-emission
     ($\gamma_\mathrm{iso}$, green) for $q=0.75$. However, note that
     these fractions depend on the present-day $q$ rather than
     $q^\mathrm{ZAMS}$ and therefore they are not constant throughout the evolution in our simulations.}
   \label{fig:speedup_factor}
 \end{figure*}

The key quantity to define runaway and walkaway stars is their velocity.
As the example described above demonstrates, the final velocity
obtained at the binary disruption is of the
order of its final (pre-CC) orbital velocity $v_\mathrm{dis}\simeq
v_2^\mathrm{pre-CC}$, \citep[][]{blaauw:61,
  eldridge:11}. Therefore, using a few simplifying assumptions, we can
estimate the ejection velocity for systems that have experienced stable mass transfer, and express it in terms of the initial
parameters of the binary. In particular, we will express the final
orbital velocity as a function of the initial orbital velocity in the
binary, and use it as a proxy for the ejection velocity. This gives
insight on the possible outcomes of the binary evolution since we can estimate the typical velocity at ZAMS from
observational constraints on the binary populations, \citep[e.g.,][]{sana:12}.

Using conservation of linear momentum and
Kepler's third law, we can write the orbital velocity $v_2$ of the
secondary star at any time during the binary evolution as

\begin{equation}
  \label{eq:v2}
  v_2 =
  \frac{M_1}{M_1+M_2} v_\mathrm{orb} \equiv \frac{M_1}{M_1+M_2} \sqrt{\frac{G(M_1+M_2)}{a}}
  \ \ ,
\end{equation}
where $G$ is the gravitational constant, $a$ is the semimajor axis of
the orbit (assumed to be circular), and $M_1$, $M_2$ are the two
masses.

If we assume that a constant fraction
$\beta_\mathrm{RLOF}=|\dot{M}_\mathrm{acc}|/|\dot{M}_\mathrm{don}|$ of the mass
transferred is accreted by the secondary star, and that the mass not
accreted leaves the binary system with a constant fraction
$\gamma_\mathrm{RLOF}$ of the total
specific orbital angular momentum, we can relate the final and initial
separations with,
\begin{equation}
  \label{eq:af_ai}
  \frac{a_\mathrm{pre-CC}}{a_\mathrm{ZAMS}} =
  \left(\frac{M_1^\mathrm{ZAMS}}{M_1^\mathrm{pre-CC}}\frac{M_2^\mathrm{ZAMS}}{M_2^\mathrm{pre-CC}}\right)^2\left(\frac{M_1^\mathrm{pre-CC}+M_2^\mathrm{pre-CC}}{M_1^\mathrm{ZAMS}+M_2^\mathrm{ZAMS}}\right)^{2\gamma_\mathrm{RLOF}+1}
  \ \ .
\end{equation}

The parameter $\beta_\mathrm{RLOF}$ enters implicitly in the values of
the final pre-CC masses (see \Eqref{eq:DM}). Equation~\ref{eq:af_ai} neglects for simplicity all other processes that can
remove mass from the binary, such as stellar winds \citep[see ][for a
more elaborate expression]{soberman:97}.

The most common type of binary interaction is case B RLOF \citep[][]{kippenhahn:67}.
During case B RLOF, the primary typically loses nearly all its hydrogen-rich envelope
\citep[][]{gotberg:17,yoon:17, gotberg:18}. We parametrize the
fraction of the total mass in the envelope as $\mu_\mathrm{env} =
M_\mathrm{env}/M$, or in other words $1-\mu_\mathrm{env}$ is the fraction of mass in the
helium core. The parameter $\mu_\mathrm{env}$ is dependent on
the overshooting assumed in the underlying stellar evolution
models. The single massive star models from \cite{pols:98} give 
values in the range $0.6\lesssim\mu_\mathrm{env}\lesssim 0.8$ with the larger values
corresponding to less massive (and thus more common) stars. More
recent models \citep[e.g.][]{brott:11} typically adopt larger values
for the overshooting parameter, resulting in smaller values of $\mu_\mathrm{env}$.
Using this parametrization, we can relate the final masses of the two star in
the binary to their ZAMS masses as 
\begin{multline}
  \label{eq:DM}
 M_1^\mathrm{pre-CC} \simeq M_1^\mathrm{ZAMS}-\mu_\mathrm{env} M_1^\mathrm{ZAMS} \
 \ , \\
 M_2^\mathrm{pre-CC} \simeq M_2^\mathrm{ZAMS} +
 \beta_\mathrm{RLOF}\mu_\mathrm{env}M_1^\mathrm{ZAMS}
 \ \ . \hfill
\end{multline}
We can now express the final orbital velocity of the secondary $v_2^\mathrm{pre-CC}$ in
terms of the initial $v_2^\mathrm{ZAMS}$. Introducing $q\equiv q^\mathrm{ZAMS} =
M_2^\mathrm{ZAMS}/M_1^\mathrm{ZAMS}$,
$\beta\equiv\beta_\mathrm{RLOF}$, $\mu\equiv\mu_\mathrm{env}$, and
$\gamma\equiv\gamma_\mathrm{RLOF}$ to simplify the notation, we can write
\begin{equation}
  \label{eq:speedupdown}
  \frac{v_2^\mathrm{pre-CC}}{v_2^\mathrm{ZAMS}}=\left(\frac{q+\beta\mu-\mu
      q-\beta\mu^2}{q}\right)^2
  \left(\frac{1+q}{q+1+\beta\mu-\mu}\right)^{\gamma+1} \ \ .
\end{equation}

 In \Figref{fig:speedup_factor} we visualize the outcome of these
 analytical estimates for different assumptions for the mass transfer
 efficiency $\beta\equiv\beta_\mathrm{RLOF}$ and the angular momentum loss
 parameter $\gamma\equiv\gamma_\mathrm{RLOF}$. To make this figure, we adopt the parameters of the example system of
 \Secref{sec:example} ($q=0.75,\ \mu_\mathrm{env}=0.75$).

 The left panel of \Figref{fig:speedup_factor} shows the relative
 variation in the specific orbital angular momentum \mbox{$h\udef
 J_\mathrm{orb}/(M_1+M_2)=M_1M_2\sqrt{Ga}/(M_1+M_2)^{3/2}$}. For fully conservative evolution
 ($\beta=1$, white region on the right of this panel), $h$ stays
 constant. Also assuming $\gamma=1$ \citep[e.g.,][]{dominik:12} keeps
 the specific angular momentum constant, since this corresponds to
 losing mass and angular momentum at the same relative
 rate. Conversely, low values $\gamma\lesssim1$ result in a
 net increase of the specific orbital angular momentum because mass is
 lost relatively faster than angular momentum, while very large
 $\gamma\gtrsim 1.5$ values result in a loss of specific
 angular momentum.

 The central panel of \Figref{fig:speedup_factor} shows how the
 evolution affects the ratio of the final-to-initial separation (cf.~\Eqref{eq:af_ai}). The
 vast majority of the parameter space results in significant orbital
 widening (red colors), with only rather extreme angular momentum
 losses ($\gamma\gtrsim2$) in non-conservative
 ($\beta\lesssim0.6$) systems result in orbital shrinking.

 The right panel in \Figref{fig:speedup_factor} shows the main
 parameter of interest, the ratio of the final-to-initial orbital
 velocity of the initially less massive star in the system (cf.~\Eqref{eq:speedupdown}). Most
 assumptions for $\beta$ and $\gamma$ result in a decrease in the
 orbital velocity of the secondary (blue colors). This is the combined
 effect of (i) the orbital widening, and (ii) the increase mass of the secondary
 because of mass transfer. Only for very non-conservative systems
 ($\beta\lesssim0.3$) experiencing large
 angular momentum losses ($\gamma\gtrsim4$, corresponding to
 non-accreted mass removing four times the specific orbital angular
 momentum of the binary) the final velocity of the secondary is higher
 than its initial value.

 In our numerical simulations we adopt physically motivated model for the
 efficiency of mass transfer instead of constant parameters, $\beta$
 and $\gamma$, as we assumed in these analytical considerations. As we
 will argue later, most of the massive unbound companions result from
 nearly conservative mass transfer. Therefore, it is instructive to
 consider the analytical solution for $\beta = 1$. The expression for the final orbital velocity of the secondary (\Eqref{eq:speedupdown}) simplifies to 
 \begin{equation}
   \label{eq:cons}
   \frac{v_2^\mathrm{pre-CC}}{v_2^\mathrm{ZAMS}} =
   \left(\frac{(1-\mu)(q+\mu)}{q}\right)^2 = \left(\frac{3+4q}{16
       q}\right)^2 \ \ ,
 \end{equation}
 where we use as a typical value $\mu=3/4=0.75$ in the last
 step. \Eqref{eq:cons} gives a final-to-initial orbital velocity ratio for the
 secondary always smaller than one for initial mass
 ratios $q>0.25$. In other words, for all systems with initial mass
 ratios of interest evolving through stable and nearly conservative
 mass transfer, the secondary always slows down
 substantially, by up to a factor of 5 for an initially equal mass
 system. We emphasize that we have ignored the additional effect of mass loss by
 stellar winds. % (which typically reach terminal velocities much higher
 % than $v_\mathrm{orb}$)
 These cause the orbit to widen further and lead to a
 further slow down of the secondary star.

\begin{figure*}[tpb]
  \centering
  \tikzset{font=\small,
edge from parent fork down,
level distance=2cm,
sibling distance=3pt,
every node/.style={top color=blue!15,
  bottom color=blue!15,
  rectangle,rounded corners,
  minimum height=0mm,
  align=center,
},
level 4/.style={sibling distance=20pt},
level 5/.style={sibling distance=10pt},
edge from parent/.style={
  draw=blue!50,
  very thick,
},
blank/.style={draw=none},
}

\begin{tikzpicture}
  \Tree [
          .{Binaries with\\$M_1^\mathrm{ZAMS}\geq7.5\,M_\odot$\\ $q^\mathrm{ZAMS}\geq0.1$\\$0.15\leq \log_{10}(P^\mathrm{ZAMS}/\mathrm{days})\leq 5.5$}
               \edge [] node [draw=none, top color=white, bottom
               color=white, above, xshift=-10pt] {$22_{-9}^{+26}\%$};
               [
                   .{Stellar\\mergers}
               ]
               \edge [] node [draw=none, top color=white, bottom
               color=white, above, xshift=10pt] {$78_{-22}^{+9}\%$};
               [
                   .{Non mergers}
                   \edge [] node [draw=none, top color=white, bottom color=white, midway, above] {$1-\mathcal{D}=14_{-10}^{+22}\%$};
                   [
                       .{Non Disrupted\\ {\tiny
                           (cf.~\Secref{sec:bound})}}
                                          \edge [] node [draw=none, top color=white, bottom color=white, midway, left] {$36_{-15}^{+42}\%$};
                       [
                       .{NS companion}
                       ]
                       \edge [] node [draw=none, top color=white, bottom color=white, midway, right] {$64_{-42}^{+15}\%$};
                       [
                       .{BH companion}
                       ]
                   ]                       
                   \edge [] node [draw=none, top color=white, bottom color=white, midway, above] {\textcolor{red}{$\mathcal{D}=86_{-22}^{+10}\%$}};
                   [
                       .{Disrupted}
                       \edge [] node [draw=none, top color=white, bottom color=white, midway, above] {$25_{-10}^{+8}\%$};
                       [
                           .{post-MS companion \\ {\tiny
                               (cf.~\Secref{sec:non-MS-companions})}}
                           \edge [] node [draw=none, top color=white,
                           bottom color=white,  above, xshift=-6pt] {$53_{-46}^{+12}\%$};
                           [
                              .{WD}
                           ]
                           \edge [] node [draw=none, top color=white,
                           bottom color=white,
                           below, yshift=-40pt] {$23_{-20}^{+38}\%$};
                           [
                               .{He star}
                           ]
                           \edge [] node [draw=none, top color=white,
                           bottom color=white,  above, xshift=6pt] {$25_{-6}^{+23}\%$};
                           [
                               .{H-rich}
                           ]                        
                      ]
                      \edge [] node [draw=none, top color=white, bottom color=white, midway, above] {$75_{-8}^{+10}\%$};
                      [
                           .{MS companion \\ {\tiny (cf.~\Secref{sec:MS})}}
                           \edge [] node [draw=none, top color=white,
                           bottom color=white, above, xshift=-3pt] {$5_{-2}^{+14}\%$};
                           [
                              .{Runaway\\$v_\mathrm{dis}\geq30\,\mathrm{km\ s^{-1}}$}
                           ]
                           \edge [] node [draw=none, top color=white,
                           bottom color=white, above, xshift=+3pt] {$95_{-14}^{+2}\%$};
                           [
                               .{Walkaway\\$v_\mathrm{dis}<30\,\mathrm{km\ s^{-1}}$}
                           ]                              
                     ]
                   ]       
               ]
        ]    
\end{tikzpicture}
  % \input{figures/fig_tree/tree}% 
  \caption{Overview of the possible binary evolution scenarios up to
    the first CC event. The
    branching ratios shown are from our fiducial simulation, and we
    highlight in red the disruption fraction
    $\mathcal{D}$. The errors on each fraction exclude the run without
  SN kicks ($\sigma_\mathrm{kick}=0\,\mathrm{km\ s^{-1}}$), which
  produces an unrealistically low disruption fraction
  (cf.~\Tabref{tab:parameters} and \Secref{sec:param_variation}).}
  \label{fig:tree}
\end{figure*}

Using the typical values of our example system discussed in
\Secref{sec:example} ($q = 0.75$), we find $v_2^\mathrm{pre-CC}
\simeq 0.25 \, v_2^\mathrm{ZAMS} \simeq  20\,\mathrm{km\ s^{-1}}$.
Our numerical model for this system gives an even lower value,
primarily because of the additional effect of widening as a result of
mass loss through stellar winds. As a sanity check we recomputed the
evolution of the example system, but artificially switching off the
stellar wind mass loss for both stars. This gives
$v_2^\mathrm{pre-CC}\simeq22\,\mathrm{km\ s^{-1}}$, in good agreement
with the analitycal estimate. The remaining difference is due to the fact that the mass transfer is not fully conservative.

For the analytic estimates presented in this section, we only
considered the case of stable mass transfer. Binary evolution through
unstable mass transfer takes place for a limited range of the initial
distributions \citep[e.g.,][]{soberman:97,schneider:15}. It is
expected to result in a common envelope phase, followed by substantial
shrinking of the orbit either leading to a merger or the formation of
a compact binary system if the envelope is ejected successfully.
While this is a very interesting pathway to create very fast runaway
stars, the numerical simulations presented later indicate that this
channel produces primarily low mass runaway stars. This is because it typically concerns system with extreme initial mass ratios in which the secondary does not significantly gains mass. This means that the secondaries in these systems are typically not very massive and therefore this channel does not significantly contribute to the production of massive early type runaways, at least in our simulations. We return to this in \Secref{sec:discussion}.

\section{Ejected companions}
\label{sec:fiducial_res}

In this section, present the results obtained from our numerical simulations for a full population computed with our fiducial assumptions (cf.~\Secref{sec:methods}). We describe the robustness of our findings against model assumptions in \Secref{sec:param_variation}.

Our main goal is to characterize the
velocity acquired by the companion stars ejected at the time of the first CC
in the binary. Therefore, we ignore all binaries that merge, which constitute $\sim$$22\%$ of our fiducial
population (cf.~\Figref{fig:tree}), in good agreement
with~\cite{sana:12,demink:14,zapartas:17,zapartas:17b}; see also \cite{kochanek:14}.

Once the merging systems are excluded, the first CC happens in the presence
of a companion. The CC can result in either the disruption of the
binary ($\mathcal{D}=86\%$ in our fiducial run), or the newly formed
compact object remains bound to the companion star. In
\Secref{sec:MS} we focus on the population of ejected MS companions,
which are the $75\%$ of all the ejected companions in our fiducial run. We consider companions
that have already evolved off the MS at the time of the first CC in \Secref{sec:non-MS-companions}.

For completeness, we describe the systems
with a MS companion remaining bound in \Secref{sec:bound}.

\subsection{Ejected main sequence companions}
\label{sec:MS}

\begin{figure*}[htp]
  \centering
  \includegraphics[width=\textwidth]{./figures/v2_dist_fiducial}
  \caption{Velocity distribution of MS stars ejected
    from a binary system at the time of the first CC. The top panel shows
    the corresponding cumulative distributions. The 
    walkaways are one order of magnitude more numerous than the
    runaways. The gray, blue, and red histograms show all MS
    secondaries, only MS stars more massive than $7.5\,M_\odot$ (i.e.,
    roughly those that might experience CC), and only MS stars more
    massive than $15\,M_\odot$ (i.e., roughly the O-type stars),
    respectively. The mass considered here is taken right after the CC of the primary star. The dashed
    lines show the distribution of walkaways and runaways that have
    gone through RLOF (or common envelope evolution) before being
    ejected. Almost all massive runaways and walkaways, and the
    majority of the ejected stars at large velocities, have gone
    through RLOF. The inset plot magnifies the runaway regime
    $v_\mathrm{dis}\geq30\,\mathrm{km\ s^{-1}}$. See also
    \Figref{fig:v_dist_log} for a wider velocity range and
    \Figref{fig:v_dist_obs} for a figure accounting for the remaining
    lifetime as walkaway or runaway star.}
  \label{fig:v_dist}
\end{figure*}

The vast majority of binaries hosting a MS companion is disrupted at the time of
the first CC. Figure~\ref{fig:v_dist}
shows the distribution of velocities of the ejected companions. The three different colors subdivide the population based on
the mass of the ejected companion. The gray histogram shows all
ejected MS companions. The blue histogram shows only massive MS
companions, $M_2^\mathrm{post-CC}\geq 7.5\,M_\odot$, where
$M_2^\mathrm{post-CC}$ is the mass of the secondary just after the
collapse of the primary. This group roughly corresponds to all
secondaries that will experience CC at the end of their lifetime. The red histogram shows the velocity
distribution of companions with mass larger than $M_2^\mathrm{post-CC}\geq15\,M_\odot$, corresponding roughly to O-type
stars. In the rest of this section, we quote the ratios for the
population of massive stars (i.e.~$M_2^\mathrm{post-CC}\geq 7.5\,M_\odot$) unless
otherwise stated, see also \Tabref{tab:parameters} for more
information.

Although we do produce runaways, the majority of the systems ejects a star slower than $30\,\mathrm{km \
  s^{-1}}$, i.e., a walkaway star \citep[][]{de-Mink+2012,lennon:16}. We find that the CC of a star
with a MS binary companion is about twenty times more likely to produce a
walkaway than a runaway. In the first line of \Tabref{tab:parameters},
we list the ratio $\mathcal{R}$ of walkaway to runaway stars produced per CC event in a binary. From our
fiducial simulation, we obtain $\mathcal{R}_\mathrm{7.5}\simeq20$ for
massive companions.
As shown in
\Tabref{tab:parameters}, considering progressively larger masses, the number of walkaways
produced per each runaway from a binary increases.

Runaways resulting from the disruption of binaries rarely exceed $60\,\mathrm{km \
  s^{-1}}$: about 99.8\% of the massive unbound companions we simulate are slower than
this threshold. We note that higher velocity are not strictly
forbidden, but extremely unlikely (cf.~\Figref{fig:v_dist_log} which
shows the velocity distribution on a logarithmic scale).
Low mass runaways can reach much
tighter pre-CC orbits through common envelope evolution (without
significant accretion), and thus reach higher ejection velocities, but typically
not in excess of $400\,\mathrm{km\ s^{-1}}$. Stars significantly
faster than this are likely produced by different mechanisms
(see \citealt{boubert:17b}, and \citealt{brown:15} for a review). However, in extremely rare cases
where a very short pre-CC binary is formed and also successfully disrupted,
much higher velocities can be achieved.  The
absolute maximum velocity we obtain is about $1110\,\mathrm{km\
  s^{-1}}$, in good agreement with \cite{tauris:15}.

The slowest stars come from the widest pre-CC binaries:
the low-velocity drop-off of the distribution in \Figref{fig:v_dist}
is thus
an effect of the upper-limit on our period distribution (corresponding
to an upper-limit on the pre-CC period).

The velocity distribution peaks at $v_\mathrm{dis}\simeq6\,\mathrm{km\
  s^{-1}}$. Ejected MS companions with such velocity typically
originate from systems with $q_\mathrm{ZAMS}\gtrsim0.5$ and long initial
periods ($\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})\simeq3$),
regardless of the stellar masses. These systems experience late case B
or case C mass transfer \citep[][]{lauterborn:70}, but owing to the
sufficiently high mass ratio, the mass transfer remains conservative
and the system widens.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/Dhist}
  \caption{Distribution of the maximum distance $L$ that ejected
    MS secondaries can reach, neglecting any gravitational
    potential. The blue and red histograms are for ejected stars more
    massive than $7.5\,M_\odot$ and $15\,M_\odot$, respectively. The
    vertical dashed lines mark the mean values of
    the distributions.}
  \label{fig:Dhist}
\end{figure}

Among low and intermediate mass walkaways with
$v_\mathrm{dis}\simeq6\,\mathrm{km\ s^{-1}}$, there is also a contribution from
wide systems which do not experience mass transfer. This contribution
is almost negligible for ejected stars more massive than
$15\,M_\odot$. This can be seen from the dashed distributions in
\Figref{fig:v_dist}, which show that the majority
of massive runaways and walkaways have accreted mass from
their companion before its CC. Over the entire velocity range shown
(i.e., considering both runaways and walkaways), $71\%$ of ejected MS
secondaries more massive than $7.5\,M_\odot$, and $91\%$ of those more
massive than $15\,M_\odot$ have accreted from their companion in
the previous evolution. Among low mass ejected companions, $50\%$ have
experienced either RLOF or common envelope evolution before their
ejection from the binary. From the inset in \Figref{fig:v_dist}, where
the blue and red dashed lines representing post-interaction systems
overlap with the full distribution, it is clear that all massive
runaways from binary disruption have experienced mass transfer.

The majority of the ejected massive MS
companions are not ``usual'' single stars, and this remains true for a
very large fraction of the low mass MS ejected companions. Before the binary disruption,
they have accreted mass from their companion. Mass transfer also causes the convective core of the accretor to grow,
resulting in rejuvenation of the accreting star
\citep[e.g.,][]{hellings:83,schneider:16}. The fact that most ejected
MS stars are accretors is the result of a
combination of assumptions. For disruption to be possible,
the primary needs to be massive enough to collapse. We assume a flat
mass ratio distribution, implying that on average $M_2^\mathrm{ZAMS}\simeq
0.5\,M_1^\mathrm{ZAMS}$, and we also assume that mass transfer is
unstable and leads to mergers if the mass ratio is too extreme
($M_\mathrm{acc}/M_\mathrm{don}<q_\mathrm{crit}$), therefore binaries
that can eject a star have initial mass ratios closer to one than an
average system. Finally, when stable mass transfer occurs, our fiducial
assumption for the accretion efficiency results in rather conservative
mass transfer over most of the period range considered
\citep[][]{schneider:15}.

To make an estimate of the
ratio of walkaways  per runaway {\it existing at a given time}
requires us to
consider the finite MS lifetime of the ejected stars. Assuming that the star formation rate in the galaxy is
constant and therefore that the Galactic population is in a steady state,
the ratio of walkaways per runaway is
$\sim$13 for masses larger $7.5\,M_\odot$.

Accounting for the durations of each evolutionary phase, we can also
quantify the fraction of MS stars more massive than $15\,M_\odot$ that are
runaway or walkaway stars. In our fiducial simulation, we find a
runaway fraction for masses larger than $15\,M_\odot$ of about $\sim$$0.5\%$ and a
walkaway fraction of $\sim$$10\%$. We only simulate binaries with primaries more
massive than $7.5\,M_\odot$, therefore we do not have in our
population intermediate-mass binaries which could contribute through
mass transfer and/or mergers  to the
normalization factor for the runaway and walkaway fraction at lower
masses \citep[][]{zapartas:17}. We show in \Figref{fig:v_dist_obs} the distribution in
velocity accounting for the lifetime of the stars in each bin. This
can be directly compared to observations as long as the effects of the
Galactic potential can be neglected.

\begin{figure*}[!htbp]
  \centering
  \includegraphics[width=0.75\textwidth]{figures/veq_vrw}
  \caption{Equatorial velocity at the time of ejection for massive ($M_\mathrm{dis}\geq7.5\,M_\odot$) companions as a function
    of the ejection velocity. Brighter colors indicate location of the
    parameter space more populated by our fiducial simulation. All the massive runaway stars spin with
    an equatorial velocity of $\sim$$500\,\mathrm{km\ s^{-1}}$, close to
  breakup rotation, since they have accreted mass from their companion
  before being ejected. The spread is due to wind spin down before the
  binary disruption. We do not include the effect of projecting on
  the line of sight, or post-ejection wind spin down in this
  plot.  The
    vertical dot-dashed line marks the threshold to define runaway
    stars.}
  \label{fig:vrot_vrw}
\end{figure*}

Figure~\ref{fig:Dhist} shows the distribution of distances traveled by
massive, unbound stars, calculated by multiplying their velocity
$v_\mathrm{dis}$ by their remaining lifetime on the MS, which includes
rejuvenation if the star has accreted mass, plus 10\% of the
MS duration of a star with the same helium core mass, to account for
the duration of the helium core burning phase in the rejuvenated
star\footnote{The duration of the helium core burning depends on the helium core mass, and thus the overshooting
parameter. The value of $10\%$ of the MS duration is typical for overshooting values
larger than what assumed in \cite{pols:98}.}.
This approach neglects the effect of an external potential on the trajectory of the
stars \citep[see][for how this could be done]{boubert:17a,boubert:18}, so it is effectively an upper limit to the distance they can travel. Considering both runaway and walkaway stars with masses
$M_\mathrm{dis} \geq7.5\,M_\odot$, we find
the mean distance $\langle L \rangle$ they travel before experiencing
CC to be 126\,pc, see the blue vertical dashed line in
\Figref{fig:Dhist} and also \Tabref{tab:parameters}. This number
rises to $\langle L_\mathrm{run}\rangle=584$\,pc considering only 
runaways (faster than $30\,\mathrm{km \ s^{-1}}$) more massive than
$7.5\,M_\odot$, but it remains $\langle
L_\mathrm{walk}\rangle=103$\,pc considering massive walkaways only
(slower than $30\,\mathrm{km \ s^{-1}}$). For comparison, the typical
size of OB-associations is of the order of tens of pc,
while the Galactic thin and thick disks have a vertical scale height
of $\sim$300 and $\sim$1500\,pc, respectively
\citep[e.g.,][]{gilmore:83}.

\subsection{Spin of the main sequence ejected companions}

Figure~\ref{fig:vrot_vrw} shows the equatorial rotational velocity
$v_\mathrm{dis}^\mathrm{eq}$ for
the ejected MS companions with $M_\mathrm{dis}\geq7.5\,M_\odot$. The top panels show the normalized one-dimensional
distribution of ejection velocity for the whole sample (same as the
blue distribution in \Figref{fig:v_dist}), and for the runaways only,
respectively. The side panels show the normalized equatorial
rotational velocity distribution
for the whole sample, and for the runaways only, respectively.
We do not include
in \Figref{fig:vrot_vrw} any projection effects due to the inclination
of the stellar spin axis with respect to the line of sight, nor the
wind spin down due to the evolution after the binary disruption. The
combination of these two effects would smear out the distribution in
the vertical direction, allowing any physically possible
\emph{projected} equatorial rotational velocity to be observable.

The main feature at the top of the central panel (also visible in the rightmost panel) shows that almost all the massive runaways are spinning
nearly at breakup rotation at the time of the ejection, because of the
relatively recent spin up during the mass transfer phase(s).

The walkaways ($v_\mathrm{dis}<30\,\mathrm{km\ s^{-1}}$) instead can
have a broader range of rotational velocities: this is because a
subset of them comes from wide, non-interacting binaries. The right
panel including both the walkaway and runaway population shows two
peaks for $v_\mathrm{dis}^\mathrm{eq}\simeq200$ and $500\,\mathrm{km\
  s^{-1}}$. The former is due to the combination of the accretion-induced spin up and the wind spin down
before the ejection, which can be more efficient for walkaways
experiencing early mass transfer and with longer delay between the
mass transfer and the disruption of the binary. The former peak is \emph{not} related to tidal locking, which
is efficient only for tight orbits that would result in much larger ejection
velocities $v_\mathrm{dis}$. The latter is populated by secondaries ejected after a
mass transfer phase (analogous to the peak in the equatorial velocity
distribution for the runaways).

\subsection{Ejected post-main sequence companions}
\label{sec:non-MS-companions}

Figure \ref{fig:v2_postMS} shows the velocity distribution of companion stars which have evolved off
the MS by the time of the first CC in the binary. All
together, these correspond roughly to 25\% of the population of
disrupted binaries (cf.~\Figref{fig:tree}). We distinguish three sub-populations:
hydrogen-rich non-degenerate secondaries (mostly of red-supergiant
secondaries, corresponding to $\sim$$4\%$
of the entire simulated population, including mergers and
non-disrupted binaries), white dwarfs (WD, $\sim$8\% of
the total), and helium stars which have lost their hydrogen-rich envelope because of previous
binary interactions ($\sim4\%$ of the total).

For the first group, characterized by large stellar radii, the velocity distribution resembles
closely that of the MS ejected companions, with a large number
of walkaways (roughly corresponding to $96\%$ of the hydrogen-rich
non-degenerate ejected companions) compared to the runaways.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/v2_dist_postMS.pdf}
  \caption{Velocity distribution of post-MS ejected companions. Helium
    stars and white dwarfs (WD) can have much shorter pre-CC orbits
    because of their smaller radii, and thus they can reach much
    higher ejection velocities. Conversely, the number of non-degenerate, hydrogen-rich
    post-MS stars is dominated by red supergiants with very large
    radii, which can only exist at large separation from the exploding
    companion, and thus typically have much slower velocities. The top
    panel shows the corresponding cumulative distributions.}
  \label{fig:v2_postMS}
\end{figure}

Ejected WDs and helium stars come from different evolutionary paths. Mass transfer during the primary MS
can reduce the helium core mass thus slowing its evolution, while
accretion of mass speeds up the evolution of the companion, although this is compensated by the rejuvenation effect. In
$\sim23\%$ of our full population (excluding stellar mergers), this leads
to a reversal of the CC order, with the secondary exploding first \citep[][]{pols:94}. In
these cases the star that is ejected is the initially more massive, which
has become, by the time of the CC, a helium star or a WD.

Because of their smaller radii, WDs and helium stars can
have much shorter pre-CC orbits and therefore reach velocities much
larger than (super-)giant stars. More importantly, these stars are what
is left of the initial donor and they are thus less massive than
their companion after the mass transfer phase.

This results in ejection velocities of
the WDs spanning a range  $15 \lesssim v_\mathrm{dis}/\mathrm{[km\
  s^{-1}]}\lesssim 90$, and helium stars are even faster with $70 \lesssim v_\mathrm{dis}/\mathrm{[km\ s^{-1}]}\lesssim
120$. This difference arises because it takes longer for a star to
become a WD than its helium core burning duration. Therefore, WDs are
ejected by systems where the rejuvenation of the 
secondary is not extreme. This means that these systems have
a less conservative mass transfer phase than binaries ejecting a
helium star, which corresponds to less
rejuvenation of the secondary and more orbital widening, and thus
lower velocity of the WD at ekection. Moreover, if
the companion star at pre-CC stage is a WD,  there is a longer time
for the binary to widen, e.g.,~because of winds, than if it is a
helium star.

Conversely, if the ejected star is a non-degenerate helium
star, then the entire evolution of the rejuvenated secondary needs to
be faster than the remaining lifetime of the primary (which now
has a reduced mass): the time delay between the RLOF phase and the
first CC ejecting the helium star is much shorter and there is not enough
time for significant orbital widening.

Figure~\ref{fig:v2_postMS} also shows a small population of slow ($v_\mathrm{dis} <
30\,\mathrm{km\ s^{-1}}$) helium stars. These are very massive
($M_2\gtrsim 60\,M_\odot$) stars characterized by large Wolf-Rayet
wind mass loss rates, which cause significant orbital widening before
the first CC, resulting in the slow ejection velocity. We emphasize that \Figref{fig:v2_postMS} shows the velocity of
stars ejected while in the corresponding evolutionary stage, and stars
ejected during their MS will then evolve into post-MS hydrogen-rich,
or possibly Wolf-Rayet stars, creating a more complicated velocity distribution of
massive Wolf-Rayet stars \citep[see, e.g.,][]{dray:05, eldridge:11}.

\section{Impact of model variations}
\label{sec:param_variation}

Our predictions depend on a set of parametrized assumptions. We
perform a systematic study of the impact of these uncertainties by
varying the values of the free parameters one-by-one, as we describe in
\Secref{sec:methods}. In each variation, all the other parameters are set to their
fiducial value.

Table \ref{tab:parameters} summarizes our results for binaries with a
collapsing star and a MS companion. If our extreme assumptions
do not change much our synthetic population, then the outcome is
robust and independent of the particular physical process. Conversely,
strong dependence on any one parameter indicates that such parameter could be constrained by
comparing to observed samples, and therefore allow for physical tests
of the process it represents. Overall, our results are robust and relatively insensitive to
uncertain physical processes, at least within the framework of
parametrizations we assume.

\subsection{Robust  predictions}

The fraction of binaries that are disrupted by the first CC is always larger
than $75\%$, with two notable exceptions. These concern the natal kick
distribution assumed. With a double Maxwellian kick distribution with
$\sigma_\mathrm{kick}=30\,\mathrm{km\ s^{-1}}$ for NSs less massive
than $1.35\,M_\odot$ (and our fiducial kick for more massive
remnants), the disruption fraction
decreases to $\mathcal{D}=65\%$ even though this changes the kick only
for a small subset of the systems. Low mass NSs are the
typical result of progenitors with small total and core masses, which
are favored by the IMF.

If instead we make the extreme assumption of zero
natal kick in all CC ($\sigma_\mathrm{kick}=0\,\mathrm{km\ s^{-1}}$), then only $\mathcal{D}=16\%$ of the
binaries are disrupted. In these cases the disruption is
only due to the mass loss at the moment of the explosion (Blaauw
kick). Since the ejecta mass must be larger than half of the total
mass of the system to disrupt the binary \citep[][]{blaauw:61}, and since after mass
transfer the primary is typically less massive than the companion, only very wide binaries which did not experience mass transfer
can be disrupted in such a way. Therefore, in this extreme assumption,
the ejected stars are on average
less massive and slower compared to calculations including a natal
kick. We emphasize that this parameter variation is in contrast with
the observed distribution of pulsar velocities, and we
include it for insight only. This variation does not produce any massive
runaways, which is also in contrast with observations.

We obtain the largest disruption
fraction when we do not rescale the natal kick with the fallback
fraction: in this model, up to $97\%$ of the systems are
disrupted. This assumption is equivalent to assuming a ``BH velocity
kick'' \mbox{\citep[e.g.,][]{belczynski:08,fryer:12}}: BHs and NSs receive
the same kick amplitude at birth, despite the differences in their masses. Therefore, many CC events
resulting in the formation of a massive BH with large fallback
fractions (up to $f_b=1$) still result in the disruption of the binary.

Despite the overall large disruption fraction in all our parameter
variations, the predicted runaway fraction $f_{15}^\mathrm{RW}$ for masses larger than $15\,M_\odot$ is low for all our
parameter variations and never exceeds a few percent. Similarly, the walkaway
fraction $f_{15}^\mathrm{WA}$ is also robust, and is generally around
$10\%$. In other words, the walkaways outnumber the runaways by
a factor of 10-30 (smaller for larger masses), regardless of the
physical assumptions we make. The
most notable exceptions are computations at lower metallicity (see
\Secref{sec:Z}). Another case producing slightly smaller $\mathcal{R}$ is
fully non-conservative mass transfer ($\beta_\mathrm{RLOF}=0$): more
systems are in a very close pre-CC orbits and the average velocity of
an ejected secondary is higher than in our fiducial assumptions.

The average mass of the walkaways, runaways, and all ejected
secondaries ($\langle M_2^\mathrm{walk} \rangle$, $\langle
M_2^\mathrm{run} \rangle$, and $\langle M_2 \rangle$, respectively) exceeds $7.5\,M_\odot$ in almost all our parameter variations. With
fully non-conservative mass transfer ($\beta_\mathrm{RLOF}=0$),
the secondary accretes no mass. This makes them on average less
massive than in our fiducial assumptions at the moment of the first CC, and also allows
for less orbital widening (the donor has to lose more mass before the
mass ratio reverts), resulting in a faster population of disrupted
stars with lower masses compared to our fiducial case. If we assume that case B mass transfer is
unstable whenever the accretor is less massive than the donor
($q_\mathrm{crit, B}=1$), or if we take a negative slope of the mass ratio
distribution ($\kappa=-1$), then the fraction of binary systems that
merge after the end of the primary MS increases compared to our
fiducial case. 
When a binary system merges, it cannot eject a walkaway or runaway.
Finally, assuming no natal kicks ($\sigma_\mathrm{kick}=0$), the only
systems disrupted are those where the SN ejecta consist of more than
half of the total mass of the binary at the pre-explosion stage. This
requirement naturally biases the population of ejected companions
towards low masses. If we assume that the kick velocity is not scaled down by the amount
of fallback (``velocity kick'' for BHs), CC events from more massive primaries can also disrupt
the system. On average, this produces more massive ejected
secondaries compared to our fiducial case. Since the remaining lifetimes are shorter, the
distance traveled by the population of ejected secondaries is
also decreased, despite their higher velocity.

Our assumptions on the stability of case B RLOF (assumed to be stable
if the accretor-to-donor mass ratio is smaller than $q_\mathrm{crit,B}$) impact the
velocity and mass distribution of the ejected stars. The sequence from
more stable to more unstable
$q_\mathrm{crit,B}=1,\,0.5,\,0.4$\,(fiducial
value),\,0 shows progressively higher $\mathcal{R}$ indicating
relatively less runaways, but higher average velocity $\langle v\rangle$ (and
consequently also a larger $\langle L \rangle$). The increase in velocity and
decrease in average mass of the ejected companions for higher
$q_\mathrm{crit,B}$ indicates an increased importance of evolutionary
channels involving a common-envelope evolution leading to tighter
pre-CC orbits without significant accretion of mass from the companion.

The average distance traveled by walkaways and runaways is also a
robust prediction. The entire population of ejected stars typically
travels slightly farther than $\langle L\rangle\simeq120$\,pc, which
corresponds to an average distance for the runaways of $\langle
L\rangle_\mathrm{run}\gtrsim500$\,pc, while the walkaways alone reach $\langle L\rangle_\mathrm{run}\gtrsim100$\,pc. The
effect of rejuvenation, allowing for increased lifetimes of
these stars, is largely responsible for this.

These distances are
significantly shortened only assuming fully conservative mass transfer
($\beta_\mathrm{RLOF}=1$), or assuming higher angular momentum loss
(e.g., through a circumbinary disk, $\gamma_\mathrm{RLOF}=\gamma_\mathrm{disk}$). The former results in
higher masses of the ejected secondaries and thus smaller velocities
and shorter lifetimes,
while the latter results in more orbital shrinking and mergers. Another variation
resulting in small $\langle
L\rangle_\mathrm{walk}\simeq66\,\mathrm{pc}$ is that with zero natal kicks
($\sigma_\mathrm{kick}=0\,\kms$). As mentioned above, only the widest pre-CC binary can be
disrupted in this case producing a much slower population of ejected
MS stars.

The predicted rotational velocity of the ejected companions is also a
robust prediction. The vast majority of massive
($M_\mathrm{dis}\geq7.5\,M_\odot$) main sequence runaways accrete mass
from their companions, and thus spin up to critical rotation, before
being ejected. Visual inspection of plots like the one shown in
\Figref{fig:vrot_vrw} for our fiducial simulation show no variation
when changing the initial rotation rate of the stars, confirming that
mass transfer in binaries overwrites the initial rotation of the
accretor. %% beta=0 might have an issue with wind spin up.

\subsection{The mass function of massive runaways as probes for black hole kicks}
\label{sec:bh_kicks_mf_rw}
\begin{figure}[!bp]
  \centering
  \includegraphics[width=.5\textwidth]{./figures/BH_MF_sigma100}%BH_kick_MF}
  \caption{The mass function of massive ($M_\mathrm{dis}\geq7.5\,M_\odot$), runaway
    (faster than $\geq30\,\mathrm{km \ s^{-1}}$) stars depends on the
    assumptions for the BH kick. The bottom panel
    shows our fiducial simulation
    with $\sigma_\mathrm{kick}=265\,\mathrm{km \ s^{-1}}$, including the fallback
    downscaling of the kick both for NSs and BHs. The fallback is an
    important effect only for BHs. The central panel
    shows the mass function for intermediate BH kicks
    ($\sigma_\mathrm{kick}=100\,\mathrm{km\ s^{-1}}$, no fallback
    downscaling) and our fiducial kick for NSs. The top panel shows
    the distribution with our fiducial kick amplitude, but without fallback downscaling (i.e., BH
    velocity kick). The dashed lines indicate the $90^\mathrm{th}$
    percentiles of the mass distributions.}
  \label{fig:bh_kicks}
\end{figure}

\begin{figure*}[htbp]
  \centering
   \includegraphics[width=0.33\textwidth]{figures/M_v_fiducial}
   \includegraphics[width=0.33\textwidth]{figures/M_v_sigma100}
   \includegraphics[width=0.33\textwidth]{figures/M_v_no_fb}
   \caption{Mass-velocity correlation for the three BH kick variations
     discussed in \Secref{sec:bh_kicks_mf_rw}. The brighter colors indicate
     where on this plane we predict more ejected companions, the
     colors have a logarithmic scale. The left panel shows
     our fiducial distribution, the central panel shows the case for
     intermediate BH kicks ($\sigma_\mathrm{kick}=100\,\mathrm{km\
       s^{-1}}$ and no fallback), and the right panel shows large BH
     kicks (same as for NSs). The NS kick is effectively the same in
     all panels, since the amount of fallback is very small for SNe
     producing a NS.}
   \label{fig:M_vrw}
\end{figure*}

The mass distribution of runaways carries
information about the BH kicks. With a sufficiently large sample, it might be used to
discriminate between BH velocity or momentum kicks, at least in a
statistical sense. We focus on the mass function of massive
($M_\mathrm{dis}\geq7.5\,M_\odot$) and fast runaways
($v_\mathrm{dis}\geq30\,\mathrm{km \ s^{-1}}$)
only, i.e., not including the more common walkaways, since we expect
upcoming observational distributions to provide cleaner sample for
these. Ejected companions with high masses are more likely to come from
binaries with an initially very massive primary resulting in BH
formation at the end of its evolution.
The constraints on the amplitude of BH kicks from unbound companions are
complimentary to those from binaries surviving the first CC, and Gaia
might be able to probe both populations of bound and ejected
companions, \citep[][]{breivik:17}.

\Figref{fig:bh_kicks}
shows the mass distribution of runaways
($v_\mathrm{dis}>30\,\mathrm{km\ s^{-1}}$) more massive than $7.5\,M_\odot$ for three BH
formation scenario. The bottom panel shows our fiducial
simulation, in which natal kicks are damped by fallback (effectively,
this is equivalent to a BH ``momentum kick''). In this case, the $90^\mathrm{th}$
percentile of the massive runaway mass function is $\sim$$25\,M_\odot$.
In this model, the amount of fallback and the amplitude of the BH natal kicks
are degenerate parameters. Because of the large fallback fraction
prescribed by the \cite{fryer:12} algorithm, the runaway mass function
produced by this model is practically indistinguishable from that with
no BH natal kicks.

The central panel of \Figref{fig:bh_kicks} shows the mass distribution
assuming no fallback downscaling, but a smaller kick amplitude
distribution characterized by $\sigma_\mathrm{kick}=100\,\mathrm{km \
  s^{-1}}$ for the BHs (while we keep the same fiducial assumptions
for NSs). Compared to our fiducial
simulation, BHs receive larger kicks, and thus the runaways produced
are also generally more massive, and the $90^\mathrm{th}$ percentile
of the mass distribution shifts to $\sim$$30\,M_\odot$.

The histogram in the top panel of \Figref{fig:bh_kicks} instead corresponds
to a scenario in which the natal kick is \emph{not} decreased by the amount of
fallback (e.g., if the natal kick is entirely due to asymmetries in the neutrino
emission). In this case BHs and NSs receive the same kick amplitude
(BH ``velocity kick'') and this results in a larger number of massive
runaways characterized by a mass function skewed towards larger
values. In this scenario, the $90^\mathrm{th}$ percentile of the mass
distribution is as high as $\sim$$50\,M_\odot$.

In summary, the tail of the mass function of massive runaways carries
information about the BH natal kicks: finding runaways of large masses
would support the possibility of large BH natal kicks. We do not
expect large contamination of such tail from dynamical ejection
of runaway stars, since dynamical channels tend to eject the least
massive star among those interacting \citep[although see
also][]{fujii:11, oh:16}.

Our result allow us to check also for a correlation between the mass
of the ejected companions and its velocity. \Figref{fig:M_vrw} shows
this comparison for the three BH kick variations of
\Figref{fig:bh_kicks}. As expected, there is a correlation between the
maximum velocity that can be achieved because of
the binary disruption, and the mass of the ejected MS companion. This
can be seen by the shape of the colors in each panel. More massive companions have more
inertia and are harder to accelerate, or, in other words, more massive
companion have slower pre-CC orbital velocity because of the mass
dependent factor in \Eqref{eq:v2}.

From left to right, the BH kick amplitude increases (while effectively
keeping the same NS kick, since the fallback fraction is small for
CCSNe resulting in NS formation), and so does the maximum mass of the
ejected companions. There is an sharp drop in the number of ejected
companions for $M_\mathrm{dis}\gtrsim70\,M_\odot$. This
arises because of the combination of several effects. First of all,
stars in this mass range are intrinsically rare because of the
IMF. Also, the lifetimes of very
massive stars become very similar ($\sim3\,\mathrm{Myr}$) above $\sim$$50\,M_\odot$. This
means that by the time the primary collapses, the secondary is also an
evolved star that has finished burning hydrogen in its core, and it
does not appear in the panels of \Figref{fig:M_vrw}. Finally,
these stars tend to have large wind mass loss rates, which can be
enhanced significantly when the star is spun up by the accretion: the
combination of the wind enhancement with already large wind mass loss
rates limits the growth in mass of the companions.

In an observed
population, the short post-ejection lifetime would also make very
massive ejected companion rare. However, the upperlimit of
$\sim$$70\,M_\odot$ is not strict, and it is possible to produce
walkaways in this mass range through binary interactions, although at
a negligible rate compared to less massive walkaways.

\subsection{Effects of metallicity }
\label{sec:Z}

We find that, other than the natal kick distribution, metallicity is the most important parameter
influencing our results. Decreasing the metallicity from super-solar,
$Z=0.03$, to $Z=0.02\simeq Z_\odot$ (fiducial run), to
$Z=0.008$, $Z=0.0047$ to $Z=0.0002\simeq10^{-2}Z_\odot$ a clear trend
emerges from \Tabref{tab:parameters}. At lower metallicity, the
fraction $\mathcal{D}$ of binaries disrupted by the first CC
is lower ($77\%$ for the lowest metallicity, cf. $88\%$ at the highest), but the stars ejected are on average
faster. This can be seen from the average velocity $\langle v\rangle$,
which goes from $21.6\,\mathrm{km \ s^{-1}}$ to $11.2\,\mathrm{km\ s^{-1}}$, but also from the progressive decrease of the ratio of walkaways per runaway
$\mathcal{R}$, regardless of the lower mass cut. Nevertheless,
walkaways outnumber runaways at all the metallicities we explore.

In our lowest metallicity run, the runaways travel, on average, as far as 706\,pc from their birth
location, and the walkaways reach, on average, 163\,pc. The mean
distance travelled by population of disrupted stars is 279\,pc for
the lowest metallicity.

The reduced likelihood of disrupting a system and increased velocity
of the stars ejected can be understood in
terms of the effect of metallicity on stellar radii. At lower $Z$
stars of a given mass and evolutionary stage have smaller radii. This means that a given
binary will enter in a mass transfer phase later, resulting in less conservative mass transfer and 
less orbital widening, making the binary harder to
disrupt and the pre-CC orbital velocity of the secondary
faster. This is corroborated by the fact that the average mass of
walkaways and runaways decreases slightly with metallicity, which is indicative of
a less conservative mass transfer. Moreover, at lower metallicity, the post-interaction orbital widening
due to stellar winds is less important.

Another important physical effect, presently not included in
our simulations, is that, at the end of the RLOF, the primary remains more massive at
lower metallicity \citep[owing again to the smaller radii, e.g.][]{gotberg:17,yoon:17}. This
should contribute to the reduction of orbital widening for interacting binaries,
and make the collapsing star more massive and thus the disruption harder.

\input{./Table_BHmid_kick.tex}

\begin{figure*}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{./figures/bound_vsys}
  \caption{Normalized distribution of systemic velocities of binaries
    with a MS star remaining bound after the CC of the first star. The
    distributions in both panels are normalized to unity. The left panel shows
    the distribution for NS companions, the right panel shows only
    binaries with a BH companion. The blue histogram
    represents our fiducial simulation. The orange solid line gives the
    distribution when the kick is not rescaled because of
    fallback, which changes the BH kicks, but not the NS kicks. The red distribution corresponds to a double Maxwellian
    kick distribution, which changes the NS kicks but not the BH kicks. The top panels show the corresponding cumulative distributions.}
  \label{fig:bound_v_dist}
\end{figure*}

\section{Systemic velocities of NS and BH binaries}
\label{sec:bound}

Our simulations also provide predictions for the systemic velocities of binary systems that remain bound after the first CC  \citep[e.g.,][]{vanoijen:89}. This concerns only \mbox{$1-\mathcal{D} = 14_{-10}^{14}\%$} of the systems (see \Secref{sec:param_variation} for a discussion of model variations).  These bound systems are of large interest since they can give rise to X-ray binaries if the compact object starts to accrete from the companion.  They typically originate from systems where the newly formed compact object received a small natal kick. This bias places  limitations on the extent to which these special systems can be used to infer conclusions about the properties of the general population of NS and BH.  For comparison, the average effective natal kick for the compact objects formed in systems that remain bound is  $\sim$$66\,\mathrm{km\ s^{-1}}$, much lower than the average of $\sim$$330\,\mathrm{km\ s^{-1}}$ (as drawn from our standard Maxwellian after rescaling to account for fallback).

Systems that remain bound also obtain a systemic velocity, because of
mass lost from the progenitor star during the supernova explosion
while orbiting around the center of mass.  We define the newly
obtained systemic velocity $v_\mathrm{sys}$ as the velocity of the new
center of mass (of the compact object plus the secondary star) in the
frame of the old center of mass (e.g.,
\citealt{brandt:95}, \citealt{kalogera:96}, \citetalias{tauris:98}).

Figure~\ref{fig:bound_v_dist} shows the normalized distribution of
systemic velocities $v_\mathrm{sys}$ for different natal kick
assumptions. The left (right) panel shows NS+MS (BH+MS) binaries, with
the cumulative distributions in the top panels, respectively.  We find that the distribution of systemic velocities are very insensitive to the model variations that we considered, except when adopting extreme assumptions for the natal kicks. We only show the distributions for variations that show large deviations from the fiducial simulation.

For NS+MS binaries  our fiducial simulation shows a median systemic velocity of $\sim 20\kms$, for nearly all assumptions that we consider. The strongest deviation is found with a run in which we adopt a double Maxwellian distribution, where low mass progenitors are assumed to form neutrons stars with very low kick velocities.  This substantially increases the fraction of systems that remain bound to 35\%. It primarily adds systems with low systemic velocities. The median of the normalized distribution for this simulation lies near $7 \kms$.

In our fiducial simulations for BH+MS binaries we obtain a bimodal
distribution with a large peak at $0 \kms$ accounting for nearly 80\%
of all systems and a second broader component peaking at $\sim
10\kms$.  The peak near zero results from progenitors that experience
complete fallback ($f_b=1$ according to the algorithm from
\citealt{fryer:12}). They receive no kick at all, and no mass is lost from
the system (we neglect here, as elsewhere, the mass possibly lost to
neutrinos during the CC). For these systems we also find no changes in
the separation and/or eccentricity after the CC of the first star.
The sharp peak at $v_\mathrm{sys}\simeq0\,\kms$ would likely be
smeared out by the velocity dispersion of the region where the system
formed.

Nearly all  variations considered give indistinguishable results. However, we find large deviations with the simulation where we used the extreme assumption that compact objects get kicks of a similar velocity amplitude regardless of their mass (no fall back down scaling, orange in \Figref{fig:bound_v_dist}). This effectively results in much larger kicks for black holes and is sometimes referred to as ``velocity kick''. Only $1-\mathcal{D}\simeq3\%$  of systems remain bound in this simulation.  The few resulting BH binaries typically have large systemic velocities: we find a broad distribution with a median at $50\kms$.

Our simulations also provide predictions for the distributions of further properties of the bound systems.  
If we exclude systems with $f_b=1$, i.e.\ if we consider only NSs and
BHs formed via fallback but still producing ejecta, we find
an anti-correlation between the maximum post-CC separation and systemic
velocity: the wider the system, the slower the maximum systemic
velocity it can reach \citep[see also][]{brandt:95}.

Most systems remaining bound come from initially 
relatively tight pre-CC orbits, with typical periods $\log_{10}(P_\mathrm{pre-CC}/\mathrm{days})\simeq -2$. Typically, these very close systems have a reversed mass ratio
(i.e., at the time of the first CC $M_2^\mathrm{pre-CC}>M_1^\mathrm{pre-CC}$)
because of previous mass transfer phase. In our fiducial run, for the
systems remaining bound the average pre-CC mass ratio is
$\langle q_\mathrm{pre-CC}\rangle\equiv\langle
M_2^\mathrm{pre-CC}/M_1^\mathrm{pre-CC}\rangle\simeq 2.9$.
The majority of the bound systems, about $60\%$ for all variations
with non-zero natal kicks, have separation of less than
about $500\,R_\odot$. This suggests that these systems will likely
evolve through a phase during which they might be detectable as X-ray
sources, and possibly even through a mass transfer phase. The tail of the distribution at large separations can reach
post-CC separation as large as $\sim6000\,R_\odot$.

We do not consider cases in which the compact object is shot within the
Roche lobe of the secondary. A prompt collision between the
newly formed compact object and the companion star could lead to the
formation of a Thore-Zytkow object \citep[if the compact object is a
NS,][]{thorne:75,thorne:77, leonard:94}, or some other transient
event, possibly involving the disintegration of the companion. For a given binary system, the probability of a prompt collision
caused by the natal kick (whose direction is isotropically distributed
in the frame of the collapsing star) can be estimated as the solid
angle subtended by the companion at the position of the collapsing
star. For our fiducial population, excluding systems that do not
receive a natal kick because of the fallback rescaling, we obtain that
roughly 1 out of 10\,000 CC events in a binary would result in a
prompt collision (i.e., a probability per binary system of about
$10^{-4}$). We exclude from the normalization systems merging before the first CC.

\section{Astrophysical Implications}
\label{sec:discussion}

\subsection{How to identify walkaway stars}

The velocity dispersion of OB stars in our galaxy is
$\lesssim$10\,$\mathrm{km\ s^{-1}}$ (e.g.,
\citealt{blaauw:56}; \citealt{gies:87}; \citetalias{tauris:98};
\citealt{hoogerwerf:01}), and it is typically lower for OB
associations 
\citep[][]{debruijne:99,steenbrugge:03, kiminki:18}. We emphasize also that the velocity dispersion
increases going from samples of OB stars in clusters, in associations,
and in the field \citep[][]{gies:87}. The average velocity of ejected
secondaries is slightly higher than this value
(cf.~\Tabref{tab:parameters}), and if considering only the
post-binary interaction secondaries (i.e.,~removing all the ejected
secondaries from very wide and non-interacting binaries, cf.~dashed
distributions in \Figref{fig:v_dist}) it increases slightly,
especially for $M_\mathrm{dis}\geq7.5\,M_\odot$. Therefore, the
companions ejected after a mass transfer episode might form a distinct
population in the Gaia data. Moreover, regardless of their previous history, the walkaway
and runaway stars from binary disruptions do not need to follow the
local galactic rotation curve, and their motion can bring them in
locations of lower extinction making them easier to see
\citep[e.g.,][]{boubert:18,maiz-appellaniz:18}.

\citet{tetzlaff:11}, following \cite{stone:91}, propose to fit the
velocity distribution of young ($\lesssim50\,\mathrm{Myr}$) stars
within 3\,Kpc of the Solar neighborhood using
two Maxwellian: one for the ``low-velocity'' ($v\lesssim30\,\mathrm{km\
  s^{-1}}$) and one for the ``high-velocity'' group. Our findings
suggest that $\sim$$10\%$ of the O-stars in the low velocity group might still be the
result of binary disruptions.

Because of the binary interactions taking
place before the first CC, these stars carry observational signatures which might
make them recognizable even if they do not stand out from a kinematic
point of view. For instance, if there is a mass transfer phase during
the previous binary evolution, the ejected star will be spun up
\citep[e.g.,][]{packet:81, pols:91,boubert:18} and
possibly chemically polluted with He-rich and/or N-rich material \citep[e.g.,][]{blaauw:93}. Mass transfer
might also be responsible for the presence of magnetic fields,
\citep[e.g.][]{schneider:16}. These features should in principle leave
observable signatures in the spectra of these stars \citep[e.g.,][]{maiz-appellaniz:18}. However, the
effects of the metallicity gradient in the radial direction of the
galaxy complicates the spectral identification of binary products, as pointed out by \cite{mcevoy:17}.

Characterizing the population of stars ejected from isolated binary
disruptions is necessary to reduce the number of false positive detections of hypervelocity stars in current and
upcoming astrometric catalogues \citep[e.g.,][]{marchetti:17b}.

Stars ejected by a successful SN explosion recently (i.e., less time
ago than the lifetime of the SN remnant) can be connected to the SN
location, and the associated NS can sometimes also be found. Assuming
a typical visible lifetime of the SN remnant of
$\tau_\mathrm{SNR} \simeq10^4$\,years, the distance traveled by
an ejected companion is of order $D\equiv
v_\mathrm{dis}\times\tau_\mathrm{SNR}\simeq
0.1\,\mathrm{pc}$. Therefore runaways, but especially walkaways
are expected to
still reside within the SN remnant while it is still observable. This
strategy has been successfully applied to search for runaways by
\cite{vandenbergh:80,guseinov:05,tetzlaff:13,tetzlaff:14,dincel:15,boubert:17a}. Recently,
\cite{kerzendorf:17} also used the predicted kinematics of disrupted
binaries to search for companions of the star that exploded producing
the supernova remnant Cas A. \cite{tetzlaff:11} also suggested that
with precise astrometry, it might be possible to relate the ejected
companion to the remnant of the companion up to $\sim$5\,Myr after the
SN explosion, provided that the compact remnant of the former companion is visible (e.g., as a pulsar).

The results of \cite{banerjee:12} and \cite{perets:12} suggest there
might be a spatial distinction between dynamically ejected stars and
post-binary runaways and walkaways. Dynamical interactions in a
cluster are most efficient early in the cluster evolution \citep[e.g.,][]{oh:16},
well before the first stellar CC event happens. Moreover, the median velocity
of cluster ejection is higher than the typical walkaway
velocity: \cite{banerjee:12} derived 
a median velocity of dynamically ejected stars of
$50\,\mathrm{km\ s^{-1}}$ after 3\,Myr of evolution. This value
 is significantly higher than what we find from binary disruptions. 
The combination of these two effects would suggest that
walkaways and runaways from binary disruptions would generally be closer to their parent cluster than
dynamically ejected stars. However, this simplistic prediction is
complicated by the fact that binaries can eject stars from the parent
cluster outskirts, while most dynamical interactions produce runaways
from the center of the cluster. Another complication is the possible
ejection of stars as tidal tales in a cluster merger process, \citep[e.g.,][]{lucas:17}.

\cite{evans:15} and \cite{neugent:18} have reported the first
discovery of extragalactic post-MS massive runaways. They reported the
observation of a red super giant in M31 and a yellow super giant in the SMC,
respectively. They inferred peculiar space velocities larger than
$\sim$$400\,\mathrm{km\ s^{-1}}$ and $\sim$$150\,\mathrm{km\ s^{-1}}$ for
these stars, and proposed a binary origin at least for the yellow
supergiant in the SMC. Our fiducial population can hardly reach such
high peculiar velocities, unless the velocity of the system as a whole binary itself was already 
high. Even if these stars were ejected from a binary during their main
sequence, velocities in excess of a few hundred $\mathrm{km\ s^{-1}}$
are only marginally reached in our simulations
(cf.~\Figref{fig:v_dist},~\Figref{fig:v2_postMS}, and
\Figref{fig:v_dist_log}), and this might suggest a different ejection
mechanism (or combination of mechanisms) for these two stars. \cite{evans:15} also report that
vast majority of O/B stars and red super giant stars in the Milky-Way have radial
velocities lower than $\sim70\,\mathrm{km\ s^{-1}}$, in very good
agreement with our results.

\subsection{On the runaway fraction}
\label{sec:howtofind}
Our results predict a runaway fraction among O type stars of at best a
few percent, in agreement with \cite{eldridge:11}, but in tension with
the observational result that 
$\sim$$10-20\%$ of O type stars are runaways \citep[][Sana et al., in
preparation]{blaauw:61, gies:87, stone:91, tetzlaff:11, maiz-appellaniz:18}, and the claim
that roughly two-thirds of runaways come from the
disruption of binaries \citep[][]{hoogerwerf:01}. The latter has also been
challenged by the more recent observations from \cite{jilinski:10}.
Our results are roughly consistent with the following back of the
envelope estimate for the runaway fraction:
\begin{equation}
  \label{eq:ratios}
  f^\mathrm{RW} \simeq
  f_\mathrm{bin}\times(1-f_\mathrm{mergers})\times\mathcal{D}\times
  (1+f_\mathrm{acc})^{\alpha}\times f_{v>30} \ \ ,  %% the time is in
                                %% the 1+f_acc factor!!
\end{equation}
where we can assume a binary fraction $f_\mathrm{bin}=1$, reasonable
for the population of O type stars considering initial periods as
large as $10^{5.5}$\,days, the fraction of binaries that do not merge is
$1-f_\mathrm{mergers}\simeq0.8$ \citep[][]{sana:12}, the disruption
fraction is $\mathcal{D}\simeq0.8$ \citep[this study, but also,
e.g.,][]{eldridge:11}, the fraction of material accreted, which is not
likely to exceed $f_\mathrm{acc}\simeq 0.2$ \citep[][]{packet:81}, and the
fraction of ejected companions fast enough to be a runaway which we
can estimate\footnote{We neglect here that $\mathcal{D}$ is the
  disruption fraction including also binaries with no MS stars at the
  pre-CC stage} from $\mathcal{D}$ and $\mathcal{R}_{15}$, and is
$f_{v>30}\simeq \mathcal{D}/(1+\mathcal{R}_{15})\simeq0.04$ for our fiducial run. Assuming an IMF slope $\alpha=-2.3$ we obtain a
runaway fraction of $f^\mathrm{RW}\simeq 0.01$, in reasonable
agreement with our results.

The tension between our results and the observed runaway fraction suggests that either (i)
binaries tend to evolve towards shorter pre-CC orbits, corresponding
to higher ejection velocities, but are still disrupted easily by the
CC event, (ii) the contribution of dynamical ejections to the runaway
population is presently underestimated, (iii) other ejection
mechanisms exists and are presently overlooked, (iv) the velocities of
massive runaways are presently overestimated in observations (e.g.,
because of observational bias which hide more easily slow moving walkaways), or any
combination of the previous.

Since
$v_\mathrm{dis}\simeq v_2^\mathrm{pre-CC}\propto
v_\mathrm{orb}^\mathrm{pre-CC}\propto J_\mathrm{orb}$, with
$J_\mathrm{orb}$ orbital angular momentum, to shift the
peak of the velocity distribution in \Figref{fig:v_dist} from
$\sim$$6\,\mathrm{km\ s^{-1}}$ to
$v_\mathrm{dis}\gtrsim30\,\mathrm{km\ s^{-1}}$, the binaries would
need to lose about five times more orbital angular momentum during the evolution
before the collapse of the primary. This discrepancy
deserves to be revisited once the runaway fraction can be evaluated from the
homogeneous Gaia datasets. If the contribution from
dynamical ejections can robustly be quantified, the 
population of runaways might provide new constraints on the mass
and angular momentum losses during mass transfer in
binaries.

A mechanism to remove angular momentum and mass from the
binary presently not included in our population are eruptive mass loss
events for which there is growing evidence from early observations of
SNe \citep[e.g.][]{khazov:16}. Such mass loss events might be much
more common than previously thought. One possible physical cause for
these event are gravity waves
excited by late shell burning depositing energy at the base of the
envelope \citep[][]{quataert:12, fuller:17, fuller:18}. It is possible
that the mechanism driving these mass loss events might also result in
pre-CC binary interactions which might significantly change the pre-CC
orbit. SN-impostor events happening earlier in the evolution might
also lead to similar effects on the orbital evolution, which we
presently do not model.

Another possibility to increase the angular momentum losses could be
magnetic braking. If non conservative mass transfer can simultaneously
generate a magnetic field and eject some mass from the system, this
might result in torques on the binary much larger than we consider here.

\subsection{X-ray binaries and gravitational wave sources}

Binaries remaining bound after the CC of the primary are a minority.

Selecting from our results only bound systems consisting of a BH and a
MS star, we obtain the systemic velocity distributions for BH X-ray
binaries with different BH kick scenarios (see right panel of
\Figref{fig:bound_v_dist}). Gaia will give astrometric constraints on the 19 known BH X-ray binary in the Galaxy: if the
effect of the Galactic potential can be singled out from these data,
the comparison with our distribution might shed light onto the typical
amplitude of BH kicks \citep[e.g.,][]{fragos:09,repetto:12,Fragos+2013a,mandel:16}.

The binaries surviving the first CC can
become X-ray sources if their separation is
short enough for the compact object to accrete during the subsequent evolution. For all our natal kick
assumptions, the post-CC separation peaks at $\sim300\,R_\odot$,
and most bound system have separations smaller than
$\sim500\,R_\odot$, indicating that the majority of the systems
remaining bound will become X-ray sources, and potentially even go
through a common envelope evolution phase because of the large mass
ratio between the compact object and the secondary star (which has
accreted mass, cf.~\Secref{sec:example}).

However, observed
X-ray sources are necessarily on relatively short period orbits, or, in
other words, the energy released at the CC of the first star did not
widen significantly the binary and instead contributed mostly to the
kinetic energy of the system. This could mean that the observed sample
of galactic BH X-ray binaries is biased towards the high velocity tail
of our distributions.

System that might become X-ray sources have systemic velocities of a few tens of
$\mathrm{km\ s^{-1}}$, and since the secondary star is also an
accretor and rejuvenated, they are able to travel a
distance comparable to $\langle L_\mathrm{walk}\rangle$,
which can also increase their contribution as feedback engines
\citep[e.g.,][]{justham:12,Fragos+2013a}.

If a bound system is not disrupted by the first CC and does not merge
during the evolution of the secondary star, it is less likely to be
disrupted by the second SN \citep[e.g.][]{stevenson:17,oshaughnessy:17}. This is because the impact of the kick on the orbit
scales as $v_k/v_\mathrm{orb}$, and the orbit of bound systems is much
more likely to shrink during the evolution after the first CC: the
donor star needs to become less massive than the compact object for
the system to widen as described in
\Secref{sec:example} and \Secref{sec:analytics}. Consequently, it is likely that most of the
systems remaining bound after the first CC will also remain bound as
double compact objects (and potential gravitational wave sources if
the separation is small enough).

\subsection{Runaways and walkaways as feedback engines}
\label{sec:disrupted}

The large majority of binary systems is disrupted at the first CC
(with a few notable exceptions, see also \Secref{sec:param_variation}):
systems that can become X-ray sources and/or gravitational wave
sources are the exception rather than the rule. The constraints on binary evolution coming from the much more
common evolutionary path should be considered together with those
coming from the more rare channels.

Binary interactions increase the lifetime of the ejected star because
of rejuvenation: both walkaways
and runaways can reach significant distances from their birth
location, see \Tabref{tab:parameters} and \Figref{fig:Dhist}. The typical distances traveled
by the ejected secondaries (cf.~\Figref{fig:Dhist}) are sufficient to
get massive stars out of their parent cluster or association, but only
the fastest runaways would get beyond the thick disk in our galaxy,
especially if considering the relative inclination between the
trajectory of the star and the disk itself. As noted by
\cite{boubert:18}, this introduces a bias in observed samples that
favors the detection of the fastest runaways.

The motion of the ejected stars
effectively spreads the massive stars and their (possible) SN explosions in a larger
volume, possibly enhancing their impact on the composition of galaxies (more efficient mixing of
nuclear yields), on the ionization of their surroundings (enhanced
escape fraction, e.g.~\citealt{conroy:12, kimm:14, ma:16}), but it might possibly
decrease the efficiency of their mechanical feedback because of the
lower density of the surrounding gas.
The large number of walkaways
formed per each runaway $\mathcal{R}$ suggests that walkaway stars might
enhance massive stars feedback more significantly than the faster runaways, although they cannot
reach as far out as the latter because of their lower velocity.

As a consequence of the typical evolutionary scenario
described in \Secref{sec:example}, the first CC event in a
binary system will result either in a stripped SN Ib/Ic or a failed SN
with little or no electromagnetic signature. The second CC will
instead be from the accretor star, which could again result in a
SNIb/Ic (if the star becomes a Wolf-Rayet because of its own stellar wind),
but, because of the IMF and the possibly non-fully conservative
accretion rate, the secondary is more likely to remain H-rich until
its CC. As already pointed out by
\cite{eldridge:11}, this suggests that SNIb/Ic should preferentially
happen in star forming regions, while type II SN should show a larger
spatial spread (because of the motion of the ejected star).

In this study, we did not consider the effects of the background
cluster and/or galactic potential on the motion of the ejected
secondaries. These can in principle be used as test particles to probe
the potential itself, and consequently allow for testing the local
and/or galactic dark matter distribution \citep[e.g.,][]{rossi:14,marchetti:17}.

While runaways have peculiar kinematic %(and possibly spectal)
characteristics that make them recognizable, walkaway stars might more
easily be mistaken for genuine single stars. Especially masses larger
than $7.5\,M_\odot$ and velocities larger than $v_\mathrm{dis}\gtrsim20\,\mathrm{km\
  s^{-1}}$, the walkaways and runaways accrete mass before being shot
out. Therefore, the disruption
of binaries can
pollute observed samples of
present-day single stars with binary evolution by-products, by populating the field of the galaxy
\citep[][]{gvaramadze:12,demink:14}.

As an example, the runaway star $\zeta$ Puppis has been considered to be
a ``canonical'' O-type star \citep[e.g.][]{ramiaramanantsoa:17}, and
it was used to calibrate the free parameters in the \cite{castor:75}
(CAK) theory of stellar winds \citep[][]{pauldrach:94} widely applied
to model single stars. However, this object is
likely to be a accretor ejected from a binary, \cite{vanrensbergen:96,ramiaramanantsoa:17}.

Massive walkaway ejected from a binary might also appear as isolated
massive stars. For massive stars observed in isolation, the question
of their origin poses: did they form in isolation? or could they
possibly have reached that location, if ejected from a binary? \cite{bestenlehner:11} identify an isolated
$\sim150\,M_\odot$ star (VFTS682) in the 30 Doradus field. Such
massive object is unlikely to be the result of a binary disruption,
but might have been ejected from the massive young cluster
R136. \cite{bestenlehner:11} suggest that if this star has not been
ejected from this cluster, it might be a direct indication of isolated
star formation. Walkaway massive stars can pose similar problems if
not recognized as rejuvenated binary products.

This might also connect to the evolutionary path of luminous blue
variable (LBV) stars. These are massive stars which experience
extreme and yet unexplained outbursts of mass loss. In the classical
picture of massive stellar evolution (so-called ``Conti scenario''),
LBVs are an intermediate evolutionary stage between O-type and Wolf-Rayet
stars, during which the bulk of the H-rich envelope is lost \citep[e.g.][]{conti:75,
  maeder:94,maeder:96}. However, \cite{smith:15}
pointed out that LBV and LBV-candidates are farther from the nearest
O-type star than Wolf-Rayet stars, and suggested that LBVs might be rejuvenated products of binary
evolution\footnote{This idea has also be questioned by
  \cite{humphreys:16,davidson:16}, see also \cite{smith:16}.} (accretors or merger
products) with an increased lifetime to reach the observed isolation. The followup study of \cite{aghakhanlootakanloo:17}
suggested that partial rejuvenation during mass transfer coupled to a walkaway (or
runaway) velocity might be sufficient to explain the isolation
observed by \cite{smith:15}. Alternatively, fully rejuvenated mergers
might reach isolation just because of their velocity from the local dispersion.

\section{Summary and conclusion}
\label{sec:conclusion}

In anticipation of the new data releases by the Gaia satellite we have
carried out an extensive suite of numerical simulations to provide
predictions for the kinematics and properties of massive stars ejected
from binary systems disrupted by the supernova explosion of the
companion star. Our aim with this study is three-fold:  (i) identify
which theoretical predictions are robust against model uncertainties
and discuss their astrophysical implications (ii)  investigate
theoretically which uncertain physical processes affect the kinematics
of the population of unbound stars in such a way that their imprints
would be observable and  (iii) provide a framework of models that can
be used for comparison against observations or as input for
other simulations.

We summarize our main findings below. When we quote percentages or fractions, the error bars given indicate the maximum variations that we encountered in our model variations for different physical assumptions (excluding the unphysical variation where we ignored all natal kicks for newly born compact objects).

\begin{itemize}

\item[$\bullet$] Nearly a quarter,   $22_{-8}^{+26}$\%, of all binary
  systems with at least one star more massive initially than
  $7.5\Msun$ merges prior to the explosion of the first star (see \Figref{fig:tree} and \Tabref{tab:parameters}).

\item[$\bullet$] The large majority of binary systems is disrupted at the moment of core collapse of the first star, $\mathcal{D}=86_{-9}^{+11}\%$ in our simulations, consistent with earlier studies. This prediction is robust against variations in the treatment of the evolutionary and interaction processes, but it is sensitive to the choices concerning the natal kicks of neutron stars and black holes.

\item[$\bullet$] The disruption of the binary system produces an
  unbound stellar companion. Remarkably, we find that these events
  only rarely produce runaway stars (i.e. stars with peculiar motions
  larger than $30\,\kms$).  The velocity distribution peaks at
  $\sim6\,\mathrm{km\  s^{-1}}$, with a $90^\mathrm{th}$ percentile
  around $20\kms$ well below the typical threshold for runaway stars.

\item[$\bullet$] The slow moving unbound companions, to which we refer
  as walkaway stars, outnumber the runaway stars by at least an order
  of magnitude in nearly all our models. This is a robust outcome of
  the models and is primarily due to the reversal of the mass ratio
  prior to the explosion as we show numerically and analytically.
  This finding has been noted in earlier studies \citep[e.g., ][]{dedonder:97,eldridge:11},  but we believe this results (and its potential implications) are under appreciated.

\item[$\bullet$] We also produce runaway stars but we rarely find them
  to be faster than $60\,\kms$, which corresponds to
  $98.9^\mathrm{th}$ of the population. This is much lower than the
  theoretical maximum set by the finite size of the stars and their
  orbit, which is in range 400-1000$\,\kms$ depending on their masses
  and evolutionary state. We expect that stars with larger peculiar
  velocities result from dynamical ejections from clusters, capture
  from another galaxy, or most exotically, interaction with a
  intermediate mass or supermassive black hole.

\item[$\bullet$] We find a mild trend for runaway stars to be faster
  and more common at lower metallicity, caused by the fact that stars are on average smaller and the reduced effects of stellar winds.   Gaia should in principle be able to test these predictions by identifying the fastest runaway stars in the Magellanic clouds. In our most metal poor simulations for $Z=0.0002$, which may have relevance for the earliest stellar populations, we find 5 times more runaway stars, if we keep all other assumptions for the initial distributions the same.

\item[$\bullet$] Both runaways and walkaways typically accrete mass from their
companion prior to the disruption, This is especially true for those more massive than
$7.5\,M_\odot$ at the time they are ejected. They typically also gain angular momentum in the process. These features might make walkaway stars recognizable as binary product even if they do not stand out from a kinematic point of view. Their further evolution and structure may also therefore differ from that of non rotating single stars, as has been speculated in the context of the LBV phenomenon.

\item[$\bullet$] We suggest that the high mass tail of the runaway
  mass function might provide insight on the formation mechanism of
  black holes from the core-collapse of the primary star, and in particular on whether these receive a ``velocity'' or a ``momentum''
kick. Studies so far focussed on the surviving X-ray binaries seeking for constraints, but we propose to use their unbound counter parts since they constitute the majority and should be identified by Gaia.

\item[$\bullet$] By ejecting secondaries, the disruption of binaries can enhance the role of
massive stars as feedback engines. The average distance they travel,
calculated neglecting any external potential, is of order
$\gtrsim100$\,pc, and up to $\gtrsim 500$\,pc if considering only the
faster runaways. These suggests that both runaways and walkaways
can exit the dust cloud embedding their birth location, thus enhancing
the escape fraction for their ionizing radiation. If they are massive
enough, the ejected star will explode far from other stars and gas
overdensities, with potential implications for the driving of
turbulence and star formation.

\end{itemize}

Our simulations will be made available upon final publication of this study.  Early requests for access can be addressed to the lead author.

\begin{acknowledgements}
  We acknowledge for the many helpful discussions
  F.~Broekgaarden, A~.de~Koter, D.~Hendriks, L.~Kaper, N.~Langer, I.~Mandel, C.~J.~Neijssel, Ph.~Podsiadlowski,
  K.~Sakar, F.~N.~Schneider, S.~N.~Shore, E.~van~den~Heuvel, A.~Vigna-Gomez.
  MR is grateful to the KITP for the visiting graduate fellowship during which a large portion of this
  study has been carried out. This research was supported in part by
  the National Science Foundation under Grant No. NSF PHY11-25915. SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
Council (ERC) (Grant agreement No. 715063).
  RGI thanks the Science and Technology Facilities Council (STFC) for funding his Rutherford fellowship under grant ST/L003910/1.

\end{acknowledgements}

\bibliographystyle{aa}
%\bibliography{bibliography/walkaway,bibliography/selma_bib}
\begin{thebibliography}{242}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{Abbott {et~al.}(2016)Abbott, Abbott, Abbott, Abernathy, Acernese,
  Ackley, Adams, Adams, Addesso, Adhikari, Adya, Affeldt, Agathos, Agatsuma,
  Aggarwal, Aguiar, Aiello, Ain, Ajith, Allen, Allocca, Altin, Anderson,
  Anderson, Arai, Araya, Arceneaux, Areeda, Arnaud, Arun, Ascenzi, Ashton, Ast,
  Aston, Astone, Aufmuth, Aulbert, Babak, Bacon, Bader, Baker, Baldaccini,
  Ballardin, Ballmer, Barayoga, Barclay, Barish, Barker, Barone, Barr,
  Barsotti, Barsuglia, Barta, Bartlett, Bartos, Bassiri, Basti, Batch, Baune,
  Bavigadda, Bazzan, Behnke, Bejger, Belczynski, Bell, Bell, Berger, Bergman,
  Bergmann, Berry, Bersanetti, Bertolini, Betzwieser, Bhagwat, Bhandare,
  Bilenko, Billingsley, Birch, Birney, Biscans, Bisht, Bitossi, Biwer,
  Bizouard, Blackburn, Blair, Blair, Blair, Bloemen, Bock, Bodiya, Boer,
  Bogaert, Bogan, Bohe, Bojtos, Bond, Bondu, Bonnand, Boom, Bork, Boschi, Bose,
  Bouffanais, Bozzi, Bradaschia, Brady, Braginsky, Branchesi, Brau, Briant,
  Brillet, Brinkmann, Brisson, Brockill, Brooks, Brown, Brown, Brown, Buchanan,
  Buikema, Bulik, Bulten, Buonanno, Buskulic, Buy, Byer, Cadonati, Cagnoli,
  Cahillane, Bustillo, Callister, Calloni, Camp, Cannon, Cao, Capano, Capocasa,
  Carbognani, Caride, Diaz, Casentini, Caudill, Cavaglià, Cavalier, Cavalieri,
  Cella, Cepeda, Baiardi, Cerretani, Cesarini, Chakraborty, Chalermsongsak,
  Chamberlin, Chan, Chao, Charlton, Chassande-Mottin, Chen, Chen, Cheng,
  Chincarini, Chiummo, Cho, Cho, Chow, Christensen, Chu, Chua, Chung, Ciani,
  Clara, Clark, Cleva, Coccia, Cohadon, Colla, Collette, Cominsky, Jr., Conte,
  Conti, Cook, Corbitt, Cornish, Corsi, Cortese, Costa, Coughlin, Coughlin,
  Coulon, Countryman, Couvares, Cowan, Coward, Cowart, Coyne, Coyne, Craig,
  Creighton, Cripe, Crowder, Cumming, Cunningham, Cuoco, Canton, Danilishin,
  D’Antonio, Danzmann, Darman, Dattilo, Dave, Daveloza, Davier, Davies, Daw,
  Day, DeBra, Debreczeni, Degallaix, Laurentis, Deléglise, Pozzo, Denker,
  Dent, Dereli, Dergachev, DeRosa, DeRosa, DeSalvo, Dhurandhar, Díaz, Fiore,
  Giovanni, Lieto, Pace, Palma, Virgilio, Dojcinoski, Dolique, Donovan, Dooley,
  Doravari, Douglas, Downes, Drago, Drever, Driggers, Du, Ducrot, Dwyer, Edo,
  Edwards, Effler, Eggenstein, Ehrens, Eichholz, Eikenberry, Engels, Essick,
  Etzel, Evans, Evans, Everett, Factourovich, Fafone, Fair, Fairhurst, Fan,
  Fang, Farinon, Farr, Farr, Favata, Fays, Fehrmann, Fejer, Ferrante, Ferreira,
  Ferrini, Fidecaro, Fiori, Fiorucci, Fisher, Flaminio, Fletcher, Fournier,
  Franco, Frasca, Frasconi, Frei, Freise, Frey, Frey, Fricke, Fritschel,
  Frolov, Fulda, Fyffe, Gabbard, Gair, Gammaitoni, Gaonkar, Garufi, Gatto,
  Gaur, Gehrels, Gemme, Gendre, Genin, Gennai, George, Gergely, Germain, Ghosh,
  Ghosh, Giaime, Giardina, Giazotto, Gill, Glaefke, Goetz, Goetz, Gondan,
  González, Castro, Gopakumar, Gordon, Gorodetsky, Gossan, Gosselin, Gouaty,
  Graef, Graff, Granata, Grant, Gras, Gray, Greco, Green, Groot, Grote,
  Grunewald, Guidi, Guo, Gupta, Gupta, Gushwa, Gustafson, Gustafson, Hacker,
  Hall, Hall, Hammond, Haney, Hanke, Hanks, Hanna, Hannam, Hanson, Hardwick,
  Harms, Harry, Harry, Hart, Hartman, Haster, Haughian, Heidmann, Heintze,
  Heitmann, Hello, Hemming, Hendry, Heng, Hennig, Heptonstall, Heurs, Hild,
  Hoak, Hodge, Hofman, Hollitt, Holt, Holz, Hopkins, Hosken, Hough, Houston,
  Howell, Hu, Huang, Huerta, Huet, Hughey, Husa, Huttner, Huynh-Dinh, Idrisy,
  Indik, Ingram, Inta, Isa, Isac, Isi, Islas, Isogai, Iyer, Izumi, Jacqmin,
  Jang, Jani, Jaranowski, Jawahar, Jiménez-Forteza, Johnson, Jones, Jones,
  Jonker, Ju, K, Kalaghatgi, Kalogera, Kandhasamy, Kang, Kanner, Karki,
  Kasprzack, Katsavounidis, Katzman, Kaufer, Kaur, Kawabe, Kawazoe, Kéfélian,
  Kehl, Keitel, Kelley, Kells, Kennedy, Key, Khalaidovski, Khalili, Khan, Khan,
  Khan, Khazanov, Kijbunchoo, Kim, Kim, Kim, Kim, Kim, Kim, King, King, Kinzel,
  Kissel, Kleybolte, Klimenko, Koehlenbeck, Kokeyama, Koley, Kondrashov,
  Kontos, Korobko, Korth, Kowalska, Kozak, Kringel, Krishnan, Królak, Krueger,
  Kuehn, Kumar, Kuo, Kutynia, Lackey, Landry, Lange, Lantz, Lasky, Lazzarini,
  Lazzaro, Leaci, Leavey, Lebigot, Lee, Lee, Lee, Lee, Lenon, Leonardi, Leong,
  Leroy, Letendre, Levin, Levine, Li, Libson, Littenberg, Lockerbie, Logue,
  Lombardi, Lord, Lorenzini, Loriette, Lormand, Losurdo, Lough, Lück,
  Lundgren, Luo, Lynch, Ma, MacDonald, Machenschalk, MacInnis, Macleod,
  Magaña-Sandoval, Magee, Mageswaran, Majorana, Maksimovic, Malvezzi, Man,
  Mandel, Mandic, Mangano, Mansell, Manske, Mantovani, Marchesoni, Marion,
  Márka, Márka, Markosyan, Maros, Martelli, Martellini, Martin, Martin,
  Martynov, Marx, Mason, Masserot, Massinger, Masso-Reid, Matichard, Matone,
  Mavalvala, Mazumder, Mazzolo, McCarthy, McClelland, McCormick, McGuire,
  McIntyre, McIver, McManus, McWilliams, Meacher, Meadors, Meidam, Melatos,
  Mendell, Mendoza-Gandara, Mercer, Merilh, Merzougui, Meshkov, Messenger,
  Messick, Meyers, Mezzani, Miao, Michel, Middleton, Mikhailov, Milano, Miller,
  Millhouse, Minenkov, Ming, Mirshekari, Mishra, Mitra, Mitrofanov,
  Mitselmakher, Mittleman, Moggi, Mohan, Mohapatra, Montani, Moore, Moore,
  Moraru, Moreno, Morriss, Mossavi, Mours, Mow-Lowry, Mueller, Mueller, Muir,
  Mukherjee, Mukherjee, Mukherjee, Mukund, Mullavey, Munch, Murphy, Murray,
  Mytidis, Nardecchia, Naticchioni, Nayak, Necula, Nedkova, Nelemans, Neri,
  Neunzert, Newton, Nguyen, Nielsen, Nissanke, Nitz, Nocera, Nolting,
  Normandin, Nuttall, Oberling, Ochsner, O’Dell, Oelker, Ogin, Oh, Oh, Ohme,
  Oliver, Oppermann, Oram, O’Reilly, O’Shaughnessy, Ott, Ottaway, Ottens,
  Overmier, Owen, Pai, Pai, Palamos, Palashov, Palomba, Pal-Singh, Pan, Pankow,
  Pannarale, Pant, Paoletti, Paoli, Papa, Paris, Parker, Pascucci, Pasqualetti,
  Passaquieti, Passuello, Patricelli, Patrick, Pearlstone, Pedraza, Pedurand,
  Pekowsky, Pele, Penn, Perreca, Phelps, Piccinni, Pichot, Piergiovanni,
  Pierro, Pillant, Pinard, Pinto, Pitkin, Poggiani, Popolizio, Post, Powell,
  Prasad, Predoi, Premachandra, Prestegard, Price, Prijatelj, Principe,
  Privitera, Prix, Prodi, Prokhorov, Puncken, Punturo, Puppo, Pürrer, Qi, Qin,
  Quetschke, Quintero, Quitzow-James, Raab, Rabeling, Radkins, Raffai, Raja,
  Rakhmanov, Rapagnani, Raymond, Razzano, Re, Read, Reed, Regimbau, Rei, Reid,
  Reitze, Rew, Reyes, Ricci, Riles, Robertson, Robie, Robinet, Rocchi, Rolland,
  Rollins, Roma, Romano, Romano, Romanov, Romie, Rosińska, Rowan, Rüdiger,
  Ruggi, Ryan, Sachdev, Sadecki, Sadeghian, Salconi, Saleem, Salemi, Samajdar,
  Sammut, Sanchez, Sandberg, Sandeen, Sanders, Sassolas, Sathyaprakash,
  Saulson, Sauter, Savage, Sawadsky, Schale, Schilling, Schmidt, Schmidt,
  Schnabel, Schofield, Schönbeck, Schreiber, Schuette, Schutz, Scott, Scott,
  Sellers, Sentenac, Sequino, Sergeev, Serna, Setyawati, Sevigny, Shaddock,
  Shah, Shahriar, Shaltev, Shao, Shapiro, Shawhan, Sheperd, Shoemaker,
  Shoemaker, Siellez, Siemens, Sigg, Silva, Simakov, Singer, Singer, Singh,
  Singh, Singhal, Sintes, Slagmolen, Smith, Smith, Smith, Son, Sorazu,
  Sorrentino, Souradeep, Srivastava, Staley, Steinke, Steinlechner,
  Steinlechner, Steinmeyer, Stephens, Stevenson, Stone, Strain, Straniero,
  Stratta, Strauss, Strigin, Sturani, Stuver, Summerscales, Sun, Sutton,
  Swinkels, Szczepańczyk, Tacca, Talukder, Tanner, Tápai, Tarabrin,
  Taracchini, Taylor, Theeg, Thirugnanasambandam, Thomas, Thomas, Thomas,
  Thorne, Thorne, Thrane, Tiwari, Tiwari, Tokmakov, Tomlinson, Tonelli, Torres,
  Torrie, Töyrä, Travasso, Traylor, Trifirò, Tringali, Trozzo, Tse, Turconi,
  Tuyenbayev, Ugolini, Unnikrishnan, Urban, Usman, Vahlbruch, Vajente, Valdes,
  van Bakel, van Beuzekom, van~den Brand, van~den Broeck, Vander-Hyde, van~der
  Schaaf, van Heijningen, van Veggel, Vardaro, Vass, Vasúth, Vaulin, Vecchio,
  Vedovato, Veitch, Veitch, Venkateswara, Verkindt, Vetrano, Viceré,
  Vinciguerra, Vine, Vinet, Vitale, Vo, Vocca, Vorvick, Voss, Vousden,
  Vyatchanin, Wade, Wade, Wade, Walker, Wallace, Walsh, Wang, Wang, Wang, Wang,
  Wang, Ward, Warner, Was, Weaver, Wei, Weinert, Weinstein, Weiss, Welborn,
  Wen, Weßels, Westphal, Wette, Whelan, White, Whiting, Williams, Williamson,
  Willis, Willke, Wimmer, Winkler, Wipf, Wittel, Woan, Worden, Wright, Wu,
  Yablon, Yam, Yamamoto, Yancey, Yap, Yu, Yvert, Zadrożny, Zangrando, Zanolin,
  Zendri, Zevin, Zhang, Zhang, Zhang, Zhang, Zhao, Zhou, Zhou, Zhu, Zucker,
  Zuraw, Zweizig, Collaboration, \& Collaboration}]{LVC:16b}
Abbott, B.~P., Abbott, R., Abbott, T.~D., {et~al.} 2016, \apjl, 818, L22

\bibitem[{{Abbott} {et~al.}(2017{\natexlab{a}}){Abbott}, {Abbott}, {Abbott},
  {Acernese}, {Ackley}, {Adams}, {Adams}, {Addesso}, {Adhikari}, {Adya}, \&
  et~al.}]{LVC:17:grb}
{Abbott}, B.~P., {Abbott}, R., {Abbott}, T.~D., {et~al.} 2017{\natexlab{a}},
  \apjl, 848, L13

\bibitem[{{Abbott} {et~al.}(2017{\natexlab{b}}){Abbott}, {Abbott}, {Abbott},
  {Acernese}, {Ackley}, {Adams}, {Adams}, {Addesso}, {Adhikari}, {Adya}, \&
  et~al.}]{BNSmerger:17}
{Abbott}, B.~P., {Abbott}, R., {Abbott}, T.~D., {et~al.} 2017{\natexlab{b}},
  Physical Review Letters, 119, 161101

\bibitem[{{Aghakhanloo} {et~al.}(2017){Aghakhanloo}, {Murphy}, {Smith}, \&
  {Hlo{\v z}ek}}]{aghakhanlootakanloo:17}
{Aghakhanloo}, M., {Murphy}, J.~W., {Smith}, N., \& {Hlo{\v z}ek}, R. 2017,
  \mnras, 472, 591

\bibitem[{{Allison}(2012)}]{allison:12}
{Allison}, R.~J. 2012, \mnras, 421, 3338

\bibitem[{{Almeida} {et~al.}(2017){Almeida}, {Sana}, {Taylor}, {Barb{\'a}},
  {Bonanos}, {Crowther}, {Damineli}, {de Koter}, {de Mink}, {Evans}, {Gieles},
  {Grin}, {H{\'e}nault-Brunet}, {Langer}, {Lennon}, {Lockwood}, {Ma{\'{\i}}z
  Apell{\'a}niz}, {Moffat}, {Neijssel}, {Norman}, {Ram{\'{\i}}rez-Agudelo},
  {Richardson}, {Schootemeijer}, {Shenar}, {Soszy{\'n}ski}, {Tramper}, \&
  {Vink}}]{almeida:17}
{Almeida}, L.~A., {Sana}, H., {Taylor}, W., {et~al.} 2017, \aap, 598, A84

\bibitem[{{Anders} \& {Grevesse}(1989)}]{anders:89}
{Anders}, E. \& {Grevesse}, N. 1989, \gca, 53, 197

\bibitem[{{Artymowicz} \& {Lubow}(1994)}]{artymowicz:94}
{Artymowicz}, P. \& {Lubow}, S.~H. 1994, \apj, 421, 651

\bibitem[{{Arzoumanian} {et~al.}(2002){Arzoumanian}, {Chernoff}, \&
  {Cordes}}]{arzoumanian:02}
{Arzoumanian}, Z., {Chernoff}, D.~F., \& {Cordes}, J.~M. 2002, \apj, 568, 289

\bibitem[{{Asplund} {et~al.}(2009){Asplund}, {Grevesse}, {Sauval}, \&
  {Scott}}]{asplund:09}
{Asplund}, M., {Grevesse}, N., {Sauval}, A.~J., \& {Scott}, P. 2009, \araa, 47,
  481

\bibitem[{{Banerjee} {et~al.}(2012){Banerjee}, {Kroupa}, \& {Oh}}]{banerjee:12}
{Banerjee}, S., {Kroupa}, P., \& {Oh}, S. 2012, \apj, 746, 15

\bibitem[{{Bear} \& {Soker}(2017)}]{bear:17}
{Bear}, E. \& {Soker}, N. 2017, \mnras, 468, 140

\bibitem[{{Becerra} {et~al.}(2016){Becerra}, {Bianco}, {Fryer}, {Rueda}, \&
  {Ruffini}}]{becerra:16}
{Becerra}, L., {Bianco}, C.~L., {Fryer}, C.~L., {Rueda}, J.~A., \& {Ruffini},
  R. 2016, \apj, 833, 107

\bibitem[{{Belczy{\'n}ski} \& {Bulik}(1999)}]{belczynski:99}
{Belczy{\'n}ski}, K. \& {Bulik}, T. 1999, \aap, 346, 91

\bibitem[{{Belczynski} {et~al.}(2008){Belczynski}, {Kalogera}, {Rasio}, {Taam},
  {Zezas}, {Bulik}, {Maccarone}, \& {Ivanova}}]{belczynski:08}
{Belczynski}, K., {Kalogera}, V., {Rasio}, F.~A., {et~al.} 2008, \apjs, 174,
  223

\bibitem[{{Belczynski} {et~al.}(2017){Belczynski}, {Klencki}, {Meynet},
  {Fryer}, {Brown}, {Chruslinska}, {Gladysz}, {O'Shaughnessy}, {Bulik},
  {Berti}, {Holz}, {Gerosa}, {Giersz}, {Ekstrom}, {Georgy}, {Askar}, \&
  {Lasota}}]{belczynski:17}
{Belczynski}, K., {Klencki}, J., {Meynet}, G., {et~al.} 2017, arXiv:1706.07053

\bibitem[{{Belczynski} {et~al.}(2012){Belczynski}, {Wiktorowicz}, {Fryer},
  {Holz}, \& {Kalogera}}]{Belczynski+2012b}
{Belczynski}, K., {Wiktorowicz}, G., {Fryer}, C.~L., {Holz}, D.~E., \&
  {Kalogera}, V. 2012, \apj, 757, 91

\bibitem[{{Beniamini} \& {Piran}(2016)}]{beniamini:16}
{Beniamini}, P. \& {Piran}, T. 2016, \mnras, 456, 4089

\bibitem[{{Bestenlehner} {et~al.}(2011){Bestenlehner}, {Vink}, {Gr{\"a}fener},
  {Najarro}, {Evans}, {Bastian}, {Bonanos}, {Bressert}, {Crowther}, {Doran},
  {Friedrich}, {H{\'e}nault-Brunet}, {Herrero}, {de Koter}, {Langer}, {Lennon},
  {Ma{\'{\i}}z Apell{\'a}niz}, {Sana}, {Soszynski}, \&
  {Taylor}}]{bestenlehner:11}
{Bestenlehner}, J.~M., {Vink}, J.~S., {Gr{\"a}fener}, G., {et~al.} 2011, \aap,
  530, L14

\bibitem[{{Blaauw}(1956)}]{blaauw:56}
{Blaauw}, A. 1956, \pasp, 68, 495

\bibitem[{{Blaauw}(1961)}]{blaauw:61}
{Blaauw}, A. 1961, \bain, 15, 265

\bibitem[{{Blaauw}(1993)}]{blaauw:93}
{Blaauw}, A. 1993, in Astronomical Society of the Pacific Conference Series,
  Vol.~35, Massive Stars: Their Lives in the Interstellar Medium, ed. J.~P.
  {Cassinelli} \& E.~B. {Churchwell}, 207

\bibitem[{{Boersma}(1961)}]{boersma:61}
{Boersma}, J. 1961, \bain, 15, 291

\bibitem[{{Bolton}(1972)}]{Bolton1972}
{Bolton}, C.~T. 1972, \nat, 235, 271

\bibitem[{{Boubert} {et~al.}(2017{\natexlab{a}}){Boubert}, {Erkal}, {Evans}, \&
  {Izzard}}]{boubert:17b}
{Boubert}, D., {Erkal}, D., {Evans}, N.~W., \& {Izzard}, R.~G.
  2017{\natexlab{a}}, \mnras, 469, 2151

\bibitem[{{Boubert} \& {Evans}(2018)}]{boubert:18}
{Boubert}, D. \& {Evans}, N.~W. 2018, ArXiv:1804.05849

\bibitem[{{Boubert} {et~al.}(2017{\natexlab{b}}){Boubert}, {Fraser}, {Evans},
  {Green}, \& {Izzard}}]{boubert:17a}
{Boubert}, D., {Fraser}, M., {Evans}, N.~W., {Green}, D.~A., \& {Izzard}, R.~G.
  2017{\natexlab{b}}, \aap, 606, A14

\bibitem[{{Brandt} \& {Podsiadlowski}(1995)}]{brandt:95}
{Brandt}, N. \& {Podsiadlowski}, P. 1995, \mnras, 274, 461

\bibitem[{{Bray} \& {Eldridge}(2016)}]{bray:16}
{Bray}, J.~C. \& {Eldridge}, J.~J. 2016, \mnras, 461, 3747

\bibitem[{{Bray} \& {Eldridge}(2018)}]{bray:18}
{Bray}, J.~C. \& {Eldridge}, J.~J. 2018, ArXiv:1804.04414

\bibitem[{{Breivik} {et~al.}(2017){Breivik}, {Chatterjee}, \&
  {Larson}}]{breivik:17}
{Breivik}, K., {Chatterjee}, S., \& {Larson}, S.~L. 2017, \apjl, 850, L13

\bibitem[{{Brott} {et~al.}(2011){Brott}, {de Mink}, {Cantiello}, {Langer}, {de
  Koter}, {Evans}, {Hunter}, {Trundle}, \& {Vink}}]{brott:11}
{Brott}, I., {de Mink}, S.~E., {Cantiello}, M., {et~al.} 2011, \aap, 530, A115

\bibitem[{{Brown}(2015)}]{brown:15}
{Brown}, W.~R. 2015, \araa, 53, 15

\bibitem[{{Cacciari}(2009)}]{cacciari:09}
{Cacciari}, C. 2009, \memsai, 80, 97

\bibitem[{{Castor} {et~al.}(1975){Castor}, {Abbott}, \& {Klein}}]{castor:75}
{Castor}, J.~I., {Abbott}, D.~C., \& {Klein}, R.~I. 1975, \apj, 195, 157

\bibitem[{{Ceverino} \& {Klypin}(2009)}]{ceverino:09}
{Ceverino}, D. \& {Klypin}, A. 2009, \apj, 695, 292

\bibitem[{{Chan} {et~al.}(2018){Chan}, {M{\"u}ller}, {Heger}, {Pakmor}, \&
  {Springel}}]{chan:18}
{Chan}, C., {M{\"u}ller}, B., {Heger}, A., {Pakmor}, R., \& {Springel}, V.
  2018, \apjl, 852, L19

\bibitem[{{Chatterjee} {et~al.}(2005){Chatterjee}, {Vlemmings}, {Brisken},
  {Lazio}, {Cordes}, {Goss}, {Thorsett}, {Fomalont}, {Lyne}, \&
  {Kramer}}]{chatterjee:05}
{Chatterjee}, S., {Vlemmings}, W.~H.~T., {Brisken}, W.~F., {et~al.} 2005,
  \apjl, 630, L61

\bibitem[{{Chini} {et~al.}(2012){Chini}, {Hoffmeister}, {Nasseri}, {Stahl}, \&
  {Zinnecker}}]{chini:12}
{Chini}, R., {Hoffmeister}, V.~H., {Nasseri}, A., {Stahl}, O., \& {Zinnecker},
  H. 2012, \mnras, 424, 1925

\bibitem[{{Claeys} {et~al.}(2014){Claeys}, {Pols}, {Izzard}, {Vink}, \&
  {Verbunt}}]{claeys:14}
{Claeys}, J.~S.~W., {Pols}, O.~R., {Izzard}, R.~G., {Vink}, J., \& {Verbunt},
  F.~W.~M. 2014, \aap, 563, A83

\bibitem[{{Conroy} \& {Kratter}(2012)}]{conroy:12}
{Conroy}, C. \& {Kratter}, K.~M. 2012, \apj, 755, 123

\bibitem[{{Conti}(1975)}]{conti:75}
{Conti}, P.~S. 1975, Memoires of the Societe Royale des Sciences de Liege, 9,
  193

\bibitem[{{Cruz-Gonz{\'a}lez} {et~al.}(1974){Cruz-Gonz{\'a}lez},
  {Recillas-Cruz}, {Costero}, {Peimbert}, \&
  {Torres-Peimbert}}]{Cruz-Gonzalez+1974}
{Cruz-Gonz{\'a}lez}, C., {Recillas-Cruz}, E., {Costero}, R., {Peimbert}, M., \&
  {Torres-Peimbert}, S. 1974, \rmxaa, 1, 211

\bibitem[{{Davidson} {et~al.}(2016){Davidson}, {Humphreys}, \&
  {Weis}}]{davidson:16}
{Davidson}, K., {Humphreys}, R.~M., \& {Weis}, K. 2016, arXiv:1608.02007

\bibitem[{{de Bruijne}(1999)}]{debruijne:99}
{de Bruijne}, J.~H.~J. 1999, \mnras, 310, 585

\bibitem[{{De Donder} {et~al.}(1997){De Donder}, {Vanbeveren}, \& {van
  Bever}}]{dedonder:97}
{De Donder}, E., {Vanbeveren}, D., \& {van Bever}, J. 1997, \aap, 318, 812

\bibitem[{{de Kool}(1990)}]{dekool:90}
{de Kool}, M. 1990, \apj, 358, 189

\bibitem[{{De Marco} \& {Izzard}(2017)}]{demarco:17}
{De Marco}, O. \& {Izzard}, R.~G. 2017, \pasa, 34, e001

\bibitem[{{de Mink} \& {Belczynski}(2015)}]{demink:15}
{de Mink}, S.~E. \& {Belczynski}, K. 2015, \apj, 814, 58

\bibitem[{{de Mink} {et~al.}(2012){de Mink}, {Brott}, {Cantiello}, {Izzard},
  {Langer}, \& {Sana}}]{de-Mink+2012}
{de Mink}, S.~E., {Brott}, I., {Cantiello}, M., {et~al.} 2012, in Astronomical
  Society of the Pacific Conference Series, Vol. 465, Proceedings of a
  Scientific Meeting in Honor of Anthony F. J. Moffat, ed. L.~{Drissen},
  C.~{Robert}, N.~{St-Louis}, \& A.~F.~J. {Moffat}, 65

\bibitem[{{de Mink} {et~al.}(2009){de Mink}, {Cantiello}, {Langer}, {Pols},
  {Brott}, \& {Yoon}}]{demink:09}
{de Mink}, S.~E., {Cantiello}, M., {Langer}, N., {et~al.} 2009, \aap, 497, 243

\bibitem[{{de Mink} {et~al.}(2013){de Mink}, {Langer}, {Izzard}, {Sana}, \& {de
  Koter}}]{demink:13}
{de Mink}, S.~E., {Langer}, N., {Izzard}, R.~G., {Sana}, H., \& {de Koter}, A.
  2013, \apj, 764, 166

\bibitem[{{de Mink} {et~al.}(2007){de Mink}, {Pols}, \& {Hilditch}}]{demink:07}
{de Mink}, S.~E., {Pols}, O.~R., \& {Hilditch}, R.~W. 2007, \aap, 467, 1181

\bibitem[{{de Mink} {et~al.}(2014){de Mink}, {Sana}, {Langer}, {Izzard}, \&
  {Schneider}}]{demink:14}
{de Mink}, S.~E., {Sana}, H., {Langer}, N., {Izzard}, R.~G., \& {Schneider},
  F.~R.~N. 2014, \apj, 782, 7

\bibitem[{{de Wit} {et~al.}(2005){de Wit}, {Testi}, {Palla}, \&
  {Zinnecker}}]{dewit:05}
{de Wit}, W.~J., {Testi}, L., {Palla}, F., \& {Zinnecker}, H. 2005, \aap, 437,
  247

\bibitem[{{Dewi} \& {Tauris}(2000)}]{dewi:00}
{Dewi}, J.~D.~M. \& {Tauris}, T.~M. 2000, \aap, 360, 1043

\bibitem[{{Din{\c c}el} {et~al.}(2015){Din{\c c}el}, {Neuh{\"a}user}, {Yerli},
  {Ankay}, {Tetzlaff}, {Torres}, \& {Mugrauer}}]{dincel:15}
{Din{\c c}el}, B., {Neuh{\"a}user}, R., {Yerli}, S.~K., {et~al.} 2015, \mnras,
  448, 3196

\bibitem[{{Dominik} {et~al.}(2012){Dominik}, {Belczynski}, {Fryer}, {Holz},
  {Berti}, {Bulik}, {Mandel}, \& {O'Shaughnessy}}]{dominik:12}
{Dominik}, M., {Belczynski}, K., {Fryer}, C., {et~al.} 2012, \apj, 759, 52

\bibitem[{{Dominik} {et~al.}(2013){Dominik}, {Belczynski}, {Fryer}, {Holz},
  {Berti}, {Bulik}, {Mandel}, \& {O'Shaughnessy}}]{dominik:13}
{Dominik}, M., {Belczynski}, K., {Fryer}, C., {et~al.} 2013, \apj, 779, 72

\bibitem[{{Dray} {et~al.}(2005){Dray}, {Dale}, {Beer}, {Napiwotzki}, \&
  {King}}]{dray:05}
{Dray}, L.~M., {Dale}, J.~E., {Beer}, M.~E., {Napiwotzki}, R., \& {King}, A.~R.
  2005, \mnras, 364, 59

\bibitem[{{Eggleton}(1983)}]{eggleton:83}
{Eggleton}, P.~P. 1983, \apj, 268, 368

\bibitem[{{Eldridge}(2009)}]{eldridge:09}
{Eldridge}, J.~J. 2009, \mnras, 400, L20

\bibitem[{{Eldridge}(2012)}]{eldridge:12}
{Eldridge}, J.~J. 2012, \mnras, 422, 794

\bibitem[{{Eldridge} {et~al.}(2011){Eldridge}, {Langer}, \&
  {Tout}}]{eldridge:11}
{Eldridge}, J.~J., {Langer}, N., \& {Tout}, C.~A. 2011, \mnras, 414, 3501

\bibitem[{{Evans} {et~al.}(2010){Evans}, {Walborn}, {Crowther},
  {H{\'e}nault-Brunet}, {Massa}, {Taylor}, {Howarth}, {Sana}, {Lennon}, \& {van
  Loon}}]{Evans+2010a}
{Evans}, C.~J., {Walborn}, N.~R., {Crowther}, P.~A., {et~al.} 2010, \apjl, 715,
  L74

\bibitem[{{Evans} \& {Massey}(2015)}]{evans:15}
{Evans}, K.~A. \& {Massey}, P. 2015, \aj, 150, 149

\bibitem[{{Farr} {et~al.}(2011){Farr}, {Sravan}, {Cantrell}, {Kreidberg},
  {Bailyn}, {Mandel}, \& {Kalogera}}]{farr:11}
{Farr}, W.~M., {Sravan}, N., {Cantrell}, A., {et~al.} 2011, \apj, 741, 103

\bibitem[{{Fender} {et~al.}(2013){Fender}, {Maccarone}, \&
  {Heywood}}]{fender:13}
{Fender}, R.~P., {Maccarone}, T.~J., \& {Heywood}, I. 2013, \mnras, 430, 1538

\bibitem[{{Filippenko}(1997)}]{filippenko:97}
{Filippenko}, A.~V. 1997, \araa, 35, 309

\bibitem[{{Fragos} {et~al.}(2013){Fragos}, {Lehmer}, {Naoz}, {Zezas}, \&
  {Basu-Zych}}]{Fragos+2013a}
{Fragos}, T., {Lehmer}, B.~D., {Naoz}, S., {Zezas}, A., \& {Basu-Zych}, A.
  2013, \apjl, 776, L31

\bibitem[{{Fragos} {et~al.}(2009){Fragos}, {Willems}, {Kalogera}, {Ivanova},
  {Rockefeller}, {Fryer}, \& {Young}}]{fragos:09}
{Fragos}, T., {Willems}, B., {Kalogera}, V., {et~al.} 2009, \apj, 697, 1057

\bibitem[{{Fryer} {et~al.}(1998){Fryer}, {Burrows}, \& {Benz}}]{fryer:98}
{Fryer}, C., {Burrows}, A., \& {Benz}, W. 1998, \apj, 496, 333

\bibitem[{{Fryer} \& {Kalogera}(1997)}]{Fryer+1997}
{Fryer}, C. \& {Kalogera}, V. 1997, \apj, 489, 244

\bibitem[{{Fryer}(2004)}]{fryer:04}
{Fryer}, C.~L. 2004, \apjl, 601, L175

\bibitem[{{Fryer} {et~al.}(2012){Fryer}, {Belczynski}, {Wiktorowicz},
  {Dominik}, {Kalogera}, \& {Holz}}]{fryer:12}
{Fryer}, C.~L., {Belczynski}, K., {Wiktorowicz}, G., {et~al.} 2012, \apj, 749,
  91

\bibitem[{{Fryer} \& {Kalogera}(2001)}]{Fryer+2001}
{Fryer}, C.~L. \& {Kalogera}, V. 2001, \apj, 554, 548

\bibitem[{{Fujii} \& {Portegies Zwart}(2011)}]{fujii:11}
{Fujii}, M.~S. \& {Portegies Zwart}, S. 2011, Science, 334, 1380

\bibitem[{{Fuller}(2017)}]{fuller:17}
{Fuller}, J. 2017, \mnras, 470, 1642

\bibitem[{{Fuller} \& {Ro}(2018)}]{fuller:18}
{Fuller}, J. \& {Ro}, S. 2018, \mnras, 476, 1853

\bibitem[{{Gaggero} {et~al.}(2017){Gaggero}, {Bertone}, {Calore}, {Connors},
  {Lovell}, {Markoff}, \& {Storm}}]{gaggero:17}
{Gaggero}, D., {Bertone}, G., {Calore}, F., {et~al.} 2017, Physical Review
  Letters, 118, 241101

\bibitem[{{Gaia Collaboration} {et~al.}(2016{\natexlab{a}}){Gaia
  Collaboration}, {Brown}, {Vallenari}, {Prusti}, {de Bruijne}, {Mignard},
  {Drimmel}, {Babusiaux}, {Bailer-Jones}, {Bastian}, {Biermann}, {Evans},
  {Eyer}, {Jansen}, {Jordi}, {Katz}, {Klioner}, {Lammers}, {Lindegren}, {Luri},
  {O'Mullane}, {Panem}, {Pourbaix}, {Randich}, {Sartoretti}, {Siddiqui},
  {Soubiran}, {Valette}, {van Leeuwen}, {Walton}, {Aerts}, {Arenou}, {Cropper},
  {H{\o}g}, {Lattanzi}, {Grebel}, {Holland}, {Huc}, {Passot}, {Perryman},
  {Bramante}, {Cacciari}, {Casta{\~n}eda}, {Chaoul}, {Cheek}, {De Angeli},
  {Fabricius}, {Guerra}, {Hern{\'a}ndez}, {Jean-Antoine-Piccolo}, {Masana},
  {Messineo}, {Mowlavi}, {Nienartowicz}, {Ord{\'o}{\~n}ez-Blanco}, {Panuzzo},
  {Portell}, {Richards}, {Riello}, {Seabroke}, {Tanga}, {Th{\'e}venin},
  {Torra}, {Els}, {Gracia-Abril}, {Comoretto}, {Garcia-Reinaldos}, {Lock},
  {Mercier}, {Altmann}, {Andrae}, {Astraatmadja}, {Bellas-Velidis}, {Benson},
  {Berthier}, {Blomme}, {Busso}, {Carry}, {Cellino}, {Clementini}, {Cowell},
  {Creevey}, {Cuypers}, {Davidson}, {De Ridder}, {de Torres}, {Delchambre},
  {Dell'Oro}, {Ducourant}, {Fr{\'e}mat}, {Garc{\'{\i}}a-Torres}, {Gosset},
  {Halbwachs}, {Hambly}, {Harrison}, {Hauser}, {Hestroffer}, {Hodgkin},
  {Huckle}, {Hutton}, {Jasniewicz}, {Jordan}, {Kontizas}, {Korn}, {Lanzafame},
  {Manteiga}, {Moitinho}, {Muinonen}, {Osinde}, {Pancino}, {Pauwels}, {Petit},
  {Recio-Blanco}, {Robin}, {Sarro}, {Siopis}, {Smith}, {Smith}, {Sozzetti},
  {Thuillot}, {van Reeven}, {Viala}, {Abbas}, {Abreu Aramburu}, {Accart},
  {Aguado}, {Allan}, {Allasia}, {Altavilla}, {{\'A}lvarez}, {Alves},
  {Anderson}, {Andrei}, {Anglada Varela}, {Antiche}, {Antoja}, {Ant{\'o}n},
  {Arcay}, {Bach}, {Baker}, {Balaguer-N{\'u}{\~n}ez}, {Barache}, {Barata},
  {Barbier}, {Barblan}, {Barrado y Navascu{\'e}s}, {Barros}, {Barstow},
  {Becciani}, {Bellazzini}, {Bello Garc{\'{\i}}a}, {Belokurov}, {Bendjoya},
  {Berihuete}, {Bianchi}, {Bienaym{\'e}}, {Billebaud}, {Blagorodnova},
  {Blanco-Cuaresma}, {Boch}, {Bombrun}, {Borrachero}, {Bouquillon}, {Bourda},
  {Bouy}, {Bragaglia}, {Breddels}, {Brouillet}, {Br{\"u}semeister},
  {Bucciarelli}, {Burgess}, {Burgon}, {Burlacu}, {Busonero}, {Buzzi}, {Caffau},
  {Cambras}, {Campbell}, {Cancelliere}, {Cantat-Gaudin}, {Carlucci},
  {Carrasco}, {Castellani}, {Charlot}, {Charnas}, {Chiavassa}, {Clotet},
  {Cocozza}, {Collins}, {Costigan}, {Crifo}, {Cross}, {Crosta}, {Crowley},
  {Dafonte}, {Damerdji}, {Dapergolas}, {David}, {David}, {De Cat}, {de Felice},
  {de Laverny}, {De Luise}, {De March}, {de Martino}, {de Souza}, {Debosscher},
  {del Pozo}, {Delbo}, {Delgado}, {Delgado}, {Di Matteo}, {Diakite},
  {Distefano}, {Dolding}, {Dos Anjos}, {Drazinos}, {Duran}, {Dzigan},
  {Edvardsson}, {Enke}, {Evans}, {Eynard Bontemps}, {Fabre}, {Fabrizio},
  {Faigler}, {Falc{\~a}o}, {Farr{\`a}s Casas}, {Federici}, {Fedorets},
  {Fern{\'a}ndez-Hern{\'a}ndez}, {Fernique}, {Fienga}, {Figueras}, {Filippi},
  {Findeisen}, {Fonti}, {Fouesneau}, {Fraile}, {Fraser}, {Fuchs}, {Gai},
  {Galleti}, {Galluccio}, {Garabato}, {Garc{\'{\i}}a-Sedano}, {Garofalo},
  {Garralda}, {Gavras}, {Gerssen}, {Geyer}, {Gilmore}, {Girona}, {Giuffrida},
  {Gomes}, {Gonz{\'a}lez-Marcos}, {Gonz{\'a}lez-N{\'u}{\~n}ez},
  {Gonz{\'a}lez-Vidal}, {Granvik}, {Guerrier}, {Guillout}, {Guiraud},
  {G{\'u}rpide}, {Guti{\'e}rrez-S{\'a}nchez}, {Guy}, {Haigron},
  {Hatzidimitriou}, {Haywood}, {Heiter}, {Helmi}, {Hobbs}, {Hofmann}, {Holl},
  {Holland}, {Hunt}, {Hypki}, {Icardi}, {Irwin}, {Jevardat de Fombelle},
  {Jofr{\'e}}, {Jonker}, {Jorissen}, {Julbe}, {Karampelas}, {Kochoska},
  {Kohley}, {Kolenberg}, {Kontizas}, {Koposov}, {Kordopatis}, {Koubsky},
  {Krone-Martins}, {Kudryashova}, {Kull}, {Bachchan}, {Lacoste-Seris}, {Lanza},
  {Lavigne}, {Le Poncin-Lafitte}, {Lebreton}, {Lebzelter}, {Leccia}, {Leclerc},
  {Lecoeur-Taibi}, {Lemaitre}, {Lenhardt}, {Leroux}, {Liao}, {Licata},
  {Lindstr{\o}m}, {Lister}, {Livanou}, {Lobel}, {L{\"o}ffler}, {L{\'o}pez},
  {Lorenz}, {MacDonald}, {Magalh{\~a}es Fernandes}, {Managau}, {Mann},
  {Mantelet}, {Marchal}, {Marchant}, {Marconi}, {Marinoni}, {Marrese},
  {Marschalk{\'o}}, {Marshall}, {Mart{\'{\i}}n-Fleitas}, {Martino}, {Mary},
  {Matijevi{\v c}}, {Mazeh}, {McMillan}, {Messina}, {Michalik}, {Millar},
  {Miranda}, {Molina}, {Molinaro}, {Molinaro}, {Moln{\'a}r}, {Moniez},
  {Montegriffo}, {Mor}, {Mora}, {Morbidelli}, {Morel}, {Morgenthaler},
  {Morris}, {Mulone}, {Muraveva}, {Musella}, {Narbonne}, {Nelemans},
  {Nicastro}, {Noval}, {Ord{\'e}novic}, {Ordieres-Mer{\'e}}, {Osborne},
  {Pagani}, {Pagano}, {Pailler}, {Palacin}, {Palaversa}, {Parsons}, {Pecoraro},
  {Pedrosa}, {Pentik{\"a}inen}, {Pichon}, {Piersimoni}, {Pineau}, {Plachy},
  {Plum}, {Poujoulet}, {Pr{\v s}a}, {Pulone}, {Ragaini}, {Rago}, {Rambaux},
  {Ramos-Lerate}, {Ranalli}, {Rauw}, {Read}, {Regibo}, {Reyl{\'e}}, {Ribeiro},
  {Rimoldini}, {Ripepi}, {Riva}, {Rixon}, {Roelens}, {Romero-G{\'o}mez},
  {Rowell}, {Royer}, {Ruiz-Dern}, {Sadowski}, {Sagrist{\`a} Sell{\'e}s},
  {Sahlmann}, {Salgado}, {Salguero}, {Sarasso}, {Savietto}, {Schultheis},
  {Sciacca}, {Segol}, {Segovia}, {Segransan}, {Shih}, {Smareglia}, {Smart},
  {Solano}, {Solitro}, {Sordo}, {Soria Nieto}, {Souchay}, {Spagna}, {Spoto},
  {Stampa}, {Steele}, {Steidelm{\"u}ller}, {Stephenson}, {Stoev}, {Suess},
  {S{\"u}veges}, {Surdej}, {Szabados}, {Szegedi-Elek}, {Tapiador}, {Taris},
  {Tauran}, {Taylor}, {Teixeira}, {Terrett}, {Tingley}, {Trager}, {Turon},
  {Ulla}, {Utrilla}, {Valentini}, {van Elteren}, {Van Hemelryck}, {van
  Leeuwen}, {Varadi}, {Vecchiato}, {Veljanoski}, {Via}, {Vicente}, {Vogt},
  {Voss}, {Votruba}, {Voutsinas}, {Walmsley}, {Weiler}, {Weingrill}, {Wevers},
  {Wyrzykowski}, {Yoldas}, {{\v Z}erjal}, {Zucker}, {Zurbach}, {Zwitter},
  {Alecu}, {Allen}, {Allende Prieto}, {Amorim}, {Anglada-Escud{\'e}},
  {Arsenijevic}, {Azaz}, {Balm}, {Beck}, {Bernstein}, {Bigot}, {Bijaoui},
  {Blasco}, {Bonfigli}, {Bono}, {Boudreault}, {Bressan}, {Brown}, {Brunet},
  {Bunclark}, {Buonanno}, {Butkevich}, {Carret}, {Carrion}, {Chemin},
  {Ch{\'e}reau}, {Corcione}, {Darmigny}, {de Boer}, {de Teodoro}, {de Zeeuw},
  {Delle Luche}, {Domingues}, {Dubath}, {Fodor}, {Fr{\'e}zouls}, {Fries},
  {Fustes}, {Fyfe}, {Gallardo}, {Gallegos}, {Gardiol}, {Gebran}, {Gomboc},
  {G{\'o}mez}, {Grux}, {Gueguen}, {Heyrovsky}, {Hoar}, {Iannicola}, {Isasi
  Parache}, {Janotto}, {Joliet}, {Jonckheere}, {Keil}, {Kim}, {Klagyivik},
  {Klar}, {Knude}, {Kochukhov}, {Kolka}, {Kos}, {Kutka}, {Lainey}, {LeBouquin},
  {Liu}, {Loreggia}, {Makarov}, {Marseille}, {Martayan}, {Martinez-Rubi},
  {Massart}, {Meynadier}, {Mignot}, {Munari}, {Nguyen}, {Nordlander}, {Ocvirk},
  {O'Flaherty}, {Olias Sanz}, {Ortiz}, {Osorio}, {Oszkiewicz}, {Ouzounis},
  {Palmer}, {Park}, {Pasquato}, {Peltzer}, {Peralta}, {P{\'e}turaud},
  {Pieniluoma}, {Pigozzi}, {Poels}, {Prat}, {Prod'homme}, {Raison}, {Rebordao},
  {Risquez}, {Rocca-Volmerange}, {Rosen}, {Ruiz-Fuertes}, {Russo}, {Sembay},
  {Serraller Vizcaino}, {Short}, {Siebert}, {Silva}, {Sinachopoulos}, {Slezak},
  {Soffel}, {Sosnowska}, {Strai{\v z}ys}, {ter Linden}, {Terrell}, {Theil},
  {Tiede}, {Troisi}, {Tsalmantza}, {Tur}, {Vaccari}, {Vachier}, {Valles}, {Van
  Hamme}, {Veltz}, {Virtanen}, {Wallut}, {Wichmann}, {Wilkinson}, {Ziaeepour},
  \& {Zschocke}}]{Gaia-Collaboration+2016}
{Gaia Collaboration}, {Brown}, A.~G.~A., {Vallenari}, A., {et~al.}
  2016{\natexlab{a}}, \aap, 595, A2

\bibitem[{{Gaia Collaboration} {et~al.}(2016{\natexlab{b}}){Gaia
  Collaboration}, {Prusti}, {de Bruijne}, {Brown}, {Vallenari}, {Babusiaux},
  {Bailer-Jones}, {Bastian}, {Biermann}, {Evans}, {Eyer}, {Jansen}, {Jordi},
  {Klioner}, {Lammers}, {Lindegren}, {Luri}, {Mignard}, {Milligan}, {Panem},
  {Poinsignon}, {Pourbaix}, {Randich}, {Sarri}, {Sartoretti}, {Siddiqui},
  {Soubiran}, {Valette}, {van Leeuwen}, {Walton}, {Aerts}, {Arenou}, {Cropper},
  {Drimmel}, {H{\o}g}, {Katz}, {Lattanzi}, {O'Mullane}, {Grebel}, {Holland},
  {Huc}, {Passot}, {Bramante}, {Cacciari}, {Casta{\~n}eda}, {Chaoul}, {Cheek},
  {De Angeli}, {Fabricius}, {Guerra}, {Hern{\'a}ndez}, {Jean-Antoine-Piccolo},
  {Masana}, {Messineo}, {Mowlavi}, {Nienartowicz}, {Ord{\'o}{\~n}ez-Blanco},
  {Panuzzo}, {Portell}, {Richards}, {Riello}, {Seabroke}, {Tanga},
  {Th{\'e}venin}, {Torra}, {Els}, {Gracia-Abril}, {Comoretto},
  {Garcia-Reinaldos}, {Lock}, {Mercier}, {Altmann}, {Andrae}, {Astraatmadja},
  {Bellas-Velidis}, {Benson}, {Berthier}, {Blomme}, {Busso}, {Carry},
  {Cellino}, {Clementini}, {Cowell}, {Creevey}, {Cuypers}, {Davidson}, {De
  Ridder}, {de Torres}, {Delchambre}, {Dell'Oro}, {Ducourant}, {Fr{\'e}mat},
  {Garc{\'{\i}}a-Torres}, {Gosset}, {Halbwachs}, {Hambly}, {Harrison},
  {Hauser}, {Hestroffer}, {Hodgkin}, {Huckle}, {Hutton}, {Jasniewicz},
  {Jordan}, {Kontizas}, {Korn}, {Lanzafame}, {Manteiga}, {Moitinho},
  {Muinonen}, {Osinde}, {Pancino}, {Pauwels}, {Petit}, {Recio-Blanco}, {Robin},
  {Sarro}, {Siopis}, {Smith}, {Smith}, {Sozzetti}, {Thuillot}, {van Reeven},
  {Viala}, {Abbas}, {Abreu Aramburu}, {Accart}, {Aguado}, {Allan}, {Allasia},
  {Altavilla}, {{\'A}lvarez}, {Alves}, {Anderson}, {Andrei}, {Anglada Varela},
  {Antiche}, {Antoja}, {Ant{\'o}n}, {Arcay}, {Atzei}, {Ayache}, {Bach},
  {Baker}, {Balaguer-N{\'u}{\~n}ez}, {Barache}, {Barata}, {Barbier}, {Barblan},
  {Baroni}, {Barrado y Navascu{\'e}s}, {Barros}, {Barstow}, {Becciani},
  {Bellazzini}, {Bellei}, {Bello Garc{\'{\i}}a}, {Belokurov}, {Bendjoya},
  {Berihuete}, {Bianchi}, {Bienaym{\'e}}, {Billebaud}, {Blagorodnova},
  {Blanco-Cuaresma}, {Boch}, {Bombrun}, {Borrachero}, {Bouquillon}, {Bourda},
  {Bouy}, {Bragaglia}, {Breddels}, {Brouillet}, {Br{\"u}semeister},
  {Bucciarelli}, {Budnik}, {Burgess}, {Burgon}, {Burlacu}, {Busonero}, {Buzzi},
  {Caffau}, {Cambras}, {Campbell}, {Cancelliere}, {Cantat-Gaudin}, {Carlucci},
  {Carrasco}, {Castellani}, {Charlot}, {Charnas}, {Charvet}, {Chassat},
  {Chiavassa}, {Clotet}, {Cocozza}, {Collins}, {Collins}, {Costigan}, {Crifo},
  {Cross}, {Crosta}, {Crowley}, {Dafonte}, {Damerdji}, {Dapergolas}, {David},
  {David}, {De Cat}, {de Felice}, {de Laverny}, {De Luise}, {De March}, {de
  Martino}, {de Souza}, {Debosscher}, {del Pozo}, {Delbo}, {Delgado},
  {Delgado}, {di Marco}, {Di Matteo}, {Diakite}, {Distefano}, {Dolding}, {Dos
  Anjos}, {Drazinos}, {Dur{\'a}n}, {Dzigan}, {Ecale}, {Edvardsson}, {Enke},
  {Erdmann}, {Escolar}, {Espina}, {Evans}, {Eynard Bontemps}, {Fabre},
  {Fabrizio}, {Faigler}, {Falc{\~a}o}, {Farr{\`a}s Casas}, {Faye}, {Federici},
  {Fedorets}, {Fern{\'a}ndez-Hern{\'a}ndez}, {Fernique}, {Fienga}, {Figueras},
  {Filippi}, {Findeisen}, {Fonti}, {Fouesneau}, {Fraile}, {Fraser}, {Fuchs},
  {Furnell}, {Gai}, {Galleti}, {Galluccio}, {Garabato}, {Garc{\'{\i}}a-Sedano},
  {Gar{\'e}}, {Garofalo}, {Garralda}, {Gavras}, {Gerssen}, {Geyer}, {Gilmore},
  {Girona}, {Giuffrida}, {Gomes}, {Gonz{\'a}lez-Marcos},
  {Gonz{\'a}lez-N{\'u}{\~n}ez}, {Gonz{\'a}lez-Vidal}, {Granvik}, {Guerrier},
  {Guillout}, {Guiraud}, {G{\'u}rpide}, {Guti{\'e}rrez-S{\'a}nchez}, {Guy},
  {Haigron}, {Hatzidimitriou}, {Haywood}, {Heiter}, {Helmi}, {Hobbs},
  {Hofmann}, {Holl}, {Holland}, {Hunt}, {Hypki}, {Icardi}, {Irwin}, {Jevardat
  de Fombelle}, {Jofr{\'e}}, {Jonker}, {Jorissen}, {Julbe}, {Karampelas},
  {Kochoska}, {Kohley}, {Kolenberg}, {Kontizas}, {Koposov}, {Kordopatis},
  {Koubsky}, {Kowalczyk}, {Krone-Martins}, {Kudryashova}, {Kull}, {Bachchan},
  {Lacoste-Seris}, {Lanza}, {Lavigne}, {Le Poncin-Lafitte}, {Lebreton},
  {Lebzelter}, {Leccia}, {Leclerc}, {Lecoeur-Taibi}, {Lemaitre}, {Lenhardt},
  {Leroux}, {Liao}, {Licata}, {Lindstr{\o}m}, {Lister}, {Livanou}, {Lobel},
  {L{\"o}ffler}, {L{\'o}pez}, {Lopez-Lozano}, {Lorenz}, {Loureiro},
  {MacDonald}, {Magalh{\~a}es Fernandes}, {Managau}, {Mann}, {Mantelet},
  {Marchal}, {Marchant}, {Marconi}, {Marie}, {Marinoni}, {Marrese},
  {Marschalk{\'o}}, {Marshall}, {Mart{\'{\i}}n-Fleitas}, {Martino}, {Mary},
  {Matijevi{\v c}}, {Mazeh}, {McMillan}, {Messina}, {Mestre}, {Michalik},
  {Millar}, {Miranda}, {Molina}, {Molinaro}, {Molinaro}, {Moln{\'a}r},
  {Moniez}, {Montegriffo}, {Monteiro}, {Mor}, {Mora}, {Morbidelli}, {Morel},
  {Morgenthaler}, {Morley}, {Morris}, {Mulone}, {Muraveva}, {Musella},
  {Narbonne}, {Nelemans}, {Nicastro}, {Noval}, {Ord{\'e}novic},
  {Ordieres-Mer{\'e}}, {Osborne}, {Pagani}, {Pagano}, {Pailler}, {Palacin},
  {Palaversa}, {Parsons}, {Paulsen}, {Pecoraro}, {Pedrosa}, {Pentik{\"a}inen},
  {Pereira}, {Pichon}, {Piersimoni}, {Pineau}, {Plachy}, {Plum}, {Poujoulet},
  {Pr{\v s}a}, {Pulone}, {Ragaini}, {Rago}, {Rambaux}, {Ramos-Lerate},
  {Ranalli}, {Rauw}, {Read}, {Regibo}, {Renk}, {Reyl{\'e}}, {Ribeiro},
  {Rimoldini}, {Ripepi}, {Riva}, {Rixon}, {Roelens}, {Romero-G{\'o}mez},
  {Rowell}, {Royer}, {Rudolph}, {Ruiz-Dern}, {Sadowski}, {Sagrist{\`a}
  Sell{\'e}s}, {Sahlmann}, {Salgado}, {Salguero}, {Sarasso}, {Savietto},
  {Schnorhk}, {Schultheis}, {Sciacca}, {Segol}, {Segovia}, {Segransan},
  {Serpell}, {Shih}, {Smareglia}, {Smart}, {Smith}, {Solano}, {Solitro},
  {Sordo}, {Soria Nieto}, {Souchay}, {Spagna}, {Spoto}, {Stampa}, {Steele},
  {Steidelm{\"u}ller}, {Stephenson}, {Stoev}, {Suess}, {S{\"u}veges}, {Surdej},
  {Szabados}, {Szegedi-Elek}, {Tapiador}, {Taris}, {Tauran}, {Taylor},
  {Teixeira}, {Terrett}, {Tingley}, {Trager}, {Turon}, {Ulla}, {Utrilla},
  {Valentini}, {van Elteren}, {Van Hemelryck}, {van Leeuwen}, {Varadi},
  {Vecchiato}, {Veljanoski}, {Via}, {Vicente}, {Vogt}, {Voss}, {Votruba},
  {Voutsinas}, {Walmsley}, {Weiler}, {Weingrill}, {Werner}, {Wevers},
  {Whitehead}, {Wyrzykowski}, {Yoldas}, {{\v Z}erjal}, {Zucker}, {Zurbach},
  {Zwitter}, {Alecu}, {Allen}, {Allende Prieto}, {Amorim},
  {Anglada-Escud{\'e}}, {Arsenijevic}, {Azaz}, {Balm}, {Beck}, {Bernstein},
  {Bigot}, {Bijaoui}, {Blasco}, {Bonfigli}, {Bono}, {Boudreault}, {Bressan},
  {Brown}, {Brunet}, {Bunclark}, {Buonanno}, {Butkevich}, {Carret}, {Carrion},
  {Chemin}, {Ch{\'e}reau}, {Corcione}, {Darmigny}, {de Boer}, {de Teodoro}, {de
  Zeeuw}, {Delle Luche}, {Domingues}, {Dubath}, {Fodor}, {Fr{\'e}zouls},
  {Fries}, {Fustes}, {Fyfe}, {Gallardo}, {Gallegos}, {Gardiol}, {Gebran},
  {Gomboc}, {G{\'o}mez}, {Grux}, {Gueguen}, {Heyrovsky}, {Hoar}, {Iannicola},
  {Isasi Parache}, {Janotto}, {Joliet}, {Jonckheere}, {Keil}, {Kim},
  {Klagyivik}, {Klar}, {Knude}, {Kochukhov}, {Kolka}, {Kos}, {Kutka}, {Lainey},
  {LeBouquin}, {Liu}, {Loreggia}, {Makarov}, {Marseille}, {Martayan},
  {Martinez-Rubi}, {Massart}, {Meynadier}, {Mignot}, {Munari}, {Nguyen},
  {Nordlander}, {Ocvirk}, {O'Flaherty}, {Olias Sanz}, {Ortiz}, {Osorio},
  {Oszkiewicz}, {Ouzounis}, {Palmer}, {Park}, {Pasquato}, {Peltzer}, {Peralta},
  {P{\'e}turaud}, {Pieniluoma}, {Pigozzi}, {Poels}, {Prat}, {Prod'homme},
  {Raison}, {Rebordao}, {Risquez}, {Rocca-Volmerange}, {Rosen}, {Ruiz-Fuertes},
  {Russo}, {Sembay}, {Serraller Vizcaino}, {Short}, {Siebert}, {Silva},
  {Sinachopoulos}, {Slezak}, {Soffel}, {Sosnowska}, {Strai{\v z}ys}, {ter
  Linden}, {Terrell}, {Theil}, {Tiede}, {Troisi}, {Tsalmantza}, {Tur},
  {Vaccari}, {Vachier}, {Valles}, {Van Hamme}, {Veltz}, {Virtanen}, {Wallut},
  {Wichmann}, {Wilkinson}, {Ziaeepour}, \&
  {Zschocke}}]{Gaia-Collaboration+2016a}
{Gaia Collaboration}, {Prusti}, T., {de Bruijne}, J.~H.~J., {et~al.}
  2016{\natexlab{b}}, \aap, 595, A1

\bibitem[{{Gatto} {et~al.}(2015){Gatto}, {Walch}, {Low}, {Naab}, {Girichidis},
  {Glover}, {W{\"u}nsch}, {Klessen}, {Clark}, {Baczynski}, {Peters},
  {Ostriker}, {Ib{\'a}{\~n}ez-Mej{\'{\i}}a}, \& {Haid}}]{gatto:15}
{Gatto}, A., {Walch}, S., {Low}, M.-M.~M., {et~al.} 2015, \mnras, 449, 1057

\bibitem[{{Gies}(1987)}]{gies:87}
{Gies}, D.~R. 1987, \apjs, 64, 545

\bibitem[{{Gies} \& {Bolton}(1986)}]{gies:86}
{Gies}, D.~R. \& {Bolton}, C.~T. 1986, \apjs, 61, 419

\bibitem[{{Gilmore} \& {Reid}(1983)}]{gilmore:83}
{Gilmore}, G. \& {Reid}, N. 1983, \mnras, 202, 1025

\bibitem[{{G{\"o}tberg} {et~al.}(2017){G{\"o}tberg}, {de Mink}, \&
  {Groh}}]{gotberg:17}
{G{\"o}tberg}, Y., {de Mink}, S.~E., \& {Groh}, J.~H. 2017, \aap, 608, A11

\bibitem[{{G{\"o}tberg} {et~al.}(2018){G{\"o}tberg}, {de Mink}, {Groh},
  {Kupfer}, {Crowther}, {Zapartas}, \& {Renzo}}]{gotberg:18}
{G{\"o}tberg}, Y., {de Mink}, S.~E., {Groh}, J.~H., {et~al.} 2018,
  ArXiv:1802.03018

\bibitem[{{Gott}(1971)}]{gott:71}
{Gott}, J.~R. 1971, \nat, 234, 342

\bibitem[{{Gott} {et~al.}(1970){Gott}, {Gunn}, \& {Ostriker}}]{gott:70}
{Gott}, III, J.~R., {Gunn}, J.~E., \& {Ostriker}, J.~P. 1970, \apjl, 160, L91

\bibitem[{{Grefenstette} {et~al.}(2017){Grefenstette}, {Fryer}, {Harrison},
  {Boggs}, {DeLaney}, {Laming}, {Reynolds}, {Alexander}, {Barret},
  {Christensen}, {Craig}, {Forster}, {Giommi}, {Hailey}, {Hornstrup},
  {Kitaguchi}, {Koglin}, {Lopez}, {Mao}, {Madsen}, {Miyasaka}, {Mori}, {Perri},
  {Pivovaroff}, {Puccetti}, {Rana}, {Stern}, {Westergaard}, {Wik}, {Zhang}, \&
  {Zoglauer}}]{grefenstette:17}
{Grefenstette}, B.~W., {Fryer}, C.~L., {Harrison}, F.~A., {et~al.} 2017, \apj,
  834, 19

\bibitem[{{Grudzinska} {et~al.}(2015){Grudzinska}, {Belczynski}, {Casares}, {de
  Mink}, {Ziolkowski}, {Negueruela}, {Rib{\'o}}, {Ribas}, {Paredes}, {Herrero},
  \& {Benacquista}}]{Grudzinska+2015}
{Grudzinska}, M., {Belczynski}, K., {Casares}, J., {et~al.} 2015, \mnras, 452,
  2773

\bibitem[{{Gull} \& {Sofia}(1979)}]{gull:79}
{Gull}, T.~R. \& {Sofia}, S. 1979, \apj, 230, 782

\bibitem[{{Gunn} \& {Ostriker}(1970)}]{gunn:70}
{Gunn}, J.~E. \& {Ostriker}, J.~P. 1970, \apj, 160, 979

\bibitem[{{Guseinov} {et~al.}(2005){Guseinov}, {Ankay}, \&
  {Tagieva}}]{guseinov:05}
{Guseinov}, O.~H., {Ankay}, A., \& {Tagieva}, S.~O. 2005, Astrophysics, 48, 330

\bibitem[{{Gvaramadze} {et~al.}(2012){Gvaramadze}, {Weidner}, {Kroupa}, \&
  {Pflamm-Altenburg}}]{gvaramadze:12}
{Gvaramadze}, V.~V., {Weidner}, C., {Kroupa}, P., \& {Pflamm-Altenburg}, J.
  2012, \mnras, 424, 3037

\bibitem[{{Hamann} \& {Koesterke}(1998)}]{hamann:98}
{Hamann}, W.~R. \& {Koesterke}, L. 1998, \aap, 335, 1003

\bibitem[{{Heger} {et~al.}(2003){Heger}, {Fryer}, {Woosley}, {Langer}, \&
  {Hartmann}}]{Heger+2003}
{Heger}, A., {Fryer}, C.~L., {Woosley}, S.~E., {Langer}, N., \& {Hartmann},
  D.~H. 2003, \apj, 591, 288

\bibitem[{{Hellings}(1983)}]{hellings:83}
{Hellings}, P. 1983, \apss, 96, 37

\bibitem[{{Hills}(1988)}]{hills:88}
{Hills}, J.~G. 1988, \nat, 331, 687

\bibitem[{{Hirai} {et~al.}(2018){Hirai}, {Podsiadlowski}, \&
  {Yamada}}]{hirai:18}
{Hirai}, R., {Podsiadlowski}, P., \& {Yamada}, S. 2018, ArXiv:1803.10808

\bibitem[{{Hobbs} {et~al.}(2005){Hobbs}, {Lorimer}, {Lyne}, \&
  {Kramer}}]{hobbs:05}
{Hobbs}, G., {Lorimer}, D.~R., {Lyne}, A.~G., \& {Kramer}, M. 2005, \mnras,
  360, 974

\bibitem[{{Holland-Ashford} {et~al.}(2017){Holland-Ashford}, {Lopez},
  {Auchettl}, {Temim}, \& {Ramirez-Ruiz}}]{holland-ashford:17}
{Holland-Ashford}, T., {Lopez}, L.~A., {Auchettl}, K., {Temim}, T., \&
  {Ramirez-Ruiz}, E. 2017, \apj, 844, 84

\bibitem[{{Hoogerwerf} {et~al.}(2000){Hoogerwerf}, {de Bruijne}, \& {de
  Zeeuw}}]{hoogerwerf:00}
{Hoogerwerf}, R., {de Bruijne}, J.~H.~J., \& {de Zeeuw}, P.~T. 2000, \apjl,
  544, L133

\bibitem[{{Hoogerwerf} {et~al.}(2001){Hoogerwerf}, {de Bruijne}, \& {de
  Zeeuw}}]{hoogerwerf:01}
{Hoogerwerf}, R., {de Bruijne}, J.~H.~J., \& {de Zeeuw}, P.~T. 2001, \aap, 365,
  49

\bibitem[{{Huang}(1963)}]{huang:63}
{Huang}, S.-S. 1963, \apj, 138, 471

\bibitem[{{Hulse} \& {Taylor}(1975)}]{hulse:75}
{Hulse}, R.~A. \& {Taylor}, J.~H. 1975, \apjl, 195, L51

\bibitem[{{Humphreys} {et~al.}(2016){Humphreys}, {Weis}, {Davidson}, \&
  {Gordon}}]{humphreys:16}
{Humphreys}, R.~M., {Weis}, K., {Davidson}, K., \& {Gordon}, M.~S. 2016, \apj,
  825, 64

\bibitem[{{Hurley} {et~al.}(2000){Hurley}, {Pols}, \& {Tout}}]{hurley:00}
{Hurley}, J.~R., {Pols}, O.~R., \& {Tout}, C.~A. 2000, \mnras, 315, 543

\bibitem[{{Hurley} {et~al.}(2002){Hurley}, {Tout}, \& {Pols}}]{hurley:02}
{Hurley}, J.~R., {Tout}, C.~A., \& {Pols}, O.~R. 2002, \mnras, 329, 897

\bibitem[{{Hut}(1981)}]{hut:81}
{Hut}, P. 1981, \aap, 99, 126

\bibitem[{{Ivanova} {et~al.}(2013){Ivanova}, {Justham}, {Chen}, {De Marco},
  {Fryer}, {Gaburov}, {Ge}, {Glebbeek}, {Han}, {Li}, {Lu}, {Marsh},
  {Podsiadlowski}, {Potter}, {Soker}, {Taam}, {Tauris}, {van den Heuvel}, \&
  {Webbink}}]{ivanova:13}
{Ivanova}, N., {Justham}, S., {Chen}, X., {et~al.} 2013, \aapr, 21, 59

\bibitem[{{Izzard} {et~al.}(2006){Izzard}, {Dray}, {Karakas}, {Lugaro}, \&
  {Tout}}]{izzard:06}
{Izzard}, R.~G., {Dray}, L.~M., {Karakas}, A.~I., {Lugaro}, M., \& {Tout},
  C.~A. 2006, \aap, 460, 565

\bibitem[{{Izzard} {et~al.}(2009){Izzard}, {Glebbeek}, {Stancliffe}, \&
  {Pols}}]{izzard:09}
{Izzard}, R.~G., {Glebbeek}, E., {Stancliffe}, R.~J., \& {Pols}, O.~R. 2009,
  \aap, 508, 1359

\bibitem[{Izzard {et~al.}(2018)Izzard, Preece, Jofre, Halabi, Masseron, \&
  Tout}]{izzard:18}
Izzard, R.~G., Preece, H., Jofre, P., {et~al.} 2018, Monthly Notices of the
  Royal Astronomical Society, 473, 2984

\bibitem[{{Izzard} {et~al.}(2004{\natexlab{a}}){Izzard}, {Ramirez-Ruiz}, \&
  {Tout}}]{izzard:04b}
{Izzard}, R.~G., {Ramirez-Ruiz}, E., \& {Tout}, C.~A. 2004{\natexlab{a}},
  \mnras, 348, 1215

\bibitem[{{Izzard} {et~al.}(2004{\natexlab{b}}){Izzard}, {Tout}, {Karakas}, \&
  {Pols}}]{izzard:04}
{Izzard}, R.~G., {Tout}, C.~A., {Karakas}, A.~I., \& {Pols}, O.~R.
  2004{\natexlab{b}}, \mnras, 350, 407

\bibitem[{{Janka}(2013)}]{janka:13}
{Janka}, H.-T. 2013, \mnras, 434, 1355

\bibitem[{{Janka}(2017)}]{janka:17}
{Janka}, H.-T. 2017, \apj, 837, 84

\bibitem[{{Jeffries} {et~al.}(2014){Jeffries}, {Jackson}, {Cottaar}, {Koposov},
  {Lanzafame}, {Meyer}, {Prisinzano}, {Randich}, {Sacco}, {Brugaletta},
  {Caramazza}, {Damiani}, {Franciosini}, {Frasca}, {Gilmore}, {Feltzing},
  {Micela}, {Alfaro}, {Bensby}, {Pancino}, {Recio-Blanco}, {de Laverny},
  {Lewis}, {Magrini}, {Morbidelli}, {Costado}, {Jofr{\'e}}, {Klutsch}, {Lind},
  \& {Maiorca}}]{Jeffries+2014}
{Jeffries}, R.~D., {Jackson}, R.~J., {Cottaar}, M., {et~al.} 2014, \aap, 563,
  A94

\bibitem[{{Jilinski} {et~al.}(2010){Jilinski}, {Ortega}, {Drake}, \& {de la
  Reza}}]{jilinski:10}
{Jilinski}, E., {Ortega}, V.~G., {Drake}, N.~A., \& {de la Reza}, R. 2010,
  \apj, 721, 469

\bibitem[{{Johnston} {et~al.}(2005){Johnston}, {Hobbs}, {Vigeland}, {Kramer},
  {Weisberg}, \& {Lyne}}]{johnston:05}
{Johnston}, S., {Hobbs}, G., {Vigeland}, S., {et~al.} 2005, \mnras, 364, 1397

\bibitem[{{Justham} \& {Schawinski}(2012)}]{justham:12}
{Justham}, S. \& {Schawinski}, K. 2012, \mnras, 423, 1641

\bibitem[{{Kalogera}(1996)}]{kalogera:96}
{Kalogera}, V. 1996, \apj, 471, 352

\bibitem[{{Kaplan} {et~al.}(2008){Kaplan}, {Chatterjee}, {Gaensler}, \&
  {Anderson}}]{kaplan:08}
{Kaplan}, D.~L., {Chatterjee}, S., {Gaensler}, B.~M., \& {Anderson}, J. 2008,
  \apj, 677, 1201

\bibitem[{{Kaspi} {et~al.}(1996){Kaspi}, {Bailes}, {Manchester}, {Stappers}, \&
  {Bell}}]{kaspi:96}
{Kaspi}, V.~M., {Bailes}, M., {Manchester}, R.~N., {Stappers}, B.~W., \&
  {Bell}, J.~F. 1996, \nat, 381, 584

\bibitem[{{Katsuda} {et~al.}(2018){Katsuda}, {Morii}, {Janka},
  {Wongwathanarat}, {Nakamura}, {Kotake}, {Mori}, {M{\"u}ller}, {Takiwaki},
  {Tanaka}, {Tominaga}, \& {Tsunemi}}]{katsuda:18}
{Katsuda}, S., {Morii}, M., {Janka}, H.-T., {et~al.} 2018, \apj, 856, 18

\bibitem[{{Katz}(1975)}]{katz:75}
{Katz}, J.~I. 1975, \nat, 253, 698

\bibitem[{{Kerzendorf} {et~al.}(2017){Kerzendorf}, {Do}, {de Mink},
  {G{\"o}tberg}, {Millisaljevic}, {Zapartas}, {Renzo}, {Justham},
  {Podsiadlowski}, \& {Fesen}}]{kerzendorf:17}
{Kerzendorf}, W.~E., {Do}, T., {de Mink}, S.~E., {et~al.} 2017,
  ArXiv:1711.00055

\bibitem[{{Khazov} {et~al.}(2016){Khazov}, {Yaron}, {Gal-Yam}, {Manulis},
  {Rubin}, {Kulkarni}, {Arcavi}, {Kasliwal}, {Ofek}, {Cao}, {Perley},
  {Sollerman}, {Horesh}, {Sullivan}, {Filippenko}, {Nugent}, {Howell}, {Cenko},
  {Silverman}, {Ebeling}, {Taddia}, {Johansson}, {Laher}, {Surace},
  {Rebbapragada}, {Wozniak}, \& {Matheson}}]{khazov:16}
{Khazov}, D., {Yaron}, O., {Gal-Yam}, A., {et~al.} 2016, \apj, 818, 3

\bibitem[{{Kiminki} \& {Kobulnicky}(2012)}]{kiminki:12}
{Kiminki}, D.~C. \& {Kobulnicky}, H.~A. 2012, \apj, 751, 4

\bibitem[{{Kiminki} \& {Smith}(2018)}]{kiminki:18}
{Kiminki}, M.~M. \& {Smith}, N. 2018, \mnras

\bibitem[{{Kimm} \& {Cen}(2014)}]{kimm:14}
{Kimm}, T. \& {Cen}, R. 2014, \apj, 788, 121

\bibitem[{{Kippenhahn} \& {Weigert}(1967)}]{kippenhahn:67}
{Kippenhahn}, R. \& {Weigert}, A. 1967, \zap, 65, 251

\bibitem[{{Knigge} {et~al.}(2011){Knigge}, {Coe}, \&
  {Podsiadlowski}}]{knigge:11}
{Knigge}, C., {Coe}, M.~J., \& {Podsiadlowski}, P. 2011, \nat, 479, 372

\bibitem[{{Kobulnicky} {et~al.}(2018){Kobulnicky}, {Chick}, \&
  {Povich}}]{kobulnicky:18}
{Kobulnicky}, H.~A., {Chick}, W.~T., \& {Povich}, M.~S. 2018, \apj, 856, 74

\bibitem[{{Kobulnicky} \& {Fryer}(2007)}]{kobulnicky:07}
{Kobulnicky}, H.~A. \& {Fryer}, C.~L. 2007, \apj, 670, 747

\bibitem[{{Kobulnicky} {et~al.}(2014){Kobulnicky}, {Kiminki}, {Lundquist},
  {Burke}, {Chapman}, {Keller}, {Lester}, {Rolen}, {Topel}, {Bhattacharjee},
  {Smullen}, {Vargas {\'A}lvarez}, {Runnoe}, {Dale}, \&
  {Brotherton}}]{kobulnicky:14}
{Kobulnicky}, H.~A., {Kiminki}, D.~C., {Lundquist}, M.~J., {et~al.} 2014,
  \apjs, 213, 34

\bibitem[{{Kochanek} {et~al.}(2014){Kochanek}, {Adams}, \&
  {Belczynski}}]{kochanek:14}
{Kochanek}, C.~S., {Adams}, S.~M., \& {Belczynski}, K. 2014, \mnras, 443, 1319

\bibitem[{{Kouwenhoven} {et~al.}(2005){Kouwenhoven}, {Brown}, {Zinnecker},
  {Kaper}, \& {Portegies Zwart}}]{kouwenhoven:05}
{Kouwenhoven}, M.~B.~N., {Brown}, A.~G.~A., {Zinnecker}, H., {Kaper}, L., \&
  {Portegies Zwart}, S.~F. 2005, \aap, 430, 137

\bibitem[{{Kroupa}(2001)}]{kroupa:01}
{Kroupa}, P. 2001, \mnras, 322, 231

\bibitem[{{Kusenko} \& {Segr{\`e}}(1996)}]{kusenko:96}
{Kusenko}, A. \& {Segr{\`e}}, G. 1996, Physical Review Letters, 77, 4872

\bibitem[{{Lamers}(2013)}]{lamers:13}
{Lamers}, H.~J.~G.~L.~M. 2013, in Astronomical Society of the Pacific
  Conference Series, Vol. 470, 370 Years of Astronomy in Utrecht, ed.
  G.~{Pugliese}, A.~{de Koter}, \& M.~{Wijburg}, 97

\bibitem[{{Langer}(2012)}]{langer:12}
{Langer}, N. 2012, \araa, 50, 107

\bibitem[{{Larson}(1974)}]{larson:74}
{Larson}, R.~B. 1974, \mnras, 169, 229

\bibitem[{{Lauterborn}(1970)}]{lauterborn:70}
{Lauterborn}, D. 1970, \aap, 7, 150

\bibitem[{{Lennon} {et~al.}(2016){Lennon}, {van der Marel}, {Ramos Lerate},
  {O'Mullane}, \& {Sahlmann}}]{lennon:16}
{Lennon}, D.~J., {van der Marel}, R.~P., {Ramos Lerate}, M., {O'Mullane}, W.,
  \& {Sahlmann}, J. 2016, arXiv:1611.05504

\bibitem[{{Leonard}(1991)}]{leonard:91}
{Leonard}, P.~J.~T. 1991, \aj, 101, 562

\bibitem[{{Leonard} {et~al.}(1994){Leonard}, {Hills}, \& {Dewey}}]{leonard:94}
{Leonard}, P.~J.~T., {Hills}, J.~G., \& {Dewey}, R.~J. 1994, \apjl, 423, L19

\bibitem[{{Liu} {et~al.}(2015){Liu}, {Tauris}, {R{\"o}pke}, {Moriya},
  {Kruckow}, {Stancliffe}, \& {Izzard}}]{liu:15}
{Liu}, Z.-W., {Tauris}, T.~M., {R{\"o}pke}, F.~K., {et~al.} 2015, \aap, 584,
  A11

\bibitem[{{Livio} \& {Soker}(1988)}]{livio:88}
{Livio}, M. \& {Soker}, N. 1988, \apj, 329, 764

\bibitem[{{Lucas} {et~al.}(2018){Lucas}, {Rybak}, {Bonnell}, \&
  {Gieles}}]{lucas:17}
{Lucas}, W.~E., {Rybak}, M., {Bonnell}, I.~A., \& {Gieles}, M. 2018, \mnras,
  474, 3582

\bibitem[{{Lyne} \& {Lorimer}(1994)}]{lyne:94}
{Lyne}, A.~G. \& {Lorimer}, D.~R. 1994, \nat, 369, 127

\bibitem[{{Ma} {et~al.}(2016{\natexlab{a}}){Ma}, {Hopkins}, {Kasen},
  {Quataert}, {Faucher-Gigu{\`e}re}, {Kere{\v s}}, {Murray}, \&
  {Strom}}]{Ma+2016}
{Ma}, X., {Hopkins}, P.~F., {Kasen}, D., {et~al.} 2016{\natexlab{a}}, \mnras,
  459, 3614

\bibitem[{{Ma} {et~al.}(2016{\natexlab{b}}){Ma}, {Hopkins}, {Kasen},
  {Quataert}, {Faucher-Gigu{\`e}re}, {Kere{\v s}}, {Murray}, \&
  {Strom}}]{ma:16}
{Ma}, X., {Hopkins}, P.~F., {Kasen}, D., {et~al.} 2016{\natexlab{b}}, \mnras,
  459, 3614

\bibitem[{{Maeder}(1996)}]{maeder:96}
{Maeder}, A. 1996, in Liege International Astrophysical Colloquia, Vol.~33,
  Liege International Astrophysical Colloquia, ed. J.~M. {Vreux}, A.~{Detal},
  D.~{Fraipont-Caro}, E.~{Gosset}, \& G.~{Rauw}, 39

\bibitem[{{Maeder} \& {Conti}(1994)}]{maeder:94}
{Maeder}, A. \& {Conti}, P.~S. 1994, \araa, 32, 227

\bibitem[{{Maeder} \& {Meynet}(2000)}]{maeder:00}
{Maeder}, A. \& {Meynet}, G. 2000, \aap, 361, 159

\bibitem[{{Ma{\'{\i}}z Apell{\'a}niz} {et~al.}(2018){Ma{\'{\i}}z
  Apell{\'a}niz}, {Pantaleoni Gonz{\'a}lez}, {Barb{\'a}},
  {Sim{\'o}n-D{\'{\i}}az}, {Negueruela}, {Lennon}, {Sota}, \& {Trigueros
  P{\'a}ez}}]{maiz-appellaniz:18}
{Ma{\'{\i}}z Apell{\'a}niz}, J., {Pantaleoni Gonz{\'a}lez}, M., {Barb{\'a}},
  R.~H., {et~al.} 2018, ArXiv:1804.06915

\bibitem[{{Mandel}(2016)}]{mandel:16}
{Mandel}, I. 2016, \mnras, 456, 578

\bibitem[{{Maoil{\'e}idigh} {et~al.}(2005){Maoil{\'e}idigh}, {Meurs}, \&
  {Norci}}]{maoileidigh:05}
{Maoil{\'e}idigh}, C.~{\'O}., {Meurs}, E.~J.~A., \& {Norci}, L. 2005, \na, 10,
  591

\bibitem[{{Marchetti} {et~al.}(2017){Marchetti}, {Rossi}, {Kordopatis},
  {Brown}, {Rimoldi}, {Starkenburg}, {Youakim}, \& {Ashley}}]{marchetti:17}
{Marchetti}, T., {Rossi}, E.~M., {Kordopatis}, G., {et~al.} 2017, \mnras, 470,
  1388

\bibitem[{Marchetti {et~al.}(2017)Marchetti, Rossi, Kordopatis, Brown, Rimoldi,
  Starkenburg, Youakim, \& Ashley}]{marchetti:17b}
Marchetti, T., Rossi, E.~M., Kordopatis, G., {et~al.} 2017, Monthly Notices of
  the Royal Astronomical Society, 470, 1388

\bibitem[{{Martin} {et~al.}(2009){Martin}, {Tout}, \& {Pringle}}]{martin:09}
{Martin}, R.~G., {Tout}, C.~A., \& {Pringle}, J.~E. 2009, \mnras, 397, 1563

\bibitem[{{Mason} {et~al.}(2009){Mason}, {Hartkopf}, {Gies}, {Henry}, \&
  {Helsel}}]{mason:09}
{Mason}, B.~D., {Hartkopf}, W.~I., {Gies}, D.~R., {Henry}, T.~J., \& {Helsel},
  J.~W. 2009, \aj, 137, 3358

\bibitem[{McEvoy {et~al.}(2017)McEvoy, Dufton, Smoker, Lambert, Keenan,
  Schneider, \& de~Wit}]{mcevoy:17}
McEvoy, C.~M., Dufton, P.~L., Smoker, J.~V., {et~al.} 2017, The Astrophysical
  Journal, 842, 32

\bibitem[{{Moe} \& {Di Stefano}(2017)}]{moe:17}
{Moe}, M. \& {Di Stefano}, R. 2017, \apjs, 230, 15

\bibitem[{{Morningstar} {et~al.}(2014){Morningstar}, {Miller}, {Reis}, \&
  {Ebisawa}}]{morningstar:14}
{Morningstar}, W.~R., {Miller}, J.~M., {Reis}, R.~C., \& {Ebisawa}, K. 2014,
  \apjl, 784, L18

\bibitem[{{Neo} {et~al.}(1977){Neo}, {Miyaji}, {Nomoto}, \&
  {Sugimoto}}]{neo:77}
{Neo}, S., {Miyaji}, S., {Nomoto}, K., \& {Sugimoto}, D. 1977, \pasj, 29, 249

\bibitem[{{Neugent} {et~al.}(2018){Neugent}, {Massey}, {Morrell}, {Skiff}, \&
  {Georgy}}]{neugent:18}
{Neugent}, K., {Massey}, P., {Morrell}, N., {Skiff}, B., \& {Georgy}, C. 2018,
  ArXiv:1803.02859

\bibitem[{{Nieuwenhuijzen} \& {de Jager}(1990)}]{nieuwenhuijzen:90}
{Nieuwenhuijzen}, H. \& {de Jager}, C. 1990, \aap, 231, 134

\bibitem[{{Nomoto} {et~al.}(2013){Nomoto}, {Kobayashi}, \&
  {Tominaga}}]{nomoto:13}
{Nomoto}, K., {Kobayashi}, C., \& {Tominaga}, N. 2013, \araa, 51, 457

\bibitem[{{Oh} \& {Kroupa}(2016)}]{oh:16}
{Oh}, S. \& {Kroupa}, P. 2016, \aap, 590, A107

\bibitem[{{{\"O}pik}(1924)}]{opik:24}
{{\"O}pik}, E. 1924, Publications of the Tartu Astrofizica Observatory, 25

\bibitem[{{O'Shaughnessy} {et~al.}(2017){O'Shaughnessy}, {Gerosa}, \&
  {Wysocki}}]{oshaughnessy:17}
{O'Shaughnessy}, R., {Gerosa}, D., \& {Wysocki}, D. 2017, Physical Review
  Letters, 119, 011101

\bibitem[{{Packet}(1981)}]{packet:81}
{Packet}, W. 1981, \aap, 102, 17

\bibitem[{{Pauldrach} {et~al.}(1994){Pauldrach}, {Kudritzki}, {Puls}, {Butler},
  \& {Hunsinger}}]{pauldrach:94}
{Pauldrach}, A.~W.~A., {Kudritzki}, R.~P., {Puls}, J., {Butler}, K., \&
  {Hunsinger}, J. 1994, \aap, 283, 525

\bibitem[{{Perets} \& {{\v S}ubr}(2012)}]{perets:12}
{Perets}, H.~B. \& {{\v S}ubr}, L. 2012, \apj, 751, 133

\bibitem[{{Perryman} {et~al.}(2001){Perryman}, {de Boer}, {Gilmore}, {H{\o}g},
  {Lattanzi}, {Lindegren}, {Luri}, {Mignard}, {Pace}, \& {de
  Zeeuw}}]{perryman:01}
{Perryman}, M.~A.~C., {de Boer}, K.~S., {Gilmore}, G., {et~al.} 2001, \aap,
  369, 339

\bibitem[{{Pfahl} {et~al.}(2002){Pfahl}, {Rappaport}, {Podsiadlowski}, \&
  {Spruit}}]{pfahl:02}
{Pfahl}, E., {Rappaport}, S., {Podsiadlowski}, P., \& {Spruit}, H. 2002, \apj,
  574, 364

\bibitem[{{Podsiadlowski} {et~al.}(2004){Podsiadlowski}, {Langer},
  {Poelarends}, {Rappaport}, {Heger}, \& {Pfahl}}]{podsiadlowski:04}
{Podsiadlowski}, P., {Langer}, N., {Poelarends}, A.~J.~T., {et~al.} 2004, \apj,
  612, 1044

\bibitem[{{Pols}(1994)}]{pols:94}
{Pols}, O.~R. 1994, \aap, 290, 119

\bibitem[{{Pols} {et~al.}(1991){Pols}, {Cote}, {Waters}, \& {Heise}}]{pols:91}
{Pols}, O.~R., {Cote}, J., {Waters}, L.~B.~F.~M., \& {Heise}, J. 1991, \aap,
  241, 419

\bibitem[{{Pols} {et~al.}(1998){Pols}, {Schr{\"o}der}, {Hurley}, {Tout}, \&
  {Eggleton}}]{pols:98}
{Pols}, O.~R., {Schr{\"o}der}, K.-P., {Hurley}, J.~R., {Tout}, C.~A., \&
  {Eggleton}, P.~P. 1998, \mnras, 298, 525

\bibitem[{{Poveda} {et~al.}(1967){Poveda}, {Ruiz}, \& {Allen}}]{poveda:67}
{Poveda}, A., {Ruiz}, J., \& {Allen}, C. 1967, Boletin de los Observatorios
  Tonantzintla y Tacubaya, 4, 86

\bibitem[{{Quataert} \& {Shiode}(2012)}]{quataert:12}
{Quataert}, E. \& {Shiode}, J. 2012, \mnras, 423, L92

\bibitem[{{Ramiaramanantsoa} {et~al.}(2018){Ramiaramanantsoa}, {Moffat},
  {Harmon}, {Ignace}, {St-Louis}, {Vanbeveren}, {Shenar}, {Pablo},
  {Richardson}, {Howarth}, {Stevens}, {Piaulet}, {St-Jean}, {Eversberg},
  {Pigulski}, {Popowicz}, {Kuschnig}, {Zoc{\l}o{\'n}ska}, {Buysschaert},
  {Handler}, {Weiss}, {Wade}, {Rucinski}, {Zwintz}, {Luckas}, {Heathcote},
  {Cacella}, {Powles}, {Locke}, {Bohlsen}, {Chen{\'e}}, {Miszalski}, {Waldron},
  {Kotze}, {Kotze}, \& {B{\"o}hm}}]{ramiaramanantsoa:17}
{Ramiaramanantsoa}, T., {Moffat}, A.~F.~J., {Harmon}, R., {et~al.} 2018,
  \mnras, 473, 5532

\bibitem[{{Ram{\'{\i}}rez-Agudelo} {et~al.}(2015){Ram{\'{\i}}rez-Agudelo},
  {Sana}, {de Mink}, {H{\'e}nault-Brunet}, {de Koter}, {Langer}, {Tramper},
  {Gr{\"a}fener}, {Evans}, {Vink}, {Dufton}, \& {Taylor}}]{ramirez-agudelo:15}
{Ram{\'{\i}}rez-Agudelo}, O.~H., {Sana}, H., {de Mink}, S.~E., {et~al.} 2015,
  \aap, 580, A92

\bibitem[{{Remillard} \& {McClintock}(2006)}]{remillard:06}
{Remillard}, R.~A. \& {McClintock}, J.~E. 2006, \araa, 44, 49

\bibitem[{{Renzo} {et~al.}(2017){Renzo}, {Ott}, {Shore}, \& {de
  Mink}}]{renzo:17}
{Renzo}, M., {Ott}, C.~D., {Shore}, S.~N., \& {de Mink}, S.~E. 2017, \aap, 603,
  A118

\bibitem[{{Repetto} {et~al.}(2012){Repetto}, {Davies}, \&
  {Sigurdsson}}]{repetto:12}
{Repetto}, S., {Davies}, M.~B., \& {Sigurdsson}, S. 2012, \mnras, 425, 2799

\bibitem[{{Repetto} {et~al.}(2017){Repetto}, {Igoshev}, \&
  {Nelemans}}]{repetto:17}
{Repetto}, S., {Igoshev}, A.~P., \& {Nelemans}, G. 2017, \mnras, 467, 298

\bibitem[{{Repetto} \& {Nelemans}(2015)}]{repetto:15}
{Repetto}, S. \& {Nelemans}, G. 2015, \mnras, 453, 3341

\bibitem[{{Rimoldi} {et~al.}(2016){Rimoldi}, {Portegies Zwart}, \&
  {Rossi}}]{rimoldi:16}
{Rimoldi}, A., {Portegies Zwart}, S., \& {Rossi}, E.~M. 2016, Computational
  Astrophysics and Cosmology, 3, 2

\bibitem[{{Rossi} {et~al.}(2014){Rossi}, {Kobayashi}, \& {Sari}}]{rossi:14}
{Rossi}, E.~M., {Kobayashi}, S., \& {Sari}, R. 2014, \apj, 795, 125

\bibitem[{{Sana} {et~al.}(2012){Sana}, {de Mink}, {de Koter}, {Langer},
  {Evans}, {Gieles}, {Gosset}, {Izzard}, {Le Bouquin}, \&
  {Schneider}}]{sana:12}
{Sana}, H., {de Mink}, S.~E., {de Koter}, A., {et~al.} 2012, Science, 337, 444

\bibitem[{{Sana} \& {Evans}(2011)}]{sana:11}
{Sana}, H. \& {Evans}, C.~J. 2011, in IAU Symposium, Vol. 272, Active OB Stars:
  Structure, Evolution, Mass Loss, and Critical Limits, ed. C.~{Neiner},
  G.~{Wade}, G.~{Meynet}, \& G.~{Peters}, 474--485

\bibitem[{{Sana} {et~al.}(2014){Sana}, {Le Bouquin}, {Lacour}, {Berger},
  {Duvert}, {Gauchet}, {Norris}, {Olofsson}, {Pickel}, {Zins}, {Absil}, {de
  Koter}, {Kratter}, {Schnurr}, \& {Zinnecker}}]{sana:14}
{Sana}, H., {Le Bouquin}, J.-B., {Lacour}, S., {et~al.} 2014, \apjs, 215, 15

\bibitem[{{Schneider} {et~al.}(2015){Schneider}, {Izzard}, {Langer}, \& {de
  Mink}}]{schneider:15}
{Schneider}, F.~R.~N., {Izzard}, R.~G., {Langer}, N., \& {de Mink}, S.~E. 2015,
  \apj, 805, 20

\bibitem[{{Schneider} {et~al.}(2016){Schneider}, {Podsiadlowski}, {Langer},
  {Castro}, \& {Fossati}}]{schneider:16}
{Schneider}, F.~R.~N., {Podsiadlowski}, P., {Langer}, N., {Castro}, N., \&
  {Fossati}, L. 2016, \mnras, 457, 2355

\bibitem[{{Schneider} {et~al.}(2018){Schneider}, {Sana}, {Evans},
  {Bestenlehner}, {Castro}, {Fossati}, {Gr{\"a}fener}, {Langer},
  {Ram{\'{\i}}rez-Agudelo}, {Sab{\'{\i}}n-Sanjuli{\'a}n},
  {Sim{\'o}n-D{\'{\i}}az}, {Tramper}, {Crowther}, {de Koter}, {de Mink},
  {Dufton}, {Garcia}, {Gieles}, {H{\'e}nault-Brunet}, {Herrero}, {Izzard},
  {Kalari}, {Lennon}, {Ma{\'{\i}}z Apell{\'a}niz}, {Markova}, {Najarro},
  {Podsiadlowski}, {Puls}, {Taylor}, {van Loon}, {Vink}, \&
  {Norman}}]{schneider:18}
{Schneider}, F.~R.~N., {Sana}, H., {Evans}, C.~J., {et~al.} 2018, Science, 359,
  69

\bibitem[{{Schwab} {et~al.}(2010){Schwab}, {Podsiadlowski}, \&
  {Rappaport}}]{schwab:10}
{Schwab}, J., {Podsiadlowski}, P., \& {Rappaport}, S. 2010, \apj, 719, 722

\bibitem[{{Shklovskii}(1970)}]{shklovskii:70}
{Shklovskii}, I.~S. 1970, \sovast, 13, 562

\bibitem[{{Smith}(2016)}]{smith:16}
{Smith}, N. 2016, \mnras, 461, 3353

\bibitem[{{Smith} \& {Tombleson}(2015)}]{smith:15}
{Smith}, N. \& {Tombleson}, R. 2015, \mnras, 447, 598

\bibitem[{{Soberman} {et~al.}(1997){Soberman}, {Phinney}, \& {van den
  Heuvel}}]{soberman:97}
{Soberman}, G.~E., {Phinney}, E.~S., \& {van den Heuvel}, E.~P.~J. 1997, \aap,
  327, 620

\bibitem[{{Socrates} {et~al.}(2005){Socrates}, {Blaes}, {Hungerford}, \&
  {Fryer}}]{socrates:05}
{Socrates}, A., {Blaes}, O., {Hungerford}, A., \& {Fryer}, C.~L. 2005, \apj,
  632, 531

\bibitem[{{Steenbrugge} {et~al.}(2003){Steenbrugge}, {de Bruijne},
  {Hoogerwerf}, \& {de Zeeuw}}]{steenbrugge:03}
{Steenbrugge}, K.~C., {de Bruijne}, J.~H.~J., {Hoogerwerf}, R., \& {de Zeeuw},
  P.~T. 2003, \aap, 402, 587

\bibitem[{{Stevenson} {et~al.}(2017){Stevenson}, {Vigna-G{\'o}mez}, {Mandel},
  {Barrett}, {Neijssel}, {Perkins}, \& {de Mink}}]{stevenson:17}
{Stevenson}, S., {Vigna-G{\'o}mez}, A., {Mandel}, I., {et~al.} 2017, Nature
  Communications, 8, 14906

\bibitem[{{Stone}(1991)}]{stone:91}
{Stone}, R.~C. 1991, \aj, 102, 333

\bibitem[{{Tauris}(2015)}]{tauris:15}
{Tauris}, T.~M. 2015, \mnras, 448, L6

\bibitem[{{Tauris} {et~al.}(2017){Tauris}, {Kramer}, {Freire}, {Wex}, {Janka},
  {Langer}, {Podsiadlowski}, {Bozzo}, {Chaty}, {Kruckow}, {van den Heuvel},
  {Antoniadis}, {Breton}, \& {Champion}}]{tauris:17}
{Tauris}, T.~M., {Kramer}, M., {Freire}, P.~C.~C., {et~al.} 2017, \apj, 846,
  170

\bibitem[{{Tauris} \& {Takens}(1998)}]{tauris:98}
{Tauris}, T.~M. \& {Takens}, R.~J. 1998, \aap, 330, 1047

\bibitem[{{Tetzlaff} {et~al.}(2014){Tetzlaff}, {Din{\c c}el}, {Neuh{\"a}user},
  \& {Kovtyukh}}]{tetzlaff:14}
{Tetzlaff}, N., {Din{\c c}el}, B., {Neuh{\"a}user}, R., \& {Kovtyukh}, V.~V.
  2014, \mnras, 438, 3587

\bibitem[{{Tetzlaff} {et~al.}(2011){Tetzlaff}, {Neuh{\"a}user}, \&
  {Hohle}}]{tetzlaff:11}
{Tetzlaff}, N., {Neuh{\"a}user}, R., \& {Hohle}, M.~M. 2011, \mnras, 410, 190

\bibitem[{{Tetzlaff} {et~al.}(2013){Tetzlaff}, {Torres}, {Neuh{\"a}user}, \&
  {Hohle}}]{tetzlaff:13}
{Tetzlaff}, N., {Torres}, G., {Neuh{\"a}user}, R., \& {Hohle}, M.~M. 2013,
  \mnras, 435, 879

\bibitem[{{Thorne} \& {Zytkow}(1975)}]{thorne:75}
{Thorne}, K.~S. \& {Zytkow}, A.~N. 1975, \apjl, 199, L19

\bibitem[{{Thorne} \& {Zytkow}(1977)}]{thorne:77}
{Thorne}, K.~S. \& {Zytkow}, A.~N. 1977, \apj, 212, 832

\bibitem[{{Tout} {et~al.}(1997){Tout}, {Aarseth}, {Pols}, \&
  {Eggleton}}]{tout:97}
{Tout}, C.~A., {Aarseth}, S.~J., {Pols}, O.~R., \& {Eggleton}, P.~P. 1997,
  \mnras, 291, 732

\bibitem[{{Tutukov} \& {Yungelson}(1973)}]{tutukov:73}
{Tutukov}, A. \& {Yungelson}, L. 1973, Nauchnye Informatsii, 27, 70

\bibitem[{{van den Bergh}(1980)}]{vandenbergh:80}
{van den Bergh}, S. 1980, Journal of Astrophysics and Astronomy, 1, 67

\bibitem[{{van den Heuvel} \& {Heise}(1972)}]{van-den-Heuvel+1972}
{van den Heuvel}, E.~P.~J. \& {Heise}, J. 1972, \nat, 239, 67

\bibitem[{{van den Heuvel} {et~al.}(2017){van den Heuvel}, {Portegies Zwart},
  \& {de Mink}}]{vandenheuvel:17}
{van den Heuvel}, E.~P.~J., {Portegies Zwart}, S.~F., \& {de Mink}, S.~E. 2017,
  \mnras, 471, 4256

\bibitem[{{van Oijen}(1989)}]{vanoijen:89}
{van Oijen}, J.~G.~J. 1989, \aap, 217, 115

\bibitem[{{van Rensbergen} {et~al.}(1996){van Rensbergen}, {Vanbeveren}, \& {De
  Loore}}]{vanrensbergen:96}
{van Rensbergen}, W., {Vanbeveren}, D., \& {De Loore}, C. 1996, \aap, 305, 825

\bibitem[{{Vanbeveren}(1982)}]{vanbeveren:82}
{Vanbeveren}, D. 1982, \aap, 113, 205

\bibitem[{{Verbunt} \& {Cator}(2017)}]{verbunt:17}
{Verbunt}, F. \& {Cator}, E. 2017, Journal of Astrophysics and Astronomy, 38

\bibitem[{{Verbunt} {et~al.}(2017){Verbunt}, {Igoshev}, \&
  {Cator}}]{verbunt:17b}
{Verbunt}, F., {Igoshev}, A., \& {Cator}, E. 2017, \aap, 608, A57

\bibitem[{{Vink} {et~al.}(2000){Vink}, {de Koter}, \& {Lamers}}]{vink:00}
{Vink}, J.~S., {de Koter}, A., \& {Lamers}, H.~J.~G.~L.~M. 2000, \aap, 362, 295

\bibitem[{{Vink} {et~al.}(2001){Vink}, {de Koter}, \& {Lamers}}]{vink:01}
{Vink}, J.~S., {de Koter}, A., \& {Lamers}, H.~J.~G.~L.~M. 2001, \aap, 369, 574

\bibitem[{{Webbink}(1984)}]{webbink:84}
{Webbink}, R.~F. 1984, \apj, 277, 355

\bibitem[{{Webster} \& {Murdin}(1972)}]{Webster+1972}
{Webster}, B.~L. \& {Murdin}, P. 1972, \nat, 235, 37

\bibitem[{{Wheeler} {et~al.}(1975){Wheeler}, {Lecar}, \& {McKee}}]{wheeler:75}
{Wheeler}, J.~C., {Lecar}, M., \& {McKee}, C.~F. 1975, \apj, 200, 145

\bibitem[{{Wijers} {et~al.}(1992){Wijers}, {van Paradijs}, \& {van den
  Heuvel}}]{wijers:92}
{Wijers}, R.~A.~M.~J., {van Paradijs}, J., \& {van den Heuvel}, E.~P.~J. 1992,
  \aap, 261, 145

\bibitem[{{Wongwathanarat} {et~al.}(2013){Wongwathanarat}, {Janka}, \&
  {M{\"u}ller}}]{wongwathanarat:13}
{Wongwathanarat}, A., {Janka}, H.-T., \& {M{\"u}ller}, E. 2013, \aap, 552, A126

\bibitem[{{Wyrzykowski} {et~al.}(2016){Wyrzykowski}, {Kostrzewa-Rutkowska},
  {Skowron}, {Rybicki}, {Mr{\'o}z}, {Koz{\l}owski}, {Udalski}, {Szyma{\'n}ski},
  {Pietrzy{\'n}ski}, {Soszy{\'n}ski}, {Ulaczyk}, {Pietrukowicz}, {Poleski},
  {Pawlak}, {I{\l}kiewicz}, \& {Rattenbury}}]{wyrzykowski:16}
{Wyrzykowski}, {\L}., {Kostrzewa-Rutkowska}, Z., {Skowron}, J., {et~al.} 2016,
  \mnras, 458, 3012

\bibitem[{{Wysocki} {et~al.}(2018){Wysocki}, {Gerosa}, {O'Shaughnessy},
  {Belczynski}, {Gladysz}, {Berti}, {Kesden}, \& {Holz}}]{wysocki:18}
{Wysocki}, D., {Gerosa}, D., {O'Shaughnessy}, R., {et~al.} 2018, \prd, 97,
  043014

\bibitem[{{Yoon} {et~al.}(2017){Yoon}, {Dessart}, \& {Clocchiatti}}]{yoon:17}
{Yoon}, S.-C., {Dessart}, L., \& {Clocchiatti}, A. 2017, \apj, 840, 10

\bibitem[{{Zahn}(1977)}]{zahn:77}
{Zahn}, J.-P. 1977, \aap, 57, 383

\bibitem[{{Zapartas} {et~al.}(2017{\natexlab{a}}){Zapartas}, {de Mink},
  {Izzard}, {Yoon}, {Badenes}, {G{\"o}tberg}, {de Koter}, {Neijssel}, {Renzo},
  {Schootemeijer}, \& {Shrotriya}}]{zapartas:17}
{Zapartas}, E., {de Mink}, S.~E., {Izzard}, R.~G., {et~al.} 2017{\natexlab{a}},
  \aap, 601, A29

\bibitem[{{Zapartas} {et~al.}(2017{\natexlab{b}}){Zapartas}, {de Mink}, {Van
  Dyk}, {Fox}, {Smith}, {Bostroem}, {de Koter}, {Filippenko}, {Izzard},
  {Kelly}, {Neijssel}, {Renzo}, \& {Ryder}}]{zapartas:17b}
{Zapartas}, E., {de Mink}, S.~E., {Van Dyk}, S.~D., {et~al.}
  2017{\natexlab{b}}, \apj, 842, 125

\bibitem[{{Zwicky}(1957)}]{zwicky:57}
{Zwicky}, F. 1957, \zap, 44, 64

\end{thebibliography}

\appendix

\section{Comparison to selected previous population studies}
\label{sec:literature}

Various earlier studies have simulated the production of binaries
containing a compact object and/or the producing runaway stars. We briefly discuss a selection that focuses on estimates for unbound companions and comment briefly comment on how our findings compare.

\cite{dedonder:97} presented one of the first extensive binary evolution studies discussing predictions for the O-type runaways as well as the systems that remain bound. They consider a variety of assumptions for the initial distributions and the uncertain physical parameters concerning the efficiency of mass transfer, angular momentum loss and the common envelope ejection. Generally we find that our results agree well.  They find that 16--23\% of systems remain bound after the SN explosion of the primary, which is consistent with our findings $1- \mathcal{D} = 14^{+22}_{-10}\%$.  They find that between 6\% and 27\% of the O-type are unbound former companions that are now single, which is also agreement with our walkaway fraction (one but last column in their table 1).

The find that  2--7\% of the O-type stars have velocities large than $30\,\kms$ (final column in their Table 1), which is slightly larger but still consistent with what we find.  We expect that this difference may be in part the result of a difference the treatment of mass transfer and angular momentum loss, although the combined effect of further differences in our assumptions will also contribute. Their default assumption is that a fixed fraction $\beta_\mathrm{RLOF} = 0.5$ of the material transferred during Roche-lobe overflow leaves the system through the outer L2 Lagrangian point forming a ring around the binary system, which leads to larger angular momentum loss, shrinking the orbit further.  This is somewhat similar to our model variation where we assumed $\gamma_{\rm RLOF} = \gamma_{\rm disk} $ although our results cannot be compared one to one, since we adopt a different, physically motivate, assumption for the mass transfer efficiency. In our simulations we find that this assumption is not increasing the number of runaway stars, as the fraction of system that merges increases too.

\cite{eldridge:11} perform population synthesis simulations with a
detailed stellar evolutionary evolutionary code to investigate O and
early B
type runaway stars. Their primary aim is the predict the spatial distribution of different types of CCSN and gamma-ray bursts. They find a disruption fraction, $\mathcal{D}=80\%$, in good agreement with our results, $\mathcal{D} = 86^{+10}_{-22}\%$.  In their work they use the term runaway to refer to all unbound companions with velocities larger than $5\kms$, which encloses the large majority of what we refer to as walkaway stars, but they also quote estimates for stars faster than $30\kms$ which corresponds to our definition of runaway stars.  They predict a walkaway (runaway) fraction of $2.2\%$ ($0.5\%$) for O type stars
%, and $7.1\%$ ($2.2\%$) for B-type stars, respectively 
in their simulations for $Z=0.02$. For comparison, for stars more massive than $15\,\Msun$, which roughly corresponds to O type stars,  we estimate a walkaway (runaway) fraction of  $10^{+4.7}_{-8.5}$\% ($0.5^{+1.0}_{-0.3}$\%).  Our runaway fraction agrees very well.  We find a somewhat larger walkaway fraction, but we consider this a fairly good agreement, given the uncertainties and differences in definitions that we have adopted. Also the distance they estimate that ejected stars can reach is in good agreement with our predictions and they also note that the majority unbound companions accrete mass from their companions before the first CC.  Our predictions for the velocities of bound systems are slightly lower compared to \cite{eldridge:11}: the vast majority of bound post-CC systems is slower than $30\,\mathrm{km\ s^{-1}}$, and a high-velocity tail barely extending beyond about $100\,\mathrm{km\  s^{-1}}$.

%The authors do not comment on the We note that they do not discuss the discrepancy between the low predictions for the fraction of O star runaways and the fractions typically quoted in observational studies, that $\sim10\%$ of O-type stars are runaways resulting from the binary ejection scenario \citep[][Sana \emph{et al.}, in preparation]{blaauw:61,gies:87,stone:91}. We return to this in our discussion in \Secref{sec:howtofind}.

Very recently, \cite{boubert:18} published a study  investigating  the hypothesis that Be stars are products of mass transfer in binary systems \citep[e.g.][]{pols:91, demink:13}.  They % precise astrometry of the Tycho-Gaia Astrometric Solution (TGAS) from the first Gaia Data Release and
compare the  kinematics of a flux-limited sample of Galactic Be stars
with binary population synthesis simulations. These simulations were
obtained with a different version of the  binary evolutionary code that we use. Generally our findings are in agreement, despite the minor differences in the model assumptions. They also find a large fraction of unbound companions, many of which are rapidly rotating and moving at velocities slower than $30\,\kms$ (e.g., their Fig. 6 and 7b), in good agreement with our findings.   They also find that the natal kick distribution does not greatly affect the resulting velocity distribution and/or the runaway fraction and that the maximum distance traveled by ejected Be stars is likely to be smaller than the vertical scale height of the thin disk, in agreement with our findings.

We find some disagreement in the provided explanations of the
theoretical results.  For example, the authors state in their
section~3.2 that that mass transfer shrinks the orbit and accelerates
the secondary star. While this the case initially, upon the onset of
mass transfer, we find that the orbit generally widens after the
reversal of the mass ratio. The widening and the inversion of the mass
ratio both slow down the orbital velocity as we verified both
analytically \Secref{sec:analytics} and by detailed inspection of
representative example systems, e.g. \Secref{sec:example}.  In
section~3.3 they state that  ``whether a binary is disrupted by a
supernova is principally determined by whether the primary loses more
than half its mass \citep{blaauw:61}, and the kick on the compact
object is only a second order effect''.  We find instead that mass
loss during the explosion is rarely responsible for unbinding of a
binary system that can produce a runaway star. The amount of mass lost
needs to exceed half of the total mass of the system (and not half of
the primary star), which is rarely achieved in our simulations. This
is because the  CC progenitor loses most of its mass during the
preceding mass transfer phase. We find that the Blaauw kick due to
rapid mass loss is only important of initially very wide binaries, in which the two stars have not exchanged mass prior to CC.  This difference may be in part due to the differences in the assumptions for the range of initial orbital periods. The authors consider systems with initial orbital periods up to  $10^{10}$ days in their simulations, which means that the majority of their progenitors should effectively evolve as single stars. We consider instead systems up to $10^{5.5}$ days, which is more appropriate for the more massive progenitors that we are interested in.

Many further studies investigated the populations of binaries in the
context of the formation of X-ray binaries and binary neutron stars
and black holes. Providing a complete overview and detailed
comparison is beyond our present scope. We discuss below a comparison
to a limited set of studies.

\cite{brandt:95} focus on binary systems remaining bound at the first CC, with the aim of understanding the
effects of natal kicks on X-ray binaries.  They estimate a disruption
fraction in their calculations of $\mathcal{D}\simeq73-81\%$
(depending on the companion mass), in good agreement with our
results.

\cite{kalogera:96} presented a similar analytic study of the effects of natal kicks on the systemic velocities of X-ray binaries. Our population synthesis results agree in predicting systemic velocities of bound post-CC systems generally lower than the pre-explosion orbital velocity,
except with $\sigma_\mathrm{kick} \gg v_\mathrm{orb}^\mathrm{pre-CC}$ (corresponding to large $\xi$ in the notation used by \citealt{kalogera:96}).

 \cite{fryer:98} investigated the impact of NS natal kicks on the
 formation of NS X-ray binaries and NS-NS binaries, but also present
 results for the unbound companions. Assuming a bimodal kick
 distribution, they find that most O/B-type ejected companion move
 slower than about $50\kms$, consistent with our findings.

 \cite{dray:05} focus on high-mass runaways, which become Wolf-Rayet (WR) stars during or before their post-disruption evolution. They argue in favor of significant BH kick amplitudes to explain the rarity of BH-WR binaries and the observed velocity distribution of WR runaways. We reach similar conclusions based on our simulations for different assumptions for the BH kicks. The systemic velocities we find for bound systems are also in reasonable agreement with those found by \cite{dray:05}

Recently, \cite{tauris:17} presented a detailed study of the evolutionary processes leading to the formation of NS NS binary systems. They also find that the majority of binaries hosting a NS after the first CC have systemic velocities smaller than $\sim 30\,\mathrm{km\ s^{-1}}$, in good agreement with the results shown in the left panel of \Figref{fig:bound_v_dist}.

\section{Output files}
\label{app:data}

The outcome of our population synthesis calculations will be made
available upon publication. For early inquiries, please contact the
lead author.
%We provide the output of our population synthesis simulations at
% \todo{where?}.
Each file corresponds to one parameter variation (see
\Secref{sec:methods} and \Secref{sec:param_variation}), and logs the
following information for each binary system where a star goes
CC\footnote{We stop our computations at the first CC event.}.
\begin{itemize}
 \item primary ZAMS, pre-CC, and post-CC (corresponding to the NS or
   BH mass if the primary is the star collapsing) masses in $M_\odot$ units:
   \texttt{M1zams}, \texttt{M1preCC}, \texttt{M1postCC};
 \item secondary ZAMS, pre-CC, and post-CC masses in $M_\odot$ units:
   \texttt{M2zams}, \texttt{M2preCC}, \texttt{M2postCC};
 \item fallback fraction $f_b$ for each star (set to zero for the star
   that is not collapsing):
   \texttt{fb1}, \texttt{fb2}
 \item evolutionary stage before and after CC (stellar types listed
   according to \citealt{hurley:00}):
   \texttt{type1preCC}, \texttt{type1postCC},
   \texttt{type2preCC}, \texttt{type2postCC}
 \item post-CC velocities of the 2 stars in the original frame, in
   $\mathrm{km\ s^{-1}}$: \texttt{v1postCC}, \texttt{v2postCC};
 \item pre-CC and post-CC eccentricity (the latter is -1 for mergers
   and disrupted systems): \texttt{e\_preCC}, \texttt{e\_postCC};
 \item pre-CC and post-CC separation in $R_\odot$ units (the latter
   is set to 0 for mergers and disrupted systems):
   \texttt{a\_preCC}, \texttt{a\_postCC};
 \item ZAMS, pre-CC, and post-CC periods in
   days: \texttt{Pzams}, \texttt{PpreCC}, \texttt{PpostCC};
 \item kick amplitude in $\mathrm{km\ s^{-1}}$: \texttt{v\_kick};
 \item kick direction, with $\theta$ angle between the collapsing star
   orbital velocity and the kick (see also \citetalias{tauris:98} for notation):
   \texttt{theta}, \texttt{phi};
 \item systemic velocity in $\mathrm{km\ s^{-1}}$: \texttt{v\_sys};
 \item age of the system at the time of CC in Myr: \texttt{t\_explosion};
 \item time left in the current evolutionary stage in Myr:
   \texttt{t\_remaining};
 \item system probability (see below): \texttt{Prob}.
\end{itemize}

The system probability corresponds to the hyper-volume of the initial
parameter space ($M_1^\mathrm{ZAMS}$, $q^\mathrm{ZAMS}$,
$P^\mathrm{ZAMS}$) represented by each binary system in our model
grid. In other words, the probability of each system is the
statistical weight of the system seen as sampling point for the initial
distributions. To construct distributions of the output quantities (e.g.~those presented in
\Figref{fig:v_dist}-\ref{fig:bound_v_dist}), the properties of each
system should be weigh with the corresponding probability. Similarly,
the mean value of a quantity $\langle x \rangle$ (e.g.~$\langle v \rangle$ in
\Tabref{tab:parameters}) should be calculated using:

\begin{equation}
  \label{eq:mean}
  \langle x \rangle = \frac{\int x P(x)\,dx}{\int P(x)\,dx} \equiv
  \frac{\sum_i^{}  x_i P_i}{\sum_i^{} P_i} \ \ ,
\end{equation}
where $P$ is the probability, and the index $i$ runs over all the
binary systems in a population.

To calculate the average distance traveled by stars
ejected by the binary disruption, we use:

\begin{equation}
  \begin{aligned}
    \label{eq:meanD}
\langle L \rangle \equiv \langle v\times \Delta t\rangle = \\
  \frac{\sum_i \texttt{v2postCC}\times [\texttt{t\_remaining}+0.1\times\tau_\mathrm{MS}(M_2)]
    \times P_i}{\sum_i P_i} \ \ ,
  \end{aligned}
\end{equation}

% \begin{equation}
%   \label{eq:meanD}
%   \langle L \rangle \equiv \langle v\times \Delta t\rangle = 
%   \frac{\sum_i \texttt{v2postCC}\times [\texttt{t\_remaining}+0.1\times\tau_\mathrm{MS}(M_2)]
%     \times P_i}{\sum_i P_i} \ \ ,
% \end{equation}

where the second term in squared brackets accounts, albeit in a
simplified way, for the helium core burning duration of the
rejuvenated star.

\section{Observable velocity distribution}

Figure~\ref{fig:v_dist_obs} shows the velocity distribution of MS
stars ejected by the disruption of binaries that can
be directly compared to observations provided that (i) the
contribution of dynamical ejection can be separated in the observed
sample and (ii) the effects of the galactic potential can be
neglected, i.e.\ effectively each ejected star moves in a straight
line at constant velocity for the remaining duration of its MS. This is the same information of
\Figref{fig:v_dist}, but each bin is populated considering also the
remaining MS lifetime of the ejected star
($\tau_\mathrm{MS}$).

\begin{figure*}[htbp]
  \includegraphics[width=\textwidth]{figures/v2_dist_duration}
  \caption{Velocity distribution of ejected stars, including the
    finite MS lifetime to populate the bins (see also
    \Figref{fig:v_dist}). }
  \label{fig:v_dist_obs}
\end{figure*}

\begin{figure*}[bp]
  \includegraphics[width=\textwidth]{figures/v2_dist_log}
  \caption{Same as \Figref{fig:v_dist}, but using a logarithmic scale
    for the velocity. The use of a logarithmic scale allows for the
    display of a wider range of velocities. A minor peak in the grey
    histogram can be seen between $100\lesssim
    v_\mathrm{dis}/\mathrm{km \ s^{-1}} \lesssim
    400$, but is absent in the histograms for massive ejected
    stars. Such high ejection velocities are reached through a common
    envelope evolution without accreting mass.}
  \label{fig:v_dist_log}
\end{figure*}

\section{Pre-collapse distributions}

We present in this appendix the pre-CC distribution in separation,
mass of the collapsing star, and mass of the companions for all the
binaries with a MS companion to the collapsing star in our fiducial
simulation. Similar distributions can be derived for all our parameter
variations from the data files available that will be made
available.%\todo{where}.
These distributions can inform studies of the interaction of the
SN shock with the companion star
\citep[e.g.][]{wheeler:75,liu:15,rimoldi:16, hirai:18}.

We show in \Figref{fig:preCC_sep_dist} the pre-CC separation distribution. The colors indicate indicate the minimum mass of the MS
companion (i.e., not of the collapsing star). Roughly speaking, the
two peaks shown in \Figref{fig:preCC_sep_dist} correspond to the
orbital widening due to conservative (case A and early case B) and
non-conservative (late case B and case C) mass transfer phase. For all
pre-CC separations shorter than $10^3\,R_\odot$, the dashed and solid
histograms coincide, indicating that all these binaries have
experienced a direct interaction previously during the evolution.

\begin{figure}[htbp]
  \includegraphics[width=0.5\textwidth]{figures/Sep_dist}
  \caption{Pre-CC separation distribution for binaries with a
    collapsing star and a main sequence companion. Colors indicate the
    pre-CC mass of the MS companion according to the legend. Dashed
    histograms indicate post-interaction (RLOF or common envelope)
    binaries. The top panel shows the corresponding cumulative
    distributions.}
  \label{fig:preCC_sep_dist}
\end{figure}

Figure \ref{fig:preCC_mass_dist} shows the mass distribution for the
exploding star and the MS companion ($M_\mathrm{CC}$, and $M_2$,
respectively), at the pre-CC stage. The combination of the
distributions shown in \Figref{fig:preCC_mass_dist} and
\Figref{fig:preCC_sep_dist}, together with the effects of the natal
kick distribution results in the ejection velocities in
\Figref{fig:v_dist} which is our main result.

\begin{figure}[htbp]
  \includegraphics[width=0.5\textwidth]{figures/masses_dist}
  \caption{Pre-CC mass distribution for the exploding star
    ($M_\mathrm{CC}$) and the MS companion ($M_2$) binaries with a
    collapsing star and a main sequence companion. Dashed
    histograms indicate post-interaction (RLOF or common envelope)
    binaries. The top panel shows the corresponding cumulative
    distributions.}
  \label{fig:preCC_mass_dist}
\end{figure}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2018-04-30 VFTS682.tex

@@
 
-\defcitealias{tauris:98}{TT98}
-\defcitealias{ramirez-agudelo:15}{R-A15}
+% \defcitealias{tauris:98}{TT98}
+% \defcitealias{ramirez-agudelo:15}{R-A15}
 
@@
 
-\title{Massive Runaway and Walkaway Stars}
-\subtitle{A study of the kinematical imprints of the physical processes \\ governing the evolution and explosion of their binary progenitors}
+\title{VFTS682: a confirmed dynamical ejection?}
 
-\author{M.~Renzo\inst{1} \and E.~Zapartas\inst{1} \and S.~E.~de~Mink\inst{1} \and
- Y.~G\"otberg\inst{1} \and S.~Justham\inst{2,3} \and
- R.~J.~Farmer\inst{1} \and R.~G.~Izzard\inst{4}
- \and S.~Toonen\inst{1}
- \and H.~Sana\inst{5}.}
+\author{M.~Renzo\inst{1} \and \todo{TBD}% S.~E.~de~Mink\inst{1} \and
+  % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
+  .}
 
@@
     Sciences, Beijing 100012, China}
-  \and{Astrophysics Research Group, Faculty of Engineering and Physical Sciences, University of Surrey, Guildford, Surrey, GU2 7XH, United Kingdom}
-  \and{Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}}
+}
 
@@
 \date{}
-\abstract{Anticipating the kinematic constraints from
-  the Gaia mission, we perform an extensive numerical study of
-  the evolution of massive binary systems to predict the peculiar
-  velocities that stars obtain when their companion collapses and disrupts the system.
-Our aim is to (1) identify which predictions are robust against model uncertainties and assess their implications, (2) investigate which physical processes leave a clear imprint and may therefore be constrained observationally and (3) provide a suite of publicly available model predictions.
-We find that $22_{-8}^{+26}$\% of all massive binary systems merge
-prior to the first core collapse in the system. Of the remainder,
-$86_{-9}^{+11}\%$ become unbound because of the
-core-collapse. Remarkably, this rarely produce runaway stars (i.e.,
-stars with velocities above $30\,\kms$). These are outnumbered by more
-than an order of magnitude by slower unbound companions, or ``walkaway
-stars’’. This is a robust outcome of our simulations and is due to the
-reversal of the mass ratio prior to the explosion and widening of the
-orbit, as we show analytically and numerically. We estimate a
-$10^{+5}_{-8}$\% of massive stars to be walkaways and only $0.5^{+1.0}_{-0.4}$\% to be runaways, nearly all of which have
-accreted mass from their companion. Our findings are consistent with
-earlier studies, however the low runaway fraction we find is in
-tension with observed fractions $\sim$10\%. If Gaia confirms these high fractions of
-massive runaway stars resulting from binaries, it would imply that we
-are currently missing physics in the binary models. Finally, we show
-that high end of the mass distributions of runaway stars is very
-sensitive to the assumed black hole natal kicks and propose this as a
-potentially stringent test for the explosion mechanism. We also
-discuss companions remaining bound which can evolve into X-ray
- and gravitational wave sources.}
+\abstract{}
 
-\keywords{stars: kinematics, binaries, massive, supernovae }
+\keywords{stars: kinematics, stars:individual:VFTS682}
 \maketitle{}
@@
 
-Stars with initial mass larger than
-about $7.5\,M_\odot$ are the progenitors of black holes (BH) and
-neutron stars (NS). These stars play an important role in shaping galaxies
-through their radiative, chemical, mechanical feedback
-(e.g., \citealt{larson:74, Heger+2003, ceverino:09, lamers:13}, see
-also \citealt{langer:12, nomoto:13} for reviews).
-Most young, unevolved, massive stars have a nearby companion with
-which they form a close binary system
-\citep[e.g.,][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. 
-Binary systems that remain bound throughout the entire evolution of both
-stars can give rise to many exotic phenomena, including X-ray binaries
-\citep[e.g., ][]{gott:71,Bolton1972, Webster+1972,
-  van-den-Heuvel+1972}, binary neutron stars
-\citep[e.g.,][]{hulse:75,wijers:92}, gamma-ray bursts
-\citep[e.g.,][]{izzard:04b,becerra:16, LVC:17:grb}, and gravitational wave events
-\citep{LVC:16b, BNSmerger:17}.  
-However, only a small fraction of massive stars that are born as
-binary systems are expected to stay together their entire lives. The
-majority of systems is disrupted by the first core collapse event,
-which can separate the newly formed compact object from its former
-companion star
-\citep[][]{zwicky:57,blaauw:61, tauris:98, belczynski:99,
-  belczynski:08, eldridge:11}.
+How do stars form is one longstanding question in astrophysics
+\todo{ref?}. Answering it for massive stars is particularly difficult,
+since these are intrinsically rare
+\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
+remain enshrouded in their parent cloud during the whole formation
+process. Moreover, observations of young massive stars reveal a
+complicated multipliticy structure which requires explanation.
 
-Single NSs can be detected as pulsars or magnetars. Many of
-them are observed to have large proper motions (e.g.,
-\citealt{gunn:70,gott:70,lyne:94,kaspi:96,hobbs:05,chatterjee:05}, but see also \citealt{verbunt:17}).
-Conversely, single BHs are only detectable under special
-circumstances, for example through lensing events when passing in
-front of a background star \citep[e.g.,][]{wyrzykowski:16}, or if they
-accrete gas from the ambient medium
-\citep[e.g.,][]{fender:13, gaggero:17}. Therefore, to probe the
-population of stellar-mass black holes we are limited to X-ray or
-gravitational wave observations that can only target the rare
-cases that remain bound to their companion and are close enough for
-the BH to accrete or merge. To learn more about the black holes that
-form in less special cases, it is worth to investigate the imprints
-they may leave on their former companion star.
+Understanding massive star formation, possibly as a function
+of metallicity, is a key question given the present and upcoming
+transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{} which will reveal transients associated to their
+evolution and death.
 
-The primary focus of this study is the population of unbound main sequence companions,
-which can be identified observationally because of their peculiar
-spatial velocities compared to the surrounding population. Their velocities can either be detected as proper
-motions (i.e.,\ their motion in the plane of the sky measured directly
-from the displacement of the star) or as radial velocities (i.e.,\ their
-motion perpendicular to the plane of the sky, measured from the
-Doppler shift of the spectral lines). Large spatial velocities have
-been inferred for a significant sub-population of young massive stars
-\citep[e.g.][]{blaauw:61, Cruz-Gonzalez+1974, gies:86, gies:87,
-  hoogerwerf:00, hoogerwerf:01, tetzlaff:11, boubert:18}.
+\todo{short summary of formation scenarios goes here}.
 
-\citet[][]{blaauw:61} introduced the term ``runaway stars'' for those %stars
-in the fast tail of the velocity distribution for a given spectral type. The typical threshold
-adopted to define the tail of this distribution for O and B-type stars is $v \gtrsim 30\,\mathrm{km \ s^{-1}}$
-\citep[e.g.,][]{blaauw:56, gies:86, dedonder:97, hoogerwerf:00, hoogerwerf:01, dray:05, eldridge:11}, although sometimes other values
-have been considered \citep[e.g., $40\,\mathrm{km\ s^{-1}}$
-in][]{blaauw:61, dewit:05, boubert:18}.
-However, as we will argue based on simulations presented in this work, the majority of unbound companions is expected to exhibit velocities well below these thresholds.
-We will refer to these slow unbound former companions resulting from
-disrupted binary systems as ``walkaway stars'' \citep[to our knowledge first coined
-by][]{de-Mink+2012} to distinguish them from the faster counterparts.
+The second data release (DR2) from the Gaia satellite
+\cite[][]{brown:18,} allows us to test
+these hypothesis using one particular star, VFTS682. This star is a
+very massive ($\sim150\,M_\odot$) WNh5 star in the 30 Doradus region
+of the Large Magellanic Cloud, and it is presently observed at a
+projected distance of $\sim$$29$\,pc from the nearest cluster of
+massive stars R136 \citep[][]{bestenlehner:11}. Based on the extremely high mass of this star and
+its present day apparent isolation, \cite{bestenlehner:11} proposed it
+might be a candidate for isolated star formation, if it has not been
+ejected from R136 in the past.
 
-A proposed explanation for the large spatial velocity of runaways is that
-they originate from disrupted binary systems \citep[]{zwicky:57,
-  blaauw:61, boersma:61},
-which naturally explains the lower number of companions they have,
-compared to typical massive stars \citep[e.g.,][]{blaauw:61, gies:86, sana:14}. \Figref{fig:cartoon} sketches the
-typical evolution of such a binary system. Most binaries are disrupted
-at the time of the first core-collapse.
+In this study, we combine the radial velocity measurements from the
+VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2
+to reconstruct the three-dimensional velocity of VFTS682, and test the
+null hypothesis that this star was ejected from R136. \todo{check the
+  following} Our results indicate that R136 is the likely origin of
+this star, and therefore isolated star formation is \emph{not}
+required to explain it. However, we find a discrepancy between the
+apparent age of VFTS682, it's kinematic age, and the age estimates for
+R136.
 
-An alternative mechanism to produce stars with peculiar spatial
-velocities is dynamical ejection from a star cluster \citep[e.g.,][]{poveda:67, leonard:91, Evans+2010a,
-  fujii:11, allison:12, oh:16}. Dynamical interactions with
-a supermassive BH can also disrupt a binary, but the ejection
-velocities achieved in this scenario are typically much higher
-\citep[$\gtrsim$$10^3\,\mathrm{km\ s^{-1}}$,][]{hills:88}.
+In \Secref{sec:sample} we describe how we select stars in R136 and in
+the surroundings of VFTS682 itself to remove the mean motion of the
+region. \Secref{sec:results} presents our main findings. We conclude by
+discussing the implications for theories of
+star formation, N-body interactions, and binary evolution in
+\Secref{sec:discussion}.
 
-Both mechanisms, i.e. the ``binary
-disruption scenario'' and the ``dynamical ejection scenario'', are
-expected to act in nature, but their relative importance is not not well
-constrained \citep[e.g.,][]{hoogerwerf:00,hoogerwerf:01,guseinov:05}.
-\citet{hoogerwerf:01} analyzed the properties of a sample of 56 nearby
-runaway stars and 9 radio pulsars, and traced back the runaways to
-their most likely parent stellar group. From the sub-sample for which a
-clear identification of the parent group was possible, they estimated that the disruption of binaries is
-responsible for roughly two thirds of observed runaways \citep[see
-also][]{gies:86, gies:87, stone:91, hoogerwerf:00, dincel:15,
-  boubert:17b}. However, this claim could not be confirmed by
-\cite{jilinski:10}.
+\section{Gaia DR2 data selection}
+\label{sec:sample}
+VFTS682 is labeled in the Gaia DR2
+catalog\footnote{\url{https://gea.esac.esa.int/archive/}} with the
+source id 4657685637907503744. For this star, we retrieve from the
+Gaia catalog its position in right ascension (RA) and declination (DE)
+in the IGCS frame\cite[][]{brown:18}\todo{some 2016 Gaia paper?}, its
+proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DE}$,
+respectively). For the radial velocity of VFTS16 and of the 30 Doradus
+region as a whole, we instead use the VFTS data
+as quoted in \cite{bestenlehner:11}. Our main conclusions do not
+depend on the detailed value of the radial velocity.\todo{improve
+  phrasing} The values of all these quantities that we use are listed
+in \Tabref{tab:vfts682}.
 
-Unbound stars resulting from the disruption of a binary are of
-potential interest for several topical questions in
-astrophysics. Their kinematics and stellar properties bear imprints of the uncertain physical processes that govern the evolution of their
-binary progenitor systems, including the expected phase of mass
-transfer between the two stars (phase B. in
-\Figref{fig:cartoon}). Detailed knowledge of the spatial velocity of
-runaway stars can improve the accuracy of wind mass-loss rate
-determinations relying on their bow shocks \citep[e.g.][]{gull:79,
-  kobulnicky:18}. Of particular interest is the
-question whether they can provide any unique constraints on the physics of
-core-collapse, in particular the natal kick on the compact objects
-that they produce. Such kicks are expected either from asymmetries in
-the explosion and/or the neutrino emission
-\citep[][]{shklovskii:70,kusenko:96,fryer:04,socrates:05,wongwathanarat:13,
-  janka:13,janka:17}, and they determine which systems disrupt and
-eject the companion star and which systems remain bound and thus have a
-potential as future X-ray and gravitational wave sources.
+\begin{table}[tbp]
+  \centering
+  \begin{tabular}[htbp]{l|c|c}
+    Parameter & Value & Source\\ \hline\hline
+    RA \hfill[degree] &  84.73 $\pm$  0.036 & \multirow{4}{*}{Gaia DR2}\\
+    DE \hfill [degree] & -69.07 $\pm$  0.05  & \\
+    $\mu_\mathrm{RA}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 1.84 $\pm$ 0.07 & \\
+    $\mu_\mathrm{DE}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.78 $\pm$ 0.08& \\
+    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & -45 $\pm$ 25 & \cite{bestenlehner:11}\\
+    \hline
+  \end{tabular}
+  \caption{Astrometric parameters for VFTS682}
+  \label{tab:vfts682}
+\end{table}
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{./figures/cartoon}
-  \caption{Schematic depiction of the evolution of a close massive
-    binary through stable mass transfer. The
-    evolution forks at the first core-collapse event in the system: the vast
-  majority of the systems is disrupted and produces a runaway or
-  walkaway star which can travel far from its birth location. The
-  systems remaining bound are possible progenitors of X-ray binaries
-  and gravitational wave sources. The fraction of binaries disrupted
-  $\mathcal{D}$ comes from the simulations presented in this study.
-}
-  \label{fig:cartoon}
-\end{figure}
+To compare the astrometry of VFTS682 and derive its peculiar motion,
+we then select data from the Gaia DR2 catalog for two regions: the
+``surroundings'' of VFTS 682, and the ``R136 cluster''. The
+surrounding region is defined by all the stars within 10 arcminutes from
+VFTS682. This includes the subset of stars which we consider to be
+part of R136. We include all the stars for which a value of the proper
+motion components and the corresponding errors are available.
+The ``R136 cluster'' is effectively defined by taking all the stars
+within 25 arcseconds from R136a, one of the most massive members of
+the cluster itself \citep[][]{crowther:10}.
 
-These unbound stars have also been considered for their potential
-importance as non-canonical sources of stellar feedback
-\citep[e.g.,][]{larson:74, ceverino:09}. 
-They can travel long distances and end their lives tens to hundreds of
-parsecs away from their birth location. Because of their motion, their
-ionizing photons are
-less likely to be absorbed by their birth clouds. Therefore, in the context of the
-re-ionization of the early universe, the ionizing radiation of stars
-ejected from a binary are more likely to
-escape and become available for ionization of hydrogen in the
-intergalactic medium \citep[e.g.][]{conroy:12}. This has been
-suggested to be of importance in the 
-context of the epoch of reionization
-\citep[e.g.,][]{conroy:12,kimm:14,Ma+2016}. Furthermore they are
-expected to explode in lower density regions, which may enhance the
-impact they have as sources of turbulence in the ISM \citep[e.g.][]{gatto:15}.
+Throughout this study, we assume the same distance of $50$\,kpc to the star, and to
+the 30 Doradus region as a whole, since the parallax for VFTS682
+listed in the Gaia DR2 catalog is negative.
 
-The present and upcoming second Gaia data releases provide an important
-observational motivation for this study \citep[e.g.,][]{perryman:01}. The determination of
-distances, proper motions and radial velocities by the Gaia satellite
-is expected to drastically increase the available sample of massive
-stars with precisely known velocities
-\citep{maoileidigh:05,  cacciari:09, Jeffries+2014,
-  Gaia-Collaboration+2016, Gaia-Collaboration+2016a}.
-
-In this study, we present a systematic theoretical study of the
-kinematical signatures of stars ejected from binary systems and how
-these depend on the uncertain physical processes governing binary
-stellar evolution. For this purpose we employ a rapid binary
-population synthesis code, which we have updated to account for the main relevant processes that affect the disruption of the binary systems (see \Secref{sec:methods}).
-
-We first provide insight in our simulations by presenting the results
-for an individual system in \Secref{sec:example}.
-We present in \Secref{sec:analytics} an analytic estimate of the
-typical and maximum velocity of ejected companions expected from the
-disruption of binaries. Then, in \Secref{sec:fiducial_res} we describe the results of a fiducial simulation of a
-full population of binary stars, for which we adopt realistic
-input assumptions provided by detailed simulations and/or observations
-when available.  In addition, we present in
-\Secref{sec:param_variation} an extensive grid of models where we vary the most relevant uncertain assumptions.
-This allows us to study (i) which predictions are robust against the model uncertainties and (ii) which uncertain physical processes leave a clear imprint on the observables that can be used to test and constrain them.
-
-Various earlier studies have discussed the evolution and
-interaction of populations of massive binary systems. Examples
-include, but are not limited to, \citet[][]{vanbeveren:82, dedonder:97, Fryer+1997, Fryer+2001, fryer:12,
-  Belczynski+2012b, repetto:12, Fragos+2013a, Grudzinska+2015,
-  bray:18, boubert:18}. Most
-of these studies focus on the minor fraction of systems
-that remain bound and are the progenitors of X-ray binaries and/or
-gravitational wave sources. A few studies focus specifically on
-disrupted systems and unbound stars. Examples are
-\cite{vanrensbergen:96,dedonder:97, dray:05, eldridge:11, bray:16, boubert:17b,
-  zapartas:17b, boubert:18}. We expand on these studies by using updated physical
-assumptions, and by focusing on which physical processes could be
-constrained using the Galactic population of unbound stars. A
-comparison to previously published results is given in \Secref{sec:literature}.
-
-Our first main result is that the majority of disrupted binaries
-ejects a slow-moving walkaway star ($v < 30\,\mathrm{km\ s^{-1}}$), rather than a faster runaway. This result is robust against
-variations of uncertain parameters in the model, as we discuss in \Secref{sec:param_variation}.
-This would imply that the more easily detected runaway stars only
-reveal a small subset of the population of unbound stars that are
-former companions of disrupted binary systems.  We discuss the implications in \Secref{sec:discussion}.
-
-Our second main result is that the kinematic properties and absolute
-number of unbound companions depend sensitively on assumptions
-concerning the  natal kicks of BHs. We show the imprints this leaves on
-the mass distribution and discuss whether future observations can be
-used to constrain these processes.
-
-Section~\ref{sec:bound} briefly describes the population of systems remaining bound after the first
-core-collapse.
-
-We also find that the fraction of runaways among massive stars
-predicted by our simulations is much lower than the observed value in
-all our parameter variations. This finding agrees with the results of
-\cite{eldridge:11}, but is in potential contrast
-with the observational results of \cite{hoogerwerf:01}, and needs further
-investigation.
-
-Finally, we discuss how our results  may be useful to
-include the effect of unbound binaries in models for stellar
-feedback. We will provide our numerical results upon publication. %provide our numerical results at \todo{where?}\todo{zenodo/CDS?}.
-
-\section{Binary population synthesis calculations}
-\label{sec:methods}
-
-We carry out population synthesis calculations of isolated binaries with the rapid binary evolution code \texttt{binary\_c}
-\citep[][]{izzard:04,izzard:06,izzard:09,demink:13,schneider:15,izzard:18}. This
-code is based on the algorithms by \cite{tout:97,hurley:00,hurley:02}, which rely on the analytic fits to the
-single stellar evolution models from \cite{pols:98}.
-
-First, we compute a fiducial population using observationally
-validated assumptions for the free parameters that describe the initial
-conditions and physical assumptions.
-Then, we check the robustness of our results (or
-equivalently, to which assumptions they are most sensitive) by varying
-the free parameters that are most relevant for the
-velocity distribution of disrupted binaries. To limit
-computational costs, we explore variations in each parameter
-one-by-one while keeping the other parameters fixed to our
-fiducial choices, following the approach of
-\cite{Fragos+2013a,demink:13,demink:14,zapartas:17,zapartas:17b,
-  belczynski:17}. Effectively, we treat
-each parameter in our models as independent from the others, therefore this approach does not account for possible
-correlations between either initial distributions \citep[e.g., the one between initial period and mass ratio
-suggested by][]{moe:17} or uncertain
-physical processes (e.g., core spin and natal kick amplitude). Varying
-two (or more) parameters simultaneously, or changing the algorithmic
-representation of the uncertain physical processes, might possibly result in
-larger variations than those we present in
-\Secref{sec:param_variation}.
-
-\subsection{Initial distributions and parameters}
-
-Each binary system in our calculations is characterized by a zero age
-main sequence (ZAMS)
-mass for the primary\footnote{Throughout this study, we define the primary star
-  to be the initially more massive star, even
-  if it becomes the less massive star in the binary during the evolution.}
-$M_1^\mathrm{ZAMS}$, initial mass ratio $q\udef M_2/M_1$, and initial orbital
-period $P^\mathrm{ZAMS}$. We build a grid in this parameter space and
-weigh each system in our grid according to the initial distributions
-described below.
-
-We select $N_{M_1}$ primary stars with ZAMS mass $M_1^\mathrm{ZAMS}$
-at logarithmically spaced intervals in the range $7.5\,M_\odot\leq M_1^\mathrm{ZAMS}
-\leq 100\,M_\odot$. We weigh each primary star with an initial mass function (IMF) with
-slope $\alpha=-2.3$ 
-\citep[][]{kroupa:01}. In our model
-variations, we explore values of $\alpha$ of -1.9
-\citep[][]{schneider:18} and -3.
-
-For each primary star of mass $M_1^\mathrm{ZAMS}$, we select $N_q$ secondaries with mass
-$M_2^\mathrm{ZAMS} = qM_1^\mathrm{ZAMS}$, taken at regular intervals
-in mass ratio $q$ between 0.1 and 1 assuming a
-flat distribution \citep[slope $\kappa=0$, e.g.,][]{kouwenhoven:05,sana:12}. We consider in
-\Secref{sec:param_variation} also variations with $\kappa=\pm1$.
-
-Finally, for each pair of masses$(M_1^\mathrm{ZAMS},M_2^\mathrm{ZAMS})$ we choose $N_p$ different periods
-$P^\mathrm{ZAMS}$ equally spaced in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$
-between 0.15 and 5.5. We weigh the birth probability of each binary system with a
-mass-dependent distribution for the initial orbital period: if $M_1<15\,M_\odot$ we assume
-a flat distribution in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$
-\citep[][]{opik:24,kobulnicky:07}; while for $M_1\geq15\,M_\odot$ we assume a
-powerlaw distribution in $\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})$ with exponent
-$\pi=-0.55$, \citep[][]{sana:12}. We explore values of $\pi=0\,$and\,$\pi=-1$
-(for all values of $M_1^\mathrm{ZAMS}$)
-in our model variations. We chose our upper-limit on the initial period, $10^{5.5}$\,days, to include
-wide systems in which the stars effectively evolve as single stars \citep[][]{demink:15}.
-
-To limit the dimensions of the parameter space, we assume
-all orbits to be initially circular (i.e., eccentricity
-$e=0$). Because we focus on the population of stars resulting from the
-disruption of post-interaction binaries, this assumption is not
-critical \citep[][]{demink:15}: orbits are expected to
-circularize
-because of tides before or during mass transfer
-(\citealt{belczynski:99,hurley:02}, but see also \citealt{eldridge:09}
-for arguments against circularization of the orbit in post-main
-sequence mass transfer).
-
-In our fiducial model, we assume a canonical metallicity (i.e., mass fraction of
-elements heavier than helium) $Z=0.02$. This value is slightly above
-the most recent determination for solar neighborhood
-\cite{asplund:09}, but it should generally describe young population
-of massive stars in the Milky-Way. The models from \cite{pols:98}, which serve as
-input for our computations, adopt the
-isotopic mixture of \cite{anders:89}. In
-\Secref{sec:param_variation}, we also consider $Z=0.0002$, $0.0047$, $0.008$,
-and $0.03$.
-
-We initialize the stellar rotation rate with a mass-dependent
-equatorial velocity according to \cite{hurley:00}, assuming
-alignment between the stellar spins and the orbital angular momentum. In one of our
-parameter variations, we draw the initial spin velocity randomly from
-the distribution given by \citealt{ramirez-agudelo:15} (R15, see last
-line in \Tabref{tab:parameters}), but maintain the assumption of
-alignment. However, changing the initial rotation rate has almost no
-impact on our results because binary interaction processes overwrite it \citep[][]{demink:13}.
-
-We build a grid in the initial parameter space, and initialize a
-binary system in each cell of such grid. 
-Each system is assigned a ``probability''
-which can then be multiplied by the
-amount of time spent by the system in that cell to obtain a quantity comparable to
-observed number counts. See Appendix~\ref{app:data} for more details.
-
-For each set of parameters, the total size of our model grid is
-$N_{M_1}\times N_q \times N_p = 50 \times 50 \times 100 = 250\,000$ 
-binary systems\footnote{This number is then multiplied by the number
-  of natal kicks that we draw for each core-collapse, see \Secref{sec:SN_kick}.}.  Increasing this resolution to
-$432\,000$ introduced no
-significant variation in our results.
-
-\subsection{Physical assumptions}
-
-We follow the evolution of each binary system until the
-first core-collapse (CC) event or until they merge. 
-Stellar winds can have a dramatic effect on the evolution of single
-stars \citep[e.g.,][]{renzo:17} and impact significantly the orbital
-evolution in a binary. We include wind mass loss in our single star models as implemented by
-\cite{demink:13}, i.e.\ we use a combination of mass loss rates from
-\cite{vink:00,vink:01} for hot hydrogen-rich stars,
-\cite{nieuwenhuijzen:90} for cool stars, % (or for luminosities larger than 4\,000$L_\odot$ if the mass loss rate
-% is larger than what suggested by \cite{vink:00,vink:01})
- including a ``luminous blue variable'' enhancement as in
- \citealt{hurley:00}, and from \cite{hamann:98} reduced by a factor of 10 (for
- Wolf-Rayet stars. We also include the wind mass loss
- enhancement for fast rotating stars following \cite{maeder:00}.
-
-We include the effects
-of tides on the spins and orbital angular momentum using the
-algorithm from \cite{hurley:02} based on the calculations of \cite{zahn:77,hut:81}.
-
-When the radius of the most massive star exceeds the Roche radius,
-(calculated using \citealt{eggleton:83} fitting formula)
-we use the algorithm of \cite{claeys:14} to determine the
-mass transfer rate (see their Eq.~10).
-
-We limit the accretion rate of the
-accretor to 10 times its Kelvin-Helmholtz timescale, i.e., we assume a
-mass transfer efficiency $\beta_\mathrm{RLOF}\equiv |
-\dot{M}_\mathrm{acc}|/|\dot{M}_\mathrm{don}|\equiv
-\beta_\mathrm{th}$, with $\dot{M}_\mathrm{don}$ and
-$\dot{M}_\mathrm{acc}$ temporal variation of the donor and accretor mass, respectively. Larger rates are
-likely to drive the accretor out of thermal equilibrium\footnote{We do
-not model the internal structure of the stars, therefore, we do not
-follow its possible bloating during mass transfer, which could
-potentially enhance its mass loss and related spin down.} and lead to
-unstable mass transfer \citep[][]{neo:77,hurley:02}. However, for a
-large portion of the parameter space, $\beta_\mathrm{th}$ results in a
-rather conservative mass transfer \citep[][]{schneider:15}. Since the
-efficiency of mass transfer is a major uncertainty in binary evolution
-\citep[][]{demink:07}, we also consider the extreme parameter
-variations with $\beta=0,\,0.5,$ and 1, respectively.
-
-Matter that is not accreted is assumed to leave the system with
-the specific angular momentum of the orbit of the accretor
-\citep[e.g.,][]{soberman:97,vandenheuvel:17}, $h = \gamma_\mathrm{RLOF}
-J_\mathrm{orb}/(M_\mathrm{don}+M_\mathrm{acc})$,
-with $\gamma_\mathrm{RLOF}=M_\mathrm{don}/M_\mathrm{acc}$. In our
-model variations, we also explore a scenario where the mass that is not
-accreted leaves the binary from the outer Lagrangian point L2 and is assumed to form a circumbinary disk \citep[i.e.,
-$\gamma_\mathrm{RLOF}=\sqrt{2}(M_\mathrm{don}+M_\mathrm{acc})^2/(M_\mathrm{acc}M_\mathrm{don})\equiv
-\gamma_\mathrm{disk}$,][]{artymowicz:94}. We further consider the
-assumptions that the mass that is not accreted leaves the system with the
-orbital specific angular momentum, i.e.~$\gamma=1$, which is the standard model of \cite{belczynski:08, dominik:12, dominik:13}.
-
-%qcrits
-If, by the time the first star fills its Roche lobe, the mass
-ratio of the accretor to the donor is smaller than a certain threshold
-($M_\mathrm{acc}/M_\mathrm{don}<q_\mathrm{crit}$), we assume that the
-system enters a common envelope phase. The threshold value
-$q_\mathrm{crit}$ is uncertain and
-depends on the evolutionary stage of the donor. In our fiducial simulation $q_\mathrm{crit,A}=0.65$ for a
-main sequence (MS) donor \citep[][]{demink:07}, and $q_\mathrm{crit,B}=0.4$ for a Hertzsprung gap
-donor \citep[][]{hurley:02}.
-We also consider values of $q_\mathrm{crit,A}=0.25,\,0.8$ and $q_\mathrm{crit,B}=1,\,0.5,\,0$
-in our model variations.
-
-We treat common envelope evolution using the $\alpha_{\rm CE}
-\lambda$-formalism \citep[][]{webbink:84, livio:88, dekool:90, hurley:02}. In all
-variations, we use an analytic fit to the $\lambda_{g}$ values of
-\cite{dewi:00} for the binding energy parameter $\lambda$. Those values do not include the energy stored in thermal
-motions and the ionization state of the material within the
-envelope. It is unknown what fraction of that internal energy is
-useful in unbinding the envelope \citep[see, e.g.,][and references
-therein]{dewi:00, ivanova:13}.  We assume $\alpha_{\rm CE} = 1$ for
-our fiducial simulation, i.e., perfectly efficient use of the
-liberated gravitational potential energy from the orbit but without
-additional energy sources.  We explore variations with  $\alpha_{\rm
-  CE} = 0.1$, i.e., inefficient use of the orbital energy to eject the
-common envelope, and $\alpha_{\rm CE} = 10$, intended as indicative of
-a fairly extreme case of additional energy input. For
-binaries that interact while neither stars have a well defined core-envelope structure, such as
-MS stars and helium-MS stars, we assume that the common envelope
-phase leads to a merger as described in \cite{demink:13, schneider:15}.
-
-\subsubsection{Natal kick}
-\label{sec:SN_kick}
-
-When a CC event happens in a binary system three different physical
-ingredients contribute to possibly disrupting the system and ejecting the
-companion star.
-
-\begin{enumerate}
-\item[i.] The orbit is modified by the sudden change in gravitational potential, because
-of the mass lost through ejecta \citep[the so-called ``Blaauw
-kick'',][]{zwicky:57,blaauw:61, boersma:61}. However, this is not the
-dominant effect, because the typical evolution of a massive close binary
-system involves mass transfer through Roche lobe overflow (RLOF) before the first CC (cf.~\Figref{fig:cartoon}). RLOF
-removes the envelope of the initially more massive star, limiting
-severely the amount of mass that can be ejected by the SN. The
-``Blaauw kick'' alone rarely unbinds the system \citep[][]{huang:63,tutukov:73,leonard:94}.
-\item[ii.] The SN shock can interact with the companion star.
-  \citep[e.g.,][]{wheeler:75,liu:15, rimoldi:16, hirai:18}. The shock can dynamically
-  remove mass from the companion (stripping), heat the envelope of the
-  secondary, thus enhancing its own mass
-  loss (ablation), and deposit mass and momentum on the secondary. The injection of energy can inflate the companion and make
-  it look redder for a duration comparable to its thermal timescale. Our
-  treatment of accretion only changes the mass of the companion, without
-  checking for structural readjustments of the star. However, the
-  interaction between the ejecta and the secondary is typically a small effect \citep[e.g.,][]{liu:15}.
-\item[iii.]  The natal kick of the compact object changes %instantly
-  its kinetic energy and momentum. This kick is caused by asymmetries
-  in the SN ejecta and/or neutrino flux at the explosion, possibly
-  seeded by the late core and shell burning phases of stellar
-  evolution \citep[see e.g.,][]{shklovskii:70,kusenko:96,fryer:04,socrates:05,wongwathanarat:13,
-    janka:13,janka:17,holland-ashford:17,grefenstette:17,katsuda:18}. It
-is typically parametrized using a kick velocity $\mathbf{v}_k$, drawn
-from a distribution of amplitude $v_k\equiv |\mathbf{v}_k|$ and directions (see below). The natal
-kick is the dominant reason for the disruption of binaries.
-\end{enumerate}
-
-We have updated the treatment of binary disruptions by a CC event in
-\texttt{binary\_c}, following \citealt{tauris:98} (TT98 hereafter). To re-calculate the post-CC
-orbital parameters, their algorithm
-assumes instantaneous
-loss of the ejecta, because the ejecta velocity $v_\mathrm{ej}$ is much larger than the
-orbital velocity $v_\mathrm{orb}$ ($v_\mathrm{orb} \lesssim 10^2\,\mathrm{km \
-  s^{-1}} \ll v_\mathrm{ej}\sim10^4\,\mathrm{km \
-  s^{-1}}$). This algorithm considers the motion of the compact
-object within the pre-CC orbit, and the interaction between
-the ejecta and the companion star. 
-We set the
-relative change of mass 
-of the secondary due to stripping, ablation, and
-accretion of SN ejecta following a fit to the three-dimensional
-hydrodynamical simulations
-of \cite{liu:15} for the impact of the ejecta on a
-$M_2=3.5\,M_\odot$ star, which is the most massive companion they
-considered. We also assume an efficiency of momentum
-transfer\footnote{$\eta=0.5$ in the algorithm from \citetalias{tauris:98}.} from the
-ejecta to the companion star of 0.5 (\citetalias{tauris:98};\citealt{hirai:18}). The contribution of 
-the interactions between the SN ejecta and the secondary star to the total natal kick is
-typically small, $\lesssim10\,\mathrm{km \ s^{-1}}$
-(\citetalias{tauris:98}; \citealt{liu:15}; \citealt{hirai:18}).
-
-In our fiducial run, the kick direction is assumed to be isotropically
-distributed in the frame of the collapsing star \citep[][]{wongwathanarat:13,
-  bear:17}, although  \cite{johnston:05} and \cite{kaplan:08} suggest that there
-is weak evidence for kick-spin alignment in well observed pulsars
-such as the Crab or Vela. \cite{johnston:05} also underlined that the
-direction of the pulsar spin does not necessarily match the direction
-of the pre-CC spin of the stellar core, if significant torques
-develop during the CC. In our model
-variations, we also consider kicks constrained in a cone with an
-opening angle of
-$\alpha=10\,\mathrm{degrees}$ oriented along the spin of the exploding
-star (which we assume to be perpendicular to the orbital plane), and
-kicks constrained at angles from the orbital plane $90 - \alpha
-\leq 45\,\mathrm{degrees}$. The kick amplitude $|\mathbf{v}_k| \equiv
-v_k$ is drawn from a Maxwellian distribution with one-dimensional root
-mean squared dispersion $\sigma_\mathrm{kick}=265\,\mathrm{km \  s^{-1}}$ \citep[][]{hobbs:05}. Such
-distribution is motivated by the observation of the line of sight velocities of pulsars in
-\cite{hobbs:05} (see also \citealt{lyne:94}). 
-We also compute
-populations with $\sigma_\mathrm{kick}=0,\,300,\,1000\,\mathrm{km\
-  s^{-1}}$ in our model variations.
-
-The value of $v_k$ drawn is then reduced to consider the amount of matter that 
-falls back after the successful launch of the SN shock,
-i.e.~$v_k \rightarrow v_k(1-f_b)$, were $f_b$ is the fallback
-fraction taken from\footnote{We correct for a missing parenthesis in
-  their equations, which can be found by dimensional analysis \citep[][]{belczynski:17}.} Eq.~16 in \cite{fryer:12}, corresponding to their
-``rapid SN engine'' which reproduces the NS-BH mass gap between
-$\sim2-5\,M_\odot$ (e.g., \citealt{farr:11}, however see also \citealt{wyrzykowski:16}
-regarding the existence of this mass gap). 
-The inclusion of fallback
-also determines the mass of the compact remnant obtained after each
-CC event, and we set the mass boundary between NSs and BHs to
-$2.5\,M_\odot$.
-
-The fallback fraction is highly uncertain in BH formation. The algorithm we use
-here prescribes a large amount of fallback, implying close to zero
-BH kick amplitudes, sometimes referred to as a ``BH momentum kick'' \citep[][]{belczynski:08,stevenson:17}. Whether
-this is realistic is subject of debate in the literature. Evidence for
-non-zero BH natal kicks comes from the observed
-Galactic latitude of BH X-ray binaries
-(e.g., \citealt{fragos:09,repetto:12,repetto:15,repetto:17}, although
-see also \citealt{mandel:16}), the
-possibility of retrograde BH spin \citep[e.g.,][]{morningstar:14}, eccentric orbits in these systems
-\citep[e.g.][]{remillard:06}, the
-small number of WR+BH binaries \citep[e.g.,][]{dray:05}, and most
-recently, the gravitational wave constraints on
-BH spins \citep[e.g.,][]{oshaughnessy:17,wysocki:18}. Recently,
-multi-dimensional calculations of fallback in CC resulting in the
-formation of a BH showed that large natal kicks might be possible \citep[][]{chan:18}. Therefore, we
-also consider a model without the rescaling of the kick amplitude, but still
-including fallback for the remnant mass calculation. This variation is
-effectively equivalent to a so-called ``velocity kick'' for the
-BHs. To test an ``intermediate'' BH kick, we run a simulation using
-$\sigma_\mathrm{kick}=100\,\mathrm{km\ s^{-1}}$ and no fallback
-down scaling of the kick amplitude for
-the CC resulting in BH formation, and out fiducial kick
-($\sigma_\mathrm{kick}=265\,\mathrm{km\ s^{-1}}$, including fallback scaling)
-for CC resulting in NS formation.
-
-Another debated issue is whether the CC of low-mass iron
-(or oxygen-neon-magnesium) cores produces small natal
-kicks, possibly resulting in a bi-modal kick distribution
-\citep[][]{katz:75, arzoumanian:02, pfahl:02,
-  podsiadlowski:04, martin:09, knigge:11, wongwathanarat:13,
-  beniamini:16, tauris:17, verbunt:17, verbunt:17b}. To
-explore the possibility of low kicks for low mass collapsing cores, we
-run one model variation reducing the natal kick for NSs less massive
-than $1.35\,M_\odot$ \citep[][]{schwab:10,knigge:11} drawing
-the natal kicks for these NSs from a Maxwellian with
-$\sigma_\mathrm{kick}=30\,\mathrm{km\ s^{-1}}$, effectively creating a
-double peaked distribution. For more massive NSs and BHs the kick is drawn from a
-Maxwellian with $\sigma_\mathrm{kick}=265\,\mathrm{km\ s^{-1}}$.
-
-In all our simulations, we draw $N_\mathrm{kick}=20$ intrinsic natal
-kick directions and amplitudes for each CC in our population, effectively increasing the size of our sample by this
-factor. Convergence tests showed no significant variations
-of our results increasing $N_\mathrm{kick}$ to 50.
-
-\subsection{Caveats}
-\label{sec:caveats}
-In addition to the limitations discussed  earlier in this section there are a few more caveats that should be kept in mind when comparing our simulations with observations (aside from the possible effects of biases).
-
-We only consider runaways and walkaways resulting from the disruption of binary systems at the first stellar collapse.
-Observational samples of runaway stars also include another sub-population, coming from cluster ejection \citep[e.g.,][]{poveda:67, leonard:91, fujii:11,  allison:12, perets:12, banerjee:12, oh:16}.
-Similarly, walkaway stars produced by the disruption of binaries will add to slow moving stars generated by the dissolution of clusters which did not go through binary interactions, see e.g.~\cite{allison:12}.
-In the Discussion we consider how spectroscopic measurements may allow to disentangle binary products from the outcome of purely dynamical processes.
-
-All the calculations presented here implicitly assume a constant star formation history (SFH) for a duration longer than the longest stellar lifetime of interest.
-However, a recent changes in the the SFH of a region will affect the ratio of runaways and walkaway stars over the number of normal stars.
-For example, a recent increase in the SFH leads to an increase of systems that did not yet have time to interact.
-For a binary to be disrupted, a CC event is needed, and it can only happen after the lifetime of the primary star is over.
-This  introduces a delay corresponding to at least the lifetime of the most massive stars \citep[$\sim$3\,Myr,][]{zapartas:17}, between a SFH event and the first binary disruption.
-A recent increase in the SFH can therefore decrease the fraction of runaway and walkaway stars in a population.
-
-Finally we warn for the possible effects of stochasticity \citep[][]{eldridge:12}.
-We have aimed to simulate sufficient binary systems such that are results are converged, but it should  be kept in mind that the fastest runaways come from relatively rare channels.
-In a relatively low mass stellar population will not fully sample these rare events and stochastic effects can be large.
-
-\section{Example of the evolution of a binary system}
-\label{sec:example}
-
-\begin{figure*}[!htbp]
-  \centering  
-\includegraphics[width=\textwidth]{./figures/evol_example}  
-  \caption{Evolution of an example binary system with 
-    $M_1^\mathrm{ZAMS}=20\,M_\odot$, $M_2^\mathrm{ZAMS}=15\,M_\odot$
-    binary with $P^\mathrm{ZAMS}=100\,\mathrm{days}$, described in
-    \Secref{sec:example} (see also the cartoon in \Figref{fig:cartoon}). The top panel shows the
-    primary, secondary, and total mass. The bottom panel shows the
-    orbital velocity of the secondary (left
-    y-axis) and period of the binary (right y-axis). The labels
-    indicate the following phases: A. the MS of the primary, B. case B
-    RLOF, emphasized by the vertical dashed line, C. the binary is
-    made of a He star (possibly looking like a WR) and a
-    rejuvenated MS secondary; D. (and the vertical dot-dashed line)
-    marks the explosion of the primary in a SNIb/Ic which unbinds the
-    NS and its companion. During the phase E1. the secondary is a slow
-    moving walkaway star, traveling for a duration of about half the
-    main sequence lifetime of the primary.}
-  \label{fig:example}
-\end{figure*}
-
-We first describe the typical evolution of a binary system that
-produces an unbound companion as a result of the disruption of the
-system by the first CC event. The aim is to provide the reader with
-some insight in our simulations. As a representative example, we consider
-a  $M_1^\mathrm{ZAMS}=20\,M_\odot$ primary star,
-a $M_2^\mathrm{ZAMS}=15\,M_\odot$ secondary with
-$P^\mathrm{ZAMS}=100$\,days (corresponding to a separation
-$a_\mathrm{ZAMS}\simeq300\,R_\odot$). \Figref{fig:example} shows the
-evolution of the masses, orbital period, and orbital (and ejection,
-after the CC of the primary) velocity of the secondary
-as a function of time. The labels on the top axis correspond to the phases depicted in the cartoon shown in \Figref{fig:cartoon}.
-
-\subsection{Binary evolution until first CC}
-\label{sec:example_evol}
-
-Our example system remains detached during the MS evolution
-of the primary (phase A. in \Figref{fig:example}). Because of wind mass loss, the period increases by
-about $7\%$.
-
-After $\sim$8.8\,Myr, the primary leaves
-the MS, and starts expanding on a thermal timescale as it burns hydrogen
-in a shell. Shortly after, at the vertical dashed line B. in \Figref{fig:example}, the primary fills its Roche lobe
-initiating case B
-RLOF \citep[][]{kippenhahn:67}.
-For a system which has initial mass ratio close to one
-($q_\mathrm{ZAMS}=0.75$ in this example),
-RLOF is almost conservative in our calculations, i.e.~nearly all the
-mass lost by the primary is accreted by the companion (cf.~\citealt{schneider:15}). The
-secondary star accretes  $\Delta M_2\simeq14\,M_\odot$, and becomes the
-more massive star, now with a mass of about $28\,M_\odot$ (red line in the top
-panel of \Figref{fig:example}). It is also spun up to
-near to breakup rotation \citep[][]{packet:81,demink:09}, and it is rejuvenated because of the increased
-mass and consequent growth of its convective core
-\citep[][]{hellings:83,schneider:16}. The primary  becomes
-a $\sim$$5.3\,M_\odot$ helium star \citep[][]{gotberg:17, gotberg:18}, and
-the orbits widens during mass transfer, reaching a period
-longer than 600\,days (phase C. in \Figref{fig:example}).
-
-After about $10$\,Myr, the primary reaches CC (dot-dashed vertical line
-D. in \Figref{fig:example}). Just before the core-collapse
-of the primary, the orbit is circular with a period of
- about 700\,days. The secondary has a mass of about $28\,M_\odot$ and a pre-SN orbital velocity of
-about $v_2 = (M_1/(M_1+M_2))v_\mathrm{orb} \simeq 12.5\,\mathrm{km\ s^{-1}}$.
-
-If the system is disrupted at the first CC event, the final spatial velocity of the
-secondary is nearly equal to its pre-explosion orbital velocity
-\citep[e.g.,][]{blaauw:61,eldridge:11}. For comparison, similar
-initial masses but a shorter initial period of $P^\mathrm{ZAMS}=7$\,days would experience a very similar
-evolutionary path, but it would only widen to a pre-CC period of
-$\sim$$45$\,days, resulting in a pre-CC orbital velocity of the
-secondary slightly larger than $30\,\mathrm{km\  s^{-1}}$.
-
-Because of the rejuvenating effect of
-mass accretion,
-the secondary remains on the MS for another $\sim$$3.6\,\mathrm{Myr}$ after
-the CC of the primary.
-
-\subsection{Effects of the natal kick on the post-CC orbit}
-\label{sec:example_kick}
-
-In our example system, the CC of
-the primary results in a stripped envelope SN of type\footnote{We do not attempt to distinguish between stripped SN types (IIb, Ib or Ic depending on
-the presence or lack of He lines in the spectrum). For a discussion of
-how this can tentatively be done in population synthesis calculations,
-and the related uncertainties, see \cite{zapartas:17b}.} Ib/Ic 
-\citep[][]{filippenko:97}, which forms a NS of about
-$M_\mathrm{NS}=1.6\,M_\odot$, with a fallback fraction $f_b=0.13$. The SN ejecta mass is $\Delta M_\mathrm{SN} = M_1^\mathrm{pre-CC}-M_\mathrm{NS}
-\simeq 3.7\,M_\odot$, where $M_1^\mathrm{pre-CC}$ is the primary mass just
-before the CC, and $M_\mathrm{NS}$ is the mass of the resulting
-NS. Because the ejecta mass is less than half of the total pre-CC mass
-of the system, $M_1^\mathrm{pre-CC}+M_2^\mathrm{pre-CC}\simeq33\,M_\odot$,
-the loss of the SN ejecta alone is insufficient to
-unbind the system \citep[][]{blaauw:61}. Whether the system is
-disrupted depends on the amplitude and direction of the natal kick
-imparted to the newly formed NS.
-
-We distinguish three cases: (i) a large kick with $v_\mathrm{k}\gg v_\mathrm{orb}$ results in disruption of the binary regardless of the
-direction, (ii) an intermediate kick $v_\mathrm{k} \simeq v_\mathrm{orb}$ disrupts the system only if the kick
-orientation is favorable, while (iii) a small kick, $v_\mathrm{k}
-\ll v_\mathrm{orb}$, never unbinds the binary, although it
-can still modify its orbit (e.g., \citealt{brandt:95, kalogera:96}, \citetalias{tauris:98}).
-
-To illustrate each case, we consider here three different fixed natal kick
-amplitudes $v_k\simeq 870, 230$, and $43\,\mathrm{km\ s^{-1}}$ (respectively
-corresponding to $1000,265$, and $50\,\mathrm{km \ s^{-1}}$ before the
-rescaling of the kick amplitude with the fallback fraction). For comparison, the pre-CC orbital
-velocity in our example system is $v_\mathrm{orb}^\mathrm{pre-CC}\simeq
-100\,\mathrm{km\ s^{-1}}$ (corresponding to
-$v_2^\mathrm{pre-CC}\simeq12.5\,\mathrm{km\ s^{-1}}$). For each kick amplitude $v_k$, we draw 10\,000 different directions,
-isotropically distributed in the frame of the exploding star \citep[e.g.,][]{wongwathanarat:13}.
-
-For large natal kicks, our example system is disrupted in 99.9\% of
-all the directions drawn. In the remaining 0.1\%, the
-compact object is shot in the envelope of the secondary (see
-\Secref{sec:bound} for a brief discussion).
-In the case of disruption, the ejected star acquires a velocity
-$v_\mathrm{dis}\simeq v_2^\mathrm{pre-SN}=12.5\,\mathrm{km\
-  s^{-1}}$. With $v_\mathrm{dis} < 30\,\mathrm{km\
-  s^{-1}}$, it is below the threshold to
-be classified as runaway as defined by \cite{blaauw:56,
-  hoogerwerf:00, hoogerwerf:01}, and thus it is
-an example of what we refer to as a walkaway star.
-
-In the case of intermediate kick amplitudes (here we use
-$230\,\mathrm{km\ s^{-1}}$), the angle $\theta$ between the kick direction and the
-pre-CC orbital velocity of the primary, given by $\cos\theta = \mathbf{v_k}\cdot\mathbf{v}_1^\mathrm{pre-SN}/(v_k
-\ v_1^\mathrm{pre-SN})$, determines whether the kick disrupts the
-system. Kicks that are roughly aligned with the pre-CC orbital
-velocity, i.e.\ $0 \lesssim \theta\lesssim\pi/4$, successfully disrupt the
-binary. Conversely, kicks that are roughly oriented opposite to the
-pre-CC orbital velocity of the exploding star just tend to slow the
-orbital motion: the system will remain bound on an eccentric
-orbit. In our example, we find that about $84\%$ of the kick
-directions result in disruptions. All cases produce a walkaway
-star with $v_\mathrm{dis}\simeq v_2^\mathrm{pre-SN}$.
-
-With small kick amplitudes (here $v_k\lesssim50\,\mathrm{km \ s^{-1}}$),
-our example system is never disrupted. The post-SN eccentricity, separation, and systemic velocity of the
-binary vary with the kick direction \citep[e.g.,][]{brandt:95,
-  kalogera:96, tauris:98}. We discuss the population of binaries surviving the
-first CC (and hosting a compact object) in \Secref{sec:bound}.
-
-In wide pre-CC systems, small kicks have a larger relative impact on the
-velocity of the unbound secondary. The reason is that in wide pre-CC
-orbits, smaller kicks result in a longer time for the compact object
-to exit the orbit and exert a gravitational pull on the other
-star. Conversely, for short pre-SN period systems (unlike the example
-considered here) the effects of large kicks 
-are more important. This is because the time for the compact object to
-exit the pre-CC orbit is short, but larger kicks are more
-likely to result in the disruption of the system than small kicks.
-
-\section{Analytical  estimates}
-\label{sec:analytics}
-
-\begin{figure*}[htb]
-  \centering
-  \includegraphics[width=0.33\textwidth]{figures/specificAM}
-  \includegraphics[width=0.33\textwidth]{figures/widening}
-  \includegraphics[width=0.33\textwidth]{figures/speedup}
-  \caption{Analytic estimates of the final-to-ZAMS ratio of the
-     specific orbital angular momentum $h$ (left panel), the orbital
-     separation $a$ (center panel), and orbital velocity of the
-     secondary $v_2$ (right panel), assuming $q^\mathrm{ZAMS}=0.75$ and
-     $\mu_\mathrm{env}=0.75$, which are representative for the example
-     system of \Secref{sec:example}. Blue colors correspond to a
-     decrease, red to an increase. The mass transfer efficiency
-     $\beta_\mathrm{RLOF}$ and the angular momentum parameter
-     $\gamma_\mathrm{RLOF}$ are assumed constant throughout the
-     evolution. The horizontal dashed lines show the fraction of the specific
-     orbital angular momentum removed for circumbinary disk mass loss
-     ($\gamma_\mathrm{disk}$, yellow) and for isotropic re-emission
-     ($\gamma_\mathrm{iso}$, green) for $q=0.75$. However, note that
-     these fractions depend on the present-day $q$ rather than
-     $q^\mathrm{ZAMS}$ and therefore they are not constant throughout the evolution in our simulations.}
-   \label{fig:speedup_factor}
- \end{figure*}
-
-The key quantity to define runaway and walkaway stars is their velocity.
-As the example described above demonstrates, the final velocity
-obtained at the binary disruption is of the
-order of its final (pre-CC) orbital velocity $v_\mathrm{dis}\simeq
-v_2^\mathrm{pre-CC}$, \citep[][]{blaauw:61,
-  eldridge:11}. Therefore, using a few simplifying assumptions, we can
-estimate the ejection velocity for systems that have experienced stable mass transfer, and express it in terms of the initial
-parameters of the binary. In particular, we will express the final
-orbital velocity as a function of the initial orbital velocity in the
-binary, and use it as a proxy for the ejection velocity. This gives
-insight on the possible outcomes of the binary evolution since we can estimate the typical velocity at ZAMS from
-observational constraints on the binary populations, \citep[e.g.,][]{sana:12}.
-
-Using conservation of linear momentum and
-Kepler's third law, we can write the orbital velocity $v_2$ of the
-secondary star at any time during the binary evolution as
-
-\begin{equation}
-  \label{eq:v2}
-  v_2 =
-  \frac{M_1}{M_1+M_2} v_\mathrm{orb} \equiv \frac{M_1}{M_1+M_2} \sqrt{\frac{G(M_1+M_2)}{a}}
-  \ \ ,
-\end{equation}
-where $G$ is the gravitational constant, $a$ is the semimajor axis of
-the orbit (assumed to be circular), and $M_1$, $M_2$ are the two
-masses.
-
-If we assume that a constant fraction
-$\beta_\mathrm{RLOF}=|\dot{M}_\mathrm{acc}|/|\dot{M}_\mathrm{don}|$ of the mass
-transferred is accreted by the secondary star, and that the mass not
-accreted leaves the binary system with a constant fraction
-$\gamma_\mathrm{RLOF}$ of the total
-specific orbital angular momentum, we can relate the final and initial
-separations with,
-\begin{equation}
-  \label{eq:af_ai}
-  \frac{a_\mathrm{pre-CC}}{a_\mathrm{ZAMS}} =
-  \left(\frac{M_1^\mathrm{ZAMS}}{M_1^\mathrm{pre-CC}}\frac{M_2^\mathrm{ZAMS}}{M_2^\mathrm{pre-CC}}\right)^2\left(\frac{M_1^\mathrm{pre-CC}+M_2^\mathrm{pre-CC}}{M_1^\mathrm{ZAMS}+M_2^\mathrm{ZAMS}}\right)^{2\gamma_\mathrm{RLOF}+1}
-  \ \ .
-\end{equation}
-
-The parameter $\beta_\mathrm{RLOF}$ enters implicitly in the values of
-the final pre-CC masses (see \Eqref{eq:DM}). Equation~\ref{eq:af_ai} neglects for simplicity all other processes that can
-remove mass from the binary, such as stellar winds \citep[see ][for a
-more elaborate expression]{soberman:97}.
-
-The most common type of binary interaction is case B RLOF \citep[][]{kippenhahn:67}.
-During case B RLOF, the primary typically loses nearly all its hydrogen-rich envelope
-\citep[][]{gotberg:17,yoon:17, gotberg:18}. We parametrize the
-fraction of the total mass in the envelope as $\mu_\mathrm{env} =
-M_\mathrm{env}/M$, or in other words $1-\mu_\mathrm{env}$ is the fraction of mass in the
-helium core. The parameter $\mu_\mathrm{env}$ is dependent on
-the overshooting assumed in the underlying stellar evolution
-models. The single massive star models from \cite{pols:98} give 
-values in the range $0.6\lesssim\mu_\mathrm{env}\lesssim 0.8$ with the larger values
-corresponding to less massive (and thus more common) stars. More
-recent models \citep[e.g.][]{brott:11} typically adopt larger values
-for the overshooting parameter, resulting in smaller values of $\mu_\mathrm{env}$.
-Using this parametrization, we can relate the final masses of the two star in
-the binary to their ZAMS masses as 
-\begin{multline}
-  \label{eq:DM}
- M_1^\mathrm{pre-CC} \simeq M_1^\mathrm{ZAMS}-\mu_\mathrm{env} M_1^\mathrm{ZAMS} \
- \ , \\
- M_2^\mathrm{pre-CC} \simeq M_2^\mathrm{ZAMS} +
- \beta_\mathrm{RLOF}\mu_\mathrm{env}M_1^\mathrm{ZAMS}
- \ \ . \hfill
-\end{multline}
-We can now express the final orbital velocity of the secondary $v_2^\mathrm{pre-CC}$ in
-terms of the initial $v_2^\mathrm{ZAMS}$. Introducing $q\equiv q^\mathrm{ZAMS} =
-M_2^\mathrm{ZAMS}/M_1^\mathrm{ZAMS}$,
-$\beta\equiv\beta_\mathrm{RLOF}$, $\mu\equiv\mu_\mathrm{env}$, and
-$\gamma\equiv\gamma_\mathrm{RLOF}$ to simplify the notation, we can write
-\begin{equation}
-  \label{eq:speedupdown}
-  \frac{v_2^\mathrm{pre-CC}}{v_2^\mathrm{ZAMS}}=\left(\frac{q+\beta\mu-\mu
-      q-\beta\mu^2}{q}\right)^2
-  \left(\frac{1+q}{q+1+\beta\mu-\mu}\right)^{\gamma+1} \ \ .
-\end{equation}
-
- In \Figref{fig:speedup_factor} we visualize the outcome of these
- analytical estimates for different assumptions for the mass transfer
- efficiency $\beta\equiv\beta_\mathrm{RLOF}$ and the angular momentum loss
- parameter $\gamma\equiv\gamma_\mathrm{RLOF}$. To make this figure, we adopt the parameters of the example system of
- \Secref{sec:example} ($q=0.75,\ \mu_\mathrm{env}=0.75$).
-
- The left panel of \Figref{fig:speedup_factor} shows the relative
- variation in the specific orbital angular momentum \mbox{$h\udef
- J_\mathrm{orb}/(M_1+M_2)=M_1M_2\sqrt{Ga}/(M_1+M_2)^{3/2}$}. For fully conservative evolution
- ($\beta=1$, white region on the right of this panel), $h$ stays
- constant. Also assuming $\gamma=1$ \citep[e.g.,][]{dominik:12} keeps
- the specific angular momentum constant, since this corresponds to
- losing mass and angular momentum at the same relative
- rate. Conversely, low values $\gamma\lesssim1$ result in a
- net increase of the specific orbital angular momentum because mass is
- lost relatively faster than angular momentum, while very large
- $\gamma\gtrsim 1.5$ values result in a loss of specific
- angular momentum.
-
- The central panel of \Figref{fig:speedup_factor} shows how the
- evolution affects the ratio of the final-to-initial separation (cf.~\Eqref{eq:af_ai}). The
- vast majority of the parameter space results in significant orbital
- widening (red colors), with only rather extreme angular momentum
- losses ($\gamma\gtrsim2$) in non-conservative
- ($\beta\lesssim0.6$) systems result in orbital shrinking.
-
- The right panel in \Figref{fig:speedup_factor} shows the main
- parameter of interest, the ratio of the final-to-initial orbital
- velocity of the initially less massive star in the system (cf.~\Eqref{eq:speedupdown}). Most
- assumptions for $\beta$ and $\gamma$ result in a decrease in the
- orbital velocity of the secondary (blue colors). This is the combined
- effect of (i) the orbital widening, and (ii) the increase mass of the secondary
- because of mass transfer. Only for very non-conservative systems
- ($\beta\lesssim0.3$) experiencing large
- angular momentum losses ($\gamma\gtrsim4$, corresponding to
- non-accreted mass removing four times the specific orbital angular
- momentum of the binary) the final velocity of the secondary is higher
- than its initial value.
-
- In our numerical simulations we adopt physically motivated model for the
- efficiency of mass transfer instead of constant parameters, $\beta$
- and $\gamma$, as we assumed in these analytical considerations. As we
- will argue later, most of the massive unbound companions result from
- nearly conservative mass transfer. Therefore, it is instructive to
- consider the analytical solution for $\beta = 1$. The expression for the final orbital velocity of the secondary (\Eqref{eq:speedupdown}) simplifies to 
- \begin{equation}
-   \label{eq:cons}
-   \frac{v_2^\mathrm{pre-CC}}{v_2^\mathrm{ZAMS}} =
-   \left(\frac{(1-\mu)(q+\mu)}{q}\right)^2 = \left(\frac{3+4q}{16
-       q}\right)^2 \ \ ,
- \end{equation}
- where we use as a typical value $\mu=3/4=0.75$ in the last
- step. \Eqref{eq:cons} gives a final-to-initial orbital velocity ratio for the
- secondary always smaller than one for initial mass
- ratios $q>0.25$. In other words, for all systems with initial mass
- ratios of interest evolving through stable and nearly conservative
- mass transfer, the secondary always slows down
- substantially, by up to a factor of 5 for an initially equal mass
- system. We emphasize that we have ignored the additional effect of mass loss by
- stellar winds. % (which typically reach terminal velocities much higher
- % than $v_\mathrm{orb}$)
- These cause the orbit to widen further and lead to a
- further slow down of the secondary star.
-
-\begin{figure*}[tpb]
-  \centering
-  \tikzset{font=\small,
-edge from parent fork down,
-level distance=2cm,
-sibling distance=3pt,
-every node/.style={top color=blue!15,
-  bottom color=blue!15,
-  rectangle,rounded corners,
-  minimum height=0mm,
-  align=center,
-},
-level 4/.style={sibling distance=20pt},
-level 5/.style={sibling distance=10pt},
-edge from parent/.style={
-  draw=blue!50,
-  very thick,
-},
-blank/.style={draw=none},
-}
-
-\begin{tikzpicture}
-  \Tree [
-          .{Binaries with\\$M_1^\mathrm{ZAMS}\geq7.5\,M_\odot$\\ $q^\mathrm{ZAMS}\geq0.1$\\$0.15\leq \log_{10}(P^\mathrm{ZAMS}/\mathrm{days})\leq 5.5$}
-               \edge [] node [draw=none, top color=white, bottom
-               color=white, above, xshift=-10pt] {$22_{-9}^{+26}\%$};
-               [
-                   .{Stellar\\mergers}
-               ]
-               \edge [] node [draw=none, top color=white, bottom
-               color=white, above, xshift=10pt] {$78_{-22}^{+9}\%$};
-               [
-                   .{Non mergers}
-                   \edge [] node [draw=none, top color=white, bottom color=white, midway, above] {$1-\mathcal{D}=14_{-10}^{+22}\%$};
-                   [
-                       .{Non Disrupted\\ {\tiny
-                           (cf.~\Secref{sec:bound})}}
-                                          \edge [] node [draw=none, top color=white, bottom color=white, midway, left] {$36_{-15}^{+42}\%$};
-                       [
-                       .{NS companion}
-                       ]
-                       \edge [] node [draw=none, top color=white, bottom color=white, midway, right] {$64_{-42}^{+15}\%$};
-                       [
-                       .{BH companion}
-                       ]
-                   ]                       
-                   \edge [] node [draw=none, top color=white, bottom color=white, midway, above] {\textcolor{red}{$\mathcal{D}=86_{-22}^{+10}\%$}};
-                   [
-                       .{Disrupted}
-                       \edge [] node [draw=none, top color=white, bottom color=white, midway, above] {$25_{-10}^{+8}\%$};
-                       [
-                           .{post-MS companion \\ {\tiny
-                               (cf.~\Secref{sec:non-MS-companions})}}
-                           \edge [] node [draw=none, top color=white,
-                           bottom color=white,  above, xshift=-6pt] {$53_{-46}^{+12}\%$};
-                           [
-                              .{WD}
-                           ]
-                           \edge [] node [draw=none, top color=white,
-                           bottom color=white,
-                           below, yshift=-40pt] {$23_{-20}^{+38}\%$};
-                           [
-                               .{He star}
-                           ]
-                           \edge [] node [draw=none, top color=white,
-                           bottom color=white,  above, xshift=6pt] {$25_{-6}^{+23}\%$};
-                           [
-                               .{H-rich}
-                           ]                        
-                      ]
-                      \edge [] node [draw=none, top color=white, bottom color=white, midway, above] {$75_{-8}^{+10}\%$};
-                      [
-                           .{MS companion \\ {\tiny (cf.~\Secref{sec:MS})}}
-                           \edge [] node [draw=none, top color=white,
-                           bottom color=white, above, xshift=-3pt] {$5_{-2}^{+14}\%$};
-                           [
-                              .{Runaway\\$v_\mathrm{dis}\geq30\,\mathrm{km\ s^{-1}}$}
-                           ]
-                           \edge [] node [draw=none, top color=white,
-                           bottom color=white, above, xshift=+3pt] {$95_{-14}^{+2}\%$};
-                           [
-                               .{Walkaway\\$v_\mathrm{dis}<30\,\mathrm{km\ s^{-1}}$}
-                           ]                              
-                     ]
-                   ]       
-               ]
-        ]    
-\end{tikzpicture}
-  % \input{figures/fig_tree/tree}% 
-  \caption{Overview of the possible binary evolution scenarios up to
-    the first CC event. The
-    branching ratios shown are from our fiducial simulation, and we
-    highlight in red the disruption fraction
-    $\mathcal{D}$. The errors on each fraction exclude the run without
-  SN kicks ($\sigma_\mathrm{kick}=0\,\mathrm{km\ s^{-1}}$), which
-  produces an unrealistically low disruption fraction
-  (cf.~\Tabref{tab:parameters} and \Secref{sec:param_variation}).}
-  \label{fig:tree}
-\end{figure*}
-
-Using the typical values of our example system discussed in
-\Secref{sec:example} ($q = 0.75$), we find $v_2^\mathrm{pre-CC}
-\simeq 0.25 \, v_2^\mathrm{ZAMS} \simeq  20\,\mathrm{km\ s^{-1}}$.
-Our numerical model for this system gives an even lower value,
-primarily because of the additional effect of widening as a result of
-mass loss through stellar winds. As a sanity check we recomputed the
-evolution of the example system, but artificially switching off the
-stellar wind mass loss for both stars. This gives
-$v_2^\mathrm{pre-CC}\simeq22\,\mathrm{km\ s^{-1}}$, in good agreement
-with the analitycal estimate. The remaining difference is due to the fact that the mass transfer is not fully conservative.
-
-For the analytic estimates presented in this section, we only
-considered the case of stable mass transfer. Binary evolution through
-unstable mass transfer takes place for a limited range of the initial
-distributions \citep[e.g.,][]{soberman:97,schneider:15}. It is
-expected to result in a common envelope phase, followed by substantial
-shrinking of the orbit either leading to a merger or the formation of
-a compact binary system if the envelope is ejected successfully.
-While this is a very interesting pathway to create very fast runaway
-stars, the numerical simulations presented later indicate that this
-channel produces primarily low mass runaway stars.
-This is because it typically concerns system with extreme initial mass ratios in which the secondary does not significantly gains mass.
-This means that the secondaries in these systems are typically not very massive and therefore this channel does not significantly contribute to the production of massive early type runaways, at least in our simulations.
-We return to this in \Secref{sec:discussion}.
-
-\section{Ejected companions}
-\label{sec:fiducial_res}
-
-In this section, present the results obtained from our numerical simulations for a full population computed with our fiducial assumptions (cf.~\Secref{sec:methods}).
-We describe the robustness of our findings against model assumptions in \Secref{sec:param_variation}.
-
-Our main goal is to characterize the
-velocity acquired by the companion stars ejected at the time of the first CC
-in the binary. Therefore, we ignore all binaries that merge, which constitute $\sim$$22\%$ of our fiducial
-population (cf.~\Figref{fig:tree}), in good agreement
-with~\cite{sana:12,demink:14,zapartas:17,zapartas:17b}; see also \cite{kochanek:14}.
-
-Once the merging systems are excluded, the first CC happens in the presence
-of a companion. The CC can result in either the disruption of the
-binary ($\mathcal{D}=86\%$ in our fiducial run), or the newly formed
-compact object remains bound to the companion star. In
-\Secref{sec:MS} we focus on the population of ejected MS companions,
-which are the $75\%$ of all the ejected companions in our fiducial run. We consider companions
-that have already evolved off the MS at the time of the first CC in \Secref{sec:non-MS-companions}.
-
-For completeness, we describe the systems
-with a MS companion remaining bound in \Secref{sec:bound}.
-
-\subsection{Ejected main sequence companions}
-\label{sec:MS}
-
-\begin{figure*}[htp]
-  \centering
-  \includegraphics[width=\textwidth]{./figures/v2_dist_fiducial}
-  \caption{Velocity distribution of MS stars ejected
-    from a binary system at the time of the first CC. The top panel shows
-    the corresponding cumulative distributions. The 
-    walkaways are one order of magnitude more numerous than the
-    runaways. The gray, blue, and red histograms show all MS
-    secondaries, only MS stars more massive than $7.5\,M_\odot$ (i.e.,
-    roughly those that might experience CC), and only MS stars more
-    massive than $15\,M_\odot$ (i.e., roughly the O-type stars),
-    respectively. The mass considered here is taken right after the CC of the primary star. The dashed
-    lines show the distribution of walkaways and runaways that have
-    gone through RLOF (or common envelope evolution) before being
-    ejected. Almost all massive runaways and walkaways, and the
-    majority of the ejected stars at large velocities, have gone
-    through RLOF. The inset plot magnifies the runaway regime
-    $v_\mathrm{dis}\geq30\,\mathrm{km\ s^{-1}}$. See also
-    \Figref{fig:v_dist_log} for a wider velocity range and
-    \Figref{fig:v_dist_obs} for a figure accounting for the remaining
-    lifetime as walkaway or runaway star.}
-  \label{fig:v_dist}
-\end{figure*}
-
-The vast majority of binaries hosting a MS companion is disrupted at the time of
-the first CC. Figure~\ref{fig:v_dist}
-shows the distribution of velocities of the ejected companions. The three different colors subdivide the population based on
-the mass of the ejected companion. The gray histogram shows all
-ejected MS companions. The blue histogram shows only massive MS
-companions, $M_2^\mathrm{post-CC}\geq 7.5\,M_\odot$, where
-$M_2^\mathrm{post-CC}$ is the mass of the secondary just after the
-collapse of the primary. This group roughly corresponds to all
-secondaries that will experience CC at the end of their lifetime. The red histogram shows the velocity
-distribution of companions with mass larger than $M_2^\mathrm{post-CC}\geq15\,M_\odot$, corresponding roughly to O-type
-stars. In the rest of this section, we quote the ratios for the
-population of massive stars (i.e.~$M_2^\mathrm{post-CC}\geq 7.5\,M_\odot$) unless
-otherwise stated, see also \Tabref{tab:parameters} for more
-information.
-
-Although we do produce runaways, the majority of the systems ejects a star slower than $30\,\mathrm{km \
-  s^{-1}}$, i.e., a walkaway star \citep[][]{de-Mink+2012,lennon:16}. We find that the CC of a star
-with a MS binary companion is about twenty times more likely to produce a
-walkaway than a runaway. In the first line of \Tabref{tab:parameters},
-we list the ratio $\mathcal{R}$ of walkaway to runaway stars produced per CC event in a binary. From our
-fiducial simulation, we obtain $\mathcal{R}_\mathrm{7.5}\simeq20$ for
-massive companions.
-As shown in
-\Tabref{tab:parameters}, considering progressively larger masses, the number of walkaways
-produced per each runaway from a binary increases.
-
-Runaways resulting from the disruption of binaries rarely exceed $60\,\mathrm{km \
-  s^{-1}}$: about 99.8\% of the massive unbound companions we simulate are slower than
-this threshold. We note that higher velocity are not strictly
-forbidden, but extremely unlikely (cf.~\Figref{fig:v_dist_log} which
-shows the velocity distribution on a logarithmic scale).
-Low mass runaways can reach much
-tighter pre-CC orbits through common envelope evolution (without
-significant accretion), and thus reach higher ejection velocities, but typically
-not in excess of $400\,\mathrm{km\ s^{-1}}$. Stars significantly
-faster than this are likely produced by different mechanisms
-(see \citealt{boubert:17b}, and \citealt{brown:15} for a review). However, in extremely rare cases
-where a very short pre-CC binary is formed and also successfully disrupted,
-much higher velocities can be achieved.  The
-absolute maximum velocity we obtain is about $1110\,\mathrm{km\
-  s^{-1}}$, in good agreement with \cite{tauris:15}.
-
-The slowest stars come from the widest pre-CC binaries:
-the low-velocity drop-off of the distribution in \Figref{fig:v_dist}
-is thus
-an effect of the upper-limit on our period distribution (corresponding
-to an upper-limit on the pre-CC period).
-
-The velocity distribution peaks at $v_\mathrm{dis}\simeq6\,\mathrm{km\
-  s^{-1}}$. Ejected MS companions with such velocity typically
-originate from systems with $q_\mathrm{ZAMS}\gtrsim0.5$ and long initial
-periods ($\log_{10}(P^\mathrm{ZAMS}/\mathrm{days})\simeq3$),
-regardless of the stellar masses. These systems experience late case B
-or case C mass transfer \citep[][]{lauterborn:70}, but owing to the
-sufficiently high mass ratio, the mass transfer remains conservative
-and the system widens.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figures/Dhist}
-  \caption{Distribution of the maximum distance $L$ that ejected
-    MS secondaries can reach, neglecting any gravitational
-    potential. The blue and red histograms are for ejected stars more
-    massive than $7.5\,M_\odot$ and $15\,M_\odot$, respectively. The
-    vertical dashed lines mark the mean values of
-    the distributions.}
-  \label{fig:Dhist}
-\end{figure}
-
-Among low and intermediate mass walkaways with
-$v_\mathrm{dis}\simeq6\,\mathrm{km\ s^{-1}}$, there is also a contribution from
-wide systems which do not experience mass transfer. This contribution
-is almost negligible for ejected stars more massive than
-$15\,M_\odot$. This can be seen from the dashed distributions in
-\Figref{fig:v_dist}, which show that the majority
-of massive runaways and walkaways have accreted mass from
-their companion before its CC. Over the entire velocity range shown
-(i.e., considering both runaways and walkaways), $71\%$ of ejected MS
-secondaries more massive than $7.5\,M_\odot$, and $91\%$ of those more
-massive than $15\,M_\odot$ have accreted from their companion in
-the previous evolution. Among low mass ejected companions, $50\%$ have
-experienced either RLOF or common envelope evolution before their
-ejection from the binary. From the inset in \Figref{fig:v_dist}, where
-the blue and red dashed lines representing post-interaction systems
-overlap with the full distribution, it is clear that all massive
-runaways from binary disruption have experienced mass transfer.
-
-The majority of the ejected massive MS
-companions are not ``usual'' single stars, and this remains true for a
-very large fraction of the low mass MS ejected companions. Before the binary disruption,
-they have accreted mass from their companion. Mass transfer also causes the convective core of the accretor to grow,
-resulting in rejuvenation of the accreting star
-\citep[e.g.,][]{hellings:83,schneider:16}. The fact that most ejected
-MS stars are accretors is the result of a
-combination of assumptions. For disruption to be possible,
-the primary needs to be massive enough to collapse. We assume a flat
-mass ratio distribution, implying that on average $M_2^\mathrm{ZAMS}\simeq
-0.5\,M_1^\mathrm{ZAMS}$, and we also assume that mass transfer is
-unstable and leads to mergers if the mass ratio is too extreme
-($M_\mathrm{acc}/M_\mathrm{don}<q_\mathrm{crit}$), therefore binaries
-that can eject a star have initial mass ratios closer to one than an
-average system. Finally, when stable mass transfer occurs, our fiducial
-assumption for the accretion efficiency results in rather conservative
-mass transfer over most of the period range considered
-\citep[][]{schneider:15}.
-
-To make an estimate of the
-ratio of walkaways  per runaway {\it existing at a given time}
-requires us to
-consider the finite MS lifetime of the ejected stars. Assuming that the star formation rate in the galaxy is
-constant and therefore that the Galactic population is in a steady state,
-the ratio of walkaways per runaway is
-$\sim$13 for masses larger $7.5\,M_\odot$.
-
-Accounting for the durations of each evolutionary phase, we can also
-quantify the fraction of MS stars more massive than $15\,M_\odot$ that are
-runaway or walkaway stars. In our fiducial simulation, we find a
-runaway fraction for masses larger than $15\,M_\odot$ of about $\sim$$0.5\%$ and a
-walkaway fraction of $\sim$$10\%$. We only simulate binaries with primaries more
-massive than $7.5\,M_\odot$, therefore we do not have in our
-population intermediate-mass binaries which could contribute through
-mass transfer and/or mergers  to the
-normalization factor for the runaway and walkaway fraction at lower
-masses \citep[][]{zapartas:17}. We show in \Figref{fig:v_dist_obs} the distribution in
-velocity accounting for the lifetime of the stars in each bin. This
-can be directly compared to observations as long as the effects of the
-Galactic potential can be neglected.
-
-\begin{figure*}[!htbp]
-  \centering
-  \includegraphics[width=0.75\textwidth]{figures/veq_vrw}
-  \caption{Equatorial velocity at the time of ejection for massive ($M_\mathrm{dis}\geq7.5\,M_\odot$) companions as a function
-    of the ejection velocity. Brighter colors indicate location of the
-    parameter space more populated by our fiducial simulation. All the massive runaway stars spin with
-    an equatorial velocity of $\sim$$500\,\mathrm{km\ s^{-1}}$, close to
-  breakup rotation, since they have accreted mass from their companion
-  before being ejected. The spread is due to wind spin down before the
-  binary disruption. We do not include the effect of projecting on
-  the line of sight, or post-ejection wind spin down in this
-  plot.  The
-    vertical dot-dashed line marks the threshold to define runaway
-    stars.}
-  \label{fig:vrot_vrw}
-\end{figure*}
-
-Figure~\ref{fig:Dhist} shows the distribution of distances traveled by
-massive, unbound stars, calculated by multiplying their velocity
-$v_\mathrm{dis}$ by their remaining lifetime on the MS, which includes
-rejuvenation if the star has accreted mass, plus 10\% of the
-MS duration of a star with the same helium core mass, to account for
-the duration of the helium core burning phase in the rejuvenated
-star\footnote{The duration of the helium core burning depends on the helium core mass, and thus the overshooting
-parameter. The value of $10\%$ of the MS duration is typical for overshooting values
-larger than what assumed in \cite{pols:98}.}.
-This approach neglects the effect of an external potential on the trajectory of the
-stars \citep[see][for how this could be done]{boubert:17a,boubert:18}, so it is effectively an upper limit to the distance they can travel. Considering both runaway and walkaway stars with masses
-$M_\mathrm{dis} \geq7.5\,M_\odot$, we find
-the mean distance $\langle L \rangle$ they travel before experiencing
-CC to be 126\,pc, see the blue vertical dashed line in
-\Figref{fig:Dhist} and also \Tabref{tab:parameters}. This number
-rises to $\langle L_\mathrm{run}\rangle=584$\,pc considering only 
-runaways (faster than $30\,\mathrm{km \ s^{-1}}$) more massive than
-$7.5\,M_\odot$, but it remains $\langle
-L_\mathrm{walk}\rangle=103$\,pc considering massive walkaways only
-(slower than $30\,\mathrm{km \ s^{-1}}$). For comparison, the typical
-size of OB-associations is of the order of tens of pc,
-while the Galactic thin and thick disks have a vertical scale height
-of $\sim$300 and $\sim$1500\,pc, respectively
-\citep[e.g.,][]{gilmore:83}.
-
-\subsection{Spin of the main sequence ejected companions}
-
-Figure~\ref{fig:vrot_vrw} shows the equatorial rotational velocity
-$v_\mathrm{dis}^\mathrm{eq}$ for
-the ejected MS companions with $M_\mathrm{dis}\geq7.5\,M_\odot$. The top panels show the normalized one-dimensional
-distribution of ejection velocity for the whole sample (same as the
-blue distribution in \Figref{fig:v_dist}), and for the runaways only,
-respectively. The side panels show the normalized equatorial
-rotational velocity distribution
-for the whole sample, and for the runaways only, respectively.
-We do not include
-in \Figref{fig:vrot_vrw} any projection effects due to the inclination
-of the stellar spin axis with respect to the line of sight, nor the
-wind spin down due to the evolution after the binary disruption. The
-combination of these two effects would smear out the distribution in
-the vertical direction, allowing any physically possible
-\emph{projected} equatorial rotational velocity to be observable.
-
-The main feature at the top of the central panel (also visible in the rightmost panel) shows that almost all the massive runaways are spinning
-nearly at breakup rotation at the time of the ejection, because of the
-relatively recent spin up during the mass transfer phase(s).
-
-The walkaways ($v_\mathrm{dis}<30\,\mathrm{km\ s^{-1}}$) instead can
-have a broader range of rotational velocities: this is because a
-subset of them comes from wide, non-interacting binaries. The right
-panel including both the walkaway and runaway population shows two
-peaks for $v_\mathrm{dis}^\mathrm{eq}\simeq200$ and $500\,\mathrm{km\
-  s^{-1}}$. The former is due to the combination of the accretion-induced spin up and the wind spin down
-before the ejection, which can be more efficient for walkaways
-experiencing early mass transfer and with longer delay between the
-mass transfer and the disruption of the binary. The former peak is \emph{not} related to tidal locking, which
-is efficient only for tight orbits that would result in much larger ejection
-velocities $v_\mathrm{dis}$. The latter is populated by secondaries ejected after a
-mass transfer phase (analogous to the peak in the equatorial velocity
-distribution for the runaways).
-
-\subsection{Ejected post-main sequence companions}
-\label{sec:non-MS-companions}
-
-Figure \ref{fig:v2_postMS} shows the velocity distribution of companion stars which have evolved off
-the MS by the time of the first CC in the binary. All
-together, these correspond roughly to 25\% of the population of
-disrupted binaries (cf.~\Figref{fig:tree}). We distinguish three sub-populations:
-hydrogen-rich non-degenerate secondaries (mostly of red-supergiant
-secondaries, corresponding to $\sim$$4\%$
-of the entire simulated population, including mergers and
-non-disrupted binaries), white dwarfs (WD, $\sim$8\% of
-the total), and helium stars which have lost their hydrogen-rich envelope because of previous
-binary interactions ($\sim4\%$ of the total).
-
-For the first group, characterized by large stellar radii, the velocity distribution resembles
-closely that of the MS ejected companions, with a large number
-of walkaways (roughly corresponding to $96\%$ of the hydrogen-rich
-non-degenerate ejected companions) compared to the runaways.
-
-\begin{figure}[tp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figures/v2_dist_postMS.pdf}
-  \caption{Velocity distribution of post-MS ejected companions. Helium
-    stars and white dwarfs (WD) can have much shorter pre-CC orbits
-    because of their smaller radii, and thus they can reach much
-    higher ejection velocities. Conversely, the number of non-degenerate, hydrogen-rich
-    post-MS stars is dominated by red supergiants with very large
-    radii, which can only exist at large separation from the exploding
-    companion, and thus typically have much slower velocities. The top
-    panel shows the corresponding cumulative distributions.}
-  \label{fig:v2_postMS}
-\end{figure}
-
-Ejected WDs and helium stars come from different evolutionary paths. Mass transfer during the primary MS
-can reduce the helium core mass thus slowing its evolution, while
-accretion of mass speeds up the evolution of the companion, although this is compensated by the rejuvenation effect. In
-$\sim23\%$ of our full population (excluding stellar mergers), this leads
-to a reversal of the CC order, with the secondary exploding first \citep[][]{pols:94}. In
-these cases the star that is ejected is the initially more massive, which
-has become, by the time of the CC, a helium star or a WD.
-
-Because of their smaller radii, WDs and helium stars can
-have much shorter pre-CC orbits and therefore reach velocities much
-larger than (super-)giant stars. More importantly, these stars are what
-is left of the initial donor and they are thus less massive than
-their companion after the mass transfer phase.
-
-This results in ejection velocities of
-the WDs spanning a range  $15 \lesssim v_\mathrm{dis}/\mathrm{[km\
-  s^{-1}]}\lesssim 90$, and helium stars are even faster with $70 \lesssim v_\mathrm{dis}/\mathrm{[km\ s^{-1}]}\lesssim
-120$. This difference arises because it takes longer for a star to
-become a WD than its helium core burning duration. Therefore, WDs are
-ejected by systems where the rejuvenation of the 
-secondary is not extreme. This means that these systems have
-a less conservative mass transfer phase than binaries ejecting a
-helium star, which corresponds to less
-rejuvenation of the secondary and more orbital widening, and thus
-lower velocity of the WD at ekection. Moreover, if
-the companion star at pre-CC stage is a WD,  there is a longer time
-for the binary to widen, e.g.,~because of winds, than if it is a
-helium star.
-
-Conversely, if the ejected star is a non-degenerate helium
-star, then the entire evolution of the rejuvenated secondary needs to
-be faster than the remaining lifetime of the primary (which now
-has a reduced mass): the time delay between the RLOF phase and the
-first CC ejecting the helium star is much shorter and there is not enough
-time for significant orbital widening.
-
-Figure~\ref{fig:v2_postMS} also shows a small population of slow ($v_\mathrm{dis} <
-30\,\mathrm{km\ s^{-1}}$) helium stars. These are very massive
-($M_2\gtrsim 60\,M_\odot$) stars characterized by large Wolf-Rayet
-wind mass loss rates, which cause significant orbital widening before
-the first CC, resulting in the slow ejection velocity. We emphasize that \Figref{fig:v2_postMS} shows the velocity of
-stars ejected while in the corresponding evolutionary stage, and stars
-ejected during their MS will then evolve into post-MS hydrogen-rich,
-or possibly Wolf-Rayet stars, creating a more complicated velocity distribution of
-massive Wolf-Rayet stars \citep[see, e.g.,][]{dray:05, eldridge:11}.
-
-\section{Impact of model variations}
-\label{sec:param_variation}
-
-Our predictions depend on a set of parametrized assumptions. We
-perform a systematic study of the impact of these uncertainties by
-varying the values of the free parameters one-by-one, as we describe in
-\Secref{sec:methods}. In each variation, all the other parameters are set to their
-fiducial value.
-
-Table \ref{tab:parameters} summarizes our results for binaries with a
-collapsing star and a MS companion. If our extreme assumptions
-do not change much our synthetic population, then the outcome is
-robust and independent of the particular physical process. Conversely,
-strong dependence on any one parameter indicates that such parameter could be constrained by
-comparing to observed samples, and therefore allow for physical tests
-of the process it represents. Overall, our results are robust and relatively insensitive to
-uncertain physical processes, at least within the framework of
-parametrizations we assume.
-
-\subsection{Robust  predictions}
-
-The fraction of binaries that are disrupted by the first CC is always larger
-than $75\%$, with two notable exceptions. These concern the natal kick
-distribution assumed. With a double Maxwellian kick distribution with
-$\sigma_\mathrm{kick}=30\,\mathrm{km\ s^{-1}}$ for NSs less massive
-than $1.35\,M_\odot$ (and our fiducial kick for more massive
-remnants), the disruption fraction
-decreases to $\mathcal{D}=65\%$ even though this changes the kick only
-for a small subset of the systems. Low mass NSs are the
-typical result of progenitors with small total and core masses, which
-are favored by the IMF.
-
-If instead we make the extreme assumption of zero
-natal kick in all CC ($\sigma_\mathrm{kick}=0\,\mathrm{km\ s^{-1}}$), then only $\mathcal{D}=16\%$ of the
-binaries are disrupted. In these cases the disruption is
-only due to the mass loss at the moment of the explosion (Blaauw
-kick). Since the ejecta mass must be larger than half of the total
-mass of the system to disrupt the binary \citep[][]{blaauw:61}, and since after mass
-transfer the primary is typically less massive than the companion, only very wide binaries which did not experience mass transfer
-can be disrupted in such a way. Therefore, in this extreme assumption,
-the ejected stars are on average
-less massive and slower compared to calculations including a natal
-kick. We emphasize that this parameter variation is in contrast with
-the observed distribution of pulsar velocities, and we
-include it for insight only. This variation does not produce any massive
-runaways, which is also in contrast with observations.
-
-We obtain the largest disruption
-fraction when we do not rescale the natal kick with the fallback
-fraction: in this model, up to $97\%$ of the systems are
-disrupted. This assumption is equivalent to assuming a ``BH velocity
-kick'' \mbox{\citep[e.g.,][]{belczynski:08,fryer:12}}: BHs and NSs receive
-the same kick amplitude at birth, despite the differences in their masses. Therefore, many CC events
-resulting in the formation of a massive BH with large fallback
-fractions (up to $f_b=1$) still result in the disruption of the binary.
-
-Despite the overall large disruption fraction in all our parameter
-variations, the predicted runaway fraction $f_{15}^\mathrm{RW}$ for masses larger than $15\,M_\odot$ is low for all our
-parameter variations and never exceeds a few percent. Similarly, the walkaway
-fraction $f_{15}^\mathrm{WA}$ is also robust, and is generally around
-$10\%$. In other words, the walkaways outnumber the runaways by
-a factor of 10-30 (smaller for larger masses), regardless of the
-physical assumptions we make. The
-most notable exceptions are computations at lower metallicity (see
-\Secref{sec:Z}). Another case producing slightly smaller $\mathcal{R}$ is
-fully non-conservative mass transfer ($\beta_\mathrm{RLOF}=0$): more
-systems are in a very close pre-CC orbits and the average velocity of
-an ejected secondary is higher than in our fiducial assumptions.
-
-The average mass of the walkaways, runaways, and all ejected
-secondaries ($\langle M_2^\mathrm{walk} \rangle$, $\langle
-M_2^\mathrm{run} \rangle$, and $\langle M_2 \rangle$, respectively) exceeds $7.5\,M_\odot$ in almost all our parameter variations. With
-fully non-conservative mass transfer ($\beta_\mathrm{RLOF}=0$),
-the secondary accretes no mass. This makes them on average less
-massive than in our fiducial assumptions at the moment of the first CC, and also allows
-for less orbital widening (the donor has to lose more mass before the
-mass ratio reverts), resulting in a faster population of disrupted
-stars with lower masses compared to our fiducial case. If we assume that case B mass transfer is
-unstable whenever the accretor is less massive than the donor
-($q_\mathrm{crit, B}=1$), or if we take a negative slope of the mass ratio
-distribution ($\kappa=-1$), then the fraction of binary systems that
-merge after the end of the primary MS increases compared to our
-fiducial case. 
-When a binary system merges, it cannot eject a walkaway or runaway.
-Finally, assuming no natal kicks ($\sigma_\mathrm{kick}=0$), the only
-systems disrupted are those where the SN ejecta consist of more than
-half of the total mass of the binary at the pre-explosion stage. This
-requirement naturally biases the population of ejected companions
-towards low masses. If we assume that the kick velocity is not scaled down by the amount
-of fallback (``velocity kick'' for BHs), CC events from more massive primaries can also disrupt
-the system. On average, this produces more massive ejected
-secondaries compared to our fiducial case. Since the remaining lifetimes are shorter, the
-distance traveled by the population of ejected secondaries is
-also decreased, despite their higher velocity.
-
-Our assumptions on the stability of case B RLOF (assumed to be stable
-if the accretor-to-donor mass ratio is smaller than $q_\mathrm{crit,B}$) impact the
-velocity and mass distribution of the ejected stars. The sequence from
-more stable to more unstable
-$q_\mathrm{crit,B}=1,\,0.5,\,0.4$\,(fiducial
-value),\,0 shows progressively higher $\mathcal{R}$ indicating
-relatively less runaways, but higher average velocity $\langle v\rangle$ (and
-consequently also a larger $\langle L \rangle$). The increase in velocity and
-decrease in average mass of the ejected companions for higher
-$q_\mathrm{crit,B}$ indicates an increased importance of evolutionary
-channels involving a common-envelope evolution leading to tighter
-pre-CC orbits without significant accretion of mass from the companion.
-
-The average distance traveled by walkaways and runaways is also a
-robust prediction. The entire population of ejected stars typically
-travels slightly farther than $\langle L\rangle\simeq120$\,pc, which
-corresponds to an average distance for the runaways of $\langle
-L\rangle_\mathrm{run}\gtrsim500$\,pc, while the walkaways alone reach $\langle L\rangle_\mathrm{run}\gtrsim100$\,pc. The
-effect of rejuvenation, allowing for increased lifetimes of
-these stars, is largely responsible for this.
-
-These distances are
-significantly shortened only assuming fully conservative mass transfer
-($\beta_\mathrm{RLOF}=1$), or assuming higher angular momentum loss
-(e.g., through a circumbinary disk, $\gamma_\mathrm{RLOF}=\gamma_\mathrm{disk}$). The former results in
-higher masses of the ejected secondaries and thus smaller velocities
-and shorter lifetimes,
-while the latter results in more orbital shrinking and mergers. Another variation
-resulting in small $\langle
-L\rangle_\mathrm{walk}\simeq66\,\mathrm{pc}$ is that with zero natal kicks
-($\sigma_\mathrm{kick}=0\,\kms$). As mentioned above, only the widest pre-CC binary can be
-disrupted in this case producing a much slower population of ejected
-MS stars.
-
-The predicted rotational velocity of the ejected companions is also a
-robust prediction. The vast majority of massive
-($M_\mathrm{dis}\geq7.5\,M_\odot$) main sequence runaways accrete mass
-from their companions, and thus spin up to critical rotation, before
-being ejected. Visual inspection of plots like the one shown in
-\Figref{fig:vrot_vrw} for our fiducial simulation show no variation
-when changing the initial rotation rate of the stars, confirming that
-mass transfer in binaries overwrites the initial rotation of the
-accretor. %% beta=0 might have an issue with wind spin up.
-
-\subsection{The mass function of massive runaways as probes for black hole kicks}
-\label{sec:bh_kicks_mf_rw}
-\begin{figure}[!bp]
-  \centering
-  \includegraphics[width=.5\textwidth]{./figures/BH_MF_sigma100}%BH_kick_MF}
-  \caption{The mass function of massive ($M_\mathrm{dis}\geq7.5\,M_\odot$), runaway
-    (faster than $\geq30\,\mathrm{km \ s^{-1}}$) stars depends on the
-    assumptions for the BH kick. The bottom panel
-    shows our fiducial simulation
-    with $\sigma_\mathrm{kick}=265\,\mathrm{km \ s^{-1}}$, including the fallback
-    downscaling of the kick both for NSs and BHs. The fallback is an
-    important effect only for BHs. The central panel
-    shows the mass function for intermediate BH kicks
-    ($\sigma_\mathrm{kick}=100\,\mathrm{km\ s^{-1}}$, no fallback
-    downscaling) and our fiducial kick for NSs. The top panel shows
-    the distribution with our fiducial kick amplitude, but without fallback downscaling (i.e., BH
-    velocity kick). The dashed lines indicate the $90^\mathrm{th}$
-    percentiles of the mass distributions.}
-  \label{fig:bh_kicks}
-\end{figure}
+\section{The kinematics of VFTS682}
 
@@
   \centering
-   \includegraphics[width=0.33\textwidth]{figures/M_v_fiducial}
-   \includegraphics[width=0.33\textwidth]{figures/M_v_sigma100}
-   \includegraphics[width=0.33\textwidth]{figures/M_v_no_fb}
-   \caption{Mass-velocity correlation for the three BH kick variations
-     discussed in \Secref{sec:bh_kicks_mf_rw}. The brighter colors indicate
-     where on this plane we predict more ejected companions, the
-     colors have a logarithmic scale. The left panel shows
-     our fiducial distribution, the central panel shows the case for
-     intermediate BH kicks ($\sigma_\mathrm{kick}=100\,\mathrm{km\
-       s^{-1}}$ and no fallback), and the right panel shows large BH
-     kicks (same as for NSs). The NS kick is effectively the same in
-     all panels, since the amount of fallback is very small for SNe
-     producing a NS.}
-   \label{fig:M_vrw}
+  \includegraphics[width=\textwidth]{./figures/main_plot}  
+  \caption{position and projected relative velocity to R136}
+  \label{fig:main}
 \end{figure*}
 
-The mass distribution of runaways carries
-information about the BH kicks. With a sufficiently large sample, it might be used to
-discriminate between BH velocity or momentum kicks, at least in a
-statistical sense. We focus on the mass function of massive
-($M_\mathrm{dis}\geq7.5\,M_\odot$) and fast runaways
-($v_\mathrm{dis}\geq30\,\mathrm{km \ s^{-1}}$)
-only, i.e., not including the more common walkaways, since we expect
-upcoming observational distributions to provide cleaner sample for
-these. Ejected companions with high masses are more likely to come from
-binaries with an initially very massive primary resulting in BH
-formation at the end of its evolution.
-The constraints on the amplitude of BH kicks from unbound companions are
-complimentary to those from binaries surviving the first CC, and Gaia
-might be able to probe both populations of bound and ejected
-companions, \citep[][]{breivik:17}.
-
-\Figref{fig:bh_kicks}
-shows the mass distribution of runaways
-($v_\mathrm{dis}>30\,\mathrm{km\ s^{-1}}$) more massive than $7.5\,M_\odot$ for three BH
-formation scenario. The bottom panel shows our fiducial
-simulation, in which natal kicks are damped by fallback (effectively,
-this is equivalent to a BH ``momentum kick''). In this case, the $90^\mathrm{th}$
-percentile of the massive runaway mass function is $\sim$$25\,M_\odot$.
-In this model, the amount of fallback and the amplitude of the BH natal kicks
-are degenerate parameters. Because of the large fallback fraction
-prescribed by the \cite{fryer:12} algorithm, the runaway mass function
-produced by this model is practically indistinguishable from that with
-no BH natal kicks.
-
-The central panel of \Figref{fig:bh_kicks} shows the mass distribution
-assuming no fallback downscaling, but a smaller kick amplitude
-distribution characterized by $\sigma_\mathrm{kick}=100\,\mathrm{km \
-  s^{-1}}$ for the BHs (while we keep the same fiducial assumptions
-for NSs). Compared to our fiducial
-simulation, BHs receive larger kicks, and thus the runaways produced
-are also generally more massive, and the $90^\mathrm{th}$ percentile
-of the mass distribution shifts to $\sim$$30\,M_\odot$.
-
-The histogram in the top panel of \Figref{fig:bh_kicks} instead corresponds
-to a scenario in which the natal kick is \emph{not} decreased by the amount of
-fallback (e.g., if the natal kick is entirely due to asymmetries in the neutrino
-emission). In this case BHs and NSs receive the same kick amplitude
-(BH ``velocity kick'') and this results in a larger number of massive
-runaways characterized by a mass function skewed towards larger
-values. In this scenario, the $90^\mathrm{th}$ percentile of the mass
-distribution is as high as $\sim$$50\,M_\odot$.
-
-In summary, the tail of the mass function of massive runaways carries
-information about the BH natal kicks: finding runaways of large masses
-would support the possibility of large BH natal kicks. We do not
-expect large contamination of such tail from dynamical ejection
-of runaway stars, since dynamical channels tend to eject the least
-massive star among those interacting \citep[although see
-also][]{fujii:11, oh:16}.
-
-Our result allow us to check also for a correlation between the mass
-of the ejected companions and its velocity. \Figref{fig:M_vrw} shows
-this comparison for the three BH kick variations of
-\Figref{fig:bh_kicks}. As expected, there is a correlation between the
-maximum velocity that can be achieved because of
-the binary disruption, and the mass of the ejected MS companion. This
-can be seen by the shape of the colors in each panel. More massive companions have more
-inertia and are harder to accelerate, or, in other words, more massive
-companion have slower pre-CC orbital velocity because of the mass
-dependent factor in \Eqref{eq:v2}.
-
-From left to right, the BH kick amplitude increases (while effectively
-keeping the same NS kick, since the fallback fraction is small for
-CCSNe resulting in NS formation), and so does the maximum mass of the
-ejected companions. There is an sharp drop in the number of ejected
-companions for $M_\mathrm{dis}\gtrsim70\,M_\odot$. This
-arises because of the combination of several effects. First of all,
-stars in this mass range are intrinsically rare because of the
-IMF. Also, the lifetimes of very
-massive stars become very similar ($\sim3\,\mathrm{Myr}$) above $\sim$$50\,M_\odot$. This
-means that by the time the primary collapses, the secondary is also an
-evolved star that has finished burning hydrogen in its core, and it
-does not appear in the panels of \Figref{fig:M_vrw}. Finally,
-these stars tend to have large wind mass loss rates, which can be
-enhanced significantly when the star is spun up by the accretion: the
-combination of the wind enhancement with already large wind mass loss
-rates limits the growth in mass of the companions.
-
-In an observed
-population, the short post-ejection lifetime would also make very
-massive ejected companion rare. However, the upperlimit of
-$\sim$$70\,M_\odot$ is not strict, and it is possible to produce
-walkaways in this mass range through binary interactions, although at
-a negligible rate compared to less massive walkaways.
-
-\subsection{Effects of metallicity }
-\label{sec:Z}
-
-We find that, other than the natal kick distribution, metallicity is the most important parameter
-influencing our results. Decreasing the metallicity from super-solar,
-$Z=0.03$, to $Z=0.02\simeq Z_\odot$ (fiducial run), to
-$Z=0.008$, $Z=0.0047$ to $Z=0.0002\simeq10^{-2}Z_\odot$ a clear trend
-emerges from \Tabref{tab:parameters}. At lower metallicity, the
-fraction $\mathcal{D}$ of binaries disrupted by the first CC
-is lower ($77\%$ for the lowest metallicity, cf. $88\%$ at the highest), but the stars ejected are on average
-faster. This can be seen from the average velocity $\langle v\rangle$,
-which goes from $21.6\,\mathrm{km \ s^{-1}}$ to $11.2\,\mathrm{km\ s^{-1}}$, but also from the progressive decrease of the ratio of walkaways per runaway
-$\mathcal{R}$, regardless of the lower mass cut. Nevertheless,
-walkaways outnumber runaways at all the metallicities we explore.
-
-In our lowest metallicity run, the runaways travel, on average, as far as 706\,pc from their birth
-location, and the walkaways reach, on average, 163\,pc. The mean
-distance travelled by population of disrupted stars is 279\,pc for
-the lowest metallicity.
-
-The reduced likelihood of disrupting a system and increased velocity
-of the stars ejected can be understood in
-terms of the effect of metallicity on stellar radii. At lower $Z$
-stars of a given mass and evolutionary stage have smaller radii. This means that a given
-binary will enter in a mass transfer phase later, resulting in less conservative mass transfer and 
-less orbital widening, making the binary harder to
-disrupt and the pre-CC orbital velocity of the secondary
-faster. This is corroborated by the fact that the average mass of
-walkaways and runaways decreases slightly with metallicity, which is indicative of
-a less conservative mass transfer. Moreover, at lower metallicity, the post-interaction orbital widening
-due to stellar winds is less important.
-
-Another important physical effect, presently not included in
-our simulations, is that, at the end of the RLOF, the primary remains more massive at
-lower metallicity \citep[owing again to the smaller radii, e.g.][]{gotberg:17,yoon:17}. This
-should contribute to the reduction of orbital widening for interacting binaries,
-and make the collapsing star more massive and thus the disruption harder.
-
-\input{./Table_BHmid_kick.tex}
-
-\begin{figure*}[!htbp]
-  \centering
-  \includegraphics[width=\textwidth]{./figures/bound_vsys}
-  \caption{Normalized distribution of systemic velocities of binaries
-    with a MS star remaining bound after the CC of the first star. The
-    distributions in both panels are normalized to unity. The left panel shows
-    the distribution for NS companions, the right panel shows only
-    binaries with a BH companion. The blue histogram
-    represents our fiducial simulation. The orange solid line gives the
-    distribution when the kick is not rescaled because of
-    fallback, which changes the BH kicks, but not the NS kicks. The red distribution corresponds to a double Maxwellian
-    kick distribution, which changes the NS kicks but not the BH kicks. The top panels show the corresponding cumulative distributions.}
-  \label{fig:bound_v_dist}
-\end{figure*}
-
-\section{Systemic velocities of NS and BH binaries}
-\label{sec:bound}
-
-Our simulations also provide predictions for the systemic velocities of binary systems that remain bound after the first CC  \citep[e.g.,][]{vanoijen:89}.
-This concerns only \mbox{$1-\mathcal{D} = 14_{-10}^{14}\%$} of the systems (see \Secref{sec:param_variation} for a discussion of model variations).
-These bound systems are of large interest since they can give rise to X-ray binaries if the compact object starts to accrete from the companion.
-They typically originate from systems where the newly formed compact object received a small natal kick.
-This bias places  limitations on the extent to which these special systems can be used to infer conclusions about the properties of the general population of NS and BH.
-For comparison, the average effective natal kick for the compact objects formed in systems that remain bound is  $\sim$$66\,\mathrm{km\ s^{-1}}$, much lower than the average of $\sim$$330\,\mathrm{km\ s^{-1}}$ (as drawn from our standard Maxwellian after rescaling to account for fallback).
-
-Systems that remain bound also obtain a systemic velocity, because of
-mass lost from the progenitor star during the supernova explosion
-while orbiting around the center of mass.  We define the newly
-obtained systemic velocity $v_\mathrm{sys}$ as the velocity of the new
-center of mass (of the compact object plus the secondary star) in the
-frame of the old center of mass (e.g.,
-\citealt{brandt:95}, \citealt{kalogera:96}, \citetalias{tauris:98}).
-
-Figure~\ref{fig:bound_v_dist} shows the normalized distribution of
-systemic velocities $v_\mathrm{sys}$ for different natal kick
-assumptions. The left (right) panel shows NS+MS (BH+MS) binaries, with
-the cumulative distributions in the top panels, respectively.
-We find that the distribution of systemic velocities are very insensitive to the model variations that we considered, except when adopting extreme assumptions for the natal kicks.
-We only show the distributions for variations that show large deviations from the fiducial simulation.
-
-For NS+MS binaries  our fiducial simulation shows a median systemic velocity of $\sim 20\kms$, for nearly all assumptions that we consider.
-The strongest deviation is found with a run in which we adopt a double Maxwellian distribution, where low mass progenitors are assumed to form neutrons stars with very low kick velocities.
-This substantially increases the fraction of systems that remain bound to 35\%.
-It primarily adds systems with low systemic velocities.
-The median of the normalized distribution for this simulation lies near $7 \kms$.
-
-In our fiducial simulations for BH+MS binaries we obtain a bimodal
-distribution with a large peak at $0 \kms$ accounting for nearly 80\%
-of all systems and a second broader component peaking at $\sim
-10\kms$.  The peak near zero results from progenitors that experience
-complete fallback ($f_b=1$ according to the algorithm from
-\citealt{fryer:12}). They receive no kick at all, and no mass is lost from
-the system (we neglect here, as elsewhere, the mass possibly lost to
-neutrinos during the CC). For these systems we also find no changes in
-the separation and/or eccentricity after the CC of the first star.
-The sharp peak at $v_\mathrm{sys}\simeq0\,\kms$ would likely be
-smeared out by the velocity dispersion of the region where the system
-formed.
-
-Nearly all  variations considered give indistinguishable results.
-However, we find large deviations with the simulation where we used the extreme assumption that compact objects get kicks of a similar velocity amplitude regardless of their mass (no fall back down scaling, orange in \Figref{fig:bound_v_dist}).
-This effectively results in much larger kicks for black holes and is sometimes referred to as ``velocity kick''.
-Only $1-\mathcal{D}\simeq3\%$  of systems remain bound in this simulation.
-The few resulting BH binaries typically have large systemic velocities: we find a broad distribution with a median at $50\kms$.
-
-Our simulations also provide predictions for the distributions of further properties of the bound systems.  
-If we exclude systems with $f_b=1$, i.e.\ if we consider only NSs and
-BHs formed via fallback but still producing ejecta, we find
-an anti-correlation between the maximum post-CC separation and systemic
-velocity: the wider the system, the slower the maximum systemic
-velocity it can reach \citep[see also][]{brandt:95}.
-
-Most systems remaining bound come from initially 
-relatively tight pre-CC orbits, with typical periods $\log_{10}(P_\mathrm{pre-CC}/\mathrm{days})\simeq -2$. Typically, these very close systems have a reversed mass ratio
-(i.e., at the time of the first CC $M_2^\mathrm{pre-CC}>M_1^\mathrm{pre-CC}$)
-because of previous mass transfer phase. In our fiducial run, for the
-systems remaining bound the average pre-CC mass ratio is
-$\langle q_\mathrm{pre-CC}\rangle\equiv\langle
-M_2^\mathrm{pre-CC}/M_1^\mathrm{pre-CC}\rangle\simeq 2.9$.
-The majority of the bound systems, about $60\%$ for all variations
-with non-zero natal kicks, have separation of less than
-about $500\,R_\odot$. This suggests that these systems will likely
-evolve through a phase during which they might be detectable as X-ray
-sources, and possibly even through a mass transfer phase. The tail of the distribution at large separations can reach
-post-CC separation as large as $\sim6000\,R_\odot$.
-
-We do not consider cases in which the compact object is shot within the
-Roche lobe of the secondary. A prompt collision between the
-newly formed compact object and the companion star could lead to the
-formation of a Thore-Zytkow object \citep[if the compact object is a
-NS,][]{thorne:75,thorne:77, leonard:94}, or some other transient
-event, possibly involving the disintegration of the companion. For a given binary system, the probability of a prompt collision
-caused by the natal kick (whose direction is isotropically distributed
-in the frame of the collapsing star) can be estimated as the solid
-angle subtended by the companion at the position of the collapsing
-star. For our fiducial population, excluding systems that do not
-receive a natal kick because of the fallback rescaling, we obtain that
-roughly 1 out of 10\,000 CC events in a binary would result in a
-prompt collision (i.e., a probability per binary system of about
-$10^{-4}$). We exclude from the normalization systems merging before the first CC.
-
-\section{Astrophysical Implications}
-\label{sec:discussion}
-
-\subsection{How to identify walkaway stars}
-
-The velocity dispersion of OB stars in our galaxy is
-$\lesssim$10\,$\mathrm{km\ s^{-1}}$ (e.g.,
-\citealt{blaauw:56}; \citealt{gies:87}; \citetalias{tauris:98};
-\citealt{hoogerwerf:01}), and it is typically lower for OB
-associations 
-\citep[][]{debruijne:99,steenbrugge:03, kiminki:18}. We emphasize also that the velocity dispersion
-increases going from samples of OB stars in clusters, in associations,
-and in the field \citep[][]{gies:87}. The average velocity of ejected
-secondaries is slightly higher than this value
-(cf.~\Tabref{tab:parameters}), and if considering only the
-post-binary interaction secondaries (i.e.,~removing all the ejected
-secondaries from very wide and non-interacting binaries, cf.~dashed
-distributions in \Figref{fig:v_dist}) it increases slightly,
-especially for $M_\mathrm{dis}\geq7.5\,M_\odot$. Therefore, the
-companions ejected after a mass transfer episode might form a distinct
-population in the Gaia data. Moreover, regardless of their previous history, the walkaway
-and runaway stars from binary disruptions do not need to follow the
-local galactic rotation curve, and their motion can bring them in
-locations of lower extinction making them easier to see
-\citep[e.g.,][]{boubert:18,maiz-appellaniz:18}.
-
-\citet{tetzlaff:11}, following \cite{stone:91}, propose to fit the
-velocity distribution of young ($\lesssim50\,\mathrm{Myr}$) stars
-within 3\,Kpc of the Solar neighborhood using
-two Maxwellian: one for the ``low-velocity'' ($v\lesssim30\,\mathrm{km\
-  s^{-1}}$) and one for the ``high-velocity'' group. Our findings
-suggest that $\sim$$10\%$ of the O-stars in the low velocity group might still be the
-result of binary disruptions.
-
-Because of the binary interactions taking
-place before the first CC, these stars carry observational signatures which might
-make them recognizable even if they do not stand out from a kinematic
-point of view. For instance, if there is a mass transfer phase during
-the previous binary evolution, the ejected star will be spun up
-\citep[e.g.,][]{packet:81, pols:91,boubert:18} and
-possibly chemically polluted with He-rich and/or N-rich material \citep[e.g.,][]{blaauw:93}. Mass transfer
-might also be responsible for the presence of magnetic fields,
-\citep[e.g.][]{schneider:16}. These features should in principle leave
-observable signatures in the spectra of these stars \citep[e.g.,][]{maiz-appellaniz:18}. However, the
-effects of the metallicity gradient in the radial direction of the
-galaxy complicates the spectral identification of binary products, as pointed out by \cite{mcevoy:17}.
-
-Characterizing the population of stars ejected from isolated binary
-disruptions is necessary to reduce the number of false positive detections of hypervelocity stars in current and
-upcoming astrometric catalogues \citep[e.g.,][]{marchetti:17b}.
-
-Stars ejected by a successful SN explosion recently (i.e., less time
-ago than the lifetime of the SN remnant) can be connected to the SN
-location, and the associated NS can sometimes also be found. Assuming
-a typical visible lifetime of the SN remnant of
-$\tau_\mathrm{SNR} \simeq10^4$\,years, the distance traveled by
-an ejected companion is of order $D\equiv
-v_\mathrm{dis}\times\tau_\mathrm{SNR}\simeq
-0.1\,\mathrm{pc}$. Therefore runaways, but especially walkaways
-are expected to
-still reside within the SN remnant while it is still observable. This
-strategy has been successfully applied to search for runaways by
-\cite{vandenbergh:80,guseinov:05,tetzlaff:13,tetzlaff:14,dincel:15,boubert:17a}. Recently,
-\cite{kerzendorf:17} also used the predicted kinematics of disrupted
-binaries to search for companions of the star that exploded producing
-the supernova remnant Cas A. \cite{tetzlaff:11} also suggested that
-with precise astrometry, it might be possible to relate the ejected
-companion to the remnant of the companion up to $\sim$5\,Myr after the
-SN explosion, provided that the compact remnant of the former companion is visible (e.g., as a pulsar).
-
-The results of \cite{banerjee:12} and \cite{perets:12} suggest there
-might be a spatial distinction between dynamically ejected stars and
-post-binary runaways and walkaways. Dynamical interactions in a
-cluster are most efficient early in the cluster evolution \citep[e.g.,][]{oh:16},
-well before the first stellar CC event happens. Moreover, the median velocity
-of cluster ejection is higher than the typical walkaway
-velocity: \cite{banerjee:12} derived 
-a median velocity of dynamically ejected stars of
-$50\,\mathrm{km\ s^{-1}}$ after 3\,Myr of evolution. This value
- is significantly higher than what we find from binary disruptions. 
-The combination of these two effects would suggest that
-walkaways and runaways from binary disruptions would generally be closer to their parent cluster than
-dynamically ejected stars. However, this simplistic prediction is
-complicated by the fact that binaries can eject stars from the parent
-cluster outskirts, while most dynamical interactions produce runaways
-from the center of the cluster. Another complication is the possible
-ejection of stars as tidal tales in a cluster merger process, \citep[e.g.,][]{lucas:17}.
-
-\cite{evans:15} and \cite{neugent:18} have reported the first
-discovery of extragalactic post-MS massive runaways. They reported the
-observation of a red super giant in M31 and a yellow super giant in the SMC,
-respectively. They inferred peculiar space velocities larger than
-$\sim$$400\,\mathrm{km\ s^{-1}}$ and $\sim$$150\,\mathrm{km\ s^{-1}}$ for
-these stars, and proposed a binary origin at least for the yellow
-supergiant in the SMC. Our fiducial population can hardly reach such
-high peculiar velocities, unless the velocity of the system as a whole binary itself was already 
-high. Even if these stars were ejected from a binary during their main
-sequence, velocities in excess of a few hundred $\mathrm{km\ s^{-1}}$
-are only marginally reached in our simulations
-(cf.~\Figref{fig:v_dist},~\Figref{fig:v2_postMS}, and
-\Figref{fig:v_dist_log}), and this might suggest a different ejection
-mechanism (or combination of mechanisms) for these two stars. \cite{evans:15} also report that
-vast majority of O/B stars and red super giant stars in the Milky-Way have radial
-velocities lower than $\sim70\,\mathrm{km\ s^{-1}}$, in very good
-agreement with our results.
-
-\subsection{On the runaway fraction}
-\label{sec:howtofind}
-Our results predict a runaway fraction among O type stars of at best a
-few percent, in agreement with \cite{eldridge:11}, but in tension with
-the observational result that 
-$\sim$$10-20\%$ of O type stars are runaways \citep[][Sana et al., in
-preparation]{blaauw:61, gies:87, stone:91, tetzlaff:11, maiz-appellaniz:18}, and the claim
-that roughly two-thirds of runaways come from the
-disruption of binaries \citep[][]{hoogerwerf:01}. The latter has also been
-challenged by the more recent observations from \cite{jilinski:10}.
-Our results are roughly consistent with the following back of the
-envelope estimate for the runaway fraction:
-\begin{equation}
-  \label{eq:ratios}
-  f^\mathrm{RW} \simeq
-  f_\mathrm{bin}\times(1-f_\mathrm{mergers})\times\mathcal{D}\times
-  (1+f_\mathrm{acc})^{\alpha}\times f_{v>30} \ \ ,  %% the time is in
-                                %% the 1+f_acc factor!!
-\end{equation}
-where we can assume a binary fraction $f_\mathrm{bin}=1$, reasonable
-for the population of O type stars considering initial periods as
-large as $10^{5.5}$\,days, the fraction of binaries that do not merge is
-$1-f_\mathrm{mergers}\simeq0.8$ \citep[][]{sana:12}, the disruption
-fraction is $\mathcal{D}\simeq0.8$ \citep[this study, but also,
-e.g.,][]{eldridge:11}, the fraction of material accreted, which is not
-likely to exceed $f_\mathrm{acc}\simeq 0.2$ \citep[][]{packet:81}, and the
-fraction of ejected companions fast enough to be a runaway which we
-can estimate\footnote{We neglect here that $\mathcal{D}$ is the
-  disruption fraction including also binaries with no MS stars at the
-  pre-CC stage} from $\mathcal{D}$ and $\mathcal{R}_{15}$, and is
-$f_{v>30}\simeq \mathcal{D}/(1+\mathcal{R}_{15})\simeq0.04$ for our fiducial run. Assuming an IMF slope $\alpha=-2.3$ we obtain a
-runaway fraction of $f^\mathrm{RW}\simeq 0.01$, in reasonable
-agreement with our results.
-
-The tension between our results and the observed runaway fraction suggests that either (i)
-binaries tend to evolve towards shorter pre-CC orbits, corresponding
-to higher ejection velocities, but are still disrupted easily by the
-CC event, (ii) the contribution of dynamical ejections to the runaway
-population is presently underestimated, (iii) other ejection
-mechanisms exists and are presently overlooked, (iv) the velocities of
-massive runaways are presently overestimated in observations (e.g.,
-because of observational bias which hide more easily slow moving walkaways), or any
-combination of the previous.
-
-Since
-$v_\mathrm{dis}\simeq v_2^\mathrm{pre-CC}\propto
-v_\mathrm{orb}^\mathrm{pre-CC}\propto J_\mathrm{orb}$, with
-$J_\mathrm{orb}$ orbital angular momentum, to shift the
-peak of the velocity distribution in \Figref{fig:v_dist} from
-$\sim$$6\,\mathrm{km\ s^{-1}}$ to
-$v_\mathrm{dis}\gtrsim30\,\mathrm{km\ s^{-1}}$, the binaries would
-need to lose about five times more orbital angular momentum during the evolution
-before the collapse of the primary. This discrepancy
-deserves to be revisited once the runaway fraction can be evaluated from the
-homogeneous Gaia datasets. If the contribution from
-dynamical ejections can robustly be quantified, the 
-population of runaways might provide new constraints on the mass
-and angular momentum losses during mass transfer in
-binaries.
-
-A mechanism to remove angular momentum and mass from the
-binary presently not included in our population are eruptive mass loss
-events for which there is growing evidence from early observations of
-SNe \citep[e.g.][]{khazov:16}. Such mass loss events might be much
-more common than previously thought. One possible physical cause for
-these event are gravity waves
-excited by late shell burning depositing energy at the base of the
-envelope \citep[][]{quataert:12, fuller:17, fuller:18}. It is possible
-that the mechanism driving these mass loss events might also result in
-pre-CC binary interactions which might significantly change the pre-CC
-orbit. SN-impostor events happening earlier in the evolution might
-also lead to similar effects on the orbital evolution, which we
-presently do not model.
-
-Another possibility to increase the angular momentum losses could be
-magnetic braking. If non conservative mass transfer can simultaneously
-generate a magnetic field and eject some mass from the system, this
-might result in torques on the binary much larger than we consider here.
-
-\subsection{X-ray binaries and gravitational wave sources}
-
-Binaries remaining bound after the CC of the primary are a minority.
-
-Selecting from our results only bound systems consisting of a BH and a
-MS star, we obtain the systemic velocity distributions for BH X-ray
-binaries with different BH kick scenarios (see right panel of
-\Figref{fig:bound_v_dist}). Gaia will give astrometric constraints on the 19 known BH X-ray binary in the Galaxy: if the
-effect of the Galactic potential can be singled out from these data,
-the comparison with our distribution might shed light onto the typical
-amplitude of BH kicks \citep[e.g.,][]{fragos:09,repetto:12,Fragos+2013a,mandel:16}.
-
-The binaries surviving the first CC can
-become X-ray sources if their separation is
-short enough for the compact object to accrete during the subsequent evolution. For all our natal kick
-assumptions, the post-CC separation peaks at $\sim300\,R_\odot$,
-and most bound system have separations smaller than
-$\sim500\,R_\odot$, indicating that the majority of the systems
-remaining bound will become X-ray sources, and potentially even go
-through a common envelope evolution phase because of the large mass
-ratio between the compact object and the secondary star (which has
-accreted mass, cf.~\Secref{sec:example}).
-
-However, observed
-X-ray sources are necessarily on relatively short period orbits, or, in
-other words, the energy released at the CC of the first star did not
-widen significantly the binary and instead contributed mostly to the
-kinetic energy of the system. This could mean that the observed sample
-of galactic BH X-ray binaries is biased towards the high velocity tail
-of our distributions.
-
-System that might become X-ray sources have systemic velocities of a few tens of
-$\mathrm{km\ s^{-1}}$, and since the secondary star is also an
-accretor and rejuvenated, they are able to travel a
-distance comparable to $\langle L_\mathrm{walk}\rangle$,
-which can also increase their contribution as feedback engines
-\citep[e.g.,][]{justham:12,Fragos+2013a}.
-
-If a bound system is not disrupted by the first CC and does not merge
-during the evolution of the secondary star, it is less likely to be
-disrupted by the second SN \citep[e.g.][]{stevenson:17,oshaughnessy:17}. This is because the impact of the kick on the orbit
-scales as $v_k/v_\mathrm{orb}$, and the orbit of bound systems is much
-more likely to shrink during the evolution after the first CC: the
-donor star needs to become less massive than the compact object for
-the system to widen as described in
-\Secref{sec:example} and \Secref{sec:analytics}. Consequently, it is likely that most of the
-systems remaining bound after the first CC will also remain bound as
-double compact objects (and potential gravitational wave sources if
-the separation is small enough).
-
-\subsection{Runaways and walkaways as feedback engines}
-\label{sec:disrupted}
-
-The large majority of binary systems is disrupted at the first CC
-(with a few notable exceptions, see also \Secref{sec:param_variation}):
-systems that can become X-ray sources and/or gravitational wave
-sources are the exception rather than the rule. The constraints on binary evolution coming from the much more
-common evolutionary path should be considered together with those
-coming from the more rare channels.
-
-Binary interactions increase the lifetime of the ejected star because
-of rejuvenation: both walkaways
-and runaways can reach significant distances from their birth
-location, see \Tabref{tab:parameters} and \Figref{fig:Dhist}. The typical distances traveled
-by the ejected secondaries (cf.~\Figref{fig:Dhist}) are sufficient to
-get massive stars out of their parent cluster or association, but only
-the fastest runaways would get beyond the thick disk in our galaxy,
-especially if considering the relative inclination between the
-trajectory of the star and the disk itself. As noted by
-\cite{boubert:18}, this introduces a bias in observed samples that
-favors the detection of the fastest runaways.
-
-The motion of the ejected stars
-effectively spreads the massive stars and their (possible) SN explosions in a larger
-volume, possibly enhancing their impact on the composition of galaxies (more efficient mixing of
-nuclear yields), on the ionization of their surroundings (enhanced
-escape fraction, e.g.~\citealt{conroy:12, kimm:14, ma:16}), but it might possibly
-decrease the efficiency of their mechanical feedback because of the
-lower density of the surrounding gas.
-The large number of walkaways
-formed per each runaway $\mathcal{R}$ suggests that walkaway stars might
-enhance massive stars feedback more significantly than the faster runaways, although they cannot
-reach as far out as the latter because of their lower velocity.
-
-As a consequence of the typical evolutionary scenario
-described in \Secref{sec:example}, the first CC event in a
-binary system will result either in a stripped SN Ib/Ic or a failed SN
-with little or no electromagnetic signature. The second CC will
-instead be from the accretor star, which could again result in a
-SNIb/Ic (if the star becomes a Wolf-Rayet because of its own stellar wind),
-but, because of the IMF and the possibly non-fully conservative
-accretion rate, the secondary is more likely to remain H-rich until
-its CC. As already pointed out by
-\cite{eldridge:11}, this suggests that SNIb/Ic should preferentially
-happen in star forming regions, while type II SN should show a larger
-spatial spread (because of the motion of the ejected star).
-
-In this study, we did not consider the effects of the background
-cluster and/or galactic potential on the motion of the ejected
-secondaries. These can in principle be used as test particles to probe
-the potential itself, and consequently allow for testing the local
-and/or galactic dark matter distribution \citep[e.g.,][]{rossi:14,marchetti:17}.
-
-While runaways have peculiar kinematic %(and possibly spectal)
-characteristics that make them recognizable, walkaway stars might more
-easily be mistaken for genuine single stars. Especially masses larger
-than $7.5\,M_\odot$ and velocities larger than $v_\mathrm{dis}\gtrsim20\,\mathrm{km\
-  s^{-1}}$, the walkaways and runaways accrete mass before being shot
-out. Therefore, the disruption
-of binaries can
-pollute observed samples of
-present-day single stars with binary evolution by-products, by populating the field of the galaxy
-\citep[][]{gvaramadze:12,demink:14}.
-
-As an example, the runaway star $\zeta$ Puppis has been considered to be
-a ``canonical'' O-type star \citep[e.g.][]{ramiaramanantsoa:17}, and
-it was used to calibrate the free parameters in the \cite{castor:75}
-(CAK) theory of stellar winds \citep[][]{pauldrach:94} widely applied
-to model single stars. However, this object is
-likely to be a accretor ejected from a binary, \cite{vanrensbergen:96,ramiaramanantsoa:17}.
-
-Massive walkaway ejected from a binary might also appear as isolated
-massive stars. For massive stars observed in isolation, the question
-of their origin poses: did they form in isolation? or could they
-possibly have reached that location, if ejected from a binary? \cite{bestenlehner:11} identify an isolated
-$\sim150\,M_\odot$ star (VFTS682) in the 30 Doradus field. Such
-massive object is unlikely to be the result of a binary disruption,
-but might have been ejected from the massive young cluster
-R136. \cite{bestenlehner:11} suggest that if this star has not been
-ejected from this cluster, it might be a direct indication of isolated
-star formation. Walkaway massive stars can pose similar problems if
-not recognized as rejuvenated binary products.
-
-This might also connect to the evolutionary path of luminous blue
-variable (LBV) stars. These are massive stars which experience
-extreme and yet unexplained outbursts of mass loss. In the classical
-picture of massive stellar evolution (so-called ``Conti scenario''),
-LBVs are an intermediate evolutionary stage between O-type and Wolf-Rayet
-stars, during which the bulk of the H-rich envelope is lost \citep[e.g.][]{conti:75,
-  maeder:94,maeder:96}. However, \cite{smith:15}
-pointed out that LBV and LBV-candidates are farther from the nearest
-O-type star than Wolf-Rayet stars, and suggested that LBVs might be rejuvenated products of binary
-evolution\footnote{This idea has also be questioned by
-  \cite{humphreys:16,davidson:16}, see also \cite{smith:16}.} (accretors or merger
-products) with an increased lifetime to reach the observed isolation. The followup study of \cite{aghakhanlootakanloo:17}
-suggested that partial rejuvenation during mass transfer coupled to a walkaway (or
-runaway) velocity might be sufficient to explain the isolation
-observed by \cite{smith:15}. Alternatively, fully rejuvenated mergers
-might reach isolation just because of their velocity from the local dispersion.
-
-\section{Summary and conclusion}
-\label{sec:conclusion}
-
-In anticipation of the new data releases by the Gaia satellite we have
-carried out an extensive suite of numerical simulations to provide
-predictions for the kinematics and properties of massive stars ejected
-from binary systems disrupted by the supernova explosion of the
-companion star. Our aim with this study is three-fold:  (i) identify
-which theoretical predictions are robust against model uncertainties
-and discuss their astrophysical implications (ii)  investigate
-theoretically which uncertain physical processes affect the kinematics
-of the population of unbound stars in such a way that their imprints
-would be observable and  (iii) provide a framework of models that can
-be used for comparison against observations or as input for
-other simulations.
-
-We summarize our main findings below.
-When we quote percentages or fractions, the error bars given indicate the maximum variations that we encountered in our model variations for different physical assumptions (excluding the unphysical variation where we ignored all natal kicks for newly born compact objects).
-
-\begin{itemize}
-
-\item[$\bullet$] Nearly a quarter,   $22_{-8}^{+26}$\%, of all binary
-  systems with at least one star more massive initially than
-  $7.5\Msun$ merges prior to the explosion of the first star (see \Figref{fig:tree} and \Tabref{tab:parameters}).
-
-\item[$\bullet$] The large majority of binary systems is disrupted at the moment of core collapse of the first star, $\mathcal{D}=86_{-9}^{+11}\%$ in our simulations, consistent with earlier studies.
-This prediction is robust against variations in the treatment of the evolutionary and interaction processes, but it is sensitive to the choices concerning the natal kicks of neutron stars and black holes.
-
-\item[$\bullet$] The disruption of the binary system produces an
-  unbound stellar companion. Remarkably, we find that these events
-  only rarely produce runaway stars (i.e. stars with peculiar motions
-  larger than $30\,\kms$).  The velocity distribution peaks at
-  $\sim6\,\mathrm{km\  s^{-1}}$, with a $90^\mathrm{th}$ percentile
-  around $20\kms$ well below the typical threshold for runaway stars.
-
-\item[$\bullet$] The slow moving unbound companions, to which we refer
-  as walkaway stars, outnumber the runaway stars by at least an order
-  of magnitude in nearly all our models. This is a robust outcome of
-  the models and is primarily due to the reversal of the mass ratio
-  prior to the explosion as we show numerically and analytically.
-  This finding has been noted in earlier studies \citep[e.g., ][]{dedonder:97,eldridge:11},  but we believe this results (and its potential implications) are under appreciated.
-
-\item[$\bullet$] We also produce runaway stars but we rarely find them
-  to be faster than $60\,\kms$, which corresponds to
-  $98.9^\mathrm{th}$ of the population. This is much lower than the
-  theoretical maximum set by the finite size of the stars and their
-  orbit, which is in range 400-1000$\,\kms$ depending on their masses
-  and evolutionary state. We expect that stars with larger peculiar
-  velocities result from dynamical ejections from clusters, capture
-  from another galaxy, or most exotically, interaction with a
-  intermediate mass or supermassive black hole.
-
-\item[$\bullet$] We find a mild trend for runaway stars to be faster
-  and more common at lower metallicity, caused by the fact that stars are on average smaller and the reduced effects of stellar winds.
-Gaia should in principle be able to test these predictions by identifying the fastest runaway stars in the Magellanic clouds.
-In our most metal poor simulations for $Z=0.0002$, which may have relevance for the earliest stellar populations, we find 5 times more runaway stars, if we keep all other assumptions for the initial distributions the same.
-
-\item[$\bullet$] Both runaways and walkaways typically accrete mass from their
-companion prior to the disruption, This is especially true for those more massive than
-$7.5\,M_\odot$ at the time they are ejected.
-They typically also gain angular momentum in the process.
-These features might make walkaway stars recognizable as binary product even if they do not stand out from a kinematic point of view.
-Their further evolution and structure may also therefore differ from that of non rotating single stars, as has been speculated in the context of the LBV phenomenon.
-
-\item[$\bullet$] We suggest that the high mass tail of the runaway
-  mass function might provide insight on the formation mechanism of
-  black holes from the core-collapse of the primary star, and in particular on whether these receive a ``velocity'' or a ``momentum''
-kick.
-Studies so far focussed on the surviving X-ray binaries seeking for constraints, but we propose to use their unbound counter parts since they constitute the majority and should be identified by Gaia.
-
-\item[$\bullet$] By ejecting secondaries, the disruption of binaries can enhance the role of
-massive stars as feedback engines. The average distance they travel,
-calculated neglecting any external potential, is of order
-$\gtrsim100$\,pc, and up to $\gtrsim 500$\,pc if considering only the
-faster runaways. These suggests that both runaways and walkaways
-can exit the dust cloud embedding their birth location, thus enhancing
-the escape fraction for their ionizing radiation. If they are massive
-enough, the ejected star will explode far from other stars and gas
-overdensities, with potential implications for the driving of
-turbulence and star formation.
-
-\end{itemize}
-
-Our simulations will be made available upon final publication of this study.  Early requests for access can be addressed to the lead author.
-
-\begin{acknowledgements}
-  We acknowledge for the many helpful discussions
-  F.~Broekgaarden, A~.de~Koter, D.~Hendriks, L.~Kaper, N.~Langer, I.~Mandel, C.~J.~Neijssel, Ph.~Podsiadlowski,
-  K.~Sakar, F.~N.~Schneider, S.~N.~Shore, E.~van~den~Heuvel, A.~Vigna-Gomez.
-  MR is grateful to the KITP for the visiting graduate fellowship during which a large portion of this
-  study has been carried out. This research was supported in part by
-  the National Science Foundation under Grant No. NSF PHY11-25915. SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
-Council (ERC) (Grant agreement No. 715063).
-  RGI thanks the Science and Technology Facilities Council (STFC) for funding his Rutherford fellowship under grant ST/L003910/1.
-
-\end{acknowledgements}
-
 \bibliographystyle{aa}
-%\bibliography{bibliography/walkaway,bibliography/selma_bib}
-\begin{thebibliography}{242}
-\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
-
-\bibitem[{Abbott {et~al.}(2016)Abbott, Abbott, Abbott, Abernathy, Acernese,
-  Ackley, Adams, Adams, Addesso, Adhikari, Adya, Affeldt, Agathos, Agatsuma,
-  Aggarwal, Aguiar, Aiello, Ain, Ajith, Allen, Allocca, Altin, Anderson,
-  Anderson, Arai, Araya, Arceneaux, Areeda, Arnaud, Arun, Ascenzi, Ashton, Ast,
-  Aston, Astone, Aufmuth, Aulbert, Babak, Bacon, Bader, Baker, Baldaccini,
-  Ballardin, Ballmer, Barayoga, Barclay, Barish, Barker, Barone, Barr,
-  Barsotti, Barsuglia, Barta, Bartlett, Bartos, Bassiri, Basti, Batch, Baune,
-  Bavigadda, Bazzan, Behnke, Bejger, Belczynski, Bell, Bell, Berger, Bergman,
-  Bergmann, Berry, Bersanetti, Bertolini, Betzwieser, Bhagwat, Bhandare,
-  Bilenko, Billingsley, Birch, Birney, Biscans, Bisht, Bitossi, Biwer,
-  Bizouard, Blackburn, Blair, Blair, Blair, Bloemen, Bock, Bodiya, Boer,
-  Bogaert, Bogan, Bohe, Bojtos, Bond, Bondu, Bonnand, Boom, Bork, Boschi, Bose,
-  Bouffanais, Bozzi, Bradaschia, Brady, Braginsky, Branchesi, Brau, Briant,
-  Brillet, Brinkmann, Brisson, Brockill, Brooks, Brown, Brown, Brown, Buchanan,
-  Buikema, Bulik, Bulten, Buonanno, Buskulic, Buy, Byer, Cadonati, Cagnoli,
-  Cahillane, Bustillo, Callister, Calloni, Camp, Cannon, Cao, Capano, Capocasa,
-  Carbognani, Caride, Diaz, Casentini, Caudill, Cavaglià, Cavalier, Cavalieri,
-  Cella, Cepeda, Baiardi, Cerretani, Cesarini, Chakraborty, Chalermsongsak,
-  Chamberlin, Chan, Chao, Charlton, Chassande-Mottin, Chen, Chen, Cheng,
-  Chincarini, Chiummo, Cho, Cho, Chow, Christensen, Chu, Chua, Chung, Ciani,
-  Clara, Clark, Cleva, Coccia, Cohadon, Colla, Collette, Cominsky, Jr., Conte,
-  Conti, Cook, Corbitt, Cornish, Corsi, Cortese, Costa, Coughlin, Coughlin,
-  Coulon, Countryman, Couvares, Cowan, Coward, Cowart, Coyne, Coyne, Craig,
-  Creighton, Cripe, Crowder, Cumming, Cunningham, Cuoco, Canton, Danilishin,
-  D’Antonio, Danzmann, Darman, Dattilo, Dave, Daveloza, Davier, Davies, Daw,
-  Day, DeBra, Debreczeni, Degallaix, Laurentis, Deléglise, Pozzo, Denker,
-  Dent, Dereli, Dergachev, DeRosa, DeRosa, DeSalvo, Dhurandhar, Díaz, Fiore,
-  Giovanni, Lieto, Pace, Palma, Virgilio, Dojcinoski, Dolique, Donovan, Dooley,
-  Doravari, Douglas, Downes, Drago, Drever, Driggers, Du, Ducrot, Dwyer, Edo,
-  Edwards, Effler, Eggenstein, Ehrens, Eichholz, Eikenberry, Engels, Essick,
-  Etzel, Evans, Evans, Everett, Factourovich, Fafone, Fair, Fairhurst, Fan,
-  Fang, Farinon, Farr, Farr, Favata, Fays, Fehrmann, Fejer, Ferrante, Ferreira,
-  Ferrini, Fidecaro, Fiori, Fiorucci, Fisher, Flaminio, Fletcher, Fournier,
-  Franco, Frasca, Frasconi, Frei, Freise, Frey, Frey, Fricke, Fritschel,
-  Frolov, Fulda, Fyffe, Gabbard, Gair, Gammaitoni, Gaonkar, Garufi, Gatto,
-  Gaur, Gehrels, Gemme, Gendre, Genin, Gennai, George, Gergely, Germain, Ghosh,
-  Ghosh, Giaime, Giardina, Giazotto, Gill, Glaefke, Goetz, Goetz, Gondan,
-  González, Castro, Gopakumar, Gordon, Gorodetsky, Gossan, Gosselin, Gouaty,
-  Graef, Graff, Granata, Grant, Gras, Gray, Greco, Green, Groot, Grote,
-  Grunewald, Guidi, Guo, Gupta, Gupta, Gushwa, Gustafson, Gustafson, Hacker,
-  Hall, Hall, Hammond, Haney, Hanke, Hanks, Hanna, Hannam, Hanson, Hardwick,
-  Harms, Harry, Harry, Hart, Hartman, Haster, Haughian, Heidmann, Heintze,
-  Heitmann, Hello, Hemming, Hendry, Heng, Hennig, Heptonstall, Heurs, Hild,
-  Hoak, Hodge, Hofman, Hollitt, Holt, Holz, Hopkins, Hosken, Hough, Houston,
-  Howell, Hu, Huang, Huerta, Huet, Hughey, Husa, Huttner, Huynh-Dinh, Idrisy,
-  Indik, Ingram, Inta, Isa, Isac, Isi, Islas, Isogai, Iyer, Izumi, Jacqmin,
-  Jang, Jani, Jaranowski, Jawahar, Jiménez-Forteza, Johnson, Jones, Jones,
-  Jonker, Ju, K, Kalaghatgi, Kalogera, Kandhasamy, Kang, Kanner, Karki,
-  Kasprzack, Katsavounidis, Katzman, Kaufer, Kaur, Kawabe, Kawazoe, Kéfélian,
-  Kehl, Keitel, Kelley, Kells, Kennedy, Key, Khalaidovski, Khalili, Khan, Khan,
-  Khan, Khazanov, Kijbunchoo, Kim, Kim, Kim, Kim, Kim, Kim, King, King, Kinzel,
-  Kissel, Kleybolte, Klimenko, Koehlenbeck, Kokeyama, Koley, Kondrashov,
-  Kontos, Korobko, Korth, Kowalska, Kozak, Kringel, Krishnan, Królak, Krueger,
-  Kuehn, Kumar, Kuo, Kutynia, Lackey, Landry, Lange, Lantz, Lasky, Lazzarini,
-  Lazzaro, Leaci, Leavey, Lebigot, Lee, Lee, Lee, Lee, Lenon, Leonardi, Leong,
-  Leroy, Letendre, Levin, Levine, Li, Libson, Littenberg, Lockerbie, Logue,
-  Lombardi, Lord, Lorenzini, Loriette, Lormand, Losurdo, Lough, Lück,
-  Lundgren, Luo, Lynch, Ma, MacDonald, Machenschalk, MacInnis, Macleod,
-  Magaña-Sandoval, Magee, Mageswaran, Majorana, Maksimovic, Malvezzi, Man,
-  Mandel, Mandic, Mangano, Mansell, Manske, Mantovani, Marchesoni, Marion,
-  Márka, Márka, Markosyan, Maros, Martelli, Martellini, Martin, Martin,
-  Martynov, Marx, Mason, Masserot, Massinger, Masso-Reid, Matichard, Matone,
-  Mavalvala, Mazumder, Mazzolo, McCarthy, McClelland, McCormick, McGuire,
-  McIntyre, McIver, McManus, McWilliams, Meacher, Meadors, Meidam, Melatos,
-  Mendell, Mendoza-Gandara, Mercer, Merilh, Merzougui, Meshkov, Messenger,
-  Messick, Meyers, Mezzani, Miao, Michel, Middleton, Mikhailov, Milano, Miller,
-  Millhouse, Minenkov, Ming, Mirshekari, Mishra, Mitra, Mitrofanov,
-  Mitselmakher, Mittleman, Moggi, Mohan, Mohapatra, Montani, Moore, Moore,
-  Moraru, Moreno, Morriss, Mossavi, Mours, Mow-Lowry, Mueller, Mueller, Muir,
-  Mukherjee, Mukherjee, Mukherjee, Mukund, Mullavey, Munch, Murphy, Murray,
-  Mytidis, Nardecchia, Naticchioni, Nayak, Necula, Nedkova, Nelemans, Neri,
-  Neunzert, Newton, Nguyen, Nielsen, Nissanke, Nitz, Nocera, Nolting,
-  Normandin, Nuttall, Oberling, Ochsner, O’Dell, Oelker, Ogin, Oh, Oh, Ohme,
-  Oliver, Oppermann, Oram, O’Reilly, O’Shaughnessy, Ott, Ottaway, Ottens,
-  Overmier, Owen, Pai, Pai, Palamos, Palashov, Palomba, Pal-Singh, Pan, Pankow,
-  Pannarale, Pant, Paoletti, Paoli, Papa, Paris, Parker, Pascucci, Pasqualetti,
-  Passaquieti, Passuello, Patricelli, Patrick, Pearlstone, Pedraza, Pedurand,
-  Pekowsky, Pele, Penn, Perreca, Phelps, Piccinni, Pichot, Piergiovanni,
-  Pierro, Pillant, Pinard, Pinto, Pitkin, Poggiani, Popolizio, Post, Powell,
-  Prasad, Predoi, Premachandra, Prestegard, Price, Prijatelj, Principe,
-  Privitera, Prix, Prodi, Prokhorov, Puncken, Punturo, Puppo, Pürrer, Qi, Qin,
-  Quetschke, Quintero, Quitzow-James, Raab, Rabeling, Radkins, Raffai, Raja,
-  Rakhmanov, Rapagnani, Raymond, Razzano, Re, Read, Reed, Regimbau, Rei, Reid,
-  Reitze, Rew, Reyes, Ricci, Riles, Robertson, Robie, Robinet, Rocchi, Rolland,
-  Rollins, Roma, Romano, Romano, Romanov, Romie, Rosińska, Rowan, Rüdiger,
-  Ruggi, Ryan, Sachdev, Sadecki, Sadeghian, Salconi, Saleem, Salemi, Samajdar,
-  Sammut, Sanchez, Sandberg, Sandeen, Sanders, Sassolas, Sathyaprakash,
-  Saulson, Sauter, Savage, Sawadsky, Schale, Schilling, Schmidt, Schmidt,
-  Schnabel, Schofield, Schönbeck, Schreiber, Schuette, Schutz, Scott, Scott,
-  Sellers, Sentenac, Sequino, Sergeev, Serna, Setyawati, Sevigny, Shaddock,
-  Shah, Shahriar, Shaltev, Shao, Shapiro, Shawhan, Sheperd, Shoemaker,
-  Shoemaker, Siellez, Siemens, Sigg, Silva, Simakov, Singer, Singer, Singh,
-  Singh, Singhal, Sintes, Slagmolen, Smith, Smith, Smith, Son, Sorazu,
-  Sorrentino, Souradeep, Srivastava, Staley, Steinke, Steinlechner,
-  Steinlechner, Steinmeyer, Stephens, Stevenson, Stone, Strain, Straniero,
-  Stratta, Strauss, Strigin, Sturani, Stuver, Summerscales, Sun, Sutton,
-  Swinkels, Szczepańczyk, Tacca, Talukder, Tanner, Tápai, Tarabrin,
-  Taracchini, Taylor, Theeg, Thirugnanasambandam, Thomas, Thomas, Thomas,
-  Thorne, Thorne, Thrane, Tiwari, Tiwari, Tokmakov, Tomlinson, Tonelli, Torres,
-  Torrie, Töyrä, Travasso, Traylor, Trifirò, Tringali, Trozzo, Tse, Turconi,
-  Tuyenbayev, Ugolini, Unnikrishnan, Urban, Usman, Vahlbruch, Vajente, Valdes,
-  van Bakel, van Beuzekom, van~den Brand, van~den Broeck, Vander-Hyde, van~der
-  Schaaf, van Heijningen, van Veggel, Vardaro, Vass, Vasúth, Vaulin, Vecchio,
-  Vedovato, Veitch, Veitch, Venkateswara, Verkindt, Vetrano, Viceré,
-  Vinciguerra, Vine, Vinet, Vitale, Vo, Vocca, Vorvick, Voss, Vousden,
-  Vyatchanin, Wade, Wade, Wade, Walker, Wallace, Walsh, Wang, Wang, Wang, Wang,
-  Wang, Ward, Warner, Was, Weaver, Wei, Weinert, Weinstein, Weiss, Welborn,
-  Wen, Weßels, Westphal, Wette, Whelan, White, Whiting, Williams, Williamson,
-  Willis, Willke, Wimmer, Winkler, Wipf, Wittel, Woan, Worden, Wright, Wu,
-  Yablon, Yam, Yamamoto, Yancey, Yap, Yu, Yvert, Zadrożny, Zangrando, Zanolin,
-  Zendri, Zevin, Zhang, Zhang, Zhang, Zhang, Zhao, Zhou, Zhou, Zhu, Zucker,
-  Zuraw, Zweizig, Collaboration, \& Collaboration}]{LVC:16b}
-Abbott, B.~P., Abbott, R., Abbott, T.~D., {et~al.} 2016, \apjl, 818, L22
-
-\bibitem[{{Abbott} {et~al.}(2017{\natexlab{a}}){Abbott}, {Abbott}, {Abbott},
-  {Acernese}, {Ackley}, {Adams}, {Adams}, {Addesso}, {Adhikari}, {Adya}, \&
-  et~al.}]{LVC:17:grb}
-{Abbott}, B.~P., {Abbott}, R., {Abbott}, T.~D., {et~al.} 2017{\natexlab{a}},
-  \apjl, 848, L13
-
-\bibitem[{{Abbott} {et~al.}(2017{\natexlab{b}}){Abbott}, {Abbott}, {Abbott},
-  {Acernese}, {Ackley}, {Adams}, {Adams}, {Addesso}, {Adhikari}, {Adya}, \&
-  et~al.}]{BNSmerger:17}
-{Abbott}, B.~P., {Abbott}, R., {Abbott}, T.~D., {et~al.} 2017{\natexlab{b}},
-  Physical Review Letters, 119, 161101
-
-\bibitem[{{Aghakhanloo} {et~al.}(2017){Aghakhanloo}, {Murphy}, {Smith}, \&
-  {Hlo{\v z}ek}}]{aghakhanlootakanloo:17}
-{Aghakhanloo}, M., {Murphy}, J.~W., {Smith}, N., \& {Hlo{\v z}ek}, R. 2017,
-  \mnras, 472, 591
-
-\bibitem[{{Allison}(2012)}]{allison:12}
-{Allison}, R.~J. 2012, \mnras, 421, 3338
-
-\bibitem[{{Almeida} {et~al.}(2017){Almeida}, {Sana}, {Taylor}, {Barb{\'a}},
-  {Bonanos}, {Crowther}, {Damineli}, {de Koter}, {de Mink}, {Evans}, {Gieles},
-  {Grin}, {H{\'e}nault-Brunet}, {Langer}, {Lennon}, {Lockwood}, {Ma{\'{\i}}z
-  Apell{\'a}niz}, {Moffat}, {Neijssel}, {Norman}, {Ram{\'{\i}}rez-Agudelo},
-  {Richardson}, {Schootemeijer}, {Shenar}, {Soszy{\'n}ski}, {Tramper}, \&
-  {Vink}}]{almeida:17}
-{Almeida}, L.~A., {Sana}, H., {Taylor}, W., {et~al.} 2017, \aap, 598, A84
-
-\bibitem[{{Anders} \& {Grevesse}(1989)}]{anders:89}
-{Anders}, E. \& {Grevesse}, N. 1989, \gca, 53, 197
-
-\bibitem[{{Artymowicz} \& {Lubow}(1994)}]{artymowicz:94}
-{Artymowicz}, P. \& {Lubow}, S.~H. 1994, \apj, 421, 651
-
-\bibitem[{{Arzoumanian} {et~al.}(2002){Arzoumanian}, {Chernoff}, \&
-  {Cordes}}]{arzoumanian:02}
-{Arzoumanian}, Z., {Chernoff}, D.~F., \& {Cordes}, J.~M. 2002, \apj, 568, 289
-
-\bibitem[{{Asplund} {et~al.}(2009){Asplund}, {Grevesse}, {Sauval}, \&
-  {Scott}}]{asplund:09}
-{Asplund}, M., {Grevesse}, N., {Sauval}, A.~J., \& {Scott}, P. 2009, \araa, 47,
-  481
-
-\bibitem[{{Banerjee} {et~al.}(2012){Banerjee}, {Kroupa}, \& {Oh}}]{banerjee:12}
-{Banerjee}, S., {Kroupa}, P., \& {Oh}, S. 2012, \apj, 746, 15
-
-\bibitem[{{Bear} \& {Soker}(2017)}]{bear:17}
-{Bear}, E. \& {Soker}, N. 2017, \mnras, 468, 140
-
-\bibitem[{{Becerra} {et~al.}(2016){Becerra}, {Bianco}, {Fryer}, {Rueda}, \&
-  {Ruffini}}]{becerra:16}
-{Becerra}, L., {Bianco}, C.~L., {Fryer}, C.~L., {Rueda}, J.~A., \& {Ruffini},
-  R. 2016, \apj, 833, 107
-
-\bibitem[{{Belczy{\'n}ski} \& {Bulik}(1999)}]{belczynski:99}
-{Belczy{\'n}ski}, K. \& {Bulik}, T. 1999, \aap, 346, 91
-
-\bibitem[{{Belczynski} {et~al.}(2008){Belczynski}, {Kalogera}, {Rasio}, {Taam},
-  {Zezas}, {Bulik}, {Maccarone}, \& {Ivanova}}]{belczynski:08}
-{Belczynski}, K., {Kalogera}, V., {Rasio}, F.~A., {et~al.} 2008, \apjs, 174,
-  223
-
-\bibitem[{{Belczynski} {et~al.}(2017){Belczynski}, {Klencki}, {Meynet},
-  {Fryer}, {Brown}, {Chruslinska}, {Gladysz}, {O'Shaughnessy}, {Bulik},
-  {Berti}, {Holz}, {Gerosa}, {Giersz}, {Ekstrom}, {Georgy}, {Askar}, \&
-  {Lasota}}]{belczynski:17}
-{Belczynski}, K., {Klencki}, J., {Meynet}, G., {et~al.} 2017, arXiv:1706.07053
-
-\bibitem[{{Belczynski} {et~al.}(2012){Belczynski}, {Wiktorowicz}, {Fryer},
-  {Holz}, \& {Kalogera}}]{Belczynski+2012b}
-{Belczynski}, K., {Wiktorowicz}, G., {Fryer}, C.~L., {Holz}, D.~E., \&
-  {Kalogera}, V. 2012, \apj, 757, 91
-
-\bibitem[{{Beniamini} \& {Piran}(2016)}]{beniamini:16}
-{Beniamini}, P. \& {Piran}, T. 2016, \mnras, 456, 4089
-
-\bibitem[{{Bestenlehner} {et~al.}(2011){Bestenlehner}, {Vink}, {Gr{\"a}fener},
-  {Najarro}, {Evans}, {Bastian}, {Bonanos}, {Bressert}, {Crowther}, {Doran},
-  {Friedrich}, {H{\'e}nault-Brunet}, {Herrero}, {de Koter}, {Langer}, {Lennon},
-  {Ma{\'{\i}}z Apell{\'a}niz}, {Sana}, {Soszynski}, \&
-  {Taylor}}]{bestenlehner:11}
-{Bestenlehner}, J.~M., {Vink}, J.~S., {Gr{\"a}fener}, G., {et~al.} 2011, \aap,
-  530, L14
-
-\bibitem[{{Blaauw}(1956)}]{blaauw:56}
-{Blaauw}, A. 1956, \pasp, 68, 495
-
-\bibitem[{{Blaauw}(1961)}]{blaauw:61}
-{Blaauw}, A. 1961, \bain, 15, 265
-
-\bibitem[{{Blaauw}(1993)}]{blaauw:93}
-{Blaauw}, A. 1993, in Astronomical Society of the Pacific Conference Series,
-  Vol.~35, Massive Stars: Their Lives in the Interstellar Medium, ed. J.~P.
-  {Cassinelli} \& E.~B. {Churchwell}, 207
-
-\bibitem[{{Boersma}(1961)}]{boersma:61}
-{Boersma}, J. 1961, \bain, 15, 291
-
-\bibitem[{{Bolton}(1972)}]{Bolton1972}
-{Bolton}, C.~T. 1972, \nat, 235, 271
-
-\bibitem[{{Boubert} {et~al.}(2017{\natexlab{a}}){Boubert}, {Erkal}, {Evans}, \&
-  {Izzard}}]{boubert:17b}
-{Boubert}, D., {Erkal}, D., {Evans}, N.~W., \& {Izzard}, R.~G.
-  2017{\natexlab{a}}, \mnras, 469, 2151
-
-\bibitem[{{Boubert} \& {Evans}(2018)}]{boubert:18}
-{Boubert}, D. \& {Evans}, N.~W. 2018, ArXiv:1804.05849
-
-\bibitem[{{Boubert} {et~al.}(2017{\natexlab{b}}){Boubert}, {Fraser}, {Evans},
-  {Green}, \& {Izzard}}]{boubert:17a}
-{Boubert}, D., {Fraser}, M., {Evans}, N.~W., {Green}, D.~A., \& {Izzard}, R.~G.
-  2017{\natexlab{b}}, \aap, 606, A14
-
-\bibitem[{{Brandt} \& {Podsiadlowski}(1995)}]{brandt:95}
-{Brandt}, N. \& {Podsiadlowski}, P. 1995, \mnras, 274, 461
-
-\bibitem[{{Bray} \& {Eldridge}(2016)}]{bray:16}
-{Bray}, J.~C. \& {Eldridge}, J.~J. 2016, \mnras, 461, 3747
-
-\bibitem[{{Bray} \& {Eldridge}(2018)}]{bray:18}
-{Bray}, J.~C. \& {Eldridge}, J.~J. 2018, ArXiv:1804.04414
-
-\bibitem[{{Breivik} {et~al.}(2017){Breivik}, {Chatterjee}, \&
-  {Larson}}]{breivik:17}
-{Breivik}, K., {Chatterjee}, S., \& {Larson}, S.~L. 2017, \apjl, 850, L13
-
-\bibitem[{{Brott} {et~al.}(2011){Brott}, {de Mink}, {Cantiello}, {Langer}, {de
-  Koter}, {Evans}, {Hunter}, {Trundle}, \& {Vink}}]{brott:11}
-{Brott}, I., {de Mink}, S.~E., {Cantiello}, M., {et~al.} 2011, \aap, 530, A115
-
-\bibitem[{{Brown}(2015)}]{brown:15}
-{Brown}, W.~R. 2015, \araa, 53, 15
-
-\bibitem[{{Cacciari}(2009)}]{cacciari:09}
-{Cacciari}, C. 2009, \memsai, 80, 97
-
-\bibitem[{{Castor} {et~al.}(1975){Castor}, {Abbott}, \& {Klein}}]{castor:75}
-{Castor}, J.~I., {Abbott}, D.~C., \& {Klein}, R.~I. 1975, \apj, 195, 157
-
-\bibitem[{{Ceverino} \& {Klypin}(2009)}]{ceverino:09}
-{Ceverino}, D. \& {Klypin}, A. 2009, \apj, 695, 292
-
-\bibitem[{{Chan} {et~al.}(2018){Chan}, {M{\"u}ller}, {Heger}, {Pakmor}, \&
-  {Springel}}]{chan:18}
-{Chan}, C., {M{\"u}ller}, B., {Heger}, A., {Pakmor}, R., \& {Springel}, V.
-  2018, \apjl, 852, L19
-
-\bibitem[{{Chatterjee} {et~al.}(2005){Chatterjee}, {Vlemmings}, {Brisken},
-  {Lazio}, {Cordes}, {Goss}, {Thorsett}, {Fomalont}, {Lyne}, \&
-  {Kramer}}]{chatterjee:05}
-{Chatterjee}, S., {Vlemmings}, W.~H.~T., {Brisken}, W.~F., {et~al.} 2005,
-  \apjl, 630, L61
-
-\bibitem[{{Chini} {et~al.}(2012){Chini}, {Hoffmeister}, {Nasseri}, {Stahl}, \&
-  {Zinnecker}}]{chini:12}
-{Chini}, R., {Hoffmeister}, V.~H., {Nasseri}, A., {Stahl}, O., \& {Zinnecker},
-  H. 2012, \mnras, 424, 1925
-
-\bibitem[{{Claeys} {et~al.}(2014){Claeys}, {Pols}, {Izzard}, {Vink}, \&
-  {Verbunt}}]{claeys:14}
-{Claeys}, J.~S.~W., {Pols}, O.~R., {Izzard}, R.~G., {Vink}, J., \& {Verbunt},
-  F.~W.~M. 2014, \aap, 563, A83
-
-\bibitem[{{Conroy} \& {Kratter}(2012)}]{conroy:12}
-{Conroy}, C. \& {Kratter}, K.~M. 2012, \apj, 755, 123
-
-\bibitem[{{Conti}(1975)}]{conti:75}
-{Conti}, P.~S. 1975, Memoires of the Societe Royale des Sciences de Liege, 9,
-  193
-
-\bibitem[{{Cruz-Gonz{\'a}lez} {et~al.}(1974){Cruz-Gonz{\'a}lez},
-  {Recillas-Cruz}, {Costero}, {Peimbert}, \&
-  {Torres-Peimbert}}]{Cruz-Gonzalez+1974}
-{Cruz-Gonz{\'a}lez}, C., {Recillas-Cruz}, E., {Costero}, R., {Peimbert}, M., \&
-  {Torres-Peimbert}, S. 1974, \rmxaa, 1, 211
-
-\bibitem[{{Davidson} {et~al.}(2016){Davidson}, {Humphreys}, \&
-  {Weis}}]{davidson:16}
-{Davidson}, K., {Humphreys}, R.~M., \& {Weis}, K. 2016, arXiv:1608.02007
-
-\bibitem[{{de Bruijne}(1999)}]{debruijne:99}
-{de Bruijne}, J.~H.~J. 1999, \mnras, 310, 585
-
-\bibitem[{{De Donder} {et~al.}(1997){De Donder}, {Vanbeveren}, \& {van
-  Bever}}]{dedonder:97}
-{De Donder}, E., {Vanbeveren}, D., \& {van Bever}, J. 1997, \aap, 318, 812
-
-\bibitem[{{de Kool}(1990)}]{dekool:90}
-{de Kool}, M. 1990, \apj, 358, 189
-
-\bibitem[{{De Marco} \& {Izzard}(2017)}]{demarco:17}
-{De Marco}, O. \& {Izzard}, R.~G. 2017, \pasa, 34, e001
-
-\bibitem[{{de Mink} \& {Belczynski}(2015)}]{demink:15}
-{de Mink}, S.~E. \& {Belczynski}, K. 2015, \apj, 814, 58
-
-\bibitem[{{de Mink} {et~al.}(2012){de Mink}, {Brott}, {Cantiello}, {Izzard},
-  {Langer}, \& {Sana}}]{de-Mink+2012}
-{de Mink}, S.~E., {Brott}, I., {Cantiello}, M., {et~al.} 2012, in Astronomical
-  Society of the Pacific Conference Series, Vol. 465, Proceedings of a
-  Scientific Meeting in Honor of Anthony F. J. Moffat, ed. L.~{Drissen},
-  C.~{Robert}, N.~{St-Louis}, \& A.~F.~J. {Moffat}, 65
-
-\bibitem[{{de Mink} {et~al.}(2009){de Mink}, {Cantiello}, {Langer}, {Pols},
-  {Brott}, \& {Yoon}}]{demink:09}
-{de Mink}, S.~E., {Cantiello}, M., {Langer}, N., {et~al.} 2009, \aap, 497, 243
-
-\bibitem[{{de Mink} {et~al.}(2013){de Mink}, {Langer}, {Izzard}, {Sana}, \& {de
-  Koter}}]{demink:13}
-{de Mink}, S.~E., {Langer}, N., {Izzard}, R.~G., {Sana}, H., \& {de Koter}, A.
-  2013, \apj, 764, 166
-
-\bibitem[{{de Mink} {et~al.}(2007){de Mink}, {Pols}, \& {Hilditch}}]{demink:07}
-{de Mink}, S.~E., {Pols}, O.~R., \& {Hilditch}, R.~W. 2007, \aap, 467, 1181
-
-\bibitem[{{de Mink} {et~al.}(2014){de Mink}, {Sana}, {Langer}, {Izzard}, \&
-  {Schneider}}]{demink:14}
-{de Mink}, S.~E., {Sana}, H., {Langer}, N., {Izzard}, R.~G., \& {Schneider},
-  F.~R.~N. 2014, \apj, 782, 7
-
-\bibitem[{{de Wit} {et~al.}(2005){de Wit}, {Testi}, {Palla}, \&
-  {Zinnecker}}]{dewit:05}
-{de Wit}, W.~J., {Testi}, L., {Palla}, F., \& {Zinnecker}, H. 2005, \aap, 437,
-  247
-
-\bibitem[{{Dewi} \& {Tauris}(2000)}]{dewi:00}
-{Dewi}, J.~D.~M. \& {Tauris}, T.~M. 2000, \aap, 360, 1043
-
-\bibitem[{{Din{\c c}el} {et~al.}(2015){Din{\c c}el}, {Neuh{\"a}user}, {Yerli},
-  {Ankay}, {Tetzlaff}, {Torres}, \& {Mugrauer}}]{dincel:15}
-{Din{\c c}el}, B., {Neuh{\"a}user}, R., {Yerli}, S.~K., {et~al.} 2015, \mnras,
-  448, 3196
-
-\bibitem[{{Dominik} {et~al.}(2012){Dominik}, {Belczynski}, {Fryer}, {Holz},
-  {Berti}, {Bulik}, {Mandel}, \& {O'Shaughnessy}}]{dominik:12}
-{Dominik}, M., {Belczynski}, K., {Fryer}, C., {et~al.} 2012, \apj, 759, 52
-
-\bibitem[{{Dominik} {et~al.}(2013){Dominik}, {Belczynski}, {Fryer}, {Holz},
-  {Berti}, {Bulik}, {Mandel}, \& {O'Shaughnessy}}]{dominik:13}
-{Dominik}, M., {Belczynski}, K., {Fryer}, C., {et~al.} 2013, \apj, 779, 72
-
-\bibitem[{{Dray} {et~al.}(2005){Dray}, {Dale}, {Beer}, {Napiwotzki}, \&
-  {King}}]{dray:05}
-{Dray}, L.~M., {Dale}, J.~E., {Beer}, M.~E., {Napiwotzki}, R., \& {King}, A.~R.
-  2005, \mnras, 364, 59
-
-\bibitem[{{Eggleton}(1983)}]{eggleton:83}
-{Eggleton}, P.~P. 1983, \apj, 268, 368
-
-\bibitem[{{Eldridge}(2009)}]{eldridge:09}
-{Eldridge}, J.~J. 2009, \mnras, 400, L20
-
-\bibitem[{{Eldridge}(2012)}]{eldridge:12}
-{Eldridge}, J.~J. 2012, \mnras, 422, 794
-
-\bibitem[{{Eldridge} {et~al.}(2011){Eldridge}, {Langer}, \&
-  {Tout}}]{eldridge:11}
-{Eldridge}, J.~J., {Langer}, N., \& {Tout}, C.~A. 2011, \mnras, 414, 3501
-
-\bibitem[{{Evans} {et~al.}(2010){Evans}, {Walborn}, {Crowther},
-  {H{\'e}nault-Brunet}, {Massa}, {Taylor}, {Howarth}, {Sana}, {Lennon}, \& {van
-  Loon}}]{Evans+2010a}
-{Evans}, C.~J., {Walborn}, N.~R., {Crowther}, P.~A., {et~al.} 2010, \apjl, 715,
-  L74
-
-\bibitem[{{Evans} \& {Massey}(2015)}]{evans:15}
-{Evans}, K.~A. \& {Massey}, P. 2015, \aj, 150, 149
-
-\bibitem[{{Farr} {et~al.}(2011){Farr}, {Sravan}, {Cantrell}, {Kreidberg},
-  {Bailyn}, {Mandel}, \& {Kalogera}}]{farr:11}
-{Farr}, W.~M., {Sravan}, N., {Cantrell}, A., {et~al.} 2011, \apj, 741, 103
-
-\bibitem[{{Fender} {et~al.}(2013){Fender}, {Maccarone}, \&
-  {Heywood}}]{fender:13}
-{Fender}, R.~P., {Maccarone}, T.~J., \& {Heywood}, I. 2013, \mnras, 430, 1538
-
-\bibitem[{{Filippenko}(1997)}]{filippenko:97}
-{Filippenko}, A.~V. 1997, \araa, 35, 309
-
-\bibitem[{{Fragos} {et~al.}(2013){Fragos}, {Lehmer}, {Naoz}, {Zezas}, \&
-  {Basu-Zych}}]{Fragos+2013a}
-{Fragos}, T., {Lehmer}, B.~D., {Naoz}, S., {Zezas}, A., \& {Basu-Zych}, A.
-  2013, \apjl, 776, L31
-
-\bibitem[{{Fragos} {et~al.}(2009){Fragos}, {Willems}, {Kalogera}, {Ivanova},
-  {Rockefeller}, {Fryer}, \& {Young}}]{fragos:09}
-{Fragos}, T., {Willems}, B., {Kalogera}, V., {et~al.} 2009, \apj, 697, 1057
-
-\bibitem[{{Fryer} {et~al.}(1998){Fryer}, {Burrows}, \& {Benz}}]{fryer:98}
-{Fryer}, C., {Burrows}, A., \& {Benz}, W. 1998, \apj, 496, 333
-
-\bibitem[{{Fryer} \& {Kalogera}(1997)}]{Fryer+1997}
-{Fryer}, C. \& {Kalogera}, V. 1997, \apj, 489, 244
-
-\bibitem[{{Fryer}(2004)}]{fryer:04}
-{Fryer}, C.~L. 2004, \apjl, 601, L175
-
-\bibitem[{{Fryer} {et~al.}(2012){Fryer}, {Belczynski}, {Wiktorowicz},
-  {Dominik}, {Kalogera}, \& {Holz}}]{fryer:12}
-{Fryer}, C.~L., {Belczynski}, K., {Wiktorowicz}, G., {et~al.} 2012, \apj, 749,
-  91
-
-\bibitem[{{Fryer} \& {Kalogera}(2001)}]{Fryer+2001}
-{Fryer}, C.~L. \& {Kalogera}, V. 2001, \apj, 554, 548
-
-\bibitem[{{Fujii} \& {Portegies Zwart}(2011)}]{fujii:11}
-{Fujii}, M.~S. \& {Portegies Zwart}, S. 2011, Science, 334, 1380
-
-\bibitem[{{Fuller}(2017)}]{fuller:17}
-{Fuller}, J. 2017, \mnras, 470, 1642
-
-\bibitem[{{Fuller} \& {Ro}(2018)}]{fuller:18}
-{Fuller}, J. \& {Ro}, S. 2018, \mnras, 476, 1853
-
-\bibitem[{{Gaggero} {et~al.}(2017){Gaggero}, {Bertone}, {Calore}, {Connors},
-  {Lovell}, {Markoff}, \& {Storm}}]{gaggero:17}
-{Gaggero}, D., {Bertone}, G., {Calore}, F., {et~al.} 2017, Physical Review
-  Letters, 118, 241101
-
-\bibitem[{{Gaia Collaboration} {et~al.}(2016{\natexlab{a}}){Gaia
-  Collaboration}, {Brown}, {Vallenari}, {Prusti}, {de Bruijne}, {Mignard},
-  {Drimmel}, {Babusiaux}, {Bailer-Jones}, {Bastian}, {Biermann}, {Evans},
-  {Eyer}, {Jansen}, {Jordi}, {Katz}, {Klioner}, {Lammers}, {Lindegren}, {Luri},
-  {O'Mullane}, {Panem}, {Pourbaix}, {Randich}, {Sartoretti}, {Siddiqui},
-  {Soubiran}, {Valette}, {van Leeuwen}, {Walton}, {Aerts}, {Arenou}, {Cropper},
-  {H{\o}g}, {Lattanzi}, {Grebel}, {Holland}, {Huc}, {Passot}, {Perryman},
-  {Bramante}, {Cacciari}, {Casta{\~n}eda}, {Chaoul}, {Cheek}, {De Angeli},
-  {Fabricius}, {Guerra}, {Hern{\'a}ndez}, {Jean-Antoine-Piccolo}, {Masana},
-  {Messineo}, {Mowlavi}, {Nienartowicz}, {Ord{\'o}{\~n}ez-Blanco}, {Panuzzo},
-  {Portell}, {Richards}, {Riello}, {Seabroke}, {Tanga}, {Th{\'e}venin},
-  {Torra}, {Els}, {Gracia-Abril}, {Comoretto}, {Garcia-Reinaldos}, {Lock},
-  {Mercier}, {Altmann}, {Andrae}, {Astraatmadja}, {Bellas-Velidis}, {Benson},
-  {Berthier}, {Blomme}, {Busso}, {Carry}, {Cellino}, {Clementini}, {Cowell},
-  {Creevey}, {Cuypers}, {Davidson}, {De Ridder}, {de Torres}, {Delchambre},
-  {Dell'Oro}, {Ducourant}, {Fr{\'e}mat}, {Garc{\'{\i}}a-Torres}, {Gosset},
-  {Halbwachs}, {Hambly}, {Harrison}, {Hauser}, {Hestroffer}, {Hodgkin},
-  {Huckle}, {Hutton}, {Jasniewicz}, {Jordan}, {Kontizas}, {Korn}, {Lanzafame},
-  {Manteiga}, {Moitinho}, {Muinonen}, {Osinde}, {Pancino}, {Pauwels}, {Petit},
-  {Recio-Blanco}, {Robin}, {Sarro}, {Siopis}, {Smith}, {Smith}, {Sozzetti},
-  {Thuillot}, {van Reeven}, {Viala}, {Abbas}, {Abreu Aramburu}, {Accart},
-  {Aguado}, {Allan}, {Allasia}, {Altavilla}, {{\'A}lvarez}, {Alves},
-  {Anderson}, {Andrei}, {Anglada Varela}, {Antiche}, {Antoja}, {Ant{\'o}n},
-  {Arcay}, {Bach}, {Baker}, {Balaguer-N{\'u}{\~n}ez}, {Barache}, {Barata},
-  {Barbier}, {Barblan}, {Barrado y Navascu{\'e}s}, {Barros}, {Barstow},
-  {Becciani}, {Bellazzini}, {Bello Garc{\'{\i}}a}, {Belokurov}, {Bendjoya},
-  {Berihuete}, {Bianchi}, {Bienaym{\'e}}, {Billebaud}, {Blagorodnova},
-  {Blanco-Cuaresma}, {Boch}, {Bombrun}, {Borrachero}, {Bouquillon}, {Bourda},
-  {Bouy}, {Bragaglia}, {Breddels}, {Brouillet}, {Br{\"u}semeister},
-  {Bucciarelli}, {Burgess}, {Burgon}, {Burlacu}, {Busonero}, {Buzzi}, {Caffau},
-  {Cambras}, {Campbell}, {Cancelliere}, {Cantat-Gaudin}, {Carlucci},
-  {Carrasco}, {Castellani}, {Charlot}, {Charnas}, {Chiavassa}, {Clotet},
-  {Cocozza}, {Collins}, {Costigan}, {Crifo}, {Cross}, {Crosta}, {Crowley},
-  {Dafonte}, {Damerdji}, {Dapergolas}, {David}, {David}, {De Cat}, {de Felice},
-  {de Laverny}, {De Luise}, {De March}, {de Martino}, {de Souza}, {Debosscher},
-  {del Pozo}, {Delbo}, {Delgado}, {Delgado}, {Di Matteo}, {Diakite},
-  {Distefano}, {Dolding}, {Dos Anjos}, {Drazinos}, {Duran}, {Dzigan},
-  {Edvardsson}, {Enke}, {Evans}, {Eynard Bontemps}, {Fabre}, {Fabrizio},
-  {Faigler}, {Falc{\~a}o}, {Farr{\`a}s Casas}, {Federici}, {Fedorets},
-  {Fern{\'a}ndez-Hern{\'a}ndez}, {Fernique}, {Fienga}, {Figueras}, {Filippi},
-  {Findeisen}, {Fonti}, {Fouesneau}, {Fraile}, {Fraser}, {Fuchs}, {Gai},
-  {Galleti}, {Galluccio}, {Garabato}, {Garc{\'{\i}}a-Sedano}, {Garofalo},
-  {Garralda}, {Gavras}, {Gerssen}, {Geyer}, {Gilmore}, {Girona}, {Giuffrida},
-  {Gomes}, {Gonz{\'a}lez-Marcos}, {Gonz{\'a}lez-N{\'u}{\~n}ez},
-  {Gonz{\'a}lez-Vidal}, {Granvik}, {Guerrier}, {Guillout}, {Guiraud},
-  {G{\'u}rpide}, {Guti{\'e}rrez-S{\'a}nchez}, {Guy}, {Haigron},
-  {Hatzidimitriou}, {Haywood}, {Heiter}, {Helmi}, {Hobbs}, {Hofmann}, {Holl},
-  {Holland}, {Hunt}, {Hypki}, {Icardi}, {Irwin}, {Jevardat de Fombelle},
-  {Jofr{\'e}}, {Jonker}, {Jorissen}, {Julbe}, {Karampelas}, {Kochoska},
-  {Kohley}, {Kolenberg}, {Kontizas}, {Koposov}, {Kordopatis}, {Koubsky},
-  {Krone-Martins}, {Kudryashova}, {Kull}, {Bachchan}, {Lacoste-Seris}, {Lanza},
-  {Lavigne}, {Le Poncin-Lafitte}, {Lebreton}, {Lebzelter}, {Leccia}, {Leclerc},
-  {Lecoeur-Taibi}, {Lemaitre}, {Lenhardt}, {Leroux}, {Liao}, {Licata},
-  {Lindstr{\o}m}, {Lister}, {Livanou}, {Lobel}, {L{\"o}ffler}, {L{\'o}pez},
-  {Lorenz}, {MacDonald}, {Magalh{\~a}es Fernandes}, {Managau}, {Mann},
-  {Mantelet}, {Marchal}, {Marchant}, {Marconi}, {Marinoni}, {Marrese},
-  {Marschalk{\'o}}, {Marshall}, {Mart{\'{\i}}n-Fleitas}, {Martino}, {Mary},
-  {Matijevi{\v c}}, {Mazeh}, {McMillan}, {Messina}, {Michalik}, {Millar},
-  {Miranda}, {Molina}, {Molinaro}, {Molinaro}, {Moln{\'a}r}, {Moniez},
-  {Montegriffo}, {Mor}, {Mora}, {Morbidelli}, {Morel}, {Morgenthaler},
-  {Morris}, {Mulone}, {Muraveva}, {Musella}, {Narbonne}, {Nelemans},
-  {Nicastro}, {Noval}, {Ord{\'e}novic}, {Ordieres-Mer{\'e}}, {Osborne},
-  {Pagani}, {Pagano}, {Pailler}, {Palacin}, {Palaversa}, {Parsons}, {Pecoraro},
-  {Pedrosa}, {Pentik{\"a}inen}, {Pichon}, {Piersimoni}, {Pineau}, {Plachy},
-  {Plum}, {Poujoulet}, {Pr{\v s}a}, {Pulone}, {Ragaini}, {Rago}, {Rambaux},
-  {Ramos-Lerate}, {Ranalli}, {Rauw}, {Read}, {Regibo}, {Reyl{\'e}}, {Ribeiro},
-  {Rimoldini}, {Ripepi}, {Riva}, {Rixon}, {Roelens}, {Romero-G{\'o}mez},
-  {Rowell}, {Royer}, {Ruiz-Dern}, {Sadowski}, {Sagrist{\`a} Sell{\'e}s},
-  {Sahlmann}, {Salgado}, {Salguero}, {Sarasso}, {Savietto}, {Schultheis},
-  {Sciacca}, {Segol}, {Segovia}, {Segransan}, {Shih}, {Smareglia}, {Smart},
-  {Solano}, {Solitro}, {Sordo}, {Soria Nieto}, {Souchay}, {Spagna}, {Spoto},
-  {Stampa}, {Steele}, {Steidelm{\"u}ller}, {Stephenson}, {Stoev}, {Suess},
-  {S{\"u}veges}, {Surdej}, {Szabados}, {Szegedi-Elek}, {Tapiador}, {Taris},
-  {Tauran}, {Taylor}, {Teixeira}, {Terrett}, {Tingley}, {Trager}, {Turon},
-  {Ulla}, {Utrilla}, {Valentini}, {van Elteren}, {Van Hemelryck}, {van
-  Leeuwen}, {Varadi}, {Vecchiato}, {Veljanoski}, {Via}, {Vicente}, {Vogt},
-  {Voss}, {Votruba}, {Voutsinas}, {Walmsley}, {Weiler}, {Weingrill}, {Wevers},
-  {Wyrzykowski}, {Yoldas}, {{\v Z}erjal}, {Zucker}, {Zurbach}, {Zwitter},
-  {Alecu}, {Allen}, {Allende Prieto}, {Amorim}, {Anglada-Escud{\'e}},
-  {Arsenijevic}, {Azaz}, {Balm}, {Beck}, {Bernstein}, {Bigot}, {Bijaoui},
-  {Blasco}, {Bonfigli}, {Bono}, {Boudreault}, {Bressan}, {Brown}, {Brunet},
-  {Bunclark}, {Buonanno}, {Butkevich}, {Carret}, {Carrion}, {Chemin},
-  {Ch{\'e}reau}, {Corcione}, {Darmigny}, {de Boer}, {de Teodoro}, {de Zeeuw},
-  {Delle Luche}, {Domingues}, {Dubath}, {Fodor}, {Fr{\'e}zouls}, {Fries},
-  {Fustes}, {Fyfe}, {Gallardo}, {Gallegos}, {Gardiol}, {Gebran}, {Gomboc},
-  {G{\'o}mez}, {Grux}, {Gueguen}, {Heyrovsky}, {Hoar}, {Iannicola}, {Isasi
-  Parache}, {Janotto}, {Joliet}, {Jonckheere}, {Keil}, {Kim}, {Klagyivik},
-  {Klar}, {Knude}, {Kochukhov}, {Kolka}, {Kos}, {Kutka}, {Lainey}, {LeBouquin},
-  {Liu}, {Loreggia}, {Makarov}, {Marseille}, {Martayan}, {Martinez-Rubi},
-  {Massart}, {Meynadier}, {Mignot}, {Munari}, {Nguyen}, {Nordlander}, {Ocvirk},
-  {O'Flaherty}, {Olias Sanz}, {Ortiz}, {Osorio}, {Oszkiewicz}, {Ouzounis},
-  {Palmer}, {Park}, {Pasquato}, {Peltzer}, {Peralta}, {P{\'e}turaud},
-  {Pieniluoma}, {Pigozzi}, {Poels}, {Prat}, {Prod'homme}, {Raison}, {Rebordao},
-  {Risquez}, {Rocca-Volmerange}, {Rosen}, {Ruiz-Fuertes}, {Russo}, {Sembay},
-  {Serraller Vizcaino}, {Short}, {Siebert}, {Silva}, {Sinachopoulos}, {Slezak},
-  {Soffel}, {Sosnowska}, {Strai{\v z}ys}, {ter Linden}, {Terrell}, {Theil},
-  {Tiede}, {Troisi}, {Tsalmantza}, {Tur}, {Vaccari}, {Vachier}, {Valles}, {Van
-  Hamme}, {Veltz}, {Virtanen}, {Wallut}, {Wichmann}, {Wilkinson}, {Ziaeepour},
-  \& {Zschocke}}]{Gaia-Collaboration+2016}
-{Gaia Collaboration}, {Brown}, A.~G.~A., {Vallenari}, A., {et~al.}
-  2016{\natexlab{a}}, \aap, 595, A2
-
-\bibitem[{{Gaia Collaboration} {et~al.}(2016{\natexlab{b}}){Gaia
-  Collaboration}, {Prusti}, {de Bruijne}, {Brown}, {Vallenari}, {Babusiaux},
-  {Bailer-Jones}, {Bastian}, {Biermann}, {Evans}, {Eyer}, {Jansen}, {Jordi},
-  {Klioner}, {Lammers}, {Lindegren}, {Luri}, {Mignard}, {Milligan}, {Panem},
-  {Poinsignon}, {Pourbaix}, {Randich}, {Sarri}, {Sartoretti}, {Siddiqui},
-  {Soubiran}, {Valette}, {van Leeuwen}, {Walton}, {Aerts}, {Arenou}, {Cropper},
-  {Drimmel}, {H{\o}g}, {Katz}, {Lattanzi}, {O'Mullane}, {Grebel}, {Holland},
-  {Huc}, {Passot}, {Bramante}, {Cacciari}, {Casta{\~n}eda}, {Chaoul}, {Cheek},
-  {De Angeli}, {Fabricius}, {Guerra}, {Hern{\'a}ndez}, {Jean-Antoine-Piccolo},
-  {Masana}, {Messineo}, {Mowlavi}, {Nienartowicz}, {Ord{\'o}{\~n}ez-Blanco},
-  {Panuzzo}, {Portell}, {Richards}, {Riello}, {Seabroke}, {Tanga},
-  {Th{\'e}venin}, {Torra}, {Els}, {Gracia-Abril}, {Comoretto},
-  {Garcia-Reinaldos}, {Lock}, {Mercier}, {Altmann}, {Andrae}, {Astraatmadja},
-  {Bellas-Velidis}, {Benson}, {Berthier}, {Blomme}, {Busso}, {Carry},
-  {Cellino}, {Clementini}, {Cowell}, {Creevey}, {Cuypers}, {Davidson}, {De
-  Ridder}, {de Torres}, {Delchambre}, {Dell'Oro}, {Ducourant}, {Fr{\'e}mat},
-  {Garc{\'{\i}}a-Torres}, {Gosset}, {Halbwachs}, {Hambly}, {Harrison},
-  {Hauser}, {Hestroffer}, {Hodgkin}, {Huckle}, {Hutton}, {Jasniewicz},
-  {Jordan}, {Kontizas}, {Korn}, {Lanzafame}, {Manteiga}, {Moitinho},
-  {Muinonen}, {Osinde}, {Pancino}, {Pauwels}, {Petit}, {Recio-Blanco}, {Robin},
-  {Sarro}, {Siopis}, {Smith}, {Smith}, {Sozzetti}, {Thuillot}, {van Reeven},
-  {Viala}, {Abbas}, {Abreu Aramburu}, {Accart}, {Aguado}, {Allan}, {Allasia},
-  {Altavilla}, {{\'A}lvarez}, {Alves}, {Anderson}, {Andrei}, {Anglada Varela},
-  {Antiche}, {Antoja}, {Ant{\'o}n}, {Arcay}, {Atzei}, {Ayache}, {Bach},
-  {Baker}, {Balaguer-N{\'u}{\~n}ez}, {Barache}, {Barata}, {Barbier}, {Barblan},
-  {Baroni}, {Barrado y Navascu{\'e}s}, {Barros}, {Barstow}, {Becciani},
-  {Bellazzini}, {Bellei}, {Bello Garc{\'{\i}}a}, {Belokurov}, {Bendjoya},
-  {Berihuete}, {Bianchi}, {Bienaym{\'e}}, {Billebaud}, {Blagorodnova},
-  {Blanco-Cuaresma}, {Boch}, {Bombrun}, {Borrachero}, {Bouquillon}, {Bourda},
-  {Bouy}, {Bragaglia}, {Breddels}, {Brouillet}, {Br{\"u}semeister},
-  {Bucciarelli}, {Budnik}, {Burgess}, {Burgon}, {Burlacu}, {Busonero}, {Buzzi},
-  {Caffau}, {Cambras}, {Campbell}, {Cancelliere}, {Cantat-Gaudin}, {Carlucci},
-  {Carrasco}, {Castellani}, {Charlot}, {Charnas}, {Charvet}, {Chassat},
-  {Chiavassa}, {Clotet}, {Cocozza}, {Collins}, {Collins}, {Costigan}, {Crifo},
-  {Cross}, {Crosta}, {Crowley}, {Dafonte}, {Damerdji}, {Dapergolas}, {David},
-  {David}, {De Cat}, {de Felice}, {de Laverny}, {De Luise}, {De March}, {de
-  Martino}, {de Souza}, {Debosscher}, {del Pozo}, {Delbo}, {Delgado},
-  {Delgado}, {di Marco}, {Di Matteo}, {Diakite}, {Distefano}, {Dolding}, {Dos
-  Anjos}, {Drazinos}, {Dur{\'a}n}, {Dzigan}, {Ecale}, {Edvardsson}, {Enke},
-  {Erdmann}, {Escolar}, {Espina}, {Evans}, {Eynard Bontemps}, {Fabre},
-  {Fabrizio}, {Faigler}, {Falc{\~a}o}, {Farr{\`a}s Casas}, {Faye}, {Federici},
-  {Fedorets}, {Fern{\'a}ndez-Hern{\'a}ndez}, {Fernique}, {Fienga}, {Figueras},
-  {Filippi}, {Findeisen}, {Fonti}, {Fouesneau}, {Fraile}, {Fraser}, {Fuchs},
-  {Furnell}, {Gai}, {Galleti}, {Galluccio}, {Garabato}, {Garc{\'{\i}}a-Sedano},
-  {Gar{\'e}}, {Garofalo}, {Garralda}, {Gavras}, {Gerssen}, {Geyer}, {Gilmore},
-  {Girona}, {Giuffrida}, {Gomes}, {Gonz{\'a}lez-Marcos},
-  {Gonz{\'a}lez-N{\'u}{\~n}ez}, {Gonz{\'a}lez-Vidal}, {Granvik}, {Guerrier},
-  {Guillout}, {Guiraud}, {G{\'u}rpide}, {Guti{\'e}rrez-S{\'a}nchez}, {Guy},
-  {Haigron}, {Hatzidimitriou}, {Haywood}, {Heiter}, {Helmi}, {Hobbs},
-  {Hofmann}, {Holl}, {Holland}, {Hunt}, {Hypki}, {Icardi}, {Irwin}, {Jevardat
-  de Fombelle}, {Jofr{\'e}}, {Jonker}, {Jorissen}, {Julbe}, {Karampelas},
-  {Kochoska}, {Kohley}, {Kolenberg}, {Kontizas}, {Koposov}, {Kordopatis},
-  {Koubsky}, {Kowalczyk}, {Krone-Martins}, {Kudryashova}, {Kull}, {Bachchan},
-  {Lacoste-Seris}, {Lanza}, {Lavigne}, {Le Poncin-Lafitte}, {Lebreton},
-  {Lebzelter}, {Leccia}, {Leclerc}, {Lecoeur-Taibi}, {Lemaitre}, {Lenhardt},
-  {Leroux}, {Liao}, {Licata}, {Lindstr{\o}m}, {Lister}, {Livanou}, {Lobel},
-  {L{\"o}ffler}, {L{\'o}pez}, {Lopez-Lozano}, {Lorenz}, {Loureiro},
-  {MacDonald}, {Magalh{\~a}es Fernandes}, {Managau}, {Mann}, {Mantelet},
-  {Marchal}, {Marchant}, {Marconi}, {Marie}, {Marinoni}, {Marrese},
-  {Marschalk{\'o}}, {Marshall}, {Mart{\'{\i}}n-Fleitas}, {Martino}, {Mary},
-  {Matijevi{\v c}}, {Mazeh}, {McMillan}, {Messina}, {Mestre}, {Michalik},
-  {Millar}, {Miranda}, {Molina}, {Molinaro}, {Molinaro}, {Moln{\'a}r},
-  {Moniez}, {Montegriffo}, {Monteiro}, {Mor}, {Mora}, {Morbidelli}, {Morel},
-  {Morgenthaler}, {Morley}, {Morris}, {Mulone}, {Muraveva}, {Musella},
-  {Narbonne}, {Nelemans}, {Nicastro}, {Noval}, {Ord{\'e}novic},
-  {Ordieres-Mer{\'e}}, {Osborne}, {Pagani}, {Pagano}, {Pailler}, {Palacin},
-  {Palaversa}, {Parsons}, {Paulsen}, {Pecoraro}, {Pedrosa}, {Pentik{\"a}inen},
-  {Pereira}, {Pichon}, {Piersimoni}, {Pineau}, {Plachy}, {Plum}, {Poujoulet},
-  {Pr{\v s}a}, {Pulone}, {Ragaini}, {Rago}, {Rambaux}, {Ramos-Lerate},
-  {Ranalli}, {Rauw}, {Read}, {Regibo}, {Renk}, {Reyl{\'e}}, {Ribeiro},
-  {Rimoldini}, {Ripepi}, {Riva}, {Rixon}, {Roelens}, {Romero-G{\'o}mez},
-  {Rowell}, {Royer}, {Rudolph}, {Ruiz-Dern}, {Sadowski}, {Sagrist{\`a}
-  Sell{\'e}s}, {Sahlmann}, {Salgado}, {Salguero}, {Sarasso}, {Savietto},
-  {Schnorhk}, {Schultheis}, {Sciacca}, {Segol}, {Segovia}, {Segransan},
-  {Serpell}, {Shih}, {Smareglia}, {Smart}, {Smith}, {Solano}, {Solitro},
-  {Sordo}, {Soria Nieto}, {Souchay}, {Spagna}, {Spoto}, {Stampa}, {Steele},
-  {Steidelm{\"u}ller}, {Stephenson}, {Stoev}, {Suess}, {S{\"u}veges}, {Surdej},
-  {Szabados}, {Szegedi-Elek}, {Tapiador}, {Taris}, {Tauran}, {Taylor},
-  {Teixeira}, {Terrett}, {Tingley}, {Trager}, {Turon}, {Ulla}, {Utrilla},
-  {Valentini}, {van Elteren}, {Van Hemelryck}, {van Leeuwen}, {Varadi},
-  {Vecchiato}, {Veljanoski}, {Via}, {Vicente}, {Vogt}, {Voss}, {Votruba},
-  {Voutsinas}, {Walmsley}, {Weiler}, {Weingrill}, {Werner}, {Wevers},
-  {Whitehead}, {Wyrzykowski}, {Yoldas}, {{\v Z}erjal}, {Zucker}, {Zurbach},
-  {Zwitter}, {Alecu}, {Allen}, {Allende Prieto}, {Amorim},
-  {Anglada-Escud{\'e}}, {Arsenijevic}, {Azaz}, {Balm}, {Beck}, {Bernstein},
-  {Bigot}, {Bijaoui}, {Blasco}, {Bonfigli}, {Bono}, {Boudreault}, {Bressan},
-  {Brown}, {Brunet}, {Bunclark}, {Buonanno}, {Butkevich}, {Carret}, {Carrion},
-  {Chemin}, {Ch{\'e}reau}, {Corcione}, {Darmigny}, {de Boer}, {de Teodoro}, {de
-  Zeeuw}, {Delle Luche}, {Domingues}, {Dubath}, {Fodor}, {Fr{\'e}zouls},
-  {Fries}, {Fustes}, {Fyfe}, {Gallardo}, {Gallegos}, {Gardiol}, {Gebran},
-  {Gomboc}, {G{\'o}mez}, {Grux}, {Gueguen}, {Heyrovsky}, {Hoar}, {Iannicola},
-  {Isasi Parache}, {Janotto}, {Joliet}, {Jonckheere}, {Keil}, {Kim},
-  {Klagyivik}, {Klar}, {Knude}, {Kochukhov}, {Kolka}, {Kos}, {Kutka}, {Lainey},
-  {LeBouquin}, {Liu}, {Loreggia}, {Makarov}, {Marseille}, {Martayan},
-  {Martinez-Rubi}, {Massart}, {Meynadier}, {Mignot}, {Munari}, {Nguyen},
-  {Nordlander}, {Ocvirk}, {O'Flaherty}, {Olias Sanz}, {Ortiz}, {Osorio},
-  {Oszkiewicz}, {Ouzounis}, {Palmer}, {Park}, {Pasquato}, {Peltzer}, {Peralta},
-  {P{\'e}turaud}, {Pieniluoma}, {Pigozzi}, {Poels}, {Prat}, {Prod'homme},
-  {Raison}, {Rebordao}, {Risquez}, {Rocca-Volmerange}, {Rosen}, {Ruiz-Fuertes},
-  {Russo}, {Sembay}, {Serraller Vizcaino}, {Short}, {Siebert}, {Silva},
-  {Sinachopoulos}, {Slezak}, {Soffel}, {Sosnowska}, {Strai{\v z}ys}, {ter
-  Linden}, {Terrell}, {Theil}, {Tiede}, {Troisi}, {Tsalmantza}, {Tur},
-  {Vaccari}, {Vachier}, {Valles}, {Van Hamme}, {Veltz}, {Virtanen}, {Wallut},
-  {Wichmann}, {Wilkinson}, {Ziaeepour}, \&
-  {Zschocke}}]{Gaia-Collaboration+2016a}
-{Gaia Collaboration}, {Prusti}, T., {de Bruijne}, J.~H.~J., {et~al.}
-  2016{\natexlab{b}}, \aap, 595, A1
-
-\bibitem[{{Gatto} {et~al.}(2015){Gatto}, {Walch}, {Low}, {Naab}, {Girichidis},
-  {Glover}, {W{\"u}nsch}, {Klessen}, {Clark}, {Baczynski}, {Peters},
-  {Ostriker}, {Ib{\'a}{\~n}ez-Mej{\'{\i}}a}, \& {Haid}}]{gatto:15}
-{Gatto}, A., {Walch}, S., {Low}, M.-M.~M., {et~al.} 2015, \mnras, 449, 1057
-
-\bibitem[{{Gies}(1987)}]{gies:87}
-{Gies}, D.~R. 1987, \apjs, 64, 545
-
-\bibitem[{{Gies} \& {Bolton}(1986)}]{gies:86}
-{Gies}, D.~R. \& {Bolton}, C.~T. 1986, \apjs, 61, 419
-
-\bibitem[{{Gilmore} \& {Reid}(1983)}]{gilmore:83}
-{Gilmore}, G. \& {Reid}, N. 1983, \mnras, 202, 1025
-
-\bibitem[{{G{\"o}tberg} {et~al.}(2017){G{\"o}tberg}, {de Mink}, \&
-  {Groh}}]{gotberg:17}
-{G{\"o}tberg}, Y., {de Mink}, S.~E., \& {Groh}, J.~H. 2017, \aap, 608, A11
-
-\bibitem[{{G{\"o}tberg} {et~al.}(2018){G{\"o}tberg}, {de Mink}, {Groh},
-  {Kupfer}, {Crowther}, {Zapartas}, \& {Renzo}}]{gotberg:18}
-{G{\"o}tberg}, Y., {de Mink}, S.~E., {Groh}, J.~H., {et~al.} 2018,
-  ArXiv:1802.03018
-
-\bibitem[{{Gott}(1971)}]{gott:71}
-{Gott}, J.~R. 1971, \nat, 234, 342
-
-\bibitem[{{Gott} {et~al.}(1970){Gott}, {Gunn}, \& {Ostriker}}]{gott:70}
-{Gott}, III, J.~R., {Gunn}, J.~E., \& {Ostriker}, J.~P. 1970, \apjl, 160, L91
-
-\bibitem[{{Grefenstette} {et~al.}(2017){Grefenstette}, {Fryer}, {Harrison},
-  {Boggs}, {DeLaney}, {Laming}, {Reynolds}, {Alexander}, {Barret},
-  {Christensen}, {Craig}, {Forster}, {Giommi}, {Hailey}, {Hornstrup},
-  {Kitaguchi}, {Koglin}, {Lopez}, {Mao}, {Madsen}, {Miyasaka}, {Mori}, {Perri},
-  {Pivovaroff}, {Puccetti}, {Rana}, {Stern}, {Westergaard}, {Wik}, {Zhang}, \&
-  {Zoglauer}}]{grefenstette:17}
-{Grefenstette}, B.~W., {Fryer}, C.~L., {Harrison}, F.~A., {et~al.} 2017, \apj,
-  834, 19
-
-\bibitem[{{Grudzinska} {et~al.}(2015){Grudzinska}, {Belczynski}, {Casares}, {de
-  Mink}, {Ziolkowski}, {Negueruela}, {Rib{\'o}}, {Ribas}, {Paredes}, {Herrero},
-  \& {Benacquista}}]{Grudzinska+2015}
-{Grudzinska}, M., {Belczynski}, K., {Casares}, J., {et~al.} 2015, \mnras, 452,
-  2773
-
-\bibitem[{{Gull} \& {Sofia}(1979)}]{gull:79}
-{Gull}, T.~R. \& {Sofia}, S. 1979, \apj, 230, 782
-
-\bibitem[{{Gunn} \& {Ostriker}(1970)}]{gunn:70}
-{Gunn}, J.~E. \& {Ostriker}, J.~P. 1970, \apj, 160, 979
-
-\bibitem[{{Guseinov} {et~al.}(2005){Guseinov}, {Ankay}, \&
-  {Tagieva}}]{guseinov:05}
-{Guseinov}, O.~H., {Ankay}, A., \& {Tagieva}, S.~O. 2005, Astrophysics, 48, 330
-
-\bibitem[{{Gvaramadze} {et~al.}(2012){Gvaramadze}, {Weidner}, {Kroupa}, \&
-  {Pflamm-Altenburg}}]{gvaramadze:12}
-{Gvaramadze}, V.~V., {Weidner}, C., {Kroupa}, P., \& {Pflamm-Altenburg}, J.
-  2012, \mnras, 424, 3037
-
-\bibitem[{{Hamann} \& {Koesterke}(1998)}]{hamann:98}
-{Hamann}, W.~R. \& {Koesterke}, L. 1998, \aap, 335, 1003
-
-\bibitem[{{Heger} {et~al.}(2003){Heger}, {Fryer}, {Woosley}, {Langer}, \&
-  {Hartmann}}]{Heger+2003}
-{Heger}, A., {Fryer}, C.~L., {Woosley}, S.~E., {Langer}, N., \& {Hartmann},
-  D.~H. 2003, \apj, 591, 288
-
-\bibitem[{{Hellings}(1983)}]{hellings:83}
-{Hellings}, P. 1983, \apss, 96, 37
-
-\bibitem[{{Hills}(1988)}]{hills:88}
-{Hills}, J.~G. 1988, \nat, 331, 687
-
-\bibitem[{{Hirai} {et~al.}(2018){Hirai}, {Podsiadlowski}, \&
-  {Yamada}}]{hirai:18}
-{Hirai}, R., {Podsiadlowski}, P., \& {Yamada}, S. 2018, ArXiv:1803.10808
-
-\bibitem[{{Hobbs} {et~al.}(2005){Hobbs}, {Lorimer}, {Lyne}, \&
-  {Kramer}}]{hobbs:05}
-{Hobbs}, G., {Lorimer}, D.~R., {Lyne}, A.~G., \& {Kramer}, M. 2005, \mnras,
-  360, 974
-
-\bibitem[{{Holland-Ashford} {et~al.}(2017){Holland-Ashford}, {Lopez},
-  {Auchettl}, {Temim}, \& {Ramirez-Ruiz}}]{holland-ashford:17}
-{Holland-Ashford}, T., {Lopez}, L.~A., {Auchettl}, K., {Temim}, T., \&
-  {Ramirez-Ruiz}, E. 2017, \apj, 844, 84
-
-\bibitem[{{Hoogerwerf} {et~al.}(2000){Hoogerwerf}, {de Bruijne}, \& {de
-  Zeeuw}}]{hoogerwerf:00}
-{Hoogerwerf}, R., {de Bruijne}, J.~H.~J., \& {de Zeeuw}, P.~T. 2000, \apjl,
-  544, L133
-
-\bibitem[{{Hoogerwerf} {et~al.}(2001){Hoogerwerf}, {de Bruijne}, \& {de
-  Zeeuw}}]{hoogerwerf:01}
-{Hoogerwerf}, R., {de Bruijne}, J.~H.~J., \& {de Zeeuw}, P.~T. 2001, \aap, 365,
-  49
-
-\bibitem[{{Huang}(1963)}]{huang:63}
-{Huang}, S.-S. 1963, \apj, 138, 471
-
-\bibitem[{{Hulse} \& {Taylor}(1975)}]{hulse:75}
-{Hulse}, R.~A. \& {Taylor}, J.~H. 1975, \apjl, 195, L51
-
-\bibitem[{{Humphreys} {et~al.}(2016){Humphreys}, {Weis}, {Davidson}, \&
-  {Gordon}}]{humphreys:16}
-{Humphreys}, R.~M., {Weis}, K., {Davidson}, K., \& {Gordon}, M.~S. 2016, \apj,
-  825, 64
-
-\bibitem[{{Hurley} {et~al.}(2000){Hurley}, {Pols}, \& {Tout}}]{hurley:00}
-{Hurley}, J.~R., {Pols}, O.~R., \& {Tout}, C.~A. 2000, \mnras, 315, 543
-
-\bibitem[{{Hurley} {et~al.}(2002){Hurley}, {Tout}, \& {Pols}}]{hurley:02}
-{Hurley}, J.~R., {Tout}, C.~A., \& {Pols}, O.~R. 2002, \mnras, 329, 897
-
-\bibitem[{{Hut}(1981)}]{hut:81}
-{Hut}, P. 1981, \aap, 99, 126
-
-\bibitem[{{Ivanova} {et~al.}(2013){Ivanova}, {Justham}, {Chen}, {De Marco},
-  {Fryer}, {Gaburov}, {Ge}, {Glebbeek}, {Han}, {Li}, {Lu}, {Marsh},
-  {Podsiadlowski}, {Potter}, {Soker}, {Taam}, {Tauris}, {van den Heuvel}, \&
-  {Webbink}}]{ivanova:13}
-{Ivanova}, N., {Justham}, S., {Chen}, X., {et~al.} 2013, \aapr, 21, 59
-
-\bibitem[{{Izzard} {et~al.}(2006){Izzard}, {Dray}, {Karakas}, {Lugaro}, \&
-  {Tout}}]{izzard:06}
-{Izzard}, R.~G., {Dray}, L.~M., {Karakas}, A.~I., {Lugaro}, M., \& {Tout},
-  C.~A. 2006, \aap, 460, 565
-
-\bibitem[{{Izzard} {et~al.}(2009){Izzard}, {Glebbeek}, {Stancliffe}, \&
-  {Pols}}]{izzard:09}
-{Izzard}, R.~G., {Glebbeek}, E., {Stancliffe}, R.~J., \& {Pols}, O.~R. 2009,
-  \aap, 508, 1359
-
-\bibitem[{Izzard {et~al.}(2018)Izzard, Preece, Jofre, Halabi, Masseron, \&
-  Tout}]{izzard:18}
-Izzard, R.~G., Preece, H., Jofre, P., {et~al.} 2018, Monthly Notices of the
-  Royal Astronomical Society, 473, 2984
-
-\bibitem[{{Izzard} {et~al.}(2004{\natexlab{a}}){Izzard}, {Ramirez-Ruiz}, \&
-  {Tout}}]{izzard:04b}
-{Izzard}, R.~G., {Ramirez-Ruiz}, E., \& {Tout}, C.~A. 2004{\natexlab{a}},
-  \mnras, 348, 1215
-
-\bibitem[{{Izzard} {et~al.}(2004{\natexlab{b}}){Izzard}, {Tout}, {Karakas}, \&
-  {Pols}}]{izzard:04}
-{Izzard}, R.~G., {Tout}, C.~A., {Karakas}, A.~I., \& {Pols}, O.~R.
-  2004{\natexlab{b}}, \mnras, 350, 407
-
-\bibitem[{{Janka}(2013)}]{janka:13}
-{Janka}, H.-T. 2013, \mnras, 434, 1355
-
-\bibitem[{{Janka}(2017)}]{janka:17}
-{Janka}, H.-T. 2017, \apj, 837, 84
-
-\bibitem[{{Jeffries} {et~al.}(2014){Jeffries}, {Jackson}, {Cottaar}, {Koposov},
-  {Lanzafame}, {Meyer}, {Prisinzano}, {Randich}, {Sacco}, {Brugaletta},
-  {Caramazza}, {Damiani}, {Franciosini}, {Frasca}, {Gilmore}, {Feltzing},
-  {Micela}, {Alfaro}, {Bensby}, {Pancino}, {Recio-Blanco}, {de Laverny},
-  {Lewis}, {Magrini}, {Morbidelli}, {Costado}, {Jofr{\'e}}, {Klutsch}, {Lind},
-  \& {Maiorca}}]{Jeffries+2014}
-{Jeffries}, R.~D., {Jackson}, R.~J., {Cottaar}, M., {et~al.} 2014, \aap, 563,
-  A94
-
-\bibitem[{{Jilinski} {et~al.}(2010){Jilinski}, {Ortega}, {Drake}, \& {de la
-  Reza}}]{jilinski:10}
-{Jilinski}, E., {Ortega}, V.~G., {Drake}, N.~A., \& {de la Reza}, R. 2010,
-  \apj, 721, 469
-
-\bibitem[{{Johnston} {et~al.}(2005){Johnston}, {Hobbs}, {Vigeland}, {Kramer},
-  {Weisberg}, \& {Lyne}}]{johnston:05}
-{Johnston}, S., {Hobbs}, G., {Vigeland}, S., {et~al.} 2005, \mnras, 364, 1397
-
-\bibitem[{{Justham} \& {Schawinski}(2012)}]{justham:12}
-{Justham}, S. \& {Schawinski}, K. 2012, \mnras, 423, 1641
-
-\bibitem[{{Kalogera}(1996)}]{kalogera:96}
-{Kalogera}, V. 1996, \apj, 471, 352
-
-\bibitem[{{Kaplan} {et~al.}(2008){Kaplan}, {Chatterjee}, {Gaensler}, \&
-  {Anderson}}]{kaplan:08}
-{Kaplan}, D.~L., {Chatterjee}, S., {Gaensler}, B.~M., \& {Anderson}, J. 2008,
-  \apj, 677, 1201
-
-\bibitem[{{Kaspi} {et~al.}(1996){Kaspi}, {Bailes}, {Manchester}, {Stappers}, \&
-  {Bell}}]{kaspi:96}
-{Kaspi}, V.~M., {Bailes}, M., {Manchester}, R.~N., {Stappers}, B.~W., \&
-  {Bell}, J.~F. 1996, \nat, 381, 584
-
-\bibitem[{{Katsuda} {et~al.}(2018){Katsuda}, {Morii}, {Janka},
-  {Wongwathanarat}, {Nakamura}, {Kotake}, {Mori}, {M{\"u}ller}, {Takiwaki},
-  {Tanaka}, {Tominaga}, \& {Tsunemi}}]{katsuda:18}
-{Katsuda}, S., {Morii}, M., {Janka}, H.-T., {et~al.} 2018, \apj, 856, 18
-
-\bibitem[{{Katz}(1975)}]{katz:75}
-{Katz}, J.~I. 1975, \nat, 253, 698
-
-\bibitem[{{Kerzendorf} {et~al.}(2017){Kerzendorf}, {Do}, {de Mink},
-  {G{\"o}tberg}, {Millisaljevic}, {Zapartas}, {Renzo}, {Justham},
-  {Podsiadlowski}, \& {Fesen}}]{kerzendorf:17}
-{Kerzendorf}, W.~E., {Do}, T., {de Mink}, S.~E., {et~al.} 2017,
-  ArXiv:1711.00055
-
-\bibitem[{{Khazov} {et~al.}(2016){Khazov}, {Yaron}, {Gal-Yam}, {Manulis},
-  {Rubin}, {Kulkarni}, {Arcavi}, {Kasliwal}, {Ofek}, {Cao}, {Perley},
-  {Sollerman}, {Horesh}, {Sullivan}, {Filippenko}, {Nugent}, {Howell}, {Cenko},
-  {Silverman}, {Ebeling}, {Taddia}, {Johansson}, {Laher}, {Surace},
-  {Rebbapragada}, {Wozniak}, \& {Matheson}}]{khazov:16}
-{Khazov}, D., {Yaron}, O., {Gal-Yam}, A., {et~al.} 2016, \apj, 818, 3
-
-\bibitem[{{Kiminki} \& {Kobulnicky}(2012)}]{kiminki:12}
-{Kiminki}, D.~C. \& {Kobulnicky}, H.~A. 2012, \apj, 751, 4
-
-\bibitem[{{Kiminki} \& {Smith}(2018)}]{kiminki:18}
-{Kiminki}, M.~M. \& {Smith}, N. 2018, \mnras
-
-\bibitem[{{Kimm} \& {Cen}(2014)}]{kimm:14}
-{Kimm}, T. \& {Cen}, R. 2014, \apj, 788, 121
-
-\bibitem[{{Kippenhahn} \& {Weigert}(1967)}]{kippenhahn:67}
-{Kippenhahn}, R. \& {Weigert}, A. 1967, \zap, 65, 251
-
-\bibitem[{{Knigge} {et~al.}(2011){Knigge}, {Coe}, \&
-  {Podsiadlowski}}]{knigge:11}
-{Knigge}, C., {Coe}, M.~J., \& {Podsiadlowski}, P. 2011, \nat, 479, 372
-
-\bibitem[{{Kobulnicky} {et~al.}(2018){Kobulnicky}, {Chick}, \&
-  {Povich}}]{kobulnicky:18}
-{Kobulnicky}, H.~A., {Chick}, W.~T., \& {Povich}, M.~S. 2018, \apj, 856, 74
-
-\bibitem[{{Kobulnicky} \& {Fryer}(2007)}]{kobulnicky:07}
-{Kobulnicky}, H.~A. \& {Fryer}, C.~L. 2007, \apj, 670, 747
-
-\bibitem[{{Kobulnicky} {et~al.}(2014){Kobulnicky}, {Kiminki}, {Lundquist},
-  {Burke}, {Chapman}, {Keller}, {Lester}, {Rolen}, {Topel}, {Bhattacharjee},
-  {Smullen}, {Vargas {\'A}lvarez}, {Runnoe}, {Dale}, \&
-  {Brotherton}}]{kobulnicky:14}
-{Kobulnicky}, H.~A., {Kiminki}, D.~C., {Lundquist}, M.~J., {et~al.} 2014,
-  \apjs, 213, 34
-
-\bibitem[{{Kochanek} {et~al.}(2014){Kochanek}, {Adams}, \&
-  {Belczynski}}]{kochanek:14}
-{Kochanek}, C.~S., {Adams}, S.~M., \& {Belczynski}, K. 2014, \mnras, 443, 1319
-
-\bibitem[{{Kouwenhoven} {et~al.}(2005){Kouwenhoven}, {Brown}, {Zinnecker},
-  {Kaper}, \& {Portegies Zwart}}]{kouwenhoven:05}
-{Kouwenhoven}, M.~B.~N., {Brown}, A.~G.~A., {Zinnecker}, H., {Kaper}, L., \&
-  {Portegies Zwart}, S.~F. 2005, \aap, 430, 137
-
-\bibitem[{{Kroupa}(2001)}]{kroupa:01}
-{Kroupa}, P. 2001, \mnras, 322, 231
-
-\bibitem[{{Kusenko} \& {Segr{\`e}}(1996)}]{kusenko:96}
-{Kusenko}, A. \& {Segr{\`e}}, G. 1996, Physical Review Letters, 77, 4872
-
-\bibitem[{{Lamers}(2013)}]{lamers:13}
-{Lamers}, H.~J.~G.~L.~M. 2013, in Astronomical Society of the Pacific
-  Conference Series, Vol. 470, 370 Years of Astronomy in Utrecht, ed.
-  G.~{Pugliese}, A.~{de Koter}, \& M.~{Wijburg}, 97
-
-\bibitem[{{Langer}(2012)}]{langer:12}
-{Langer}, N. 2012, \araa, 50, 107
-
-\bibitem[{{Larson}(1974)}]{larson:74}
-{Larson}, R.~B. 1974, \mnras, 169, 229
-
-\bibitem[{{Lauterborn}(1970)}]{lauterborn:70}
-{Lauterborn}, D. 1970, \aap, 7, 150
-
-\bibitem[{{Lennon} {et~al.}(2016){Lennon}, {van der Marel}, {Ramos Lerate},
-  {O'Mullane}, \& {Sahlmann}}]{lennon:16}
-{Lennon}, D.~J., {van der Marel}, R.~P., {Ramos Lerate}, M., {O'Mullane}, W.,
-  \& {Sahlmann}, J. 2016, arXiv:1611.05504
-
-\bibitem[{{Leonard}(1991)}]{leonard:91}
-{Leonard}, P.~J.~T. 1991, \aj, 101, 562
-
-\bibitem[{{Leonard} {et~al.}(1994){Leonard}, {Hills}, \& {Dewey}}]{leonard:94}
-{Leonard}, P.~J.~T., {Hills}, J.~G., \& {Dewey}, R.~J. 1994, \apjl, 423, L19
-
-\bibitem[{{Liu} {et~al.}(2015){Liu}, {Tauris}, {R{\"o}pke}, {Moriya},
-  {Kruckow}, {Stancliffe}, \& {Izzard}}]{liu:15}
-{Liu}, Z.-W., {Tauris}, T.~M., {R{\"o}pke}, F.~K., {et~al.} 2015, \aap, 584,
-  A11
-
-\bibitem[{{Livio} \& {Soker}(1988)}]{livio:88}
-{Livio}, M. \& {Soker}, N. 1988, \apj, 329, 764
-
-\bibitem[{{Lucas} {et~al.}(2018){Lucas}, {Rybak}, {Bonnell}, \&
-  {Gieles}}]{lucas:17}
-{Lucas}, W.~E., {Rybak}, M., {Bonnell}, I.~A., \& {Gieles}, M. 2018, \mnras,
-  474, 3582
-
-\bibitem[{{Lyne} \& {Lorimer}(1994)}]{lyne:94}
-{Lyne}, A.~G. \& {Lorimer}, D.~R. 1994, \nat, 369, 127
-
-\bibitem[{{Ma} {et~al.}(2016{\natexlab{a}}){Ma}, {Hopkins}, {Kasen},
-  {Quataert}, {Faucher-Gigu{\`e}re}, {Kere{\v s}}, {Murray}, \&
-  {Strom}}]{Ma+2016}
-{Ma}, X., {Hopkins}, P.~F., {Kasen}, D., {et~al.} 2016{\natexlab{a}}, \mnras,
-  459, 3614
-
-\bibitem[{{Ma} {et~al.}(2016{\natexlab{b}}){Ma}, {Hopkins}, {Kasen},
-  {Quataert}, {Faucher-Gigu{\`e}re}, {Kere{\v s}}, {Murray}, \&
-  {Strom}}]{ma:16}
-{Ma}, X., {Hopkins}, P.~F., {Kasen}, D., {et~al.} 2016{\natexlab{b}}, \mnras,
-  459, 3614
-
-\bibitem[{{Maeder}(1996)}]{maeder:96}
-{Maeder}, A. 1996, in Liege International Astrophysical Colloquia, Vol.~33,
-  Liege International Astrophysical Colloquia, ed. J.~M. {Vreux}, A.~{Detal},
-  D.~{Fraipont-Caro}, E.~{Gosset}, \& G.~{Rauw}, 39
-
-\bibitem[{{Maeder} \& {Conti}(1994)}]{maeder:94}
-{Maeder}, A. \& {Conti}, P.~S. 1994, \araa, 32, 227
-
-\bibitem[{{Maeder} \& {Meynet}(2000)}]{maeder:00}
-{Maeder}, A. \& {Meynet}, G. 2000, \aap, 361, 159
-
-\bibitem[{{Ma{\'{\i}}z Apell{\'a}niz} {et~al.}(2018){Ma{\'{\i}}z
-  Apell{\'a}niz}, {Pantaleoni Gonz{\'a}lez}, {Barb{\'a}},
-  {Sim{\'o}n-D{\'{\i}}az}, {Negueruela}, {Lennon}, {Sota}, \& {Trigueros
-  P{\'a}ez}}]{maiz-appellaniz:18}
-{Ma{\'{\i}}z Apell{\'a}niz}, J., {Pantaleoni Gonz{\'a}lez}, M., {Barb{\'a}},
-  R.~H., {et~al.} 2018, ArXiv:1804.06915
-
-\bibitem[{{Mandel}(2016)}]{mandel:16}
-{Mandel}, I. 2016, \mnras, 456, 578
-
-\bibitem[{{Maoil{\'e}idigh} {et~al.}(2005){Maoil{\'e}idigh}, {Meurs}, \&
-  {Norci}}]{maoileidigh:05}
-{Maoil{\'e}idigh}, C.~{\'O}., {Meurs}, E.~J.~A., \& {Norci}, L. 2005, \na, 10,
-  591
-
-\bibitem[{{Marchetti} {et~al.}(2017){Marchetti}, {Rossi}, {Kordopatis},
-  {Brown}, {Rimoldi}, {Starkenburg}, {Youakim}, \& {Ashley}}]{marchetti:17}
-{Marchetti}, T., {Rossi}, E.~M., {Kordopatis}, G., {et~al.} 2017, \mnras, 470,
-  1388
-
-\bibitem[{Marchetti {et~al.}(2017)Marchetti, Rossi, Kordopatis, Brown, Rimoldi,
-  Starkenburg, Youakim, \& Ashley}]{marchetti:17b}
-Marchetti, T., Rossi, E.~M., Kordopatis, G., {et~al.} 2017, Monthly Notices of
-  the Royal Astronomical Society, 470, 1388
-
-\bibitem[{{Martin} {et~al.}(2009){Martin}, {Tout}, \& {Pringle}}]{martin:09}
-{Martin}, R.~G., {Tout}, C.~A., \& {Pringle}, J.~E. 2009, \mnras, 397, 1563
-
-\bibitem[{{Mason} {et~al.}(2009){Mason}, {Hartkopf}, {Gies}, {Henry}, \&
-  {Helsel}}]{mason:09}
-{Mason}, B.~D., {Hartkopf}, W.~I., {Gies}, D.~R., {Henry}, T.~J., \& {Helsel},
-  J.~W. 2009, \aj, 137, 3358
-
-\bibitem[{McEvoy {et~al.}(2017)McEvoy, Dufton, Smoker, Lambert, Keenan,
-  Schneider, \& de~Wit}]{mcevoy:17}
-McEvoy, C.~M., Dufton, P.~L., Smoker, J.~V., {et~al.} 2017, The Astrophysical
-  Journal, 842, 32
-
-\bibitem[{{Moe} \& {Di Stefano}(2017)}]{moe:17}
-{Moe}, M. \& {Di Stefano}, R. 2017, \apjs, 230, 15
-
-\bibitem[{{Morningstar} {et~al.}(2014){Morningstar}, {Miller}, {Reis}, \&
-  {Ebisawa}}]{morningstar:14}
-{Morningstar}, W.~R., {Miller}, J.~M., {Reis}, R.~C., \& {Ebisawa}, K. 2014,
-  \apjl, 784, L18
-
-\bibitem[{{Neo} {et~al.}(1977){Neo}, {Miyaji}, {Nomoto}, \&
-  {Sugimoto}}]{neo:77}
-{Neo}, S., {Miyaji}, S., {Nomoto}, K., \& {Sugimoto}, D. 1977, \pasj, 29, 249
-
-\bibitem[{{Neugent} {et~al.}(2018){Neugent}, {Massey}, {Morrell}, {Skiff}, \&
-  {Georgy}}]{neugent:18}
-{Neugent}, K., {Massey}, P., {Morrell}, N., {Skiff}, B., \& {Georgy}, C. 2018,
-  ArXiv:1803.02859
-
-\bibitem[{{Nieuwenhuijzen} \& {de Jager}(1990)}]{nieuwenhuijzen:90}
-{Nieuwenhuijzen}, H. \& {de Jager}, C. 1990, \aap, 231, 134
-
-\bibitem[{{Nomoto} {et~al.}(2013){Nomoto}, {Kobayashi}, \&
-  {Tominaga}}]{nomoto:13}
-{Nomoto}, K., {Kobayashi}, C., \& {Tominaga}, N. 2013, \araa, 51, 457
-
-\bibitem[{{Oh} \& {Kroupa}(2016)}]{oh:16}
-{Oh}, S. \& {Kroupa}, P. 2016, \aap, 590, A107
-
-\bibitem[{{{\"O}pik}(1924)}]{opik:24}
-{{\"O}pik}, E. 1924, Publications of the Tartu Astrofizica Observatory, 25
-
-\bibitem[{{O'Shaughnessy} {et~al.}(2017){O'Shaughnessy}, {Gerosa}, \&
-  {Wysocki}}]{oshaughnessy:17}
-{O'Shaughnessy}, R., {Gerosa}, D., \& {Wysocki}, D. 2017, Physical Review
-  Letters, 119, 011101
-
-\bibitem[{{Packet}(1981)}]{packet:81}
-{Packet}, W. 1981, \aap, 102, 17
-
-\bibitem[{{Pauldrach} {et~al.}(1994){Pauldrach}, {Kudritzki}, {Puls}, {Butler},
-  \& {Hunsinger}}]{pauldrach:94}
-{Pauldrach}, A.~W.~A., {Kudritzki}, R.~P., {Puls}, J., {Butler}, K., \&
-  {Hunsinger}, J. 1994, \aap, 283, 525
-
-\bibitem[{{Perets} \& {{\v S}ubr}(2012)}]{perets:12}
-{Perets}, H.~B. \& {{\v S}ubr}, L. 2012, \apj, 751, 133
-
-\bibitem[{{Perryman} {et~al.}(2001){Perryman}, {de Boer}, {Gilmore}, {H{\o}g},
-  {Lattanzi}, {Lindegren}, {Luri}, {Mignard}, {Pace}, \& {de
-  Zeeuw}}]{perryman:01}
-{Perryman}, M.~A.~C., {de Boer}, K.~S., {Gilmore}, G., {et~al.} 2001, \aap,
-  369, 339
-
-\bibitem[{{Pfahl} {et~al.}(2002){Pfahl}, {Rappaport}, {Podsiadlowski}, \&
-  {Spruit}}]{pfahl:02}
-{Pfahl}, E., {Rappaport}, S., {Podsiadlowski}, P., \& {Spruit}, H. 2002, \apj,
-  574, 364
-
-\bibitem[{{Podsiadlowski} {et~al.}(2004){Podsiadlowski}, {Langer},
-  {Poelarends}, {Rappaport}, {Heger}, \& {Pfahl}}]{podsiadlowski:04}
-{Podsiadlowski}, P., {Langer}, N., {Poelarends}, A.~J.~T., {et~al.} 2004, \apj,
-  612, 1044
-
-\bibitem[{{Pols}(1994)}]{pols:94}
-{Pols}, O.~R. 1994, \aap, 290, 119
-
-\bibitem[{{Pols} {et~al.}(1991){Pols}, {Cote}, {Waters}, \& {Heise}}]{pols:91}
-{Pols}, O.~R., {Cote}, J., {Waters}, L.~B.~F.~M., \& {Heise}, J. 1991, \aap,
-  241, 419
-
-\bibitem[{{Pols} {et~al.}(1998){Pols}, {Schr{\"o}der}, {Hurley}, {Tout}, \&
-  {Eggleton}}]{pols:98}
-{Pols}, O.~R., {Schr{\"o}der}, K.-P., {Hurley}, J.~R., {Tout}, C.~A., \&
-  {Eggleton}, P.~P. 1998, \mnras, 298, 525
-
-\bibitem[{{Poveda} {et~al.}(1967){Poveda}, {Ruiz}, \& {Allen}}]{poveda:67}
-{Poveda}, A., {Ruiz}, J., \& {Allen}, C. 1967, Boletin de los Observatorios
-  Tonantzintla y Tacubaya, 4, 86
-
-\bibitem[{{Quataert} \& {Shiode}(2012)}]{quataert:12}
-{Quataert}, E. \& {Shiode}, J. 2012, \mnras, 423, L92
-
-\bibitem[{{Ramiaramanantsoa} {et~al.}(2018){Ramiaramanantsoa}, {Moffat},
-  {Harmon}, {Ignace}, {St-Louis}, {Vanbeveren}, {Shenar}, {Pablo},
-  {Richardson}, {Howarth}, {Stevens}, {Piaulet}, {St-Jean}, {Eversberg},
-  {Pigulski}, {Popowicz}, {Kuschnig}, {Zoc{\l}o{\'n}ska}, {Buysschaert},
-  {Handler}, {Weiss}, {Wade}, {Rucinski}, {Zwintz}, {Luckas}, {Heathcote},
-  {Cacella}, {Powles}, {Locke}, {Bohlsen}, {Chen{\'e}}, {Miszalski}, {Waldron},
-  {Kotze}, {Kotze}, \& {B{\"o}hm}}]{ramiaramanantsoa:17}
-{Ramiaramanantsoa}, T., {Moffat}, A.~F.~J., {Harmon}, R., {et~al.} 2018,
-  \mnras, 473, 5532
-
-\bibitem[{{Ram{\'{\i}}rez-Agudelo} {et~al.}(2015){Ram{\'{\i}}rez-Agudelo},
-  {Sana}, {de Mink}, {H{\'e}nault-Brunet}, {de Koter}, {Langer}, {Tramper},
-  {Gr{\"a}fener}, {Evans}, {Vink}, {Dufton}, \& {Taylor}}]{ramirez-agudelo:15}
-{Ram{\'{\i}}rez-Agudelo}, O.~H., {Sana}, H., {de Mink}, S.~E., {et~al.} 2015,
-  \aap, 580, A92
-
-\bibitem[{{Remillard} \& {McClintock}(2006)}]{remillard:06}
-{Remillard}, R.~A. \& {McClintock}, J.~E. 2006, \araa, 44, 49
-
-\bibitem[{{Renzo} {et~al.}(2017){Renzo}, {Ott}, {Shore}, \& {de
-  Mink}}]{renzo:17}
-{Renzo}, M., {Ott}, C.~D., {Shore}, S.~N., \& {de Mink}, S.~E. 2017, \aap, 603,
-  A118
-
-\bibitem[{{Repetto} {et~al.}(2012){Repetto}, {Davies}, \&
-  {Sigurdsson}}]{repetto:12}
-{Repetto}, S., {Davies}, M.~B., \& {Sigurdsson}, S. 2012, \mnras, 425, 2799
-
-\bibitem[{{Repetto} {et~al.}(2017){Repetto}, {Igoshev}, \&
-  {Nelemans}}]{repetto:17}
-{Repetto}, S., {Igoshev}, A.~P., \& {Nelemans}, G. 2017, \mnras, 467, 298
-
-\bibitem[{{Repetto} \& {Nelemans}(2015)}]{repetto:15}
-{Repetto}, S. \& {Nelemans}, G. 2015, \mnras, 453, 3341
-
-\bibitem[{{Rimoldi} {et~al.}(2016){Rimoldi}, {Portegies Zwart}, \&
-  {Rossi}}]{rimoldi:16}
-{Rimoldi}, A., {Portegies Zwart}, S., \& {Rossi}, E.~M. 2016, Computational
-  Astrophysics and Cosmology, 3, 2
-
-\bibitem[{{Rossi} {et~al.}(2014){Rossi}, {Kobayashi}, \& {Sari}}]{rossi:14}
-{Rossi}, E.~M., {Kobayashi}, S., \& {Sari}, R. 2014, \apj, 795, 125
-
-\bibitem[{{Sana} {et~al.}(2012){Sana}, {de Mink}, {de Koter}, {Langer},
-  {Evans}, {Gieles}, {Gosset}, {Izzard}, {Le Bouquin}, \&
-  {Schneider}}]{sana:12}
-{Sana}, H., {de Mink}, S.~E., {de Koter}, A., {et~al.} 2012, Science, 337, 444
-
-\bibitem[{{Sana} \& {Evans}(2011)}]{sana:11}
-{Sana}, H. \& {Evans}, C.~J. 2011, in IAU Symposium, Vol. 272, Active OB Stars:
-  Structure, Evolution, Mass Loss, and Critical Limits, ed. C.~{Neiner},
-  G.~{Wade}, G.~{Meynet}, \& G.~{Peters}, 474--485
-
-\bibitem[{{Sana} {et~al.}(2014){Sana}, {Le Bouquin}, {Lacour}, {Berger},
-  {Duvert}, {Gauchet}, {Norris}, {Olofsson}, {Pickel}, {Zins}, {Absil}, {de
-  Koter}, {Kratter}, {Schnurr}, \& {Zinnecker}}]{sana:14}
-{Sana}, H., {Le Bouquin}, J.-B., {Lacour}, S., {et~al.} 2014, \apjs, 215, 15
-
-\bibitem[{{Schneider} {et~al.}(2015){Schneider}, {Izzard}, {Langer}, \& {de
-  Mink}}]{schneider:15}
-{Schneider}, F.~R.~N., {Izzard}, R.~G., {Langer}, N., \& {de Mink}, S.~E. 2015,
-  \apj, 805, 20
-
-\bibitem[{{Schneider} {et~al.}(2016){Schneider}, {Podsiadlowski}, {Langer},
-  {Castro}, \& {Fossati}}]{schneider:16}
-{Schneider}, F.~R.~N., {Podsiadlowski}, P., {Langer}, N., {Castro}, N., \&
-  {Fossati}, L. 2016, \mnras, 457, 2355
-
-\bibitem[{{Schneider} {et~al.}(2018){Schneider}, {Sana}, {Evans},
-  {Bestenlehner}, {Castro}, {Fossati}, {Gr{\"a}fener}, {Langer},
-  {Ram{\'{\i}}rez-Agudelo}, {Sab{\'{\i}}n-Sanjuli{\'a}n},
-  {Sim{\'o}n-D{\'{\i}}az}, {Tramper}, {Crowther}, {de Koter}, {de Mink},
-  {Dufton}, {Garcia}, {Gieles}, {H{\'e}nault-Brunet}, {Herrero}, {Izzard},
-  {Kalari}, {Lennon}, {Ma{\'{\i}}z Apell{\'a}niz}, {Markova}, {Najarro},
-  {Podsiadlowski}, {Puls}, {Taylor}, {van Loon}, {Vink}, \&
-  {Norman}}]{schneider:18}
-{Schneider}, F.~R.~N., {Sana}, H., {Evans}, C.~J., {et~al.} 2018, Science, 359,
-  69
-
-\bibitem[{{Schwab} {et~al.}(2010){Schwab}, {Podsiadlowski}, \&
-  {Rappaport}}]{schwab:10}
-{Schwab}, J., {Podsiadlowski}, P., \& {Rappaport}, S. 2010, \apj, 719, 722
-
-\bibitem[{{Shklovskii}(1970)}]{shklovskii:70}
-{Shklovskii}, I.~S. 1970, \sovast, 13, 562
-
-\bibitem[{{Smith}(2016)}]{smith:16}
-{Smith}, N. 2016, \mnras, 461, 3353
-
-\bibitem[{{Smith} \& {Tombleson}(2015)}]{smith:15}
-{Smith}, N. \& {Tombleson}, R. 2015, \mnras, 447, 598
-
-\bibitem[{{Soberman} {et~al.}(1997){Soberman}, {Phinney}, \& {van den
-  Heuvel}}]{soberman:97}
-{Soberman}, G.~E., {Phinney}, E.~S., \& {van den Heuvel}, E.~P.~J. 1997, \aap,
-  327, 620
-
-\bibitem[{{Socrates} {et~al.}(2005){Socrates}, {Blaes}, {Hungerford}, \&
-  {Fryer}}]{socrates:05}
-{Socrates}, A., {Blaes}, O., {Hungerford}, A., \& {Fryer}, C.~L. 2005, \apj,
-  632, 531
-
-\bibitem[{{Steenbrugge} {et~al.}(2003){Steenbrugge}, {de Bruijne},
-  {Hoogerwerf}, \& {de Zeeuw}}]{steenbrugge:03}
-{Steenbrugge}, K.~C., {de Bruijne}, J.~H.~J., {Hoogerwerf}, R., \& {de Zeeuw},
-  P.~T. 2003, \aap, 402, 587
-
-\bibitem[{{Stevenson} {et~al.}(2017){Stevenson}, {Vigna-G{\'o}mez}, {Mandel},
-  {Barrett}, {Neijssel}, {Perkins}, \& {de Mink}}]{stevenson:17}
-{Stevenson}, S., {Vigna-G{\'o}mez}, A., {Mandel}, I., {et~al.} 2017, Nature
-  Communications, 8, 14906
-
-\bibitem[{{Stone}(1991)}]{stone:91}
-{Stone}, R.~C. 1991, \aj, 102, 333
-
-\bibitem[{{Tauris}(2015)}]{tauris:15}
-{Tauris}, T.~M. 2015, \mnras, 448, L6
-
-\bibitem[{{Tauris} {et~al.}(2017){Tauris}, {Kramer}, {Freire}, {Wex}, {Janka},
-  {Langer}, {Podsiadlowski}, {Bozzo}, {Chaty}, {Kruckow}, {van den Heuvel},
-  {Antoniadis}, {Breton}, \& {Champion}}]{tauris:17}
-{Tauris}, T.~M., {Kramer}, M., {Freire}, P.~C.~C., {et~al.} 2017, \apj, 846,
-  170
-
-\bibitem[{{Tauris} \& {Takens}(1998)}]{tauris:98}
-{Tauris}, T.~M. \& {Takens}, R.~J. 1998, \aap, 330, 1047
-
-\bibitem[{{Tetzlaff} {et~al.}(2014){Tetzlaff}, {Din{\c c}el}, {Neuh{\"a}user},
-  \& {Kovtyukh}}]{tetzlaff:14}
-{Tetzlaff}, N., {Din{\c c}el}, B., {Neuh{\"a}user}, R., \& {Kovtyukh}, V.~V.
-  2014, \mnras, 438, 3587
-
-\bibitem[{{Tetzlaff} {et~al.}(2011){Tetzlaff}, {Neuh{\"a}user}, \&
-  {Hohle}}]{tetzlaff:11}
-{Tetzlaff}, N., {Neuh{\"a}user}, R., \& {Hohle}, M.~M. 2011, \mnras, 410, 190
-
-\bibitem[{{Tetzlaff} {et~al.}(2013){Tetzlaff}, {Torres}, {Neuh{\"a}user}, \&
-  {Hohle}}]{tetzlaff:13}
-{Tetzlaff}, N., {Torres}, G., {Neuh{\"a}user}, R., \& {Hohle}, M.~M. 2013,
-  \mnras, 435, 879
-
-\bibitem[{{Thorne} \& {Zytkow}(1975)}]{thorne:75}
-{Thorne}, K.~S. \& {Zytkow}, A.~N. 1975, \apjl, 199, L19
-
-\bibitem[{{Thorne} \& {Zytkow}(1977)}]{thorne:77}
-{Thorne}, K.~S. \& {Zytkow}, A.~N. 1977, \apj, 212, 832
-
-\bibitem[{{Tout} {et~al.}(1997){Tout}, {Aarseth}, {Pols}, \&
-  {Eggleton}}]{tout:97}
-{Tout}, C.~A., {Aarseth}, S.~J., {Pols}, O.~R., \& {Eggleton}, P.~P. 1997,
-  \mnras, 291, 732
-
-\bibitem[{{Tutukov} \& {Yungelson}(1973)}]{tutukov:73}
-{Tutukov}, A. \& {Yungelson}, L. 1973, Nauchnye Informatsii, 27, 70
-
-\bibitem[{{van den Bergh}(1980)}]{vandenbergh:80}
-{van den Bergh}, S. 1980, Journal of Astrophysics and Astronomy, 1, 67
-
-\bibitem[{{van den Heuvel} \& {Heise}(1972)}]{van-den-Heuvel+1972}
-{van den Heuvel}, E.~P.~J. \& {Heise}, J. 1972, \nat, 239, 67
-
-\bibitem[{{van den Heuvel} {et~al.}(2017){van den Heuvel}, {Portegies Zwart},
-  \& {de Mink}}]{vandenheuvel:17}
-{van den Heuvel}, E.~P.~J., {Portegies Zwart}, S.~F., \& {de Mink}, S.~E. 2017,
-  \mnras, 471, 4256
-
-\bibitem[{{van Oijen}(1989)}]{vanoijen:89}
-{van Oijen}, J.~G.~J. 1989, \aap, 217, 115
-
-\bibitem[{{van Rensbergen} {et~al.}(1996){van Rensbergen}, {Vanbeveren}, \& {De
-  Loore}}]{vanrensbergen:96}
-{van Rensbergen}, W., {Vanbeveren}, D., \& {De Loore}, C. 1996, \aap, 305, 825
-
-\bibitem[{{Vanbeveren}(1982)}]{vanbeveren:82}
-{Vanbeveren}, D. 1982, \aap, 113, 205
-
-\bibitem[{{Verbunt} \& {Cator}(2017)}]{verbunt:17}
-{Verbunt}, F. \& {Cator}, E. 2017, Journal of Astrophysics and Astronomy, 38
-
-\bibitem[{{Verbunt} {et~al.}(2017){Verbunt}, {Igoshev}, \&
-  {Cator}}]{verbunt:17b}
-{Verbunt}, F., {Igoshev}, A., \& {Cator}, E. 2017, \aap, 608, A57
-
-\bibitem[{{Vink} {et~al.}(2000){Vink}, {de Koter}, \& {Lamers}}]{vink:00}
-{Vink}, J.~S., {de Koter}, A., \& {Lamers}, H.~J.~G.~L.~M. 2000, \aap, 362, 295
-
-\bibitem[{{Vink} {et~al.}(2001){Vink}, {de Koter}, \& {Lamers}}]{vink:01}
-{Vink}, J.~S., {de Koter}, A., \& {Lamers}, H.~J.~G.~L.~M. 2001, \aap, 369, 574
-
-\bibitem[{{Webbink}(1984)}]{webbink:84}
-{Webbink}, R.~F. 1984, \apj, 277, 355
-
-\bibitem[{{Webster} \& {Murdin}(1972)}]{Webster+1972}
-{Webster}, B.~L. \& {Murdin}, P. 1972, \nat, 235, 37
-
-\bibitem[{{Wheeler} {et~al.}(1975){Wheeler}, {Lecar}, \& {McKee}}]{wheeler:75}
-{Wheeler}, J.~C., {Lecar}, M., \& {McKee}, C.~F. 1975, \apj, 200, 145
-
-\bibitem[{{Wijers} {et~al.}(1992){Wijers}, {van Paradijs}, \& {van den
-  Heuvel}}]{wijers:92}
-{Wijers}, R.~A.~M.~J., {van Paradijs}, J., \& {van den Heuvel}, E.~P.~J. 1992,
-  \aap, 261, 145
-
-\bibitem[{{Wongwathanarat} {et~al.}(2013){Wongwathanarat}, {Janka}, \&
-  {M{\"u}ller}}]{wongwathanarat:13}
-{Wongwathanarat}, A., {Janka}, H.-T., \& {M{\"u}ller}, E. 2013, \aap, 552, A126
-
-\bibitem[{{Wyrzykowski} {et~al.}(2016){Wyrzykowski}, {Kostrzewa-Rutkowska},
-  {Skowron}, {Rybicki}, {Mr{\'o}z}, {Koz{\l}owski}, {Udalski}, {Szyma{\'n}ski},
-  {Pietrzy{\'n}ski}, {Soszy{\'n}ski}, {Ulaczyk}, {Pietrukowicz}, {Poleski},
-  {Pawlak}, {I{\l}kiewicz}, \& {Rattenbury}}]{wyrzykowski:16}
-{Wyrzykowski}, {\L}., {Kostrzewa-Rutkowska}, Z., {Skowron}, J., {et~al.} 2016,
-  \mnras, 458, 3012
-
-\bibitem[{{Wysocki} {et~al.}(2018){Wysocki}, {Gerosa}, {O'Shaughnessy},
-  {Belczynski}, {Gladysz}, {Berti}, {Kesden}, \& {Holz}}]{wysocki:18}
-{Wysocki}, D., {Gerosa}, D., {O'Shaughnessy}, R., {et~al.} 2018, \prd, 97,
-  043014
-
-\bibitem[{{Yoon} {et~al.}(2017){Yoon}, {Dessart}, \& {Clocchiatti}}]{yoon:17}
-{Yoon}, S.-C., {Dessart}, L., \& {Clocchiatti}, A. 2017, \apj, 840, 10
-
-\bibitem[{{Zahn}(1977)}]{zahn:77}
-{Zahn}, J.-P. 1977, \aap, 57, 383
-
-\bibitem[{{Zapartas} {et~al.}(2017{\natexlab{a}}){Zapartas}, {de Mink},
-  {Izzard}, {Yoon}, {Badenes}, {G{\"o}tberg}, {de Koter}, {Neijssel}, {Renzo},
-  {Schootemeijer}, \& {Shrotriya}}]{zapartas:17}
-{Zapartas}, E., {de Mink}, S.~E., {Izzard}, R.~G., {et~al.} 2017{\natexlab{a}},
-  \aap, 601, A29
-
-\bibitem[{{Zapartas} {et~al.}(2017{\natexlab{b}}){Zapartas}, {de Mink}, {Van
-  Dyk}, {Fox}, {Smith}, {Bostroem}, {de Koter}, {Filippenko}, {Izzard},
-  {Kelly}, {Neijssel}, {Renzo}, \& {Ryder}}]{zapartas:17b}
-{Zapartas}, E., {de Mink}, S.~E., {Van Dyk}, S.~D., {et~al.}
-  2017{\natexlab{b}}, \apj, 842, 125
-
-\bibitem[{{Zwicky}(1957)}]{zwicky:57}
-{Zwicky}, F. 1957, \zap, 44, 64
-
-\end{thebibliography}
-
-\appendix
-
-\section{Comparison to selected previous population studies}
-\label{sec:literature}
-
-Various earlier studies have simulated the production of binaries
-containing a compact object and/or the producing runaway stars. We briefly discuss a selection that focuses on estimates for unbound companions and comment briefly comment on how our findings compare.
-
-\cite{dedonder:97} presented one of the first extensive binary evolution studies discussing predictions for the O-type runaways as well as the systems that remain bound.
-They consider a variety of assumptions for the initial distributions and the uncertain physical parameters concerning the efficiency of mass transfer, angular momentum loss and the common envelope ejection.
-Generally we find that our results agree well.
-They find that 16--23\% of systems remain bound after the SN explosion of the primary, which is consistent with our findings $1- \mathcal{D} = 14^{+22}_{-10}\%$.
-They find that between 6\% and 27\% of the O-type are unbound former companions that are now single, which is also agreement with our walkaway fraction (one but last column in their table 1).
-
-The find that  2--7\% of the O-type stars have velocities large than $30\,\kms$ (final column in their Table 1), which is slightly larger but still consistent with what we find.
-We expect that this difference may be in part the result of a difference the treatment of mass transfer and angular momentum loss, although the combined effect of further differences in our assumptions will also contribute.
-Their default assumption is that a fixed fraction $\beta_\mathrm{RLOF} = 0.5$ of the material transferred during Roche-lobe overflow leaves the system through the outer L2 Lagrangian point forming a ring around the binary system, which leads to larger angular momentum loss, shrinking the orbit further.
-This is somewhat similar to our model variation where we assumed $\gamma_{\rm RLOF} = \gamma_{\rm disk} $ although our results cannot be compared one to one, since we adopt a different, physically motivate, assumption for the mass transfer efficiency.
-In our simulations we find that this assumption is not increasing the number of runaway stars, as the fraction of system that merges increases too.
-
-\cite{eldridge:11} perform population synthesis simulations with a
-detailed stellar evolutionary evolutionary code to investigate O and
-early B
-type runaway stars.
-Their primary aim is the predict the spatial distribution of different types of CCSN and gamma-ray bursts.
-They find a disruption fraction, $\mathcal{D}=80\%$, in good agreement with our results, $\mathcal{D} = 86^{+10}_{-22}\%$.
-In their work they use the term runaway to refer to all unbound companions with velocities larger than $5\kms$, which encloses the large majority of what we refer to as walkaway stars, but they also quote estimates for stars faster than $30\kms$ which corresponds to our definition of runaway stars.
-They predict a walkaway (runaway) fraction of $2.2\%$ ($0.5\%$) for O type stars
-%, and $7.1\%$ ($2.2\%$) for B-type stars, respectively 
-in their simulations for $Z=0.02$.
-For comparison, for stars more massive than $15\,\Msun$, which roughly corresponds to O type stars,  we estimate a walkaway (runaway) fraction of  $10^{+4.7}_{-8.5}$\% ($0.5^{+1.0}_{-0.3}$\%).
-Our runaway fraction agrees very well.
-We find a somewhat larger walkaway fraction, but we consider this a fairly good agreement, given the uncertainties and differences in definitions that we have adopted.
-Also the distance they estimate that ejected stars can reach is in good agreement with our predictions and they also note that the majority unbound companions accrete mass from their companions before the first CC.
-Our predictions for the velocities of bound systems are slightly lower compared to \cite{eldridge:11}: the vast majority of bound post-CC systems is slower than $30\,\mathrm{km\ s^{-1}}$, and a high-velocity tail barely extending beyond about $100\,\mathrm{km\  s^{-1}}$.
-
-%The authors do not comment on the We note that they do not discuss the discrepancy between the low predictions for the fraction of O star runaways and the fractions typically quoted in observational studies, that $\sim10\%$ of O-type stars are runaways resulting from the binary ejection scenario \citep[][Sana \emph{et al.}, in preparation]{blaauw:61,gies:87,stone:91}.
-We return to this in our discussion in \Secref{sec:howtofind}.
-
-Very recently, \cite{boubert:18} published a study  investigating  the hypothesis that Be stars are products of mass transfer in binary systems \citep[e.g.][]{pols:91, demink:13}.
-They % precise astrometry of the Tycho-Gaia Astrometric Solution (TGAS) from the first Gaia Data Release and
-compare the  kinematics of a flux-limited sample of Galactic Be stars
-with binary population synthesis simulations. These simulations were
-obtained with a different version of the  binary evolutionary code that we use.
-Generally our findings are in agreement, despite the minor differences in the model assumptions.
-They also find a large fraction of unbound companions, many of which are rapidly rotating and moving at velocities slower than $30\,\kms$ (e.g., their Fig.
-6 and 7b), in good agreement with our findings.
-They also find that the natal kick distribution does not greatly affect the resulting velocity distribution and/or the runaway fraction and that the maximum distance traveled by ejected Be stars is likely to be smaller than the vertical scale height of the thin disk, in agreement with our findings.
-
-We find some disagreement in the provided explanations of the
-theoretical results.  For example, the authors state in their
-section~3.2 that that mass transfer shrinks the orbit and accelerates
-the secondary star. While this the case initially, upon the onset of
-mass transfer, we find that the orbit generally widens after the
-reversal of the mass ratio. The widening and the inversion of the mass
-ratio both slow down the orbital velocity as we verified both
-analytically \Secref{sec:analytics} and by detailed inspection of
-representative example systems, e.g. \Secref{sec:example}.  In
-section~3.3 they state that  ``whether a binary is disrupted by a
-supernova is principally determined by whether the primary loses more
-than half its mass \citep{blaauw:61}, and the kick on the compact
-object is only a second order effect''.  We find instead that mass
-loss during the explosion is rarely responsible for unbinding of a
-binary system that can produce a runaway star. The amount of mass lost
-needs to exceed half of the total mass of the system (and not half of
-the primary star), which is rarely achieved in our simulations. This
-is because the  CC progenitor loses most of its mass during the
-preceding mass transfer phase. We find that the Blaauw kick due to
-rapid mass loss is only important of initially very wide binaries, in which the two stars have not exchanged mass prior to CC.
-This difference may be in part due to the differences in the assumptions for the range of initial orbital periods.
-The authors consider systems with initial orbital periods up to  $10^{10}$ days in their simulations, which means that the majority of their progenitors should effectively evolve as single stars.
-We consider instead systems up to $10^{5.5}$ days, which is more appropriate for the more massive progenitors that we are interested in.
-
-Many further studies investigated the populations of binaries in the
-context of the formation of X-ray binaries and binary neutron stars
-and black holes. Providing a complete overview and detailed
-comparison is beyond our present scope. We discuss below a comparison
-to a limited set of studies.
-
-\cite{brandt:95} focus on binary systems remaining bound at the first CC, with the aim of understanding the
-effects of natal kicks on X-ray binaries.  They estimate a disruption
-fraction in their calculations of $\mathcal{D}\simeq73-81\%$
-(depending on the companion mass), in good agreement with our
-results.
-
-\cite{kalogera:96} presented a similar analytic study of the effects of natal kicks on the systemic velocities of X-ray binaries.
-Our population synthesis results agree in predicting systemic velocities of bound post-CC systems generally lower than the pre-explosion orbital velocity,
-except with $\sigma_\mathrm{kick} \gg v_\mathrm{orb}^\mathrm{pre-CC}$ (corresponding to large $\xi$ in the notation used by \citealt{kalogera:96}).
-
- \cite{fryer:98} investigated the impact of NS natal kicks on the
- formation of NS X-ray binaries and NS-NS binaries, but also present
- results for the unbound companions. Assuming a bimodal kick
- distribution, they find that most O/B-type ejected companion move
- slower than about $50\kms$, consistent with our findings.
-
- \cite{dray:05} focus on high-mass runaways, which become Wolf-Rayet (WR) stars during or before their post-disruption evolution.
-They argue in favor of significant BH kick amplitudes to explain the rarity of BH-WR binaries and the observed velocity distribution of WR runaways.
-We reach similar conclusions based on our simulations for different assumptions for the BH kicks.
-The systemic velocities we find for bound systems are also in reasonable agreement with those found by \cite{dray:05}
-
-Recently, \cite{tauris:17} presented a detailed study of the evolutionary processes leading to the formation of NS NS binary systems.
-They also find that the majority of binaries hosting a NS after the first CC have systemic velocities smaller than $\sim 30\,\mathrm{km\ s^{-1}}$, in good agreement with the results shown in the left panel of \Figref{fig:bound_v_dist}.
-
-\section{Output files}
-\label{app:data}
-
-The outcome of our population synthesis calculations will be made
-available upon publication. For early inquiries, please contact the
-lead author.
-%We provide the output of our population synthesis simulations at
-% \todo{where?}.
-Each file corresponds to one parameter variation (see
-\Secref{sec:methods} and \Secref{sec:param_variation}), and logs the
-following information for each binary system where a star goes
-CC\footnote{We stop our computations at the first CC event.}.
-\begin{itemize}
- \item primary ZAMS, pre-CC, and post-CC (corresponding to the NS or
-   BH mass if the primary is the star collapsing) masses in $M_\odot$ units:
-   \texttt{M1zams}, \texttt{M1preCC}, \texttt{M1postCC};
- \item secondary ZAMS, pre-CC, and post-CC masses in $M_\odot$ units:
-   \texttt{M2zams}, \texttt{M2preCC}, \texttt{M2postCC};
- \item fallback fraction $f_b$ for each star (set to zero for the star
-   that is not collapsing):
-   \texttt{fb1}, \texttt{fb2}
- \item evolutionary stage before and after CC (stellar types listed
-   according to \citealt{hurley:00}):
-   \texttt{type1preCC}, \texttt{type1postCC},
-   \texttt{type2preCC}, \texttt{type2postCC}
- \item post-CC velocities of the 2 stars in the original frame, in
-   $\mathrm{km\ s^{-1}}$: \texttt{v1postCC}, \texttt{v2postCC};
- \item pre-CC and post-CC eccentricity (the latter is -1 for mergers
-   and disrupted systems): \texttt{e\_preCC}, \texttt{e\_postCC};
- \item pre-CC and post-CC separation in $R_\odot$ units (the latter
-   is set to 0 for mergers and disrupted systems):
-   \texttt{a\_preCC}, \texttt{a\_postCC};
- \item ZAMS, pre-CC, and post-CC periods in
-   days: \texttt{Pzams}, \texttt{PpreCC}, \texttt{PpostCC};
- \item kick amplitude in $\mathrm{km\ s^{-1}}$: \texttt{v\_kick};
- \item kick direction, with $\theta$ angle between the collapsing star
-   orbital velocity and the kick (see also \citetalias{tauris:98} for notation):
-   \texttt{theta}, \texttt{phi};
- \item systemic velocity in $\mathrm{km\ s^{-1}}$: \texttt{v\_sys};
- \item age of the system at the time of CC in Myr: \texttt{t\_explosion};
- \item time left in the current evolutionary stage in Myr:
-   \texttt{t\_remaining};
- \item system probability (see below): \texttt{Prob}.
-\end{itemize}
-
-The system probability corresponds to the hyper-volume of the initial
-parameter space ($M_1^\mathrm{ZAMS}$, $q^\mathrm{ZAMS}$,
-$P^\mathrm{ZAMS}$) represented by each binary system in our model
-grid. In other words, the probability of each system is the
-statistical weight of the system seen as sampling point for the initial
-distributions. To construct distributions of the output quantities (e.g.~those presented in
-\Figref{fig:v_dist}-\ref{fig:bound_v_dist}), the properties of each
-system should be weigh with the corresponding probability. Similarly,
-the mean value of a quantity $\langle x \rangle$ (e.g.~$\langle v \rangle$ in
-\Tabref{tab:parameters}) should be calculated using:
-
-\begin{equation}
-  \label{eq:mean}
-  \langle x \rangle = \frac{\int x P(x)\,dx}{\int P(x)\,dx} \equiv
-  \frac{\sum_i^{}  x_i P_i}{\sum_i^{} P_i} \ \ ,
-\end{equation}
-where $P$ is the probability, and the index $i$ runs over all the
-binary systems in a population.
-
-To calculate the average distance traveled by stars
-ejected by the binary disruption, we use:
-
-\begin{equation}
-  \begin{aligned}
-    \label{eq:meanD}
-\langle L \rangle \equiv \langle v\times \Delta t\rangle = \\
-  \frac{\sum_i \texttt{v2postCC}\times [\texttt{t\_remaining}+0.1\times\tau_\mathrm{MS}(M_2)]
-    \times P_i}{\sum_i P_i} \ \ ,
-  \end{aligned}
-\end{equation}
-
-% \begin{equation}
-%   \label{eq:meanD}
-%   \langle L \rangle \equiv \langle v\times \Delta t\rangle = 
-%   \frac{\sum_i \texttt{v2postCC}\times [\texttt{t\_remaining}+0.1\times\tau_\mathrm{MS}(M_2)]
-%     \times P_i}{\sum_i P_i} \ \ ,
-% \end{equation}
-
-where the second term in squared brackets accounts, albeit in a
-simplified way, for the helium core burning duration of the
-rejuvenated star.
-
-\section{Observable velocity distribution}
-
-Figure~\ref{fig:v_dist_obs} shows the velocity distribution of MS
-stars ejected by the disruption of binaries that can
-be directly compared to observations provided that (i) the
-contribution of dynamical ejection can be separated in the observed
-sample and (ii) the effects of the galactic potential can be
-neglected, i.e.\ effectively each ejected star moves in a straight
-line at constant velocity for the remaining duration of its MS. This is the same information of
-\Figref{fig:v_dist}, but each bin is populated considering also the
-remaining MS lifetime of the ejected star
-($\tau_\mathrm{MS}$).
-
-\begin{figure*}[htbp]
-  \includegraphics[width=\textwidth]{figures/v2_dist_duration}
-  \caption{Velocity distribution of ejected stars, including the
-    finite MS lifetime to populate the bins (see also
-    \Figref{fig:v_dist}). }
-  \label{fig:v_dist_obs}
-\end{figure*}
-
-\begin{figure*}[bp]
-  \includegraphics[width=\textwidth]{figures/v2_dist_log}
-  \caption{Same as \Figref{fig:v_dist}, but using a logarithmic scale
-    for the velocity. The use of a logarithmic scale allows for the
-    display of a wider range of velocities. A minor peak in the grey
-    histogram can be seen between $100\lesssim
-    v_\mathrm{dis}/\mathrm{km \ s^{-1}} \lesssim
-    400$, but is absent in the histograms for massive ejected
-    stars. Such high ejection velocities are reached through a common
-    envelope evolution without accreting mass.}
-  \label{fig:v_dist_log}
-\end{figure*}
-
-\section{Pre-collapse distributions}
-
-We present in this appendix the pre-CC distribution in separation,
-mass of the collapsing star, and mass of the companions for all the
-binaries with a MS companion to the collapsing star in our fiducial
-simulation. Similar distributions can be derived for all our parameter
-variations from the data files available that will be made
-available.%\todo{where}.
-These distributions can inform studies of the interaction of the
-SN shock with the companion star
-\citep[e.g.][]{wheeler:75,liu:15,rimoldi:16, hirai:18}.
-
-We show in \Figref{fig:preCC_sep_dist} the pre-CC separation distribution. The colors indicate indicate the minimum mass of the MS
-companion (i.e., not of the collapsing star). Roughly speaking, the
-two peaks shown in \Figref{fig:preCC_sep_dist} correspond to the
-orbital widening due to conservative (case A and early case B) and
-non-conservative (late case B and case C) mass transfer phase. For all
-pre-CC separations shorter than $10^3\,R_\odot$, the dashed and solid
-histograms coincide, indicating that all these binaries have
-experienced a direct interaction previously during the evolution.
-
-\begin{figure}[htbp]
-  \includegraphics[width=0.5\textwidth]{figures/Sep_dist}
-  \caption{Pre-CC separation distribution for binaries with a
-    collapsing star and a main sequence companion. Colors indicate the
-    pre-CC mass of the MS companion according to the legend. Dashed
-    histograms indicate post-interaction (RLOF or common envelope)
-    binaries. The top panel shows the corresponding cumulative
-    distributions.}
-  \label{fig:preCC_sep_dist}
-\end{figure}
-
-Figure \ref{fig:preCC_mass_dist} shows the mass distribution for the
-exploding star and the MS companion ($M_\mathrm{CC}$, and $M_2$,
-respectively), at the pre-CC stage. The combination of the
-distributions shown in \Figref{fig:preCC_mass_dist} and
-\Figref{fig:preCC_sep_dist}, together with the effects of the natal
-kick distribution results in the ejection velocities in
-\Figref{fig:v_dist} which is our main result.
-
-\begin{figure}[htbp]
-  \includegraphics[width=0.5\textwidth]{figures/masses_dist}
-  \caption{Pre-CC mass distribution for the exploding star
-    ($M_\mathrm{CC}$) and the MS companion ($M_2$) binaries with a
-    collapsing star and a main sequence companion. Dashed
-    histograms indicate post-interaction (RLOF or common envelope)
-    binaries. The top panel shows the corresponding cumulative
-    distributions.}
-  \label{fig:preCC_mass_dist}
-\end{figure}
+\bibliography{bibliography/vfts682}
 

### 2018-04-30 VFTS682.tex
started writing the paper

@@
 \renewcommand{\labelitemii}{$\bullet$}
-\newcommand{\kms}{{\mathrm{km\ s^{-1}}}}
+\newcommand{\kms}{{\,\mathrm{km\ s^{-1}}}}
 \newcommand{\Msun}{{\mathrm{M}_\odot}}
@@
 
-How do stars form is one longstanding question in astrophysics
+How do stars form is one longstanding question in astrophysics \cite{lada:03}
 \todo{ref?}. Answering it for massive stars is particularly difficult,
@@
 remain enshrouded in their parent cloud during the whole formation
-process. Moreover, observations of young massive stars reveal a
-complicated multipliticy structure which requires explanation.
-
-Understanding massive star formation, possibly as a function
-of metallicity, is a key question given the present and upcoming
-transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{} which will reveal transients associated to their
+process \todo{ref}. Moreover, observations of young massive stars reveal a
+complicated multipliticy structure which requires
+explanation \citep[][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. Understanding massive star formation, possibly as a
+function of metallicity, is a key question given the present and upcoming
+transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{} which
+will reveal transients associated to massive stars
 evolution and death.
@@
 these hypothesis using one particular star, VFTS682. This star is a
-very massive ($\sim150\,M_\odot$) WNh5 star in the 30 Doradus region
+very massive ($\sim$$150\,M_\odot$) WNh5 star in the 30 Doradus region
 of the Large Magellanic Cloud, and it is presently observed at a
@@
 its present day apparent isolation, \cite{bestenlehner:11} proposed it
-might be a candidate for isolated star formation, if it has not been
-ejected from R136 in the past.
+might be a candidate for isolated star formation, or a runaway ejected
+from R136 in the past.
+
+Massive stars can in principle be ejected from R136 as a consequence of dynamical
+interactions \citep[][]{poveda:67,leonard:91, evans:10, fujii:11,
+  allison:12, oh:16}, or by the disruption of a binary by the first
+core-collapse supernova
+\citep[][]{zwicky:57,blaauw:61,dedonder:97,eldridge:11,renzo:18}. However,
+R136 has an estimated age of $\lessim2$\,Myr \todo{ref}, which is
+shorter than the shortest stellar lifetime of about $\sim$3\,Myr
+\citep[e.g.,][]{zapartas:17}, so one would not expect the binary
+disruption scenario to be relevant for this cluster.
 
@@
 this star, and therefore isolated star formation is \emph{not}
-required to explain it. However, we find a discrepancy between the
-apparent age of VFTS682, it's kinematic age, and the age estimates for
+required to explain it. However, we find a mild discrepancy between the
+apparent age of VFTS682, its kinematic age, and the age estimates for
 R136.
 
-In \Secref{sec:sample} we describe how we select stars in R136 and in
+In \Secref{sec:sample}, we describe how we select stars in R136 and in
 the surroundings of VFTS682 itself to remove the mean motion of the
@@
 proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DE}$,
-respectively). For the radial velocity of VFTS16 and of the 30 Doradus
+respectively). For the radial velocity of VFTS682 and of the 30 Doradus
 region as a whole, we instead use the VFTS data
@@
 depend on the detailed value of the radial velocity.\todo{improve
-  phrasing} The values of all these quantities that we use are listed
-in \Tabref{tab:vfts682}.
+  phrasing} \Tabref{tab:vfts682} lists the values adopted throughout
+this work for each of these quantities.
 
@@
   \centering
+    \caption{Astrometric parameters for VFTS682. The peculiar radial
+    velocity $\delta v_\mathrm{rad}$ is obtained as the difference
+    between the average radial velocity of the 30 Doradus region
+    ($270\pm10\kms$) minus the radial velocity measured from the HeII $\lambda4686$
+    line for VFTS682 ($315\pm15\kms$).}
+
   \begin{tabular}[htbp]{l|c|c}
@@
   \end{tabular}
-  \caption{Astrometric parameters for VFTS682}
   \label{tab:vfts682}
@@
 
+The 
 \section{The kinematics of VFTS682}
+\label{sec:results}
 
@@
   \includegraphics[width=\textwidth]{./figures/main_plot}  
-  \caption{position and projected relative velocity to R136}
+  \caption{position and projected relative velocity to
+    R136. \todo{load pretty picture on background, check scale, clean
+      foreground stars with parallax, add cone of uncertainty}}
   \label{fig:main}
 \end{figure*}
+
+\section{Discussion}
+\label{sec:discussion}
 

### 2018-05-08 VFTS682.tex
started writing the paper
polishing, hardcore stuff still to be done

@@
     Amsterdam, 1098 XH Amsterdam, The Netherlands}
-  \and{School of Astronomy \& Space Science, University of the Chinese
-    Academy of Sciences, Beijing 100012, China}
-  \and{National Astronomical Observatories, Chinese Academy of
-    Sciences, Beijing 100012, China}
+  % \and{School of Astronomy \& Space Science, University of the Chinese
+  %   Academy of Sciences, Beijing 100012, China}
+  % \and{National Astronomical Observatories, Chinese Academy of
+  %   Sciences, Beijing 100012, China}
 }
@@
 
-\keywords{stars: kinematics, stars:individual:VFTS682}
+\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS682}
 \maketitle{}
@@
 
-How do stars form is one longstanding question in astrophysics \cite{lada:03}
-\todo{ref?}. Answering it for massive stars is particularly difficult,
-since these are intrinsically rare
+How do stars form is one longstanding question in astrophysics
+\cite{lada:03, zinnecker:07}. It is particularly difficult for massive stars, because these are intrinsically rare
 \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
-remain enshrouded in their parent cloud during the whole formation
-process \todo{ref}. Moreover, observations of young massive stars reveal a
+remain enshrouded in their parent cloud during the formation
+process. Moreover, observations of young massive stars reveal a
 complicated multipliticy structure which requires
@@
 
-\todo{short summary of formation scenarios goes here}.
-
 The second data release (DR2) from the Gaia satellite
-\cite[][]{brown:18,} allows us to test
-these hypothesis using one particular star, VFTS682. This star is a
-very massive ($\sim$$150\,M_\odot$) WNh5 star in the 30 Doradus region
-of the Large Magellanic Cloud, and it is presently observed at a
+\cite[][]{gaia:16,brown:18} allows us to test
+these hypothesis using one particular star, VFTS682. 
+This star is a
+very massive \citep[$M_\mathrm{ZAMS}\simeq150\,M_\odot$,][]{bestenlehner:11,schneider:18} WNh5 star in the 30 Doradus region
+of the Large Magellanic Cloud (LMC), and it is presently observed at a
 projected distance of $\sim$$29$\,pc from the nearest cluster of
@@
 its present day apparent isolation, \cite{bestenlehner:11} proposed it
-might be a candidate for isolated star formation, or a runaway ejected
-from R136 in the past.
+might be a candidate for isolated star formation, or a ``slow runaway'' ejected
+from R136 in the past. This second option is also supported by the
+N-body simulations of \cite{fujii:11, banerjee:12}.   Many other
+very massive stars are present in the surroundings of R136, and a more
+detailed analysis on the larger sample is desirable.
 
@@
 \citep[][]{zwicky:57,blaauw:61,dedonder:97,eldridge:11,renzo:18}. However,
-R136 has an estimated age of $\lessim2$\,Myr \todo{ref}, which is
-shorter than the shortest stellar lifetime of about $\sim$3\,Myr
-\citep[e.g.,][]{zapartas:17}, so one would not expect the binary
+R136 has an estimated age of $\lesssim2$\,Myr \citep[][]{sabbi:12}, which is
+shorter than the shortest stellar lifetime
+\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, so one would not expect the binary
 disruption scenario to be relevant for this cluster.
@@
 
-In \Secref{sec:sample}, we describe how we select stars in R136 and in
-the surroundings of VFTS682 itself to remove the mean motion of the
-region. \Secref{sec:results} presents our main findings. We conclude by
-discussing the implications for theories of
+In \Secref{sec:sample}, we describe the data selection and validation
+process for VFTS 682 and for stars in the surrounding used to
+calculate and remove the mean motion of the
+region. \Secref{sec:results} presents our main findings. We conclude
+by discussing the implications for theories of
 star formation, N-body interactions, and binary evolution in
@@
 \label{sec:sample}
+
 VFTS682 is labeled in the Gaia DR2
 catalog\footnote{\url{https://gea.esac.esa.int/archive/}} with the
-source id 4657685637907503744. For this star, we retrieve from the
-Gaia catalog its position in right ascension (RA) and declination (DE)
-in the IGCS frame\cite[][]{brown:18}\todo{some 2016 Gaia paper?}, its
+source id 4657685637907503744. The star has a
+\texttt{visibility\_period} = 17, which counts how many observations have
+been used to reconstruct its astrometric solution \citep[][]{lindengren:18}. Its G-band
+magnitude is 15.65, cf. the V-band magnitude of 16.08
+\citep[][]{evans:11, bestenlehner:11}, and the reported
+\texttt{astrometric\_noise} = 0. These values suggest that the Gaia
+data for VFTS682 are trustworthy. However, the effective temperature
+reported is one order of magnitude lower than what found by
+\cite{bestenlehner:11}, and the best fit parallax of this star is
+negative. We do not use the effective temperature of the star anywhere
+in this study, and we attribute the unphysical value of the parallax
+to the large distance to the LMC. Our main findings do not rely on the
+parallax value reported in the Gaia catalog.
+
+We retrieve for VFTS682 the position in right ascension (RA) and declination (DE)
+in the IGCS frame \cite[][]{brown:18}, its
 proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DE}$,
@@
 region as a whole, we instead use the VFTS data
-as quoted in \cite{bestenlehner:11}. Our main conclusions do not
-depend on the detailed value of the radial velocity.\todo{improve
-  phrasing} \Tabref{tab:vfts682} lists the values adopted throughout
+as quoted in \cite{bestenlehner:11}. \Tabref{tab:vfts682} lists the values adopted throughout
 this work for each of these quantities.
@@
 surrounding region is defined by all the stars within 10 arcminutes from
-VFTS682. This includes the subset of stars which we consider to be
+VFTS682 \todo{without a pm and with G mag<17 for VFTS completeness and
+to focus on massive stars only}. This includes the subset of stars which we consider to be
 part of R136. We include all the stars for which a value of the proper
@@
 within 25 arcseconds from R136a, one of the most massive members of
-the cluster itself \citep[][]{crowther:10}.
+the cluster itself \citep[][]{crowther:10}. \todo{improve definition
+  of stars from R136 -- check with Danny}
 
@@
 
-The 
+The data retrieved, and the ipython notebook used for the analysis
+presented here will be made available at \todo{probably git repo on bitbucket?}. 
 \section{The kinematics of VFTS682}
@@
 
-\section{Discussion}
+\section{Summary and Discussion}
 \label{sec:discussion}
+
+\begin{itemize}
+\item VFTS682 is a bona fide runaway with $v\sim60\,\mathrm{km\
+    s^{-1}}$ thrown out from R136. Both its speed and the age of the
+  cluster are consitent with a dynamical ejection.
+  \item VFTS682 comes from R136 as was expected by
+  \cite{bestenlehner:11, fujii:11, banerjee:12}, so it does not
+  require isolated SFH to be explained
+\item apparent age tension (connect to VFTS16 as well).
+\item is R136 a single young cluster or a merger
+\item estimate the influence of the gravitational potential of R136,
+  what is its total mass and relaxation time?
+\end{itemize}
+
+Random notes: vsini<200 form \cite{schneider:18}, age $1\pm 0.2$\,Myr
+from \cite{schneider:18}
 

### 2018-05-08 VFTS682.tex
downloaded new data, a bit more trustworthy

@@
 process for VFTS 682 and for stars in the surrounding used to
-calculate and remove the mean motion of the
-region. \Secref{sec:results} presents our main findings. We conclude
+define the local reference frame. \Secref{sec:results} presents our main findings. We conclude
 by discussing the implications for theories of
@@
 \citep[][]{evans:11, bestenlehner:11}, and the reported
-\texttt{astrometric\_noise} = 0. These values suggest that the Gaia
+\texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
 data for VFTS682 are trustworthy. However, the effective temperature
@@
 ``surroundings'' of VFTS 682, and the ``R136 cluster''. The
-surrounding region is defined by all the stars within 10 arcminutes from
-VFTS682 \todo{without a pm and with G mag<17 for VFTS completeness and
-to focus on massive stars only}. This includes the subset of stars which we consider to be
-part of R136. We include all the stars for which a value of the proper
-motion components and the corresponding errors are available.
+surrounding region is defined by all the stars in a target of 10 arcminutes around
+VFTS682 fulfilling the following criteria: we require a G-band
+magnitude brighter than 17, to mimic the completeness threshold for
+the VFTS survey, \texttt{visibility\_period} >= 5, and
+\texttt{astrometric\_excess\_noise} <= 1. This selection yields
+roughly 17800 stars, including the subset of stars which we consider to be
+part of R136, discussed below. We further require  these stars to have finite proper
+motion components and relative errors, reducing further the sample to
+14856 stars. % A more strict sample selection for the surroundings can
+% be found in \cite{lennon:18}.
+
 The ``R136 cluster'' is effectively defined by taking all the stars
 within 25 arcseconds from R136a, one of the most massive members of
-the cluster itself \citep[][]{crowther:10}. \todo{improve definition
-  of stars from R136 -- check with Danny}
+the cluster itself \citep[][]{crowther:10}, requiring the same
+``quality'' criteria applied above. Because of the magnitude cut
+however, this would yield an incredibly small sample of 15 stars only
+% which makes no sense at all
+\todo{improve definition
+  of stars from R136, describe accordingly -- check with Danny}
 
@@
 presented here will be made available at \todo{probably git repo on bitbucket?}.
+
 \section{The kinematics of VFTS682}
@@
 
+% \begin{acknowledgements}
+%  I am grateful to S.~Torres who helped me p
+% \end{acknowledgements}
 \end{document}

### 2018-05-09 VFTS682.tex
minor

@@
 VFTS682 is labeled in the Gaia DR2
-catalog\footnote{\url{https://gea.esac.esa.int/archive/}} with the
+catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}} with the
 source id 4657685637907503744. The star has a
 \texttt{visibility\_period} = 17, which counts how many observations have
-been used to reconstruct its astrometric solution \citep[][]{lindengren:18}. Its G-band
+been used to reconstruct its astrometric solution
+\citep[][]{lindengren:18}. Its reported G-band
 magnitude is 15.65, cf. the V-band magnitude of 16.08
@@
 data for VFTS682 are trustworthy. However, the effective temperature
-reported is one order of magnitude lower than what found by
+reported in Gaia DR2 is one order of magnitude lower than what found by
 \cite{bestenlehner:11}, and the best fit parallax of this star is
@@
 to the large distance to the LMC. Our main findings do not rely on the
-parallax value reported in the Gaia catalog.
+parallax nor the effective temperature values reported in the Gaia DR2
+catalog.
+
+%% should I really mention the line of sight velocity? Not really
+%% using it...
 
@@
 \texttt{astrometric\_excess\_noise} <= 1. This selection yields
-roughly 17800 stars, including the subset of stars which we consider to be
+roughly 1330 stars, including the subset of stars which we consider to be
 part of R136, discussed below. We further require  these stars to have finite proper
 motion components and relative errors, reducing further the sample to
-14856 stars. % A more strict sample selection for the surroundings can
+1323 stars. % A more strict sample selection for the surroundings can
 % be found in \cite{lennon:18}.
@@
 the cluster itself \citep[][]{crowther:10}, requiring the same
-``quality'' criteria applied above. Because of the magnitude cut
-however, this would yield an incredibly small sample of 15 stars only
-% which makes no sense at all
+``quality'' criteria applied above. This selection yields 62 stars.
 \todo{improve definition
@@
 the 30 Doradus region as a whole, since the parallax for VFTS682
-listed in the Gaia DR2 catalog is negative.
+listed in the Gaia DR2 catalog is negative. %% this is what makes the
+                                %% LOS velocity useless...how to improve?
 
@@
   \caption{position and projected relative velocity to
-    R136. \todo{load pretty picture on background, check scale, clean
-      foreground stars with parallax, add cone of uncertainty}}
+    R136. \todo{orient properly, load  picture on background, add cone of uncertainty}}
   \label{fig:main}
@@
 
-Random notes: vsini<200 form \cite{schneider:18}, age $1\pm 0.2$\,Myr
+Random notes: $v\sin(i)<200\,\mathrm{km\ s^{-1}}$ form \cite{schneider:18}, age $1.0\pm 0.2$\,Myr
 from \cite{schneider:18}

### 2018-05-09 VFTS682.tex
./aa.cls

@@
 
+For each of the two local rest frames (``surroundings'' and ``R136
+cluster''), we compute the average transverse velocity as:
+
+\begin{equation}
+  \label{eq:mean}
+  \langle \mu_i\rangle = \frac{\sum_\mathrm{stars}\frac{1}{\Delta
+      \mu_i}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i}} \ \ , \
+  \ \Delta \langle \mu_i\rangle = \frac{\sqrt{N}}{\sum_\mathrm{stars}
+    \frac{1}{\Delta \mu_i}} \ \ ,
+\end{equation}
+where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
+on the proper motion component reported by Gaia, and the sums run over
+all the $N$ stars of the frame considered. We evaluate each proper motion
+component separately.
+
 The data retrieved, and the ipython notebook used for the analysis
@@
 \label{sec:results}
+
+\subsection{Is it a runaway star?}
+We first address the question of whether VFTS682 is a typical star in
+its surroundings from the kinematic point of view, or whether it is a
+runaway star with a significantly large peculiar velocity. The former is what should
+be expected if it formed where we observe it today, in relative
+isolation from other massive stars. We calculate the velocity relative
+to the general population in its surroundings, shown as blue dots in
+the large panel of\Figref{fig:main}, assuming a distance of
+$D=50$\,kpc.
+
+\todo{discuss numbers found, refer to figure}
+likely conclusion looking at the general region: it is a ``bona
+  fide'' runaway star with velocity components $\left(v_\mathrm{RA},
+  v_\mathrm{DE}, v_\mathrm{rad}\right) = \left(\todo{XX},\todo{YY},
+  \todo{ZZ}\right)$, corresponding to $v\equiv\sqrt{v_\mathrm{RA}^2+v_\mathrm{DE}^2+v_\mathrm{rad}^2}=\todo{AA}\,\kms$.
+
+\subsection{Does it come from the R136 cluster?}
+The inset plot in \Figref{fig:main} shows a zoom in of the region
+around R136 which includes VFTS682. To test the hypothesis that the
+star was indeed ejected from this region, we check the relative
+orientation of its proper motion. The error cone, obtained by
+propagating the errors on the proper motion components indicates that
+R136 is indeed the most likely origin, as expected by \cite{fujii:11,
+  banerjee:12}.
+
+\todo{kinematic age}
 

### 2018-05-09 VFTS682.tex
./aa.cls

@@
     Parameter & Value & Source\\ \hline\hline
-    RA \hfill[degree] &  84.73 $\pm$  0.036 & \multirow{4}{*}{Gaia DR2}\\
+    RA \hfill[degree] &  \phantom{-}84.73 $\pm$  0.03 & \multirow{4}{*}{Gaia DR2}\\
     DE \hfill [degree] & -69.07 $\pm$  0.05  & \\
-    $\mu_\mathrm{RA}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 1.84 $\pm$ 0.07 & \\
-    $\mu_\mathrm{DE}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.78 $\pm$ 0.08& \\
-    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & -45 $\pm$ 25 & \cite{bestenlehner:11}\\
+    $\mu_\mathrm{RA}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}1.84 $\pm$ 0.07 & \\
+    $\mu_\mathrm{DE}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}0.78 $\pm$ 0.08& \\
+    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{0}-45 $\pm$ 25 & \cite{bestenlehner:11}\\
     \hline

### 2018-05-09 VFTS682.tex
./aa.cls

@@
 surrounding region is defined by all the stars in a target of 10 arcminutes around
-VFTS682 fulfilling the following criteria: we require a G-band
-magnitude brighter than 17, to mimic the completeness threshold for
-the VFTS survey, \texttt{visibility\_period} >= 5, and
-\texttt{astrometric\_excess\_noise} <= 1. This selection yields
-roughly 1330 stars, including the subset of stars which we consider to be
-part of R136, discussed below. We further require  these stars to have finite proper
-motion components and relative errors, reducing further the sample to
-1323 stars. % A more strict sample selection for the surroundings can
+VFTS682 fulfilling the following criteria: we require \texttt{visibility\_period} >= 5,
+\texttt{astrometric\_excess\_noise} <= 1, the error on the proper
+motion components to be smaller than 0.07\,$\mathrm{mas\ yr^{-1}}$,
+and the proper motion components themselves to be smaller than
+2\,$\mathrm{mas\ yr^{-1}}$ in absolute value. At the distance to the
+LMC, $1\mathrm{mas\ yr^{-1}}\simeq250\,\kms$, so the cut on the values
+of the proper motions removes stars that would have projected
+tangential velocities in excess of $\sim$500\,$\kms$, which are most
+likely to be foreground stars \todo{clean phrasing here}.
+This selection yields 182 stars.% , including the subset of stars which we consider to be
+% part of R136, discussed below. We further require  these stars to have finite proper
+% motion components and relative errors, reducing further the sample to
+% 1323 stars.
+% A more strict sample selection for the surroundings can
 % be found in \cite{lennon:18}.

### 2018-05-09 VFTS682.tex
./aa.cls

--- 2018-05-09 VFTS682.tex
@@
 surrounding region is defined by all the stars in a target of 10 arcminutes around
-VFTS682 fulfilling the following criteria: we require a G-band
-magnitude brighter than 17, to mimic the completeness threshold for
-the VFTS survey, \texttt{visibility\_period} >= 5, and
-\texttt{astrometric\_excess\_noise} <= 1. This selection yields
-roughly 1330 stars, including the subset of stars which we consider to be
-part of R136, discussed below. We further require  these stars to have finite proper
-motion components and relative errors, reducing further the sample to
-1323 stars. % A more strict sample selection for the surroundings can
-% be found in \cite{lennon:18}.
+VFTS682 fulfilling the following criteria: we require \texttt{visibility\_period} >= 5,
+\texttt{astrometric\_excess\_noise} <= 1, the error on the proper
+motion components to be smaller than 0.1\,$\mathrm{mas\ yr^{-1}}$,
+and the proper motion components themselves to be smaller than
+2\,$\mathrm{mas\ yr^{-1}}$ in absolute value. At the distance to the
+LMC, $1\mathrm{mas\ yr^{-1}}\simeq250\,\kms$ \citep[e.g.,][]{lennon:18}, so the cut on the values
+of the proper motions removes stars that would have projected
+tangential velocities in excess of $\sim$500\,$\kms$, which are most
+likely to be foreground stars \todo{clean phrasing here}.
+This selection yields 437 stars.
 
@@
 likely conclusion looking at the general region: it is a ``bona
-  fide'' runaway star with velocity components $\left(v_\mathrm{RA},
-  v_\mathrm{DE}, v_\mathrm{rad}\right) = \left(\todo{XX},\todo{YY},
-  \todo{ZZ}\right)$, corresponding to $v\equiv\sqrt{v_\mathrm{RA}^2+v_\mathrm{DE}^2+v_\mathrm{rad}^2}=\todo{AA}\,\kms$.
+  fide'' runaway star with velocity components $\left(v_\mathrm{RA}, v_\mathrm{DE}, v_\mathrm{rad}\right)$ = $\left(\todo{XX},\todo{YY},\todo{ZZ}\right)$, corresponding to $v\equiv\sqrt{v_\mathrm{RA}^2+v_\mathrm{DE}^2+v_\mathrm{rad}^2}=\todo{AA}\,\kms$.
 

### 2018-05-09 VFTS682.tex
minor stuff

@@
 tangential velocities in excess of $\sim$500\,$\kms$, which are most
-likely to be foreground stars \todo{clean phrasing here}.
+likely to be foreground stars.
 This selection yields 437 stars.
@@
 to the general population in its surroundings, shown as blue dots in
-the large panel of\Figref{fig:main}, assuming a distance of
+the large panel of \Figref{fig:main}, assuming a distance of
 $D=50$\,kpc.
@@
 likely conclusion looking at the general region: it is a ``bona
-  fide'' runaway star with velocity components $\left(v_\mathrm{RA}, v_\mathrm{DE}, v_\mathrm{rad}\right)$ = $\left(\todo{XX},\todo{YY},\todo{ZZ}\right)$, corresponding to $v\equiv\sqrt{v_\mathrm{RA}^2+v_\mathrm{DE}^2+v_\mathrm{rad}^2}=\todo{AA}\,\kms$.
+  fide'' runaway star with velocity components $\left(v_\mathrm{RA},
+    v_\mathrm{DE}, v_\mathrm{rad}\right)$ =
+  $\left(\todo{XX},\todo{YY},\todo{ZZ}\right)$, corresponding to $v\equiv\sqrt{v_\mathrm{RA}^2+v_\mathrm{DE}^2+v_\mathrm{rad}^2}=\todo{AA}\,\kms$.
 

### 2018-05-09 VFTS682.tex
section on generic surroundings done, but error calculation is wrong

@@
 Throughout this study, we assume the same distance of $50$\,kpc to the star, and to
-the 30 Doradus region as a whole, since the parallax for VFTS682
-listed in the Gaia DR2 catalog is negative. %% this is what makes the
-                                %% LOS velocity useless...how to improve?
+the 30 Doradus region as a whole.% , since the parallax for VFTS682
+% listed in the Gaia DR2 catalog is negative. %% this is what makes the
+%                                 % LOS velocity useless...how to improve?
 
@@
 \subsection{Is it a runaway star?}
-We first address the question of whether VFTS682 is a typical star in
-its surroundings from the kinematic point of view, or whether it is a
-runaway star with a significantly large peculiar velocity. The former is what should
+
+We first address the question of whether VFTS682 is a typical star
+from the kinematic point of view, or whether it is a runaway star with
+a significantly large peculiar velocity compared to its surrounding population. The former is what should
 be expected if it formed where we observe it today, in relative
-isolation from other massive stars. We calculate the velocity relative
-to the general population in its surroundings, shown as blue dots in
-the large panel of \Figref{fig:main}, assuming a distance of
-$D=50$\,kpc.
+isolation from other massive stars.
 
-\todo{discuss numbers found, refer to figure}
-likely conclusion looking at the general region: it is a ``bona
-  fide'' runaway star with velocity components $\left(v_\mathrm{RA},
-    v_\mathrm{DE}, v_\mathrm{rad}\right)$ =
-  $\left(\todo{XX},\todo{YY},\todo{ZZ}\right)$, corresponding to $v\equiv\sqrt{v_\mathrm{RA}^2+v_\mathrm{DE}^2+v_\mathrm{rad}^2}=\todo{AA}\,\kms$.
+Using the 437 stars selected as the ``surrounding population'' in
+\Secref{sec:sample}, we find averaged proper motion components of
+$\mu_\mathrm{RA} = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
+$\mu_\mathrm{DE} = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
+proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
+components of proper motion of the star relative to the surrounding region
+$\mu^\mathrm{sur}_\mathrm{RA} = 0.14\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
+0.09\pm 0.08\,\mathrm{mas\ yr^{-1}}$. These can be converted in the
+the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=35\pm17\,\kms$,
+$v^\mathrm{sur}_\mathrm{DE}=22\pm19\,\kms$, assuming a distance of
+50\,kpc (we do not account for the uncertainty in the distance
+estimate when propagating errors). The radial velocity from
+\cite{bestenlehner:11} then gives the third component along
+the line of sight, allowing us to calculate the speed of the star:
+
+\begin{equation}
+  \label{eq:speed}
+  v = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
+    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 1
+  \, \kms \ .
+\end{equation}
+This value for the peculiar 3D velocity of VFTS682 make it a ``bona
+fide'' runaway star.
+
+\todo{errors are wrong}
 

### 2018-05-09 VFTS682.tex
section on generic surroundings done, fixed errors

@@
 function of metallicity, is a key question given the present and upcoming
-transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{} which
+transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{}\todo{ref} which
 will reveal transients associated to massive stars
@@
 to reconstruct the three-dimensional velocity of VFTS682, and test the
-null hypothesis that this star was ejected from R136. \todo{check the
-  following} Our results indicate that R136 is the likely origin of
-this star, and therefore isolated star formation is \emph{not}
-required to explain it. However, we find a mild discrepancy between the
-apparent age of VFTS682, its kinematic age, and the age estimates for
-R136.
-
-In \Secref{sec:sample}, we describe the data selection and validation
-process for VFTS 682 and for stars in the surrounding used to
-define the local reference frame. \Secref{sec:results} presents our main findings. We conclude
-by discussing the implications for theories of
+hypothesis that this star was ejected from R136. We discuss in
+\Secref{sec:sample} the data for VFTS682, and the selection of stars
+used to define a local reference frame. Our results
+indicate that R136 is a bona fide runaway star (\Secref{sec:runaway}), therefore isolated star formation is \emph{not}
+required to explain it. We also find that a dynamical ejection from
+R136 is compatible with the direction of its velocity
+vector. \todo{double check the following}
+However, we find a mild discrepancy between the apparent age of
+VFTS682, its kinematic age, and the age estimates for R136 (\Secref{sec:r136_origin}).  We conclude
+with a very brief discussion on the implications for theories of
 star formation, N-body interactions, and binary evolution in
@@
 \subsection{Is it a runaway star?}
-
+\label{sec:runaway}
 We first address the question of whether VFTS682 is a typical star
@@
 Using the 437 stars selected as the ``surrounding population'' in
-\Secref{sec:sample}, we find averaged proper motion components of
-$\mu_\mathrm{RA} = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
-$\mu_\mathrm{DE} = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
+\Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
+$\langle\mu_\mathrm{RA}\rangle = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
+$\langle\mu_\mathrm{DE}\rangle = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
 proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
 components of proper motion of the star relative to the surrounding region
-$\mu^\mathrm{sur}_\mathrm{RA} = 0.14\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
-0.09\pm 0.08\,\mathrm{mas\ yr^{-1}}$. These can be converted in the
+$\mu^\mathrm{sur}_\mathrm{RA} = 0.15\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
+0.09\pm 0.08\,\mathrm{mas\ yr^{-1}}$. %% note that the error budget is
+% dominated by VFTS682
+These can be converted in the
 the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=35\pm17\,\kms$,
@@
   v = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
-    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 1
+    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 22
   \, \kms \ .
 \end{equation}
-This value for the peculiar 3D velocity of VFTS682 make it a ``bona
-fide'' runaway star.
-
-\todo{errors are wrong}
+This value for the three-dimensional speed of VFTS682 with respect the
+surrounding stars make it a ``bona fide'' runaway star.
 
 \subsection{Does it come from the R136 cluster?}
+\label{sec:r136_origin}
+
 The inset plot in \Figref{fig:main} shows a zoom in of the region
 around R136 which includes VFTS682. To test the hypothesis that the
-star was indeed ejected from this region, we check the relative
-orientation of its proper motion. The error cone, obtained by
+star was indeed ejected from this region, we check the
+orientation of its proper motion relative to the group of stars
+selected to represent the R136 cluster in \Secref{sec:sample}. We
+further consider also the kinematic age of the star. The error cone, obtained by
 propagating the errors on the proper motion components indicates that

### 2018-05-09 VFTS682.tex
more data background

@@
 surrounding region is defined by all the stars in a target of 10 arcminutes around
-VFTS682 fulfilling the following criteria: we require \texttt{visibility\_period} >= 5,
-\texttt{astrometric\_excess\_noise} <= 1, the error on the proper
+VFTS682 fulfilling the following criteria: we require \texttt{visibility\_period} $\geq$ 5,
+\texttt{astrometric\_excess\_noise} < 1, the error on the proper
 motion components to be smaller than 0.1\,$\mathrm{mas\ yr^{-1}}$,

### 2018-05-09 VFTS682.tex
more data R136 and minor
./gitignore

@@
 tangential velocities in excess of $\sim$500\,$\kms$, which are most
-likely to be foreground stars.
-This selection yields 437 stars.
+likely to be foreground stars. We checked that the additional
+requirement of having parallaxes smaller than $1\,\mathrm{mas\
+  yr^{-1}}$ does not reduces further our sample. This selection yields 437 stars.
 
@@
 
-% \begin{acknowledgements}
-%  I am grateful to S.~Torres who helped me p
-% \end{acknowledgements}
+\begin{acknowledgements}
+  The background image of \Figref{fig:main} is based on observations made with ESO Telescopes at the La Silla Observatory under programme ID 076.C-0888, processed and released by the ESO VOS/ADP group.
+  We are grateful to S.~Torres for extremely useful discussions on the
+  Gaia DR2 dataset.
+\end{acknowledgements}
 \end{document}

### 2018-05-09 VFTS682.tex
./gitignore

@@
 The ``R136 cluster'' is effectively defined by taking all the stars
-within 25 arcseconds from R136a, one of the most massive members of
+in a region of 50 arcseconds around R136a, one of the most massive members of
 the cluster itself \citep[][]{crowther:10}, requiring the same
-``quality'' criteria applied above. This selection yields 62 stars.
-\todo{improve definition
-  of stars from R136, describe accordingly -- check with Danny}
+``quality'' criteria applied above. This selection yields 42 stars.
+\todo{improve definition of stars from R136, describe accordingly -- check with Danny}
 
@@
 \Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
-$\langle\mu_\mathrm{RA}\rangle = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
-$\langle\mu_\mathrm{DE}\rangle = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
+$\langle\mu_\mathrm{RA}^\mathrm{sur}\rangle = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
+$\langle\mu_\mathrm{DE}^\mathrm{sur}\rangle = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
 proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
@@
 \begin{equation}
-  \label{eq:speed}
-  v = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
+  \label{eq:speed_around}
+  v^\mathrm{sur} = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
     +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 22
@@
 The inset plot in \Figref{fig:main} shows a zoom in of the region
-around R136 which includes VFTS682. To test the hypothesis that the
-star was indeed ejected from this region, we check the
-orientation of its proper motion relative to the group of stars
-selected to represent the R136 cluster in \Secref{sec:sample}. We
-further consider also the kinematic age of the star. The error cone, obtained by
+around R136 which includes VFTS682. Stars belonging to our ``R136
+cluster'' group are shown in green (see \Secref{sec:sample}), together with their proper motion
+as listed in the Gaia DR2 catalog.
+To test the hypothesis that VFTS682 was indeed ejected from this region, we check the
+orientation of its proper motion relative to these stars. We further
+consider also the kinematic age of the star.
+
+We begin by evaluating the average proper motion for the stars
+representing the cluster. This is necessary to constrain the actual velocity
+obtained by VFTS682 at its ejection. By using \Eqref{eq:mean} we find  
+$\langle\mu_\mathrm{RA}^{R136}\rangle = 1.750\pm0.008\,\mathrm{mas\ yr^{-1}}$
+and $\langle\mu_\mathrm{DE}^\mathrm{R136}\rangle =
+0.742\pm0.009\,\mathrm{mas\ yr^{-1}}$. Therefore, using the values
+from \Tabref{tab:vfts682}, we obtain the components of the proper
+motion for VFTS682 relative to R136 $\mu^{R136}_\mathrm{RA} = 0.09\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^{R136}_\mathrm{DE} =
+0.04\pm 0.08\,\mathrm{mas\ yr^{-1}}$.\todo{improve numbers by
+  selecting better sample}.
+\todo{discuss if direction is compatible
+The error cone, obtained by
 propagating the errors on the proper motion components indicates that
@@
   banerjee:12}.
+}
+
+Assuming again a distance of 50\,kpc and
+not considering the error on the distance, we can convert these into
+the components of the tangential velocity relative to R136,
+$v^\mathrm{R136}_\mathrm{RA}=21 \pm 18\, \kms$ and $v^\mathrm{R136}_\mathrm{DE}=10
+\pm 21\,\kms$. Finally, using the peculiar line of sight velocity
+from \cite{bestenlehner:11}, we obtain the three-dimensional peculiar
+velocity of VFTS682 with respect to R136
+
+\begin{equation}
+  \label{eq:speed_R136}
+  v^\mathrm{sur} = \sqrt{\left(v^{R136}_\mathrm{RA}\right)^2
+    +\left(v^{R136}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 51 \pm 23
+  \, \kms \ .
+\end{equation}
 

### 2018-05-09 VFTS682.tex
no idea what's going on
introduced todo, added list of VFTS stars

@@
 \label{sec:intro}
-
+\todo{go to major star formation immediately}
+\todo{sentence 2:debate of isolated or not}
+\todo{paragraph 2 about }
 How do stars form is one longstanding question in astrophysics
@@
 detailed analysis on the larger sample is desirable.
+\todo{this paper raised the debate/interest on VFTS682}
+
+\todo{vfts682 is interesting also because}
 
@@
 \citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, so one would not expect the binary
-disruption scenario to be relevant for this cluster.
+disruption scenario to be relevant for this cluster. \todo{move up}
 
@@
 
+\todo{number above not trustworthy}
+
 \todo{kinematic age}
@@
   \includegraphics[width=\textwidth]{./figures/main_plot}  
-  \caption{position and projected relative velocity to
-    R136. \todo{orient properly, load  picture on background, add cone of uncertainty}}
+  \caption{The red cross indicates the position of VFTS682. Blue stars
+    are those we use to define the generic ``surroundings'', while the
+    green dots indicate the stars we use to probe the core of R136,
+    magnified in the inset. The red arrow shows the direction of the
+    proper motion of VFTS682 relative to R136, and the gray shade
+    indicates the uncertainty in the direction. \todo{orient properly, load  picture on background, add cone
+      of uncertainty, maybe show pm for all stars in inset?}}
   \label{fig:main}
@@
 
+\todo{compare to Lennon: time of ejection is the same? what does it
+  say for cluster}
+\todo{say implications for center of R136 (same stars are there), how
+  many ejected given this one, was a binary ejected in the other direction?}
+\todo{does this imply there is a binary in the core ?}
 \begin{itemize}

### 2018-05-10 VFTS682.tex
newdataset - Lennon-like
removed images

@@
 \todo{paragraph 2 about }
-How do stars form is one longstanding question in astrophysics
-\cite{lada:03, zinnecker:07}. It is particularly difficult for massive stars, because these are intrinsically rare
+How do massive stars form is one longstanding question in astrophysics
+\cite{lada:03, zinnecker:07}. This because they are intrinsically rare
 \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
@@
 complicated multipliticy structure which requires
-explanation \citep[][]{kobulnicky:07,mason:09,sana:11,sana:12,kiminki:12,chini:12,kobulnicky:14,almeida:17,demarco:17}. Understanding massive star formation, possibly as a
+explanation \citep[][]{sana:12,sana:17}. Understanding massive star formation, possibly as a
 function of metallicity, is a key question given the present and upcoming
@@
 evolution and death.
+
+Two competing classes of models for massive star formation exist, one predicting 
+that massive stars can form in relative isolation \todo{ref}, the other that
+they always form in cluster and/or associations, and therefore massive
+stars should be rarely isolated \todo{ref}.
 
@@
 \begin{acknowledgements}
-  The background image of \Figref{fig:main} is based on observations made with ESO Telescopes at the La Silla Observatory under programme ID 076.C-0888, processed and released by the ESO VOS/ADP group.
+  The background image of \Figref{fig:main} is based on observations
+  made with ESO Telescopes at the La Silla Observatory under programme
+  ID 076.C-0888, processed and released by the ESO VOS/ADP group.
+  This research made use of APLpy, an open-source plotting package for Python \cite{robitaille:12}.
   We are grateful to S.~Torres for extremely useful discussions on the

### 2018-05-10 VFTS682.tex
fixed sec. 3.1 - numbers are ok now

@@
 \label{sec:intro}
-\todo{go to major star formation immediately}
-\todo{sentence 2:debate of isolated or not}
-\todo{paragraph 2 about }
+
 How do massive stars form is one longstanding question in astrophysics
-\cite{lada:03, zinnecker:07}. This because they are intrinsically rare
+\citep[e.g.,][]{lada:03, zinnecker:07}, because they are intrinsically rare
 \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
@@
 
-Two competing classes of models for massive star formation exist, one predicting 
-that massive stars can form in relative isolation \todo{ref}, the other that
-they always form in cluster and/or associations, and therefore massive
-stars should be rarely isolated \todo{ref}.
+Two competing classes of models for massive star formation exist. One class predicts 
+that massive stars can form in relative isolation \todo{ref}, while the other predicts
+they always form in cluster and/or associations, and therefore
+isolated massive stars should necessarily peculiar velocities to reach
+an isolated position on the sky\todo{ref}.
 
@@
 these hypothesis using one particular star, VFTS682. 
-This star is a
-very massive \citep[$M_\mathrm{ZAMS}\simeq150\,M_\odot$,][]{bestenlehner:11,schneider:18} WNh5 star in the 30 Doradus region
-of the Large Magellanic Cloud (LMC), and it is presently observed at a
+This star belongs to the 30 Doradus region in the Large Magellanic
+Cloud (LMC), and has an inferred initial mass of $M_\mathrm{ZAMS}=150.0^{+28.7}_{-17.4}\,M_\odot$
+\citep[][]{bestenlehner:11,schneider:18}. Its spectral type is WNh5, and it is presently observed at a
 projected distance of $\sim$$29$\,pc from the nearest cluster of
-massive stars R136 \citep[][]{bestenlehner:11}. Based on the extremely high mass of this star and
-its present day apparent isolation, \cite{bestenlehner:11} proposed it
-might be a candidate for isolated star formation, or a ``slow runaway'' ejected
-from R136 in the past. This second option is also supported by the
-N-body simulations of \cite{fujii:11, banerjee:12}.   Many other
-very massive stars are present in the surroundings of R136, and a more
-detailed analysis on the larger sample is desirable.
-\todo{this paper raised the debate/interest on VFTS682}
+massive stars R136 \citep[also known as NGC2070][]{bestenlehner:11}. The inferred mass loss
+rate is $\sim10^{-XX}\,M_\odot \ \mathrm{yr}^{-1}$ \todo{ref}.
 
-\todo{vfts682 is interesting also because}
-
-Massive stars can in principle be ejected from R136 as a consequence of dynamical
-interactions \citep[][]{poveda:67,leonard:91, evans:10, fujii:11,
-  allison:12, oh:16}, or by the disruption of a binary by the first
-core-collapse supernova
-\citep[][]{zwicky:57,blaauw:61,dedonder:97,eldridge:11,renzo:18}. However,
-R136 has an estimated age of $\lesssim2$\,Myr \citep[][]{sabbi:12}, which is
-shorter than the shortest stellar lifetime
-\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, so one would not expect the binary
-disruption scenario to be relevant for this cluster. \todo{move up}
+Based on the extremely high mass of this star and
+its present-day apparent isolation, \cite{bestenlehner:11} proposed it
+could be a candidate for isolated star formation, or a ``slow runaway'' ejected
+from R136 in the past. Their study raised the interest in the origin
+of this star. \cite{fujii:11, banerjee:12} carried out N-body
+simulations of massive young stellar clusters and concluded that the
+ejection of such a massive object from R136 through dynamical
+interactions \citep[e,g,][]{poveda:67} would be possible.
+The competing ejection mechanism, the disruption of a binary
+system by a core-collapse event \citep[][]{zwicky:57, blaauw:61}, is
+thought to be ruled out, since the cluster has an estimated age of
+$\lesssim2$\,Myr \citep[][]{sabbi:12}, which is shorter than the shortest stellar lifetime
+\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}.
 
@@
 used to define a local reference frame. Our results
-indicate that R136 is a bona fide runaway star (\Secref{sec:runaway}), therefore isolated star formation is \emph{not}
+indicate that R136 is a bona fide runaway star (\Secref{sec:runaway}),
+therefore isolated star formation is \emph{not}
 required to explain it. We also find that a dynamical ejection from
 R136 is compatible with the direction of its velocity
-vector. \todo{double check the following}
-However, we find a mild discrepancy between the apparent age of
-VFTS682, its kinematic age, and the age estimates for R136 (\Secref{sec:r136_origin}).  We conclude
+vector. 
+We conclude
 with a very brief discussion on the implications for theories of
@@
 
-To compare the astrometry of VFTS682 and derive its peculiar motion,
-we then select data from the Gaia DR2 catalog for two regions: the
-``surroundings'' of VFTS 682, and the ``R136 cluster''. The
-surrounding region is defined by all the stars in a target of 10 arcminutes around
-VFTS682 fulfilling the following criteria: we require \texttt{visibility\_period} $\geq$ 5,
-\texttt{astrometric\_excess\_noise} < 1, the error on the proper
+We define a local standard frame to derive the peculiar velocity
+of VFTS682 by selecting from the Gaia DR2 catalog a sample of nearby
+stars, following closely the approach of \cite{lennon:18}.
+We select all the stars in a target of 0.2 degrees around R136
+(NGC2070) fulfilling the following criteria. First, we require G-band
+magnitude brighter than 17, correspondingly roughly to the
+completeness level of the VFTS survey \citep[here we implicitly assume
+G$\sim$V,][]{evans:11}. Then we require \texttt{visibility\_period} $\geq$ 5,
+\texttt{astrometric\_excess\_noise} < 1, the errors on the proper
 motion components to be smaller than 0.1\,$\mathrm{mas\ yr^{-1}}$,
@@
 likely to be foreground stars. We checked that the additional
-requirement of having parallaxes smaller than $1\,\mathrm{mas\
-  yr^{-1}}$ does not reduces further our sample. This selection yields 437 stars.
+requirement of having parallaxes smaller than $2\,\mathrm{mas}$ does
+not reduces further our sample.
 
-The ``R136 cluster'' is effectively defined by taking all the stars
-in a region of 50 arcseconds around R136a, one of the most massive members of
-the cluster itself \citep[][]{crowther:10}, requiring the same
-``quality'' criteria applied above. This selection yields 42 stars.
-\todo{improve definition of stars from R136, describe accordingly -- check with Danny}
-
-Throughout this study, we assume the same distance of $50$\,kpc to the star, and to
-the 30 Doradus region as a whole.% , since the parallax for VFTS682
-% listed in the Gaia DR2 catalog is negative. %% this is what makes the
-%                                 % LOS velocity useless...how to improve?
-
-For each of the two local rest frames (``surroundings'' and ``R136
-cluster''), we compute the average transverse velocity as:
-
+We calculate the averaged proper motion components for the whole
+region using 
 \begin{equation}
@@
 where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
-on the proper motion component reported by Gaia, and the sums run over
-all the $N$ stars of the frame considered. We evaluate each proper motion
-component separately.
-
-The data retrieved, and the ipython notebook used for the analysis
+on the proper motion component reported by Gaia. The sums run over
+all the $N=651$ stars in our selected sample. We evaluate each proper motion
+component separately. For simplicity, throughout this study, we assume the same
+distance of $50$\,kpc to the star \citep[][]{lebouteiller:08}, and to
+the 30 Doradus region as a whole. We do not consider the error bars on
+the distance determination when converting proper motions into
+physical velocities. The data retrieved, and the ipython notebook used for the analysis
 presented here will be made available at \todo{probably git repo on bitbucket?}.
@@
 
-Using the 437 stars selected as the ``surrounding population'' in
-\Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
-$\langle\mu_\mathrm{RA}^\mathrm{sur}\rangle = 1.695\pm0.003\,\mathrm{mas\ yr^{-1}}$ and
-$\langle\mu_\mathrm{DE}^\mathrm{sur}\rangle = 0.691\pm0.003\,\mathrm{mas\ yr^{-1}}$. Subtracting these values from the
+Using the 651 stars selected as described in \Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
+$\langle\mu_\mathrm{RA}^\mathrm{sur}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
+$\langle\mu_\mathrm{DE}^\mathrm{sur}\rangle =
+0.672\pm0.003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
+in good agreement with what found by \cite{lennon:18}. Subtracting these values from the
 proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
 components of proper motion of the star relative to the surrounding region
-$\mu^\mathrm{sur}_\mathrm{RA} = 0.15\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
-0.09\pm 0.08\,\mathrm{mas\ yr^{-1}}$. %% note that the error budget is
-% dominated by VFTS682
+$\mu^\mathrm{sur}_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
+0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$.
 These can be converted in the
-the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=35\pm17\,\kms$,
-$v^\mathrm{sur}_\mathrm{DE}=22\pm19\,\kms$, assuming a distance of
+the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=38\pm17\,\kms$,
+$v^\mathrm{sur}_\mathrm{DE}=26\pm19\,\kms$, assuming a distance of
 50\,kpc (we do not account for the uncertainty in the distance
@@
   v^\mathrm{sur} = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
-    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 61 \pm 22
+    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta
+      v_\mathrm{rad}\right)^2} = 64 \pm 21 
   \, \kms \ .
@@
 This value for the three-dimensional speed of VFTS682 with respect the
-surrounding stars make it a ``bona fide'' runaway star.
+surrounding stars make it the most massive ``bona fide'' runaway star
+known to date.
 
@@
 
-The inset plot in \Figref{fig:main} shows a zoom in of the region
-around R136 which includes VFTS682. Stars belonging to our ``R136
-cluster'' group are shown in green (see \Secref{sec:sample}), together with their proper motion
-as listed in the Gaia DR2 catalog.
-To test the hypothesis that VFTS682 was indeed ejected from this region, we check the
-orientation of its proper motion relative to these stars. We further
-consider also the kinematic age of the star.
+% The inset plot in \Figref{fig:main} shows a zoom in of the region
+% around R136 which includes VFTS682. Stars belonging to our ``R136
+% cluster'' group are shown in green (see \Secref{sec:sample}), together with their proper motion
+% as listed in the Gaia DR2 catalog.
+% To test the hypothesis that VFTS682 was indeed ejected from this region, we check the
+% orientation of its proper motion relative to these stars. We further
+% consider also the kinematic age of the star.
 
-We begin by evaluating the average proper motion for the stars
-representing the cluster. This is necessary to constrain the actual velocity
-obtained by VFTS682 at its ejection. By using \Eqref{eq:mean} we find  
-$\langle\mu_\mathrm{RA}^{R136}\rangle = 1.750\pm0.008\,\mathrm{mas\ yr^{-1}}$
-and $\langle\mu_\mathrm{DE}^\mathrm{R136}\rangle =
-0.742\pm0.009\,\mathrm{mas\ yr^{-1}}$. Therefore, using the values
-from \Tabref{tab:vfts682}, we obtain the components of the proper
-motion for VFTS682 relative to R136 $\mu^{R136}_\mathrm{RA} = 0.09\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^{R136}_\mathrm{DE} =
-0.04\pm 0.08\,\mathrm{mas\ yr^{-1}}$.\todo{improve numbers by
-  selecting better sample}.
-\todo{discuss if direction is compatible
-The error cone, obtained by
-propagating the errors on the proper motion components indicates that
-R136 is indeed the most likely origin, as expected by \cite{fujii:11,
-  banerjee:12}.
-}
+% We begin by evaluating the average proper motion for the stars
+% representing the cluster. This is necessary to constrain the actual velocity
+% obtained by VFTS682 at its ejection. By using \Eqref{eq:mean} we find  
+% $\langle\mu_\mathrm{RA}^{R136}\rangle = 1.750\pm0.008\,\mathrm{mas\ yr^{-1}}$
+% and $\langle\mu_\mathrm{DE}^\mathrm{R136}\rangle =
+% 0.742\pm0.009\,\mathrm{mas\ yr^{-1}}$. Therefore, using the values
+% from \Tabref{tab:vfts682}, we obtain the components of the proper
+% motion for VFTS682 relative to R136 $\mu^{R136}_\mathrm{RA} = 0.09\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^{R136}_\mathrm{DE} =
+% 0.04\pm 0.08\,\mathrm{mas\ yr^{-1}}$.\todo{improve numbers by
+%   selecting better sample}.
+% \todo{discuss if direction is compatible
+% The error cone, obtained by
+% propagating the errors on the proper motion components indicates that
+% R136 is indeed the most likely origin, as expected by \cite{fujii:11,
+%   banerjee:12}.
+% }
 
-Assuming again a distance of 50\,kpc and
-not considering the error on the distance, we can convert these into
-the components of the tangential velocity relative to R136,
-$v^\mathrm{R136}_\mathrm{RA}=21 \pm 18\, \kms$ and $v^\mathrm{R136}_\mathrm{DE}=10
-\pm 21\,\kms$. Finally, using the peculiar line of sight velocity
-from \cite{bestenlehner:11}, we obtain the three-dimensional peculiar
-velocity of VFTS682 with respect to R136
+% Assuming again a distance of 50\,kpc and
+% not considering the error on the distance, we can convert these into
+% the components of the tangential velocity relative to R136,
+% $v^\mathrm{R136}_\mathrm{RA}=21 \pm 18\, \kms$ and $v^\mathrm{R136}_\mathrm{DE}=10
+% \pm 21\,\kms$. Finally, using the peculiar line of sight velocity
+% from \cite{bestenlehner:11}, we obtain the three-dimensional peculiar
+% velocity of VFTS682 with respect to R136
 
-\begin{equation}
-  \label{eq:speed_R136}
-  v^\mathrm{sur} = \sqrt{\left(v^{R136}_\mathrm{RA}\right)^2
-    +\left(v^{R136}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 51 \pm 23
-  \, \kms \ .
-\end{equation}
+% \begin{equation}
+%   \label{eq:speed_R136}
+%   v^\mathrm{sur} = \sqrt{\left(v^{R136}_\mathrm{RA}\right)^2
+%     +\left(v^{R136}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 51 \pm 23
+%   \, \kms \ .
+% \end{equation}
 
-\todo{number above not trustworthy}
-
-\todo{kinematic age}
+\subsection{Kinematic age}
 
@@
 
-\section{Summary and Discussion}
+\section{Discussion}
 \label{sec:discussion}
@@
 \end{itemize}
+\section{Summary}
+
+1. Gaia data confirm a 150Msun runaway with v~60, thus the most
+massive known to date
+2. directions suggest ejection from R136, confirming prediction of
+fuji+ banerjee+
+2b. This fits with the observation from crowther of M>300Msun in R136
+(and mention vfts682 has the same spectral type): the most likely
+mechanism needs encounter with 2 more massive -> additional evidence
+for Crowther claim
+3. comment on final fate (BH)
 
@@
 from \cite{schneider:18}
+
+Many other very massive stars are present in the surroundings this cluster, and a more
+detailed analysis on the larger sample is desirable.
 

### 2018-05-10 VFTS682.tex
new fig. wrote something for kinematic age

@@
 $\mu^\mathrm{sur}_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
-0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$.
+0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$. We note that the error budget is
+dominated by the errors on the proper motion components of VFTS682.
+
 These can be converted in the
@@
 
-% The inset plot in \Figref{fig:main} shows a zoom in of the region
-% around R136 which includes VFTS682. Stars belonging to our ``R136
-% cluster'' group are shown in green (see \Secref{sec:sample}), together with their proper motion
-% as listed in the Gaia DR2 catalog.
-% To test the hypothesis that VFTS682 was indeed ejected from this region, we check the
-% orientation of its proper motion relative to these stars. We further
-% consider also the kinematic age of the star.
+The red arrow in \Figref{fig:main} shows the proper motion of VFTS682
+relative to the region, and the lighter red arrows show the possible
+range of directions within the uncertainties in the measured proper
+motion. It is clear that the most likely origin of the star is R136,
+as predicted by \cite{fujii:11, banerjee:12}.
 
-% We begin by evaluating the average proper motion for the stars
-% representing the cluster. This is necessary to constrain the actual velocity
-% obtained by VFTS682 at its ejection. By using \Eqref{eq:mean} we find  
-% $\langle\mu_\mathrm{RA}^{R136}\rangle = 1.750\pm0.008\,\mathrm{mas\ yr^{-1}}$
-% and $\langle\mu_\mathrm{DE}^\mathrm{R136}\rangle =
-% 0.742\pm0.009\,\mathrm{mas\ yr^{-1}}$. Therefore, using the values
-% from \Tabref{tab:vfts682}, we obtain the components of the proper
-% motion for VFTS682 relative to R136 $\mu^{R136}_\mathrm{RA} = 0.09\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^{R136}_\mathrm{DE} =
-% 0.04\pm 0.08\,\mathrm{mas\ yr^{-1}}$.\todo{improve numbers by
-%   selecting better sample}.
-% \todo{discuss if direction is compatible
-% The error cone, obtained by
-% propagating the errors on the proper motion components indicates that
-% R136 is indeed the most likely origin, as expected by \cite{fujii:11,
-%   banerjee:12}.
-% }
+%\subsection{Kinematic age}
 
-% Assuming again a distance of 50\,kpc and
-% not considering the error on the distance, we can convert these into
-% the components of the tangential velocity relative to R136,
-% $v^\mathrm{R136}_\mathrm{RA}=21 \pm 18\, \kms$ and $v^\mathrm{R136}_\mathrm{DE}=10
-% \pm 21\,\kms$. Finally, using the peculiar line of sight velocity
-% from \cite{bestenlehner:11}, we obtain the three-dimensional peculiar
-% velocity of VFTS682 with respect to R136
+We can therefore consider the kinematic age for this star assuming
+that it originates from the cluster,
 
-% \begin{equation}
-%   \label{eq:speed_R136}
-%   v^\mathrm{sur} = \sqrt{\left(v^{R136}_\mathrm{RA}\right)^2
-%     +\left(v^{R136}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 51 \pm 23
-%   \, \kms \ .
-% \end{equation}
-
-\subsection{Kinematic age}
+\begin{equation}
+  \label{eq:kin_age}
+  \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} =
+  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.63\, \mathrm{Myr} \ \ ,
+\end{equation}
+where $d_\parallel =29$\,pc is the projected distance from VFTS682 to
+the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel = 46\pm
+17\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
+using the relative proper motion components calculated in
+\Secref{sec:runaway}, and we use the approximation $1 \kms \simeq 1
+\mathrm{pc \ Myr^{-1}}$.
 
@@
   We are grateful to S.~Torres for extremely useful discussions on the
-  Gaia DR2 dataset.
+  Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for help with APLpy.
 \end{acknowledgements}

### 2018-05-10 VFTS682.tex
new sec. origin

@@
 
-\begin{figure*}[htbp]
+The kinematic age $\tau_\mathrm{kin}$ is smaller than the apparent age
+of the star age $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
+corroborates the idea that the star is the result of a dynamical
+ejection. The value of $\tau_\mathrm{kin}$ we find is also smaller
+than the corresponding value for another massive runaway in 30 Dor,
+VFTS16 for which \cite{lennon:18} inferred a kinematic age of
+$\sim$1.5\,Myr. This means that the more
+massive VFTS682 was ejected later than VFTS16.
+
+\begin{figure}[htbp]
   \centering
-  \includegraphics[width=\textwidth]{./figures/main_plot}  
+  \includegraphics[width=0.5\textwidth]{./figures/main_plot}  
   \caption{The red cross indicates the position of VFTS682. Blue stars
@@
     proper motion of VFTS682 relative to R136, and the gray shade
-    indicates the uncertainty in the direction. \todo{orient properly, load  picture on background, add cone
-      of uncertainty, maybe show pm for all stars in inset?}}
+    indicates the uncertainty in the direction. \todo{load color picture on background}}
   \label{fig:main}
-\end{figure*}
+\end{figure}
 

### 2018-05-10 VFTS682.tex
new sec. origin

@@
 thought to be ruled out, since the cluster has an estimated age of
-$\lesssim2$\,Myr \citep[][]{sabbi:12}, which is shorter than the shortest stellar lifetime
+$\lesssim2$\,Myr \citep[][]{crowther:10,sabbi:12}, which is shorter than the shortest stellar lifetime
 \citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}.
@@
 We conclude
-with a very brief discussion on the implications for theories of
-star formation, N-body interactions, and binary evolution in
+with a very brief discussion on the implications for star formation,
+cluster evolution, and stellar physics studies in
 \Secref{sec:discussion}.
@@
 
-Using the 651 stars selected as described in \Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
+Using the 651 stars selected as described in \Secref{sec:sample} (a
+subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
 $\langle\mu_\mathrm{RA}^\mathrm{sur}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
@@
 corroborates the idea that the star is the result of a dynamical
-ejection. The value of $\tau_\mathrm{kin}$ we find is also smaller
-than the corresponding value for another massive runaway in 30 Dor,
-VFTS16 for which \cite{lennon:18} inferred a kinematic age of
-$\sim$1.5\,Myr. This means that the more
-massive VFTS682 was ejected later than VFTS16.
+ejection.
 
@@
 
-\todo{compare to Lennon: time of ejection is the same? what does it
-  say for cluster}
-\todo{say implications for center of R136 (same stars are there), how
-  many ejected given this one, was a binary ejected in the other direction?}
-\todo{does this imply there is a binary in the core ?}
+Based on our results, we can claim that VFTS682 is the most massive
+runaway known to date, with a peculiar three-dimensional spatial
+velocity of $\sim60\,\kms$. This means that isolated star formation is
+\emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
+$\sim0.6$\,Myr ago. Given the exceptionally large mass
+of this star, this raises the question of which stars must populate
+the core of the cluster.
+
+Dynamical ejections due to N-body interactions typically (although not necessarily) eject the least
+massive star among those interacting \todo{ref}. This means that, just
+based on the kinematic properties of VFTS682, we would expect several
+stars with initial masses larger than $\sim150\,M_\odot$ in R136.
+This can be considered an independent confirmation of the detection
+of extremely massive stars by \cite{crowther:10} in the core of the
+cluster. The projected rotational equatorial velocity of VFTS682
+reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
+line with the average rotation rate of massive stars in the region
+\citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 has not
+experienced binary interactions, nor it will, since the multi-epoch
+data of the VFTS survey rule out present-day binarity. Moreover, the spectral type of VFTS682
+\citep[WNh5][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
+three most massive stars detected in the core of the cluster by
+\cite{crowther:10}. Therefore,
+VFTS682 might be an ideal target to constrain the stellar physics of
+stars with masses well above $\sim100\,M_\odot$: its isolation makes
+it an easier target for observation than the crowded core of R136,
+where similar stars exist.
+
+The similarities between VFTS682 and the WNh5 stars in the core of
+R136 are also in agreement with the ``bully binary'' model of
+\cite{fujii:11}. They suggested, based on their numerical simulations,
+that dynamical interactions lead to the formation of an extremely
+massive binary in the early phases of a cluster evolution, and then it tightens by ejecting other stars passing
+by. Interpreting our results for VFTS682 through the lens of their simulations would
+suggest the presence of a binary with total mass
+$M_1+M_2\gtrsim 400\,M_\odot$ in the core of the cluster, which they
+suggested could be R145. The kinematic age of VFTS682 puts an
+upperlimit to the timescale to form such ``bully binary'' in
+R136. Unfortunately, their simulations did not include stars more
+massive than $100\,M_\odot$, so it is difficult to predict, based on
+the existence of VFTS682, how many other stars should have been
+ejected from the cluster already and what is their mass
+distribution. A comprehensive study of the kinematic properties of the
+large sample of stars with extremely large inferred masses visible
+around R136 is encouraged.
+
+\cite{lennon:18} have carried out a similar study on VFTS16, another
+massive star ($\sim90\,M_\odot$) in the 30 Doradus region, previously
+known to be a runaway from the value of
+its radial velocity. They also concluded that VFTS16 is 
+the result of a dynamical ejection from the R136 cluster. 
+The value of $\tau_\mathrm{kin}\simeq0.63$\,Myr we find for VFTS682 (see \Secref{sec:r136}) is smaller
+than the corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic age of
+$\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
+massive VFTS682 was ejected later than VFTS16 from the same cluster.
+
 \begin{itemize}
-\item VFTS682 is a bona fide runaway with $v\sim60\,\mathrm{km\
-    s^{-1}}$ thrown out from R136. Both its speed and the age of the
-  cluster are consitent with a dynamical ejection.
-  \item VFTS682 comes from R136 as was expected by
-  \cite{bestenlehner:11, fujii:11, banerjee:12}, so it does not
-  require isolated SFH to be explained
-\item apparent age tension (connect to VFTS16 as well).
 \item is R136 a single young cluster or a merger
@@
   what is its total mass and relaxation time?
+\item comment on final fate (BH)
+
 \end{itemize}
-\section{Summary}
-
-1. Gaia data confirm a 150Msun runaway with v~60, thus the most
-massive known to date
-2. directions suggest ejection from R136, confirming prediction of
-fuji+ banerjee+
-2b. This fits with the observation from crowther of M>300Msun in R136
-(and mention vfts682 has the same spectral type): the most likely
-mechanism needs encounter with 2 more massive -> additional evidence
-for Crowther claim
-3. comment on final fate (BH)
-
-Random notes: $v\sin(i)<200\,\mathrm{km\ s^{-1}}$ form \cite{schneider:18}, age $1.0\pm 0.2$\,Myr
-from \cite{schneider:18}
-
-Many other very massive stars are present in the surroundings this cluster, and a more
-detailed analysis on the larger sample is desirable.
 

### 2018-05-10 VFTS682.tex
re-read stuff

@@
 
-\author{M.~Renzo\inst{1} \and \todo{TBD}% S.~E.~de~Mink\inst{1} \and
+\author{M.~Renzo\inst{1} \and S.~E.~de~Mink\inst{1} \todo{etc...}%\and
   % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
@@
 massive stars R136 \citep[also known as NGC2070][]{bestenlehner:11}. The inferred mass loss
-rate is $\sim10^{-XX}\,M_\odot \ \mathrm{yr}^{-1}$ \todo{ref}.
+rate is $\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$ \todo{ref}.
 
@@
 
-We retrieve for VFTS682 the position in right ascension (RA) and declination (DE)
+We retrieve for VFTS682 the position in right ascension (RA) and declination (DEC)
 in the IGCS frame \cite[][]{brown:18}, its
-proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DE}$,
+proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
 respectively). For the radial velocity of VFTS682 and of the 30 Doradus
@@
     RA \hfill[degree] &  \phantom{-}84.73 $\pm$  0.03 & \multirow{4}{*}{Gaia DR2}\\
-    DE \hfill [degree] & -69.07 $\pm$  0.05  & \\
+    DEC \hfill [degree] & -69.07 $\pm$  0.05  & \\
     $\mu_\mathrm{RA}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}1.84 $\pm$ 0.07 & \\
-    $\mu_\mathrm{DE}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}0.78 $\pm$ 0.08& \\
+    $\mu_\mathrm{DEC}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}0.78 $\pm$ 0.08& \\
     $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{0}-45 $\pm$ 25 & \cite{bestenlehner:11}\\
@@
 subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
-$\langle\mu_\mathrm{RA}^\mathrm{sur}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
-$\langle\mu_\mathrm{DE}^\mathrm{sur}\rangle =
+$\langle\mu_\mathrm{RA}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
+$\langle\mu_\mathrm{DEC}\rangle =
 0.672\pm0.003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
@@
 components of proper motion of the star relative to the surrounding region
-$\mu^\mathrm{sur}_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
+$\mu_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu_\mathrm{DEC} =
 0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$. We note that the error budget is
@@
 These can be converted in the
-the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=38\pm17\,\kms$,
-$v^\mathrm{sur}_\mathrm{DE}=26\pm19\,\kms$, assuming a distance of
+the components of the relative transverse velocity $\delta v_\mathrm{RA}=38\pm17\,\kms$,
+$\delta v_\mathrm{DEC}=26\pm19\,\kms$, assuming a distance of
 50\,kpc (we do not account for the uncertainty in the distance
@@
 \cite{bestenlehner:11} then gives the third component along
-the line of sight, allowing us to calculate the speed of the star:
+the line of sight, allowing us to calculate the three-dimensional
+peculiar speed of the star:
 
@@
   \label{eq:speed_around}
-  v^\mathrm{sur} = \sqrt{\left(v^\mathrm{sur}_\mathrm{RA}\right)^2
-    +\left(v^\mathrm{sur}_\mathrm{DE}\right)^2+\left(\delta
+  v_\mathrm{pec} = \sqrt{\left(\delta v_\mathrm{RA}\right)^2
+    +\left(\delta v_\mathrm{DEC}\right)^2+\left(\delta
       v_\mathrm{rad}\right)^2} = 64 \pm 21 
@@
 where $d_\parallel =29$\,pc is the projected distance from VFTS682 to
-the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel = 46\pm
+the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel \equiv \sqrt{\left(\delta v_\mathrm{RA}\right)^2
+    +\left(\delta v_\mathrm{DEC}\right)^2} =46\pm
 17\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
@@
 runaway known to date, with a peculiar three-dimensional spatial
-velocity of $\sim60\,\kms$. This means that isolated star formation is
+velocity of $\sim$$60\,\kms$. This means that isolated star formation is
 \emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
-$\sim0.6$\,Myr ago. Given the exceptionally large mass
+$\sim$$0.6$\,Myr ago. Because of the exceptionally large mass
 of this star, this raises the question of which stars must populate
@@
 
-Dynamical ejections due to N-body interactions typically (although not necessarily) eject the least
+Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
 massive star among those interacting \todo{ref}. This means that, just
 based on the kinematic properties of VFTS682, we would expect several
-stars with initial masses larger than $\sim150\,M_\odot$ in R136.
+stars with initial masses larger than $\sim$$150\,M_\odot$ in the
+cluster R136.
 This can be considered an independent confirmation of the detection
@@
 experienced binary interactions, nor it will, since the multi-epoch
-data of the VFTS survey rule out present-day binarity. Moreover, the spectral type of VFTS682
-\citep[WNh5][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
+data of the VFTS survey rule out the presence of a companion at
+present day. Moreover, the spectral type of VFTS682
+\citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
 three most massive stars detected in the core of the cluster by
@@
 stars with masses well above $\sim100\,M_\odot$: its isolation makes
-it an easier target for observation than the crowded core of R136,
-where similar stars exist.
+it an easier target for observations compared to the similar stars present the crowded core of R136.
 
@@
 R136 are also in agreement with the ``bully binary'' model of
-\cite{fujii:11}. They suggested, based on their numerical simulations,
-that dynamical interactions lead to the formation of an extremely
-massive binary in the early phases of a cluster evolution, and then it tightens by ejecting other stars passing
-by. Interpreting our results for VFTS682 through the lens of their simulations would
-suggest the presence of a binary with total mass
-$M_1+M_2\gtrsim 400\,M_\odot$ in the core of the cluster, which they
-suggested could be R145. The kinematic age of VFTS682 puts an
+\cite{fujii:11}. Based on their numerical results, they suggested that
+early in the evolution of a cluster, dynamical interactions form an extremely
+massive binary, which then tightens its orbit by ejecting other stars passing
+by. Interpreting our results for VFTS682 through the lens of their simulations
+suggests the presence of a binary with total mass
+$M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. Such bully
+binary could be R145 according to \cite{fujii:11}.
+
+The kinematic age of VFTS682 puts an
 upperlimit to the timescale to form such ``bully binary'' in
-R136. Unfortunately, their simulations did not include stars more
+R136. The cluster must have been at the very beginning of its
+evolution, given the age estimate of $\lesssim 2$\,Myr
+\cite{crowther:10,sabbi:12} and the kinematic age of VFTS16.
+
+\todo{check \cite{banerjee} and rewrite next paragraph, they
+  have stuff} Unfortunately, their simulations did not include stars more
 massive than $100\,M_\odot$, so it is difficult to predict, based on
@@
 the result of a dynamical ejection from the R136 cluster. 
-The value of $\tau_\mathrm{kin}\simeq0.63$\,Myr we find for VFTS682 (see \Secref{sec:r136}) is smaller
+The value of $\tau_\mathrm{kin}\simeq0.63$\,Myr we find for VFTS682 (see \Secref{sec:r136_origin}) is smaller
 than the corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic age of
@@
   what is its total mass and relaxation time?
-\item comment on final fate (BH)
-
 \end{itemize}
@@
 \begin{acknowledgements}
+  We are grateful to S.~Torres for extremely useful discussions on the
+  Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
   The background image of \Figref{fig:main} is based on observations
@@
   ID 076.C-0888, processed and released by the ESO VOS/ADP group.
-  This research made use of APLpy, an open-source plotting package for Python \cite{robitaille:12}.
-  We are grateful to S.~Torres for extremely useful discussions on the
-  Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for help with APLpy.
+  This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
 \end{acknowledgements}

### 2018-05-10 VFTS682.tex
Intro: see pink

@@
 \definecolor{Wildstrawberry}{rgb}{1.0, 0.26, 0.64}
-\newcommand{\SdM}[1]{{\color{Wildstrawberry}\bf{#1}}}
+\newcommand{\SdM}[1]{{{\color{Wildstrawberry}\bf{#1}}}}
 \newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}
@@
 
-\author{M.~Renzo\inst{1} \and S.~E.~de~Mink\inst{1} \todo{etc...}%\and
+\author{M.~Renzo\inst{1} \and S.~E.~de~Mink\inst{1} \and \SdM{D.J. Lennon et al.} \todo{etc...}%\and
   % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
@@
 
-How do massive stars form is one longstanding question in astrophysics
-\citep[e.g.,][]{lada:03, zinnecker:07}, because they are intrinsically rare
-\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, evolve fast, and
-remain enshrouded in their parent cloud during the formation
-process. Moreover, observations of young massive stars reveal a
-complicated multipliticy structure which requires
-explanation \citep[][]{sana:12,sana:17}. Understanding massive star formation, possibly as a
-function of metallicity, is a key question given the present and upcoming
-transient survey \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{}\todo{ref} which
-will reveal transients associated to massive stars
+\SdM{
+
+How do massive stars form is one of the major longstanding questions in astrophysics
+\citep[e.g.,][]{zinnecker:07}. Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
+%\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
+evolve fast, typically reside in dense groups and they remain enshrouded in their parent cloud during the formation
+process.
+Major progress has been made on the theoretical side,  \citep[e.g.][]{Kuiper+2015Rosen+2016}, but the simulations are computationally very expensive.
+They can typically not yet resolve the length scales that are of interest to understand the
+high level of multiplicity among massive stars  \citep[][]{sana:12,sana:17}.
+Understanding massive star formation, and its possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies as sources of feedback, but also for understanding the transients that mark their death and the compact remnants they leave behind.
+This is therefore a key question for the present and upcoming transient surveys \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{}\todo{ref} which  will reveal transients associated to massive stars
 evolution and death.
 
-Two competing classes of models for massive star formation exist. One class predicts 
-that massive stars can form in relative isolation \todo{ref}, while the other predicts
-they always form in cluster and/or associations, and therefore
-isolated massive stars should necessarily peculiar velocities to reach
-an isolated position on the sky\todo{ref}.
+In has been conjectured that most, if not all, stars form in clusters \cite{lada:03}, where massive stars are thought to  form preferentially in the dense cores of clusters.
+In this picture, field stars are primarily the result of the dissolution of dense groups.
+ However, a significant population of massive stars has been  in relative isolation,  far from dense clusters or OB associations.
+The origin of this population still remains debated \citep{Lamb+2016}.
+One hypothesis is that they  formed in the field, but this poses a challenge for the theories of star formation.
+The alternative hypothesis is that these massive stars  formed in clusters, and then been ejected from their birth locations.
+Such ejections may result from   \citep[e,g,][]{poveda:67} or from the dsiruption of binary systems by the supernova of the companion star  \citep[][]{zwicky:57, blaauw:61}.
 
-The second data release (DR2) from the Gaia satellite
-\cite[][]{gaia:16,brown:18} allows us to test
-these hypothesis using one particular star, VFTS682. 
-This star belongs to the 30 Doradus region in the Large Magellanic
-Cloud (LMC), and has an inferred initial mass of $M_\mathrm{ZAMS}=150.0^{+28.7}_{-17.4}\,M_\odot$
-\citep[][]{bestenlehner:11,schneider:18}. Its spectral type is WNh5, and it is presently observed at a
-projected distance of $\sim$$29$\,pc from the nearest cluster of
-massive stars R136 \citep[also known as NGC2070][]{bestenlehner:11}. The inferred mass loss
-rate is $\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$ \todo{ref}.
+ A very interesting contribution to the debate whether or not massive stars an form in relative isolation was presented in recent years by  \cite{bestenlehner:11}, who discuss the case of the very massive star VFTS682.
+This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of massive stars \citep{Evans+2011}.
+With an spectral type  WNh5, an inferred initial mass of $M_\mathrm{ZAMS}=150.0^{+28.7}_{-17.4}\,M_\odot$  and a mass loss rate of  $\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$  \citep[][]{bestenlehner:11,schneider:18} it belongs to the most extreme objects in the region.
+It is reminiscent of the very  massive stars, aka monster stars, identified by  \citet{Crowther+2010} and \citet{Crowther+2016}.
 
-Based on the extremely high mass of this star and
-its present-day apparent isolation, \cite{bestenlehner:11} proposed it
-could be a candidate for isolated star formation, or a ``slow runaway'' ejected
-from R136 in the past. Their study raised the interest in the origin
-of this star. \cite{fujii:11, banerjee:12} carried out N-body
-simulations of massive young stellar clusters and concluded that the
-ejection of such a massive object from R136 through dynamical
-interactions \citep[e,g,][]{poveda:67} would be possible.
-The competing ejection mechanism, the disruption of a binary
-system by a core-collapse event \citep[][]{zwicky:57, blaauw:61}, is
-thought to be ruled out, since the cluster has an estimated age of
-$\lesssim2$\,Myr \citep[][]{crowther:10,sabbi:12}, which is shorter than the shortest stellar lifetime
-\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}.
+Most remarkable is the location of  VFTS682.
+It is located at a projected distance of $\sim$$29$\,pc from the core of the central ionizing star cluster R136.
+\citet{bestenlehner:11} discuss two possibilities for the  origin of VFTS 682.
+Quoting directly from their abstract:  "either (i) the star either formed in situ, which would have profound implications for the formation mechanism of massive stars, or (ii) VFTS 682 is a slow runaway star that originated from the dense cluster R136, which would make it the most massive runaway known to date."
 
-In this study, we combine the radial velocity measurements from the
-VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2
-to reconstruct the three-dimensional velocity of VFTS682, and test the
-hypothesis that this star was ejected from R136. We discuss in
-\Secref{sec:sample} the data for VFTS682, and the selection of stars
-used to define a local reference frame. Our results
-indicate that R136 is a bona fide runaway star (\Secref{sec:runaway}),
-therefore isolated star formation is \emph{not}
-required to explain it. We also find that a dynamical ejection from
-R136 is compatible with the direction of its velocity
-vector. 
-We conclude
-with a very brief discussion on the implications for star formation,
-cluster evolution, and stellar physics studies in
-\Secref{sec:discussion}.
+In this study we report on our analysis of Gaia data that was was part of the second data release (DR2) \cite[][]{gaia:16,brown:18}.
+We combine the radial velocity measurements from the VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2 to reconstruct the three-dimensional velocity of VFTS682, and test the hypothesis that this star was ejected from R136.
+
+Our results indicate that VFTS 682 is a bona fide runaway star.
+The direction an magnitude of the velocity vector is full consistent with dynamical ejection from R136.
+Its inferred age is consistent with the inferred travel time.
+This means that the hypothesis that VFTS 682 is formed in relative isolation is rejected.
+This makes  VFTS 682 the most massive runaway star known to date.
+
+}
+
+% make sure to cite this in the conclusion \cite{fujii:11, banerjee:12}
 

### 2018-05-11 VFTS682.tex
Obseravtions: section on vfts 682, added bib file

@@
-\documentclass{aa}
+\documentclass[apjl,twocolumn]{emulateapj}
+
 \usepackage{graphicx}
 \usepackage[varg]{txfonts}
-\usepackage[%draft, 
-colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
-\usepackage{amsmath}
+
+\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
+\renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
+%\usepackage[%draft, 
+%colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
+\usepackage{aas_macros}
+%\usepackage{amsmath} % clashes with emulateapjj for some reaons
 \usepackage[usenames]{xcolor}
@@
 \usepackage{multirow}
-\usepackage{chngpage}
-\usepackage{lscape}
+%\usepackage{chngpage}
+%\usepackage{lscape}
 \usepackage{url}
@@
 % for tree diagram
-\usepackage{forest}
-\usepackage{tikz-qtree}
-\usetikzlibrary{shadows,trees}
+%\usepackage{forest}
+%\usepackage{tikz-qtree}
+%\usetikzlibrary{shadows,trees}
 
@@
 \newcommand{\kms}{{\,\mathrm{km\ s^{-1}}}}
-\newcommand{\Msun}{{\mathrm{M}_\odot}}
+\newcommand{\Msun}{{\,\mathrm{M}_\odot}}
+\newcommand{\Lsun}{{\,\mathrm{L}_\odot}}
+\newcommand{\masyr}{{\,\mathrm{mas}\,\mathrm{yr}^{-1}}}
 
@@
 
-\title{VFTS682: a confirmed dynamical ejection?}
+\title{ \SdM{Gaia identifies VFTS682 as the most massive, dynamically ejected runaway star known to date}}
 
-\author{M.~Renzo\inst{1} \and S.~E.~de~Mink\inst{1} \and \SdM{D.J. Lennon et al.} \todo{etc...}%\and
+\author{M.~Renzo$^{1}$\and S.~E.~de~Mink$^{1}$ \and \SdM{D.J. Lennon$^{2}$ et al.} \todo{... add coauthors...}%\and
   % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
@@
 
-\institute{{Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands}
+\affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
+    Amsterdam, 1098 XH Amsterdam, The Netherlands
+    $^{2}$ ...
+    $^{3}$ ... 
   % \and{School of Astronomy \& Space Science, University of the Chinese
@@
 
-\offprints{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+\thanks{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
 \date{}
-\abstract{}
+\begin{abstract}
+...\\
+\SdM{Mathieu: please use emulate apj template for this.  We would be submitting there. }...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+...\\
+\end{abstract}
 
@@
 process.
-Major progress has been made on the theoretical side,  \citep[e.g.][]{Kuiper+2015Rosen+2016}, but the simulations are computationally very expensive.
+Major progress has been made on the theoretical side,  \citep[e.g.][]{Kuiper+2015,Rosen+2016}, but the simulations are computationally very expensive.
 They can typically not yet resolve the length scales that are of interest to understand the
@@
 This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of massive stars \citep{Evans+2011}.
-With an spectral type  WNh5, an inferred initial mass of $M_\mathrm{ZAMS}=150.0^{+28.7}_{-17.4}\,M_\odot$  and a mass loss rate of  $\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$  \citep[][]{bestenlehner:11,schneider:18} it belongs to the most extreme objects in the region.
-It is reminiscent of the very  massive stars, aka monster stars, identified by  \citet{Crowther+2010} and \citet{Crowther+2016}.
+It's spectral type WNh5,  inferred present-day mass of   $138^{+xx}_{-xx}\,M_\odot$
 
-Most remarkable is the location of  VFTS682.
-It is located at a projected distance of $\sim$$29$\,pc from the core of the central ionizing star cluster R136.
+ an inferred initial mass of $150^{+28.7}_{-17.4}\,M_\odot$  and a mass loss rate of  $\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$  \citep[][]{bestenlehner:11,schneider:18} it belongs to the most extreme objects in the region.
+It is reminiscent of the very  massive stars, a.k.a.
+``monster stars'', identified by  \citet{Crowther+2010} and \citet{Crowther+2016}.
+
+Most remarkable is the location of  VFTS682, at a projected distance of $\sim$$29$\,pc from the core of the central ionizing star cluster R136.
 \citet{bestenlehner:11} discuss two possibilities for the  origin of VFTS 682.
@@
 Our results indicate that VFTS 682 is a bona fide runaway star.
-The direction an magnitude of the velocity vector is full consistent with dynamical ejection from R136.
+The direction an magnitude of the velocity vector is fully consistent with dynamical ejection from R136.
 Its inferred age is consistent with the inferred travel time.
 This means that the hypothesis that VFTS 682 is formed in relative isolation is rejected.
-This makes  VFTS 682 the most massive runaway star known to date.
-
+This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.
 }
@@
 
-\section{Gaia DR2 data selection}
+\section{Observations}
 \label{sec:sample}
+\SdM{
+In this section we summarize the remarkable properties derived for VFTS682 \Secref{data:vfts683} in earlier studies and we discuss how we Gaia data in 
+\Secref{data:gaia}.
 
-VFTS682 is labeled in the Gaia DR2
-catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}} with the
-source id 4657685637907503744. The star has a
+\subsection{ Overview of the stellar parameters of VFTS682 derived in previous studies \label{data:vfts683}}
+
+The star VFTS682  \citep{Evans+2011}, located at  RA 05:38:55.510  Dec -69:04:26.72 \todo{format properly}
+was originally classified as a young stellar object \citep{Gruendl+2009} based on its mid-infrared excess.
+\citet{Evans+2011} classified the object a Wolf-Rayet star using spectroscopic observations taken a part of the VLT-FLAMES Tarantula survey.
+The spectra available covered $\lambda$4000--7000 and were taken at multiple epochs.
+\citep{Bestenlehner+2011} studied this object in more detail used and looked for radial velocity variations.
+This allowed them to exclude the presence of a close companion with high confidence.
+
+\citep{Bestenlehner+2011} further analyzed the spectra using the non-LTE model atmosphere code CMFGEN \citep{Hillier+1998} to derive the stellar parameters and surface abundances.
+They derive a find a peculiar extinction ($R_V \sim 4.7$), leading to a surprisingly high luminosity $\log_{10} L (\rm{L_\odot}) =  6.5\pm0.2$.
+Bayesian analysis with the Bonsai code \citep{Schneider+2017} against  evolutionary models by \citet{Brott+2011} and \citet{Kohler+2015} leads to an estimated present-day mass of  $138^{+28}_ {-16}\Msun$, an age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial masse of $150^{+29}_{-17}\Msun$.
+This places around the ``canonical upper limit'' of $150\Msun$  by \citet{Figer2005}.
+
+The star is very similar to R136a3, located inside the star cluster R136, \citep{Crowther+2010}  for which \citet{Crowther+2016} report a current mass estimate of $180^{30}_{-30\Msun}$.
+The R136 cluster hosts two further very massive WN5h stars R1361 and R136a2, whose estimated current masses are even higher.
+However, VFTS682 stands out by its location a projected offset of 32\,pc from the center of the cluster.
+
+Further worth noticing is the variability of the star.
+\citep{Bestenlehner+2011} discuss the Optical Gravitational Lensing Experiment (OGLE-III) light curves \citep{Udalski+2008} and show a variability in the V-band of the 10\% level on a timescale of years.
+The comment that this is unusual for Wolf-Rayet and more reminiscent of Luminous Blue Variables.
+
+\todo{Report mass loss rate.  Report LOS velocity.}
+
+%\citep{Parker1993}
+
+\subsection{Kinematic data from Gaia DR2  \label{data:gaia}}
+
+The Gaia Data Release 2, which became available on 25 April 2018, provides the five-parameter astrometric solution (positions, parallaxes, and proper motions) for more than a bilion sources \citep{Gaia-Collaboration+2018}.
+The median uncertainty in the proper motion components the corresponding uncertainties are $0.05\,\masyr$ for bright sources  ($G<14$ mag) and $0.2\masyr$ at $G=17\,\mathrm{mag}$ \citep{Lindegren+2018}.
+The uncertainties quoted above to $xx\kms$ and $yy\kms$ respectively.
+This unprecedented accuracy allows the identification of bright runaway stars at the distance to the LMC, 50\,kpc.
+
+VFTS682 is identified with the gaia source} with id 4657685637907503744 in the Gaia DR2
+catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia reports a G-band magnitude is 15.65. The star has a
 \texttt{visibility\_period} = 17, which counts how many observations have
 been used to reconstruct its astrometric solution
-\citep[][]{lindengren:18}. Its reported G-band
-magnitude is 15.65, cf. the V-band magnitude of 16.08
-\citep[][]{evans:11, bestenlehner:11}, and the reported
+\citep[][]{lindengren:18}, and the reported
 \texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
@@
 
-\begin{table}[tbp]
-  \centering
-    \caption{Astrometric parameters for VFTS682. The peculiar radial
-    velocity $\delta v_\mathrm{rad}$ is obtained as the difference
-    between the average radial velocity of the 30 Doradus region
-    ($270\pm10\kms$) minus the radial velocity measured from the HeII $\lambda4686$
-    line for VFTS682 ($315\pm15\kms$).}
-
-  \begin{tabular}[htbp]{l|c|c}
-    Parameter & Value & Source\\ \hline\hline
+\begin{table}
+\begin{center}
+    \caption{Astrometric parameters for VFTS682. }
+%
+  \begin{tabular}{l|c|c}
+  \tableline
+\tableline
+    Parameter & Value & Source\\ 
+    \tableline
     RA \hfill[degree] &  \phantom{-}84.73 $\pm$  0.03 & \multirow{4}{*}{Gaia DR2}\\
@@
     $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{0}-45 $\pm$ 25 & \cite{bestenlehner:11}\\
-    \hline
+\tableline
   \end{tabular}
+  \tablecomments{The peculiar radial
+    velocity $\delta v_\mathrm{rad}$ is obtained as the difference
+    between the average radial velocity of the 30 Doradus region
+    ($270\pm10\kms$) minus the radial velocity measured from the HeII $\lambda4686$
+    line for VFTS682 ($315\pm15\kms$).}
+    \end{center}
   \label{tab:vfts682}
@@
 
-We define a local standard frame to derive the peculiar velocity
+%% Any table notes must follow the \end{tabular} command.
+
+%%% change \tablenotetext{a} -> \tablenotetext{1} etc. to omit LaTeX errors
+%%% Karol Kozio? for ShareLaTeX
+
+We define a local standard frame \SdM{of reference} to derive the peculiar velocity
 of VFTS682 by selecting from the Gaia DR2 catalog a sample of nearby
@@
 
+\todo{Merge Selma's my\_bib bibtex file with Mathieu's and prevent double refs. }
 \bibliographystyle{aa}
-\bibliography{bibliography/vfts682}
+\bibliography{bibliography/vfts682,my_bib}
 

### 2018-05-11 VFTS682.tex
converted to apjletters

@@
 \definecolor{Wildstrawberry}{rgb}{1.0, 0.26, 0.64}
-\newcommand{\SdM}[1]{{{\color{Wildstrawberry}\bf{#1}}}}
+\newcommand{\SdM}[1]{{{\color{Wildstrawberry}{#1}}}}
 \newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}
@@
 ...\\
-\SdM{Mathieu: please use emulate apj template for this.  We would be submitting there. }...\\
 ...\\
@@
 Understanding massive star formation, and its possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies as sources of feedback, but also for understanding the transients that mark their death and the compact remnants they leave behind.
-This is therefore a key question for the present and upcoming transient surveys \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{}\todo{ref} which  will reveal transients associated to massive stars
+This is therefore a key question for the present and upcoming transient surveys e.g., LSST and LIGO/Virgo,
+%\citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others.  
+which  will reveal transients associated to massive stars
 evolution and death.

### 2018-05-11 VFTS682.tex
minor: authors, acknowledgements, color

@@
 %\usepackage{amsmath} % clashes with emulateapjj for some reaons
-\usepackage[usenames]{xcolor}
+\usepackage[usenames,dvipsnames]{xcolor}
 \usepackage{comment}
@@
 \definecolor{Wildstrawberry}{rgb}{1.0, 0.26, 0.64}
-\newcommand{\SdM}[1]{{{\color{Wildstrawberry}{#1}}}}
+\newcommand{\SdM}[1]{{{\color{Sepia}{#1}}}}
+%\renewcommand{\SdM}[1]{{{{#1}}}}
+
 \newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}
@@
 
-\author{M.~Renzo$^{1}$\and S.~E.~de~Mink$^{1}$ \and \SdM{D.J. Lennon$^{2}$ et al.} \todo{... add coauthors...}%\and
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, \SdM{D.J. Lennon$^{2}$ et al.} \todo{... add coauthors...}%\and
   % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
@@
 \affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands
-    $^{2}$ ...
+    Amsterdam, 1098 XH Amsterdam, The Netherlands \\
+    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
     $^{3}$ ... 
@@
 \begin{acknowledgements}
-  We are grateful to S.~Torres for extremely useful discussions on the
+  We are grateful to S.~Torres for 
+  %extremely 
+  very useful discussions on the
   Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
+  \SdM{SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
+Council (ERC) (Grant agreement No. 715063).
+   This work has made use of data from the ESA space mission Gaia (http://www.cosmos.esa.int/gaia), processed by the Gaia Data Processing and Analysis Consortium (DPAC, http://www.cosmos.esa.int/web/gaia/dpac/consortium).
+Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement.
+}
   The background image of \Figref{fig:main} is based on observations

### 2018-05-11 VFTS682.tex
very first abstract

@@
 \begin{abstract}
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
+\SdM{ The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
+%
+Previous work spectroscopic work identified VFTS682 as one of the most massive stars known.
+It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 32 pc of the star cluster R136.
+The lack of a detected offset in radial velocity derived from the spectra, lead to speculations whether such extreme stars can form in relative isolation.
+%
+We use Gaia's unprecedented proper motion measurements to investigate the kinematics of this object.
+We show that VFTS 682 is a bonafide runaway star moving with a relative velocity of about $60\kms$ away from the star cluster R136.
+The inferred age, distance, and the magnitude and direction of the velocity are consistent a dynamical ejection scenario from the cluster.
+%
+If correct, this would make VFTS 682 the most massive dynamically ejected runaway star known to date.
+It proves that star clusters are capable of ejecting their most massive members.
+We discuss the implications for the existence of stars with masses well above the canonical upper mass limit inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
+}
+
 \end{abstract}
@@
 
-% make sure to cite this in the conclusion \cite{fujii:11, banerjee:12}
+\todo{(1. Make sure we cite and disucss fujii:11, banerjee:12 in discussion. 2. Make link to dynamical interaction proposed to form BHBH systems. }
 

### 2018-05-11 VFTS682.tex
merger issues

--- 2018-05-11 VFTS682.tex
@@
-\documentclass[apjl,twocolumn]{emulateapj}
-
+\documentclass{aa}
 \usepackage{graphicx}
 \usepackage[varg]{txfonts}
-
-\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
-\renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
-%\usepackage[%draft, 
-%colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
-\usepackage{aas_macros}
-%\usepackage{amsmath} % clashes with emulateapjj for some reaons
+\usepackage[%draft, 
+colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
+\usepackage{amsmath}
 \usepackage[usenames]{xcolor}
@@
 \usepackage{multirow}
-%\usepackage{chngpage}
-%\usepackage{lscape}
+\usepackage{chngpage}
+\usepackage{lscape}
 \usepackage{url}
@@
 % for tree diagram
-%\usepackage{forest}
-%\usepackage{tikz-qtree}
-%\usetikzlibrary{shadows,trees}
+\usepackage{forest}
+\usepackage{tikz-qtree}
+\usetikzlibrary{shadows,trees}
 
@@
 \definecolor{Wildstrawberry}{rgb}{1.0, 0.26, 0.64}
-\newcommand{\SdM}[1]{{{\color{Wildstrawberry}{#1}}}}
+\newcommand{\SdM}[1]{{{\color{Wildstrawberry}\bf{#1}}}}
 \newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}
@@
 \newcommand{\kms}{{\,\mathrm{km\ s^{-1}}}}
-\newcommand{\Msun}{{\,\mathrm{M}_\odot}}
-\newcommand{\Lsun}{{\,\mathrm{L}_\odot}}
-\newcommand{\masyr}{{\,\mathrm{mas}\,\mathrm{yr}^{-1}}}
+\newcommand{\Msun}{{\mathrm{M}_\odot}}
 
@@
 
-\title{ \SdM{Gaia identifies VFTS682 as the most massive, dynamically ejected runaway star known to date}}
+\title{VFTS682: a confirmed dynamical ejection?}
 
-\author{M.~Renzo$^{1}$\and S.~E.~de~Mink$^{1}$ \and \SdM{D.J. Lennon$^{2}$ et al.} \todo{... add coauthors...}%\and
+\author{M.~Renzo\inst{1} \and S.~E.~de~Mink\inst{1} \and \SdM{D.J. Lennon et al.} \todo{etc...}%\and
   % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
@@
 
-\affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands
-    $^{2}$ ...
-    $^{3}$ ... 
+\institute{{Astronomical Institute Anton Pannekoek, University of
+    Amsterdam, 1098 XH Amsterdam, The Netherlands}
   % \and{School of Astronomy \& Space Science, University of the Chinese
@@
 
-\thanks{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+\offprints{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
 \date{}
-\begin{abstract}
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-...\\
-\end{abstract}
+\abstract{}
 
@@
 process.
-Major progress has been made on the theoretical side,  \citep[e.g.][]{Kuiper+2015,Rosen+2016}, but the simulations are computationally very expensive.
+Major progress has been made on the theoretical side,  \citep[e.g.][]{Kuiper+2015Rosen+2016}, but the simulations are computationally very expensive.
 They can typically not yet resolve the length scales that are of interest to understand the
@@
 Understanding massive star formation, and its possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies as sources of feedback, but also for understanding the transients that mark their death and the compact remnants they leave behind.
-This is therefore a key question for the present and upcoming transient surveys e.g., LSST and LIGO/Virgo,
-%\citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others.  
-which  will reveal transients associated to massive stars
+This is therefore a key question for the present and upcoming transient surveys \citep[e.g., LSST, BlackGem, LIGO/Virgo O3][]{}\todo{ref} which  will reveal transients associated to massive stars
 evolution and death.
@@
 This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of massive stars \citep{Evans+2011}.
-It's spectral type WNh5,  inferred present-day mass of   $138^{+xx}_{-xx}\,M_\odot$
+With an spectral type  WNh5, an inferred initial mass of $M_\mathrm{ZAMS}=150.0^{+28.7}_{-17.4}\,M_\odot$  and a mass loss rate of  $\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$  \citep[][]{bestenlehner:11,schneider:18} it belongs to the most extreme objects in the region.
+It is reminiscent of the very  massive stars, aka monster stars, identified by  \citet{Crowther+2010} and \citet{Crowther+2016}.
 
- an inferred initial mass of $150^{+28.7}_{-17.4}\,M_\odot$  and a mass loss rate of  $\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$  \citep[][]{bestenlehner:11,schneider:18} it belongs to the most extreme objects in the region.
-It is reminiscent of the very  massive stars, a.k.a.
-``monster stars'', identified by  \citet{Crowther+2010} and \citet{Crowther+2016}.
-
-Most remarkable is the location of  VFTS682, at a projected distance of $\sim$$29$\,pc from the core of the central ionizing star cluster R136.
+Most remarkable is the location of  VFTS682.
+It is located at a projected distance of $\sim$$29$\,pc from the core of the central ionizing star cluster R136.
 \citet{bestenlehner:11} discuss two possibilities for the  origin of VFTS 682.
@@
 Our results indicate that VFTS 682 is a bona fide runaway star.
-The direction an magnitude of the velocity vector is fully consistent with dynamical ejection from R136.
+The direction an magnitude of the velocity vector is full consistent with dynamical ejection from R136.
 Its inferred age is consistent with the inferred travel time.
 This means that the hypothesis that VFTS 682 is formed in relative isolation is rejected.
-This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.
+This makes  VFTS 682 the most massive runaway star known to date.
+
 }
@@
 % make sure to cite this in the conclusion \cite{fujii:11, banerjee:12}
+\todo{missing: R136 too young for CC, cite age $\leq2$\,Myr \citep[][]{crowther:10,sabbi:12}
 
-\section{Observations}
+\section{Gaia DR2 data selection}
 \label{sec:sample}
-\SdM{
-In this section we summarize the remarkable properties derived for VFTS682 \Secref{data:vfts683} in earlier studies and we discuss how we Gaia data in 
-\Secref{data:gaia}.
 
-\subsection{ Overview of the stellar parameters of VFTS682 derived in previous studies \label{data:vfts683}}
-
-The star VFTS682  \citep{Evans+2011}, located at  RA 05:38:55.510  Dec -69:04:26.72 \todo{format properly}
-was originally classified as a young stellar object \citep{Gruendl+2009} based on its mid-infrared excess.
-\citet{Evans+2011} classified the object a Wolf-Rayet star using spectroscopic observations taken a part of the VLT-FLAMES Tarantula survey.
-The spectra available covered $\lambda$4000--7000 and were taken at multiple epochs.
-\citep{Bestenlehner+2011} studied this object in more detail used and looked for radial velocity variations.
-This allowed them to exclude the presence of a close companion with high confidence.
-
-\citep{Bestenlehner+2011} further analyzed the spectra using the non-LTE model atmosphere code CMFGEN \citep{Hillier+1998} to derive the stellar parameters and surface abundances.
-They derive a find a peculiar extinction ($R_V \sim 4.7$), leading to a surprisingly high luminosity $\log_{10} L (\rm{L_\odot}) =  6.5\pm0.2$.
-Bayesian analysis with the Bonsai code \citep{Schneider+2017} against  evolutionary models by \citet{Brott+2011} and \citet{Kohler+2015} leads to an estimated present-day mass of  $138^{+28}_ {-16}\Msun$, an age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial masse of $150^{+29}_{-17}\Msun$.
-This places around the ``canonical upper limit'' of $150\Msun$  by \citet{Figer2005}.
-
-The star is very similar to R136a3, located inside the star cluster R136, \citep{Crowther+2010}  for which \citet{Crowther+2016} report a current mass estimate of $180^{30}_{-30\Msun}$.
-The R136 cluster hosts two further very massive WN5h stars R1361 and R136a2, whose estimated current masses are even higher.
-However, VFTS682 stands out by its location a projected offset of 32\,pc from the center of the cluster.
-
-Further worth noticing is the variability of the star.
-\citep{Bestenlehner+2011} discuss the Optical Gravitational Lensing Experiment (OGLE-III) light curves \citep{Udalski+2008} and show a variability in the V-band of the 10\% level on a timescale of years.
-The comment that this is unusual for Wolf-Rayet and more reminiscent of Luminous Blue Variables.
-
-\todo{Report mass loss rate.  Report LOS velocity.}
-
-%\citep{Parker1993}
-
-\subsection{Kinematic data from Gaia DR2  \label{data:gaia}}
-
-The Gaia Data Release 2, which became available on 25 April 2018, provides the five-parameter astrometric solution (positions, parallaxes, and proper motions) for more than a bilion sources \citep{Gaia-Collaboration+2018}.
-The median uncertainty in the proper motion components the corresponding uncertainties are $0.05\,\masyr$ for bright sources  ($G<14$ mag) and $0.2\masyr$ at $G=17\,\mathrm{mag}$ \citep{Lindegren+2018}.
-The uncertainties quoted above to $xx\kms$ and $yy\kms$ respectively.
-This unprecedented accuracy allows the identification of bright runaway stars at the distance to the LMC, 50\,kpc.
-
-VFTS682 is identified with the gaia source} with id 4657685637907503744 in the Gaia DR2
-catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia reports a G-band magnitude is 15.65. The star has a
+VFTS682 is labeled in the Gaia DR2
+catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}} with the
+source id 4657685637907503744. The star has a
 \texttt{visibility\_period} = 17, which counts how many observations have
 been used to reconstruct its astrometric solution
-\citep[][]{lindengren:18}, and the reported
+\citep[][]{lindengren:18}. Its reported G-band
+magnitude is 15.65, cf. the V-band magnitude of 16.08
+\citep[][]{evans:11, bestenlehner:11}, and the reported
 \texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
@@
 
-\begin{table}
-\begin{center}
-    \caption{Astrometric parameters for VFTS682. }
-%
-  \begin{tabular}{l|c|c}
-  \tableline
-\tableline
-    Parameter & Value & Source\\ 
-    \tableline
+\begin{table}[tbp]
+  \centering
+    \caption{Astrometric parameters for VFTS682. The peculiar radial
+    velocity $\delta v_\mathrm{rad}$ is obtained as the difference
+    between the average radial velocity of the 30 Doradus region
+    ($270\pm10\kms$) minus the radial velocity measured from the HeII $\lambda4686$
+    line for VFTS682 ($315\pm15\kms$).}
+
+  \begin{tabular}[htbp]{l|c|c}
+    Parameter & Value & Source\\ \hline\hline
     RA \hfill[degree] &  \phantom{-}84.73 $\pm$  0.03 & \multirow{4}{*}{Gaia DR2}\\
@@
     $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{0}-45 $\pm$ 25 & \cite{bestenlehner:11}\\
-\tableline
+    \hline
   \end{tabular}
-  \tablecomments{The peculiar radial
-    velocity $\delta v_\mathrm{rad}$ is obtained as the difference
-    between the average radial velocity of the 30 Doradus region
-    ($270\pm10\kms$) minus the radial velocity measured from the HeII $\lambda4686$
-    line for VFTS682 ($315\pm15\kms$).}
-    \end{center}
   \label{tab:vfts682}
@@
 
-%% Any table notes must follow the \end{tabular} command.
-
-%%% change \tablenotetext{a} -> \tablenotetext{1} etc. to omit LaTeX errors
-%%% Karol Kozio? for ShareLaTeX
-
-We define a local standard frame \SdM{of reference} to derive the peculiar velocity
+We define a local standard frame to derive the peculiar velocity
 of VFTS682 by selecting from the Gaia DR2 catalog a sample of nearby
@@
   \includegraphics[width=0.5\textwidth]{./figures/main_plot}  
-  \caption{The red cross indicates the position of VFTS682. Blue stars
-    are those we use to define the generic ``surroundings'', while the
-    green dots indicate the stars we use to probe the core of R136,
-    magnified in the inset. The red arrow shows the direction of the
-    proper motion of VFTS682 relative to R136, and the gray shade
-    indicates the uncertainty in the direction. \todo{load color picture on background}}
+  \caption{The red solid arrow indicates the proper motion of VFTS682
+    relative to the region, starting from the present day position of
+    the star. The semi-transparent arrows indicate the possible
+    directions of projected motion within the errors, and are extended
+    backwards to illustrate the uncertainty on the origin of the
+    star. Blue stars
+    are a subset of those we use to define the local rest frame. \todo{load color picture on background}}
   \label{fig:main}
@@
   what is its total mass and relaxation time?
+\item another important thing missing: with vsini(i)<200, completely average, there is no reason to think it's a CHE, contrary to the way they estimate the age in Bestenlehner:11
 \end{itemize}
 
-\todo{Merge Selma's my\_bib bibtex file with Mathieu's and prevent double refs. }
 \bibliographystyle{aa}
-\bibliography{bibliography/vfts682,my_bib}
+\bibliography{bibliography/vfts682}
 

### 2018-05-11 VFTS682.tex
major revision

--- 2018-05-11 VFTS682.tex
@@
 \documentclass[apjl,twocolumn]{emulateapj}
-
 \usepackage{graphicx}
 \usepackage[varg]{txfonts}
-
 \usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
@@
 
-% \defcitealias{tauris:98}{TT98}
-% \defcitealias{ramirez-agudelo:15}{R-A15}
-
 \interfootnotelinepenalty=10000    % brute-forces the footnote not to break over two pages
@@
 
-\title{ \SdM{Gaia identifies VFTS682 as the most massive, dynamically ejected runaway star known to date}}
+\title{Gaia identifies VFTS682 as the most massive, dynamically ejected runaway star known to date}
 
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, \SdM{D.J. Lennon$^{2}$ et al.} \todo{... add coauthors...}%\and
-  % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
-  .}
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
+  S.~Justham$^{3,4}$, J.~Vink$^{5}$, A.~de~Koter$^{1}$ \todo{... add coauthors...} }
 
@@
     $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
-    $^{3}$ ... 
+    $^{3}$ 
   % \and{School of Astronomy \& Space Science, University of the Chinese
@@
 \begin{abstract}
-\SdM{ The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
+The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
 %
-Previous work spectroscopic work identified VFTS682 as one of the most massive stars known.
-It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 32 pc of the star cluster R136.
+Previous spectroscopic studies identified VFTS682 as one of the most massive stars known.
+It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 29\,pc of the star cluster R136.
 The lack of a detected offset in radial velocity derived from the spectra, lead to speculations whether such extreme stars can form in relative isolation.
@@
 We discuss the implications for the existence of stars with masses well above the canonical upper mass limit inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
-}
 
@@
 
-\SdM{
-
 How do massive stars form is one of the major longstanding questions in astrophysics
@@
 evolve fast, typically reside in dense groups and they remain enshrouded in their parent cloud during the formation
-process.
-Major progress has been made on the theoretical side,  \citep[e.g.][]{Kuiper+2015,Rosen+2016}, but the simulations are computationally very expensive.
-They can typically not yet resolve the length scales that are of interest to understand the
+process.  Major progress has been made on the theoretical side,
+\citep[e.g.][]{kuiper:15,rosen:16}, but simulations cannot yet resolve
+the smallest length scales that are of interest to understand the 
 high level of multiplicity among massive stars  \citep[][]{sana:12,sana:17}.
@@
 
-In has been conjectured that most, if not all, stars form in clusters \cite{lada:03}, where massive stars are thought to  form preferentially in the dense cores of clusters.
+In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to  form preferentially in the dense cores of clusters.
 In this picture, field stars are primarily the result of the dissolution of dense groups.
- However, a significant population of massive stars has been  in relative isolation,  far from dense clusters or OB associations.
-The origin of this population still remains debated \citep{Lamb+2016}.
-One hypothesis is that they  formed in the field, but this poses a challenge for the theories of star formation.
-The alternative hypothesis is that these massive stars  formed in clusters, and then been ejected from their birth locations.
-Such ejections may result from   \citep[e,g,][]{poveda:67} or from the dsiruption of binary systems by the supernova of the companion star  \citep[][]{zwicky:57, blaauw:61}.
+However, a significant population of massive stars exists in relative
+isolation,  far from dense clusters or OB associations. The origin of
+this population still remains debated \citep{lamb:16}.   One
+hypothesis is that they  formed in the field, but this poses a
+challenge for the theories of star formation.  The alternative
+hypothesis is that these massive stars  formed in clusters, and then
+been ejected from their birth locations.  Such ejections may result
+from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems by the core-collapse of the companion star \citep[e.g.,][]{zwicky:57, blaauw:61}.
 
- A very interesting contribution to the debate whether or not massive stars an form in relative isolation was presented in recent years by  \cite{bestenlehner:11}, who discuss the case of the very massive star VFTS682.
-This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of massive stars \citep{Evans+2011}.
-It's spectral type WNh5,  inferred present-day mass of   $138^{+xx}_{-xx}\,M_\odot$
+An interesting contribution to the debate on whether or not massive
+stars an form in relative isolation was presented by
+\cite{bestenlehner:11}, who discuss the case of the very massive star
+VFTS682.   This star is located in the field of the 30 Doradus region
+in the Large Magellanic Cloud (LMC) and was studied as part of the
+multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of
+massive stars \citep{evans:11}.  Its spectral type is WNh5 and its 
+present-day mass is $137.8^{+27.5}_{-15.9}\,M_\odot$,
+\citep{schneider:18}. These correspond to an inferred initial mass of
+$150.0^{+28.7}_{-17.4}\,M_\odot$ and a mass loss rate of
+$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$
+\citep[][]{bestenlehner:11,schneider:18}, which make it one of the
+most extreme objects in the region.  It is reminiscent of the very
+massive stars, a.k.a. ``monster stars'', identified by
+\citet{crowther:10, crowther:16} in the core of the
+R136 cluster.
 
- an inferred initial mass of $150^{+28.7}_{-17.4}\,M_\odot$  and a mass loss rate of  $\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$  \citep[][]{bestenlehner:11,schneider:18} it belongs to the most extreme objects in the region.
-It is reminiscent of the very  massive stars, a.k.a.
-``monster stars'', identified by  \citet{Crowther+2010} and \citet{Crowther+2016}.
-
-Most remarkable is the location of  VFTS682, at a projected distance of $\sim$$29$\,pc from the core of the central ionizing star cluster R136.
-\citet{bestenlehner:11} discuss two possibilities for the  origin of VFTS 682.
+Most remarkable is the location of VFTS682, at a projected distance of
+$\sim$$29$\,pc from the core of the star cluster
+R136. \citet{bestenlehner:11} discuss two possibilities for the
+origin of VFTS 682.
 Quoting directly from their abstract: "either (i) the star either formed in situ, which would have profound implications for the formation mechanism of massive stars, or (ii) VFTS 682 is a slow runaway star that originated from the dense cluster R136, which would make it the most massive runaway known to date."
 
-In this study we report on our analysis of Gaia data that was was part of the second data release (DR2) \cite[][]{gaia:16,brown:18}.
+In this study we report on our analysis the data from the second Gaia data release (DR2) \cite[][]{gaia:16,brown:18}.
 We combine the radial velocity measurements from the VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2 to reconstruct the three-dimensional velocity of VFTS682, and test the hypothesis that this star was ejected from R136.
@@
 This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.
-}
-
-\todo{(1. Make sure we cite and disucss fujii:11, banerjee:12 in discussion. 2. Make link to dynamical interaction proposed to form BHBH systems. }
 
@@
 \label{sec:sample}
-\SdM{
-In this section we summarize the remarkable properties derived for VFTS682 \Secref{data:vfts683} in earlier studies and we discuss how we Gaia data in 
-\Secref{data:gaia}.
 
-\subsection{ Overview of the stellar parameters of VFTS682 derived in previous studies \label{data:vfts683}}
+In this section we summarize the remarkable properties derived for VFTS682 \todo{?}\Secref{data:vfts683}
+in earlier studies and we discuss we Gaia data in \Secref{data:gaia}.
 
-The star VFTS682  \citep{Evans+2011}, located at  RA 05:38:55.510  Dec -69:04:26.72 \todo{format properly}
-was originally classified as a young stellar object \citep{Gruendl+2009} based on its mid-infrared excess.
-\citet{Evans+2011} classified the object a Wolf-Rayet star using spectroscopic observations taken a part of the VLT-FLAMES Tarantula survey.
-The spectra available covered $\lambda$4000--7000 and were taken at multiple epochs.
-\citep{Bestenlehner+2011} studied this object in more detail used and looked for radial velocity variations.
-This allowed them to exclude the presence of a close companion with high confidence.
+\subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
 
-\citep{Bestenlehner+2011} further analyzed the spectra using the non-LTE model atmosphere code CMFGEN \citep{Hillier+1998} to derive the stellar parameters and surface abundances.
-They derive a find a peculiar extinction ($R_V \sim 4.7$), leading to a surprisingly high luminosity $\log_{10} L (\rm{L_\odot}) =  6.5\pm0.2$.
-Bayesian analysis with the Bonsai code \citep{Schneider+2017} against  evolutionary models by \citet{Brott+2011} and \citet{Kohler+2015} leads to an estimated present-day mass of  $138^{+28}_ {-16}\Msun$, an age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial masse of $150^{+29}_{-17}\Msun$.
-This places around the ``canonical upper limit'' of $150\Msun$  by \citet{Figer2005}.
+The star VFTS682  \citep{evans:11}, located at right ascension (RA)
+05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
+(DEC) -69:04:26.72 in the ICGS frame \citep[][]{brown:18}
+was originally classified as a young stellar object \citep{gruendl:09}
+based on its mid-infrared excess.  \citet{evans:11} classified the
+object as Wolf-Rayet star of spectral type WNh5 using spectra from the
+VFTS survey.  The spectra available covered $\lambda$4000--7000 and
+were taken at multiple epochs. Using the same dataset,
+\citet{bestenlehner:11} excluded the presence of a close companion
+with high confidence from the absence of radial velocity variations.
 
-The star is very similar to R136a3, located inside the star cluster R136, \citep{Crowther+2010}  for which \citet{Crowther+2016} report a current mass estimate of $180^{30}_{-30\Msun}$.
-The R136 cluster hosts two further very massive WN5h stars R1361 and R136a2, whose estimated current masses are even higher.
-However, VFTS682 stands out by its location a projected offset of 32\,pc from the center of the cluster.
+\citet{bestenlehner:11} further analyzed the spectra using the non-LTE
+model atmosphere code CMFGEN \citep{hillier:98} to derive the stellar
+parameters and surface abundances. They inferred a peculiar extinction
+($R_V\simeq4.7$), leading to a surprisingly high luminosity
+$\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citep{schneider:18} estimated
+the present-day mass of VFTS682 at  $137.8^{+27.5}_ {-15.9}\Msun$, an
+apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
+of $150.0^{+28.7}_{-17.4}\Msun$, using the Bayesian analysis tool Bonsai
+\citep{schneider:17}, based on the evolutionary models from
+\citet{brott:11, kohler:15}. These values place VFTS682 around the ``canonical upper limit'' of $\sim$$150\Msun$  by \citet{figer:05}.
+
+The star is very similar from the spectral point of view to R136a3, located inside the star cluster
+R136, \citep{crowther:10}  for which \citet{crowther:16} report a
+current mass estimate of $180^{+30}_{-30}\Msun$. The R136 cluster hosts
+at least two more very massive WN5h stars R136a1, R136a2 whose
+estimated current masses are even higher. However, VFTS682 stands
+out by its isolation at a projected offset of 29\,pc from the center
+of the cluster \citep{bestenlehner:11}.
 
 Further worth noticing is the variability of the star.
-\citep{Bestenlehner+2011} discuss the Optical Gravitational Lensing Experiment (OGLE-III) light curves \citep{Udalski+2008} and show a variability in the V-band of the 10\% level on a timescale of years.
-The comment that this is unusual for Wolf-Rayet and more reminiscent of Luminous Blue Variables.
+\citep{bestenlehner:11} discuss the Optical Gravitational Lensing
+Experiment (OGLE-III) light curves \citep{udalski:08} and show a
+variability in the V-band of the $\sim$10\% level on a timescale of years.
+The comment that this is unusual for Wolf-Rayet and more reminiscent
+of Luminous Blue Variable (LBV) stars.
 
 \todo{Report mass loss rate.  Report LOS velocity.}
+\citep{parker:93}
 
-%\citep{Parker1993}
+\subsection{New data from Gaia DR2  \label{data:gaia}}
 
-\subsection{Kinematic data from Gaia DR2  \label{data:gaia}}
-
-The Gaia Data Release 2, which became available on 25 April 2018, provides the five-parameter astrometric solution (positions, parallaxes, and proper motions) for more than a bilion sources \citep{Gaia-Collaboration+2018}.
-The median uncertainty in the proper motion components the corresponding uncertainties are $0.05\,\masyr$ for bright sources  ($G<14$ mag) and $0.2\masyr$ at $G=17\,\mathrm{mag}$ \citep{Lindegren+2018}.
-The uncertainties quoted above to $xx\kms$ and $yy\kms$ respectively.
-This unprecedented accuracy allows the identification of bright runaway stars at the distance to the LMC, 50\,kpc.
-
-VFTS682 is identified with the gaia source} with id 4657685637907503744 in the Gaia DR2
-catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia reports a G-band magnitude is 15.65. The star has a
-\texttt{visibility\_period} = 17, which counts how many observations have
-been used to reconstruct its astrometric solution
-\citep[][]{lindengren:18}, and the reported
-\texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
-data for VFTS682 are trustworthy. However, the effective temperature
-reported in Gaia DR2 is one order of magnitude lower than what found by
-\cite{bestenlehner:11}, and the best fit parallax of this star is
-negative. We do not use the effective temperature of the star anywhere
-in this study, and we attribute the unphysical value of the parallax
-to the large distance to the LMC. Our main findings do not rely on the
-parallax nor the effective temperature values reported in the Gaia DR2
-catalog.
-
-%% should I really mention the line of sight velocity? Not really
-%% using it...
-
-We retrieve for VFTS682 the position in right ascension (RA) and declination (DEC)
-in the IGCS frame \cite[][]{brown:18}, its
-proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
-respectively). For the radial velocity of VFTS682 and of the 30 Doradus
-region as a whole, we instead use the VFTS data
-as quoted in \cite{bestenlehner:11}. \Tabref{tab:vfts682} lists the values adopted throughout
-this work for each of these quantities.
-
-\begin{table}
+\begin{table}[t]
 \begin{center}
@@
 
-%% Any table notes must follow the \end{tabular} command.
+% %% Any table notes must follow the \end{tabular} command.
 
-%%% change \tablenotetext{a} -> \tablenotetext{1} etc. to omit LaTeX errors
-%%% Karol Kozio? for ShareLaTeX
+% %%% change \tablenotetext{a} -> \tablenotetext{1} etc. to omit LaTeX errors
+% %%% Karol Kozio? for ShareLaTeX
 
-We define a local standard frame \SdM{of reference} to derive the peculiar velocity
+The Gaia Data Release 2, which became available on 25 April 2018,
+provides the five-parameter astrometric solution (positions,
+parallaxes, and proper motion components) for more than a billion
+sources \citep{brown:18}.%  \todo{why do we care about the
+%   median? I'm
+%   selecting data with errors smaller than this, and I'm not using
+%   fainter stars!I think this is not useful here} The median uncertainty in the proper motion
+% components the corresponding uncertainties are $0.05\,\masyr$ for
+% bright sources  ($G<14$ mag) and $0.2\masyr$ at $G=17\,\mathrm{mag}$
+% \citep{lindegren:18}.
+The uncertainties quoted above correspond to $xx\kms$ and $yy\kms$ respectively.
+This unprecedented accuracy allows the identification of bright runaway stars at the distance to the LMC, 50\,kpc.
+
+VFTS682 is identified with the id 4657685637907503744 in the Gaia DR2
+catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia
+reports a G-band magnitude of 15.65. The star has a
+\texttt{visibility\_period} = 17, which counts how many observations have
+been used to reconstruct its astrometric solution
+\citep[][]{lindengren:18}, and the reported
+\texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
+data for VFTS682 are trustworthy. However, the effective temperature
+reported in Gaia DR2 is one order of magnitude lower than what found by
+\cite{bestenlehner:11}, and the best fit parallax of this star is
+negative. We do not use the effective temperature of the star anywhere
+in this study, and we attribute the unphysical value of the parallax
+to the large distance to the LMC. Our main findings do not rely on the
+parallax nor the effective temperature values reported in the Gaia DR2
+catalog.
+
+We retrieve for VFTS682 the position in RA and DEC
+in the IGCS frame \cite[][]{brown:18}, its
+proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
+respectively). For the radial velocity of VFTS682 and of the 30 Doradus
+region as a whole, we instead use the VFTS data
+as quoted in \cite{bestenlehner:11}. \Tabref{tab:vfts682} lists the values adopted throughout
+this work for each of these quantities.
+
+We define a local standard frame of reference to derive the peculiar velocity
 of VFTS682 by selecting from the Gaia DR2 catalog a sample of nearby
@@
 G$\sim$V,][]{evans:11}. Then we require \texttt{visibility\_period} $\geq$ 5,
-\texttt{astrometric\_excess\_noise} < 1, the errors on the proper
+\texttt{astrometric\_excess\_noise} $< 1$, the errors on the proper
 motion components to be smaller than 0.1\,$\mathrm{mas\ yr^{-1}}$,
@@
 
-Using the 651 stars selected as described in \Secref{sec:sample} (a
+Using the 651 stars selected as described in \Secref{data:gaia} (a
 subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
@@
 motion. It is clear that the most likely origin of the star is R136,
-as predicted by \cite{fujii:11, banerjee:12}.
+as predicted from numerical N-body simulations by \cite{fujii:11, banerjee:12}.
 
@@
   \label{eq:kin_age}
-  \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} =
+  \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} \simeq
   \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.63\, \mathrm{Myr} \ \ ,
@@
 The kinematic age $\tau_\mathrm{kin}$ is smaller than the apparent age
-of the star age $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
+of the star age of $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
 corroborates the idea that the star is the result of a dynamical
@@
   \includegraphics[width=0.5\textwidth]{./figures/main_plot}  
-  \caption{The red cross indicates the position of VFTS682. Blue stars
-    are those we use to define the generic ``surroundings'', while the
-    green dots indicate the stars we use to probe the core of R136,
-    magnified in the inset. The red arrow shows the direction of the
-    proper motion of VFTS682 relative to R136, and the gray shade
-    indicates the uncertainty in the direction. \todo{load color picture on background}}
+  \caption{The red solid arrow indicates the proper motion of VFTS682
+    relative to the region, starting from the present day position of
+    the star. The semi-transparent arrows indicate the possible
+    directions of projected motion within the errors, and are extended
+    backwards to illustrate the uncertainty on the origin of the
+    star. The blue stars
+    are a subset of those we use to define the local rest frame. \todo{load color picture on background}}
   \label{fig:main}
@@
 Based on our results, we can claim that VFTS682 is the most massive
-runaway known to date, with a peculiar three-dimensional spatial
-velocity of $\sim$$60\,\kms$. This means that isolated star formation is
+runaway known to date, with a peculiar three-dimensional speed of $64\pm21\,\kms$. This means that isolated star formation is
 \emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
@@
 line with the average rotation rate of massive stars in the region
-\citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 has not
-experienced binary interactions, nor it will, since the multi-epoch
+\citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
+experienced rotationally induced chemically homogeneous evolution
+\citep[][]{maeder:00,demink:09,marchant:16}, and (ii) it has not
+accreted mass from a binary companion, nor it will, since the multi-epoch
 data of the VFTS survey rule out the presence of a companion at
@@
 three most massive stars detected in the core of the cluster by
-\cite{crowther:10}. Therefore,
+\cite{crowther:10}, with an astonishing similarity in particular with
+the spectrum of R136a3. Therefore,
 VFTS682 might be an ideal target to constrain the stellar physics of
-stars with masses well above $\sim100\,M_\odot$: its isolation makes
-it an easier target for observations compared to the similar stars present the crowded core of R136.
+stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
+it an easier target for observations compared to the similar stars
+present in the crowded core of R136.
 
@@
 evolution, given the age estimate of $\lesssim 2$\,Myr
-\cite{crowther:10,sabbi:12} and the kinematic age of VFTS16.
+\cite{crowther:10,sabbi:12} and the kinematic age of VFTS16. If the
+cluster is indeed younger than the shortest stellar lifetime
+\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, ejection from a binary
+system of VFTS682 is excluded since the region is to young for stars
+to have experience core-collapse already.
 
@@
   what is its total mass and relaxation time?
+\item Make link to dynamical interaction proposed to form BHBH systems.
 \end{itemize}
 
-\todo{Merge Selma's my\_bib bibtex file with Mathieu's and prevent double refs. }
+% \todo{Merge Selma's my\_bib bibtex file with Mathieu's and prevent
+% double refs. } %% copied everything in one file
+
 \bibliographystyle{aa}
-\bibliography{bibliography/vfts682,my_bib}
+\bibliography{bibliography/vfts682}
 
@@
   We are grateful to S.~Torres for 
-  %extremely 
   very useful discussions on the
   Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
-  \SdM{SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
+  SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
   Council (ERC) (Grant agreement No. 715063).
@@
 Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement.
-}
   The background image of \Figref{fig:main} is based on observations
@@
 \end{acknowledgements}
+
 \end{document}

### 2018-05-11 VFTS682.tex
incorporated Fabian's comments

@@
 
-\title{Gaia identifies VFTS682 as the most massive, dynamically ejected runaway star known to date}
+\title{Gaia identifies VFTS682 as the most massive % , dynamically
+                                % ejected %% to fit in one line
+  runaway star known to date}
 
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  S.~Justham$^{3,4}$, J.~Vink$^{5}$, A.~de~Koter$^{1}$ \todo{... add coauthors...} }
+  S.~Justham$^{3,4}$, J.~Vink$^{5}$, F.~R.~Schneider$^6$, A.~de~Koter$^{1}$ \todo{... add coauthors...} }
 
@@
     $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
-    $^{3}$ 
-  % \and{School of Astronomy \& Space Science, University of the Chinese
-  %   Academy of Sciences, Beijing 100012, China}
-  % \and{National Astronomical Observatories, Chinese Academy of
-  %   Sciences, Beijing 100012, China}
+    $^{3}$ {School of Astronomy \& Space Science, University of the Chinese
+    Academy of Sciences, Beijing 100012, China}\\
+     $^4$ {National Astronomical Observatories, Chinese Academy of
+       Sciences, Beijing 100012, China}\\
+     $^5$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
+     $^6$ {Department of Physics, University of Oxford, Keble Road, Oxford OX1 3RH, UK}  
 }
@@
 %
-We use Gaia's unprecedented proper motion measurements to investigate the kinematics of this object.
-We show that VFTS 682 is a bonafide runaway star moving with a relative velocity of about $60\kms$ away from the star cluster R136.
-The inferred age, distance, and the magnitude and direction of the velocity are consistent a dynamical ejection scenario from the cluster.
+We use the unprecedented proper motions from the second Gaia data
+release (DR2) to investigate the kinematics of this object. We show
+that VFTS 682 is a bonafide runaway star moving with a relative
+velocity of about $64\pm21\kms$ away from the star cluster R136.  The
+inferred age, distance, and the magnitude and direction of the
+velocity are consistent with a dynamical ejection scenario from the cluster. 
 %
-If correct, this would make VFTS 682 the most massive dynamically ejected runaway star known to date.
+This would make VFTS 682 the most massive runaway star known to date.
 It proves that star clusters are capable of ejecting their most massive members.
-We discuss the implications for the existence of stars with masses well above the canonical upper mass limit inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
+We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
 
@@
 the smallest length scales that are of interest to understand the 
-high level of multiplicity among massive stars  \citep[][]{sana:12,sana:17}.
-Understanding massive star formation, and its possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies as sources of feedback, but also for understanding the transients that mark their death and the compact remnants they leave behind.
-This is therefore a key question for the present and upcoming transient surveys e.g., LSST and LIGO/Virgo,
+high level of multiplicity among massive stars  \citep[e.g.,][]{bate:09, sana:17}.  Understanding massive star formation, and its
+possible dependence on environment and metallicity, is crucial for
+understanding the role massive stars play within their host galaxies
+as sources of feedback, but also for understanding the transients that
+mark their death and the compact remnants they leave behind.  This is
+therefore a key question for the present and upcoming transient
+surveys e.g., Large Synoptic Spectroscopic Survey and LIGO/Virgo,
 %\citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others.  
@@
 isolation,  far from dense clusters or OB associations. The origin of
-this population still remains debated \citep{lamb:16}.   One
+this population still remains debated \citep{lamb:16,ward:18}.   One
 hypothesis is that they  formed in the field, but this poses a
@@
 stars an form in relative isolation was presented by
-\cite{bestenlehner:11}, who discuss the case of the very massive star
+\cite{bestenlehner:11, bressert:12}, who discuss the case of the very massive star
 VFTS682.   This star is located in the field of the 30 Doradus region
@@
 \citep{schneider:18}. These correspond to an inferred initial mass of
-$150.0^{+28.7}_{-17.4}\,M_\odot$ and a mass loss rate of
-$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$
-\citep[][]{bestenlehner:11,schneider:18}, which make it one of the
+$150.0^{+28.7}_{-17.4}\,M_\odot$ and a wind mass loss rate of
+$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$ inferred from the
+spectral analysis not accounting for clumping
+\citep[][]{bestenlehner:11}, which make it one of the
 most extreme objects in the region.  It is reminiscent of the very
-massive stars, a.k.a. ``monster stars'', identified by
+massive stars % , a.k.a. ``monster stars'',
+identified by
 \citet{crowther:10, crowther:16} in the core of the
@@
 Our results indicate that VFTS 682 is a bona fide runaway star.
-The direction an magnitude of the velocity vector is fully consistent with dynamical ejection from R136.
+The direction and magnitude of the velocity vector is fully consistent with dynamical ejection from R136.
 Its inferred age is consistent with the inferred travel time.
@@
 
-In this section we summarize the remarkable properties derived for VFTS682 \todo{?}\Secref{data:vfts683}
-in earlier studies and we discuss we Gaia data in \Secref{data:gaia}.
+% feels like unnecessary here...
+% In this section we summarize the remarkable properties derived for VFTS682 \todo{?}\Secref{data:vfts683}
+% in earlier studies and we discuss we Gaia data in \Secref{data:gaia}.
 
@@
 apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
-of $150.0^{+28.7}_{-17.4}\Msun$, using the Bayesian analysis tool Bonsai
+of $150.0^{+28.7}_{-17.4}\Msun$, using the Bayesian analysis tool Bonnsai
 \citep{schneider:17}, based on the evolutionary models from
-\citet{brott:11, kohler:15}. These values place VFTS682 around the ``canonical upper limit'' of $\sim$$150\Msun$  by \citet{figer:05}.
+\citet{brott:11, kohler:15}. These values place VFTS682 around the
+``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
 
@@
 \texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
-data for VFTS682 are trustworthy. However, the effective temperature
+data for VFTS682 are reliable. However, the effective temperature
 reported in Gaia DR2 is one order of magnitude lower than what found by
@@
 component separately. For simplicity, throughout this study, we assume the same
-distance of $50$\,kpc to the star \citep[][]{lebouteiller:08}, and to
+distance of $50$\,kpc to the star \citep[][]{pietrzynski:13}, and to
 the 30 Doradus region as a whole. We do not consider the error bars on
@@
 0.672\pm0.003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
-in good agreement with what found by \cite{lennon:18}. Subtracting these values from the
+in good agreement with what is found by \cite{lennon:18}. Subtracting these values from the
 proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
@@
 $\delta v_\mathrm{DEC}=26\pm19\,\kms$, assuming a distance of
-50\,kpc (we do not account for the uncertainty in the distance
-estimate when propagating errors). The radial velocity from
+50\,kpc. % (we do not account for the uncertainty in the distance
+% estimate when propagating errors).
+The radial velocity from
 \cite{bestenlehner:11} then gives the third component along
@@
 motion. It is clear that the most likely origin of the star is R136,
-as predicted from numerical N-body simulations by \cite{fujii:11, banerjee:12}.
+as suggested based on numerical N-body simulations by \cite{fujii:11, banerjee:12}.
 
@@
   \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} \simeq
-  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.63\, \mathrm{Myr} \ \ ,
+  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.6\pm\,0.2 \mathrm{Myr} \ \ ,
 \end{equation}
@@
 \Secref{sec:runaway}, and we use the approximation $1 \kms \simeq 1
-\mathrm{pc \ Myr^{-1}}$.
+\mathrm{pc \ Myr^{-1}}$. As in the rest of this study, we neglect for
+simplicity the error on the distance estimates.
 
@@
 \emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
-$\sim$$0.6$\,Myr ago. Because of the exceptionally large mass
+$0.6\pm0.2$\,Myr ago. Because of the exceptionally large mass
 of this star, this raises the question of which stars must populate
@@
 cluster R136.
-This can be considered an independent confirmation of the detection
+This is consistent with the detection
 of extremely massive stars by \cite{crowther:10} in the core of the
@@
 \begin{acknowledgements}
+  \small
   We are grateful to S.~Torres for 
@@
   Council (ERC) (Grant agreement No. 715063).
-  This work has made use of data from the ESA space mission Gaia (http://www.cosmos.esa.int/gaia), processed by the Gaia Data Processing and Analysis Consortium (DPAC, http://www.cosmos.esa.int/web/gaia/dpac/consortium).
+  This work has made use of data from the ESA space mission Gaia (\url{http://www.cosmos.esa.int/gaia}), processed by the Gaia Data Processing and Analysis Consortium (DPAC, \url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}).
 Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement.

### 2018-05-11 VFTS682.tex
small

--- 2018-05-11 VFTS682.tex
@@
-\documentclass{aa}
+\documentclass[apjl,twocolumn]{emulateapj}
+
 \usepackage{graphicx}
 \usepackage[varg]{txfonts}
-\usepackage[%draft, 
-colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
-\usepackage{amsmath}
-\usepackage[usenames]{xcolor}
+
+\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
+\renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
+%\usepackage[%draft, 
+%colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
+\usepackage{aas_macros}
+%\usepackage{amsmath} % clashes with emulateapjj for some reaons
+\usepackage[usenames,dvipsnames]{xcolor}
 \usepackage{comment}
 \usepackage{multirow}
-\usepackage{chngpage}
-\usepackage{lscape}
+%\usepackage{chngpage}
+%\usepackage{lscape}
 \usepackage{url}
@@
 \definecolor{Wildstrawberry}{rgb}{1.0, 0.26, 0.64}
-\newcommand{\SdM}[1]{{{\color{Wildstrawberry}\bf{#1}}}}
+\newcommand{\SdM}[1]{{{\color{Sepia}{#1}}}}
+\renewcommand{\SdM}[1]{{{{#1}}}}
+
 \newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}
@@
 
-\title{VFTS682: a confirmed dynamical ejection?}
+\title{ \SdM{Gaia identifies VFTS682 as the most massive, dynamically ejected runaway star known to date}}
 
-\author{M.~Renzo\inst{1} \and S.~E.~de~Mink\inst{1} \and \SdM{D.J. Lennon et al.} \todo{etc...}%\and
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, \SdM{D.J. Lennon$^{2}$ et al.} \todo{... add coauthors...}%\and
   % S.~Justham\inst{2,3} \and A.~de~Koter\inst{1} \todo{etc..order to be decided}
@@
 
-\institute{{Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands}
+\affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
+    Amsterdam, 1098 XH Amsterdam, The Netherlands \\
+    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
+    $^{3}$ ... 
   % \and{School of Astronomy \& Space Science, University of the Chinese
@@
 
-\offprints{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+\thanks{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
 \date{}
-\abstract{}
+\begin{abstract}
+\SdM{ The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
+%
+Previous work spectroscopic work identified VFTS682 as one of the most massive stars known.
+It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 32 pc of the star cluster R136.
+The lack of a detected offset in radial velocity derived from the spectra, lead to speculations whether such extreme stars can form in relative isolation.
+%
+We use Gaia's unprecedented proper motion measurements to investigate the kinematics of this object.
+We show that VFTS 682 is a bonafide runaway star moving with a relative velocity of about $60\kms$ away from the star cluster R136.
+The inferred age, distance, and the magnitude and direction of the velocity are consistent a dynamical ejection scenario from the cluster.
+%
+If correct, this would make VFTS 682 the most massive dynamically ejected runaway star known to date.
+It proves that star clusters are capable of ejecting their most massive members.
+We discuss the implications for the existence of stars with masses well above the canonical upper mass limit inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
+}
+
+\end{abstract}
 
@@
 Our results indicate that VFTS 682 is a bona fide runaway star.
-The direction an magnitude of the velocity vector is full consistent with dynamical ejection from R136.
+The direction an magnitude of the velocity vector is fully consistent with dynamical ejection from R136.
 Its inferred age is consistent with the inferred travel time.
 This means that the hypothesis that VFTS 682 is formed in relative isolation is rejected.
-This makes  VFTS 682 the most massive runaway star known to date.
-
+This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.
 }
 
-% make sure to cite this in the conclusion \cite{fujii:11, banerjee:12}
-\todo{missing: R136 too young for CC, cite age $\leq2$\,Myr \citep[][]{crowther:10,sabbi:12}
+\todo{(1. Make sure we cite and disucss fujii:11, banerjee:12 in discussion. 2. Make link to dynamical interaction proposed to form BHBH systems. }
 
-\section{Gaia DR2 data selection}
+\section{Observations}
 \label{sec:sample}
+\SdM{
+In this section we summarize the remarkable properties derived for VFTS682 \Secref{data:vfts683} in earlier studies and we discuss how we Gaia data in 
+\Secref{data:gaia}.
 
-VFTS682 is labeled in the Gaia DR2
-catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}} with the
-source id 4657685637907503744. The star has a
+\subsection{ Overview of the stellar parameters of VFTS682 derived in previous studies \label{data:vfts683}}
+
+The star VFTS682  \citep{Evans+2011}, located at  RA 05:38:55.510  Dec -69:04:26.72 \todo{format properly}
+was originally classified as a young stellar object \citep{Gruendl+2009} based on its mid-infrared excess.
+\citet{Evans+2011} classified the object a Wolf-Rayet star using spectroscopic observations taken a part of the VLT-FLAMES Tarantula survey.
+The spectra available covered $\lambda$4000--7000 and were taken at multiple epochs.
+\citep{Bestenlehner+2011} studied this object in more detail used and looked for radial velocity variations.
+This allowed them to exclude the presence of a close companion with high confidence.
+
+\citep{Bestenlehner+2011} further analyzed the spectra using the non-LTE model atmosphere code CMFGEN \citep{Hillier+1998} to derive the stellar parameters and surface abundances.
+They derive a find a peculiar extinction ($R_V \sim 4.7$), leading to a surprisingly high luminosity $\log_{10} L (\rm{L_\odot}) =  6.5\pm0.2$.
+Bayesian analysis with the Bonsai code \citep{Schneider+2017} against  evolutionary models by \citet{Brott+2011} and \citet{Kohler+2015} leads to an estimated present-day mass of  $138^{+28}_ {-16}\Msun$, an age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial masse of $150^{+29}_{-17}\Msun$.
+This places around the ``canonical upper limit'' of $150\Msun$  by \citet{Figer2005}.
+
+The star is very similar to R136a3, located inside the star cluster R136, \citep{Crowther+2010}  for which \citet{Crowther+2016} report a current mass estimate of $180^{30}_{-30\Msun}$.
+The R136 cluster hosts two further very massive WN5h stars R1361 and R136a2, whose estimated current masses are even higher.
+However, VFTS682 stands out by its location a projected offset of 32\,pc from the center of the cluster.
+
+Further worth noticing is the variability of the star.
+\citep{Bestenlehner+2011} discuss the Optical Gravitational Lensing Experiment (OGLE-III) light curves \citep{Udalski+2008} and show a variability in the V-band of the 10\% level on a timescale of years.
+The comment that this is unusual for Wolf-Rayet and more reminiscent of Luminous Blue Variables.
+
+\todo{Report mass loss rate.  Report LOS velocity.}
+
+%\citep{Parker1993}
+
+\subsection{Kinematic data from Gaia DR2  \label{data:gaia}}
+
+The Gaia Data Release 2, which became available on 25 April 2018, provides the five-parameter astrometric solution (positions, parallaxes, and proper motions) for more than a bilion sources \citep{Gaia-Collaboration+2018}.
+The median uncertainty in the proper motion components the corresponding uncertainties are $0.05\,\masyr$ for bright sources  ($G<14$ mag) and $0.2\masyr$ at $G=17\,\mathrm{mag}$ \citep{Lindegren+2018}.
+The uncertainties quoted above to $xx\kms$ and $yy\kms$ respectively.
+This unprecedented accuracy allows the identification of bright runaway stars at the distance to the LMC, 50\,kpc.
+
+VFTS682 is identified with the gaia source} with id 4657685637907503744 in the Gaia DR2
+catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia reports a G-band magnitude is 15.65. The star has a
 \texttt{visibility\_period} = 17, which counts how many observations have
 been used to reconstruct its astrometric solution
-\citep[][]{lindengren:18}. Its reported G-band
-magnitude is 15.65, cf. the V-band magnitude of 16.08
-\citep[][]{evans:11, bestenlehner:11}, and the reported
+\citep[][]{lindengren:18}, and the reported
 \texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
@@
   \includegraphics[width=0.5\textwidth]{./figures/main_plot}  
-  \caption{The red solid arrow indicates the proper motion of VFTS682
-    relative to the region, starting from the present day position of
-    the star. The semi-transparent arrows indicate the possible
-    directions of projected motion within the errors, and are extended
-    backwards to illustrate the uncertainty on the origin of the
-    star. Blue stars
-    are a subset of those we use to define the local rest frame. \todo{load color picture on background}}
+  \caption{The red cross indicates the position of VFTS682. Blue stars
+    are those we use to define the generic ``surroundings'', while the
+    green dots indicate the stars we use to probe the core of R136,
+    magnified in the inset. The red arrow shows the direction of the
+    proper motion of VFTS682 relative to R136, and the gray shade
+    indicates the uncertainty in the direction. \todo{load color picture on background}}
   \label{fig:main}
@@
 \label{sec:discussion}
+\todo{Selma: Should add caveat that it may be a spurious detection.  That reconfirmation in following epochs is warranted. }
 
@@
   what is its total mass and relaxation time?
-\item another important thing missing: with vsini(i)<200, completely average, there is no reason to think it's a CHE, contrary to the way they estimate the age in Bestenlehner:11
 \end{itemize}
 
+\todo{Merge Selma's my\_bib bibtex file with Mathieu's and prevent double refs. }
 \bibliographystyle{aa}
-\bibliography{bibliography/vfts682}
+\bibliography{bibliography/vfts682,my_bib}
 
 \begin{acknowledgements}
-  We are grateful to S.~Torres for extremely useful discussions on the
+  We are grateful to S.~Torres for 
+  %extremely 
+  very useful discussions on the
   Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
+  \SdM{SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
+Council (ERC) (Grant agreement No. 715063).
+   This work has made use of data from the ESA space mission Gaia (http://www.cosmos.esa.int/gaia), processed by the Gaia Data Processing and Analysis Consortium (DPAC, http://www.cosmos.esa.int/web/gaia/dpac/consortium).
+Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement.
+}
   The background image of \Figref{fig:main} is based on observations

### 2018-05-11 VFTS682.tex
trying to recover access without damages

--- 2018-05-11 VFTS682.tex
@@
 \newcommand{\SdM}[1]{{{\color{Sepia}{#1}}}}
-%\renewcommand{\SdM}[1]{{{{#1}}}}
+\renewcommand{\SdM}[1]{{{{#1}}}}
 
@@
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  S.~Justham$^{3,4}$, J.~Vink$^{5}$, F.~R.~Schneider$^6$, A.~de~Koter$^{1}$ \todo{... add coauthors...} }
+  S.~Justham$^{3,4}$, J.~Vink$^{5}$, F.~R.~Schneider$^6$,
+  J.~Bestenlehner, A.~de~Koter$^{1}$ \todo{... add coauthors...} }
 
@@
 
-% %%% change \tablenotetext{a} -> \tablenotetext{1} etc. to omit LaTeX errors
+% %%% change \tablenotetext{a} - \tablenotetext{1} etc. to omit LaTeX errors
 % %%% Karol Kozio? for ShareLaTeX
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{./figures/main_plot}  
+  \includegraphics[width=0.5\textwidth]{./figures/main_plot_good}  
   \caption{The red solid arrow indicates the proper motion of VFTS682
@@
 of extremely massive stars by \cite{crowther:10} in the core of the
-cluster. The projected rotational equatorial velocity of VFTS682
+cluster. The projected rotational equatorial
+velocity\footnote{However, the determination of the rotational
+  velocity for stars showing lines in emission is complicated because
+  of the optically thick wind screening the surface of the star, and should be
+  considered with caution.} of VFTS682
 reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
@@
   what is its total mass and relaxation time?
-\item Make link to dynamical interaction proposed to form BHBH systems.
+\item Make link to dynamical interaction proposed to form BHBH
+  systems.
+  \item from Fabian's email (see comment in tex):% One last aspect: I am always wondering what happens to the star/binary that ejected 682.
+There is a huge linear momentum in this star of ~150 Msun * 60 km/s.
+Even if the bully was a 300 Msun total object, it would obtain a kick of 30 km/s.
+Is this enough to eject the bully star?
+This velocity would already be 100 km/s for a 90 Msun bully (I was thinking of 016 here...
+'Cause I do not know what ejected this object either).
+Of course, if there is a third body, these velocities are no longer correct.
+The escape velocity from R136 at 0.5pc for a total mass of 10^5 Msun would only be of order 20 km/s...
+So a 300 Msun bully is at the edge of what is possible and one might need something even more massive or another very massive object will also be ejected.
+Or am I wrong here?
+Do the Banerjee and Co simulations say something about this?
 \end{itemize}

### 2018-05-11 VFTS682.tex
added joachim and minor

@@
 \usepackage[varg]{txfonts}
-\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
+%\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
+\usepackage[%draft, 
+pagebackref,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
 \renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
-%\usepackage[%draft, 
-%colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
 \usepackage{aas_macros}
@@
   S.~Justham$^{3,4}$, J.~Vink$^{5}$, F.~R.~Schneider$^6$,
-  J.~Bestenlehner, A.~de~Koter$^{1}$ \todo{... add coauthors...} }
+  J.~Bestenlehner$^7$, A.~de~Koter$^{1}$}
 
@@
      $^5$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
-     $^6$ {Department of Physics, University of Oxford, Keble Road, Oxford OX1 3RH, UK}  
+     $^6$ {Department of Physics, University of Oxford, Keble Road,
+       Oxford OX1 3RH, UK} \\
+     $^7$ {Department of Physics and Astronomy, Hicks Building, Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
 }
@@
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) -69:04:26.72 in the ICGS frame \citep[][]{brown:18}
+(DEC) -69:04:26.72 in the ICGS frame \todo{check coordinates} \citep[][]{brown:18}
 was originally classified as a young stellar object \citep{gruendl:09}
@@
 the line of sight, allowing us to calculate the three-dimensional
-peculiar speed of the star:
+peculiar speed of the star\footnote{We implicitly assume that the use
+  of a different reference frame for $\delta v_\mathrm{rad}$ does not
+  introduce significant errors.}:
 

### 2018-05-11 VFTS682.tex
need merging

@@
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) -69:04:26.72 in the ICGS frame \todo{check coordinates} \citep[][]{brown:18}
+(DEC) -69:04:26.72 \citep[][]{evans:11}
 was originally classified as a young stellar object \citep{gruendl:09}
@@
 \cite{bestenlehner:11}, and the best fit parallax of this star is
-negative. We do not use the effective temperature of the star anywhere
+negative \citep[see, e.g.,][]{hogg:18}. We do not use the effective temperature of the star anywhere
 in this study, and we attribute the unphysical value of the parallax
@@
     relative to the region, starting from the present day position of
-    the star. The semi-transparent arrows indicate the possible
+    the star. The orange arrows indicate the possible
     directions of projected motion within the errors, and are extended
@@
     star. The blue stars
-    are a subset of those we use to define the local rest frame. \todo{load color picture on background}}
+    are a subset of those we use to define the local rest
+    frame. \todo{load color picture on background, clip away outliers}}
+
   \label{fig:main}

### 2018-05-11 VFTS682.tex
authors mostly

--- 2018-05-11 VFTS682.tex
@@
 \definecolor{Wildstrawberry}{rgb}{1.0, 0.26, 0.64}
-\newcommand{\SdM}[1]{{{\color{Sepia}{#1}}}}
-\renewcommand{\SdM}[1]{{{{#1}}}}
+%\newcommand{\SdM}[1]{{{\color{Sepia}{#1}}}}
+\newcommand{\SdM}[1]{{{\color{brown}{#1}}}}
+%\renewcommand{\SdM}[1]{{{{#1}}}}
 
@@
 
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  S.~Justham$^{3,4}$, J.~Vink$^{5}$, F.~R.~Schneider$^6$,
-  J.~Bestenlehner, A.~de~Koter$^{1}$ \todo{... add coauthors...} }
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, J.
+Bestenlehner$^{x}$,  F.~R.~Schneider$^6$, J.~Vink$^{5}$ \SdM{et al.?} \\ \SdM{[ Alphabetical after Lennon.
+Possibly add further last-minute coauthors]},  }
+
+  % S.~Justham$^{3,4}$,
+%  A.~de~Koter$^{1}$ \todo{... add coauthors...} }
 
@@
     $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
-    $^{3}$ {School of Astronomy \& Space Science, University of the Chinese
-    Academy of Sciences, Beijing 100012, China}\\
-     $^4$ {National Astronomical Observatories, Chinese Academy of
-       Sciences, Beijing 100012, China}\\
+  %  $^{3}$ {School of Astronomy \& Space Science, University of the Chinese
+   % Academy of Sciences, Beijing 100012, China}\\
+    % $^4$ {National Astronomical Observatories, Chinese Academy of
+     %  Sciences, Beijing 100012, China}\\
      $^5$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
@@
 
-\thanks{M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+%\thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
 \date{}
@@
 
-% %%% change \tablenotetext{a} - \tablenotetext{1} etc. to omit LaTeX errors
+% %%% change \tablenotetext{a} -> \tablenotetext{1} etc. to omit LaTeX errors
 % %%% Karol Kozio? for ShareLaTeX
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{./figures/main_plot_good}  
+  \includegraphics[width=0.5\textwidth]{./figures/main_plot}  
   \caption{The red solid arrow indicates the proper motion of VFTS682
@@
 of extremely massive stars by \cite{crowther:10} in the core of the
-cluster. The projected rotational equatorial
-velocity\footnote{However, the determination of the rotational
-  velocity for stars showing lines in emission is complicated because
-  of the optically thick wind screening the surface of the star, and should be
-  considered with caution.} of VFTS682
+cluster. The projected rotational equatorial velocity of VFTS682
 reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
@@
   what is its total mass and relaxation time?
-\item Make link to dynamical interaction proposed to form BHBH
-  systems.
-  \item from Fabian's email (see comment in tex):% One last aspect: I am always wondering what happens to the star/binary that ejected 682.
-There is a huge linear momentum in this star of ~150 Msun * 60 km/s.
-Even if the bully was a 300 Msun total object, it would obtain a kick of 30 km/s.
-Is this enough to eject the bully star?
-This velocity would already be 100 km/s for a 90 Msun bully (I was thinking of 016 here...
-'Cause I do not know what ejected this object either).
-Of course, if there is a third body, these velocities are no longer correct.
-The escape velocity from R136 at 0.5pc for a total mass of 10^5 Msun would only be of order 20 km/s...
-So a 300 Msun bully is at the edge of what is possible and one might need something even more massive or another very massive object will also be ejected.
-Or am I wrong here?
-Do the Banerjee and Co simulations say something about this?
+\item Make link to dynamical interaction proposed to form BHBH systems.
 \end{itemize}

### 2018-05-11 VFTS682.tex

--- 2018-05-11 VFTS682.tex
@@
 
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  S.~Justham$^{3,4}$, J.~Vink$^{5}$, F.~R.~Schneider$^6$,
-  J.~Bestenlehner$^7$, A.~de~Koter$^{1}$}
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, J.
+Bestenlehner$^{x}$,  F.~R.~Schneider$^6$, J.~Vink$^{5}$ \SdM{et al.?} \\ \SdM{[ Alphabetical after Lennon.
+Possibly add further last-minute coauthors]},  }
+
+  % S.~Justham$^{3,4}$,
+%  A.~de~Koter$^{1}$ \todo{... add coauthors...} }
 

### 2018-05-11 VFTS682.tex
text polishing

@@
 release (DR2) to investigate the kinematics of this object. We show
-that VFTS 682 is a bonafide runaway star moving with a relative
-velocity of about $64\pm21\kms$ away from the star cluster R136.  The
+that VFTS 682 is a bonafide runaway star moving away from the star
+cluster R136 with a relative
+projected two-dimensional speed of about $\sim$$40\pm20 \kms$velocity,
+implying a three-dimensional speed of about $\sim$$60\pm20\kms$ when
+accounting for the velocity component along the line of sight.  The
 inferred age, distance, and the magnitude and direction of the
@@
 
+\todo{recheck all numbers}
 \section{Introduction}
@@
 
-% feels like unnecessary here...
-% In this section we summarize the remarkable properties derived for VFTS682 \todo{?}\Secref{data:vfts683}
-% in earlier studies and we discuss we Gaia data in \Secref{data:gaia}.
-
 \subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
@@
 model atmosphere code CMFGEN \citep{hillier:98} to derive the stellar
-parameters and surface abundances. They inferred a peculiar extinction
-($R_V\simeq4.7$), leading to a surprisingly high luminosity
+parameters and surface abundances. They inferred an extinction
+of $R_V\simeq4.7$, leading to the high luminosity
 $\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citep{schneider:18} estimated
@@
 
-Further worth noticing is the variability of the star.
-\citep{bestenlehner:11} discuss the Optical Gravitational Lensing
+Further worth noticing is the variability of the
+star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,dex
+lower than what found by \citet{evans:11}. \citet{bestenlehner:11} discuss the Optical Gravitational Lensing
 Experiment (OGLE-III) light curves \citep{udalski:08} and show a
 variability in the V-band of the $\sim$10\% level on a timescale of years.
-The comment that this is unusual for Wolf-Rayet and more reminiscent
+They commented that this is unusual for Wolf-Rayet and more reminiscent
 of Luminous Blue Variable (LBV) stars.
 
-\todo{Report mass loss rate.  Report LOS velocity.}
-\citep{parker:93}
+We obtain the peculiar radial velocity $\delta v_\mathrm{rad}$ of VFTS682 as the difference
+between the average radial velocity of the 30 Doradus region
+($270\pm10\kms$) minus the radial velocity measured from the NV $\lambda4944$
+line for VFTS682  \citep[$300\pm10\kms$, ][]{bestenlehner:11}. We implicitly assume that the use
+of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not
+introduce significant errors. We also note that the determination of
+radial velocities for emission-lines stars is difficult because of the
+presence of an optically thick, likely non-homogeneous, and variable
+wind hiding the stellar surface.
 
@@
     $\mu_\mathrm{DEC}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}0.78 $\pm$ 0.08& \\
-    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{0}-45 $\pm$ 25 & \cite{bestenlehner:11}\\
+    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{0}-30 $\pm$ 20 & \cite{bestenlehner:11}\\
 \tableline
   \end{tabular}
-  \tablecomments{The peculiar radial
-    velocity $\delta v_\mathrm{rad}$ is obtained as the difference
-    between the average radial velocity of the 30 Doradus region
-    ($270\pm10\kms$) minus the radial velocity measured from the HeII $\lambda4686$
-    line for VFTS682 ($315\pm15\kms$).}
+  % \tablecomments{}
     \end{center}
@@
 parallaxes, and proper motion components) for more than a billion
-sources \citep{brown:18}.%  \todo{why do we care about the
-%   median? I'm
-%   selecting data with errors smaller than this, and I'm not using
-%   fainter stars!I think this is not useful here} The median uncertainty in the proper motion
-% components the corresponding uncertainties are $0.05\,\masyr$ for
-% bright sources  ($G<14$ mag) and $0.2\masyr$ at $G=17\,\mathrm{mag}$
-% \citep{lindegren:18}.
-The uncertainties quoted above correspond to $xx\kms$ and $yy\kms$ respectively.
-This unprecedented accuracy allows the identification of bright runaway stars at the distance to the LMC, 50\,kpc.
-
-VFTS682 is identified with the id 4657685637907503744 in the Gaia DR2
+sources \citep{brown:18}. VFTS682 is identified with the id 4657685637907503744 in the Gaia DR2
 catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia
@@
 requirement of having parallaxes smaller than $2\,\mathrm{mas}$ does
-not reduces further our sample.
+not reduces further our sample. Finally, we remove clear outliers from
+the proper motion distribution
 
@@
 the line of sight, allowing us to calculate the three-dimensional
-peculiar speed of the star\footnote{We implicitly assume that the use
-  of a different reference frame for $\delta v_\mathrm{rad}$ does not
-  introduce significant errors.}:
+peculiar speed of the star:
 
@@
     +\left(\delta v_\mathrm{DEC}\right)^2+\left(\delta
-      v_\mathrm{rad}\right)^2} = 64 \pm 21 
-  \, \kms \ .
+      v_\mathrm{rad}\right)^2} = 55 \pm 18 
+  \kms \ .
 \end{equation}
@@
   \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} \simeq
-  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.6\pm\,0.2 \mathrm{Myr} \ \ ,
+  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.6\pm\,0.2\, \mathrm{Myr} \ \ ,
 \end{equation}
@@
     +\left(\delta v_\mathrm{DEC}\right)^2} =46\pm
-17\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
+18\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
 using the relative proper motion components calculated in
-\Secref{sec:runaway}, and we use the approximation $1 \kms \simeq 1
-\mathrm{pc \ Myr^{-1}}$. As in the rest of this study, we neglect for
+\Secref{sec:runaway}.% , and we use the approximation $1 \kms \simeq 1
+% \mathrm{pc \ Myr^{-1}}$.
+As in the rest of this study, we neglect for
 simplicity the error on the distance estimates.
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{./figures/main_plot_good}  
+  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good}  
   \caption{The red solid arrow indicates the proper motion of VFTS682
@@
     directions of projected motion within the errors, and are extended
-    backwards to illustrate the uncertainty on the origin of the
+    backwards (dashed) to illustrate the uncertainty on the origin of the
     star. The blue stars
     are a subset of those we use to define the local rest
-    frame. \todo{load color picture on background, clip away outliers}}
+    frame. \todo{load color picture on background}}
 
@@
 
-\todo{check \cite{banerjee} and rewrite next paragraph, they
+\todo{check \cite{banerjee:12} and rewrite next paragraph, they
   have stuff} Unfortunately, their simulations did not include stars more

### 2018-05-11 VFTS682.tex
incorporated comments on text from van der Marel

@@
 
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, J.
-Bestenlehner$^{x}$,  F.~R.~Schneider$^6$, J.~Vink$^{5}$ \SdM{et al.?} \\ \SdM{[ Alphabetical after Lennon.
-Possibly add further last-minute coauthors]},  }
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
+  J. Bestenlehner$^{3}$,  F.~R.~Schneider$^5$, R.~.P.~van~der~Marel$^{6,7}$, J.~Vink$^{8}$ \SdM{et al.?} \\ \SdM{[ Alphabetical after Lennon. Possibly add further last-minute coauthors]},  }
 
@@
     Amsterdam, 1098 XH Amsterdam, The Netherlands \\
-    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78, E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
+    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78,
+    E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
+    $^3$ {Department of Physics and Astronomy, Hicks Building,
+      Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
   %  $^{3}$ {School of Astronomy \& Space Science, University of the Chinese
@@
      %  Sciences, Beijing 100012, China}\\
-     $^5$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
-     $^6$ {Department of Physics, University of Oxford, Keble Road,
+     $^5$ {Department of Physics, University of Oxford, Keble Road,
        Oxford OX1 3RH, UK} \\
-     $^7$ {Department of Physics and Astronomy, Hicks Building, Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
+     $^{6}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+       Baltimore, MD 21218, USA}\\
+     $^{7}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
+     $^8$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}
 }
@@
 It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 29\,pc of the star cluster R136.
-The lack of a detected offset in radial velocity derived from the spectra, lead to speculations whether such extreme stars can form in relative isolation.
+The lack of a detected offset in radial velocity derived from the spectra, led to speculations whether such extreme stars can form in relative isolation.
 %
@@
 
-How do massive stars form is one of the major longstanding questions in astrophysics
+How massive stars form is one of the major longstanding questions in astrophysics
 \citep[e.g.,][]{zinnecker:07}. Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
@@
 
-In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to  form preferentially in the dense cores of clusters.
+In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to  form preferentially in the dense cores.
 In this picture, field stars are primarily the result of the dissolution of dense groups.
@@
 hypothesis is that these massive stars formed in clusters, and then
-been ejected from their birth locations. Such ejections may result
+were ejected from their birth locations. Such ejections may result
 from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems by the core-collapse of the companion star \citep[e.g.,][]{zwicky:57, blaauw:61}.
@@
 An interesting contribution to the debate on whether or not massive
-stars an form in relative isolation was presented by
-\cite{bestenlehner:11, bressert:12}, who discuss the case of the very massive star
+stars form in relative isolation was presented by
+\cite{bestenlehner:11} and \cite{bressert:12}, who discuss the case of the very massive star
 VFTS682.   This star is located in the field of the 30 Doradus region
@@
 origin of VFTS 682.
-Quoting directly from their abstract: "either (i) the star either formed in situ, which would have profound implications for the formation mechanism of massive stars, or (ii) VFTS 682 is a slow runaway star that originated from the dense cluster R136, which would make it the most massive runaway known to date."
+Quoting directly from their abstract: "(i) the star either formed in situ, which would have profound implications for the formation mechanism of massive stars, or (ii) VFTS 682 is a slow runaway star that originated from the dense cluster R136, which would make it the most massive runaway known to date."
 
-In this study we report on our analysis the data from the second Gaia data release (DR2) \cite[][]{gaia:16,brown:18}.
+Recently, \citet{platais:18} used multi-epoch Hubble Space
+Telescope imaging to identify and shed new light on runaway stars from
+the 30 Doradus region based on proper motion measurements.
+In this study we report on our analysis of the data from the second Gaia data release (DR2) \cite[][]{gaia:16,brown:18}.
 We combine the radial velocity measurements from the VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2 to reconstruct the three-dimensional velocity of VFTS682, and test the hypothesis that this star was ejected from R136.
@@
 Our results indicate that VFTS 682 is a bona fide runaway star.
-The direction and magnitude of the velocity vector is fully consistent with dynamical ejection from R136.
+The direction and magnitude of the velocity vector are fully consistent with dynamical ejection from R136.
 Its inferred age is consistent with the inferred travel time.
@@
 of $R_V\simeq4.7$, leading to the high luminosity
-$\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citep{schneider:18} estimated
-the present-day mass of VFTS682 at  $137.8^{+27.5}_ {-15.9}\Msun$, an
+$\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citet{schneider:18} estimated
+a present-day mass for VFTS682 of $137.8^{+27.5}_ {-15.9}\Msun$, an
 apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
@@
 \citep{schneider:17}, based on the evolutionary models from
-\citet{brott:11, kohler:15}. These values place VFTS682 around the
+\citet{brott:11} and \cite{kohler:15}. These values place VFTS682 around the
 ``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
@@
 The star is very similar from the spectral point of view to R136a3, located inside the star cluster
-R136, \citep{crowther:10}  for which \citet{crowther:16} report a
+R136 \citep{crowther:10}  for which \citet{crowther:16} report a
 current mass estimate of $180^{+30}_{-30}\Msun$. The R136 cluster hosts
-at least two more very massive WN5h stars R136a1, R136a2 whose
+at least two more very massive WN5h stars, R136a1 and R136a2, whose
 estimated current masses are even higher. However, VFTS682 stands
@@
 Experiment (OGLE-III) light curves \citep{udalski:08} and show a
-variability in the V-band of the $\sim$10\% level on a timescale of years.
-They commented that this is unusual for Wolf-Rayet and more reminiscent
+variability in the V-band at $\sim$10\% level on a timescale of years.
+They commented that this is unusual for Wolf-Rayet stars and more reminiscent
 of Luminous Blue Variable (LBV) stars.
@@
     $\mu_\mathrm{DEC}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}0.78 $\pm$ 0.08& \\
-    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{0}-30 $\pm$ 20 & \cite{bestenlehner:11}\\
+    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{-0}30 $\pm$ 20 & \cite{bestenlehner:11}\\
 \tableline
@@
 of VFTS682 by selecting from the Gaia DR2 catalog a sample of nearby
-stars, following closely the approach of \cite{lennon:18}.
+stars, following closely the approach of \cite{vandermarel:02,lennon:18}.
 We select all the stars in a target of 0.2 degrees around R136
@@
 requirement of having parallaxes smaller than $2\,\mathrm{mas}$ does
-not reduces further our sample. Finally, we remove clear outliers from
-the proper motion distribution
+not reduces further our sample. This yields a sample of 651 stars,
+from which remove six clear outliers from
+the proper motion distribution, ending up with 645 stars to define the
+local rest frame.
 
@@
 on the proper motion component reported by Gaia. The sums run over
-all the $N=651$ stars in our selected sample. We evaluate each proper motion
+all the $N=646$ stars in our selected sample. We evaluate each proper motion
 component separately. For simplicity, throughout this study, we assume the same
@@
 in good agreement with what is found by \cite{lennon:18}. Subtracting these values from the
-proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
+proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
 components of proper motion of the star relative to the surrounding region
@@
 
-These can be converted in the
+These can be converted into
 the components of the relative transverse velocity $\delta v_\mathrm{RA}=38\pm17\,\kms$,
@@
 motion. It is clear that the most likely origin of the star is R136,
-as suggested based on numerical N-body simulations by \cite{fujii:11, banerjee:12}.
+as suggested based on numerical N-body simulations by \cite{fujii:11}
+and \cite{banerjee:12}.
 
@@
 The kinematic age $\tau_\mathrm{kin}$ is smaller than the apparent age
-of the star age of $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
+of the star of $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
 corroborates the idea that the star is the result of a dynamical
@@
 
-Based on our results, we can claim that VFTS682 is the most massive
+Based on our results, we claim that VFTS682 is the most massive
 runaway known to date, with a peculiar three-dimensional speed of $64\pm21\,\kms$. This means that isolated star formation is
@@
 \citep[][]{maeder:00,demink:09,marchant:16}, and (ii) it has not
-accreted mass from a binary companion, nor it will, since the multi-epoch
-data of the VFTS survey rule out the presence of a companion at
+accreted mass from a binary companion, nor will it, since the multi-epoch
+data of the VFTS survey rule out the presence of a close companion at the
 present day. Moreover, the spectral type of VFTS682
@@
 The kinematic age of VFTS682 puts an
-upperlimit to the timescale to form such ``bully binary'' in
+upperlimit to the timescale to form such a ``bully binary'' in
 R136. The cluster must have been at the very beginning of its
 evolution, given the age estimate of $\lesssim 2$\,Myr
-\cite{crowther:10,sabbi:12} and the kinematic age of VFTS16. If the
+\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS16. If the
 cluster is indeed younger than the shortest stellar lifetime
-\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, ejection from a binary
-system of VFTS682 is excluded since the region is to young for stars
+\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, then the alternative
+explanation for ejection of VFTS682 from a binary
+system is excluded since the region is too young for stars
 to have experience core-collapse already.
@@
 the existence of VFTS682, how many other stars should have been
-ejected from the cluster already and what is their mass
-distribution. A comprehensive study of the kinematic properties of the
+ejected from the cluster already and what their mass
+distribution is. A comprehensive study of the kinematic properties of the
 large sample of stars with extremely large inferred masses visible

### 2018-05-12 VFTS682.tex
incorporated comments on text from van der Marel more
added stephen

@@
 
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  J. Bestenlehner$^{3}$,  F.~R.~Schneider$^5$, R.~.P.~van~der~Marel$^{6,7}$, J.~Vink$^{8}$ \SdM{et al.?} \\ \SdM{[ Alphabetical after Lennon. Possibly add further last-minute coauthors]},  }
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, J.
+Bestenlehner$^{3}$, S.~Justham$^{4,5}$, F.~R.~Schneider$^6$, R.~P.~van~der~Marel$^{7,8}$, J.~Vink$^{9}$ \SdM{et al.?} \\ \SdM{[ Alphabetical after Lennon.
+Possibly add further last-minute coauthors]},  }
 
-  % S.~Justham$^{3,4}$,
+  % ,
 %  A.~de~Koter$^{1}$ \todo{... add coauthors...} }
@@
       Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
-  %  $^{3}$ {School of Astronomy \& Space Science, University of the Chinese
-   % Academy of Sciences, Beijing 100012, China}\\
-    % $^4$ {National Astronomical Observatories, Chinese Academy of
-     %  Sciences, Beijing 100012, China}\\
-     $^5$ {Department of Physics, University of Oxford, Keble Road,
+   $^{4}$ {School of Astronomy \& Space Science, University of the Chinese
+   Academy of Sciences, Beijing 100012, China}\\
+    $^5$ {National Astronomical Observatories, Chinese Academy of
+      Sciences, Beijing 100012, China}\\
+     $^6$ {Department of Physics, University of Oxford, Keble Road,
        Oxford OX1 3RH, UK} \\
-     $^{6}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+     $^{7}$ {Space Telescope Science Institute, 3700 San Martin Drive,
        Baltimore, MD 21218, USA}\\
-     $^{7}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-     $^8$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}
+     $^{8}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
+     $^9$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}
 }
@@
   \end{tabular}
-  % \tablecomments{}
+  \tablecomments{We truncate the errors on RA and DEC for the purpose
+    of display. Readers are encouraged to consult the Gaia DR2 catalog
+  for accurate values.}
     \end{center}

### 2018-05-12 VFTS682.tex
minor

@@
 
-Recently, \citet{platais:18} used multi-epoch Hubble Space
+Recently, \citet{platais:15,platais:18} used multi-epoch Hubble Space
 Telescope imaging to identify and shed new light on runaway stars from
 the 30 Doradus region based on proper motion measurements.
-In this study we report on our analysis of the data from the second Gaia data release (DR2) \cite[][]{gaia:16,brown:18}.
+In this study we report on our analysis of the data from the second Gaia data release \cite[DR2,][]{gaia:16,brown:18}.
 We combine the radial velocity measurements from the VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2 to reconstruct the three-dimensional velocity of VFTS682, and test the hypothesis that this star was ejected from R136.
 
-Our results indicate that VFTS 682 is a bona fide runaway star.
-The direction and magnitude of the velocity vector are fully consistent with dynamical ejection from R136.
-Its inferred age is consistent with the inferred travel time.
-This means that the hypothesis that VFTS 682 is formed in relative isolation is rejected.
-This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.
+Our results indicate that VFTS 682 is a bona fide runaway star. The
+direction and magnitude of the velocity vector are consistent with
+dynamical ejection from R136. The age inferred from the spectral
+analysis \citep[from][]{schneider:18} is consistent with the travel
+time we calculate for this star. Therefore, the hypothesis that VFTS
+682 is formed in relative isolation is rejected.  This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.
 

### 2018-05-12 VFTS682.tex
newer dataset limiting parallaxes to range -0.1, 0.1 - I think lennon_like_small has a bad star in it

@@
 
-% %% Any table notes must follow the \end{tabular} command.
-
-% %%% change \tablenotetext{a} - \tablenotetext{1} etc. to omit LaTeX errors
-% %%% Karol Kozio? for ShareLaTeX
-
 The Gaia Data Release 2, which became available on 25 April 2018,
@@
 We define a local standard frame of reference to derive the peculiar velocity
-of VFTS682 by selecting from the Gaia DR2 catalog a sample of nearby
-stars, following closely the approach of \cite{vandermarel:02,lennon:18}.
-We select all the stars in a target of 0.2 degrees around R136
+of VFTS682 with respect to the cluster R136 by selecting from the Gaia DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
+We select all the stars in a target of 0.05 degrees around R136
 (NGC2070) fulfilling the following criteria. First, we require G-band
@@
 \texttt{astrometric\_excess\_noise} $< 1$, the errors on the proper
-motion components to be smaller than 0.1\,$\mathrm{mas\ yr^{-1}}$,
-and the proper motion components themselves to be smaller than
-2\,$\mathrm{mas\ yr^{-1}}$ in absolute value. At the distance to the
-LMC, $1\mathrm{mas\ yr^{-1}}\simeq250\,\kms$ \citep[e.g.,][]{lennon:18}, so the cut on the values
-of the proper motions removes stars that would have projected
-tangential velocities in excess of $\sim$500\,$\kms$, which are most
-likely to be foreground stars. We checked that the additional
-requirement of having parallaxes smaller than $2\,\mathrm{mas}$ does
-not reduces further our sample. This yields a sample of 651 stars,
-from which remove six clear outliers from
-the proper motion distribution, ending up with 645 stars to define the
-local rest frame.
+motion components to be smaller than 0.07\,$\mathrm{mas\
+  yr^{-1}}$. At the distance to the LMC, $0.1\mathrm{mas\
+  yr^{-1}}\simeq25\,\kms$ \citep[e.g.,][]{platais:18}, therefore this
+corresponds to requiring transverse velocities smaller than $\sim$$17\kms$. We
+remove obvious foreground stars by requiring parallaxes smaller than $0.1\,\mathrm{mas\ yr^{-1}}$.
+% At the distance to the
+% LMC, $1\mathrm{mas\ yr^{-1}}\simeq250\,\kms$ \citep[e.g.,][]{lennon:18}, so the cut on the values
+% of the proper motions removes stars that would have projected
+% tangential velocities in excess of $\sim$500\,$\kms$, which are most
+% likely to be foreground stars. We checked that the additional
+% requirement of having parallaxes smaller than $2\,\mathrm{mas}$ does
+% not reduces further our sample.
+This selection yields a sample of 69 stars.% ,
 
@@
   \langle \mu_i\rangle = \frac{\sum_\mathrm{stars}\frac{1}{\Delta
-      \mu_i}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i}} \ \ , \
+      \mu_i^2}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i^2}} \ \ , \
   \ \Delta \langle \mu_i\rangle = \frac{\sqrt{N}}{\sum_\mathrm{stars}
-    \frac{1}{\Delta \mu_i}} \ \ ,
+    \frac{1}{\Delta \mu_i^2}} \ \ ,
 \end{equation}
@@
 on the proper motion component reported by Gaia. The sums run over
-all the $N=646$ stars in our selected sample. We evaluate each proper motion
+all the $N=69$ stars in our selected sample. We evaluate each proper motion
 component separately. For simplicity, throughout this study, we assume the same
@@
 physical velocities. The data retrieved, and the ipython notebook used for the analysis
-presented here will be made available at \todo{probably git repo on bitbucket?}.
+presented here will be made available at \todo{git repo bitbucket?}.
 

### 2018-05-12 VFTS682.tex
updated numbers in draft

@@
 \subsection{New data from Gaia DR2  \label{data:gaia}}
+%% using
+%% /scratch/mathieu/MMPS/SCRIPTS_dirty/my_script/binaries/binary_c/30Dor/vfts682/data/asu.tsv to find average
 
@@
     \caption{Astrometric parameters for VFTS682. }
-%
   \begin{tabular}{l|c|c}
@@
   yr^{-1}}$. At the distance to the LMC, $0.1\mathrm{mas\
-  yr^{-1}}\simeq25\,\kms$ \citep[e.g.,][]{platais:18}, therefore this
-corresponds to requiring transverse velocities smaller than $\sim$$17\kms$. We
-remove obvious foreground stars by requiring parallaxes smaller than $0.1\,\mathrm{mas\ yr^{-1}}$.
+  yr^{-1}}\simeq25\,\kms$ \citep[e.g.,][]{platais:18}. We
+remove obvious foreground stars by requiring the absolute value of
+the reported parallaxes smaller than $0.1\,\mathrm{mas\ yr^{-1}}$.
 % At the distance to the
@@
 % not reduces further our sample.
-This selection yields a sample of 69 stars.% ,
+This selection yields a sample of 66 stars.% ,
 
@@
 on the proper motion component reported by Gaia. The sums run over
-all the $N=69$ stars in our selected sample. We evaluate each proper motion
+all the $N=66$ stars in our selected sample. We evaluate each proper motion
 component separately. For simplicity, throughout this study, we assume the same
@@
 subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
-$\langle\mu_\mathrm{RA}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
-$\langle\mu_\mathrm{DEC}\rangle =
-0.672\pm0.003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
+$\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
+$\langle\mu_\mathrm{DEC}\rangle =0.6986\pm0.0003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
 in good agreement with what is found by \cite{lennon:18}. Subtracting these values from the
@@
 components of proper motion of the star relative to the surrounding region
-$\mu_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu_\mathrm{DEC} =
-0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$. We note that the error budget is
+$\mu_\mathrm{RA} = 0.12\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu_\mathrm{DEC} =
+0.08\pm 0.08\,\mathrm{mas\ yr^{-1}}$. We note that the error budget is
 dominated by the errors on the proper motion components of VFTS682.
@@
 These can be converted into
-the components of the relative transverse velocity $\delta v_\mathrm{RA}=38\pm17\,\kms$,
-$\delta v_\mathrm{DEC}=26\pm19\,\kms$, assuming a distance of
+the components of the relative transverse velocity $\delta v_\mathrm{RA}=28\pm16\,\kms$,
+$\delta v_\mathrm{DEC}=20\pm19\,\kms$, assuming a distance of
 50\,kpc. % (we do not account for the uncertainty in the distance
@@
     +\left(\delta v_\mathrm{DEC}\right)^2+\left(\delta
-      v_\mathrm{rad}\right)^2} = 55 \pm 18 
+      v_\mathrm{rad}\right)^2} = 46 \pm 18 
   \kms \ .
@@
 and \cite{banerjee:12}.
-
-%\subsection{Kinematic age}
 

### 2018-05-12 VFTS682.tex
updated numbers in draft done

@@
 cluster R136 with a relative
-projected two-dimensional speed of about $\sim$$40\pm20 \kms$velocity,
-implying a three-dimensional speed of about $\sim$$60\pm20\kms$ when
+projected two-dimensional speed of about $\sim$$34\pm1 \kms$velocity,
+implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
 accounting for the velocity component along the line of sight.  The
@@
   \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} \simeq
-  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.6\pm\,0.2\, \mathrm{Myr} \ \ ,
+  \frac{29\,\mathrm{pc}}{34\,\kms} \simeq 0.85\pm\,0.02\, \mathrm{Myr} \ \ ,
 \end{equation}
@@
 the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel \equiv \sqrt{\left(\delta v_\mathrm{RA}\right)^2
-    +\left(\delta v_\mathrm{DEC}\right)^2} =46\pm
-18\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
+    +\left(\delta v_\mathrm{DEC}\right)^2} =34\pm
+1\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
 using the relative proper motion components calculated in
@@
 Based on our results, we claim that VFTS682 is the most massive
-runaway known to date, with a peculiar three-dimensional speed of $64\pm21\,\kms$. This means that isolated star formation is
+runaway known to date, with a peculiar three-dimensional speed of $46\pm18\,\kms$. This means that isolated star formation is
 \emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
-$0.6\pm0.2$\,Myr ago. Because of the exceptionally large mass
+$0.85\pm0.02$\,Myr ago. Because of the exceptionally large mass
 of this star, this raises the question of which stars must populate
@@
 the result of a dynamical ejection from the R136 cluster. 
-The value of $\tau_\mathrm{kin}\simeq0.63$\,Myr we find for VFTS682 (see \Secref{sec:r136_origin}) is smaller
+The value of $\tau_\mathrm{kin}\simeq0.85$\,Myr we find for VFTS682 (see \Secref{sec:r136_origin}) is smaller
 than the corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic age of

### 2018-05-12 VFTS682.tex
text done to end

@@
 
-Using the 651 stars selected as described in \Secref{data:gaia} (a
+Using the stars selected as described in \Secref{data:gaia} (a
 subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
 $\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
-$\langle\mu_\mathrm{DEC}\rangle =0.6986\pm0.0003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
-in good agreement with what is found by \cite{lennon:18}. Subtracting these values from the
+$\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
+  yr^{-1}}$. These values are
+in good agreement with what is found by \cite{lennon:18}. However, we
+emphasize their sensitivity to the choice of stars adopted to define
+the reference frame. Subtracting these values from the
 proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
 components of proper motion of the star relative to the surrounding region
-$\mu_\mathrm{RA} = 0.12\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu_\mathrm{DEC} =
-0.08\pm 0.08\,\mathrm{mas\ yr^{-1}}$. We note that the error budget is
+$\delta\mu_\mathrm{RA} = 0.12\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\delta\mu_\mathrm{DEC} =
+0.08\pm 0.08\,\mathrm{mas\ yr^{-1}}$. The error budget is
 dominated by the errors on the proper motion components of VFTS682.
@@
 \label{sec:discussion}
+\todo{subsection on validation with HST data?Discussion on
+  correlations between pmRA and pmDEC?Not sure where to start. HST
+  data for VFTS682 dont seem to be public.}
 \todo{Selma: Should add caveat that it may be a spurious detection.  That reconfirmation in following epochs is warranted. }
@@
 Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
-massive star among those interacting \todo{ref}. This means that, just
+massive star among those interacting \cite[e.g.,][]{banerjee:12}. This means that, just
 based on the kinematic properties of VFTS682, we would expect several
@@
 \citep[][]{maeder:00,demink:09,marchant:16}, and (ii) it has not
-accreted mass from a binary companion, nor will it, since the multi-epoch
+accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
 data of the VFTS survey rule out the presence of a close companion at the
@@
 
+\citet{banerjee:12} used N-body simulations of fully segregated
+clusters with all massive stars in binaries to suggest that VFTS682
+was ejected from R136, in agreement with our findings. They
+demonstrated that the cluster potential does not significantly change
+the velocity of the star after the ejection. In their
+model, they relied on (dynamically driven) stellar mergers to explain the high masses of
+VFTS682 and the massive members of R136. While it is not possible to
+robustly exclude that VFTS682 is itself a merger, as already
+mentioned, its spectrum does not show any of the obvious signatures
+like fast rotation. To eject such massive object, the cluster must have produced
+a large number of massive runaways. Indeed, several relatively
+isolated massive stars are observed in the region, some with known
+large radial velocities. A comprehensive study of the kinematic
+properties of all the massive stars surrounding R136 might shed light
+on whether some can be unequivocally identified as merger products. It
+is also possible that the star or binary that caused the ejection of
+VFTS682 might have been ejected in the opposite direction, and is also
+in isolation at present day.
+
 The similarities between VFTS682 and the WNh5 stars in the core of
@@
 by. Interpreting our results for VFTS682 through the lens of their simulations
-suggests the presence of a binary with total mass
+suggests the presence of a close binary with total mass
 $M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. Such bully
-binary could be R145 according to \cite{fujii:11}.
+binary could be R145 according to \cite{fujii:11}, and it might be an
+ideal observational candidate for a dynamically formed progenitor system of
+a binary black-hole, provided that stars this massive can avoid a
+pair-instability supernova \cite[][]{fowler:64, rakavy:67} at LMC
+metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS682 could be either a
+pair-instability supernova without compact remnant formation, or
+possibly direct collapse to a black hole above the $2^\mathrm{nd}$
+mass gap. The amount of mass loss of these stars will determine their final core
+mass and thus their final fate.
 
 The kinematic age of VFTS682 puts an
-upperlimit to the timescale to form such a ``bully binary'' in
+upperlimit to the timescale to form the ``bully binary'' in
 R136. The cluster must have been at the very beginning of its
 evolution, given the age estimate of $\lesssim 2$\,Myr
-\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS16. If the
+\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS682. If the
 cluster is indeed younger than the shortest stellar lifetime
@@
 
-\todo{check \cite{banerjee:12} and rewrite next paragraph, they
-  have stuff} Unfortunately, their simulations did not include stars more
-massive than $100\,M_\odot$, so it is difficult to predict, based on
-the existence of VFTS682, how many other stars should have been
-ejected from the cluster already and what their mass
-distribution is. A comprehensive study of the kinematic properties of the
-large sample of stars with extremely large inferred masses visible
-around R136 is encouraged.
+The variability of VFTS682, reminiscent of LBV stars, also suggests
+that VFTS682 (and therefore its analogs in the core of R136) might
+experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
+debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
+claim that LBV stars are typically isolated form O-type stars. The fact that VFTS682 is a dynamically
+ejected runaway which might evolve into an LBV star suggests that
+dynamical ejections can also play a role in explaining the apparent
+isolation of at least some LBV stars.
 
-\cite{lennon:18} have carried out a similar study on VFTS16, another
-massive star ($\sim90\,M_\odot$) in the 30 Doradus region, previously
+\citet{lennon:18} carried out a study similar to ours on VFTS16, another
+massive star ($\sim$$90\,M_\odot$) in the 30 Doradus region, previously
 known to be a runaway from the value of
@@
 
-\begin{itemize}
-\item is R136 a single young cluster or a merger
-\item estimate the influence of the gravitational potential of R136,
-  what is its total mass and relaxation time?
-\item Make link to dynamical interaction proposed to form BHBH
-  systems.
-  \item from Fabian's email (see comment in tex):% One last aspect: I am always wondering what happens to the star/binary that ejected 682.
-There is a huge linear momentum in this star of ~150 Msun * 60 km/s.
-Even if the bully was a 300 Msun total object, it would obtain a kick of 30 km/s.
-Is this enough to eject the bully star?
-This velocity would already be 100 km/s for a 90 Msun bully (I was thinking of 016 here...
-'Cause I do not know what ejected this object either).
-Of course, if there is a third body, these velocities are no longer correct.
-The escape velocity from R136 at 0.5pc for a total mass of 10^5 Msun would only be of order 20 km/s...
-So a 300 Msun bully is at the edge of what is possible and one might need something even more massive or another very massive object will also be ejected.
-Or am I wrong here?
-Do the Banerjee and Co simulations say something about this?
-\end{itemize}
+The numerical simulations from \cite{oh:16} suggest that dynamical
+interaction eject the majority of the stars during of shortly after the cluster
+core-collapse. The large number of isolated massive stars around it
+suggest that R136 has already evolved past the
+time of maximum stellar density. This might have implications for the
+question on whether the cluster formed via a monolithic collapse, or
+as a (potentially ongoing) merger of several sub-clusters.
 
-% \todo{Merge Selma's my\_bib bibtex file with Mathieu's and prevent
-% double refs. } %% copied everything in one file
+\citet{oh:16} also showed that the mass
+function of the ejected star depends on the cluster initial conditions
+(whether it is segregated, what is the binary fraction, what period
+distribution is assumed for the initial binaries), therefore studies on
+the population of isolated massive stars in the surroundings of R136
+might shed light on its initial stellar population and dynamical
+state.
+
+VFTS682 is the most massive runaway known to date, and its ejection
+from the cluster R136 implies that it is only the ``tip of the
+iceberg'' of possibly extremely massive runaways in the
+region. Studies of this population, enabled by recent Hubble Space
+Telescope and Gaia observations will put constraints on the evolution
+of these extreme stars, together with the formation and evolution of
+the central cluster itself.
 

### 2018-05-12 VFTS682.tex
minor here and there

@@
 cluster R136 with a relative
-projected two-dimensional speed of about $\sim$$34\pm1 \kms$velocity,
+projected two-dimensional speed of about $\sim$$34\pm1 \kms$,
 implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
@@
 %
-This would make VFTS 682 the most massive runaway star known to date.
+This makes VFTS 682 the most massive runaway star known to date.
 It proves that star clusters are capable of ejecting their most massive members.
@@
 
-\todo{recheck all numbers}
+%\todo{recheck all numbers}
 \section{Introduction}
@@
 
+% don't know where to go from here without giving away the main
+% result...Stephen wants to add a citation to fujii and banerjee in
+% the intro, which I also think would be fair...
+% The case of VFTS682 has also spawned significant theoretical efforts
+% to explain its isolation via dynamical ejection from R136, e.g.,
+% \citet{fujii:11} and \citet{banerjee:12}. Both these studies suggested
+
 Recently, \citet{platais:15,platais:18} used multi-epoch Hubble Space
@@
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) -69:04:26.72 \citep[][]{evans:11}
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} \citep[][]{evans:11}
 was originally classified as a young stellar object \citep{gruendl:09}
@@
 
-\citet{oh:16} also showed that the mass
-function of the ejected star depends on the cluster initial conditions
+\citet{oh:16} also showed that the mass and velocity distribution of the ejected star depends on the cluster initial conditions
 (whether it is segregated, what is the binary fraction, what period

### 2018-05-13 VFTS682.tex
fixed bibtex entries without arxiv number
added comments from Vincent, added author

@@
 
-\title{Gaia identifies VFTS682 as the most massive % , dynamically
+\title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
                                 % ejected %% to fit in one line
@@
 
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, J.
-Bestenlehner$^{3}$, S.~Justham$^{4,5}$, F.~R.~Schneider$^6$, R.~P.~van~der~Marel$^{7,8}$, J.~Vink$^{9}$ \SdM{et al.?} \\ \SdM{[ Alphabetical after Lennon.
-Possibly add further last-minute coauthors]},  }
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
+  R.~P.~van~der~Marel$^{3,4}$, J. Bestenlehner$^{5}$, V.~H\'enault-Brunet$^{6}$, S.~Justham$^{7,8}$, F.~R.~Schneider$^9$, J.~Vink$^{10}$}
 
@@
 
-\affil{$^{1}$Astronomical Institute Anton Pannekoek, University of
+\affil{$^{}$Astronomical Institute Anton Pannekoek, University of
     Amsterdam, 1098 XH Amsterdam, The Netherlands \\
-    $^{2}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78,
+    $^{}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78,
     E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
-    $^3$ {Department of Physics and Astronomy, Hicks Building,
+      $^{}${Department of Physics and Astronomy, Hicks Building,
       Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
-   $^{4}$ {School of Astronomy \& Space Science, University of the Chinese
+     $^{}$ {School of Astronomy \& Space Science, University of the Chinese
    Academy of Sciences, Beijing 100012, China}\\
-    $^5$ {National Astronomical Observatories, Chinese Academy of
+    $^{}$ {National Astronomical Observatories, Chinese Academy of
       Sciences, Beijing 100012, China}\\
-     $^6$ {Department of Physics, University of Oxford, Keble Road,
+     $^{}$ {Department of Physics, University of Oxford, Keble Road,
        Oxford OX1 3RH, UK} \\
-     $^{7}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+    $^{}$ {Space Telescope Science Institute, 3700 San Martin Drive,
        Baltimore, MD 21218, USA}\\
-     $^{8}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-     $^9$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}
+      {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
+     $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}
+     $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada}     
 }
@@
 %
-Previous spectroscopic studies identified VFTS682 as one of the most massive stars known.
+Previous spectroscopic studies identified VFTS 682 as one of the most massive stars known.
 It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 29\,pc of the star cluster R136.
@@
 %
-We use the unprecedented proper motions from the second Gaia data
+We use the unprecedented proper motions from the second \emph{Gaia} data
 release (DR2) to investigate the kinematics of this object. We show
@@
 
-\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS682}
+\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS 682}
 \maketitle{}
@@
 \cite{bestenlehner:11} and \cite{bressert:12}, who discuss the case of the very massive star
-VFTS682.   This star is located in the field of the 30 Doradus region
+VFTS 682.  This star is located in the field of the 30 Doradus region
 in the Large Magellanic Cloud (LMC) and was studied as part of the
@@
 
-Most remarkable is the location of VFTS682, at a projected distance of
+Most remarkable is the location of VFTS 682, at a projected distance of
 $\sim$$29$\,pc from the core of the star cluster
@@
 % the intro, which I also think would be fair...
-% The case of VFTS682 has also spawned significant theoretical efforts
+% The case of VFTS 682 has also spawned significant theoretical efforts
 % to explain its isolation via dynamical ejection from R136, e.g.,
@@
 the 30 Doradus region based on proper motion measurements.
-In this study we report on our analysis of the data from the second Gaia data release \cite[DR2,][]{gaia:16,brown:18}.
-We combine the radial velocity measurements from the VFTS survey \citep[][]{evans:11} with the proper motion from Gaia DR2 to reconstruct the three-dimensional velocity of VFTS682, and test the hypothesis that this star was ejected from R136.
+In this study we report on our analysis of the data from the second \emph{Gaia} data release \cite[DR2,][]{gaia:16,brown:18}.
+We combine the radial velocity measurements from the VFTS survey \citep[][]{evans:11} with the proper motion from \emph{Gaia} DR2 to reconstruct the three-dimensional velocity of VFTS 682, and test the hypothesis that this star was ejected from R136.
 
@@
 
-\subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
+\subsection{ Overview of VFTS 682 from previous studies \label{data:vfts683}}
 
-The star VFTS682  \citep{evans:11}, located at right ascension (RA)
+The star VFTS 682  \citep{evans:11}, located at right ascension (RA)
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
@@
 $\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citet{schneider:18} estimated
-a present-day mass for VFTS682 of $137.8^{+27.5}_ {-15.9}\Msun$, an
+a present-day mass for VFTS 682 of $137.8^{+27.5}_ {-15.9}\Msun$, an
 apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
@@
 \citep{schneider:17}, based on the evolutionary models from
-\citet{brott:11} and \cite{kohler:15}. These values place VFTS682 around the
+\citet{brott:11} and \cite{kohler:15}. These values place VFTS 682 around the
 ``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
@@
 at least two more very massive WN5h stars, R136a1 and R136a2, whose
-estimated current masses are even higher. However, VFTS682 stands
+estimated current masses are even higher. However, VFTS 682 stands
 out by its isolation at a projected offset of 29\,pc from the center
@@
 
-We obtain the peculiar radial velocity $\delta v_\mathrm{rad}$ of VFTS682 as the difference
+We obtain the peculiar radial velocity $\delta v_\mathrm{rad}$ of VFTS 682 as the difference
 between the average radial velocity of the 30 Doradus region
 ($270\pm10\kms$) minus the radial velocity measured from the NV $\lambda4944$
-line for VFTS682  \citep[$300\pm10\kms$, ][]{bestenlehner:11}. We implicitly assume that the use
+line for VFTS 682  \citep[$300\pm10\kms$, ][]{bestenlehner:11}. We implicitly assume that the use
 of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not
@@
 
-\subsection{New data from Gaia DR2  \label{data:gaia}}
+\subsection{New data from \emph{Gaia} DR2  \label{data:gaia}}
 %% using
@@
 \begin{center}
-    \caption{Astrometric parameters for VFTS682. }
+    \caption{Astrometric parameters for VFTS 682. }
   \begin{tabular}{l|c|c}
@@
     \tableline
-    RA \hfill[degree] &  \phantom{-}84.73 $\pm$  0.03 & \multirow{4}{*}{Gaia DR2}\\
+    RA \hfill[degree] &  \phantom{-}84.73 $\pm$  0.03 & \multirow{4}{*}{\emph{Gaia} DR2}\\
     DEC \hfill [degree] & -69.07 $\pm$  0.05  & \\
@@
   \end{tabular}
-  \tablecomments{We truncate the errors on RA and DEC for the purpose
-    of display. Readers are encouraged to consult the Gaia DR2 catalog
+  \tablecomments{We truncate the decimals on RA and DEC and the
+    relative errors for the purpose
+    of display. Readers are encouraged to consult the \emph{Gaia} DR2 catalog
   for accurate values.}
@@
 
-The Gaia Data Release 2, which became available on 25 April 2018,
+The \emph{Gaia} Data Release 2, which became available on 25 April 2018,
 provides the five-parameter astrometric solution (positions,
 parallaxes, and proper motion components) for more than a billion
-sources \citep{brown:18}. VFTS682 is identified with the id 4657685637907503744 in the Gaia DR2
-catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Gaia
+sources \citep{brown:18}. VFTS 682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2
+catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. \emph{Gaia}
 reports a G-band magnitude of 15.65. The star has a
@@
 \citep[][]{lindengren:18}, and the reported
-\texttt{astrometric\_excess\_noise} = 0. These values suggest that the Gaia
-data for VFTS682 are reliable. However, the effective temperature
-reported in Gaia DR2 is one order of magnitude lower than what found by
+\texttt{astrometric\_excess\_noise} = 0. These values suggest that the \emph{Gaia}
+data for VFTS 682 are reliable. However, the effective temperature
+reported in \emph{Gaia} DR2 is one order of magnitude lower than what found by
 \cite{bestenlehner:11}, and the best fit parallax of this star is
@@
 to the large distance to the LMC. Our main findings do not rely on the
-parallax nor the effective temperature values reported in the Gaia DR2
+parallax nor the effective temperature values reported in the \emph{Gaia} DR2
 catalog.
 
-We retrieve for VFTS682 the position in RA and DEC
+We retrieve for VFTS 682 the position in RA and DEC
 in the IGCS frame \cite[][]{brown:18}, its
 proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
-respectively). For the radial velocity of VFTS682 and of the 30 Doradus
+respectively). For the radial velocity of VFTS 682 and of the 30 Doradus
 region as a whole, we instead use the VFTS data
@@
 We define a local standard frame of reference to derive the peculiar velocity
-of VFTS682 with respect to the cluster R136 by selecting from the Gaia DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
+of VFTS 682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
 We select all the stars in a target of 0.05 degrees around R136
@@
 where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
-on the proper motion component reported by Gaia. The sums run over
+on the proper motion component reported by \emph{Gaia}. The sums run over
 all the $N=66$ stars in our selected sample. We evaluate each proper motion
@@
 
-\section{The kinematics of VFTS682}
+\section{The kinematics of VFTS 682}
 \label{sec:results}
@@
 \label{sec:runaway}
-We first address the question of whether VFTS682 is a typical star
+We first address the question of whether VFTS 682 is a typical star
 from the kinematic point of view, or whether it is a runaway star with
@@
 the reference frame. Subtracting these values from the
-proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
+proper motion of VFTS 682 (see \Tabref{tab:vfts682}), we obtain the
 components of proper motion of the star relative to the surrounding region
@@
 0.08\pm 0.08\,\mathrm{mas\ yr^{-1}}$. The error budget is
-dominated by the errors on the proper motion components of VFTS682.
+dominated by the errors on the proper motion components of VFTS 682.
 
@@
 \end{equation}
-This value for the three-dimensional speed of VFTS682 with respect the
+This value for the three-dimensional speed of VFTS 682 with respect the
 surrounding stars make it the most massive ``bona fide'' runaway star
@@
 
-The red arrow in \Figref{fig:main} shows the proper motion of VFTS682
+The red arrow in \Figref{fig:main} shows the proper motion of VFTS 682
 relative to the region, and the lighter red arrows show the possible
@@
 \end{equation}
-where $d_\parallel =29$\,pc is the projected distance from VFTS682 to
+where $d_\parallel =29$\,pc is the projected distance from VFTS 682 to
 the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel \equiv \sqrt{\left(\delta v_\mathrm{RA}\right)^2
@@
 
-\begin{figure}[htbp]
+\begin{figure}[tbp]
   \centering
   \includegraphics[width=0.48\textwidth]{./figures/main_plot_good}  
-  \caption{The red solid arrow indicates the proper motion of VFTS682
+  \caption{The red solid arrow indicates the proper motion of VFTS 682
     relative to the region, starting from the present day position of
@@
   correlations between pmRA and pmDEC?Not sure where to start. HST
-  data for VFTS682 dont seem to be public.}
+  data for VFTS 682 dont seem to be public.}
 \todo{Selma: Should add caveat that it may be a spurious detection. That reconfirmation in following epochs is warranted. }
 
-Based on our results, we claim that VFTS682 is the most massive
+Based on our results, we claim that VFTS 682 is the most massive
 runaway known to date, with a peculiar three-dimensional speed of $46\pm18\,\kms$. This means that isolated star formation is
@@
 massive star among those interacting \cite[e.g.,][]{banerjee:12}. This means that, just
-based on the kinematic properties of VFTS682, we would expect several
+based on the kinematic properties of VFTS 682, we would expect several
 stars with initial masses larger than $\sim$$150\,M_\odot$ in the
@@
   of the optically thick wind screening the surface of the star, and should be
-  considered with caution.} of VFTS682
+  considered with caution.} of VFTS 682
 reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
 line with the average rotation rate of massive stars in the region
-\citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
+\citep[][]{ramirez-agudelo:15}. This suggests that VFTS 682 (i) has not
 experienced rotationally induced chemically homogeneous evolution
@@
 data of the VFTS survey rule out the presence of a close companion at the
-present day. Moreover, the spectral type of VFTS682
+present day. Moreover, the spectral type of VFTS 682
 \citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
@@
 the spectrum of R136a3. Therefore,
-VFTS682 might be an ideal target to constrain the stellar physics of
+VFTS 682 might be an ideal target to constrain the stellar physics of
 stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
@@
 \citet{banerjee:12} used N-body simulations of fully segregated
-clusters with all massive stars in binaries to suggest that VFTS682
+clusters with all massive stars in binaries to suggest that VFTS 682
 was ejected from R136, in agreement with our findings. They
@@
 model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS682 and the massive members of R136. While it is not possible to
-robustly exclude that VFTS682 is itself a merger, as already
+VFTS 682 and the massive members of R136. While it is not possible to
+robustly exclude that VFTS 682 is itself a merger, as already
 mentioned, its spectrum does not show any of the obvious signatures
@@
 is also possible that the star or binary that caused the ejection of
-VFTS682 might have been ejected in the opposite direction, and is also
+VFTS 682 might have been ejected in the opposite direction, and is also
 in isolation at present day.
 
-The similarities between VFTS682 and the WNh5 stars in the core of
+The similarities between VFTS 682 and the WNh5 stars in the core of
 R136 are also in agreement with the ``bully binary'' model of
@@
 massive binary, which then tightens its orbit by ejecting other stars passing
-by. Interpreting our results for VFTS682 through the lens of their simulations
+by. Interpreting our results for VFTS 682 through the lens of their simulations
 suggests the presence of a close binary with total mass
@@
 pair-instability supernova \cite[][]{fowler:64, rakavy:67} at LMC
-metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS682 could be either a
+metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS 682 could be either a
 pair-instability supernova without compact remnant formation, or
@@
 
-The kinematic age of VFTS682 puts an
-upperlimit to the timescale to form the ``bully binary'' in
+The kinematic age of VFTS 682 puts an
+upper limit to the timescale to form the ``bully binary'' in
 R136. The cluster must have been at the very beginning of its
 evolution, given the age estimate of $\lesssim 2$\,Myr
-\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS682. If the
+\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS 682. If the
 cluster is indeed younger than the shortest stellar lifetime
 \citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, then the alternative
-explanation for ejection of VFTS682 from a binary
-system is excluded since the region is too young for stars
+explanation for ejection of VFTS 682 from the disruption of a binary
+by a core-collapse event is excluded since the region is too young for stars
 to have experience core-collapse already.
 
-The variability of VFTS682, reminiscent of LBV stars, also suggests
-that VFTS682 (and therefore its analogs in the core of R136) might
+The variability of VFTS 682, reminiscent of LBV stars, also suggests
+that VFTS 682 (and therefore its analogs in the core of R136) might
 experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
 debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
-claim that LBV stars are typically isolated form O-type stars. The fact that VFTS682 is a dynamically
+claim that LBV stars are typically isolated form O-type stars. The fact that VFTS 682 is a dynamically
 ejected runaway which might evolve into an LBV star suggests that
@@
 the result of a dynamical ejection from the R136 cluster. 
-The value of $\tau_\mathrm{kin}\simeq0.85$\,Myr we find for VFTS682 (see \Secref{sec:r136_origin}) is smaller
+The value of $\tau_\mathrm{kin}\simeq0.85$\,Myr we find for VFTS 682 (see \Secref{sec:r136_origin}) is smaller
 than the corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic age of
 $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
-massive VFTS682 was ejected later than VFTS16 from the same cluster.
+massive VFTS 682 was ejected later than VFTS16 from the same cluster.
 
@@
 
-VFTS682 is the most massive runaway known to date, and its ejection
+VFTS 682 is the most massive runaway known to date, and its ejection
 from the cluster R136 implies that it is only the ``tip of the
@@
 region. Studies of this population, enabled by recent Hubble Space
-Telescope and Gaia observations will put constraints on the evolution
+Telescope and \emph{Gaia} observations will put constraints on the evolution
 of these extreme stars, together with the formation and evolution of
@@
   very useful discussions on the
-  Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
+  \emph{Gaia} DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
   SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
-  Council (ERC) (Grant agreement No. 715063).
-  This work has made use of data from the ESA space mission Gaia (\url{http://www.cosmos.esa.int/gaia}), processed by the Gaia Data Processing and Analysis Consortium (DPAC, \url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}).
-Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the Gaia Multilateral Agreement.
+  Council (ERC) (Grant agreement No. 715063). VHB acknowledges support from the NRC-Canada Plaskett Fellowship.
+  This work has made use of data from the ESA space mission \emph{Gaia} (\url{http://www.cosmos.esa.int/gaia}), processed by the \emph{Gaia} Data Processing and Analysis Consortium (DPAC, \url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}).
+Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \emph{Gaia} Multilateral Agreement.
   The background image of \Figref{fig:main} is based on observations

### 2018-05-13 VFTS682.tex
minor: re-added fujii and banerjee papers in intro with one sentence from Stephen

@@
 R136. \citet{bestenlehner:11} discuss two possibilities for the
-origin of VFTS 682.
-Quoting directly from their abstract: "(i) the star either formed in situ, which would have profound implications for the formation mechanism of massive stars, or (ii) VFTS 682 is a slow runaway star that originated from the dense cluster R136, which would make it the most massive runaway known to date."
+origin of VFTS 682. Quoting directly from their abstract: "(i) the
+star either formed in situ, which would have profound implications for
+the formation mechanism of massive stars, or (ii) VFTS 682 is a slow
+runaway star that originated from the dense cluster R136, which would
+make it the most massive runaway known to date." Based on their
+dynamical models, \citet{fujii:11} and \citet{banerjee:12} argued that
+ejection of VFTS682 from R136 would be natural. Testing the origin of
+VFTS682 is not only important for constraining isolated star
+formation; it should also help to improve understanding of early star-cluster dynamics.
 

### 2018-05-14 VFTS682.tex
minor
started shaping HST section. Requires some significant re-writing of Sec. 3 I think

@@
 components of proper motion of the star relative to the surrounding region
-$\delta\mu_\mathrm{RA} = 0.12\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\delta\mu_\mathrm{DEC} =
+$\delta\mu_\mathrm{RA}^{Gaia} = 0.12\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\delta\mu_\mathrm{DEC}^{Gaia} =
 0.08\pm 0.08\,\mathrm{mas\ yr^{-1}}$. The error budget is
@@
 These can be converted into
-the components of the relative transverse velocity $\delta v_\mathrm{RA}=28\pm16\,\kms$,
-$\delta v_\mathrm{DEC}=20\pm19\,\kms$, assuming a distance of
+the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=28\pm16\,\kms$,
+$\delta v_\mathrm{DEC}^{Gaia}=20\pm19\,\kms$, assuming a distance of
 50\,kpc. % (we do not account for the uncertainty in the distance
@@
   \label{eq:speed_around}
-  v_\mathrm{pec} = \sqrt{\left(\delta v_\mathrm{RA}\right)^2
-    +\left(\delta v_\mathrm{DEC}\right)^2+\left(\delta
-      v_\mathrm{rad}\right)^2} = 46 \pm 18 
+  v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
+    +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
+      v_\mathrm{rad}^{VFTS}\right)^2} = 46 \pm 18 
   \kms \ .
@@
 
+\section{Validation with HST proper motions}
+
+Recently, \citet{platais:18} used HST observations (based on GO-12499
+\& 13359 observations with WFC3/UVIS and on archival WFPC2 and
+GO-12499 with both cameras, \citealt{platais:15}) to determine the
+proper motion of massive stars in the 30 Doradus region. For
+large distances ($\gtrsim50$\,kpc), HST data can compete or even
+exceed Gaia data in quality.
+
+In the catalog from \citet{platais:18} they listed the proper motion
+components of VFTS682 relative to the field, providing a
+measure of the motion of this star completely independent from
+ours. HST images of VFTS682 are not saturated in any epoch \todo{?},
+thanks to the large extinction toward this star. They found proper
+motion components relative to the field
+$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
+  yr^{-1}}$ and
+$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
+  yr^{-1}}$.
+
+Since these measures are completely independent from ours, we can take
+an average weighted with $1/\delta \mu_i^2$ (cf.~notation in
+\Eqref{eq:mean}), obtaining $\delta\mu_\mathrm{RA}=\pm\,\mathrm{mas\
+  yr^{-1}}$ and $\delta\mu_\mathrm{DEC}=\pm\,\mathrm{mas\
+  yr^{-1}}$.
+
+% %% pure HST physical velocities
+% \todo{done in the train: to be checked} Converting these into physical velocities we obtain 
+% $\delta v_\mathrm{RA}=2.3\pm30\,\kms$, $\delta v_\mathrm{DEC}=47\pm\,23\kms$, corresponding to
+% $v_\parallel^\mathrm{HST} = 47\pm\,1\kms$.
+
+\todo{redo below with average HST+Gaia?Actually redo also calculation
+  of $v_\mathrm{pec}$...}
 \subsection{Does it come from the R136 cluster?}
@@
     are a subset of those we use to define the local rest
-    frame. \todo{load color picture on background}}
+    frame. \todo{load color picture on background}\todo{show
+      separately arrows fro Gaia/HST and combination? Or show only
+      combination? Right now it's Gaia only}}
 
@@
 \label{sec:discussion}
-\todo{subsection on validation with HST data?Discussion on
-  correlations between pmRA and pmDEC?Not sure where to start. HST
-  data for VFTS 682 dont seem to be public.}
 \todo{Selma: Should add caveat that it may be a spurious detection. That reconfirmation in following epochs is warranted. }

### 2018-05-14 VFTS682.tex
minor

@@
 \section{Validation with HST proper motions}
-
+\todo{maybe rename to Astrometry from HST and make sec. 2.1.1 or 2.2
+  but above Gaia data?}
 Recently, \citet{platais:18} used HST observations (based on GO-12499
 \& 13359 observations with WFC3/UVIS and on archival WFPC2 and
-GO-12499 with both cameras, \citealt{platais:15}) to determine the
-proper motion of massive stars in the 30 Doradus region. For
+GO-12499 with both cameras, see \citealt{platais:15}) to determine the
+proper motion of massive stars in the 30 Doradus region. Our star of
+interest is not saturated in any HST image \todo{?},
+because of the large extinction toward this star. For
 large distances ($\gtrsim50$\,kpc), HST data can compete or even
@@
 
-In the catalog from \citet{platais:18} they listed the proper motion
+The catalog from \citet{platais:18} lists the proper motion
 components of VFTS682 relative to the field, providing a
 measure of the motion of this star completely independent from
-ours. HST images of VFTS682 are not saturated in any epoch \todo{?},
-thanks to the large extinction toward this star. They found proper
+ours. They found proper
 motion components relative to the field
@@
 \todo{redo below with average HST+Gaia?Actually redo also calculation
-  of $v_\mathrm{pec}$...}
+  of $v_\mathrm{pec}$}
 \subsection{Does it come from the R136 cluster?}
@@
 velocity\footnote{However, the determination of the rotational
-  velocity for stars showing lines in emission is complicated because
-  of the optically thick wind screening the surface of the star, and should be
+  velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
   considered with caution.} of VFTS 682

### 2018-05-15 VFTS682.tex
errorbar on positions is wrong! units
Jorick's initials

@@
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  R.~P.~van~der~Marel$^{3,4}$, J. Bestenlehner$^{5}$, V.~H\'enault-Brunet$^{6}$, S.~Justham$^{7,8}$, F.~R.~Schneider$^9$, J.~Vink$^{10}$}
+  R.~P.~van~der~Marel$^{3,4}$, J. Bestenlehner$^{5}$,
+  A.~de~Koter$^{1}$, V.~H\'enault-Brunet$^{6}$, S.~Justham$^{7,8}$,
+  I.~Platais$^{9}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
 
@@
   \end{tabular}
-  \tablecomments{We truncate the decimals on RA and DEC and the
-    relative errors for the purpose
-    of display. Readers are encouraged to consult the \emph{Gaia} DR2 catalog
-  for accurate values.}
+  \tablecomments{\todo{the error on position is in milliarcsec, not
+      degrees! Fix this, Roeland is right!}}
     \end{center}

### 2018-05-15 VFTS682.tex
minor?

@@
   A.~de~Koter$^{1}$, V.~H\'enault-Brunet$^{6}$, S.~Justham$^{7,8}$,
-  I.~Platais$^{9}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
+  I.~Platais$^{9}$, H.~Sana$^{}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
 
@@
 cluster R136 with a relative
-projected two-dimensional speed of about $\sim$$34\pm1 \kms$,
+projected two-dimensional speed of about $\sim$$34\pm1 \kms$ \todo{fix
+  error!},
 implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
@@
   \tablecomments{\todo{the error on position is in milliarcsec, not
-      degrees! Fix this, Roeland is right!}}
+      degrees! Fix this, Roeland is right! Also, report relative pm,
+      not absolute}}
     \end{center}
@@
 ejected runaway which might evolve into an LBV star suggests that
-dynamical ejections can also play a role in explaining the apparent
+N-body interactions also play a role in explaining the apparent
 isolation of at least some LBV stars.

### 2018-05-16 VFTS682.tex
minor

@@
       {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-     $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}
+     $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
      $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada}     
@@
 cluster R136 with a relative
-projected two-dimensional speed of about $\sim$$34\pm1 \kms$ \todo{fix
-  error!},
+projected two-dimensional speed of about $\sim$$34\pm17\kms$ ,
 implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when

### 2018-05-16 VFTS682.tex
fixed table according to Danny's comments, added Danny's bit of text for HST description

@@
     \tableline
-    RA \hfill[degree] &  \phantom{-}84.73 $\pm$  0.03 & \multirow{4}{*}{\emph{Gaia} DR2}\\
-    DEC \hfill [degree] & -69.07 $\pm$  0.05  & \\
-    $\mu_\mathrm{RA}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}1.84 $\pm$ 0.07 & \\
-    $\mu_\mathrm{DEC}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & \phantom{-0}0.78 $\pm$ 0.08& \\
-    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & \phantom{-0}30 $\pm$ 20 & \cite{bestenlehner:11}\\
+    RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
+                      & \multirow{4}{*}{\emph{Gaia} DR2}\\[5pt]
+    DEC \hfill [degree] & -69.07 % $\pm$  0.05
+                      & \\[5pt]
+    \cline{1-2}
+    $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.12$\pm$0.07 & \\[5pt]
+    $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.08 & \\[5pt]
+    \hline
+    $\delta\mu_\mathrm{RA}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.01$\pm$0.13 & \multirow{2}{*}{\cite{platais:18}}\\[5pt]
+    $\delta\mu_\mathrm{DEC}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] &
+                                                                      0.2$\pm$0.1 &
+    \\[5pt]
+    \hline
+    $\delta v_\mathrm{rad}$  \hfill[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
     \tableline
   \end{tabular}
-  \tablecomments{\todo{the error on position is in milliarcsec, not
-      degrees! Fix this, Roeland is right! Also, report relative pm,
-      not absolute}}
+  \tablecomments{We do not list the error on the RA and DEC positions,
+    which can be retrieved from the Gaia DR2 catalog and are of order
+    $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$. We adopt the peculiar
+    radial velocity calculated using the  NV $\lambda4944$ line, which is
+    less sensitive to the wind velocity structure.}
     \end{center}
@@
   but above Gaia data?}
-Recently, \citet{platais:18} used HST observations (based on GO-12499
-\& 13359 observations with WFC3/UVIS and on archival WFPC2 and
-GO-12499 with both cameras, see \citealt{platais:15}) to determine the
-proper motion of massive stars in the 30 Doradus region. Our star of
-interest is not saturated in any HST image \todo{?},
-because of the large extinction toward this star. For
+
+Recently, \citet{platais:18} presented Hubble Space Telescope (HST)  astrometry and proper
+motions in the 30 Doradus region based on two epochs of
+observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
+in October 3-8 2011 and the second epoch (GO-13359) in
+October 6-11 2014, both using the F775W filter.
+The second epoch was designed to match the pointing and
+orientation of the first epoch as exactly as possible so
+that the displacement measurements could be purely differential.
+While most bright massive stars (V$<$14) were saturated in
+their data and missing from that catalog, due to its
+high extinction, E(B-V)=0.9, this star is faint enough
+to be included in that catalog. Its proper motions are
+listed in \Tabref{tab:vfts682} \todo{add}. For
 large distances ($\gtrsim50$\,kpc), HST data can compete or even
 exceed Gaia data in quality.
+
+% Recently, \citet{platais:18} used HST observations (based on GO-12499
+% \& 13359 observations with WFC3/UVIS and on archival WFPC2 and
+% GO-12499 with both cameras, see \citealt{platais:15}) to determine the
+% proper motion of massive stars in the 30 Doradus region. Our star of
+% interest is not saturated in any HST image \todo{?},
+% because of the large extinction toward this star.
 

### 2018-05-16 VFTS682.tex
minor

--- 2018-05-16 VFTS682.tex
@@
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  R.~P.~van~der~Marel$^{3,4}$, J. Bestenlehner$^{5}$,
-  A.~de~Koter$^{1}$, V.~H\'enault-Brunet$^{6}$, S.~Justham$^{7,8}$,
-  I.~Platais$^{9}$, H.~Sana$^{}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
+  R.~P.~van~der~Marel$^{3,4}$,   I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
+   V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
+ , H.~Sana$^{9}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
 
   % ,
-%  A.~de~Koter$^{1}$ \todo{... add coauthors...} }
+%
 
-\affil{$^{}$Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands \\
-    $^{}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78,
-    E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
+\affil{$^{1}${Astronomical Institute Anton Pannekoek, University of
+    Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
+    $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
+    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
+       $^{3}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+       Baltimore, MD 21218, USA}\\
+        $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
+\SdM{Check Affiliations numbers \\}
       $^{}${Department of Physics and Astronomy, Hicks Building,
       Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
+         $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada} 
      $^{}$ {School of Astronomy \& Space Science, University of the Chinese
@@
        Oxford OX1 3RH, UK} \\
-    $^{}$ {Space Telescope Science Institute, 3700 San Martin Drive,
-       Baltimore, MD 21218, USA}\\
-      {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-     $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
-     $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada}     
+       $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}   
 }
@@
 cluster R136 with a relative
-projected two-dimensional speed of about $\sim$$34\pm17\kms$ ,
+projected two-dimensional speed of about $\sim$$34\pm17 \kms$ \todo{Double Check
+  error!},
 implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when

### 2018-05-16 VFTS682.tex
minor

--- 2018-05-16 VFTS682.tex
@@
        Oxford OX1 3RH, UK} \\
+       $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}   
     $^{}$ {Space Telescope Science Institute, 3700 San Martin Drive,
@@
 cluster R136 with a relative
-projected two-dimensional speed of about $\sim$$34\pm17\kms$ ,
-implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
+projected two-dimensional speed of about $\sim$$34\pm17\kms$, implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
 accounting for the velocity component along the line of sight. The

### 2018-05-18 VFTS682.tex
dealt with Alex comments on the text

--- 2018-05-16 VFTS682.tex
@@
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  R.~P.~van~der~Marel$^{3,4}$, J. Bestenlehner$^{5}$,
-  A.~de~Koter$^{1}$, V.~H\'enault-Brunet$^{6}$, S.~Justham$^{7,8}$,
-  I.~Platais$^{9}$, H.~Sana$^{}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
+  R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{5}$, J. Bestenlehner$^{6}$,
+  A.~de~Koter$^{1}$, V.~H\'enault-Brunet$^{7}$, S.~Justham$^{8}$, H.~Sana$^{9}$, F.~R.~N.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
 
-  % ,
-%  A.~de~Koter$^{1}$ \todo{... add coauthors...} }
-
-\affil{$^{}$Astronomical Institute Anton Pannekoek, University of
+\affil{\todo{check affil. num}$^{}$Astronomical Institute Anton Pannekoek, University of
     Amsterdam, 1098 XH Amsterdam, The Netherlands \\
@@
 \begin{abstract}
-The formation and fates of the most massive stars remains one of the most elusive questions in stellar astrophysics. 
+The formation and fates of the most massive stars remains among the most elusive questions in stellar astrophysics. 
 %
-Previous spectroscopic studies identified VFTS 682 as one of the most massive stars known.
-It resides in the field of the 30 Dor region in the Large Magellanic Cloud at a projected distance of 29\,pc of the star cluster R136.
-The lack of a detected offset in radial velocity derived from the spectra, led to speculations whether such extreme stars can form in relative isolation.
+Previous spectroscopic studies identified VFTS 682 as one of the most
+massive stars known. It resides in the field of the 30 Dor region in
+the Large Magellanic Cloud at a projected distance of 29\,pc of the
+star cluster Radcliffe 136 (R136). The lack of a detected offset in radial velocity\todo{can we really say that?RV
+  is about 30kms...}
+derived from the spectra, led to speculations whether such extreme
+stars can form in relative isolation . 
 %
-We use the unprecedented proper motions from the second \emph{Gaia} data
+\todo{add mention of HST}We use the unprecedented proper motions from the second \emph{Gaia} data
 release (DR2) to investigate the kinematics of this object. We show
@@
 %\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
-evolve fast, typically reside in dense groups and they remain enshrouded in their parent cloud during the formation
+evolve fast, typically reside in dense groups and remain enshrouded in
+their parent cloud during the entirety of their formation
 process. Major progress has been made on the theoretical side,
@@
 therefore a key question for the present and upcoming transient
-surveys e.g., Large Synoptic Spectroscopic Survey and LIGO/Virgo,
+surveys, e.g., Large Synoptic Spectroscopic Survey and LIGO/Virgo,
 %\citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others. 
@@
 isolation,  far from dense clusters or OB associations. The origin of
-this population still remains debated \citep{lamb:16,ward:18}.  One
+this population still remains debated \citep{gvaramadze:12, lamb:16,ward:18}.  One
 hypothesis is that they  formed in the field, but this poses a
@@
 were ejected from their birth locations. Such ejections may result
-from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems by the core-collapse of the companion star \citep[e.g.,][]{zwicky:57, blaauw:61}.
+from dynamical interactions \citep[e.g.,][]{poveda:67} or from the
+disruption of binary systems by the core-collapse of the companion
+star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.
 
@@
 stars form in relative isolation was presented by
-\cite{bestenlehner:11} and \cite{bressert:12}, who discuss the case of the very massive star
+\cite{bestenlehner:11} and \cite{bressert:12}, who discussed the case of the very massive star
 VFTS 682.  This star is located in the field of the 30 Doradus region
@@
 \citep{schneider:18}. These correspond to an inferred initial mass of
-$150.0^{+28.7}_{-17.4}\,M_\odot$ and a wind mass loss rate of
-$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$ inferred from the
-spectral analysis not accounting for clumping
-\citep[][]{bestenlehner:11}, which make it one of the
+$150.0^{+28.7}_{-17.4}\,M_\odot$. It features a wind mass loss rate of
+$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, inferred from
+spectral analysis assuming a smooth outflow, i.e., not accounting for clumping
+\citep[][]{bestenlehner:11}. These properties make it one of the
 most extreme objects in the region. It is reminiscent of the very
@@
 runaway star that originated from the dense cluster R136, which would
-make it the most massive runaway known to date." Based on their
+make it the most massive runaway known to date". Based on their
 dynamical models, \citet{fujii:11} and \citet{banerjee:12} argued that
-ejection of VFTS682 from R136 would be natural. Testing the origin of
-VFTS682 is not only important for constraining isolated star
+ejection of VFTS 682 from R136 would be natural. Testing the origin of
+VFTS 682 is not only important for constraining isolated star
 formation; it should also help to improve understanding of early star-cluster dynamics.
@@
 Recently, \citet{platais:15,platais:18} used multi-epoch Hubble Space
-Telescope imaging to identify and shed new light on runaway stars from
-the 30 Doradus region based on proper motion measurements.
-In this study we report on our analysis of the data from the second \emph{Gaia} data release \cite[DR2,][]{gaia:16,brown:18}.
-We combine the radial velocity measurements from the VFTS survey \citep[][]{evans:11} with the proper motion from \emph{Gaia} DR2 to reconstruct the three-dimensional velocity of VFTS 682, and test the hypothesis that this star was ejected from R136.
+Telescope (HST) imaging to identify and shed new light on runaway stars from
+the 30 Doradus region based on proper motion
+measurements. \citet{lennon:18} used data from the second \emph{Gaia}
+data release \cite[DR2,][]{gaia:16,brown:18} to further investigate
+the isolated O-type stars in the vicinity of R136. In this study, we
+report on our analysis of the \emph{Gaia} DR2 data for VFTS 682 which
+was not part of the sample of \citet{lennon:18} because of its WNh5
+spectral type. We combine the radial velocity measurements from the
+VFTS survey \citep[][]{evans:11} with the proper motion from
+\emph{Gaia} DR2 to reconstruct the three-dimensional velocity of VFTS
+682, and test the hypothesis that this star was ejected from R136. We
+use the HST proper motions from \citet{platais:18} to validate our
+analysis of the \emph{Gaia} DR2 data.
 
@@
 
-The star VFTS 682  \citep{evans:11}, located at right ascension (RA)
+The star VFTS 682, located at right ascension (RA)
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} \citep[][]{evans:11}
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} \citep[][see also
+\Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2 in the
+ICGS frame]{evans:11}
 was originally classified as a young stellar object \citep{gruendl:09}
@@
 parameters and surface abundances. They inferred an extinction
-of $R_V\simeq4.7$, leading to the high luminosity
+of $A_V=4.45\pm0.12$, leading to the high luminosity
 $\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citet{schneider:18} estimated
@@
 apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
-of $150.0^{+28.7}_{-17.4}\Msun$, using the Bayesian analysis tool Bonnsai
+of $150.0^{+28.7}_{-17.4}\Msun$, using the Bayesian analysis tool BONNSAI
 \citep{schneider:17}, based on the evolutionary models from
-\citet{brott:11} and \cite{kohler:15}. These values place VFTS 682 around the
-``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
+\citet{brott:11} and \cite{kohler:15}. % These values place VFTS 682 around the
+% ``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
 
@@
 Further worth noticing is the variability of the
-star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,dex
+star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
 lower than what found by \citet{evans:11}. \citet{bestenlehner:11} discuss the Optical Gravitational Lensing
@@
 presence of an optically thick, likely non-homogeneous, and variable
-wind hiding the stellar surface.
+wind obscuring the stellar surface.
+
+\subsection{HST astrometry for VFTS682}
+Recently, \citet{platais:18} presented HST  astrometry and proper
+motions in the 30 Doradus region based on two epochs of
+observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
+in October 3-8 2011 and the second epoch (GO-13359) in
+October 6-11 2014, both using the F775W filter.
+The second epoch was designed to match the pointing and
+orientation of the first epoch as exactly as possible so
+that the displacement measurements could be purely differential.
+While most bright massive stars (V$<$14) were saturated in
+their data and missing from that catalog, due to its
+high extinction, E(B-V)=0.9, this star is faint enough
+to be included in that catalog. Its proper motions are
+listed in \Tabref{tab:vfts682}. For
+large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even
+exceed \emph{Gaia} data in quality.
+
+% Recently, \citet{platais:18} used HST observations (based on GO-12499
+% \& 13359 observations with WFC3/UVIS and on archival WFPC2 and
+% GO-12499 with both cameras, see \citealt{platais:15}) to determine the
+% proper motion of massive stars in the 30 Doradus region. Our star of
+% interest is not saturated in any HST image \todo{?},
+% because of the large extinction toward this star.
+
+The catalog from \citet{platais:18} lists the proper motion
+components of VFTS 682 relative to the field, providing a
+measure of the motion of this star completely independent from
+ours. They found proper
+motion components relative to the field
+$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
+  yr^{-1}}$ and
+$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
+  yr^{-1}}$.
+
+Since these measures are completely independent from ours, we can take
+an average weighted with $1/\delta \mu_i^2$ (cf.~notation in
+\Eqref{eq:mean}), obtaining $\delta\mu_\mathrm{RA}=\pm\,\mathrm{mas\
+  yr^{-1}}$ and $\delta\mu_\mathrm{DEC}=\pm\,\mathrm{mas\
+  yr^{-1}}$.
+
+% %% pure HST physical velocities
+% \todo{done in the train: to be checked} Converting these into physical velocities we obtain 
+% $\delta v_\mathrm{RA}=2.3\pm30\,\kms$, $\delta v_\mathrm{DEC}=47\pm\,23\kms$, corresponding to
+% $v_\parallel^\mathrm{HST} = 47\pm\,1\kms$.
+
+\todo{redo below with average HST+Gaia?Actually redo also calculation
+  of $v_\mathrm{pec}$}
 
@@
 We select all the stars in a target of 0.05 degrees around R136
-(NGC2070) fulfilling the following criteria. First, we require G-band
-magnitude brighter than 17, correspondingly roughly to the
+(NGC2070) fulfilling the following criteria. First, we require a G-band
+magnitude brighter than 17, corresponding roughly to the
 completeness level of the VFTS survey \citep[here we implicitly assume
 G$\sim$V,][]{evans:11}. Then we require \texttt{visibility\_period} $\geq$ 5,
-\texttt{astrometric\_excess\_noise} $< 1$, the errors on the proper
+\texttt{astrometric\_excess\_noise} $< 1$, and the errors on the proper
 motion components to be smaller than 0.07\,$\mathrm{mas\
@@
 presented here will be made available at \todo{git repo bitbucket?}.
+
+\todo{nebular lines to compare velocity of the gas?}
+\todo{fix numbers}
 
@@
 \end{equation}
-This value for the three-dimensional speed of VFTS 682 with respect the
-surrounding stars make it the most massive ``bona fide'' runaway star
+This value for the three-dimensional speed of VFTS 682 with respect to
+the surrounding stars make it the most massive ``bona fide'' runaway star
 known to date.
 
-\section{Validation with HST proper motions}
-\todo{maybe rename to Astrometry from HST and make sec. 2.1.1 or 2.2
-  but above Gaia data?}
-
-Recently, \citet{platais:18} presented Hubble Space Telescope (HST)  astrometry and proper
-motions in the 30 Doradus region based on two epochs of
-observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
-in October 3-8 2011 and the second epoch (GO-13359) in
-October 6-11 2014, both using the F775W filter.
-The second epoch was designed to match the pointing and
-orientation of the first epoch as exactly as possible so
-that the displacement measurements could be purely differential.
-While most bright massive stars (V$<$14) were saturated in
-their data and missing from that catalog, due to its
-high extinction, E(B-V)=0.9, this star is faint enough
-to be included in that catalog. Its proper motions are
-listed in \Tabref{tab:vfts682} \todo{add}. For
-large distances ($\gtrsim50$\,kpc), HST data can compete or even
-exceed Gaia data in quality.
-
-% Recently, \citet{platais:18} used HST observations (based on GO-12499
-% \& 13359 observations with WFC3/UVIS and on archival WFPC2 and
-% GO-12499 with both cameras, see \citealt{platais:15}) to determine the
-% proper motion of massive stars in the 30 Doradus region. Our star of
-% interest is not saturated in any HST image \todo{?},
-% because of the large extinction toward this star.
-
-The catalog from \citet{platais:18} lists the proper motion
-components of VFTS682 relative to the field, providing a
-measure of the motion of this star completely independent from
-ours. They found proper
-motion components relative to the field
-$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
-  yr^{-1}}$ and
-$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
-  yr^{-1}}$.
-
-Since these measures are completely independent from ours, we can take
-an average weighted with $1/\delta \mu_i^2$ (cf.~notation in
-\Eqref{eq:mean}), obtaining $\delta\mu_\mathrm{RA}=\pm\,\mathrm{mas\
-  yr^{-1}}$ and $\delta\mu_\mathrm{DEC}=\pm\,\mathrm{mas\
-  yr^{-1}}$.
-
-% %% pure HST physical velocities
-% \todo{done in the train: to be checked} Converting these into physical velocities we obtain 
-% $\delta v_\mathrm{RA}=2.3\pm30\,\kms$, $\delta v_\mathrm{DEC}=47\pm\,23\kms$, corresponding to
-% $v_\parallel^\mathrm{HST} = 47\pm\,1\kms$.
-
-\todo{redo below with average HST+Gaia?Actually redo also calculation
-  of $v_\mathrm{pec}$}
 \subsection{Does it come from the R136 cluster?}
@@
 As in the rest of this study, we neglect for
-simplicity the error on the distance estimates.
+simplicity the error on the distance estimates, which amounts to about
+$\sim$$2.2\%$, and is negligible compared to other uncertainties.
 
-The kinematic age $\tau_\mathrm{kin}$ is smaller than the apparent age
+The kinematic age $\tau_\mathrm{kin}$ is similar or smaller than the apparent age
 of the star of $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
@@
 three most massive stars detected in the core of the cluster by
-\cite{crowther:10}, with an astonishing similarity in particular with
+\cite{dekoter:97,crowther:10,crowther:16}, with an astonishing similarity in particular with
 the spectrum of R136a3. Therefore,
@@
 clusters with all massive stars in binaries to suggest that VFTS 682
-was ejected from R136, in agreement with our findings. They
+was ejected from R136. They
 demonstrated that the cluster potential does not significantly change
@@
 mentioned, its spectrum does not show any of the obvious signatures
-like fast rotation. To eject such massive object, the cluster must have produced
+like fast rotation. To eject such a massive object, the cluster is
+expected to have produced
 a large number of massive runaways. Indeed, several relatively
@@
 cluster is indeed younger than the shortest stellar lifetime
-\citep[$\sim$3\,Myr, e.g.,][]{zapartas:17}, then the alternative
+\citep[$\sim$3\,Myr, e.g.,][]{brott:11, zapartas:17}, then the alternative
 explanation for ejection of VFTS 682 from the disruption of a binary
 by a core-collapse event is excluded since the region is too young for stars
-to have experience core-collapse already.
+to have experienced core-collapse already.
 
-The variability of VFTS 682, reminiscent of LBV stars, also suggests
+The variability of VFTS 682, reminiscent of LBV stars, suggests
 that VFTS 682 (and therefore its analogs in the core of R136) might
@@
 The numerical simulations from \cite{oh:16} suggest that dynamical
-interaction eject the majority of the stars during of shortly after the cluster
+interaction eject the majority of the stars during or shortly after the cluster
 core-collapse. The large number of isolated massive stars around it
@@
 \citet{oh:16} also showed that the mass and velocity distribution of the ejected star depends on the cluster initial conditions
-(whether it is segregated, what is the binary fraction, what period
-distribution is assumed for the initial binaries), therefore studies on
+(whether it is segregated, its primordial binary fraction and initial period
+distribution of the binary population), therefore studies on
 the population of isolated massive stars in the surroundings of R136
@@
 VFTS 682 is the most massive runaway known to date, and its ejection
-from the cluster R136 implies that it is only the ``tip of the
+from the cluster R136 likely implies that it is only the ``tip of the
 iceberg'' of possibly extremely massive runaways in the
-region. Studies of this population, enabled by recent Hubble Space
-Telescope and \emph{Gaia} observations will put constraints on the evolution
+region. Studies of this population, enabled by recent HST and \emph{Gaia} observations will put constraints on the evolution
 of these extreme stars, together with the formation and evolution of

### 2018-05-18 VFTS682.tex

@@
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{5}$, J. Bestenlehner$^{6}$,
-  A.~de~Koter$^{1}$, V.~H\'enault-Brunet$^{7}$, S.~Justham$^{8}$, H.~Sana$^{9}$, F.~R.~N.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
-
-\affil{\todo{check affil. num}$^{}$Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands \\
-    $^{}$ ESA, European Space Astronomy Centre, Apdo. de Correos 78,
-    E-28691 Villanueva de la Ca\~nada, Madrid, Spain \\
+  R.~P.~van~der~Marel$^{3,4}$,   I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
+   V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
+ , H.~Sana$^{9}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
+\affil{$^{1}${\todo{check affil. num}Astronomical Institute Anton Pannekoek, University of
+    Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
+    $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
+    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
+       $^{3}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+       Baltimore, MD 21218, USA}\\
+        $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
       $^{}${Department of Physics and Astronomy, Hicks Building,
       Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
+         $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada} 
      $^{}$ {School of Astronomy \& Space Science, University of the Chinese

### 2018-05-22 VFTS682.tex
tuned down abstract

@@
 
-\title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
-                                % ejected %% to fit in one line
-  runaway star known to date}
-
+% \title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
+%                                 % ejected %% to fit in one line
+%   runaway star known to date}
+\title{\emph{Gaia} DR2 adds clues for dynamical ejection of
+  a $\sim$$150\,M_\odot$ star}
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
@@
 \begin{abstract}
-The formation and fates of the most massive stars remains among the most elusive questions in stellar astrophysics. 
+The formation and fates of the most massive stars remain among the most elusive questions in stellar astrophysics. 
 %
@@
 the Large Magellanic Cloud at a projected distance of 29\,pc of the
-star cluster Radcliffe 136 (R136). The lack of a detected offset in radial velocity\todo{can we really say that?RV
-  is about 30kms...}
-derived from the spectra, led to speculations whether such extreme
-stars can form in relative isolation . 
+star cluster Radcliffe 136 (R136). The inconclusiveness of the radial
+velocity analysis for this emission line star led to speculations whether such extreme
+stars can form in relative isolation. 
 %
-\todo{add mention of HST}We use the unprecedented proper motions from the second \emph{Gaia} data
-release (DR2) to investigate the kinematics of this object. We show
-that VFTS 682 is a bonafide runaway star moving away from the star
-cluster R136 with a relative
-projected two-dimensional speed of about $\sim$$34\pm17\kms$, implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
-accounting for the velocity component along the line of sight. The
-inferred age, distance, and the magnitude and direction of the
-velocity are consistent with a dynamical ejection scenario from the cluster. 
+We use the unprecedented proper motions from the second \emph{Gaia} data
+release (DR2) together with multiepoch \emph{Hubble Space Telescope} proper motions to investigate the kinematics of this object. We show
+that the projected velocity of VFTS 682 is consistent with it being a
+runaway star dynamically ejected from R136. Its
+projected two-dimensional speed relative to the cluster is $\sim$$34\pm17\kms$, implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
+accounting for the (uncertain) velocity component along the line of sight. The
+inferred kinematic age is also  consistent with the dynamical ejection scenario from the cluster. 
 %
-This makes VFTS 682 the most massive runaway star known to date.
+If confirmed by future astrometric data releases, this would make VFTS 682 the most massive runaway star known to date.
 It proves that star clusters are capable of ejecting their most massive members.

### 2018-05-22 VFTS682.tex
tuned down intro

@@
 identified by
-\citet{crowther:10, crowther:16} in the core of the
+\citet{dekoter:97,crowther:10, crowther:16} in the core of the
 R136 cluster.
@@
 
-Recently, \citet{platais:15,platais:18} used multi-epoch Hubble Space
-Telescope (HST) imaging to identify and shed new light on runaway stars from
+Recently, \citet{platais:15,platais:18} used multi-epoch \emph{Hubble Space
+Telescope} (HST) imaging to identify and shed new light on runaway stars from
 the 30 Doradus region based on proper motion
@@
 
-Our results indicate that VFTS 682 is a bona fide runaway star. The
+Our results indicate that VFTS 682 is a runaway star, although with
+low statistical significance. The
 direction and magnitude of the velocity vector are consistent with
@@
 analysis \citep[from][]{schneider:18} is consistent with the travel
-time we calculate for this star. Therefore, the hypothesis that VFTS
-682 is formed in relative isolation is rejected. This makes VFTS 682, with an inferred present day mass well above a hundred solar masses, the most massive runaway star known to date.
+time we calculate for this star. % Therefore, the hypothesis that VFTS
+% 682 is formed in relative isolation is rejected.
+If our results are confirmed by future astrometric data releases, VFTS
+682, with an inferred present day mass well above a hundred solar
+masses, is the most massive runaway star known to date.
 

### 2018-05-22 VFTS682.tex
minor in Sec. 2.1

@@
 (DEC) \mbox{-69$^\mathrm{o}$04'26.72''} \citep[][see also
-\Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2 in the
-ICGS frame]{evans:11}
+\Tabref{tab:vfts682} for updated coordinates from \emph{Gaia} DR2]{evans:11}
 was originally classified as a young stellar object \citep{gruendl:09}
@@
 presence of an optically thick, likely non-homogeneous, and variable
-wind obscuring the stellar surface.
+wind obscuring the stellar surface. The error quoted above on the
+radial velocity only include statistical uncertainties from the
+spectral analysis, and not the systematic errors arising from the
+modeling of emission line stars.
 

### 2018-05-22 VFTS682.tex
minor in Sec. 2.2

@@
 their data and missing from that catalog, due to its
-high extinction, E(B-V)=0.9, this star is faint enough
+high extinction, this star is faint enough
 to be included in that catalog. Its proper motions are
@@
 large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even
-exceed \emph{Gaia} data in quality.
+exceed \emph{Gaia} DR2 data in quality.
 
@@
 
-Since these measures are completely independent from ours, we can take
-an average weighted with $1/\delta \mu_i^2$ (cf.~notation in
-\Eqref{eq:mean}), obtaining $\delta\mu_\mathrm{RA}=\pm\,\mathrm{mas\
-  yr^{-1}}$ and $\delta\mu_\mathrm{DEC}=\pm\,\mathrm{mas\
-  yr^{-1}}$.
+% Since these measures are completely independent from ours, we can take
+% an average weighted with $1/\delta \mu_i^2$ (cf.~notation in
+% \Eqref{eq:mean}), obtaining $\delta\mu_\mathrm{RA}=\pm\,\mathrm{mas\
+%   yr^{-1}}$ and $\delta\mu_\mathrm{DEC}=\pm\,\mathrm{mas\
+%   yr^{-1}}$.
 
@@
 
-\todo{redo below with average HST+Gaia?Actually redo also calculation
+\todo{redo below with HST only and Gaia only?Redo also calculation
   of $v_\mathrm{pec}$}

### 2018-05-22 VFTS682.tex
minor in Sec. 2.2

@@
 of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not
-introduce significant errors. We also note that the determination of
+introduce significant errors. \cite{bressert:12} also noted that VFTS
+682 shows a significant radial velocity offset compared to the nebular
+lines from the gas surrounding it, suggesting peculiar motion in the
+line of sight direction. However, the determination of
 radial velocities for emission-lines stars is difficult because of the
 presence of an optically thick, likely non-homogeneous, and variable
-wind obscuring the stellar surface. The error quoted above on the
-radial velocity only include statistical uncertainties from the
+wind obscuring the stellar surface. The errors quoted above on the
+radial velocities include only statistical uncertainties from the
 spectral analysis, and not the systematic errors arising from the
@@
     radial velocity calculated using the  NV $\lambda4944$ line, which is
-    less sensitive to the wind velocity structure.}
+    less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).}
     \end{center}
@@
 
-The \emph{Gaia} Data Release 2, which became available on 25 April 2018,
+The second \emph{Gaia} Data Release (DR2), which became available on 25 April 2018,
 provides the five-parameter astrometric solution (positions,
@@
 data for VFTS 682 are reliable. However, the effective temperature
-reported in \emph{Gaia} DR2 is one order of magnitude lower than what found by
+reported in \emph{Gaia} DR2, based on the spectral fitting around the
+CaII triplet, is one order of magnitude lower than what found by
 \cite{bestenlehner:11}, and the best fit parallax of this star is
@@
 We retrieve for VFTS 682 the position in RA and DEC
-in the IGCS frame \cite[][]{brown:18}, its
+in the International Celestial Reference System \cite[][]{brown:18}, its
 proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
@@
 
-We define a local standard frame of reference to derive the peculiar velocity
-of VFTS 682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
-We select all the stars in a target of 0.05 degrees around R136
-(NGC2070) fulfilling the following criteria. First, we require a G-band
-magnitude brighter than 17, corresponding roughly to the
-completeness level of the VFTS survey \citep[here we implicitly assume
-G$\sim$V,][]{evans:11}. Then we require \texttt{visibility\_period} $\geq$ 5,
-\texttt{astrometric\_excess\_noise} $< 1$, and the errors on the proper
-motion components to be smaller than 0.07\,$\mathrm{mas\
-  yr^{-1}}$. At the distance to the LMC, $0.1\mathrm{mas\
-  yr^{-1}}\simeq25\,\kms$ \citep[e.g.,][]{platais:18}. We
-remove obvious foreground stars by requiring the absolute value of
-the reported parallaxes smaller than $0.1\,\mathrm{mas\ yr^{-1}}$.
+We follow \citet{lennon:18} to define a local frame of reference and derive the peculiar velocity
+of VFTS 682 with respect to the cluster R136. Their selection (see
+their Sec.~2.1) of 153
+bright ($G<17$) stars around R136 with reliable astrometric data from
+\emph{Gaia} DR2 yields $\mu_\mathrm{RA}=1.74\pm0.13\,\mathrm{mas\
+  yr^{-1}}$ and $\mu_\mathrm{DEC}=0.70\pm0.20\,\mathrm{mas\ yr^{-1}}$
+for the components of the mean motion of the region projected on the sky.
+
+% We define a local standard frame of reference to derive the peculiar velocity
+% of VFTS 682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
+% We select all the stars in a target of 0.05 degrees around R136
+% (NGC2070) fulfilling the following criteria. First, we require a G-band
+% magnitude brighter than 17, corresponding roughly to the
+% completeness level of the VFTS survey \citep[here we implicitly assume
+% G$\sim$V,][]{evans:11}. Then we require \texttt{visibility\_period} $\geq$ 5,
+% \texttt{astrometric\_excess\_noise} $< 1$, and the errors on the proper
+% motion components to be smaller than 0.07\,$\mathrm{mas\
+%   yr^{-1}}$. At the distance to the LMC, $0.1\mathrm{mas\
+%   yr^{-1}}\simeq25\,\kms$ \citep[e.g.,][]{platais:18}. We
+% remove obvious foreground stars by requiring the absolute value of
+% the reported parallaxes smaller than $0.1\,\mathrm{mas\ yr^{-1}}$.
 % At the distance to the
@@
 % not reduces further our sample.
-This selection yields a sample of 66 stars.% ,
+% This selection yields a sample of 66 stars.% ,
 
-We calculate the averaged proper motion components for the whole
-region using 
-\begin{equation}
-  \label{eq:mean}
-  \langle \mu_i\rangle = \frac{\sum_\mathrm{stars}\frac{1}{\Delta
-      \mu_i^2}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i^2}} \ \ , \
-  \ \Delta \langle \mu_i\rangle = \frac{\sqrt{N}}{\sum_\mathrm{stars}
-    \frac{1}{\Delta \mu_i^2}} \ \ ,
-\end{equation}
-where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
-on the proper motion component reported by \emph{Gaia}. The sums run over
-all the $N=66$ stars in our selected sample. We evaluate each proper motion
-component separately. For simplicity, throughout this study, we assume the same
+% We calculate the averaged proper motion components for the whole
+% region using 
+% \begin{equation}
+%   \label{eq:mean}
+%   \langle \mu_i\rangle = \frac{\sum_\mathrm{stars}\frac{1}{\Delta
+%       \mu_i^2}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i^2}} \ \ , \
+%   \ \Delta \langle \mu_i\rangle = \frac{\sqrt{N}}{\sum_\mathrm{stars}
+%     \frac{1}{\Delta \mu_i^2}} \ \ ,
+% \end{equation}
+% where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
+% on the proper motion component reported by \emph{Gaia}. The sums run over
+% all the $N=66$ stars in our selected sample. We evaluate each proper motion
+% component separately.
+For simplicity, throughout this study, we assume the same
 distance of $50$\,kpc to the star \citep[][]{pietrzynski:13}, and to
@@
 the distance determination when converting proper motions into
-physical velocities. The data retrieved, and the ipython notebook used for the analysis
-presented here will be made available at \todo{git repo bitbucket?}.
-
-\todo{nebular lines to compare velocity of the gas?}
-\todo{fix numbers}
+physical velocities. % The data retrieved, and the ipython notebook used for the analysis
+% presented here will be made available at \todo{git repo bitbucket?}.
 

### 2018-05-22 VFTS682.tex
minor in Sec. 3

@@
 They commented that this is unusual for Wolf-Rayet stars and more reminiscent
-of Luminous Blue Variable (LBV) stars.
+of Luminous Blue Variable (LBV) stars \citep[e.g.,][]{langer:94}.
 
@@
 proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
-respectively). For the radial velocity of VFTS 682 and of the 30 Doradus
+respectively). While \emph{Gaia} DR2 data can be correlated, this is
+not an issue for VFTS682, and for simplicity we treat them as
+uncorrelated in the following.
+
+For the radial velocity of VFTS 682 and of the 30 Doradus
 region as a whole, we instead use the VFTS data
@@
 distance of $50$\,kpc to the star \citep[][]{pietrzynski:13}, and to
-the 30 Doradus region as a whole. We do not consider the error bars on
-the distance determination when converting proper motions into
+the 30 Doradus region as a whole. We do not consider the error on
+the distance determination ($\lesssim2\%$) when converting proper motions into
 physical velocities. % The data retrieved, and the ipython notebook used for the analysis
@@
 
-Using the stars selected as described in \Secref{data:gaia} (a
-subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
-$\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
-$\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
-  yr^{-1}}$. These values are
-in good agreement with what is found by \cite{lennon:18}. However, we
-emphasize their sensitivity to the choice of stars adopted to define
-the reference frame. Subtracting these values from the
+% Using the stars selected as described in \Secref{data:gaia} (a
+% subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
+% $\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
+% $\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
+%   yr^{-1}}$. These values are
+% in good agreement with what is found by \cite{lennon:18}. However, we
+% emphasize their sensitivity to the choice of stars adopted to define
+% the reference frame.
+Subtracting the mean proper motion components given by
+\citep{lennon:18} from the
 proper motion of VFTS 682 (see \Tabref{tab:vfts682}), we obtain the
 components of proper motion of the star relative to the surrounding region
-$\delta\mu_\mathrm{RA}^{Gaia} = 0.12\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\delta\mu_\mathrm{DEC}^{Gaia} =
-0.08\pm 0.08\,\mathrm{mas\ yr^{-1}}$. The error budget is
+$\delta\mu_\mathrm{RA}^{Gaia} = \pm \,\mathrm{mas\ yr^{-1}}$ and $\delta\mu_\mathrm{DEC}^{Gaia} =
+\pm \,\mathrm{mas\ yr^{-1}}$. The error budget is
 dominated by the errors on the proper motion components of VFTS 682.
@@
 that it originates from the cluster,
-
+\todo{lennon: use pm for kin age and not kms}
 \begin{equation}

### 2018-05-23 VFTS682.tex
major, including danny's fig

@@
 
-% Recently, \citet{platais:18} used HST observations (based on GO-12499
-% \& 13359 observations with WFC3/UVIS and on archival WFPC2 and
-% GO-12499 with both cameras, see \citealt{platais:15}) to determine the
-% proper motion of massive stars in the 30 Doradus region. Our star of
-% interest is not saturated in any HST image \todo{?},
-% because of the large extinction toward this star.
-
 The catalog from \citet{platais:18} lists the proper motion
@@
 $\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
-  yr^{-1}}$.
+  yr^{-1}}$, corresponding to a relative proper motion
 
-% Since these measures are completely independent from ours, we can take
-% an average weighted with $1/\delta \mu_i^2$ (cf.~notation in
-% \Eqref{eq:mean}), obtaining $\delta\mu_\mathrm{RA}=\pm\,\mathrm{mas\
-%   yr^{-1}}$ and $\delta\mu_\mathrm{DEC}=\pm\,\mathrm{mas\
-%   yr^{-1}}$.
+\begin{equation}
+  \label{eq:pm_around}
+  \delta \mu^{HST} = \sqrt{\left(\delta\mu_\mathrm{RA}^{HST}\right)^2+\left(\delta\mu_\mathrm{DEC}^{HST}\right)^2}
+  = 0.20 \pm 0.15 \ \ .
+\end{equation}
 
-% %% pure HST physical velocities
-% \todo{done in the train: to be checked} Converting these into physical velocities we obtain 
-% $\delta v_\mathrm{RA}=2.3\pm30\,\kms$, $\delta v_\mathrm{DEC}=47\pm\,23\kms$, corresponding to
-% $v_\parallel^\mathrm{HST} = 47\pm\,1\kms$.
+These proper motion components can also be converted into projected
+velocities assuming a distance of 50\,kpc, obtaining $\delta
+v_\mathrm{RA}^{HST}=2\pm31\kms$ and $\delta
+v_\mathrm{RA}^{HST}=47\pm24\kms$, corresponding to a two-dimensional
+projected velocity of $47\pm24\kms$. Including the radial velocity
+measured by the VFTS survey we obtain a three-dimensional velocity of
 
-\todo{redo below with HST only and Gaia only?Redo also calculation
-  of $v_\mathrm{pec}$}
+\begin{equation}
+  \label{eq:speed_around_HST}
+  v_\mathrm{pec}^{HST} = \sqrt{\left(\delta v_\mathrm{RA}^{HST}\right)^2
+    +\left(\delta v_\mathrm{DEC}^{HST}\right)^2+\left(\delta
+      v_\mathrm{rad}^{VFTS}\right)^2} = 56 \pm 23
+  \kms \ .
+\end{equation}
 
 \subsection{New data from \emph{Gaia} DR2  \label{data:gaia}}
-%% using
-%% /scratch/mathieu/MMPS/SCRIPTS_dirty/my_script/binaries/binary_c/30Dor/vfts682/data/asu.tsv to find average
 
@@
     \cline{1-2}
-    $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.12$\pm$0.07 & \\[5pt]
-    $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.08 & \\[5pt]
+    $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \\[5pt]
+    $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
     \hline
@@
 components of proper motion of the star relative to the surrounding region
-$\delta\mu_\mathrm{RA}^{Gaia} = \pm \,\mathrm{mas\ yr^{-1}}$ and $\delta\mu_\mathrm{DEC}^{Gaia} =
-\pm \,\mathrm{mas\ yr^{-1}}$. The error budget is
-dominated by the errors on the proper motion components of VFTS 682.
+$\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\,\mathrm{mas\ yr^{-1}}$
+and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08
+\pm 0.09\,\mathrm{mas\ yr^{-1}}$. These components result in a
+two-dimensional relative proper motion of
+\begin{equation}
+  \label{eq:pm_gaia_around}
+  \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
+  = 0.13\pm 0.09 \ \ .
+\end{equation}
+
+Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
+of the stars used in \cite{lennon:18} to define the frame of reference
+we also adopt here, cf.~the right panel in their Fig.~2. The
+proper motions are decomposed in the radial and tangential direction
+from R136, to highlight the likelihood of it being the origin of fast
+moving stars. The plus sign marks VFTS 682, which sits at the edge of
+the proper motion distribution, but is not a clear outlier. Most
+relevant is the direction of the proper motion of VFTS 682, which we
+discuss in \Secref{sec:r136_origin}.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
+  \caption{Polar plot of the relative proper motion components for the stars
+    defining our reference frame. VFTS682 is indicated by the plus sign, and
+    other notable outliers (see \citet{lennon:18}) are also labeled. Concentric dashed-lines denote relative proper motions of
+    0.1, 0.2, and 0.4$\,\mathrm{mas\ yr^{-1}}$. Positive angle
+    indicate a tangential component pointing in the counterclockwise
+    direction, with 0 degrees corresponding to proper motion pointing
+    radially outward from R136.
+}
+  \label{fig:pm_polar}
+\end{figure}
 
 These can be converted into
-the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=28\pm16\,\kms$,
-$\delta v_\mathrm{DEC}^{Gaia}=20\pm19\,\kms$, assuming a distance of
-50\,kpc. % (we do not account for the uncertainty in the distance
+the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=24\pm19\,\kms$,
+$\delta v_\mathrm{DEC}^{Gaia}=20\pm23\,\kms$, assuming a distance of
+50\,kpc. These can be converted in a projected two-dimensional
+velocity of $31\pm21\kms$. % (we do not account for the uncertainty in the distance
 % estimate when propagating errors).
@@
 \begin{equation}
-  \label{eq:speed_around}
+  \label{eq:speed_around_Gaia}
   v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
     +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
-      v_\mathrm{rad}^{VFTS}\right)^2} = 46 \pm 18 
+      v_\mathrm{rad}^{VFTS}\right)^2} = 44 \pm 21
   \kms \ .
 \end{equation}
-This value for the three-dimensional speed of VFTS 682 with respect to
-the surrounding stars make it the most massive ``bona fide'' runaway star
-known to date.
+
+Therefore, two completely independent measures of the proper motion of
+VFTS 682 relative to the surrounding field, one from HST and one from
+\emph{Gaia} DR2, yield values of the peculiar three-dimensional
+velocity of VFTS 682 (\Eqref{eq:speed_around_HST} and
+\Eqref{eq:speed_around_Gaia}, respectively) which would make it the most massive runaway star
+known to date. However, the large errors on both the proper motion measures
+require confirmation with future astrometric data.
 

### 2018-05-23 VFTS682.tex

@@
   V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
- , H.~Sana$^{9}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
+  N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
 \affil{$^{1}${\todo{check affil. num}Astronomical Institute Anton Pannekoek, University of
@@
   $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
-     $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada}     
+  $^{}$ {National Research Council, Herzberg Astronomy \&
+    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
+    Canada}\\
+  $^{}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn, Auf dem H\"ugel 71, 53121, Bonn, Germany}
 }
 
-%\thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+% \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
 \date{}
@@
 
-%\todo{recheck all numbers}
+% \todo{recheck all numbers}
 \section{Introduction}
@@
 \citep[e.g.,][]{zinnecker:07}. Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
-%\citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
+% \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
 evolve fast, typically reside in dense groups and remain enshrouded in
@@
 surveys, e.g., Large Synoptic Spectroscopic Survey and LIGO/Virgo,
-%\citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others. 
+% \citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others. 
 which  will reveal transients associated to massive stars
@@
 Most remarkable is the location of VFTS 682, at a projected distance of
-$\sim$$29$\,pc from the core of the star cluster
+119.4 arcseconds, corresponding to $\sim$$29$\,pc from the core of the star cluster
 R136. \citet{bestenlehner:11} discuss two possibilities for the
@@
 bright ($G<17$) stars around R136 with reliable astrometric data from
-\emph{Gaia} DR2 yields $\mu_\mathrm{RA}=1.74\pm0.13\,\mathrm{mas\
-  yr^{-1}}$ and $\mu_\mathrm{DEC}=0.70\pm0.20\,\mathrm{mas\ yr^{-1}}$
+\emph{Gaia} DR2 yields $\mu_\mathrm{RA}=1.74\pm0.01\,\mathrm{mas\
+  yr^{-1}}$ and $\mu_\mathrm{DEC}=0.70\pm0.02\,\mathrm{mas\ yr^{-1}}$
 for the components of the mean motion of the region projected on the sky.
@@
 
-These can be converted into
+The proper motion components can be converted into
 the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=24\pm19\,\kms$,
 $\delta v_\mathrm{DEC}^{Gaia}=20\pm23\,\kms$, assuming a distance of
-50\,kpc. These can be converted in a projected two-dimensional
+50\,kpc. These can be combined obtaining a projected two-dimensional
 velocity of $31\pm21\kms$. % (we do not account for the uncertainty in the distance

### 2018-05-23 VFTS682.tex

@@
 \begin{equation}
-  \label{eq:pm_around}
+  \label{eq:pm_around_HST}
   \delta \mu^{HST} = \sqrt{\left(\delta\mu_\mathrm{RA}^{HST}\right)^2+\left(\delta\mu_\mathrm{DEC}^{HST}\right)^2}
-  = 0.20 \pm 0.15 \ \ .
+  = 0.20 \pm 0.15\,\mathrm{mas\
+  yr^{-1}} \ \ .
 \end{equation}
@@
   \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
-  = 0.13\pm 0.09 \ \ .
+  = 0.13\pm 0.09\,\mathrm{mas\
+  yr^{-1}} \ \ .
 \end{equation}
@@
 
-The red arrow in \Figref{fig:main} shows the proper motion of VFTS 682
-relative to the region, and the lighter red arrows show the possible
-range of directions within the uncertainties in the measured proper
-motion. It is clear that the most likely origin of the star is R136,
+Assuming that the best estimate of the peculiar velocity of VFTS 682
+are reliable, we now address the question of its likely origin.
+
+The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
+VFTS 682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the orange arrows show the possible
+range of directions within the uncertainties in the measured relative proper
+motion. These arrows cross at the present-day location of VFTS 682 and
+are prolonged in the direction opposite to the motion to illustrate
+the possible range of origins. The green arrow shows the direction of
+the relative proper motion from HST (see \Eqref{eq:pm_around_HST}),
+but we omit to show the possible range of directions from HST data, since it is
+insufficiently constrained to be useful.
+
+Although the large uncertainties on the relative proper motion
+components results in a wide range of possible directions, it is clear that the most likely origin of the star is R136,
 as suggested based on numerical N-body simulations by \cite{fujii:11}
@@
 that it originates from the cluster,
-\todo{lennon: use pm for kin age and not kms}
+
 \begin{equation}
   \label{eq:kin_age}
-  \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} \simeq
-  \frac{29\,\mathrm{pc}}{34\,\kms} \simeq 0.85\pm\,0.02\, \mathrm{Myr} \ \ ,
+  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu_\mathrm{RA}^\mathrm{Gaia}} \simeq
+  \frac{119.4\,\mathrm{arcsec}}{0.13\,\mathrm{mas\ yr^{-1}}} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
 \end{equation}
-where $d_\parallel =29$\,pc is the projected distance from VFTS 682 to
-the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel \equiv \sqrt{\left(\delta v_\mathrm{RA}\right)^2
-  +\left(\delta v_\mathrm{DEC}\right)^2} =34\pm
-1\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
-using the relative proper motion components calculated in
-\Secref{sec:runaway}.% , and we use the approximation $1 \kms \simeq 1
-% \mathrm{pc \ Myr^{-1}}$.
+where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS 682 to
+the core of R136 \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
 As in the rest of this study, we neglect for
-simplicity the error on the distance estimates, which amounts to about
-$\sim$$2.2\%$, and is negligible compared to other uncertainties.
+simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
 
@@
 corroborates the idea that the star is the result of a dynamical
-ejection.
+ejection very soon in the cluster evolution. \todo{address Norbert comments on age determination}.
 

### 2018-05-23 VFTS682.tex
got to the end again

@@
 spectral analysis assuming a smooth outflow, i.e., not accounting for clumping
-\citep[][]{bestenlehner:11}. These properties make it one of the
+\citep[][]{bestenlehner:11}. The present-day surface helium mass
+fraction is $Y=0.45\, (0.49)$ from the spectral analysis of
+\cite{bestenlehner:11} (\citealt{rubio-diez:17}). This value can be
+used to constrain the age of the star and its past evolution. This set
+of properties make VFTS 682 one of the
 most extreme objects in the region. It is reminiscent of the very
@@
 corroborates the idea that the star is the result of a dynamical
-ejection very soon in the cluster evolution. \todo{address Norbert comments on age determination}.
+ejection very soon in the cluster evolution. We emphasize that the
+present-day surface helium abundance
+\citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
+limit on the age corresponding to the time needed for the star to
+synthesize that amount of helium, and such lower limit corresponds to
+$\sim$0.9\,Myr based on the models from \cite{kohler:15}.
 
@@
   \includegraphics[width=0.48\textwidth]{./figures/main_plot_good}  
-  \caption{The red solid arrow indicates the proper motion of VFTS 682
-    relative to the region, starting from the present day position of
+  \caption{The red and green solid arrow indicate the proper motion of VFTS 682
+    relative to the region from \emph{Gaia} DR2 and HST data respectively, starting from the present day position of
     the star. The orange arrows indicate the possible
-    directions of projected motion within the errors, and are extended
+    directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
     backwards (dashed) to illustrate the uncertainty on the origin of the
-    star. The blue stars
-    are a subset of those we use to define the local rest
-    frame. \todo{load color picture on background}\todo{show
-      separately arrows fro Gaia/HST and combination? Or show only
-      combination? Right now it's Gaia only}}
+    star. \todo{load color picture on background?}}
 
@@
 \label{sec:discussion}
-\todo{Selma: Should add caveat that it may be a spurious detection. That reconfirmation in following epochs is warranted. }
 
-Based on our results, we claim that VFTS 682 is the most massive
-runaway known to date, with a peculiar three-dimensional speed of $46\pm18\,\kms$. This means that isolated star formation is
+Based on our results, we tentatively claim that VFTS 682 is the most massive
+runaway known to date, with a peculiar three-dimensional speed of
+$44\pm21\,\kms$. This result will need to be revisited as updated
+astrometric parameters from future \emph{Gaia} data releases become
+available. If this result will be confirmed, it means that isolated star formation is
 \emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
-$0.85\pm0.02$\,Myr ago. Because of the exceptionally large mass
+$0.9\pm0.6$\,Myr ago. Because of the exceptionally large mass
 of this star, this raises the question of which stars must populate
@@
 VFTS 682 and the massive members of R136. While it is not possible to
-robustly exclude that VFTS 682 is itself a merger, as already
-mentioned, its spectrum does not show any of the obvious signatures
-like fast rotation. To eject such a massive object, the cluster is
+robustly exclude that VFTS 682 is itself a merger, its spectrum does not show any of the obvious signatures
+like fast rotation, which however could have already slowed down
+because of the wind angular momentum losses during and after the
+merger.
+
+To eject such a massive object, the cluster is
 expected to have produced
@@
 isolated massive stars are observed in the region, some with known
-large radial velocities. A comprehensive study of the kinematic
+large radial velocities and/or proper motion (see
+\Figref{fig:pm_polar} and Sana et al., in prep.). A comprehensive study of the kinematic
 properties of all the massive stars surrounding R136 might shed light
@@
 VFTS 682 might have been ejected in the opposite direction, and is also
-in isolation at present day.
+isolated at present day. If the ejection was caused by an interaction
+with a binary, however, it is likely that the binary scattered in the
+opposite direction will experience further dynamical interactions on
+its way, modifying its trajectory and making it difficult to find.  %absence of proof would not undermine our conclusions
 
@@
 
-\citet{lennon:18} carried out a study similar to ours on VFTS16, another
-massive star ($\sim$$90\,M_\odot$) in the 30 Doradus region, previously
-known to be a runaway from the value of
-its radial velocity. They also concluded that VFTS16 is 
-the result of a dynamical ejection from the R136 cluster. 
-The value of $\tau_\mathrm{kin}\simeq0.85$\,Myr we find for VFTS 682 (see \Secref{sec:r136_origin}) is smaller
-than the corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic age of
-$\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
+\citet{lennon:18} carried out a study similar to ours on the fast
+moving O-type stars
+in the region, and found two massive runaway stars
+($\sim$$90\,M_\odot$) in the 30 Doradus region. One of them (VFTS 16)
+was previously known as a runaway star from its line of sight velocity
+\citep[][]{evans:10}. \citet{lennon:18} also concluded that VFTS16 is 
+the result of a dynamical ejection from the R136 cluster, while the
+origin of the other star (VFTS 72) is less clear given its direction
+of motion. The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
+VFTS 682 (see \Secref{sec:r136_origin}) is smaller than the
+corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic
+age of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
 massive VFTS 682 was ejected later than VFTS16 from the same cluster.
@@
 time of maximum stellar density. This might have implications for the
-question on whether the cluster formed via a monolithic collapse, or
-as a (potentially ongoing) merger of several sub-clusters.
+question of whether the cluster formed via a monolithic collapse, or
+as a (potentially ongoing) merger of several sub-structures \citep[e.g.,]{sabbi:12}.
 

### 2018-05-24 VFTS682.tex
typo
added todo
fixed affiliations

@@
   N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
-\affil{$^{1}${\todo{check affil. num}Astronomical Institute Anton Pannekoek, University of
+\affil{$^{1}${Astronomical Institute Anton Pannekoek, University of
     Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
@@
   $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-  $^{}${Department of Physics and Astronomy, Hicks Building,
+  $^{5}${Department of Physics and Astronomy, Hicks Building,
     Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
-  $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada} 
-  $^{}$ {School of Astronomy \& Space Science, University of the Chinese
-    Academy of Sciences, Beijing 100012, China}\\
-  $^{}$ {National Astronomical Observatories, Chinese Academy of
-    Sciences, Beijing 100012, China}\\
-  $^{}$ {Department of Physics, University of Oxford, Keble Road,
-    Oxford OX1 3RH, UK} \\
-  $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}   
-  $^{}$ {Space Telescope Science Institute, 3700 San Martin Drive,
-    Baltimore, MD 21218, USA}\\
-  {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-  $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
-  $^{}$ {National Research Council, Herzberg Astronomy \&
+  $^{6}$ {National Research Council, Herzberg Astronomy \&
     Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
     Canada}\\
-  $^{}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn, Auf dem H\"ugel 71, 53121, Bonn, Germany}
+  $^{7}$ {School of Astronomy \& Space Science, University of the Chinese
+    Academy of Sciences, Beijing 100012, China}\\
+  $^{8}$ {National Astronomical Observatories, Chinese Academy of
+    Sciences, Beijing 100012, China}\\
+  $^{9}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
+    Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
+  $^{10}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
+  $^{11}$ {Department of Physics, University of Oxford, Keble Road,
+    Oxford OX1 3RH, UK} \\
+  $^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
 }
@@
 
-% \todo{recheck all numbers}
 \section{Introduction}
@@
   \label{eq:kin_age}
-  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu_\mathrm{RA}^\mathrm{Gaia}} \simeq
+  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu^\mathrm{Gaia}} \simeq
   \frac{119.4\,\mathrm{arcsec}}{0.13\,\mathrm{mas\ yr^{-1}}} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
@@
 $\sim$0.9\,Myr based on the models from \cite{kohler:15}.
+\todo{discuss position angle}
 

### 2018-05-24 VFTS682.tex
minor everywhere

@@
 
-% \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+ \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
 \date{}
@@
   runaway star dynamically ejected from R136. Its
-  projected two-dimensional speed relative to the cluster is $\sim$$34\pm17\kms$, implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
+  projected two-dimensional speed relative to the cluster is $\sim$$30\pm20\kms$, implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
   accounting for the (uncertain) velocity component along the line of sight. The

### 2018-05-24 VFTS682.tex
./aa.cls

@@
 The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
-VFTS 682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the orange arrows show the possible
+VFTS 682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the cyan arrows show the possible
 range of directions within the uncertainties in the measured relative proper
@@
 are prolonged in the direction opposite to the motion to illustrate
-the possible range of origins. The green arrow shows the direction of
-the relative proper motion from HST (see \Eqref{eq:pm_around_HST}),
-but we omit to show the possible range of directions from HST data, since it is
-insufficiently constrained to be useful.
+the possible range of origins. % The green arrow shows the direction of
+% the relative proper motion from HST (see \Eqref{eq:pm_around_HST}),
+% but we omit to show the possible range of directions from HST data, since it is
+% insufficiently constrained to be useful.
 
@@
   \includegraphics[width=0.48\textwidth]{./figures/main_plot_good}  
-  \caption{The red and green solid arrow indicate the proper motion of VFTS 682
-    relative to the region from \emph{Gaia} DR2 and HST data respectively, starting from the present day position of
-    the star. The orange arrows indicate the possible
+  \caption{The red solid arrow indicates the proper motion of VFTS 682
+    relative to the region from \emph{Gaia} DR2, starting from the present day position of
+    the star. The cyan arrows indicate the possible
     directions of projected motion within the \emph{Gaia} DR2 errors, and are extended

### 2018-05-25 VFTS682.tex
need new fig
new fig and caption: paper ready to be circulated with co-authors

@@
 $\sim$0.9\,Myr based on the models from \cite{kohler:15}.
-\todo{discuss position angle}
+% \todo{discuss position angle}
 
@@
   \centering
-  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good}  
+  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
   \caption{The red solid arrow indicates the proper motion of VFTS 682
     relative to the region from \emph{Gaia} DR2, starting from the present day position of
-    the star. The cyan arrows indicate the possible
+    the star. The orange arrows indicate the possible
     directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
     backwards (dashed) to illustrate the uncertainty on the origin of the
-    star. \todo{load color picture on background?}}
+    star. The length of the prolongations is proportional to the relative proper motion
+    times the kinetic age of VFTS 682. \todo{load color picture on background? Fig with HST data
+      available, but confusing...}}
 

### 2018-05-27 VFTS682.tex
some trimming to satisfy letter format, still 1000 words too long

@@
   % 
-  If confirmed by future astrometric data releases, this would make VFTS 682 the most massive runaway star known to date.
+  If confirmed by future astrometric data, this would make VFTS 682 the most massive runaway star known to date.
 It proves that star clusters are capable of ejecting their most massive members.
@@
 \maketitle{}
+
+\todo{trim ~1085 words}
 
@@
 
-\citet{bestenlehner:11} further analyzed the spectra using the non-LTE
-model atmosphere code CMFGEN \citep{hillier:98} to derive the stellar
+\citet{bestenlehner:11} further analyzed the spectra % using the non-LTE
+% model atmosphere code CMFGEN \citep{hillier:98}
+to derive the stellar
 parameters and surface abundances. They inferred an extinction
@@
 Further worth noticing is the variability of the
-star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
-lower than what found by \citet{evans:11}. \citet{bestenlehner:11} discuss the Optical Gravitational Lensing
+star. % \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
+% lower than what found by \citet{evans:11}.
+\citet{bestenlehner:11} discuss the Optical Gravitational Lensing
 Experiment (OGLE-III) light curves \citep{udalski:08} and show a
@@
 They commented that this is unusual for Wolf-Rayet stars and more reminiscent
-of Luminous Blue Variable (LBV) stars \citep[e.g.,][]{langer:94}.
+of Luminous Blue Variable (LBV) stars. % \citep[e.g.,][]{langer:94}
+.
 
@@
 experienced rotationally induced chemically homogeneous evolution
-\citep[][]{maeder:00,demink:09,marchant:16}, and (ii) it has not
+\citep[][]{maeder:00,demink:09}, and (ii) it has not
 accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
@@
 a binary black-hole, provided that stars this massive can avoid a
-pair-instability supernova \cite[][]{fowler:64, rakavy:67} at LMC
+pair-instability supernova \cite[e.g.,][]{rakavy:67} at LMC
 metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS 682 could be either a
@@
 Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \emph{Gaia} Multilateral Agreement.
-  The background image of \Figref{fig:main} is based on observations
-  made with ESO Telescopes at the La Silla Observatory under programme
-  ID 076.C-0888, processed and released by the ESO VOS/ADP group.
-  This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
+  % The background image of \Figref{fig:main} is based on observations
+  % made with ESO Telescopes at the La Silla Observatory under programme
+  % ID 076.C-0888, processed and released by the ESO VOS/ADP group.
+  % This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
 \end{acknowledgements}

### 2018-05-27 VFTS682.tex
50 refs

@@
 Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \emph{Gaia} Multilateral Agreement.
-  % The background image of \Figref{fig:main} is based on observations
-  % made with ESO Telescopes at the La Silla Observatory under programme
-  % ID 076.C-0888, processed and released by the ESO VOS/ADP group.
-  % This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
+  The background image of \Figref{fig:main} is based on observations
+  made with ESO Telescopes at the La Silla Observatory under programme
+  ID 076.C-0888, processed and released by the ESO VOS/ADP group.
+  This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
 \end{acknowledgements}

### 2018-05-27 VFTS682.tex
some words cutting, more to do, flow to be checked

@@
 possible dependence on environment and metallicity, is crucial for
-understanding the role massive stars play within their host galaxies
-as sources of feedback, but also for understanding the transients that
-mark their death and the compact remnants they leave behind. This is
-therefore a key question for the present and upcoming transient
-surveys, e.g., Large Synoptic Spectroscopic Survey and LIGO/Virgo,
-% \citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others. 
-which  will reveal transients associated to massive stars
-evolution and death.
+understanding the role massive stars play within their host galaxies,
+% as sources of feedback,
+but also for understanding the transients that
+mark their death and the compact remnants they leave behind. % This is
+% therefore a key question for the present and upcoming transient
+% surveys, e.g., Large Synoptic Spectroscopic Survey and LIGO/Virgo,
+% % \citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others. 
+% which  will reveal transients associated to massive stars
+% evolution and death.
 
-In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to  form preferentially in the dense cores.
-In this picture, field stars are primarily the result of the dissolution of dense groups.
+In has been proposed that most, if not all, stars form in clusters
+\citep{lada:03}, where massive stars are thought to  form
+preferentially in the dense cores. In this picture, field stars are primarily the result of the dissolution of dense groups. 
 However, a significant population of massive stars exists in relative
@@
 from dynamical interactions \citep[e.g.,][]{poveda:67} or from the
-disruption of binary systems by the core-collapse of the companion
+disruption of binary systems at the death of the companion %by the core-collapse of the companion
 star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.
@@
 \citep{schneider:18}. These correspond to an inferred initial mass of
-$150.0^{+28.7}_{-17.4}\,M_\odot$. It features a wind mass loss rate of
-$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, inferred from
-spectral analysis assuming a smooth outflow, i.e., not accounting for clumping
-\citep[][]{bestenlehner:11}. The present-day surface helium mass
-fraction is $Y=0.45\, (0.49)$ from the spectral analysis of
-\cite{bestenlehner:11} (\citealt{rubio-diez:17}). This value can be
-used to constrain the age of the star and its past evolution. This set
-of properties make VFTS 682 one of the
-most extreme objects in the region. It is reminiscent of the very
+$150.0^{+28.7}_{-17.4}\,M_\odot$. This makes VFTS 682 one of the
+most extreme objects in the region.
+
+From the spectral point of view, it is reminiscent of the very
 massive stars % , a.k.a. ``monster stars'',
@@
 \citet{dekoter:97,crowther:10, crowther:16} in the core of the
-R136 cluster.
+R136 cluster. In particular, a remarkable similarity exist between the
+spectrum of VFTS 682 and R136a3, located in the core of the cluster
+\citep{crowther:10} for which \citet{crowther:16} report a
+current mass estimate of $180^{+30}_{-30}\Msun$. The R136 cluster hosts
+at least two more very massive WN5h stars, R136a1 and R136a2, whose
+estimated current masses are even higher. However, VFTS 682 stands
+out by its isolation at a projected offset of 119.4 arcseconds, corresponding to $\sim$$29$\,pc from the core of the star cluster
+R136.
 
-Most remarkable is the location of VFTS 682, at a projected distance of
-119.4 arcseconds, corresponding to $\sim$$29$\,pc from the core of the star cluster
-R136. \citet{bestenlehner:11} discuss two possibilities for the
-origin of VFTS 682. Quoting directly from their abstract: "(i) the
+\citet{bestenlehner:11} discuss two possibilities for the
+origin of VFTS 682. % Quoting directly 
+From their abstract: "(i) the
 star either formed in situ, which would have profound implications for
@@
 
-% don't know where to go from here without giving away the main
-% result...Stephen wants to add a citation to fujii and banerjee in
-% the intro, which I also think would be fair...
-% The case of VFTS 682 has also spawned significant theoretical efforts
-% to explain its isolation via dynamical ejection from R136, e.g.,
-% \citet{fujii:11} and \citet{banerjee:12}. Both these studies suggested
-
 Recently, \citet{platais:15,platais:18} used multi-epoch \emph{Hubble Space
-  Telescope} (HST) imaging to identify and shed new light on runaway stars from
-the 30 Doradus region based on proper motion
+  Telescope} (HST) imaging to identify % and shed new light on
+runaway stars in
+ 30 Doradus based on proper motion
 measurements. \citet{lennon:18} used data from the second \emph{Gaia}
@@
 report on our analysis of the \emph{Gaia} DR2 data for VFTS 682 which
-was not part of the sample of \citet{lennon:18} because of its WNh5
+was not part of the sample of \citet{lennon:18} because of its 
 spectral type. We combine the radial velocity measurements from the
@@
 
-Our results indicate that VFTS 682 is a runaway star, although with
-low statistical significance. The
-direction and magnitude of the velocity vector are consistent with
-dynamical ejection from R136. The age inferred from the spectral
-analysis \citep[from][]{schneider:18} is consistent with the travel
-time we calculate for this star. % Therefore, the hypothesis that VFTS
-% 682 is formed in relative isolation is rejected.
-If our results are confirmed by future astrometric data releases, VFTS
-682, with an inferred present day mass well above a hundred solar
-masses, is the most massive runaway star known to date.
+%%% cut for length reasons: all is said below
+% Our results indicate that VFTS 682 is a runaway star, although with
+% low statistical significance. The
+% direction and magnitude of the velocity vector are consistent with
+% dynamical ejection from R136. The age inferred from the spectral
+% analysis \citep[from][]{schneider:18} is consistent with the travel
+% time we calculate for this star. % Therefore, the hypothesis that VFTS
+% % 682 is formed in relative isolation is rejected.
+% If our results are confirmed by future astrometric data releases, VFTS
+% 682, with an inferred present day mass well above a hundred solar
+% masses, is the most massive runaway star known to date.
 
@@
 (DEC) \mbox{-69$^\mathrm{o}$04'26.72''} \citep[][see also
-\Tabref{tab:vfts682} for updated coordinates from \emph{Gaia} DR2]{evans:11}
+\Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2]{evans:11}
 was originally classified as a young stellar object \citep{gruendl:09}
@@
 
-The star is very similar from the spectral point of view to R136a3, located inside the star cluster
-R136 \citep{crowther:10}  for which \citet{crowther:16} report a
-current mass estimate of $180^{+30}_{-30}\Msun$. The R136 cluster hosts
-at least two more very massive WN5h stars, R136a1 and R136a2, whose
-estimated current masses are even higher. However, VFTS 682 stands
-out by its isolation at a projected offset of 29\,pc from the center
-of the cluster \citep{bestenlehner:11}.
+\citet{bestenlehner:11} reported for VFTS 682 a wind mass loss rate of
+$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, inferred from
+spectral analysis assuming a smooth outflow, i.e., not accounting for
+clumping. The present-day surface helium mass
+fraction is $Y=0.45\, (0.49)$ from the spectral analysis of
+\cite{bestenlehner:11} (\citealt{rubio-diez:17}). This value can be
+used to constrain the age of the star and its past evolution.
+
+% The star is very similar from the spectral point of view to R136a3, located inside the star cluster
+% R136 \citep{crowther:10}  for which \citet{crowther:16} report a
+% current mass estimate of $180^{+30}_{-30}\Msun$. The R136 cluster hosts
+% at least two more very massive WN5h stars, R136a1 and R136a2, whose
+% estimated current masses are even higher. However, VFTS 682 stands
+% out by its isolation at a projected offset of 29\,pc from the center
+% of the cluster \citep{bestenlehner:11}.
 
@@
 They commented that this is unusual for Wolf-Rayet stars and more reminiscent
-of Luminous Blue Variable (LBV) stars. % \citep[e.g.,][]{langer:94}
-.
+of Luminous Blue Variable (LBV) stars. % \citep[e.g.,][]{langer:94}.
 

### 2018-05-28 VFTS682.tex
Chris Evans comments incorporated

@@
   % 
-  If confirmed by future astrometric data, this would make VFTS 682 the most massive runaway star known to date.
-It proves that star clusters are capable of ejecting their most massive members.
+  If confirmed by future astrometric data, this would make VFTS 682
+  the most massive runaway star known to date.
+It would prove that star clusters are capable of ejecting their most massive members.
 We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
@@
 682, and test the hypothesis that this star was ejected from R136. We
-use the HST proper motions from \citet{platais:18} to validate our
+use the HST proper motions from \citet{platais:18} to test our
 analysis of the \emph{Gaia} DR2 data.
@@
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} \citep[][see also
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000 \citep[][see also
 \Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2]{evans:11}
@@
 Although the large uncertainties on the relative proper motion
-components results in a wide range of possible directions, it is clear that the most likely origin of the star is R136,
+components results in a wide range of possible directions,we argue that the most likely origin of the star is R136,
 as suggested based on numerical N-body simulations by \cite{fujii:11}
@@
 
-VFTS 682 is the most massive runaway known to date, and its ejection
+VFTS 682 is potentially the most massive runaway known to date, and its ejection
 from the cluster R136 likely implies that it is only the ``tip of the
@@
 \bibliographystyle{aa}
-\bibliography{bibliography/vfts682}
+%\bibliography{bibliography/vfts682}
+\begin{thebibliography}{50}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Banerjee} {et~al.}(2012){Banerjee}, {Kroupa}, \& {Oh}}]{banerjee:12}
+{Banerjee}, S., {Kroupa}, P., \& {Oh}, S. 2012, \apj, 746, 15
+
+\bibitem[{{Bate}(2009)}]{bate:09}
+{Bate}, M.~R. 2009, \mnras, 392, 590
+
+\bibitem[{{Bestenlehner} {et~al.}(2011){Bestenlehner}, {Vink}, {Gr{\"a}fener},
+  {Najarro}, {Evans}, {Bastian}, {Bonanos}, {Bressert}, {Crowther}, {Doran},
+  {Friedrich}, {H{\'e}nault-Brunet}, {Herrero}, {de Koter}, {Langer}, {Lennon},
+  {Ma{\'{\i}}z Apell{\'a}niz}, {Sana}, {Soszynski}, \&
+  {Taylor}}]{bestenlehner:11}
+{Bestenlehner}, J.~M., {Vink}, J.~S., {Gr{\"a}fener}, G., {et~al.} 2011, \aap,
+  530, L14
+
+\bibitem[{{Blaauw}(1961)}]{blaauw:61}
+{Blaauw}, A. 1961, \bain, 15, 265
+
+\bibitem[{{Bressert} {et~al.}(2012){Bressert}, {Bastian}, {Evans}, {Sana},
+  {H{\'e}nault-Brunet}, {Goodwin}, {Parker}, {Gieles}, {Bestenlehner}, {Vink},
+  {Taylor}, {Crowther}, {Longmore}, {Gr{\"a}fener}, {Ma{\'{\i}}z
+  Apell{\'a}niz}, {de Koter}, {Cantiello}, \& {Kruijssen}}]{bressert:12}
+{Bressert}, E., {Bastian}, N., {Evans}, C.~J., {et~al.} 2012, \aap, 542, A49
+
+\bibitem[{{Brott} {et~al.}(2011){Brott}, {de Mink}, {Cantiello}, {Langer}, {de
+  Koter}, {Evans}, {Hunter}, {Trundle}, \& {Vink}}]{brott:11}
+{Brott}, I., {de Mink}, S.~E., {Cantiello}, M., {et~al.} 2011, \aap, 530, A115
+
+\bibitem[{{Crowther} {et~al.}(2016){Crowther}, {Caballero-Nieves}, {Bostroem},
+  {Ma{\'{\i}}z Apell{\'a}niz}, {Schneider}, {Walborn}, {Angus}, {Brott},
+  {Bonanos}, {de Koter}, {de Mink}, {Evans}, {Gr{\"a}fener}, {Herrero},
+  {Howarth}, {Langer}, {Lennon}, {Puls}, {Sana}, \& {Vink}}]{crowther:16}
+{Crowther}, P.~A., {Caballero-Nieves}, S.~M., {Bostroem}, K.~A., {et~al.} 2016,
+  \mnras, 458, 624
+
+\bibitem[{{Crowther} {et~al.}(2010){Crowther}, {Schnurr}, {Hirschi}, {Yusof},
+  {Parker}, {Goodwin}, \& {Kassim}}]{crowther:10}
+{Crowther}, P.~A., {Schnurr}, O., {Hirschi}, R., {et~al.} 2010, \mnras, 408,
+  731
+
+\bibitem[{{Davidson} {et~al.}(2016){Davidson}, {Humphreys}, \&
+  {Weis}}]{davidson:16}
+{Davidson}, K., {Humphreys}, R.~M., \& {Weis}, K. 2016, arXiv:1608.02007
+
+\bibitem[{{de Koter} {et~al.}(1997){de Koter}, {Heap}, \&
+  {Hubeny}}]{dekoter:97}
+{de Koter}, A., {Heap}, S.~R., \& {Hubeny}, I. 1997, \apj, 477, 792
+
+\bibitem[{{de Mink} {et~al.}(2009){de Mink}, {Cantiello}, {Langer}, {Pols},
+  {Brott}, \& {Yoon}}]{demink:09}
+{de Mink}, S.~E., {Cantiello}, M., {Langer}, N., {et~al.} 2009, \aap, 497, 243
+
+\bibitem[{{Evans} {et~al.}(2011){Evans}, {Taylor}, {H{\'e}nault-Brunet},
+  {Sana}, {de Koter}, {Sim{\'o}n-D{\'{\i}}az}, {Carraro}, {Bagnoli}, {Bastian},
+  {Bestenlehner}, {Bonanos}, {Bressert}, {Brott}, {Campbell}, {Cantiello},
+  {Clark}, {Costa}, {Crowther}, {de Mink}, {Doran}, {Dufton}, {Dunstall},
+  {Friedrich}, {Garcia}, {Gieles}, {Gr{\"a}fener}, {Herrero}, {Howarth},
+  {Izzard}, {Langer}, {Lennon}, {Ma{\'{\i}}z Apell{\'a}niz}, {Markova},
+  {Najarro}, {Puls}, {Ramirez}, {Sab{\'{\i}}n-Sanjuli{\'a}n}, {Smartt},
+  {Stroud}, {van Loon}, {Vink}, \& {Walborn}}]{evans:11}
+{Evans}, C.~J., {Taylor}, W.~D., {H{\'e}nault-Brunet}, V., {et~al.} 2011, \aap,
+  530, A108
+
+\bibitem[{{Evans} {et~al.}(2010){Evans}, {Walborn}, {Crowther},
+  {H{\'e}nault-Brunet}, {Massa}, {Taylor}, {Howarth}, {Sana}, {Lennon}, \& {van
+  Loon}}]{evans:10}
+{Evans}, C.~J., {Walborn}, N.~R., {Crowther}, P.~A., {et~al.} 2010, \apjl, 715,
+  L74
+
+\bibitem[{{Fujii} \& {Portegies Zwart}(2011)}]{fujii:11}
+{Fujii}, M.~S. \& {Portegies Zwart}, S. 2011, Science, 334, 1380
+
+\bibitem[{{Gaia Collaboration} {et~al.}(2018){Gaia Collaboration}, {Brown},
+  {Vallenari}, {Prusti}, {de Bruijne}, {Babusiaux}, \&
+  {Bailer-Jones}}]{brown:18}
+{Gaia Collaboration}, {Brown}, A.~G.~A., {Vallenari}, A., {et~al.} 2018,
+  ArXiv:1804.09365
+
+\bibitem[{{Gaia Collaboration} {et~al.}(2016){Gaia Collaboration}, {Prusti},
+  {de Bruijne}, {Brown}, {Vallenari}, {Babusiaux}, {Bailer-Jones}, {Bastian},
+  {Biermann}, {Evans}, \& et~al.}]{gaia:16}
+{Gaia Collaboration}, {Prusti}, T., {de Bruijne}, J.~H.~J., {et~al.} 2016,
+  \aap, 595, A1
+
+\bibitem[{{Gruendl} \& {Chu}(2009)}]{gruendl:09}
+{Gruendl}, R.~A. \& {Chu}, Y.-H. 2009, \apjs, 184, 172
+
+\bibitem[{{Gvaramadze} {et~al.}(2012){Gvaramadze}, {Weidner}, {Kroupa}, \&
+  {Pflamm-Altenburg}}]{gvaramadze:12}
+{Gvaramadze}, V.~V., {Weidner}, C., {Kroupa}, P., \& {Pflamm-Altenburg}, J.
+  2012, \mnras, 424, 3037
+
+\bibitem[{{Hogg}(2018)}]{hogg:18}
+{Hogg}, D.~W. 2018, ArXiv:1804.07766
+
+\bibitem[{{Humphreys} {et~al.}(2016){Humphreys}, {Weis}, {Davidson}, \&
+  {Gordon}}]{humphreys:16}
+{Humphreys}, R.~M., {Weis}, K., {Davidson}, K., \& {Gordon}, M.~S. 2016, \apj,
+  825, 64
+
+\bibitem[{{K{\"o}hler} {et~al.}(2015){K{\"o}hler}, {Langer}, {de Koter}, {de
+  Mink}, {Crowther}, {Evans}, {Gr{\"a}fener}, {Sana}, {Sanyal}, {Schneider}, \&
+  {Vink}}]{kohler:15}
+{K{\"o}hler}, K., {Langer}, N., {de Koter}, A., {et~al.} 2015, \aap, 573, A71
+
+\bibitem[{{Kuiper} {et~al.}(2015){Kuiper}, {Yorke}, \& {Turner}}]{kuiper:15}
+{Kuiper}, R., {Yorke}, H.~W., \& {Turner}, N.~J. 2015, \apj, 800, 86
+
+\bibitem[{{Lada} \& {Lada}(2003)}]{lada:03}
+{Lada}, C.~J. \& {Lada}, E.~A. 2003, \araa, 41, 57
+
+\bibitem[{{Lamb} {et~al.}(2016){Lamb}, {Oey}, {Segura-Cox}, {Graus}, {Kiminki},
+  {Golden-Marx}, \& {Parker}}]{lamb:16}
+{Lamb}, J.~B., {Oey}, M.~S., {Segura-Cox}, D.~M., {et~al.} 2016, \apj, 817, 113
+
+\bibitem[{{Langer} {et~al.}(2007){Langer}, {Norman}, {de Koter}, {Vink},
+  {Cantiello}, \& {Yoon}}]{langer:07}
+{Langer}, N., {Norman}, C.~A., {de Koter}, A., {et~al.} 2007, \aap, 475, L19
+
+\bibitem[{{Lennon} {et~al.}(2018){Lennon}, {Evans}, {van der Marel},
+  {Anderson}, {Platais}, {Herrero}, {de Mink}, {Sana}, {Sabbi}, {Crowther},
+  {Langer}, {Ramos Lerate}, {del Pino}, {Renzo}, {Sim{\'o}n-D{\'{\i}}az}, \&
+  {Schneider}}]{lennon:18}
+{Lennon}, D.~J., {Evans}, C.~J., {van der Marel}, R.~P., {et~al.} 2018,
+  ArXiv:1805.08277
+
+\bibitem[{{Lindegren} {et~al.}(2018){Lindegren}, {Hernandez}, {Bombrun},
+  {Klioner}, {Bastian}, {Ramos-Lerate}, {de Torres}, {Steidelmuller},
+  {Stephenson}, {Hobbs}, {Lammers}, {Biermann}, {Geyer}, {Hilger}, {Michalik},
+  {Stampa}, {McMillan}, {Castaneda}, {Clotet}, {Comoretto}, {Davidson},
+  {Fabricius}, {Gracia}, {Hambly}, {Hutton}, {Mora}, {Portell}, {van Leeuwen},
+  {Abbas}, {Abreu}, {Altmann}, {Andrei}, {Anglada}, {Balaguer-Nunez},
+  {Barache}, {Becciani}, {Bertone}, {Bianchi}, {Bouquillon}, {Bourda},
+  {Brusemeister}, {Bucciarelli}, {Busonero}, {Buzzi}, {Cancelliere},
+  {Carlucci}, {Charlot}, {Cheek}, {Crosta}, {Crowley}, {de Bruijne}, {de
+  Felice}, {Drimmel}, {Esquej}, {Fienga}, {Fraile}, {Gai}, {Garralda},
+  {Gonzalez-Vidal}, {Guerra}, {Hauser}, {Hofmann}, {Holl}, {Jordan},
+  {Lattanzi}, {Lenhardt}, {Liao}, {Licata}, {Lister}, {Loffler}, {Marchant},
+  {Martin-Fleitas}, {Messineo}, {Mignard}, {Morbidelli}, {Poggio}, {Riva},
+  {Rowell}, {Salguero}, {Sarasso}, {Sciacca}, {Siddiqui}, {Smart}, {Spagna},
+  {Steele}, {Taris}, {Torra}, {van Elteren}, {van Reeven}, \&
+  {Vecchiato}}]{lindengren:18}
+{Lindegren}, L., {Hernandez}, J., {Bombrun}, A., {et~al.} 2018,
+  ArXiv:1804.09366
+
+\bibitem[{{Maeder} \& {Meynet}(2000)}]{maeder:00}
+{Maeder}, A. \& {Meynet}, G. 2000, \aap, 361, 159
+
+\bibitem[{{Oh} \& {Kroupa}(2016)}]{oh:16}
+{Oh}, S. \& {Kroupa}, P. 2016, \aap, 590, A107
+
+\bibitem[{{Pietrzy{\'n}ski} {et~al.}(2013){Pietrzy{\'n}ski}, {Graczyk},
+  {Gieren}, {Thompson}, {Pilecki}, {Udalski}, {Soszy{\'n}ski}, {Koz{\l}owski},
+  {Konorski}, {Suchomska}, {Bono}, {Moroni}, {Villanova}, {Nardetto},
+  {Bresolin}, {Kudritzki}, {Storm}, {Gallenne}, {Smolec}, {Minniti}, {Kubiak},
+  {Szyma{\'n}ski}, {Poleski}, {Wyrzykowski}, {Ulaczyk}, {Pietrukowicz},
+  {G{\'o}rski}, \& {Karczmarek}}]{pietrzynski:13}
+{Pietrzy{\'n}ski}, G., {Graczyk}, D., {Gieren}, W., {et~al.} 2013, \nat, 495,
+  76
+
+\bibitem[{{Platais} {et~al.}(2018){Platais}, {Lennon}, {van der Marel},
+  {Bellini}, {Sabbi}, {Watkins}, {Sohn}, {Walborn}, {Bedin}, {Evans}, {de
+  Mink}, {Sana}, {Herrero}, {Langer}, \& {Crowther}}]{platais:18}
+{Platais}, I., {Lennon}, D.~J., {van der Marel}, R.~P., {et~al.} 2018,
+  ArXiv:1804.08678
+
+\bibitem[{{Platais} {et~al.}(2015){Platais}, {van der Marel}, {Lennon},
+  {Anderson}, {Bellini}, {Sabbi}, {Sana}, \& {Bedin}}]{platais:15}
+{Platais}, I., {van der Marel}, R.~P., {Lennon}, D.~J., {et~al.} 2015, \aj,
+  150, 89
+
+\bibitem[{{Poveda} {et~al.}(1967){Poveda}, {Ruiz}, \& {Allen}}]{poveda:67}
+{Poveda}, A., {Ruiz}, J., \& {Allen}, C. 1967, Boletin de los Observatorios
+  Tonantzintla y Tacubaya, 4, 86
+
+\bibitem[{{Rakavy} \& {Shaviv}(1967)}]{rakavy:67}
+{Rakavy}, G. \& {Shaviv}, G. 1967, \apj, 148, 803
+
+\bibitem[{{Ram{\'{\i}}rez-Agudelo} {et~al.}(2015){Ram{\'{\i}}rez-Agudelo},
+  {Sana}, {de Mink}, {H{\'e}nault-Brunet}, {de Koter}, {Langer}, {Tramper},
+  {Gr{\"a}fener}, {Evans}, {Vink}, {Dufton}, \& {Taylor}}]{ramirez-agudelo:15}
+{Ram{\'{\i}}rez-Agudelo}, O.~H., {Sana}, H., {de Mink}, S.~E., {et~al.} 2015,
+  \aap, 580, A92
+
+\bibitem[{{Renzo} {et~al.}(2018){Renzo}, {Zapartas}, {de Mink}, {G{\"o}tberg},
+  {Justham}, {Farmer}, {Izzard}, {Toonen}, \& {Sana}}]{renzo:18}
+{Renzo}, M., {Zapartas}, E., {de Mink}, S.~E., {et~al.} 2018, ArXiv:1804.09164
+
+\bibitem[{{Robitaille} \& {Bressert}(2012)}]{robitaille:12}
+{Robitaille}, T. \& {Bressert}, E. 2012, {APLpy: Astronomical Plotting Library
+  in Python}, Astrophysics Source Code Library
+
+\bibitem[{{Rosen} {et~al.}(2016){Rosen}, {Krumholz}, {McKee}, \&
+  {Klein}}]{rosen:16}
+{Rosen}, A.~L., {Krumholz}, M.~R., {McKee}, C.~F., \& {Klein}, R.~I. 2016,
+  \mnras, 463, 2553
+
+\bibitem[{{Rubio-D{\'{\i}}ez} {et~al.}(2017){Rubio-D{\'{\i}}ez}, {Najarro},
+  {Garc{\'{\i}}a}, \& {Sundqvist}}]{rubio-diez:17}
+{Rubio-D{\'{\i}}ez}, M.~M., {Najarro}, F., {Garc{\'{\i}}a}, M., \& {Sundqvist},
+  J.~O. 2017, in IAU Symposium, Vol. 329, The Lives and Death-Throes of Massive
+  Stars, ed. J.~J. {Eldridge}, J.~C. {Bray}, L.~A.~S. {McClelland}, \&
+  L.~{Xiao}, 131--135
+
+\bibitem[{{Sabbi} {et~al.}(2012){Sabbi}, {Lennon}, {Gieles}, {de Mink},
+  {Walborn}, {Anderson}, {Bellini}, {Panagia}, {van der Marel}, \& {Ma{\'{\i}}z
+  Apell{\'a}niz}}]{sabbi:12}
+{Sabbi}, E., {Lennon}, D.~J., {Gieles}, M., {et~al.} 2012, \apjl, 754, L37
+
+\bibitem[{{Sana} {et~al.}(2017){Sana}, {Ram{\'{\i}}rez-Tannus}, {de Koter},
+  {Kaper}, {Tramper}, \& {Bik}}]{sana:17}
+{Sana}, H., {Ram{\'{\i}}rez-Tannus}, M.~C., {de Koter}, A., {et~al.} 2017,
+  \aap, 599, L9
+
+\bibitem[{{Schneider} {et~al.}(2017){Schneider}, {Castro}, {Fossati}, {Langer},
+  \& {de Koter}}]{schneider:17}
+{Schneider}, F.~R.~N., {Castro}, N., {Fossati}, L., {Langer}, N., \& {de
+  Koter}, A. 2017, \aap, 598, A60
+
+\bibitem[{{Schneider} {et~al.}(2018){Schneider}, {Sana}, {Evans},
+  {Bestenlehner}, {Castro}, {Fossati}, {Gr{\"a}fener}, {Langer},
+  {Ram{\'{\i}}rez-Agudelo}, {Sab{\'{\i}}n-Sanjuli{\'a}n},
+  {Sim{\'o}n-D{\'{\i}}az}, {Tramper}, {Crowther}, {de Koter}, {de Mink},
+  {Dufton}, {Garcia}, {Gieles}, {H{\'e}nault-Brunet}, {Herrero}, {Izzard},
+  {Kalari}, {Lennon}, {Ma{\'{\i}}z Apell{\'a}niz}, {Markova}, {Najarro},
+  {Podsiadlowski}, {Puls}, {Taylor}, {van Loon}, {Vink}, \&
+  {Norman}}]{schneider:18}
+{Schneider}, F.~R.~N., {Sana}, H., {Evans}, C.~J., {et~al.} 2018, Science, 359,
+  69
+
+\bibitem[{{Smith}(2016)}]{smith:16}
+{Smith}, N. 2016, \mnras, 461, 3353
+
+\bibitem[{{Smith} \& {Tombleson}(2015)}]{smith:15}
+{Smith}, N. \& {Tombleson}, R. 2015, \mnras, 447, 598
+
+\bibitem[{{Udalski} {et~al.}(2008){Udalski}, {Soszy{\'n}ski}, {Szyma{\'n}ski},
+  {Kubiak}, {Pietrzy{\'n}ski}, {Wyrzykowski}, {Szewczyk}, {Ulaczyk}, \&
+  {Poleski}}]{udalski:08}
+{Udalski}, A., {Soszy{\'n}ski}, I., {Szyma{\'n}ski}, M.~K., {et~al.} 2008,
+  \actaa, 58, 329
+
+\bibitem[{{Ward} \& {Kruijssen}(2018)}]{ward:18}
+{Ward}, J.~L. \& {Kruijssen}, J.~M.~D. 2018, \mnras, 475, 5659
+
+\bibitem[{{Zapartas} {et~al.}(2017){Zapartas}, {de Mink}, {Izzard}, {Yoon},
+  {Badenes}, {G{\"o}tberg}, {de Koter}, {Neijssel}, {Renzo}, {Schootemeijer},
+  \& {Shrotriya}}]{zapartas:17}
+{Zapartas}, E., {de Mink}, S.~E., {Izzard}, R.~G., {et~al.} 2017, \aap, 601,
+  A29
+
+\bibitem[{{Zinnecker} \& {Yorke}(2007)}]{zinnecker:07}
+{Zinnecker}, H. \& {Yorke}, H.~W. 2007, \araa, 45, 481
+
+\bibitem[{{Zwicky}(1957)}]{zwicky:57}
+{Zwicky}, F. 1957, \zap, 44, 64
+\end{thebibliography}
 
@@
   \small
-  We are grateful to S.~Torres for 
+  We are grateful to S.~Torres and C.~J.~Evans for 
   very useful discussions on the

### 2018-05-29 scripts/plot_defaults.py
moved to MNRAS letter format, still too long and authors list looks ugly

@@ -1,430 +1,0 @@
-#/usr/bin/env python
-
-import scipy as sp
-import scipy.optimize as spo
-import matplotlib as mpl
-import matplotlib.pyplot as plt
-import matplotlib.ticker as mticker
-import itertools
-from matplotlib import rc
-from matplotlib import rcParams
-
-rc('text', usetex=True)
-rc('font', family='serif')
-rc('font', serif='palatino')
-# rc('font', weight='bolder')
-rc('mathtext', default='sf')
-rc("lines", markeredgewidth=2)
-rc("lines", linewidth=3)
-rc('axes', labelsize=30)  # 24
-rc("axes", linewidth=2)  # 2)
-rc('xtick', labelsize=28)
-rc('ytick', labelsize=28)
-rc('legend', fontsize=20)  # 16
-rc('xtick.major', pad=8)  # 8)
-rc('ytick.major', pad=8)  # 8)
-rc('xtick.major', size=13)
-rc('ytick.major', size=13)
-rc('xtick.minor', size=7)
-rc('ytick.minor', size=7)
-rcParams['text.latex.preamble'] = [r"\usepackage{color}"]
-rcParams['text.latex.preamble'] = [r"\usepackage{xcolor}"]
-rcParams['text.latex.preamble'] = [r"\usepackage{amsmath}"]
-
-# some colors
-green1 = '#00ff00'
-orange = '#ffba00'
-
-
-def setup_plot_with_vertical_panels(fig, left, bottom, heights, width, panels):
-    """left: left starting point of panels
-       bottom: bottom starting point of panels
-       width: width of panels
-       hieghts: an array of individual panel heights
-    """
-
-    # setup plotting rectangles
-    rects = []
-    total_height = 0
-    for ii in range(0, panels):
-        current_left = left
-        current_bottom = bottom + total_height
-        current_height = heights[ii]
-        current_width = width
-        total_height += current_height
-        rects.append([current_left, current_bottom,
-                      current_width, current_height])
-
-    # Add axes and
-    # share x-axis of bottom panel and hide x-axis labeling of upper panels
-    ax = []
-    ax.append(fig.add_axes(rects[0]))
-    for ii in range(1, panels):
-        ax.append(fig.add_axes(rects[ii], sharex=ax[0]))
-        for label in ax[ii].get_xticklabels():
-            label.set_visible(False)
-
-    return rects, ax
-
-
[…]

### 2018-06-05 VFTS682.tex
text trimming

--- 2018-05-27 VFTS682.tex
@@
-\documentclass[apjl,twocolumn]{emulateapj}
-\usepackage{graphicx}
-\usepackage[varg]{txfonts}
-%\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
+\documentclass[a4paper,fleqn,usenatbib]{mnras}
+\usepackage{newtxtext,newtxmath}
+\usepackage[T1]{fontenc}
+\usepackage{ae,aecompl}
+
+%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
+
+% Only include extra packages if you really need them. Common packages are:
+\usepackage{graphicx}	% Including figure files
+\usepackage{amsmath}	% Advanced maths commands
+\usepackage{amssymb}	% Extra maths symbols
 \usepackage[%draft, 
-pagebackref,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
-\renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
-\usepackage{aas_macros}
-%\usepackage{amsmath} % clashes with emulateapjj for some reaons
+% pagebackref,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue
+]{hyperref}
 \usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{comment}
+\usepackage{url}
 \usepackage{multirow}
-%\usepackage{chngpage}
-%\usepackage{lscape}
-\usepackage{url}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
 \newcommand{\todo}[1]{{\large $\blacksquare$~\textbf{\color{red}[#1]}}~$\blacksquare$}
 \newcommand{\udef}{\stackrel{\mathrm{def}}{=}}
-% for tree diagram
-%\usepackage{forest}
-%\usepackage{tikz-qtree}
-%\usetikzlibrary{shadows,trees}
 
@@
 
-\begin{document}
+% Please keep new commands to a minimum, and use \newcommand not \def to avoid
+% overwriting existing commands. Example:
+%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
 
-% \title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
-%                                 % ejected %% to fit in one line
-% runaway star known to date}
-\title{\emph{Gaia} DR2 adds clues for dynamical ejection of
-  a $\sim$$150\,M_\odot$ star}
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  R.~P.~van~der~Marel$^{3,4}$,   I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
-  V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
-  N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
-\affil{$^{1}${Astronomical Institute Anton Pannekoek, University of
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
+
+% Title of the paper, and the short title which is used in the headers.
+% Keep the title short and informative.
+\title[\emph{Gaia} DR2 adds clues for dynamical ejections of a $\sim$$150\,M_\odot$ star]{\emph{Gaia} DR2 adds clues for dynamical ejection of a $\sim$$150\,M_\odot$ star}
+
+% The list of authors, and the short list which is used in the headers.
+% If you need two or more lines of authors, add an extra line using \newauthor
+\author[Renzo et al.]{
+M.~Renzo$^{1}$\thanks{Corresponding author: M.~Renzo
+  \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}},
+S.~E.~de~Mink$^{1}$,
+D.~J.~Lennon$^{2}$,
+R.~P.~van~der~Marel$^{3,4}$,
+I.~Platais$^{4}$,
+J. Bestenlehner$^{5}$,
+V.~H\'enault-Brunet$^{6}$,
+S.~Justham$^{7,8}$,
+A.~de~Koter$^{1}$,
+N.~Langer$^{9}$,
+H.~Sana$^{10}$,
+F.~R.~Schneider$^{11}$,
+J.~S.~Vink$^{12}$ \\
+% List of institutions
+$^{1}${Astronomical Institute Anton Pannekoek, University of
   Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
@@
   Oxford OX1 3RH, UK} \\
-  $^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
+$^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}
 }
 
- \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
-\date{}
+% These dates will be filled out by the publisher
+\date{Accepted XXX. Received YYY; in original form ZZZ}
+
+% Enter the current year, for the copyright statements etc.
+\pubyear{2018}
+
+% Don't change these lines
+\begin{document}
+\label{firstpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\maketitle
+
+% Abstract of the paper
 \begin{abstract}
+
   The formation and fates of the most massive stars remain among the most elusive questions in stellar astrophysics. 
   % 
-  Previous spectroscopic studies identified VFTS 682 as one of the most
+  Previous spectroscopic studies identified VFTS682 as one of the most
   massive stars known. It resides in the field of the 30 Dor region in
@@
   % 
-  We use the unprecedented proper motions from the second \emph{Gaia} data
+  We use the proper motions from the second \emph{Gaia} data
   release (DR2) together with multiepoch \emph{Hubble Space Telescope} proper motions to investigate the kinematics of this object. We show
-  that the projected velocity of VFTS 682 is consistent with it being a
+  that the projected velocity of VFTS682 is consistent with it being a
   runaway star dynamically ejected from R136. Its
@@
   % 
-  If confirmed by future astrometric data, this would make VFTS 682 the most massive runaway star known to date.
-It proves that star clusters are capable of ejecting their most massive members.
+  If confirmed by future astrometric data, this would make VFTS682
+  the most massive runaway star known to date.
+It would prove that star clusters are capable of ejecting their most massive members.
 We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
-
 \end{abstract}
 
-\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS 682}
-\maketitle{}
+% Select between one and six entries from the list of approved keywords.
+% Don't make up new ones.
+\begin{keywords}
+  stars: kinematics, stars: runaways, stars: individual: VFTS682
+\end{keywords}
 
-\todo{trim ~1085 words}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
 
@@
 \citep[e.g.][]{kuiper:15,rosen:16}, but simulations cannot yet resolve
-the smallest length scales that are of interest to understand the 
-high level of multiplicity among massive stars  \citep[e.g.,][]{bate:09, sana:17}. Understanding massive star formation, and its
+the length scales  of interest to explain the 
+high multiplicity of massive stars  \citep[e.g.,][]{bate:09, sana:17}. Understanding massive star formation, and its
 possible dependence on environment and metallicity, is crucial for
@@
 
-An interesting contribution to the debate on whether or not massive
+A contribution to the debate on whether or not massive
 stars form in relative isolation was presented by
 \cite{bestenlehner:11} and \cite{bressert:12}, who discussed the case of the very massive star
-VFTS 682.  This star is located in the field of the 30 Doradus region
+VFTS682.  This star is located in the field of the 30 Doradus region
 in the Large Magellanic Cloud (LMC) and was studied as part of the
-multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of
-massive stars \citep{evans:11}. Its spectral type is WNh5 and its 
-present-day mass is $137.8^{+27.5}_{-15.9}\,M_\odot$,
+multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}. Its spectral type is WNh5 and its present-day mass is $137.8^{+27.5}_{-15.9}\,M_\odot$,
 \citep{schneider:18}. These correspond to an inferred initial mass of
-$150.0^{+28.7}_{-17.4}\,M_\odot$. This makes VFTS 682 one of the
+$150.0^{+28.7}_{-17.4}\,M_\odot$ and make VFTS682 one of the
 most extreme objects in the region.
-
+%
 From the spectral point of view, it is reminiscent of the very
@@
 R136 cluster. In particular, a remarkable similarity exist between the
-spectrum of VFTS 682 and R136a3, located in the core of the cluster
+spectrum of VFTS682 and R136a3, located in the core of the cluster
 \citep{crowther:10} for which \citet{crowther:16} report a
@@
 at least two more very massive WN5h stars, R136a1 and R136a2, whose
-estimated current masses are even higher. However, VFTS 682 stands
+estimated current masses are even higher. However, VFTS682 stands
 out by its isolation at a projected offset of 119.4 arcseconds, corresponding to $\sim$$29$\,pc from the core of the star cluster
@@
 \citet{bestenlehner:11} discuss two possibilities for the
-origin of VFTS 682. % Quoting directly 
+origin of VFTS682. % Quoting directly 
 From their abstract: "(i) the
 star either formed in situ, which would have profound implications for
-the formation mechanism of massive stars, or (ii) VFTS 682 is a slow
+the formation mechanism of massive stars, or (ii) VFTS682 is a slow
 runaway star that originated from the dense cluster R136, which would
@@
 dynamical models, \citet{fujii:11} and \citet{banerjee:12} argued that
-ejection of VFTS 682 from R136 would be natural. Testing the origin of
-VFTS 682 is not only important for constraining isolated star
+ejection of VFTS682 from R136 would be natural. Testing the origin of
+VFTS682 is not only important for constraining isolated star
 formation; it should also help to improve understanding of early star-cluster dynamics.
@@
 runaway stars in
- 30 Doradus based on proper motion
-measurements. \citet{lennon:18} used data from the second \emph{Gaia}
+ 30 Doradus based on their proper motions. \citet{lennon:18} used data from the second \emph{Gaia}
 data release \cite[DR2,][]{gaia:16,brown:18} to further investigate
-the isolated O-type stars in the vicinity of R136. In this study, we
-report on our analysis of the \emph{Gaia} DR2 data for VFTS 682 which
+the isolated O-type stars around R136. In this study, we
+report on our analysis of the \emph{Gaia} DR2 data for VFTS682 which
 was not part of the sample of \citet{lennon:18} because of its 
@@
 VFTS survey \citep[][]{evans:11} with the proper motion from
-\emph{Gaia} DR2 to reconstruct the three-dimensional velocity of VFTS
-682, and test the hypothesis that this star was ejected from R136. We
-use the HST proper motions from \citet{platais:18} to validate our
+\emph{Gaia} DR2 to reconstruct the three-dimensional velocity of VFTS682, and test the hypothesis that this star was ejected from R136. We
+use the HST proper motions from \citet{platais:18} to test our
 analysis of the \emph{Gaia} DR2 data.
@@
 %%% cut for length reasons: all is said below
-% Our results indicate that VFTS 682 is a runaway star, although with
+% Our results indicate that VFTS682 is a runaway star, although with
 % low statistical significance. The
@@
 
-\subsection{ Overview of VFTS 682 from previous studies \label{data:vfts683}}
+\subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
 
-The star VFTS 682, located at right ascension (RA)
+The star VFTS682, located at right ascension (RA)
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} \citep[][see also
-\Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2]{evans:11}
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000 \citep[][% see also
+% \Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2
+]{evans:11}
 was originally classified as a young stellar object \citep{gruendl:09}
-based on its mid-infrared excess. \citet{evans:11} classified the
-object as Wolf-Rayet star of spectral type WNh5 using spectra from the
-VFTS survey. The spectra available covered $\lambda$4000--7000 and
-were taken at multiple epochs. Using the same dataset,
+based on its mid-infrared excess. \citet{evans:11} reclassified the
+object as Wolf-Rayet star of spectral type WNh5 using multi-epoch
+spectra covering $\lambda$4000--7000 from the
+VFTS survey. % The spectra available covered $\lambda$4000--7000 and
+% were taken at multiple epochs.
+
+Using the same dataset,
 \citet{bestenlehner:11} excluded the presence of a close companion
 with high confidence from the absence of radial velocity variations.
-
-\citet{bestenlehner:11} further analyzed the spectra % using the non-LTE
-% model atmosphere code CMFGEN \citep{hillier:98}
-to derive the stellar
+\citet{bestenlehner:11} % further analyzed the spectra % using the non-LTE
+% % model atmosphere code CMFGEN \citep{hillier:98}
+also derived stellar
 parameters and surface abundances. They inferred an extinction
@@
 $\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citet{schneider:18} estimated
-a present-day mass for VFTS 682 of $137.8^{+27.5}_ {-15.9}\Msun$, an
+a present-day mass for VFTS682 of $137.8^{+27.5}_ {-15.9}\Msun$, an
 apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
@@
 \citep{schneider:17}, based on the evolutionary models from
-\citet{brott:11} and \cite{kohler:15}. % These values place VFTS 682 around the
+\citet{brott:11} and \cite{kohler:15}. % These values place VFTS682 around the
 % ``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
 
-\citet{bestenlehner:11} reported for VFTS 682 a wind mass loss rate of
-$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, inferred from
-spectral analysis assuming a smooth outflow, i.e., not accounting for
+\citet{bestenlehner:11} reported for VFTS682 a wind mass loss rate of
+$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, % inferred from
+% spectral analysis assuming a smooth outflow, i.e.,
+not accounting for
 clumping. The present-day surface helium mass
@@
 % at least two more very massive WN5h stars, R136a1 and R136a2, whose
-% estimated current masses are even higher. However, VFTS 682 stands
+% estimated current masses are even higher. However, VFTS682 stands
 % out by its isolation at a projected offset of 29\,pc from the center
@@
 Further worth noticing is the variability of the
-star. % \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
-% lower than what found by \citet{evans:11}.
+star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
+lower than what found by \citet{evans:11}.
 \citet{bestenlehner:11} discuss the Optical Gravitational Lensing
@@
 
-We obtain the peculiar radial velocity $\delta v_\mathrm{rad}$ of VFTS 682 as the difference
+We obtain the peculiar radial velocity $\delta v_\mathrm{rad}$ of VFTS682 as the difference
 between the average radial velocity of the 30 Doradus region
 ($270\pm10\kms$) minus the radial velocity measured from the NV $\lambda4944$
-line for VFTS 682  \citep[$300\pm10\kms$, ][]{bestenlehner:11}. We implicitly assume that the use
+line for VFTS682  \citep[$300\pm10\kms$, ][]{bestenlehner:11}. We implicitly assume that the use
 of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not
@@
 The catalog from \citet{platais:18} lists the proper motion
-components of VFTS 682 relative to the field, providing a
+components of VFTS682 relative to the field, providing a
 measure of the motion of this star completely independent from
@@
   \begin{center}
-    \caption{Astrometric parameters for VFTS 682. }
+    \caption{Astrometric parameters for VFTS682. }
     \begin{tabular}{l|c|c}
-      \tableline
-      \tableline
+      \hline
+      \hline
       Parameter & Value & Source\\
-      \tableline
+      \hline
       RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
-                        & \multirow{4}{*}{\emph{Gaia} DR2}\\[5pt]
+                        & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
       DEC \hfill [degree] & -69.07 % $\pm$  0.05
                         & \\[5pt]
-      \cline{1-2}
-      $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \\[5pt]
+      \hline
+      $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \multirow{2}{*}{\emph{Gaia} DR2}\\[5pt]
       $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
@@
       $\delta v_\mathrm{rad}$  \hfill[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
-      \tableline
+      \hline
     \end{tabular}
-    \tablecomments{We do not list the error on the RA and DEC positions,
+    % \tablecomments
+    {We do not list the error on the RA and DEC positions,
       which can be retrieved from the Gaia DR2 catalog and are of order
@@
 parallaxes, and proper motion components) for more than a billion
-sources \citep{brown:18}. VFTS 682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2
+sources \citep{brown:18}. VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2
 catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. \emph{Gaia}
@@
 \texttt{astrometric\_excess\_noise} = 0. These values suggest that the \emph{Gaia}
-data for VFTS 682 are reliable. However, the effective temperature
+data for VFTS682 are reliable. However, the effective temperature
 reported in \emph{Gaia} DR2, based on the spectral fitting around the
@@
 
-We retrieve for VFTS 682 the position in RA and DEC
+We retrieve for VFTS682 the position in RA and DEC
 in the International Celestial Reference System \cite[][]{brown:18}, its
@@
 
-For the radial velocity of VFTS 682 and of the 30 Doradus
+For the radial velocity of VFTS682 and of the 30 Doradus
 region as a whole, we instead use the VFTS data
@@
 We follow \citet{lennon:18} to define a local frame of reference and derive the peculiar velocity
-of VFTS 682 with respect to the cluster R136. Their selection (see
+of VFTS682 with respect to the cluster R136. Their selection (see
 their Sec.~2.1) of 153
@@
 % We define a local standard frame of reference to derive the peculiar velocity
-% of VFTS 682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
+% of VFTS682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
 % We select all the stars in a target of 0.05 degrees around R136
@@
 
-\section{The kinematics of VFTS 682}
+\section{The kinematics of VFTS682}
 \label{sec:results}
@@
 \label{sec:runaway}
-We first address the question of whether VFTS 682 is a typical star
+We first address the question of whether VFTS682 is a typical star
 from the kinematic point of view, or whether it is a runaway star with
@@
 \citep{lennon:18} from the
-proper motion of VFTS 682 (see \Tabref{tab:vfts682}), we obtain the
+proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
 components of proper motion of the star relative to the surrounding region
@@
 from R136, to highlight the likelihood of it being the origin of fast
-moving stars. The plus sign marks VFTS 682, which sits at the edge of
+moving stars. The plus sign marks VFTS682, which sits at the edge of
 the proper motion distribution, but is not a clear outlier. Most
-relevant is the direction of the proper motion of VFTS 682, which we
+relevant is the direction of the proper motion of VFTS682, which we
 discuss in \Secref{sec:r136_origin}.
 
-\begin{figure}[htbp]
+\begin{figure}%[htbp]
   \centering
@@
 Therefore, two completely independent measures of the proper motion of
-VFTS 682 relative to the surrounding field, one from HST and one from
+VFTS682 relative to the surrounding field, one from HST and one from
 \emph{Gaia} DR2, yield values of the peculiar three-dimensional
-velocity of VFTS 682 (\Eqref{eq:speed_around_HST} and
+velocity of VFTS682 (\Eqref{eq:speed_around_HST} and
 \Eqref{eq:speed_around_Gaia}, respectively) which would make it the most massive runaway star
@@
 
-Assuming that the best estimate of the peculiar velocity of VFTS 682
+Assuming that the best estimate of the peculiar velocity of VFTS682
 are reliable, we now address the question of its likely origin.
@@
 The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
-VFTS 682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the cyan arrows show the possible
+VFTS682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the cyan arrows show the possible
 range of directions within the uncertainties in the measured relative proper
-motion. These arrows cross at the present-day location of VFTS 682 and
+motion. These arrows cross at the present-day location of VFTS682 and
 are prolonged in the direction opposite to the motion to illustrate
@@
 \end{equation}
-where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS 682 to
+where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
 the core of R136 \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
@@
 
-\begin{figure}[tbp]
+\begin{figure}%[tbp]
   \centering
   \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
-  \caption{The red solid arrow indicates the proper motion of VFTS 682
+  \caption{The red solid arrow indicates the proper motion of VFTS682
     relative to the region from \emph{Gaia} DR2, starting from the present day position of
@@
     star. The length of the prolongations is proportional to the relative proper motion
-    times the kinetic age of VFTS 682. \todo{load color picture on background? Fig with HST data
-      available, but confusing...}}
+    times the kinetic age of VFTS682. % \todo{load color picture on background? Fig with HST data
+      % available, but confusing...}
+  }
 
@@
 
-Based on our results, we tentatively claim that VFTS 682 is the most massive
+Based on our results, we tentatively claim that VFTS682 is the most massive
 runaway known to date, with a peculiar three-dimensional speed of
@@
 massive star among those interacting \cite[e.g.,][]{banerjee:12}. This means that, just
-based on the kinematic properties of VFTS 682, we would expect several
+based on the kinematic properties of VFTS682, we would expect several
 stars with initial masses larger than $\sim$$150\,M_\odot$ in the
@@
   velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
-  considered with caution.} of VFTS 682
+  considered with caution.} of VFTS682
 reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
 line with the average rotation rate of massive stars in the region
-\citep[][]{ramirez-agudelo:15}. This suggests that VFTS 682 (i) has not
+\citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
 experienced rotationally induced chemically homogeneous evolution
@@
 data of the VFTS survey rule out the presence of a close companion at the
-present day. Moreover, the spectral type of VFTS 682
+present day. Moreover, the spectral type of VFTS682
 \citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
@@
 the spectrum of R136a3. Therefore,
-VFTS 682 might be an ideal target to constrain the stellar physics of
+VFTS682 might be an ideal target to constrain the stellar physics of
 stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
@@
 \citet{banerjee:12} used N-body simulations of fully segregated
-clusters with all massive stars in binaries to suggest that VFTS 682
+clusters with all massive stars in binaries to suggest that VFTS682
 was ejected from R136. They
@@
 model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS 682 and the massive members of R136. While it is not possible to
-robustly exclude that VFTS 682 is itself a merger, its spectrum does not show any of the obvious signatures
+VFTS682 and the massive members of R136. While it is not possible to
+robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
 like fast rotation, which however could have already slowed down
@@
 is also possible that the star or binary that caused the ejection of
-VFTS 682 might have been ejected in the opposite direction, and is also
+VFTS682 might have been ejected in the opposite direction, and is also
 isolated at present day. If the ejection was caused by an interaction
@@
 
-The similarities between VFTS 682 and the WNh5 stars in the core of
+The similarities between VFTS682 and the WNh5 stars in the core of
 R136 are also in agreement with the ``bully binary'' model of
@@
 massive binary, which then tightens its orbit by ejecting other stars passing
-by. Interpreting our results for VFTS 682 through the lens of their simulations
+by. Interpreting our results for VFTS682 through the lens of their simulations
 suggests the presence of a close binary with total mass
@@
 pair-instability supernova \cite[e.g.,][]{rakavy:67} at LMC
-metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS 682 could be either a
+metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS682 could be either a
 pair-instability supernova without compact remnant formation, or
@@
 
-The kinematic age of VFTS 682 puts an
+The kinematic age of VFTS682 puts an
 upper limit to the timescale to form the ``bully binary'' in
@@
 evolution, given the age estimate of $\lesssim 2$\,Myr
-\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS 682. If the
+\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS682. If the
 cluster is indeed younger than the shortest stellar lifetime
 \citep[$\sim$3\,Myr, e.g.,][]{brott:11, zapartas:17}, then the alternative
-explanation for ejection of VFTS 682 from the disruption of a binary
+explanation for ejection of VFTS682 from the disruption of a binary
 by a core-collapse event is excluded since the region is too young for stars
@@
 
-The variability of VFTS 682, reminiscent of LBV stars, suggests
-that VFTS 682 (and therefore its analogs in the core of R136) might
+The variability of VFTS682, reminiscent of LBV stars, suggests
+that VFTS682 (and therefore its analogs in the core of R136) might
 experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
 debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
-claim that LBV stars are typically isolated form O-type stars. The fact that VFTS 682 is a dynamically
+claim that LBV stars are typically isolated form O-type stars. The fact that VFTS682 is a dynamically
 ejected runaway which might evolve into an LBV star suggests that
@@
 of motion. The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
-VFTS 682 (see \Secref{sec:r136_origin}) is smaller than the
+VFTS682 (see \Secref{sec:r136_origin}) is smaller than the
 corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic
 age of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
-massive VFTS 682 was ejected later than VFTS16 from the same cluster.
+massive VFTS682 was ejected later than VFTS16 from the same cluster.
 
@@
 
-VFTS 682 is the most massive runaway known to date, and its ejection
+VFTS682 is potentially the most massive runaway known to date, and its ejection
 from the cluster R136 likely implies that it is only the ``tip of the
@@
 
-\begin{acknowledgements}
 \small
-  We are grateful to S.~Torres for 
-  very useful discussions on the
-  \emph{Gaia} DR2 dataset, and to M.~C.~Ramirez-Tannus for invaluable help.
-  SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research
-  Council (ERC) (Grant agreement No. 715063). VHB acknowledges support from the NRC-Canada Plaskett Fellowship.
+We acknowledge invaluable help from S.~Torres, M.~C.~Ramirez-Tannus,
+and C.~J.~Evans.
+SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC) (Grant agreement No.
+715063).
+VHB acknowledges support from the NRC-Canada Plaskett Fellowship.
 This work has made use of data from the ESA space mission \emph{Gaia} (\url{http://www.cosmos.esa.int/gaia}), processed by the \emph{Gaia} Data Processing and Analysis Consortium (DPAC, \url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}).
@@
 This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
-\end{acknowledgements}
 
+% Don't change these lines
+\bsp	% typesetting comment
+\label{lastpage}
 \end{document}
 
+% End of mnras_template.tex
 %%% Local Variables:

### 2018-06-05 VFTS682.tex
text trimming 2

@@
 star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
-lower than what found by \citet{evans:11}.
+lower than found by \citet{evans:11}.
 \citet{bestenlehner:11} discuss the Optical Gravitational Lensing
@@
 of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not
-introduce significant errors. \cite{bressert:12} also noted that VFTS
-682 shows a significant radial velocity offset compared to the nebular
+introduce significant errors. \cite{bressert:12} also noted that VFTS682 shows a significant radial velocity offset compared to the nebular
 lines from the gas surrounding it, suggesting peculiar motion in the
@@
 presence of an optically thick, likely non-homogeneous, and variable
-wind obscuring the stellar surface. The errors quoted above on the
-radial velocities include only statistical uncertainties from the
+wind obscuring the stellar surface. The radial velocity errors quoted
+above quantify only statistical uncertainties from the
 spectral analysis, and not the systematic errors arising from the
@@
 Recently, \citet{platais:18} presented HST  astrometry and proper
-motions in the 30 Doradus region based on two epochs of
-observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
-in October 3-8 2011 and the second epoch (GO-13359) in
-October 6-11 2014, both using the F775W filter.
-The second epoch was designed to match the pointing and
-orientation of the first epoch as exactly as possible so
-that the displacement measurements could be purely differential.
-While most bright massive stars (V$<$14) were saturated in
+motions in the 30 Doradus region based on two epochs. %  of
+% observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
+% in October 3-8 2011 and the second epoch (GO-13359) in
+% October 6-11 2014, both using the F775W filter.
+% The second epoch was designed to match the pointing and
+% orientation of the first epoch as exactly as possible so
+% that the displacement measurements could be purely differential.
+While the most bright stars (V$<$14) were saturated in
 their data and missing from that catalog, due to its
-high extinction, this star is faint enough
+high extinction, VFTS682 is faint enough
 to be included in that catalog. Its proper motions are
@@
 v_\mathrm{RA}^{HST}=2\pm31\kms$ and $\delta
-v_\mathrm{RA}^{HST}=47\pm24\kms$, corresponding to a two-dimensional
-projected velocity of $47\pm24\kms$. Including the radial velocity
+v_\mathrm{RA}^{HST}=47\pm24\kms$. % , corresponding to a two-dimensional
+% projected velocity of $47\pm24\kms$.
+Including the radial velocity
 measured by the VFTS survey we obtain a three-dimensional velocity of
@@
 
-The second \emph{Gaia} Data Release (DR2), which became available on 25 April 2018,
+The second \emph{Gaia} Data Release (DR2)% , which
+became available on 25 April 2018 and 
 provides the five-parameter astrometric solution (positions,
@@
 in this study, and we attribute the unphysical value of the parallax
-to the large distance to the LMC. Our main findings do not rely on the
-parallax nor the effective temperature values reported in the \emph{Gaia} DR2
-catalog.
+to the large distance to the LMC.%  Our main findings do not rely on the
+% parallax nor the effective temperature values reported in the \emph{Gaia} DR2
+% catalog.
 
-We retrieve for VFTS682 the position in RA and DEC
-in the International Celestial Reference System \cite[][]{brown:18}, its
-proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
-respectively). While \emph{Gaia} DR2 data can be correlated, this is
-not an issue for VFTS682, and for simplicity we treat them as
-uncorrelated in the following.
-
+While \emph{Gaia} DR2 data can be correlated, this is
+not an issue for VFTS682, and we treat the proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
+respectively) of the star as uncorrelated for simplicity.
+% We retrieve for VFTS682 the position in RA and DEC
+% in the International Celestial Reference System \cite[][]{brown:18},
+% its
+% proper motion components .
+%  and for simplicity we treat them as
+% uncorrelated in the following.
 For the radial velocity of VFTS682 and of the 30 Doradus
@@
 bright ($G<17$) stars around R136 with reliable astrometric data from
-\emph{Gaia} DR2 yields $\mu_\mathrm{RA}=1.74\pm0.01\,\mathrm{mas\
-  yr^{-1}}$ and $\mu_\mathrm{DEC}=0.70\pm0.02\,\mathrm{mas\ yr^{-1}}$
+\emph{Gaia} DR2 yields $\langle\mu_\mathrm{RA}\rangle=1.74\pm0.01\,\mathrm{mas\
+  yr^{-1}}$ and $\langle\mu_\mathrm{DEC}\rangle=0.70\pm0.02\,\mathrm{mas\ yr^{-1}}$
 for the components of the mean motion of the region projected on the sky.
@@
 a significantly large peculiar velocity compared to its surrounding population. The former is what should
-be expected if it formed where we observe it today, in relative
-isolation from other massive stars.
+be expected if it formed in the relative isolation that we observe today.% , in relative
+% isolation from other massive stars.
 
@@
 Subtracting the mean proper motion components given by
-\citep{lennon:18} from the
+\citet{lennon:18} from the
 proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
@@
 The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
-VFTS682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the cyan arrows show the possible
-range of directions within the uncertainties in the measured relative proper
-motion. These arrows cross at the present-day location of VFTS682 and
+VFTS682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the
+orange arrows illustrate the uncertainty. % show the possible
+% range of directions within the uncertainties in the measured relative proper
+% motion.
+These arrows cross at the present-day location of VFTS682 and
 are prolonged in the direction opposite to the motion to illustrate
@@
 components results in a wide range of possible directions,we argue that the most likely origin of the star is R136,
-as suggested based on numerical N-body simulations by \cite{fujii:11}
+as suggested by the N-body simulations of \cite{fujii:11}
 and \cite{banerjee:12}.
 
-We can therefore consider the kinematic age for this star assuming
-that it originates from the cluster,
+The kinematic age of this star, assuming it originates from the
+cluster, is
 
@@
 
-The kinematic age $\tau_\mathrm{kin}$ is similar or smaller than the apparent age
-of the star of $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
-corroborates the idea that the star is the result of a dynamical
-ejection very soon in the cluster evolution. We emphasize that the
+The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
+ejection from the cluster, given is apparent age
+of $1.0\pm 0.2$\,Myr \citep{schneider:18}. % , which
+% corroborates the idea that the star is the result of a dynamical
+% ejection very soon in the cluster evolution.
+Note that the
 present-day surface helium abundance
 \citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
-limit on the age corresponding to the time needed for the star to
-synthesize that amount of helium, and such lower limit corresponds to
-$\sim$0.9\,Myr based on the models from \cite{kohler:15}.
+limit on the age of the star of $\sim$0.9\,Myr, corresponding to the time needed to
+synthesize this amount of helium in the models from \cite{kohler:15}.
 % \todo{discuss position angle}
@@
 of extremely massive stars by \cite{crowther:10} in the core of the
-cluster. The projected rotational equatorial
-velocity\footnote{However, the determination of the rotational
-  velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
-  considered with caution.} of VFTS682
-reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
-line with the average rotation rate of massive stars in the region
-\citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
-experienced rotationally induced chemically homogeneous evolution
-\citep[][]{maeder:00,demink:09}, and (ii) it has not
-accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
-data of the VFTS survey rule out the presence of a close companion at the
-present day. Moreover, the spectral type of VFTS682
+cluster. % The projected rotational equatorial
+% velocity\footnote{However, the determination of the rotational
+%   velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
+%   considered with caution.} of VFTS682
+% reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
+% line with the average rotation rate of massive stars in the region
+% \citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
+% experienced rotationally induced chemically homogeneous evolution
+% \citep[][]{maeder:00,demink:09}, and (ii) it has not
+% accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
+% data of the VFTS survey rule out the presence of a close companion at the
+% present day. Moreover,
+
+The spectral type of VFTS682
 \citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
@@
 \cite{dekoter:97,crowther:10,crowther:16}, with an astonishing similarity in particular with
-the spectrum of R136a3. Therefore,
-VFTS682 might be an ideal target to constrain the stellar physics of
-stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
-it an easier target for observations compared to the similar stars
-present in the crowded core of R136.
+the spectrum of R136a3. Therefore, the isolation of
+VFTS682 makes it an ideal target to constrain the stellar physics of
+stars with masses well above $\sim$$100\,M_\odot$ while avoiding
+crowding issues. %  : its isolation makes
+% it an easier target for observations compared to the similar stars
+% present in the crowded core of R136.
 
@@
 expected to have produced
-a large number of massive runaways. Indeed, several relatively
+a large number of massive runaways. Indeed, several %relatively
 isolated massive stars are observed in the region, some with known
-large radial velocities and/or proper motion (see
-\Figref{fig:pm_polar} and Sana et al., in prep.). A comprehensive study of the kinematic
+large radial velocities and/or proper motion. % (see
+% \Figref{fig:pm_polar} and Sana et al., in prep.).
+A comprehensive study of the kinematic
 properties of all the massive stars surrounding R136 might shed light
@@
 question of whether the cluster formed via a monolithic collapse, or
-as a (potentially ongoing) merger of several sub-structures \citep[e.g.,]{sabbi:12}.
+as a (potentially ongoing) merger of several sub-structures \citep[e.g.,][]{sabbi:12}.
 
@@
 
-\bibliographystyle{aa}
+\bibliographystyle{mnras}
 \bibliography{bibliography/vfts682}
@@
 \small
-We acknowledge invaluable help from S.~Torres, M.~C.~Ramirez-Tannus,
-and C.~J.~Evans.
+We are grateful to S.~Torres, M.~C.~Ramirez-Tannus,
+and C.~J.~Evans for help and discussions.
 SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC) (Grant agreement No.

### 2018-06-06 VFTS682.tex
backup
changed figure and caption

@@
 respectively) of the star as uncorrelated for simplicity.
-% We retrieve for VFTS682 the position in RA and DEC
-% in the International Celestial Reference System \cite[][]{brown:18},
-% its
-% proper motion components .
-%  and for simplicity we treat them as
-% uncorrelated in the following.
 For the radial velocity of VFTS682 and of the 30 Doradus
@@
 VFTS682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the
-orange arrows illustrate the uncertainty. % show the possible
+green arrows illustrate the uncertainty. % show the possible
 % range of directions within the uncertainties in the measured relative proper
@@
     relative to the region from \emph{Gaia} DR2, starting from the present day position of
-    the star. The orange arrows indicate the possible
+    the star. The green arrows indicate the possible
     directions of projected motion within the \emph{Gaia} DR2 errors, and are extended

### 2018-06-06 VFTS682.tex
./bibliography

@@
 $\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
-  yr^{-1}}$, corresponding to a relative proper motion
-
-\begin{equation}
-  \label{eq:pm_around_HST}
-  \delta \mu^{HST} = \sqrt{\left(\delta\mu_\mathrm{RA}^{HST}\right)^2+\left(\delta\mu_\mathrm{DEC}^{HST}\right)^2}
-  = 0.20 \pm 0.15\,\mathrm{mas\
-  yr^{-1}} \ \ .
-\end{equation}
+  yr^{-1}}$, which added in quadrature give a relative proper motion of
+$\delta \mu^{HST} =0.20 \pm 0.15\,\mathrm{mas\ yr^{-1}}$.
+% \begin{equation}
+%   \label{eq:pm_around_HST}
+%   \delta \mu^{HST} = \sqrt{\left(\delta\mu_\mathrm{RA}^{HST}\right)^2+\left(\delta\mu_\mathrm{DEC}^{HST}\right)^2}
+%   = 0.20 \pm 0.15\,\mathrm{mas\
+%   yr^{-1}} \ \ .
+% \end{equation}
 
@@
 % projected velocity of $47\pm24\kms$.
-Including the radial velocity
-measured by the VFTS survey we obtain a three-dimensional velocity of
-
-\begin{equation}
-  \label{eq:speed_around_HST}
-  v_\mathrm{pec}^{HST} = \sqrt{\left(\delta v_\mathrm{RA}^{HST}\right)^2
-    +\left(\delta v_\mathrm{DEC}^{HST}\right)^2+\left(\delta
-      v_\mathrm{rad}^{VFTS}\right)^2} = 56 \pm 23
-  \kms \ .
-\end{equation}
+Adding them in quadrature to the radial velocity
+measured by \citet{bestenlehner:11} we obtain a three-dimensional velocity of
+$v_\mathrm{pec}^{HST}=56 \pm 23  \kms$.
+% \begin{equation}
+%   \label{eq:speed_around_HST}
+%   v_\mathrm{pec}^{HST} = \sqrt{\left(\delta v_\mathrm{RA}^{HST}\right)^2
+%     +\left(\delta v_\mathrm{DEC}^{HST}\right)^2+\left(\delta
+%       v_\mathrm{rad}^{VFTS}\right)^2} = 56 \pm 23
+%   \kms \ .
+% \end{equation}
 
@@
     % \tablecomments
-    {We do not list the error on the RA and DEC positions,
-      which can be retrieved from the Gaia DR2 catalog and are of order
-      $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$. We adopt the peculiar
-      radial velocity calculated using the  NV $\lambda4944$ line, which is
-      less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).}
+    {The error on the RA and DEC positions, are of order
+      $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$ in \emph{Gaia}
+      DR2. %REPETITION: %  We adopt the peculiar
+      % radial velocity calculated using the  NV $\lambda4944$ line, which is
+      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
+    }
   \end{center}
@@
 \pm 0.09\,\mathrm{mas\ yr^{-1}}$. These components result in a
-two-dimensional relative proper motion of
-\begin{equation}
-  \label{eq:pm_gaia_around}
-  \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
-  = 0.13\pm 0.09\,\mathrm{mas\
-  yr^{-1}} \ \ .
-\end{equation}
+two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\
+  yr^{-1}}$.
+% \begin{equation}
+%   \label{eq:pm_gaia_around}
+%   \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
+%   = 0.13\pm 0.09\,\mathrm{mas\
+%   yr^{-1}} \ \ .
+% \end{equation}
 
@@
 \cite{bestenlehner:11} then gives the third component along
-the line of sight, allowing us to calculate the three-dimensional
-peculiar speed of the star:
-
-\begin{equation}
-  \label{eq:speed_around_Gaia}
-  v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
-    +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
-      v_\mathrm{rad}^{VFTS}\right)^2} = 44 \pm 21
-  \kms \ .
-\end{equation}
+the line of sight, % allowing us to calculate the three-dimensional
+% peculiar speed of the star:
+which added in quadrature to the transverse components results in a
+three-dimensional velocity of $44 \pm 21\kms$.
+% \begin{equation}
+%   \label{eq:speed_around_Gaia}
+%   v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
+%     +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
+%       v_\mathrm{rad}^{VFTS}\right)^2} = 44 \pm 21
+%   \kms \ .
+% \end{equation}
 
@@
 \emph{Gaia} DR2, yield values of the peculiar three-dimensional
-velocity of VFTS682 (\Eqref{eq:speed_around_HST} and
-\Eqref{eq:speed_around_Gaia}, respectively) which would make it the most massive runaway star
+velocity of VFTS682 % (\Eqref{eq:speed_around_HST} and
+% \Eqref{eq:speed_around_Gaia}, respectively)
+which would make it the most massive runaway star
 known to date. However, the large errors on both the proper motion measures
@@
 The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
-VFTS682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the
+VFTS682 from \emph{Gaia} DR2, and the
 green arrows illustrate the uncertainty. % show the possible
@@
 
-\bibliographystyle{mnras}
-\bibliography{bibliography/vfts682}
-
 \small
@@
 
+\bibliographystyle{mnras}
+\bibliography{bibliography/vfts682}
+
 % Don't change these lines

### 2018-06-06 VFTS682_APJL.tex
moved back to apjl template for all authors to be in the paper

--- 2018-06-06 VFTS682.tex
@@
-\documentclass[a4paper,fleqn,usenatbib]{mnras}
-\usepackage{newtxtext,newtxmath}
-\usepackage[T1]{fontenc}
-\usepackage{ae,aecompl}
-
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
-
-% Only include extra packages if you really need them. Common packages are:
-\usepackage{graphicx}	% Including figure files
-\usepackage{amsmath}	% Advanced maths commands
-\usepackage{amssymb}	% Extra maths symbols
+\documentclass[apjl,twocolumn]{emulateapj}
+\usepackage{graphicx}
+\usepackage[varg]{txfonts}
+%\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
 \usepackage[%draft, 
-% pagebackref,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue
-]{hyperref}
+pagebackref,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
+\renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
+\usepackage{aas_macros}
+%\usepackage{amsmath} % clashes with emulateapjj for some reaons
 \usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{comment}
+\usepackage{multirow}
+%\usepackage{chngpage}
+%\usepackage{lscape}
 \usepackage{url}
-\usepackage{multirow}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
 \newcommand{\todo}[1]{{\large $\blacksquare$~\textbf{\color{red}[#1]}}~$\blacksquare$}
 \newcommand{\udef}{\stackrel{\mathrm{def}}{=}}
+% for tree diagram
+%\usepackage{forest}
+%\usepackage{tikz-qtree}
+%\usetikzlibrary{shadows,trees}
 
@@
 
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
+\begin{document}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
-
-% Title of the paper, and the short title which is used in the headers.
-% Keep the title short and informative.
-\title[\emph{Gaia} DR2 adds clues for dynamical ejections of a $\sim$$150\,M_\odot$ star]{\emph{Gaia} DR2 adds clues for dynamical ejection of a $\sim$$150\,M_\odot$ star}
-
-% The list of authors, and the short list which is used in the headers.
-% If you need two or more lines of authors, add an extra line using \newauthor
-\author[Renzo et al.]{
-M.~Renzo$^{1}$\thanks{Corresponding author: M.~Renzo
-  \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}},
-S.~E.~de~Mink$^{1}$,
-D.~J.~Lennon$^{2}$,
-R.~P.~van~der~Marel$^{3,4}$,
-I.~Platais$^{4}$,
-J. Bestenlehner$^{5}$,
-V.~H\'enault-Brunet$^{6}$,
-S.~Justham$^{7,8}$,
-A.~de~Koter$^{1}$,
-N.~Langer$^{9}$,
-H.~Sana$^{10}$,
-F.~R.~Schneider$^{11}$,
-J.~S.~Vink$^{12}$ \\
-% List of institutions
-$^{1}${Astronomical Institute Anton Pannekoek, University of
+% \title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
+%                                 % ejected %% to fit in one line
+% runaway star known to date}
+\title{\emph{Gaia} DR2 adds clues for dynamical ejection of
+  a $\sim$$150\,M_\odot$ star}
+\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
+  R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
+  V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
+  N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
+\affil{$^{1}${Astronomical Institute Anton Pannekoek, University of
     Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
@@
     Oxford OX1 3RH, UK} \\
-$^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}
+  $^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
 }
 
-% These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ}
-
-% Enter the current year, for the copyright statements etc.
-\pubyear{2018}
-
-% Don't change these lines
-\begin{document}
-\label{firstpage}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\maketitle
-
-% Abstract of the paper
+ \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+\date{}
 \begin{abstract}
-
   The formation and fates of the most massive stars remain among the most elusive questions in stellar astrophysics. 
@@
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
-\begin{keywords}
-  stars: kinematics, stars: runaways, stars: individual: VFTS682
-\end{keywords}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS 682}
+\maketitle{}
 
@@
 Further worth noticing is the variability of the
-star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
-lower than found by \citet{evans:11}.
+star. % \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
+% lower than found by \citet{evans:11}.
 \citet{bestenlehner:11} discuss the Optical Gravitational Lensing
@@
 
-\begin{table}[t]
+\begin{table}
   \begin{center}
@@
 VFTS682 from \emph{Gaia} DR2, and the
-green arrows illustrate the uncertainty. % show the possible
+yellow arrows illustrate the uncertainty. % show the possible
 % range of directions within the uncertainties in the measured relative proper
@@
 are prolonged in the direction opposite to the motion to illustrate
-the possible range of origins. % The green arrow shows the direction of
+the possible range of origins. % The yellow arrow shows the direction of
 % the relative proper motion from HST (see \Eqref{eq:pm_around_HST}),
@@
     relative to the region from \emph{Gaia} DR2, starting from the present day position of
-    the star. The green arrows indicate the possible
+    the star. The yellow arrows indicate the possible
     directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
@@
     star. The length of the prolongations is proportional to the relative proper motion
-    times the kinetic age of VFTS682. % \todo{load color picture on background? Fig with HST data
+    times the kinetic age of VFTS682.
+    % \todo{load color picture on background? Fig with HST data
     % available, but confusing...}
@@
 
+\bibliographystyle{aa}
+\bibliography{bibliography/vfts682}
+
+\begin{acknowledgements}
   \small
@@
   This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
+\end{acknowledgements}
 
-\bibliographystyle{mnras}
-\bibliography{bibliography/vfts682}
-
-% Don't change these lines
-\bsp	% typesetting comment
-\label{lastpage}
 \end{document}
 
-% End of mnras_template.tex
 %%% Local Variables:

### 2018-06-06 mnras_files/VFTS682.tex
moved back to apjl template for all authors to be in the paper

--- 2018-06-06 VFTS682.tex
@@
 
-\newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}
+\newcommand{\newtext}[1]{{\color{ForestYellow}\bf{#1}}}
 
@@
   \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}},
-S.~E.~de~Mink$^{1}$,
-D.~J.~Lennon$^{2}$,
-R.~P.~van~der~Marel$^{3,4}$,
-I.~Platais$^{4}$,
-J. Bestenlehner$^{5}$,
-V.~H\'enault-Brunet$^{6}$,
-S.~Justham$^{7,8}$,
-A.~de~Koter$^{1}$,
-N.~Langer$^{9}$,
-H.~Sana$^{10}$,
-F.~R.~Schneider$^{11}$,
-J.~S.~Vink$^{12}$ \\
+S.~E.~de~Mink$^{1}$, \linebreak
+D.~J.~Lennon$^{2}$,\linebreak
+R.~P.~van~der~Marel$^{3,4}$, \linebreak
+I.~Platais$^{4}$, \linebreak
+J. Bestenlehner$^{5}$, \linebreak
+V.~H\'enault-Brunet$^{6}$, \linebreak
+S.~Justham$^{7,8}$, \linebreak
+A.~de~Koter$^{1}$, \linebreak
+N.~Langer$^{9}$, \linebreak
+H.~Sana$^{10}$, \linebreak
+F.~R.~Schneider$^{11}$, \linebreak
+J.~S.~Vink$^{12}$ \linebreak
 % List of institutions
@@
 \begin{document}
-\label{firstpage}
+\label{firsxtpage}
 \pagerange{\pageref{firstpage}--\pageref{lastpage}}
@@
 \begin{abstract}
-
   The formation and fates of the most massive stars remain among the most elusive questions in stellar astrophysics. 

### 2018-06-06 VFTS682_APJL.tex
getting yellow warning from ApJL word count: excess is ~500 words

@@
 runaway known to date, with a peculiar three-dimensional speed of
-$44\pm21\,\kms$. This result will need to be revisited as updated
-astrometric parameters from future \emph{Gaia} data releases become
-available. If this result will be confirmed, it means that isolated star formation is
-\emph{not} required to explain this star. Its proper motion suggests that it was ejected from the cluster R136
+$44\pm21\,\kms$. Due to the large error bars, this result will need
+to be revisited with future astrometric data. % to be revisited as updated
+% astrometric parameters from future \emph{Gaia} data releases become
+% available.
+If confirmed, it means that isolated star formation is
+\emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
 $0.9\pm0.6$\,Myr ago. Because of the exceptionally large mass
@@
 This is consistent with the detection
-of extremely massive stars by \cite{crowther:10} in the core of the
+of extremely massive stars in the core of the
 cluster. % The projected rotational equatorial
@@
 \citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
-three most massive stars detected in the core of the cluster by
-\cite{dekoter:97,crowther:10,crowther:16}, with an astonishing similarity in particular with
+three most massive stars detected in the core of the cluster%  by
+% \cite{dekoter:97,crowther:10,crowther:16}
+, with an astonishing similarity in particular with
 the spectrum of R136a3. Therefore, the isolation of
@@
 model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS682 and the massive members of R136. While it is not possible to
-robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
-like fast rotation, which however could have already slowed down
-because of the wind angular momentum losses during and after the
-merger.
+VFTS682 and the massive members of R136.%  While it is not possible to
+% robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
+% like fast rotation, which however could have already slowed down
+% because of the wind angular momentum losses during and after the
+% merger.
 
@@
 
-\bibliographystyle{aa}
-\bibliography{bibliography/vfts682}
+% \bibliographystyle{aa}
+% \bibliography{bibliography/vfts682}
+\begin{thebibliography}{}
+\makeatletter
+\relax
+\def\mn@urlcharsother{\let\do\@makeother \do\$\do\&\do\#\do\^\do\_\do\%\do\~}
+\def\mn@doi{\begingroup\mn@urlcharsother \@ifnextchar [ {\mn@doi@}
+  {\mn@doi@[]}}
+\def\mn@doi@[#1]#2{\def\@tempa{#1}\ifx\@tempa\@empty \href
+  {http://dx.doi.org/#2} {doi:#2}\else \href {http://dx.doi.org/#2} {#1}\fi
+  \endgroup}
+\def\mn@eprint#1#2{\mn@eprint@#1:#2::\@nil}
+\def\mn@eprint@arXiv#1{\href {http://arxiv.org/abs/#1} {{\tt arXiv:#1}}}
+\def\mn@eprint@dblp#1{\href {http://dblp.uni-trier.de/rec/bibtex/#1.xml}
+  {dblp:#1}}
+\def\mn@eprint@#1:#2:#3:#4\@nil{\def\@tempa {#1}\def\@tempb {#2}\def\@tempc
+  {#3}\ifx \@tempc \@empty \let \@tempc \@tempb \let \@tempb \@tempa \fi \ifx
+  \@tempb \@empty \def\@tempb {arXiv}\fi \@ifundefined
+  {mn@eprint@\@tempb}{\@tempb:\@tempc}{\expandafter \expandafter \csname
+  mn@eprint@\@tempb\endcsname \expandafter{\@tempc}}}
 
-\begin{acknowledgements}
-  \small
-  We are grateful to S.~Torres, M.~C.~Ramirez-Tannus,
-  and C.~J.~Evans for help and discussions.
+\bibitem[\protect\citeauthoryear{{Banerjee}, {Kroupa}  \& {Oh}}{{Banerjee}
+  et~al.}{2012}]{banerjee:12}
+{Banerjee} S.,  {Kroupa} P.,   {Oh} S.,  2012, \mn@doi [\apj]
+  {10.1088/0004-637X/746/1/15}, \href
+  {http://adsabs.harvard.edu/abs/2012ApJ...746...15B} {746, 15}
+
+\bibitem[\protect\citeauthoryear{{Bate}}{{Bate}}{2009}]{bate:09}
+{Bate} M.~R.,  2009, \mn@doi [\mnras] {10.1111/j.1365-2966.2008.14106.x}, \href
+  {http://adsabs.harvard.edu/abs/2009MNRAS.392..590B} {392, 590}
+
+\bibitem[\protect\citeauthoryear{{Bestenlehner} et~al.,}{{Bestenlehner}
+  et~al.}{2011}]{bestenlehner:11}
+{Bestenlehner} J.~M.,  et~al., 2011, \mn@doi [\aap]
+  {10.1051/0004-6361/201117043}, \href
+  {http://adsabs.harvard.edu/abs/2011A\%26A...530L..14B} {530, L14}
+
+\bibitem[\protect\citeauthoryear{{Blaauw}}{{Blaauw}}{1961}]{blaauw:61}
+{Blaauw} A.,  1961, \bain, \href
+  {http://adsabs.harvard.edu/abs/1961BAN....15..265B} {15, 265}
+
+\bibitem[\protect\citeauthoryear{{Bressert} et~al.,}{{Bressert}
+  et~al.}{2012}]{bressert:12}
+{Bressert} E.,  et~al., 2012, \mn@doi [\aap] {10.1051/0004-6361/201117247},
+  \href {http://adsabs.harvard.edu/abs/2012A\%26A...542A..49B} {542, A49}
+
+\bibitem[\protect\citeauthoryear{{Brott} et~al.,}{{Brott}
+  et~al.}{2011}]{brott:11}
+{Brott} I.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201016113}, \href
+  {http://adsabs.harvard.edu/abs/2011A\%26A...530A.115B} {530, A115}
+
+\bibitem[\protect\citeauthoryear{{Crowther}, {Schnurr}, {Hirschi}, {Yusof},
+  {Parker}, {Goodwin}  \& {Kassim}}{{Crowther} et~al.}{2010}]{crowther:10}
+{Crowther} P.~A.,  {Schnurr} O.,  {Hirschi} R.,  {Yusof} N.,  {Parker} R.~J.,
+  {Goodwin} S.~P.,   {Kassim} H.~A.,  2010, \mn@doi [\mnras]
+  {10.1111/j.1365-2966.2010.17167.x}, \href
+  {http://adsabs.harvard.edu/abs/2010MNRAS.408..731C} {408, 731}
+
+\bibitem[\protect\citeauthoryear{{Crowther} et~al.,}{{Crowther}
+  et~al.}{2016}]{crowther:16}
+{Crowther} P.~A.,  et~al., 2016, \mn@doi [\mnras] {10.1093/mnras/stw273}, \href
+  {http://adsabs.harvard.edu/abs/2016MNRAS.458..624C} {458, 624}
+
+\bibitem[\protect\citeauthoryear{{Davidson}, {Humphreys}  \& {Weis}}{{Davidson}
+  et~al.}{2016}]{davidson:16}
+{Davidson} K.,  {Humphreys} R.~M.,   {Weis} K.,  2016, arXiv:1608.02007, \href
+  {http://adsabs.harvard.edu/abs/2016arXiv160802007D} {}
+
+\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
+  et~al.}{2010}]{evans:10}
+{Evans} C.~J.,  et~al., 2010, \mn@doi [\apjl] {10.1088/2041-8205/715/2/L74},
+  \href {http://adsabs.harvard.edu/abs/2010ApJ...715L..74E} {715, L74}
+
+\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
+  et~al.}{2011}]{evans:11}
+{Evans} C.~J.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201116782},
+  \href {http://adsabs.harvard.edu/abs/2011A\%26A...530A.108E} {530, A108}
+
+\bibitem[\protect\citeauthoryear{{Fujii} \& {Portegies Zwart}}{{Fujii} \&
+  {Portegies Zwart}}{2011}]{fujii:11}
+{Fujii} M.~S.,  {Portegies Zwart} S.,  2011, \mn@doi [Science]
+  {10.1126/science.1211927}, \href
+  {http://adsabs.harvard.edu/abs/2011Sci...334.1380F} {334, 1380}
+
+\bibitem[\protect\citeauthoryear{{Gaia Collaboration} et~al.,}{{Gaia
+  Collaboration} et~al.}{2016}]{gaia:16}
+{Gaia Collaboration} et~al., 2016, \mn@doi [\aap]
+  {10.1051/0004-6361/201629272}, \href
+  {http://adsabs.harvard.edu/abs/2016A\%26A...595A...1G} {595, A1}
+
+\bibitem[\protect\citeauthoryear{{Gaia Collaboration}, {Brown}, {Vallenari},
+  {Prusti}, {de Bruijne}, {Babusiaux}  \& {Bailer-Jones}}{{Gaia Collaboration}
+  et~al.}{2018}]{brown:18}
+{Gaia Collaboration} {Brown} A.~G.~A.,  {Vallenari} A.,  {Prusti} T.,  {de
+  Bruijne} J.~H.~J.,  {Babusiaux} C.,   {Bailer-Jones} C.~A.~L.,  2018,
+  ArXiv:1804.09365, \href {http://adsabs.harvard.edu/abs/2018arXiv180409365G}
+  {}
+
+\bibitem[\protect\citeauthoryear{{Gruendl} \& {Chu}}{{Gruendl} \&
+  {Chu}}{2009}]{gruendl:09}
+{Gruendl} R.~A.,  {Chu} Y.-H.,  2009, \mn@doi [\apjs]
+  {10.1088/0067-0049/184/1/172}, \href
+  {http://adsabs.harvard.edu/abs/2009ApJS..184..172G} {184, 172}
+
+\bibitem[\protect\citeauthoryear{{Gvaramadze}, {Weidner}, {Kroupa}  \&
+  {Pflamm-Altenburg}}{{Gvaramadze} et~al.}{2012}]{gvaramadze:12}
+{Gvaramadze} V.~V.,  {Weidner} C.,  {Kroupa} P.,   {Pflamm-Altenburg} J.,
+  2012, \mn@doi [\mnras] {10.1111/j.1365-2966.2012.21452.x}, \href
+  {http://adsabs.harvard.edu/abs/2012MNRAS.424.3037G} {424, 3037}
+
+\bibitem[\protect\citeauthoryear{{Hogg}}{{Hogg}}{2018}]{hogg:18}
+{Hogg} D.~W.,  2018, ArXiv:1804.07766, \href
+  {http://adsabs.harvard.edu/abs/2018arXiv180407766H} {}
+
+\bibitem[\protect\citeauthoryear{{Humphreys}, {Weis}, {Davidson}  \&
+  {Gordon}}{{Humphreys} et~al.}{2016}]{humphreys:16}
+{Humphreys} R.~M.,  {Weis} K.,  {Davidson} K.,   {Gordon} M.~S.,  2016, \mn@doi
+  [\apj] {10.3847/0004-637X/825/1/64}, \href
+  {http://adsabs.harvard.edu/abs/2016ApJ...825...64H} {825, 64}
+
+\bibitem[\protect\citeauthoryear{{K{\"o}hler} et~al.,}{{K{\"o}hler}
+  et~al.}{2015}]{kohler:15}
+{K{\"o}hler} K.,  et~al., 2015, \mn@doi [\aap] {10.1051/0004-6361/201424356},
+  \href {http://adsabs.harvard.edu/abs/2015A\%26A...573A..71K} {573, A71}
+
+\bibitem[\protect\citeauthoryear{{Kuiper}, {Yorke}  \& {Turner}}{{Kuiper}
+  et~al.}{2015}]{kuiper:15}
+{Kuiper} R.,  {Yorke} H.~W.,   {Turner} N.~J.,  2015, \mn@doi [\apj]
+  {10.1088/0004-637X/800/2/86}, \href
+  {http://adsabs.harvard.edu/abs/2015ApJ...800...86K} {800, 86}
+
+\bibitem[\protect\citeauthoryear{{Lada} \& {Lada}}{{Lada} \&
+  {Lada}}{2003}]{lada:03}
+{Lada} C.~J.,  {Lada} E.~A.,  2003, \mn@doi [\araa]
+  {10.1146/annurev.astro.41.011802.094844}, \href
+  {http://adsabs.harvard.edu/abs/2003ARA\%26A..41...57L} {41, 57}
+
+\bibitem[\protect\citeauthoryear{{Lamb}, {Oey}, {Segura-Cox}, {Graus},
+  {Kiminki}, {Golden-Marx}  \& {Parker}}{{Lamb} et~al.}{2016}]{lamb:16}
+{Lamb} J.~B.,  {Oey} M.~S.,  {Segura-Cox} D.~M.,  {Graus} A.~S.,  {Kiminki}
+  D.~C.,  {Golden-Marx} J.~B.,   {Parker} J.~W.,  2016, \mn@doi [\apj]
+  {10.3847/0004-637X/817/2/113}, \href
+  {http://adsabs.harvard.edu/abs/2016ApJ...817..113L} {817, 113}
+
+\bibitem[\protect\citeauthoryear{{Langer}, {Norman}, {de Koter}, {Vink},
+  {Cantiello}  \& {Yoon}}{{Langer} et~al.}{2007}]{langer:07}
+{Langer} N.,  {Norman} C.~A.,  {de Koter} A.,  {Vink} J.~S.,  {Cantiello} M.,
+  {Yoon} S.-C.,  2007, \mn@doi [\aap] {10.1051/0004-6361:20078482}, \href
+  {http://adsabs.harvard.edu/abs/2007A\%26A...475L..19L} {475, L19}
+
+\bibitem[\protect\citeauthoryear{{Lennon} et~al.,}{{Lennon}
+  et~al.}{2018}]{lennon:18}
+{Lennon} D.~J.,  et~al., 2018, ArXiv:1805.08277, \href
+  {http://adsabs.harvard.edu/abs/2018arXiv180508277L} {}
+
+\bibitem[\protect\citeauthoryear{{Lindegren} et~al.,}{{Lindegren}
+  et~al.}{2018}]{lindengren:18}
+{Lindegren} L.,  et~al., 2018, ArXiv:1804.09366, \href
+  {http://adsabs.harvard.edu/abs/2018arXiv180409366L} {}
+
+\bibitem[\protect\citeauthoryear{{Maeder} \& {Meynet}}{{Maeder} \&
+  {Meynet}}{2000}]{maeder:00}
+{Maeder} A.,  {Meynet} G.,  2000, \aap, \href
+  {http://adsabs.harvard.edu/abs/2000A\%26A...361..159M} {361, 159}
+
+\bibitem[\protect\citeauthoryear{{Oh} \& {Kroupa}}{{Oh} \&
+  {Kroupa}}{2016}]{oh:16}
+{Oh} S.,  {Kroupa} P.,  2016, \mn@doi [\aap] {10.1051/0004-6361/201628233},
+  \href {http://adsabs.harvard.edu/abs/2016A\%26A...590A.107O} {590, A107}
+
+% \bibitem[\protect\citeauthoryear{{Parker}}{{Parker}}{1993}]{parker:93}
+% {Parker} J.~W.,  1993, \mn@doi [\aj] {10.1086/116661}, \href
+%   {http://adsabs.harvard.edu/abs/1993AJ....106..560P} {106, 560}
+
+\bibitem[\protect\citeauthoryear{{Pietrzy{\'n}ski} et~al.,}{{Pietrzy{\'n}ski}
+  et~al.}{2013}]{pietrzynski:13}
+{Pietrzy{\'n}ski} G.,  et~al., 2013, \mn@doi [\nat] {10.1038/nature11878},
+  \href {http://adsabs.harvard.edu/abs/2013Natur.495...76P} {495, 76}
+
+\bibitem[\protect\citeauthoryear{{Platais}, {van der Marel}, {Lennon},
+  {Anderson}, {Bellini}, {Sabbi}, {Sana}  \& {Bedin}}{{Platais}
+  et~al.}{2015}]{platais:15}
+{Platais} I.,  {van der Marel} R.~P.,  {Lennon} D.~J.,  {Anderson} J.,
+  {Bellini} A.,  {Sabbi} E.,  {Sana} H.,   {Bedin} L.~R.,  2015, \mn@doi [\aj]
+  {10.1088/0004-6256/150/3/89}, \href
+  {http://adsabs.harvard.edu/abs/2015AJ....150...89P} {150, 89}
+
+\bibitem[\protect\citeauthoryear{{Platais} et~al.,}{{Platais}
+  et~al.}{2018}]{platais:18}
+{Platais} I.,  et~al., 2018, ArXiv:1804.08678, \href
+  {http://adsabs.harvard.edu/abs/2018arXiv180408678P} {}
+
+\bibitem[\protect\citeauthoryear{{Poveda}, {Ruiz}  \& {Allen}}{{Poveda}
+  et~al.}{1967}]{poveda:67}
+{Poveda} A.,  {Ruiz} J.,   {Allen} C.,  1967, Boletin de los Observatorios
+  Tonantzintla y Tacubaya, \href
+  {http://adsabs.harvard.edu/abs/1967BOTT....4...86P} {4, 86}
+
+\bibitem[\protect\citeauthoryear{{Rakavy} \& {Shaviv}}{{Rakavy} \&
+  {Shaviv}}{1967}]{rakavy:67}
+{Rakavy} G.,  {Shaviv} G.,  1967, \mn@doi [\apj] {10.1086/149204}, \href
+  {http://adsabs.harvard.edu/abs/1967ApJ...148..803R} {148, 803}
+
+\bibitem[\protect\citeauthoryear{{Ram{\'{\i}}rez-Agudelo}
+  et~al.,}{{Ram{\'{\i}}rez-Agudelo} et~al.}{2015}]{ramirez-agudelo:15}
+{Ram{\'{\i}}rez-Agudelo} O.~H.,  et~al., 2015, \mn@doi [\aap]
+  {10.1051/0004-6361/201425424}, \href
+  {http://adsabs.harvard.edu/abs/2015A\%26A...580A..92R} {580, A92}
+
+\bibitem[\protect\citeauthoryear{{Renzo} et~al.,}{{Renzo}
+  et~al.}{2018}]{renzo:18}
+{Renzo} M.,  et~al., 2018, ArXiv:1804.09164, \href
+  {http://adsabs.harvard.edu/abs/2018arXiv180409164R} {}
+
+\bibitem[\protect\citeauthoryear{{Robitaille} \& {Bressert}}{{Robitaille} \&
+  {Bressert}}{2012}]{robitaille:12}
+{Robitaille} T.,  {Bressert} E.,  2012, {APLpy: Astronomical Plotting Library
+  in Python}, Astrophysics Source Code Library (\mn@eprint {ascl} {1208.017})
+
+\bibitem[\protect\citeauthoryear{{Rosen}, {Krumholz}, {McKee}  \&
+  {Klein}}{{Rosen} et~al.}{2016}]{rosen:16}
+{Rosen} A.~L.,  {Krumholz} M.~R.,  {McKee} C.~F.,   {Klein} R.~I.,  2016,
+  \mn@doi [\mnras] {10.1093/mnras/stw2153}, \href
+  {http://adsabs.harvard.edu/abs/2016MNRAS.463.2553R} {463, 2553}
+
+\bibitem[\protect\citeauthoryear{{Rubio-D{\'{\i}}ez}, {Najarro},
+  {Garc{\'{\i}}a}  \& {Sundqvist}}{{Rubio-D{\'{\i}}ez}
+  et~al.}{2017}]{rubio-diez:17}
+{Rubio-D{\'{\i}}ez} M.~M.,  {Najarro} F.,  {Garc{\'{\i}}a} M.,   {Sundqvist}
+  J.~O.,  2017, in {Eldridge} J.~J.,  {Bray} J.~C.,  {McClelland} L.~A.~S.,
+  {Xiao} L.,  eds,  IAU Symposium Vol. 329, The Lives and Death-Throes of
+  Massive Stars. pp 131--135, \mn@doi{10.1017/S1743921317002447}
+
+\bibitem[\protect\citeauthoryear{{Sabbi} et~al.,}{{Sabbi}
+  et~al.}{2012}]{sabbi:12}
+{Sabbi} E.,  et~al., 2012, \mn@doi [\apjl] {10.1088/2041-8205/754/2/L37}, \href
+  {http://adsabs.harvard.edu/abs/2012ApJ...754L..37S} {754, L37}
+
+\bibitem[\protect\citeauthoryear{{Sana}, {Ram{\'{\i}}rez-Tannus}, {de Koter},
+  {Kaper}, {Tramper}  \& {Bik}}{{Sana} et~al.}{2017}]{sana:17}
+{Sana} H.,  {Ram{\'{\i}}rez-Tannus} M.~C.,  {de Koter} A.,  {Kaper} L.,
+  {Tramper} F.,   {Bik} A.,  2017, \mn@doi [\aap]
+  {10.1051/0004-6361/201630087}, \href
+  {http://adsabs.harvard.edu/abs/2017A\%26A...599L...9S} {599, L9}
+
+\bibitem[\protect\citeauthoryear{{Schneider}, {Castro}, {Fossati}, {Langer}  \&
+  {de Koter}}{{Schneider} et~al.}{2017}]{schneider:17}
+{Schneider} F.~R.~N.,  {Castro} N.,  {Fossati} L.,  {Langer} N.,   {de Koter}
+  A.,  2017, \mn@doi [\aap] {10.1051/0004-6361/201628409}, \href
+  {http://adsabs.harvard.edu/abs/2017A\%26A...598A..60S} {598, A60}
+
+\bibitem[\protect\citeauthoryear{{Schneider} et~al.,}{{Schneider}
+  et~al.}{2018}]{schneider:18}
+{Schneider} F.~R.~N.,  et~al., 2018, \mn@doi [Science]
+  {10.1126/science.aan0106}, \href
+  {http://adsabs.harvard.edu/abs/2018Sci...359...69S} {359, 69}
+
+\bibitem[\protect\citeauthoryear{{Smith}}{{Smith}}{2016}]{smith:16}
+{Smith} N.,  2016, \mn@doi [\mnras] {10.1093/mnras/stw1533}, \href
+  {http://adsabs.harvard.edu/abs/2016MNRAS.461.3353S} {461, 3353}
+
+\bibitem[\protect\citeauthoryear{{Smith} \& {Tombleson}}{{Smith} \&
+  {Tombleson}}{2015}]{smith:15}
+{Smith} N.,  {Tombleson} R.,  2015, \mn@doi [\mnras] {10.1093/mnras/stu2430},
+  \href {http://adsabs.harvard.edu/abs/2015MNRAS.447..598S} {447, 598}
+
+\bibitem[\protect\citeauthoryear{{Udalski} et~al.,}{{Udalski}
+  et~al.}{2008}]{udalski:08}
+{Udalski} A.,  et~al., 2008, \actaa, \href
+  {http://adsabs.harvard.edu/abs/2008AcA....58..329U} {58, 329}
+
+\bibitem[\protect\citeauthoryear{{Ward} \& {Kruijssen}}{{Ward} \&
+  {Kruijssen}}{2018}]{ward:18}
+{Ward} J.~L.,  {Kruijssen} J.~M.~D.,  2018, \mn@doi [\mnras]
+  {10.1093/mnras/sty117}, \href
+  {http://adsabs.harvard.edu/abs/2018MNRAS.475.5659W} {475, 5659}
+
+\bibitem[\protect\citeauthoryear{{Zapartas} et~al.,}{{Zapartas}
+  et~al.}{2017}]{zapartas:17}
+{Zapartas} E.,  et~al., 2017, \mn@doi [\aap] {10.1051/0004-6361/201629685},
+  \href {http://adsabs.harvard.edu/abs/2017A\%26A...601A..29Z} {601, A29}
+
+\bibitem[\protect\citeauthoryear{{Zinnecker} \& {Yorke}}{{Zinnecker} \&
+  {Yorke}}{2007}]{zinnecker:07}
+{Zinnecker} H.,  {Yorke} H.~W.,  2007, \mn@doi [\araa]
+  {10.1146/annurev.astro.44.051905.092549}, \href
+  {http://adsabs.harvard.edu/abs/2007ARA\%26A..45..481Z} {45, 481}
+
+\bibitem[\protect\citeauthoryear{{Zwicky}}{{Zwicky}}{1957}]{zwicky:57}
+{Zwicky} F.,  1957, \zap, \href
+  {http://adsabs.harvard.edu/abs/1957ZA.....44...64Z} {44, 64}
+
+\bibitem[\protect\citeauthoryear{{de Koter}, {Heap}  \& {Hubeny}}{{de Koter}
+  et~al.}{1997}]{dekoter:97}
+{de Koter} A.,  {Heap} S.~R.,   {Hubeny} I.,  1997, \mn@doi [\apj]
+  {10.1086/303736}, \href {http://adsabs.harvard.edu/abs/1997ApJ...477..792D}
+  {477, 792}
+
+\bibitem[\protect\citeauthoryear{{de Mink}, {Cantiello}, {Langer}, {Pols},
+  {Brott}  \& {Yoon}}{{de Mink} et~al.}{2009}]{demink:09}
+{de Mink} S.~E.,  {Cantiello} M.,  {Langer} N.,  {Pols} O.~R.,  {Brott} I.,
+  {Yoon} S.-C.,  2009, \mn@doi [\aap] {10.1051/0004-6361/200811439}, \href
+  {http://adsabs.harvard.edu/abs/2009A\%26A...497..243D} {497, 243}
+
+\makeatother
+\end{thebibliography}
+
+% \begin{acknowledgements}
+%   \small
+%   We are grateful to S.~Torres, M.~C.~Ramirez-Tannus,
+%   and C.~J.~Evans for help and discussions.
 SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC) (Grant agreement No.
@@
 Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \emph{Gaia} Multilateral Agreement.
-The background image of \Figref{fig:main} is based on observations
-  made with ESO Telescopes at the La Silla Observatory under programme
-  ID 076.C-0888, processed and released by the ESO VOS/ADP group.
-  This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
-\end{acknowledgements}
+% The background image of \Figref{fig:main} is based on observations
+%   % made with ESO Telescopes at the La Silla Observatory under programme
+%   % ID 076.C-0888, processed and released by the ESO VOS/ADP group.
+%   % This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
+% \end{acknowledgements}
 

### 2018-06-06 VFTS682.tex
none

--- 2018-06-05 VFTS682.tex
@@
   % 
-  Previous spectroscopic studies identified VFTS682 as one of the most
+  Previous spectroscopic studies identified VFTS 682 as one of the most
   massive stars known. It resides in the field of the 30 Dor region in
@@
   % 
-  We use the proper motions from the second \emph{Gaia} data
+  We use the unprecedented proper motions from the second \emph{Gaia} data
   release (DR2) together with multiepoch \emph{Hubble Space Telescope} proper motions to investigate the kinematics of this object. We show
-  that the projected velocity of VFTS682 is consistent with it being a
+  that the projected velocity of VFTS 682 is consistent with it being a
   runaway star dynamically ejected from R136. Its
@@
   % 
-  If confirmed by future astrometric data, this would make VFTS682
+  If confirmed by future astrometric data, this would make VFTS 682
   the most massive runaway star known to date.
@@
 \begin{keywords}
-  stars: kinematics, stars: runaways, stars: individual: VFTS682
+  stars: kinematics, stars: runaways, stars: individual: VFTS 682
 \end{keywords}
@@
 \citep[e.g.][]{kuiper:15,rosen:16}, but simulations cannot yet resolve
-the length scales  of interest to explain the 
-high multiplicity of massive stars  \citep[e.g.,][]{bate:09, sana:17}. Understanding massive star formation, and its
+the smallest length scales that are of interest to understand the 
+high level of multiplicity among massive stars  \citep[e.g.,][]{bate:09, sana:17}. Understanding massive star formation, and its
 possible dependence on environment and metallicity, is crucial for
@@
 
-A contribution to the debate on whether or not massive
+An interesting contribution to the debate on whether or not massive
 stars form in relative isolation was presented by
 \cite{bestenlehner:11} and \cite{bressert:12}, who discussed the case of the very massive star
-VFTS682.  This star is located in the field of the 30 Doradus region
+VFTS 682.  This star is located in the field of the 30 Doradus region
 in the Large Magellanic Cloud (LMC) and was studied as part of the
-multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}. Its spectral type is WNh5 and its present-day mass is $137.8^{+27.5}_{-15.9}\,M_\odot$,
+multi-epoch spectroscopic VLT-FLAMES Tarantula Survey (VFTS) of
+massive stars \citep{evans:11}. Its spectral type is WNh5 and its 
+present-day mass is $137.8^{+27.5}_{-15.9}\,M_\odot$,
 \citep{schneider:18}. These correspond to an inferred initial mass of
-$150.0^{+28.7}_{-17.4}\,M_\odot$ and make VFTS682 one of the
+$150.0^{+28.7}_{-17.4}\,M_\odot$. This makes VFTS 682 one of the
 most extreme objects in the region.
-%
+
 From the spectral point of view, it is reminiscent of the very
@@
 R136 cluster. In particular, a remarkable similarity exist between the
-spectrum of VFTS682 and R136a3, located in the core of the cluster
+spectrum of VFTS 682 and R136a3, located in the core of the cluster
 \citep{crowther:10} for which \citet{crowther:16} report a
@@
 at least two more very massive WN5h stars, R136a1 and R136a2, whose
-estimated current masses are even higher. However, VFTS682 stands
+estimated current masses are even higher. However, VFTS 682 stands
 out by its isolation at a projected offset of 119.4 arcseconds, corresponding to $\sim$$29$\,pc from the core of the star cluster
@@
 \citet{bestenlehner:11} discuss two possibilities for the
-origin of VFTS682. % Quoting directly 
+origin of VFTS 682. % Quoting directly 
 From their abstract: "(i) the
 star either formed in situ, which would have profound implications for
-the formation mechanism of massive stars, or (ii) VFTS682 is a slow
+the formation mechanism of massive stars, or (ii) VFTS 682 is a slow
 runaway star that originated from the dense cluster R136, which would
@@
 dynamical models, \citet{fujii:11} and \citet{banerjee:12} argued that
-ejection of VFTS682 from R136 would be natural. Testing the origin of
-VFTS682 is not only important for constraining isolated star
+ejection of VFTS 682 from R136 would be natural. Testing the origin of
+VFTS 682 is not only important for constraining isolated star
 formation; it should also help to improve understanding of early star-cluster dynamics.
@@
 runaway stars in
- 30 Doradus based on their proper motions. \citet{lennon:18} used data from the second \emph{Gaia}
+ 30 Doradus based on proper motion
+measurements. \citet{lennon:18} used data from the second \emph{Gaia}
 data release \cite[DR2,][]{gaia:16,brown:18} to further investigate
-the isolated O-type stars around R136. In this study, we
-report on our analysis of the \emph{Gaia} DR2 data for VFTS682 which
+the isolated O-type stars in the vicinity of R136. In this study, we
+report on our analysis of the \emph{Gaia} DR2 data for VFTS 682 which
 was not part of the sample of \citet{lennon:18} because of its 
@@
 VFTS survey \citep[][]{evans:11} with the proper motion from
-\emph{Gaia} DR2 to reconstruct the three-dimensional velocity of VFTS682, and test the hypothesis that this star was ejected from R136. We
+\emph{Gaia} DR2 to reconstruct the three-dimensional velocity of VFTS
+682, and test the hypothesis that this star was ejected from R136. We
 use the HST proper motions from \citet{platais:18} to test our
@@
 %%% cut for length reasons: all is said below
-% Our results indicate that VFTS682 is a runaway star, although with
+% Our results indicate that VFTS 682 is a runaway star, although with
 % low statistical significance. The
@@
 
-\subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
+\subsection{ Overview of VFTS 682 from previous studies \label{data:vfts683}}
 
-The star VFTS682, located at right ascension (RA)
+The star VFTS 682, located at right ascension (RA)
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000 \citep[][% see also
-% \Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2
-]{evans:11}
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000 \citep[][see also
+\Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2]{evans:11}
 was originally classified as a young stellar object \citep{gruendl:09}
-based on its mid-infrared excess. \citet{evans:11} reclassified the
-object as Wolf-Rayet star of spectral type WNh5 using multi-epoch
-spectra covering $\lambda$4000--7000 from the
-VFTS survey. % The spectra available covered $\lambda$4000--7000 and
-% were taken at multiple epochs.
-
-Using the same dataset,
+based on its mid-infrared excess. \citet{evans:11} classified the
+object as Wolf-Rayet star of spectral type WNh5 using spectra from the
+VFTS survey. The spectra available covered $\lambda$4000--7000 and
+were taken at multiple epochs. Using the same dataset,
 \citet{bestenlehner:11} excluded the presence of a close companion
 with high confidence from the absence of radial velocity variations.
-\citet{bestenlehner:11} % further analyzed the spectra % using the non-LTE
-% % model atmosphere code CMFGEN \citep{hillier:98}
-also derived stellar
+
+\citet{bestenlehner:11} further analyzed the spectra % using the non-LTE
+% model atmosphere code CMFGEN \citep{hillier:98}
+to derive the stellar
 parameters and surface abundances. They inferred an extinction
@@
 \citep{schneider:17}, based on the evolutionary models from
-\citet{brott:11} and \cite{kohler:15}. % These values place VFTS682 around the
+\citet{brott:11} and \cite{kohler:15}. % These values place VFTS 682 around the
 % ``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
 
-\citet{bestenlehner:11} reported for VFTS682 a wind mass loss rate of
-$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, % inferred from
-% spectral analysis assuming a smooth outflow, i.e.,
-not accounting for
+\citet{bestenlehner:11} reported for VFTS 682 a wind mass loss rate of
+$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, inferred from
+spectral analysis assuming a smooth outflow, i.e., not accounting for
 clumping. The present-day surface helium mass
@@
 % at least two more very massive WN5h stars, R136a1 and R136a2, whose
-% estimated current masses are even higher. However, VFTS682 stands
+% estimated current masses are even higher. However, VFTS 682 stands
 % out by its isolation at a projected offset of 29\,pc from the center
@@
 Further worth noticing is the variability of the
-star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
-lower than found by \citet{evans:11}.
+star. % \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
+% lower than what found by \citet{evans:11}.
 \citet{bestenlehner:11} discuss the Optical Gravitational Lensing
@@
 
-We obtain the peculiar radial velocity $\delta v_\mathrm{rad}$ of VFTS682 as the difference
+We obtain the peculiar radial velocity $\delta v_\mathrm{rad}$ of VFTS 682 as the difference
 between the average radial velocity of the 30 Doradus region
 ($270\pm10\kms$) minus the radial velocity measured from the NV $\lambda4944$
-line for VFTS682  \citep[$300\pm10\kms$, ][]{bestenlehner:11}. We implicitly assume that the use
+line for VFTS 682  \citep[$300\pm10\kms$, ][]{bestenlehner:11}. We implicitly assume that the use
 of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not
-introduce significant errors. \cite{bressert:12} also noted that VFTS682 shows a significant radial velocity offset compared to the nebular
+introduce significant errors. \cite{bressert:12} also noted that VFTS
+682 shows a significant radial velocity offset compared to the nebular
 lines from the gas surrounding it, suggesting peculiar motion in the
@@
 presence of an optically thick, likely non-homogeneous, and variable
-wind obscuring the stellar surface. The radial velocity errors quoted
-above quantify only statistical uncertainties from the
+wind obscuring the stellar surface. The errors quoted above on the
+radial velocities include only statistical uncertainties from the
 spectral analysis, and not the systematic errors arising from the
@@
 Recently, \citet{platais:18} presented HST  astrometry and proper
-motions in the 30 Doradus region based on two epochs. %  of
-% observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
-% in October 3-8 2011 and the second epoch (GO-13359) in
-% October 6-11 2014, both using the F775W filter.
-% The second epoch was designed to match the pointing and
-% orientation of the first epoch as exactly as possible so
-% that the displacement measurements could be purely differential.
-While the most bright stars (V$<$14) were saturated in
+motions in the 30 Doradus region based on two epochs of
+observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
+in October 3-8 2011 and the second epoch (GO-13359) in
+October 6-11 2014, both using the F775W filter.
+The second epoch was designed to match the pointing and
+orientation of the first epoch as exactly as possible so
+that the displacement measurements could be purely differential.
+While most bright massive stars (V$<$14) were saturated in
 their data and missing from that catalog, due to its
-high extinction, VFTS682 is faint enough
+high extinction, this star is faint enough
 to be included in that catalog. Its proper motions are
@@
 The catalog from \citet{platais:18} lists the proper motion
-components of VFTS682 relative to the field, providing a
+components of VFTS 682 relative to the field, providing a
 measure of the motion of this star completely independent from
@@
 v_\mathrm{RA}^{HST}=2\pm31\kms$ and $\delta
-v_\mathrm{RA}^{HST}=47\pm24\kms$. % , corresponding to a two-dimensional
-% projected velocity of $47\pm24\kms$.
-Including the radial velocity
+v_\mathrm{RA}^{HST}=47\pm24\kms$, corresponding to a two-dimensional
+projected velocity of $47\pm24\kms$. Including the radial velocity
 measured by the VFTS survey we obtain a three-dimensional velocity of
@@
   \begin{center}
-    \caption{Astrometric parameters for VFTS682. }
+    \caption{Astrometric parameters for VFTS 682. }
     \begin{tabular}{l|c|c}
@@
       RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
-                        & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
+                        & \multirow{4}{*}{\emph{Gaia} DR2}\\[5pt]
       DEC \hfill [degree] & -69.07 % $\pm$  0.05
                         & \\[5pt]
-      \hline
-      $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \multirow{2}{*}{\emph{Gaia} DR2}\\[5pt]
+      \cline{1-2}
+      $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \\[5pt]
       $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
@@
 
-The second \emph{Gaia} Data Release (DR2)% , which
-became available on 25 April 2018 and 
+The second \emph{Gaia} Data Release (DR2), which became available on 25 April 2018,
 provides the five-parameter astrometric solution (positions,
 parallaxes, and proper motion components) for more than a billion
-sources \citep{brown:18}. VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2
+sources \citep{brown:18}. VFTS 682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2
 catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. \emph{Gaia}
@@
 \texttt{astrometric\_excess\_noise} = 0. These values suggest that the \emph{Gaia}
-data for VFTS682 are reliable. However, the effective temperature
+data for VFTS 682 are reliable. However, the effective temperature
 reported in \emph{Gaia} DR2, based on the spectral fitting around the
@@
 in this study, and we attribute the unphysical value of the parallax
-to the large distance to the LMC.%  Our main findings do not rely on the
-% parallax nor the effective temperature values reported in the \emph{Gaia} DR2
-% catalog.
+to the large distance to the LMC. Our main findings do not rely on the
+parallax nor the effective temperature values reported in the \emph{Gaia} DR2
+catalog.
 
-While \emph{Gaia} DR2 data can be correlated, this is
-not an issue for VFTS682, and we treat the proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
-respectively) of the star as uncorrelated for simplicity.
-% We retrieve for VFTS682 the position in RA and DEC
-% in the International Celestial Reference System \cite[][]{brown:18},
-% its
-% proper motion components .
-%  and for simplicity we treat them as
-% uncorrelated in the following.
-For the radial velocity of VFTS682 and of the 30 Doradus
+We retrieve for VFTS 682 the position in RA and DEC
+in the International Celestial Reference System \cite[][]{brown:18}, its
+proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
+respectively). While \emph{Gaia} DR2 data can be correlated, this is
+not an issue for VFTS682, and for simplicity we treat them as
+uncorrelated in the following.
+
+For the radial velocity of VFTS 682 and of the 30 Doradus
 region as a whole, we instead use the VFTS data
@@
 We follow \citet{lennon:18} to define a local frame of reference and derive the peculiar velocity
-of VFTS682 with respect to the cluster R136. Their selection (see
+of VFTS 682 with respect to the cluster R136. Their selection (see
 their Sec.~2.1) of 153
 bright ($G<17$) stars around R136 with reliable astrometric data from
-\emph{Gaia} DR2 yields $\langle\mu_\mathrm{RA}\rangle=1.74\pm0.01\,\mathrm{mas\
-  yr^{-1}}$ and $\langle\mu_\mathrm{DEC}\rangle=0.70\pm0.02\,\mathrm{mas\ yr^{-1}}$
+\emph{Gaia} DR2 yields $\mu_\mathrm{RA}=1.74\pm0.01\,\mathrm{mas\
+  yr^{-1}}$ and $\mu_\mathrm{DEC}=0.70\pm0.02\,\mathrm{mas\ yr^{-1}}$
 for the components of the mean motion of the region projected on the sky.
@@
 % We define a local standard frame of reference to derive the peculiar velocity
-% of VFTS682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
+% of VFTS 682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
 % We select all the stars in a target of 0.05 degrees around R136
@@
 \label{sec:runaway}
-We first address the question of whether VFTS682 is a typical star
+We first address the question of whether VFTS 682 is a typical star
 from the kinematic point of view, or whether it is a runaway star with
 a significantly large peculiar velocity compared to its surrounding population. The former is what should
-be expected if it formed in the relative isolation that we observe today.% , in relative
-% isolation from other massive stars.
+be expected if it formed where we observe it today, in relative
+isolation from other massive stars.
 
@@
 Subtracting the mean proper motion components given by
-\citet{lennon:18} from the
-proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
+\citep{lennon:18} from the
+proper motion of VFTS 682 (see \Tabref{tab:vfts682}), we obtain the
 components of proper motion of the star relative to the surrounding region
@@
 from R136, to highlight the likelihood of it being the origin of fast
-moving stars. The plus sign marks VFTS682, which sits at the edge of
+moving stars. The plus sign marks VFTS 682, which sits at the edge of
 the proper motion distribution, but is not a clear outlier. Most
-relevant is the direction of the proper motion of VFTS682, which we
+relevant is the direction of the proper motion of VFTS 682, which we
 discuss in \Secref{sec:r136_origin}.
@@
 Therefore, two completely independent measures of the proper motion of
-VFTS682 relative to the surrounding field, one from HST and one from
+VFTS 682 relative to the surrounding field, one from HST and one from
 \emph{Gaia} DR2, yield values of the peculiar three-dimensional
-velocity of VFTS682 (\Eqref{eq:speed_around_HST} and
+velocity of VFTS 682 (\Eqref{eq:speed_around_HST} and
 \Eqref{eq:speed_around_Gaia}, respectively) which would make it the most massive runaway star
@@
 
-Assuming that the best estimate of the peculiar velocity of VFTS682
+Assuming that the best estimate of the peculiar velocity of VFTS 682
 are reliable, we now address the question of its likely origin.
@@
 The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
-VFTS682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the
-orange arrows illustrate the uncertainty. % show the possible
-% range of directions within the uncertainties in the measured relative proper
-% motion.
-These arrows cross at the present-day location of VFTS682 and
+VFTS 682 from \emph{Gaia} DR2 (see \Eqref{eq:pm_gaia_around}), and the cyan arrows show the possible
+range of directions within the uncertainties in the measured relative proper
+motion. These arrows cross at the present-day location of VFTS 682 and
 are prolonged in the direction opposite to the motion to illustrate
@@
 components results in a wide range of possible directions,we argue that the most likely origin of the star is R136,
-as suggested by the N-body simulations of \cite{fujii:11}
+as suggested based on numerical N-body simulations by \cite{fujii:11}
 and \cite{banerjee:12}.
 
-The kinematic age of this star, assuming it originates from the
-cluster, is
+We can therefore consider the kinematic age for this star assuming
+that it originates from the cluster,
 
@@
 \end{equation}
-where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
+where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS 682 to
 the core of R136 \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
@@
 
-The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
-ejection from the cluster, given is apparent age
-of $1.0\pm 0.2$\,Myr \citep{schneider:18}. % , which
-% corroborates the idea that the star is the result of a dynamical
-% ejection very soon in the cluster evolution.
-Note that the
+The kinematic age $\tau_\mathrm{kin}$ is similar or smaller than the apparent age
+of the star of $1.0\pm 0.2$\,Myr from \cite{schneider:18}, which
+corroborates the idea that the star is the result of a dynamical
+ejection very soon in the cluster evolution. We emphasize that the
 present-day surface helium abundance
 \citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
-limit on the age of the star of $\sim$0.9\,Myr, corresponding to the time needed to
-synthesize this amount of helium in the models from \cite{kohler:15}.
+limit on the age corresponding to the time needed for the star to
+synthesize that amount of helium, and such lower limit corresponds to
+$\sim$0.9\,Myr based on the models from \cite{kohler:15}.
 % \todo{discuss position angle}
@@
   \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
-  \caption{The red solid arrow indicates the proper motion of VFTS682
+  \caption{The red solid arrow indicates the proper motion of VFTS 682
     relative to the region from \emph{Gaia} DR2, starting from the present day position of
@@
     star. The length of the prolongations is proportional to the relative proper motion
-    times the kinetic age of VFTS682. % \todo{load color picture on background? Fig with HST data
+    times the kinetic age of VFTS 682. % \todo{load color picture on background? Fig with HST data
       % available, but confusing...}
@@
 
-Based on our results, we tentatively claim that VFTS682 is the most massive
+Based on our results, we tentatively claim that VFTS 682 is the most massive
 runaway known to date, with a peculiar three-dimensional speed of
@@
 of extremely massive stars by \cite{crowther:10} in the core of the
-cluster. % The projected rotational equatorial
-% velocity\footnote{However, the determination of the rotational
-%   velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
-%   considered with caution.} of VFTS682
-% reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
-% line with the average rotation rate of massive stars in the region
-% \citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
-% experienced rotationally induced chemically homogeneous evolution
-% \citep[][]{maeder:00,demink:09}, and (ii) it has not
-% accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
-% data of the VFTS survey rule out the presence of a close companion at the
-% present day. Moreover,
-
-The spectral type of VFTS682
+cluster. The projected rotational equatorial
+velocity\footnote{However, the determination of the rotational
+  velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
+  considered with caution.} of VFTS 682
+reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
+line with the average rotation rate of massive stars in the region
+\citep[][]{ramirez-agudelo:15}. This suggests that VFTS 682 (i) has not
+experienced rotationally induced chemically homogeneous evolution
+\citep[][]{maeder:00,demink:09}, and (ii) it has not
+accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
+data of the VFTS survey rule out the presence of a close companion at the
+present day. Moreover, the spectral type of VFTS 682
 \citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
@@
 \cite{dekoter:97,crowther:10,crowther:16}, with an astonishing similarity in particular with
-the spectrum of R136a3. Therefore, the isolation of
-VFTS682 makes it an ideal target to constrain the stellar physics of
-stars with masses well above $\sim$$100\,M_\odot$ while avoiding
-crowding issues. %  : its isolation makes
-% it an easier target for observations compared to the similar stars
-% present in the crowded core of R136.
+the spectrum of R136a3. Therefore,
+VFTS 682 might be an ideal target to constrain the stellar physics of
+stars with masses well above $\sim$$100\,M_\odot$: its isolation makes
+it an easier target for observations compared to the similar stars
+present in the crowded core of R136.
 
 \citet{banerjee:12} used N-body simulations of fully segregated
-clusters with all massive stars in binaries to suggest that VFTS682
+clusters with all massive stars in binaries to suggest that VFTS 682
 was ejected from R136. They
@@
 model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS682 and the massive members of R136. While it is not possible to
-robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
+VFTS 682 and the massive members of R136. While it is not possible to
+robustly exclude that VFTS 682 is itself a merger, its spectrum does not show any of the obvious signatures
 like fast rotation, which however could have already slowed down
@@
 expected to have produced
-a large number of massive runaways. Indeed, several %relatively
+a large number of massive runaways. Indeed, several relatively
 isolated massive stars are observed in the region, some with known
-large radial velocities and/or proper motion. % (see
-% \Figref{fig:pm_polar} and Sana et al., in prep.).
-A comprehensive study of the kinematic
+large radial velocities and/or proper motion (see
+\Figref{fig:pm_polar} and Sana et al., in prep.). A comprehensive study of the kinematic
 properties of all the massive stars surrounding R136 might shed light
@@
 is also possible that the star or binary that caused the ejection of
-VFTS682 might have been ejected in the opposite direction, and is also
+VFTS 682 might have been ejected in the opposite direction, and is also
 isolated at present day. If the ejection was caused by an interaction
@@
 
-The similarities between VFTS682 and the WNh5 stars in the core of
+The similarities between VFTS 682 and the WNh5 stars in the core of
 R136 are also in agreement with the ``bully binary'' model of
@@
 pair-instability supernova \cite[e.g.,][]{rakavy:67} at LMC
-metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS682 could be either a
+metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS 682 could be either a
 pair-instability supernova without compact remnant formation, or
@@
 
-The kinematic age of VFTS682 puts an
+The kinematic age of VFTS 682 puts an
 upper limit to the timescale to form the ``bully binary'' in
@@
 evolution, given the age estimate of $\lesssim 2$\,Myr
-\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS682. If the
+\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS 682. If the
 cluster is indeed younger than the shortest stellar lifetime
 \citep[$\sim$3\,Myr, e.g.,][]{brott:11, zapartas:17}, then the alternative
-explanation for ejection of VFTS682 from the disruption of a binary
+explanation for ejection of VFTS 682 from the disruption of a binary
 by a core-collapse event is excluded since the region is too young for stars
@@
 
-The variability of VFTS682, reminiscent of LBV stars, suggests
-that VFTS682 (and therefore its analogs in the core of R136) might
+The variability of VFTS 682, reminiscent of LBV stars, suggests
+that VFTS 682 (and therefore its analogs in the core of R136) might
 experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
 debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
-claim that LBV stars are typically isolated form O-type stars. The fact that VFTS682 is a dynamically
+claim that LBV stars are typically isolated form O-type stars. The fact that VFTS 682 is a dynamically
 ejected runaway which might evolve into an LBV star suggests that
@@
 of motion. The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
-VFTS682 (see \Secref{sec:r136_origin}) is smaller than the
+VFTS 682 (see \Secref{sec:r136_origin}) is smaller than the
 corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic
 age of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
-massive VFTS682 was ejected later than VFTS16 from the same cluster.
+massive VFTS 682 was ejected later than VFTS16 from the same cluster.
 
@@
 question of whether the cluster formed via a monolithic collapse, or
-as a (potentially ongoing) merger of several sub-structures \citep[e.g.,][]{sabbi:12}.
+as a (potentially ongoing) merger of several sub-structures \citep[e.g.,]{sabbi:12}.
 
@@
 
-VFTS682 is potentially the most massive runaway known to date, and its ejection
+VFTS 682 is potentially the most massive runaway known to date, and its ejection
 from the cluster R136 likely implies that it is only the ``tip of the
@@
 
-\bibliographystyle{mnras}
-\bibliography{bibliography/vfts682}
+\bibliographystyle{aa}
+%\bibliography{bibliography/vfts682}
+\begin{thebibliography}{50}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Banerjee} {et~al.}(2012){Banerjee}, {Kroupa}, \& {Oh}}]{banerjee:12}
+{Banerjee}, S., {Kroupa}, P., \& {Oh}, S. 2012, \apj, 746, 15
+
+\bibitem[{{Bate}(2009)}]{bate:09}
+{Bate}, M.~R. 2009, \mnras, 392, 590
+
+\bibitem[{{Bestenlehner} {et~al.}(2011){Bestenlehner}, {Vink}, {Gr{\"a}fener},
+  {Najarro}, {Evans}, {Bastian}, {Bonanos}, {Bressert}, {Crowther}, {Doran},
+  {Friedrich}, {H{\'e}nault-Brunet}, {Herrero}, {de Koter}, {Langer}, {Lennon},
+  {Ma{\'{\i}}z Apell{\'a}niz}, {Sana}, {Soszynski}, \&
+  {Taylor}}]{bestenlehner:11}
+{Bestenlehner}, J.~M., {Vink}, J.~S., {Gr{\"a}fener}, G., {et~al.} 2011, \aap,
+  530, L14
+
+\bibitem[{{Blaauw}(1961)}]{blaauw:61}
+{Blaauw}, A. 1961, \bain, 15, 265
+
+\bibitem[{{Bressert} {et~al.}(2012){Bressert}, {Bastian}, {Evans}, {Sana},
+  {H{\'e}nault-Brunet}, {Goodwin}, {Parker}, {Gieles}, {Bestenlehner}, {Vink},
+  {Taylor}, {Crowther}, {Longmore}, {Gr{\"a}fener}, {Ma{\'{\i}}z
+  Apell{\'a}niz}, {de Koter}, {Cantiello}, \& {Kruijssen}}]{bressert:12}
+{Bressert}, E., {Bastian}, N., {Evans}, C.~J., {et~al.} 2012, \aap, 542, A49
+
+\bibitem[{{Brott} {et~al.}(2011){Brott}, {de Mink}, {Cantiello}, {Langer}, {de
+  Koter}, {Evans}, {Hunter}, {Trundle}, \& {Vink}}]{brott:11}
+{Brott}, I., {de Mink}, S.~E., {Cantiello}, M., {et~al.} 2011, \aap, 530, A115
+
+\bibitem[{{Crowther} {et~al.}(2016){Crowther}, {Caballero-Nieves}, {Bostroem},
+  {Ma{\'{\i}}z Apell{\'a}niz}, {Schneider}, {Walborn}, {Angus}, {Brott},
+  {Bonanos}, {de Koter}, {de Mink}, {Evans}, {Gr{\"a}fener}, {Herrero},
+  {Howarth}, {Langer}, {Lennon}, {Puls}, {Sana}, \& {Vink}}]{crowther:16}
+{Crowther}, P.~A., {Caballero-Nieves}, S.~M., {Bostroem}, K.~A., {et~al.} 2016,
+  \mnras, 458, 624
+
+\bibitem[{{Crowther} {et~al.}(2010){Crowther}, {Schnurr}, {Hirschi}, {Yusof},
+  {Parker}, {Goodwin}, \& {Kassim}}]{crowther:10}
+{Crowther}, P.~A., {Schnurr}, O., {Hirschi}, R., {et~al.} 2010, \mnras, 408,
+  731
+
+\bibitem[{{Davidson} {et~al.}(2016){Davidson}, {Humphreys}, \&
+  {Weis}}]{davidson:16}
+{Davidson}, K., {Humphreys}, R.~M., \& {Weis}, K. 2016, arXiv:1608.02007
+
+\bibitem[{{de Koter} {et~al.}(1997){de Koter}, {Heap}, \&
+  {Hubeny}}]{dekoter:97}
+{de Koter}, A., {Heap}, S.~R., \& {Hubeny}, I. 1997, \apj, 477, 792
+
+\bibitem[{{de Mink} {et~al.}(2009){de Mink}, {Cantiello}, {Langer}, {Pols},
+  {Brott}, \& {Yoon}}]{demink:09}
+{de Mink}, S.~E., {Cantiello}, M., {Langer}, N., {et~al.} 2009, \aap, 497, 243
+
+\bibitem[{{Evans} {et~al.}(2011){Evans}, {Taylor}, {H{\'e}nault-Brunet},
+  {Sana}, {de Koter}, {Sim{\'o}n-D{\'{\i}}az}, {Carraro}, {Bagnoli}, {Bastian},
+  {Bestenlehner}, {Bonanos}, {Bressert}, {Brott}, {Campbell}, {Cantiello},
+  {Clark}, {Costa}, {Crowther}, {de Mink}, {Doran}, {Dufton}, {Dunstall},
+  {Friedrich}, {Garcia}, {Gieles}, {Gr{\"a}fener}, {Herrero}, {Howarth},
+  {Izzard}, {Langer}, {Lennon}, {Ma{\'{\i}}z Apell{\'a}niz}, {Markova},
+  {Najarro}, {Puls}, {Ramirez}, {Sab{\'{\i}}n-Sanjuli{\'a}n}, {Smartt},
+  {Stroud}, {van Loon}, {Vink}, \& {Walborn}}]{evans:11}
+{Evans}, C.~J., {Taylor}, W.~D., {H{\'e}nault-Brunet}, V., {et~al.} 2011, \aap,
+  530, A108
+
+\bibitem[{{Evans} {et~al.}(2010){Evans}, {Walborn}, {Crowther},
+  {H{\'e}nault-Brunet}, {Massa}, {Taylor}, {Howarth}, {Sana}, {Lennon}, \& {van
+  Loon}}]{evans:10}
+{Evans}, C.~J., {Walborn}, N.~R., {Crowther}, P.~A., {et~al.} 2010, \apjl, 715,
+  L74
+
+\bibitem[{{Fujii} \& {Portegies Zwart}(2011)}]{fujii:11}
+{Fujii}, M.~S. \& {Portegies Zwart}, S. 2011, Science, 334, 1380
+
+\bibitem[{{Gaia Collaboration} {et~al.}(2018){Gaia Collaboration}, {Brown},
+  {Vallenari}, {Prusti}, {de Bruijne}, {Babusiaux}, \&
+  {Bailer-Jones}}]{brown:18}
+{Gaia Collaboration}, {Brown}, A.~G.~A., {Vallenari}, A., {et~al.} 2018,
+  ArXiv:1804.09365
+
+\bibitem[{{Gaia Collaboration} {et~al.}(2016){Gaia Collaboration}, {Prusti},
+  {de Bruijne}, {Brown}, {Vallenari}, {Babusiaux}, {Bailer-Jones}, {Bastian},
+  {Biermann}, {Evans}, \& et~al.}]{gaia:16}
+{Gaia Collaboration}, {Prusti}, T., {de Bruijne}, J.~H.~J., {et~al.} 2016,
+  \aap, 595, A1
+
+\bibitem[{{Gruendl} \& {Chu}(2009)}]{gruendl:09}
+{Gruendl}, R.~A. \& {Chu}, Y.-H. 2009, \apjs, 184, 172
+
+\bibitem[{{Gvaramadze} {et~al.}(2012){Gvaramadze}, {Weidner}, {Kroupa}, \&
+  {Pflamm-Altenburg}}]{gvaramadze:12}
+{Gvaramadze}, V.~V., {Weidner}, C., {Kroupa}, P., \& {Pflamm-Altenburg}, J.
+  2012, \mnras, 424, 3037
+
+\bibitem[{{Hogg}(2018)}]{hogg:18}
+{Hogg}, D.~W. 2018, ArXiv:1804.07766
+
+\bibitem[{{Humphreys} {et~al.}(2016){Humphreys}, {Weis}, {Davidson}, \&
+  {Gordon}}]{humphreys:16}
+{Humphreys}, R.~M., {Weis}, K., {Davidson}, K., \& {Gordon}, M.~S. 2016, \apj,
+  825, 64
+
+\bibitem[{{K{\"o}hler} {et~al.}(2015){K{\"o}hler}, {Langer}, {de Koter}, {de
+  Mink}, {Crowther}, {Evans}, {Gr{\"a}fener}, {Sana}, {Sanyal}, {Schneider}, \&
+  {Vink}}]{kohler:15}
+{K{\"o}hler}, K., {Langer}, N., {de Koter}, A., {et~al.} 2015, \aap, 573, A71
+
+\bibitem[{{Kuiper} {et~al.}(2015){Kuiper}, {Yorke}, \& {Turner}}]{kuiper:15}
+{Kuiper}, R., {Yorke}, H.~W., \& {Turner}, N.~J. 2015, \apj, 800, 86
+
+\bibitem[{{Lada} \& {Lada}(2003)}]{lada:03}
+{Lada}, C.~J. \& {Lada}, E.~A. 2003, \araa, 41, 57
+
+\bibitem[{{Lamb} {et~al.}(2016){Lamb}, {Oey}, {Segura-Cox}, {Graus}, {Kiminki},
+  {Golden-Marx}, \& {Parker}}]{lamb:16}
+{Lamb}, J.~B., {Oey}, M.~S., {Segura-Cox}, D.~M., {et~al.} 2016, \apj, 817, 113
+
+\bibitem[{{Langer} {et~al.}(2007){Langer}, {Norman}, {de Koter}, {Vink},
+  {Cantiello}, \& {Yoon}}]{langer:07}
+{Langer}, N., {Norman}, C.~A., {de Koter}, A., {et~al.} 2007, \aap, 475, L19
+
+\bibitem[{{Lennon} {et~al.}(2018){Lennon}, {Evans}, {van der Marel},
+  {Anderson}, {Platais}, {Herrero}, {de Mink}, {Sana}, {Sabbi}, {Crowther},
+  {Langer}, {Ramos Lerate}, {del Pino}, {Renzo}, {Sim{\'o}n-D{\'{\i}}az}, \&
+  {Schneider}}]{lennon:18}
+{Lennon}, D.~J., {Evans}, C.~J., {van der Marel}, R.~P., {et~al.} 2018,
+  ArXiv:1805.08277
+
+\bibitem[{{Lindegren} {et~al.}(2018){Lindegren}, {Hernandez}, {Bombrun},
+  {Klioner}, {Bastian}, {Ramos-Lerate}, {de Torres}, {Steidelmuller},
+  {Stephenson}, {Hobbs}, {Lammers}, {Biermann}, {Geyer}, {Hilger}, {Michalik},
+  {Stampa}, {McMillan}, {Castaneda}, {Clotet}, {Comoretto}, {Davidson},
+  {Fabricius}, {Gracia}, {Hambly}, {Hutton}, {Mora}, {Portell}, {van Leeuwen},
+  {Abbas}, {Abreu}, {Altmann}, {Andrei}, {Anglada}, {Balaguer-Nunez},
+  {Barache}, {Becciani}, {Bertone}, {Bianchi}, {Bouquillon}, {Bourda},
+  {Brusemeister}, {Bucciarelli}, {Busonero}, {Buzzi}, {Cancelliere},
+  {Carlucci}, {Charlot}, {Cheek}, {Crosta}, {Crowley}, {de Bruijne}, {de
+  Felice}, {Drimmel}, {Esquej}, {Fienga}, {Fraile}, {Gai}, {Garralda},
+  {Gonzalez-Vidal}, {Guerra}, {Hauser}, {Hofmann}, {Holl}, {Jordan},
+  {Lattanzi}, {Lenhardt}, {Liao}, {Licata}, {Lister}, {Loffler}, {Marchant},
+  {Martin-Fleitas}, {Messineo}, {Mignard}, {Morbidelli}, {Poggio}, {Riva},
+  {Rowell}, {Salguero}, {Sarasso}, {Sciacca}, {Siddiqui}, {Smart}, {Spagna},
+  {Steele}, {Taris}, {Torra}, {van Elteren}, {van Reeven}, \&
+  {Vecchiato}}]{lindengren:18}
+{Lindegren}, L., {Hernandez}, J., {Bombrun}, A., {et~al.} 2018,
+  ArXiv:1804.09366
+
+\bibitem[{{Maeder} \& {Meynet}(2000)}]{maeder:00}
+{Maeder}, A. \& {Meynet}, G. 2000, \aap, 361, 159
+
+\bibitem[{{Oh} \& {Kroupa}(2016)}]{oh:16}
+{Oh}, S. \& {Kroupa}, P. 2016, \aap, 590, A107
+
+\bibitem[{{Pietrzy{\'n}ski} {et~al.}(2013){Pietrzy{\'n}ski}, {Graczyk},
+  {Gieren}, {Thompson}, {Pilecki}, {Udalski}, {Soszy{\'n}ski}, {Koz{\l}owski},
+  {Konorski}, {Suchomska}, {Bono}, {Moroni}, {Villanova}, {Nardetto},
+  {Bresolin}, {Kudritzki}, {Storm}, {Gallenne}, {Smolec}, {Minniti}, {Kubiak},
+  {Szyma{\'n}ski}, {Poleski}, {Wyrzykowski}, {Ulaczyk}, {Pietrukowicz},
+  {G{\'o}rski}, \& {Karczmarek}}]{pietrzynski:13}
+{Pietrzy{\'n}ski}, G., {Graczyk}, D., {Gieren}, W., {et~al.} 2013, \nat, 495,
+  76
+
+\bibitem[{{Platais} {et~al.}(2018){Platais}, {Lennon}, {van der Marel},
+  {Bellini}, {Sabbi}, {Watkins}, {Sohn}, {Walborn}, {Bedin}, {Evans}, {de
+  Mink}, {Sana}, {Herrero}, {Langer}, \& {Crowther}}]{platais:18}
+{Platais}, I., {Lennon}, D.~J., {van der Marel}, R.~P., {et~al.} 2018,
+  ArXiv:1804.08678
+
+\bibitem[{{Platais} {et~al.}(2015){Platais}, {van der Marel}, {Lennon},
+  {Anderson}, {Bellini}, {Sabbi}, {Sana}, \& {Bedin}}]{platais:15}
+{Platais}, I., {van der Marel}, R.~P., {Lennon}, D.~J., {et~al.} 2015, \aj,
+  150, 89
+
+\bibitem[{{Poveda} {et~al.}(1967){Poveda}, {Ruiz}, \& {Allen}}]{poveda:67}
+{Poveda}, A., {Ruiz}, J., \& {Allen}, C. 1967, Boletin de los Observatorios
+  Tonantzintla y Tacubaya, 4, 86
+
+\bibitem[{{Rakavy} \& {Shaviv}(1967)}]{rakavy:67}
+{Rakavy}, G. \& {Shaviv}, G. 1967, \apj, 148, 803
+
+\bibitem[{{Ram{\'{\i}}rez-Agudelo} {et~al.}(2015){Ram{\'{\i}}rez-Agudelo},
+  {Sana}, {de Mink}, {H{\'e}nault-Brunet}, {de Koter}, {Langer}, {Tramper},
+  {Gr{\"a}fener}, {Evans}, {Vink}, {Dufton}, \& {Taylor}}]{ramirez-agudelo:15}
+{Ram{\'{\i}}rez-Agudelo}, O.~H., {Sana}, H., {de Mink}, S.~E., {et~al.} 2015,
+  \aap, 580, A92
+
+\bibitem[{{Renzo} {et~al.}(2018){Renzo}, {Zapartas}, {de Mink}, {G{\"o}tberg},
+  {Justham}, {Farmer}, {Izzard}, {Toonen}, \& {Sana}}]{renzo:18}
+{Renzo}, M., {Zapartas}, E., {de Mink}, S.~E., {et~al.} 2018, ArXiv:1804.09164
+
+\bibitem[{{Robitaille} \& {Bressert}(2012)}]{robitaille:12}
+{Robitaille}, T. \& {Bressert}, E. 2012, {APLpy: Astronomical Plotting Library
+  in Python}, Astrophysics Source Code Library
+
+\bibitem[{{Rosen} {et~al.}(2016){Rosen}, {Krumholz}, {McKee}, \&
+  {Klein}}]{rosen:16}
+{Rosen}, A.~L., {Krumholz}, M.~R., {McKee}, C.~F., \& {Klein}, R.~I. 2016,
+  \mnras, 463, 2553
+
+\bibitem[{{Rubio-D{\'{\i}}ez} {et~al.}(2017){Rubio-D{\'{\i}}ez}, {Najarro},
+  {Garc{\'{\i}}a}, \& {Sundqvist}}]{rubio-diez:17}
+{Rubio-D{\'{\i}}ez}, M.~M., {Najarro}, F., {Garc{\'{\i}}a}, M., \& {Sundqvist},
+  J.~O. 2017, in IAU Symposium, Vol. 329, The Lives and Death-Throes of Massive
+  Stars, ed. J.~J. {Eldridge}, J.~C. {Bray}, L.~A.~S. {McClelland}, \&
+  L.~{Xiao}, 131--135
+
+\bibitem[{{Sabbi} {et~al.}(2012){Sabbi}, {Lennon}, {Gieles}, {de Mink},
+  {Walborn}, {Anderson}, {Bellini}, {Panagia}, {van der Marel}, \& {Ma{\'{\i}}z
+  Apell{\'a}niz}}]{sabbi:12}
+{Sabbi}, E., {Lennon}, D.~J., {Gieles}, M., {et~al.} 2012, \apjl, 754, L37
+
+\bibitem[{{Sana} {et~al.}(2017){Sana}, {Ram{\'{\i}}rez-Tannus}, {de Koter},
+  {Kaper}, {Tramper}, \& {Bik}}]{sana:17}
+{Sana}, H., {Ram{\'{\i}}rez-Tannus}, M.~C., {de Koter}, A., {et~al.} 2017,
+  \aap, 599, L9
+
+\bibitem[{{Schneider} {et~al.}(2017){Schneider}, {Castro}, {Fossati}, {Langer},
+  \& {de Koter}}]{schneider:17}
+{Schneider}, F.~R.~N., {Castro}, N., {Fossati}, L., {Langer}, N., \& {de
+  Koter}, A. 2017, \aap, 598, A60
+
+\bibitem[{{Schneider} {et~al.}(2018){Schneider}, {Sana}, {Evans},
+  {Bestenlehner}, {Castro}, {Fossati}, {Gr{\"a}fener}, {Langer},
+  {Ram{\'{\i}}rez-Agudelo}, {Sab{\'{\i}}n-Sanjuli{\'a}n},
+  {Sim{\'o}n-D{\'{\i}}az}, {Tramper}, {Crowther}, {de Koter}, {de Mink},
+  {Dufton}, {Garcia}, {Gieles}, {H{\'e}nault-Brunet}, {Herrero}, {Izzard},
+  {Kalari}, {Lennon}, {Ma{\'{\i}}z Apell{\'a}niz}, {Markova}, {Najarro},
+  {Podsiadlowski}, {Puls}, {Taylor}, {van Loon}, {Vink}, \&
+  {Norman}}]{schneider:18}
+{Schneider}, F.~R.~N., {Sana}, H., {Evans}, C.~J., {et~al.} 2018, Science, 359,
+  69
+
+x\bibitem[{{Smith}(2016)}]{smith:16}
+{Smith}, N. 2016, \mnras, 461, 3353
+
+\bibitem[{{Smith} \& {Tombleson}(2015)}]{smith:15}
+{Smith}, N. \& {Tombleson}, R. 2015, \mnras, 447, 598
+
+\bibitem[{{Udalski} {et~al.}(2008){Udalski}, {Soszy{\'n}ski}, {Szyma{\'n}ski},
+  {Kubiak}, {Pietrzy{\'n}ski}, {Wyrzykowski}, {Szewczyk}, {Ulaczyk}, \&
+  {Poleski}}]{udalski:08}
+{Udalski}, A., {Soszy{\'n}ski}, I., {Szyma{\'n}ski}, M.~K., {et~al.} 2008,
+  \actaa, 58, 329
+
+\bibitem[{{Ward} \& {Kruijssen}(2018)}]{ward:18}
+{Ward}, J.~L. \& {Kruijssen}, J.~M.~D. 2018, \mnras, 475, 5659
+
+\bibitem[{{Zapartas} {et~al.}(2017){Zapartas}, {de Mink}, {Izzard}, {Yoon},
+  {Badenes}, {G{\"o}tberg}, {de Koter}, {Neijssel}, {Renzo}, {Schootemeijer},
+  \& {Shrotriya}}]{zapartas:17}
+{Zapartas}, E., {de Mink}, S.~E., {Izzard}, R.~G., {et~al.} 2017, \aap, 601,
+  A29
+
+\bibitem[{{Zinnecker} \& {Yorke}(2007)}]{zinnecker:07}
+{Zinnecker}, H. \& {Yorke}, H.~W. 2007, \araa, 45, 481
+
+\bibitem[{{Zwicky}(1957)}]{zwicky:57}
+{Zwicky}, F. 1957, \zap, 44, 64
+\end{thebibliography}
 
 \small
-We are grateful to S.~Torres, M.~C.~Ramirez-Tannus,
-and C.~J.~Evans for help and discussions.
+We acknowledge invaluable help from S.~Torres, M.~C.~Ramirez-Tannus,
+and C.~J.~Evans.
 SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC) (Grant agreement No.

### 2018-06-06 VFTS682_APJL.tex
minor: title change

@@
 % runaway star known to date}
-\title{\emph{Gaia} DR2 adds clues for dynamical ejection of
-  a $\sim$$150\,M_\odot$ star}
+% \title{\emph{Gaia} DR2 adds clues for dynamical ejection of
+%   a $\sim$$150\,M_\odot$ star}
+\title{Gaia's constraints on the origin of the  $\sim$$150\,M_\odot$ candidate runaway star VFTS682}
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
@@
   We use the proper motions from the second \emph{Gaia} data
-  release (DR2) together with multiepoch \emph{Hubble Space Telescope} proper motions to investigate the kinematics of this object. We show
+  release together with multiepoch \emph{Hubble Space Telescope} proper motions to investigate the kinematics of this object. We show
   that the projected velocity of VFTS682 is consistent with it being a
@@
 
-A contribution to the debate on whether or not massive
-stars form in relative isolation was presented by
-\cite{bestenlehner:11} and \cite{bressert:12}, who discussed the case of the very massive star
-VFTS682.  This star is located in the field of the 30 Doradus region
+% A contribution to the debate on whether or not massive
+% stars form in relative isolation was presented by
+% \cite{bestenlehner:11} and \cite{bressert:12}, who discussed the case of the very massive star
+% VFTS682.
+\cite{bestenlehner:11} and \cite{bressert:12} discussed in this
+context the case of the very massive star VFTS682.
+This star is located in the field of the 30 Doradus region
 in the Large Magellanic Cloud (LMC) and was studied as part of the
@@
 R136 cluster. In particular, a remarkable similarity exist between the
-spectrum of VFTS682 and R136a3, located in the core of the cluster
-\citep{crowther:10} for which \citet{crowther:16} report a
-current mass estimate of $180^{+30}_{-30}\Msun$. The R136 cluster hosts
+spectrum of VFTS682 and R136a3 for which \citet{crowther:16} report a
+current mass estimate of $180^{+30}_{-30}\Msun$. R136 hosts
 at least two more very massive WN5h stars, R136a1 and R136a2, whose
@@
 \citet{bestenlehner:11} discuss two possibilities for the
-origin of VFTS682. % Quoting directly 
-From their abstract: "(i) the
-star either formed in situ, which would have profound implications for
-the formation mechanism of massive stars, or (ii) VFTS682 is a slow
-runaway star that originated from the dense cluster R136, which would
-make it the most massive runaway known to date". Based on their
+origin of VFTS682: it either formed in situ as an isolated massive
+star, or it is a slow runaway originating from R136. % Quoting directly 
+% From their abstract: "(i) the
+% star either formed in situ, which would have profound implications for
+% the formation mechanism of massive stars, or (ii) VFTS682 is a slow
+% runaway star that originated from the dense cluster R136, which would
+% make it the most massive runaway known to date".
+Based on their
 dynamical models, \citet{fujii:11} and \citet{banerjee:12} argued that
@@
 the isolated O-type stars around R136. In this study, we
-report on our analysis of the \emph{Gaia} DR2 data for VFTS682 which
+analyze the \emph{Gaia} DR2 data for VFTS682 which
 was not part of the sample of \citet{lennon:18} because of its 
@@
 ]{evans:11}
-was originally classified as a young stellar object \citep{gruendl:09}
-based on its mid-infrared excess. \citet{evans:11} reclassified the
+was originally classified as a young stellar object based on its
+mid-infrared excess \citep{gruendl:09}. \citet{evans:11} reclassified the
 object as Wolf-Rayet star of spectral type WNh5 using multi-epoch
@@
 apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
-of $150.0^{+28.7}_{-17.4}\Msun$, using the Bayesian analysis tool BONNSAI
-\citep{schneider:17}, based on the evolutionary models from
-\citet{brott:11} and \cite{kohler:15}. % These values place VFTS682 around the
+of $150.0^{+28.7}_{-17.4}\Msun$.% , using the Bayesian analysis tool BONNSAI
+% \citep{schneider:17}, based on the evolutionary models from
+% \citet{brott:11} and \cite{kohler:15}.
+% These values place VFTS682 around the
 % ``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
@@
 Further worth noticing is the variability of the
-star. % \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
-% lower than found by \citet{evans:11}.
+star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
+lower than found by \citet{evans:11}.
 \citet{bestenlehner:11} discuss the Optical Gravitational Lensing
@@
 radial velocities for emission-lines stars is difficult because of the
-presence of an optically thick, likely non-homogeneous, and variable
-wind obscuring the stellar surface. The radial velocity errors quoted
-above quantify only statistical uncertainties from the
-spectral analysis, and not the systematic errors arising from the
-modeling of emission line stars.
+presence of an optically thick, non-homogeneous, and variable
+wind obscuring the photosphere. The radial velocity errors quoted
+above are only statistical, and do not quantify the (possibly larger)
+systematic errors.% quantify only statistical uncertainties from the
+% spectral analysis, and not the systematic errors arising from the
+% modeling of emission line stars.
 
@@
 
-The second \emph{Gaia} Data Release (DR2)% , which
-became available on 25 April 2018 and 
-provides the five-parameter astrometric solution (positions,
-parallaxes, and proper motion components) for more than a billion
+The second \emph{Gaia} Data Release (DR2) % , which
+% became available on 25 April 2018 and 
+provides positions,
+parallaxes, and proper motion components for more than a billion
 sources \citep{brown:18}. VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2
-catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. \emph{Gaia}
-reports a G-band magnitude of 15.65. The star has a
+catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Its
+G-band magnitude from \emph{Gaia}
+% reports a G-band magnitude of
+is 15.65. The star has a
 \texttt{visibility\_period} = 17, which counts how many observations have
@@
 \texttt{astrometric\_excess\_noise} = 0. These values suggest that the \emph{Gaia}
-data for VFTS682 are reliable. However, the effective temperature
-reported in \emph{Gaia} DR2, based on the spectral fitting around the
-CaII triplet, is one order of magnitude lower than what found by
-\cite{bestenlehner:11}, and the best fit parallax of this star is
-negative \citep[see, e.g.,][]{hogg:18}. We do not use the effective temperature of the star anywhere
-in this study, and we attribute the unphysical value of the parallax
-to the large distance to the LMC.%  Our main findings do not rely on the
+data for VFTS682 are reliable. % However, the effective temperature
+% reported in \emph{Gaia} DR2, based on the spectral fitting around the
+% CaII triplet, is one order of magnitude lower than what found by
+% \cite{bestenlehner:11}, and the best fit parallax of this star is
+% negative \citep[see, e.g.,][]{hogg:18}. We do not use the effective temperature of the star anywhere
+% in this study, and we attribute the unphysical value of the parallax
+% to the large distance to the LMC.
+% Our main findings do not rely on the
 % parallax nor the effective temperature values reported in the \emph{Gaia} DR2
@@
     defining our reference frame. VFTS682 is indicated by the plus sign, and
-    other notable outliers (see \citet{lennon:18}) are also labeled. Concentric dashed-lines denote relative proper motions of
+    other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
     0.1, 0.2, and 0.4$\,\mathrm{mas\ yr^{-1}}$. Positive angle
@@
     direction, with 0 degrees corresponding to proper motion pointing
-    radially outward from R136.
+    radially outward from R136. The proper motion for VFTS682
+    corresponds to almost radial motion away from the cluster.
   }
@@
 
-% \bibitem[\protect\citeauthoryear{{Parker}}{{Parker}}{1993}]{parker:93}
-% {Parker} J.~W.,  1993, \mn@doi [\aj] {10.1086/116661}, \href
-%   {http://adsabs.harvard.edu/abs/1993AJ....106..560P} {106, 560}
+\bibitem[\protect\citeauthoryear{{Parker}}{{Parker}}{1993}]{parker:93}
+{Parker} J.~W.,  1993, \mn@doi [\aj] {10.1086/116661}, \href
+  {http://adsabs.harvard.edu/abs/1993AJ....106..560P} {106, 560}
 
@@
 
-\bibitem[\protect\citeauthoryear{{Schneider}, {Castro}, {Fossati}, {Langer}  \&
-  {de Koter}}{{Schneider} et~al.}{2017}]{schneider:17}
-{Schneider} F.~R.~N.,  {Castro} N.,  {Fossati} L.,  {Langer} N.,   {de Koter}
-  A.,  2017, \mn@doi [\aap] {10.1051/0004-6361/201628409}, \href
-  {http://adsabs.harvard.edu/abs/2017A\%26A...598A..60S} {598, A60}
+% \bibitem[\protect\citeauthoryear{{Schneider}, {Castro}, {Fossati}, {Langer}  \&
+%   {de Koter}}{{Schneider} et~al.}{2017}]{schneider:17}
+% {Schneider} F.~R.~N.,  {Castro} N.,  {Fossati} L.,  {Langer} N.,   {de Koter}
+%   A.,  2017, \mn@doi [\aap] {10.1051/0004-6361/201628409}, \href
+%   {http://adsabs.harvard.edu/abs/2017A\%26A...598A..60S} {598, A60}
 

### 2018-06-06 VFTS682_APJL.tex
none

--- 2018-06-06 VFTS682_APJL.tex
@@
 % runaway star known to date}
-\title{\emph{Gaia} DR2 adds clues for dynamical ejection of
-  a $\sim$$150\,M_\odot$ star}
+%\title{\emph{Gaia} DR2 adds clues for dynamical ejection of a $\sim$$150\,M_\odot$ star}
+
+%  \title{Gaia's constraints on kinematics of the $\sim$$150\,M_\odot$  star VFTS 682 and the implications for its nature as  a possible runaway star ejected dynamically from R136?}
+%  \title{Is the $\sim$$150\,M_\odot$  star VFTS 682 the most massive dynamically ejected star known to date?}
+
+\title{Gaia's constraints on the origin of the  $\sim$$150\,M_\odot$ candidate runaway star VFTS 682}
+
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
@@
 \begin{abstract}
+
   The formation and fates of the most massive stars remain among the most elusive questions in stellar astrophysics. 

### 2018-06-07 VFTS682_APJL.tex

--- 2018-06-06 VFTS682_APJL.tex
@@
 % runaway star known to date}
-% \title{\emph{Gaia} DR2 adds clues for dynamical ejection of
-%   a $\sim$$150\,M_\odot$ star}
-\title{Gaia's constraints on the origin of the  $\sim$$150\,M_\odot$ candidate runaway star VFTS682}
+\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
+
 \author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
@@
 \begin{abstract}
-  The formation and fates of the most massive stars remain among the most elusive questions in stellar astrophysics. 
-  % 
-  Previous spectroscopic studies identified VFTS682 as one of the most
-  massive stars known. It resides in the field of the 30 Dor region in
-  the Large Magellanic Cloud at a projected distance of 29\,pc of the
-  star cluster Radcliffe 136 (R136). The inconclusiveness of the radial
-  velocity analysis for this emission line star led to speculations whether such extreme
-  stars can form in relative isolation. 
-  % 
-  We use the proper motions from the second \emph{Gaia} data
-  release together with multiepoch \emph{Hubble Space Telescope} proper motions to investigate the kinematics of this object. We show
-  that the projected velocity of VFTS682 is consistent with it being a
-  runaway star dynamically ejected from R136. Its
-  projected two-dimensional speed relative to the cluster is $\sim$$30\pm20\kms$, implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
-  accounting for the (uncertain) velocity component along the line of sight. The
-  inferred kinematic age is also  consistent with the dynamical ejection scenario from the cluster. 
-  % 
-  If confirmed by future astrometric data, this would make VFTS682
-  the most massive runaway star known to date.
-It would prove that star clusters are capable of ejecting their most massive members.
-We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
+
+ How very massive stars form is still among the most intriguing open questions in stellar astrophysics.
+VFTS 682 is among the most extreme massive stars known, with an inferred initial mass of about $150\,M_\odot$.
+It is located in the Tarantula nebula at a projected distance of 29\,pc from the central cluster R136.
+The absence of other massive stars in its immediate vicinity lead to two intriguing possible hypotheses posed in earlier work.
+Either it formed in relative isolation through a new mode of star formation or it was ejected dynamically from the central cluster.
+
+Aiming to shine light on this debate, we investigate the kinematics of  VFTS 682 as obtained by \emph{Gaia} and a \emph{Hubble Space Telescope} proper motion campaign.
+We derive a projected velocity relative of to the cluster of $\sim$$30\pm20\kms$.
+The direction and magnitude of the proper motion are consistent with ejection from the central cluster considering its inferred age.
+However, the error bars and spread in proper motions derived for other sources in the region are substantial, which makes us cautious to derive hard conclusions.
+
+If future data confirms the runaway nature, this would make the VFTS682 the most massive runaway star known to date. 
+While we cannot prove this solidly from the current data, we do consider this hypothesis plausible.
+The central cluster is known to harbor several other ``monster stars'' similar spectral types and high inferred masses.
+Moreover, the current record holders are O type stars including VFTS 16 which provides direct evidence that R136 is indeed capable of ejecting some of its most massive members.
+We dicuss further implications and caveats.
+
 \end{abstract}
@@
 How massive stars form is one of the major longstanding questions in astrophysics
-\citep[e.g.,][]{zinnecker:07}. Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
+\citep[e.g.,][]{zinnecker:07}. Improving our understanding of massive star formation, and its
+possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies, but also for understanding the transients that mark their death and the compact remnants they leave behind.
+%
+Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
 % \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
@@
 their parent cloud during the entirety of their formation
-process. Major progress has been made on the theoretical side,
-\citep[e.g.][]{kuiper:15,rosen:16}, but simulations cannot yet resolve
-the length scales  of interest to explain the 
-high multiplicity of massive stars  \citep[e.g.,][]{bate:09, sana:17}. Understanding massive star formation, and its
-possible dependence on environment and metallicity, is crucial for
-understanding the role massive stars play within their host galaxies,
-% as sources of feedback,
-but also for understanding the transients that
-mark their death and the compact remnants they leave behind. % This is
-% therefore a key question for the present and upcoming transient
-% surveys, e.g., Large Synoptic Spectroscopic Survey and LIGO/Virgo,
-% % \citep[][]{}\todo{ref} . Blackgem is still hypothetical and a very tiny project compared to the others. 
-% which  will reveal transients associated to massive stars
-% evolution and death.
-
-In has been proposed that most, if not all, stars form in clusters
-\citep{lada:03}, where massive stars are thought to  form
-preferentially in the dense cores. In this picture, field stars are primarily the result of the dissolution of dense groups. 
-However, a significant population of massive stars exists in relative
-isolation,  far from dense clusters or OB associations. The origin of
-this population still remains debated \citep{gvaramadze:12, lamb:16,ward:18}.  One
-hypothesis is that they  formed in the field, but this poses a
-challenge for the theories of star formation. The alternative
-hypothesis is that these massive stars formed in clusters, and then
-were ejected from their birth locations. Such ejections may result
-from dynamical interactions \citep[e.g.,][]{poveda:67} or from the
-disruption of binary systems at the death of the companion %by the core-collapse of the companion
-star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.
+process. Important progress has been made on the theoretical side,
+\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally very expensive and therefore remain challenging.
+
+In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to  form preferentially in the dense cores.
+In this picture, field stars are primarily the result of the dissolution of dense groups.
+However, a significant population of massive stars exists in relative isolation, far from dense clusters or OB associations and their origin remains matter of debate \citep{gvaramadze:12, lamb:16,ward:18}.
+One hypothesis to explain the population of relatively isolated massive stars is that they formed in the field.
+The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
+Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.
 
@@
 % VFTS682.
-\cite{bestenlehner:11} and \cite{bressert:12} discussed in this
-context the case of the very massive star VFTS682.
-This star is located in the field of the 30 Doradus region
-in the Large Magellanic Cloud (LMC) and was studied as part of the
-multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}. Its spectral type is WNh5 and its present-day mass is $137.8^{+27.5}_{-15.9}\,M_\odot$,
-\citep{schneider:18}. These correspond to an inferred initial mass of
-$150.0^{+28.7}_{-17.4}\,M_\odot$ and make VFTS682 one of the
-most extreme objects in the region.
+
+One of the most extreme examples that has been considered in this debate is the very massive star VFTS682  \cite{bestenlehner:11, bressert:12}.
+This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}.
+It is a hydrogen-rich Wolf-Rayet star of spectral type is WNh5.
+Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $140^{+30}_{-16}\,M_\odot$ corresponding to an initial mass of $150.0^{+30}_{-17}\,M_\odot$
+\citep{schneider:18}. 
+%
+%$137.8^{+27.5}_{-15.9}\,M_\odot$,
+%
+This makes VFTS682 one of the most massive stars known and one of the most extreme objects in the region.
 %
@@
 massive stars % , a.k.a. ``monster stars'',
-identified by
-\citet{dekoter:97,crowther:10, crowther:16} in the core of the
-R136 cluster. In particular, a remarkable similarity exist between the
+ in the core of the R136 cluster \citep{dekoter:97,crowther:10, crowther:16}. 
+ %
+ In particular, a remarkable similarity exist between the
 spectrum of VFTS682 and R136a3 for which \citet{crowther:16} report a
@@
 at least two more very massive WN5h stars, R136a1 and R136a2, whose
-estimated current masses are even higher. However, VFTS682 stands
-out by its isolation at a projected offset of 119.4 arcseconds, corresponding to $\sim$$29$\,pc from the core of the star cluster
-R136.
-
-\citet{bestenlehner:11} discuss two possibilities for the
-origin of VFTS682: it either formed in situ as an isolated massive
-star, or it is a slow runaway originating from R136. % Quoting directly 
-% From their abstract: "(i) the
-% star either formed in situ, which would have profound implications for
-% the formation mechanism of massive stars, or (ii) VFTS682 is a slow
-% runaway star that originated from the dense cluster R136, which would
-% make it the most massive runaway known to date".
-Based on their
-dynamical models, \citet{fujii:11} and \citet{banerjee:12} argued that
-ejection of VFTS682 from R136 would be natural. Testing the origin of
-VFTS682 is not only important for constraining isolated star
-formation; it should also help to improve understanding of early star-cluster dynamics.
-
-Recently, \citet{platais:15,platais:18} used multi-epoch \emph{Hubble Space
-  Telescope} (HST) imaging to identify % and shed new light on
-runaway stars in
- 30 Doradus based on their proper motions. \citet{lennon:18} used data from the second \emph{Gaia}
-data release \cite[DR2,][]{gaia:16,brown:18} to further investigate
-the isolated O-type stars around R136. In this study, we
-analyze the \emph{Gaia} DR2 data for VFTS682 which
-was not part of the sample of \citet{lennon:18} because of its 
-spectral type. We combine the radial velocity measurements from the
-VFTS survey \citep[][]{evans:11} with the proper motion from
-\emph{Gaia} DR2 to reconstruct the three-dimensional velocity of VFTS682, and test the hypothesis that this star was ejected from R136. We
-use the HST proper motions from \citet{platais:18} to test our
-analysis of the \emph{Gaia} DR2 data.
+estimated current masses are even higher.
+
+%at a projected offset of 119.4 arcseconds, corresponding to
+
+VFTS682 stands out by its relative isolation at a projected distance of $\sim$$29$\,pc from  the star cluster R136.
+\citet{bestenlehner:11} discuss the possible explanation for the offset.
+They consider formation in situ as an isolated massive star, but remark that this poses a challenge for starformation theories.
+The alternative, ejection from  R136, appears to be reasonable.
+N-body simulations of dynamical interaction in star clusters indicate that the ejection of very massive stars like VFTS682  is expected \citep[e.g.][]{fujii:11, banerjee:12}.
+This is supported by the recent findings of other massive runaway stars in the region based on proper motion studies.
+\citet{platais:15,platais:18} analyze multi-epoch \emph{Hubble Space Telescope} (HST) photometry and identify 10 stars, which appear to be ejected from R136.
+\citet{lennon:18} investigate the kinematics of  isolated O-type stars in the region using the second \emph{Gaia} data release \cite[DR2,][]{gaia:16,brown:18} and show that the proper motion, postion and direction of the $\sim100\Msun$ star VFTS16 is consistent with a runaway origin from R136.
+
+%Testing the origin of VFTS682 is not only important for constraining isolated star formation; it should also help to improve understanding of early star-cluster dynamics.
+
+In this paper we present an analysis of the new kinematical constraints for VFTS682 provided by \emph{Gaia} DR2 and constraints from HST proper data  by \citet{platais:18}.
+We discuss the implications for the hypothesis that VFTS682 is a slow runaway star ejected from R136 and discuss the caveats and implications.
+
+% which was not part of the sample of \citet{lennon:18} because of its  spectral type.
 
@@
 \citet{bestenlehner:11} reported for VFTS682 a wind mass loss rate of
-$\sim10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, % inferred from
+$10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, % inferred from
 % spectral analysis assuming a smooth outflow, i.e.,
@@
 
-We obtain the peculiar radial velocity $\delta v_\mathrm{rad}$ of VFTS682 as the difference
-between the average radial velocity of the 30 Doradus region
-($270\pm10\kms$) minus the radial velocity measured from the NV $\lambda4944$
-line for VFTS682  \citep[$300\pm10\kms$, ][]{bestenlehner:11}. We implicitly assume that the use
-of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not
-introduce significant errors. \cite{bressert:12} also noted that VFTS682 shows a significant radial velocity offset compared to the nebular
-lines from the gas surrounding it, suggesting peculiar motion in the
-line of sight direction. However, the determination of
-radial velocities for emission-lines stars is difficult because of the
-presence of an optically thick, non-homogeneous, and variable
-wind obscuring the photosphere. The radial velocity errors quoted
-above are only statistical, and do not quantify the (possibly larger)
-systematic errors.% quantify only statistical uncertainties from the
-% spectral analysis, and not the systematic errors arising from the
-% modeling of emission line stars.
-
-\subsection{HST astrometry for VFTS682}
-Recently, \citet{platais:18} presented HST  astrometry and proper
-motions in the 30 Doradus region based on two epochs. %  of
+\cite{bressert:12} note that VFTS682 shows a significant radial velocity offset compared to the nebular lines from the gas surrounding it, suggesting peculiar motion in the line of sight direction.
+\citet[ ][]{bestenlehner:11} quote a radial velocity measured from the NV $\lambda4944$ of   $300\pm10\kms$.
+The average radial velocity of the 30 Doradus region is
+$270\pm10\kms$.
+Whether this measurement truly reflects a peculiar radial velocity is debated.
+The determination of radial velocities for emission-lines stars is difficult because of the presence of an optically thick, non-homogeneous, and variable wind obscuring the photosphere.
+The radial velocity errors quoted above are only statistical, and do not quantify the (possibly larger)
+systematic errors.  \SdM{TODO: Possibly mention and cite paper in prep by Paco and student on Xshooter data?  Possibly mention (or even use) on their measurement for the RV?}
+
+%We implicitly assume that the use of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not introduce significant errors.
+
 % observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
@@
 % that the displacement measurements could be purely differential.
-While the most bright stars (V$<$14) were saturated in
-their data and missing from that catalog, due to its
-high extinction, VFTS682 is faint enough
-to be included in that catalog. Its proper motions are
-listed in \Tabref{tab:vfts682}. For
-large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even
-exceed \emph{Gaia} DR2 data in quality.
+
+\subsection{HST astrometry for VFTS682}
+For large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even exceed \emph{Gaia} DR2 data in quality.
+
+Recently, \citet{platais:18} presented HST astrometry and proper motions in the 30 Doradus region (GO-12499; P.I.: D.~J.~Lennon).
+The brightest stars (V$<$14) are missing from the catalogue because of saturation, but the high extinction towards VFTS682 makes is fain enough to be included in this catalogue.
+The  proper motions are
+listed in \Tabref{tab:vfts682}.
 

### 2018-06-08 VFTS682_APJL.tex
minor, likely needs merger

@@
   R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
-  V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
+  V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{1,7,8}$,  A.~de~Koter$^{1}$,
   N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
@@
     \end{tabular}
-    % \tablecomments
+    \tablecomments
     {The error on the RA and DEC positions, are of order
@@
 
-\bibitem[\protect\citeauthoryear{{Hogg}}{{Hogg}}{2018}]{hogg:18}
-{Hogg} D.~W.,  2018, ArXiv:1804.07766, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180407766H} {}
+% \bibitem[\protect\citeauthoryear{{Hogg}}{{Hogg}}{2018}]{hogg:18}
+% {Hogg} D.~W.,  2018, ArXiv:1804.07766, \href
+%   {http://adsabs.harvard.edu/abs/2018arXiv180407766H} {}
 
@@
 
-\bibitem[\protect\citeauthoryear{{Maeder} \& {Meynet}}{{Maeder} \&
-  {Meynet}}{2000}]{maeder:00}
-{Maeder} A.,  {Meynet} G.,  2000, \aap, \href
-  {http://adsabs.harvard.edu/abs/2000A\%26A...361..159M} {361, 159}
+% \bibitem[\protect\citeauthoryear{{Maeder} \& {Meynet}}{{Maeder} \&
+%   {Meynet}}{2000}]{maeder:00}
+% {Maeder} A.,  {Meynet} G.,  2000, \aap, \href
+%   {http://adsabs.harvard.edu/abs/2000A\%26A...361..159M} {361, 159}
 
@@
 
-\bibitem[\protect\citeauthoryear{{Ram{\'{\i}}rez-Agudelo}
-  et~al.,}{{Ram{\'{\i}}rez-Agudelo} et~al.}{2015}]{ramirez-agudelo:15}
-{Ram{\'{\i}}rez-Agudelo} O.~H.,  et~al., 2015, \mn@doi [\aap]
-  {10.1051/0004-6361/201425424}, \href
-  {http://adsabs.harvard.edu/abs/2015A\%26A...580A..92R} {580, A92}
+% \bibitem[\protect\citeauthoryear{{Ram{\'{\i}}rez-Agudelo}
+%   et~al.,}{{Ram{\'{\i}}rez-Agudelo} et~al.}{2015}]{ramirez-agudelo:15}
+% {Ram{\'{\i}}rez-Agudelo} O.~H.,  et~al., 2015, \mn@doi [\aap]
+%   {10.1051/0004-6361/201425424}, \href
+%   {http://adsabs.harvard.edu/abs/2015A\%26A...580A..92R} {580, A92}
 
@@
 
-\bibitem[\protect\citeauthoryear{{Robitaille} \& {Bressert}}{{Robitaille} \&
-  {Bressert}}{2012}]{robitaille:12}
-{Robitaille} T.,  {Bressert} E.,  2012, {APLpy: Astronomical Plotting Library
-  in Python}, Astrophysics Source Code Library (\mn@eprint {ascl} {1208.017})
+% \bibitem[\protect\citeauthoryear{{Robitaille} \& {Bressert}}{{Robitaille} \&
+%   {Bressert}}{2012}]{robitaille:12}
+% {Robitaille} T.,  {Bressert} E.,  2012, {APLpy: Astronomical Plotting Library
+%   in Python}, Astrophysics Source Code Library (\mn@eprint {ascl} {1208.017})
 
@@
 
-\bibitem[\protect\citeauthoryear{{de Mink}, {Cantiello}, {Langer}, {Pols},
-  {Brott}  \& {Yoon}}{{de Mink} et~al.}{2009}]{demink:09}
-{de Mink} S.~E.,  {Cantiello} M.,  {Langer} N.,  {Pols} O.~R.,  {Brott} I.,
-  {Yoon} S.-C.,  2009, \mn@doi [\aap] {10.1051/0004-6361/200811439}, \href
-  {http://adsabs.harvard.edu/abs/2009A\%26A...497..243D} {497, 243}
+% \bibitem[\protect\citeauthoryear{{de Mink}, {Cantiello}, {Langer}, {Pols},
+%   {Brott}  \& {Yoon}}{{de Mink} et~al.}{2009}]{demink:09}
+% {de Mink} S.~E.,  {Cantiello} M.,  {Langer} N.,  {Pols} O.~R.,  {Brott} I.,
+%   {Yoon} S.-C.,  2009, \mn@doi [\aap] {10.1051/0004-6361/200811439}, \href
+%   {http://adsabs.harvard.edu/abs/2009A\%26A...497..243D} {497, 243}
 

### 2018-06-08 VFTS682_APJL.tex
editing

@@
 
- How very massive stars form is still among the most intriguing open questions in stellar astrophysics.
-VFTS 682 is among the most extreme massive stars known, with an inferred initial mass of about $150\,M_\odot$.
-It is located in the Tarantula nebula at a projected distance of 29\,pc from the central cluster R136.
+ How very massive stars form is still among the most intriguing open
+ questions in stellar astrophysics.  VFTS 682 is among the most
+ extreme massive stars known, with an inferred initial mass of about
+ $150\,M_\odot$.
+It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.
 The absence of other massive stars in its immediate vicinity lead to two intriguing possible hypotheses posed in earlier work.
@@
 If future data confirms the runaway nature, this would make the VFTS682 the most massive runaway star known to date. 
-While we cannot prove this solidly from the current data, we do consider this hypothesis plausible.
-The central cluster is known to harbor several other ``monster stars'' similar spectral types and high inferred masses.
-Moreover, the current record holders are O type stars including VFTS 16 which provides direct evidence that R136 is indeed capable of ejecting some of its most massive members.
-We dicuss further implications and caveats.
-
+While we cannot prove this solidly from the current data, we do
+consider this hypothesis plausible. The central cluster is known to
+harbor several other stars more massive than $150\,M_\odot$ similar in
+spectral type.  Moreover, the current record holders are O type stars
+including VFTS 16 which provides direct evidence that R136 is indeed
+capable of ejecting some of its most massive members. 
 \end{abstract}
@@
 
-In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to  form preferentially in the dense cores.
+In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to reside in the innermost cores.
 In this picture, field stars are primarily the result of the dissolution of dense groups.
@@
 
-One of the most extreme examples that has been considered in this debate is the very massive star VFTS682  \cite{bestenlehner:11, bressert:12}.
+One of the most extreme examples that has been considered in this debate is the very massive star VFTS682  \citep[][]{bestenlehner:11, bressert:12}.
 This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}.
-It is a hydrogen-rich Wolf-Rayet star of spectral type is WNh5.
-Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $140^{+30}_{-16}\,M_\odot$ corresponding to an initial mass of $150.0^{+30}_{-17}\,M_\odot$
+It is a hydrogen-rich Wolf-Rayet star of spectral type WNh5.
+Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $\sim$$140^{+30}_{-16}\,M_\odot$ corresponding to an initial mass of $\sim$$150^{+30}_{-17}\,M_\odot$
 \citep{schneider:18}. 
@@
 
-VFTS682 stands out by its relative isolation at a projected distance of $\sim$$29$\,pc from  the star cluster R136.
-\citet{bestenlehner:11} discuss the possible explanation for the offset.
-They consider formation in situ as an isolated massive star, but remark that this poses a challenge for starformation theories.
-The alternative, ejection from  R136, appears to be reasonable.
-N-body simulations of dynamical interaction in star clusters indicate that the ejection of very massive stars like VFTS682  is expected \citep[e.g.][]{fujii:11, banerjee:12}.
+VFTS682 stands out by its relative isolation at a projected distance of 119.4 arcseconds, corresponding to 
+$\sim$$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
+considered two possible explanation for the offset: either
+the star formed in situ as an isolated massive star, % , but remark that this
+                                % poses a challenge for starformation
+                                % theories (They don't)
+or it was ejected from  R136. N-body simulations % of dynamical interaction in star clusters
+indicate that the ejection of very massive stars like VFTS682 is expected \citep[e.g.][]{fujii:11, banerjee:12}.
 This is supported by the recent findings of other massive runaway stars in the region based on proper motion studies.
@@
 
-In this paper we present an analysis of the new kinematical constraints for VFTS682 provided by \emph{Gaia} DR2 and constraints from HST proper data  by \citet{platais:18}.
-We discuss the implications for the hypothesis that VFTS682 is a slow runaway star ejected from R136 and discuss the caveats and implications.
-
+In this paper we present an analysis of the new kinematical
+constraints for VFTS682 provided by \emph{Gaia} DR2 and constraints
+from HST proper data  by \citet{platais:18}.   We discuss the
+implications for the hypothesis that VFTS682 is a slow runaway star
+ejected from R136.
 % which was not part of the sample of \citet{lennon:18} because of its  spectral type.
@@
 
-\cite{bressert:12} note that VFTS682 shows a significant radial velocity offset compared to the nebular lines from the gas surrounding it, suggesting peculiar motion in the line of sight direction.
-\citet[ ][]{bestenlehner:11} quote a radial velocity measured from the NV $\lambda4944$ of   $300\pm10\kms$.
-The average radial velocity of the 30 Doradus region is
+\cite{bressert:12} note that VFTS682 shows a significant radial
+velocity offset compared to the nebular lines from the gas surrounding
+it, suggesting peculiar motion in the line of sight direction.
+\citet{bestenlehner:11} quote a radial velocity measured from the
+N{\footnotesize V} $\lambda4944$ of   $300\pm10\kms$. The average radial velocity of the 30 Doradus region is 
 $270\pm10\kms$.
@@
 The radial velocity errors quoted above are only statistical, and do not quantify the (possibly larger)
-systematic errors.  \SdM{TODO: Possibly mention and cite paper in prep by Paco and student on Xshooter data?  Possibly mention (or even use) on their measurement for the RV?}
-
+systematic errors.  \todo{Possibly mention and cite paper in prep by Paco and student on Xshooter data?  Possibly mention (or even use) on their measurement for the RV?}
+
+For simplicity, throughout this study, we assume the same
+distance of $50$\,kpc to the star \citep[][]{pietrzynski:13}, and to
+the 30 Doradus region as a whole. We do not consider the error on
+the distance determination ($\lesssim2\%$) when converting proper motions into
+physical velocities.
 %We implicitly assume that the use of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not introduce significant errors.
 
+\subsection{HST astrometry for VFTS682}
 % observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
@@
 
-\subsection{HST astrometry for VFTS682}
 For large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even exceed \emph{Gaia} DR2 data in quality.
 
-Recently, \citet{platais:18} presented HST astrometry and proper motions in the 30 Doradus region (GO-12499; P.I.: D.~J.~Lennon).
-The brightest stars (V$<$14) are missing from the catalogue because of saturation, but the high extinction towards VFTS682 makes is fain enough to be included in this catalogue.
+Recently, \citet{platais:18} presented HST astrometry and proper
+motions in the 30 Doradus region (GO-12499; P.I.: D.~J.~Lennon). The
+brightest stars (V$<$14) are missing from the catalogue because of
+saturation, but the high extinction towards VFTS682 makes is faint
+enough to be included. % in this catalogue.
 The  proper motions are
@@
 measure of the motion of this star completely independent from
-ours. They found proper
+\emph{Gaia}. They found proper
 motion components relative to the field
@@
 v_\mathrm{RA}^{HST}=2\pm31\kms$ and $\delta
-v_\mathrm{RA}^{HST}=47\pm24\kms$. % , corresponding to a two-dimensional
-% projected velocity of $47\pm24\kms$.
+v_\mathrm{RA}^{HST}=47\pm24\kms$% , corresponding to a two-dimensional
+% projected velocity of $47\pm24\kms$
+.
 Adding them in quadrature to the radial velocity
@@
 % component separately.
-For simplicity, throughout this study, we assume the same
-distance of $50$\,kpc to the star \citep[][]{pietrzynski:13}, and to
-the 30 Doradus region as a whole. We do not consider the error on
-the distance determination ($\lesssim2\%$) when converting proper motions into
-physical velocities. % The data retrieved, and the ipython notebook used for the analysis
-% presented here will be made available at \todo{git repo bitbucket?}.
 

### 2018-06-08 VFTS682_APJL.tex
minor bibliography
better caption
new paragraph and figure

@@
   yr^{-1}}$.
+
 % \begin{equation}
@@
 
-Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
-of the stars used in \cite{lennon:18} to define the frame of reference
-we also adopt here, cf.~the right panel in their Fig.~2. The
-proper motions are decomposed in the radial and tangential direction
-from R136, to highlight the likelihood of it being the origin of fast
-moving stars. The plus sign marks VFTS682, which sits at the edge of
-the proper motion distribution, but is not a clear outlier. Most
-relevant is the direction of the proper motion of VFTS682, which we
-discuss in \Secref{sec:r136_origin}.
-
-\begin{figure}%[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
-  \caption{Polar plot of the relative proper motion components for the stars
-    defining our reference frame. VFTS682 is indicated by the plus sign, and
-    other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
-    0.1, 0.2, and 0.4$\,\mathrm{mas\ yr^{-1}}$. Positive angle
-    indicate a tangential component pointing in the counterclockwise
-    direction, with 0 degrees corresponding to proper motion pointing
-    radially outward from R136. The proper motion for VFTS682
-    corresponds to almost radial motion away from the cluster.
-  }
-  \label{fig:pm_polar}
-\end{figure}
+% Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
+% of the stars used in \cite{lennon:18} to define the frame of reference
+% we also adopt here, cf.~the right panel in their Fig.~2. The
+% proper motions are decomposed in the radial and tangential direction
+% from R136, to highlight the likelihood of it being the origin of fast
+% moving stars. The plus sign marks VFTS682, which sits at the edge of
+% the proper motion distribution, but is not a clear outlier. Most
+% relevant is the direction of the proper motion of VFTS682, which we
+% discuss in \Secref{sec:r136_origin}.
+
+% \begin{figure}%[htbp]
+%   \centering
+%   \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
+%   \caption{Polar plot of the relative proper motion components for the stars
+%     defining our reference frame. VFTS682 is indicated by the plus sign, and
+%     other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
+%     0.1, 0.2, and 0.4$\,\mathrm{mas\ yr^{-1}}$. Positive angle
+%     indicate a tangential component pointing in the counterclockwise
+%     direction, with 0 degrees corresponding to proper motion pointing
+%     radially outward from R136. The proper motion for VFTS682
+%     corresponds to almost radial motion away from the cluster.
+%   }
+%   \label{fig:pm_polar}
+% \end{figure}
 
@@
 
+Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
+of OB type and Wolf-Rayet stars included both in the VFTS survey and
+\emph{Gaia DR2} with reliable astrometric solutions. VFTS682 is not an
+extreme outlier compared to the two O-type runaways (re-)identified by
+\cite{lennon:18}. VFTS682 was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
+the dynamical ejection scenario. The green shade in
+\Figref{fig:mu_dist} shows the range of relative proper motions within
+the uncertainties on the apparent age of the star, assuming it was
+ejected from R136 very early in its life. The proper motion
+measurements from \emph{Gaia} DR2 are consistent with the hypothesis
+of dynamical ejection.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/dist_mu_region.pdf}
+  \caption{Distribution of OB-type and Wolf-Rayet stars in proper
+    motion relative to R136. VFTS682 is not an extreme outlier, but
+    its relative proper motion matches the expected value if it were indeed
+    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The dark blue histogram contains 317 
+    stars with error smaller than $0.1\,\mathrm{mas \
+      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
+    lighter blue histogram contains 36 stars with errors smaller than $0.05\,\mathrm{mas \
+      yr^{-1}}$.}
+  \label{fig:mu_dist}
+\end{figure}
+
 The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
@@
 Although the large uncertainties on the relative proper motion
-components results in a wide range of possible directions,we argue that the most likely origin of the star is R136,
-as suggested by the N-body simulations of \cite{fujii:11}
-and \cite{banerjee:12}.
+components results in a wide range of possible directions,we argue that the most likely origin of the star is R136.
 
@@
     star. The length of the prolongations is proportional to the relative proper motion
-    times the kinetic age of VFTS682.
+    times the age of VFTS682 \citep[$1.0\pm0.2$\,Myr,][]{schneider:18}.
     % \todo{load color picture on background? Fig with HST data
@@
 
+\bibitem[\protect\citeauthoryear{{de Koter}, {Heap}  \& {Hubeny}}{{de Koter}
+  et~al.}{1997}]{dekoter:97}
+{de Koter} A.,  {Heap} S.~R.,   {Hubeny} I.,  1997, \mn@doi [\apj]
+  {10.1086/303736}, \href {http://adsabs.harvard.edu/abs/1997ApJ...477..792D}
+  {477, 792}
+
 \bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
@@
   {http://adsabs.harvard.edu/abs/1957ZA.....44...64Z} {44, 64}
-
-\bibitem[\protect\citeauthoryear{{de Koter}, {Heap}  \& {Hubeny}}{{de Koter}
-  et~al.}{1997}]{dekoter:97}
-{de Koter} A.,  {Heap} S.~R.,   {Hubeny} I.,  1997, \mn@doi [\apj]
-  {10.1086/303736}, \href {http://adsabs.harvard.edu/abs/1997ApJ...477..792D}
-  {477, 792}
 

### 2018-06-09 VFTS682_APJL.tex

@@
     lighter blue histogram contains 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$.}
+      yr^{-1}}$. The top axis shows the conversion to physical units
+    assuming a distance of 50\,kpc.}
   \label{fig:mu_dist}

### 2018-06-09 VFTS682_APJL.tex
minor
new fig.1 with thicker hatches and table
added figure on angles and text: the fig needs to be redone with errors properly

--- 2018-06-08 VFTS682_APJL.tex
@@
   \begin{center}
-    \caption{Astrometric parameters for VFTS682. }
+    \caption{Observed and derived parameters for VFTS682. We also list
+    the average proper motion components of R136 derived by
+    \cite{lennon:18} for reference.}
     \begin{tabular}{l|c|c}
@@
       \hline
+      present day $M$ \hfill $[M_\odot]$ & $137.8^{+27.5}_
+                                           {-15.9}$ &
+                                                     \multirow{3}{*}{\cite{schneider:18}}\\
+      $M_\mathrm{ZAMS}$\hfill $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ &
+      \\
+      age \hfill [Myr] & $1.0\pm0.2$ & \\
+      \hline
+      $\langle\mu_\mathrm{RA}\rangle$\hfill[$\mathrm{mas\ yr^{-1}}$] & $1.74\pm0.01$
+                        &  \multirow{2}{*}{\cite{lennon:18}}\\[5pt]
+      $\langle\mu_\mathrm{DEC}\rangle$\hfill[$\mathrm{mas\ yr^{-1}}$]
+                & $0.70\pm0.02$ & \\
+      \hline
+
     \end{tabular}
@@
 \end{table}
+
+% \begin{table}
+%   \begin{center}
+%     \caption{Astrometric parameters for VFTS682. }
+%     \begin{tabular}{l|c|c}
+%       \hline
+%       \hline
+%       Parameter & Value & Source\\
+%       \hline
+%       RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
+%                         & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
+%       DEC \hfill [degree] & -69.07 % $\pm$  0.05
+%                         & \\[5pt]
+%       \hline
+%       $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \multirow{2}{*}{\emph{Gaia} DR2}\\[5pt]
+%       $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
+%       \hline
+%       $\delta\mu_\mathrm{RA}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.01$\pm$0.13 & \multirow{2}{*}{\cite{platais:18}}\\[5pt]
+%       $\delta\mu_\mathrm{DEC}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] &
+%                                                                         0.2$\pm$0.1 &
+%       \\[5pt]
+%       \hline
+%       $\delta v_\mathrm{rad}$  \hfill[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
+%       \hline
+%     \end{tabular}
+%     \tablecomments
+%     {The error on the RA and DEC positions, are of order
+%       $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$ in \emph{Gaia}
+%       DR2. %REPETITION: %  We adopt the peculiar
+%       % radial velocity calculated using the  NV $\lambda4944$ line, which is
+%       % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
+%     }
+%   \end{center}
+%   \label{tab:vfts682}
+% \end{table}
 
@@
 
-Assuming that the best estimate of the peculiar velocity of VFTS682
-are reliable, we now address the question of its likely origin.
+% Assuming that the best estimate of the peculiar velocity of VFTS682
+% are reliable, we now address the question of its likely origin.
 
 Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
-of OB type and Wolf-Rayet stars included both in the VFTS survey and
-\emph{Gaia DR2} with reliable astrometric solutions. VFTS682 is not an
-extreme outlier compared to the two O-type runaways (re-)identified by
-\cite{lennon:18}. VFTS682 was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
+of OB-type and Wolf-Rayet stars included both in the VFTS survey and
+\emph{Gaia DR2} with reliable astrometric solutions. The bottom panel
+shows the distribution in angles $\theta$ between the relative proper motion
+direction and the direction from the core of R136 to the star. We
+emphasize that our subset of stars with reliable proper motion
+measurements is biased towards the fast moving objects, which results in a
+distribution of angles mildly peaked at small angles.
+
+VFTS682 is not an outlier in relative proper motion%  compared
+% to the two O-type runaways (re-)identified by \cite{lennon:18} (see
+% also below)
+. However, the star was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
 the dynamical ejection scenario. The green shade in
-\Figref{fig:mu_dist} shows the range of relative proper motions within
+\Figref{fig:mu_dist} shows the range of relative proper motions
+required to reach the present day location within
 the uncertainties on the apparent age of the star, assuming it was
-ejected from R136 very early in its life. The proper motion
-measurements from \emph{Gaia} DR2 are consistent with the hypothesis
+ejected from R136 very early in its life.
+
+Because of the large
+error bars, the angle between the relative proper motion is not very
+constraining, but it is suggestive that the best value is close to zero.
+Therefore, the relative proper motion from \emph{Gaia} DR2 are consistent with the hypothesis
 of dynamical ejection.
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{figures/dist_mu_region.pdf}
-  \caption{Distribution of OB-type and Wolf-Rayet stars in proper
-    motion relative to R136. VFTS682 is not an extreme outlier, but
+  \includegraphics[width=0.5\textwidth]{figures/dist_mu_region.pdf}\\
+  \includegraphics[width=0.5\textwidth]{figures/angle_dist}
+  \caption{Top panel: distribution of OB-type and Wolf-Rayet stars in proper
+    motion relative to R136. VFTS682 is not an outlier, but
     its relative proper motion matches the expected value if it were indeed
-    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The dark blue histogram contains 317 
+    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
+    assuming a distance of 50\,kpc. Bottom panel: \todo{redo with
+      errorbars} distribution of
+    angles between the relative proper motion direction and the radial
+    direction to the star. The error bars for VFTS682 are large, but
+    the best value is in agreement with the hypothesis of dynamical
+    ejection. In both
+    panels, the dark blue histograms contain 317 
     stars with error smaller than $0.1\,\mathrm{mas \
       yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
-    lighter blue histogram contains 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$.}
+    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
+      yr^{-1}}$. }
   \label{fig:mu_dist}
@@
 
-Although the large uncertainties on the relative proper motion
-components results in a wide range of possible directions,we argue that the most likely origin of the star is R136.
-
-The kinematic age of this star, assuming it originates from the
-cluster, is
+% Although the large uncertainties on the relative proper motion
+% components results in a wide range of possible directions,we argue that the most likely origin of the star is R136.
+
+The kinematic age of this star, assuming it originates from R136, is
 
@@
 where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
-the core of R136 \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
+the core of the cluster \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
 As in the rest of this study, we neglect for
 simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
-
 The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
@@
 % ejection very soon in the cluster evolution.
-Note that the
-present-day surface helium abundance
-\citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
-limit on the age of the star of $\sim$0.9\,Myr, corresponding to the time needed to
-synthesize this amount of helium in the models from \cite{kohler:15}.
-% \todo{discuss position angle}
+
+% the thing below is not relevant if it is a merger, better cut out
+% for now, since we are too long anyway
+
+%Note that the
+% present-day surface helium abundance
+% \citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
+% limit on the age of the star of $\sim$0.9\,Myr, corresponding to the time needed to
+% synthesize this amount of helium in the models from \cite{kohler:15}.
 
@@
 
+\todo{maybe move around fig. 1}
 \citet{lennon:18} carried out a study similar to ours on the fast

### 2018-06-09 _region_.tex

--- 2018-06-09 VFTS682_APJL.tex
@@
-\documentclass[apjl,twocolumn]{emulateapj}
+\message{ !name(VFTS682_APJL.tex)}\documentclass[apjl,twocolumn]{emulateapj}
 \usepackage{graphicx}
@@
 
-% \title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
-%                                 % ejected %% to fit in one line
-% runaway star known to date}
-\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
+\message{ !name(VFTS682_APJL.tex) !offset(490) }
 
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
-  V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{1,7,8}$,  A.~de~Koter$^{1}$,
-  N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
-\affil{$^{1}${Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
-  $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
-    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
-  $^{3}$ {Space Telescope Science Institute, 3700 San Martin Drive,
-    Baltimore, MD 21218, USA}\\
-  $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-  $^{5}${Department of Physics and Astronomy, Hicks Building,
-    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
-  $^{6}$ {National Research Council, Herzberg Astronomy \&
-    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
-    Canada}\\
-  $^{7}$ {School of Astronomy \& Space Science, University of the Chinese
-    Academy of Sciences, Beijing 100012, China}\\
-  $^{8}$ {National Astronomical Observatories, Chinese Academy of
-    Sciences, Beijing 100012, China}\\
-  $^{9}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
-    Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
-  $^{10}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
-  $^{11}$ {Department of Physics, University of Oxford, Keble Road,
-    Oxford OX1 3RH, UK} \\
-  $^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
-}
-
- \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
-\date{}
-\begin{abstract}
-
- How very massive stars form is still among the most intriguing open
- questions in stellar astrophysics.  VFTS 682 is among the most
- extreme massive stars known, with an inferred initial mass of about
- $150\,M_\odot$.
-It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.
-The absence of other massive stars in its immediate vicinity lead to two intriguing possible hypotheses posed in earlier work.
-Either it formed in relative isolation through a new mode of star formation or it was ejected dynamically from the central cluster.
-
-Aiming to shine light on this debate, we investigate the kinematics of  VFTS 682 as obtained by \emph{Gaia} and a \emph{Hubble Space Telescope} proper motion campaign.
-We derive a projected velocity relative of to the cluster of $\sim$$30\pm20\kms$.
-The direction and magnitude of the proper motion are consistent with ejection from the central cluster considering its inferred age.
-However, the error bars and spread in proper motions derived for other sources in the region are substantial, which makes us cautious to derive hard conclusions.
-
-If future data confirms the runaway nature, this would make the VFTS682 the most massive runaway star known to date. 
-While we cannot prove this solidly from the current data, we do
-consider this hypothesis plausible. The central cluster is known to
-harbor several other stars more massive than $150\,M_\odot$ similar in
-spectral type.  Moreover, the current record holders are O type stars
-including VFTS 16 which provides direct evidence that R136 is indeed
-capable of ejecting some of its most massive members. 
-\end{abstract}
-
-\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS 682}
-\maketitle{}
-
-\section{Introduction}
-\label{sec:intro}
-
-How massive stars form is one of the major longstanding questions in astrophysics
-\citep[e.g.,][]{zinnecker:07}. Improving our understanding of massive star formation, and its
-possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies, but also for understanding the transients that mark their death and the compact remnants they leave behind.
-%
-Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
-% \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
-evolve fast, typically reside in dense groups and remain enshrouded in
-their parent cloud during the entirety of their formation
-process. Important progress has been made on the theoretical side,
-\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally very expensive and therefore remain challenging.
-
-In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to reside in the innermost cores.
-In this picture, field stars are primarily the result of the dissolution of dense groups.
-However, a significant population of massive stars exists in relative isolation, far from dense clusters or OB associations and their origin remains matter of debate \citep{gvaramadze:12, lamb:16,ward:18}.
-One hypothesis to explain the population of relatively isolated massive stars is that they formed in the field.
-The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
-Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.
-
-% A contribution to the debate on whether or not massive
-% stars form in relative isolation was presented by
-% \cite{bestenlehner:11} and \cite{bressert:12}, who discussed the case of the very massive star
-% VFTS682.
-
-One of the most extreme examples that has been considered in this debate is the very massive star VFTS682  \citep[][]{bestenlehner:11, bressert:12}.
-This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}.
-It is a hydrogen-rich Wolf-Rayet star of spectral type WNh5.
-Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $\sim$$140^{+30}_{-16}\,M_\odot$ corresponding to an initial mass of $\sim$$150^{+30}_{-17}\,M_\odot$
-\citep{schneider:18}. 
-%
-%$137.8^{+27.5}_{-15.9}\,M_\odot$,
-%
-This makes VFTS682 one of the most massive stars known and one of the most extreme objects in the region.
-%
-From the spectral point of view, it is reminiscent of the very
-massive stars % , a.k.a. ``monster stars'',
- in the core of the R136 cluster \citep{dekoter:97,crowther:10, crowther:16}. 
- %
- In particular, a remarkable similarity exist between the
-spectrum of VFTS682 and R136a3 for which \citet{crowther:16} report a
-current mass estimate of $180^{+30}_{-30}\Msun$. R136 hosts
-at least two more very massive WN5h stars, R136a1 and R136a2, whose
-estimated current masses are even higher.
-
-%at a projected offset of 119.4 arcseconds, corresponding to
-
-VFTS682 stands out by its relative isolation at a projected distance of 119.4 arcseconds, corresponding to 
-$\sim$$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
-considered two possible explanation for the offset: either
-the star formed in situ as an isolated massive star, % , but remark that this
-                                % poses a challenge for starformation
-                                % theories (They don't)
-or it was ejected from  R136. N-body simulations % of dynamical interaction in star clusters
-indicate that the ejection of very massive stars like VFTS682 is expected \citep[e.g.][]{fujii:11, banerjee:12}.
-This is supported by the recent findings of other massive runaway stars in the region based on proper motion studies.
-\citet{platais:15,platais:18} analyze multi-epoch \emph{Hubble Space Telescope} (HST) photometry and identify 10 stars, which appear to be ejected from R136.
-\citet{lennon:18} investigate the kinematics of  isolated O-type stars in the region using the second \emph{Gaia} data release \cite[DR2,][]{gaia:16,brown:18} and show that the proper motion, postion and direction of the $\sim100\Msun$ star VFTS16 is consistent with a runaway origin from R136.
-
-%Testing the origin of VFTS682 is not only important for constraining isolated star formation; it should also help to improve understanding of early star-cluster dynamics.
-
-In this paper we present an analysis of the new kinematical
-constraints for VFTS682 provided by \emph{Gaia} DR2 and constraints
-from HST proper data  by \citet{platais:18}.   We discuss the
-implications for the hypothesis that VFTS682 is a slow runaway star
-ejected from R136.
-% which was not part of the sample of \citet{lennon:18} because of its  spectral type.
-
-%%% cut for length reasons: all is said below
-% Our results indicate that VFTS682 is a runaway star, although with
-% low statistical significance. The
-% direction and magnitude of the velocity vector are consistent with
-% dynamical ejection from R136. The age inferred from the spectral
-% analysis \citep[from][]{schneider:18} is consistent with the travel
-% time we calculate for this star. % Therefore, the hypothesis that VFTS
-% % 682 is formed in relative isolation is rejected.
-% If our results are confirmed by future astrometric data releases, VFTS
-% 682, with an inferred present day mass well above a hundred solar
-% masses, is the most massive runaway star known to date.
-
-\section{Observations}
-\label{sec:sample}
-
-\subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
-
-The star VFTS682, located at right ascension (RA)
-05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000 \citep[][% see also
-% \Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2
-]{evans:11}
-was originally classified as a young stellar object based on its
-mid-infrared excess \citep{gruendl:09}. \citet{evans:11} reclassified the
-object as Wolf-Rayet star of spectral type WNh5 using multi-epoch
-spectra covering $\lambda$4000--7000 from the
-VFTS survey. % The spectra available covered $\lambda$4000--7000 and
-% were taken at multiple epochs.
-
-Using the same dataset,
-\citet{bestenlehner:11} excluded the presence of a close companion
-with high confidence from the absence of radial velocity variations.
-\citet{bestenlehner:11} % further analyzed the spectra % using the non-LTE
-% % model atmosphere code CMFGEN \citep{hillier:98}
-also derived stellar
-parameters and surface abundances. They inferred an extinction
-of $A_V=4.45\pm0.12$, leading to the high luminosity
-$\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citet{schneider:18} estimated
-a present-day mass for VFTS682 of $137.8^{+27.5}_ {-15.9}\Msun$, an
-apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
-of $150.0^{+28.7}_{-17.4}\Msun$.% , using the Bayesian analysis tool BONNSAI
-% \citep{schneider:17}, based on the evolutionary models from
-% \citet{brott:11} and \cite{kohler:15}.
-% These values place VFTS682 around the
-% ``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
-
-\citet{bestenlehner:11} reported for VFTS682 a wind mass loss rate of
-$10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, % inferred from
-% spectral analysis assuming a smooth outflow, i.e.,
-not accounting for
-clumping. The present-day surface helium mass
-fraction is $Y=0.45\, (0.49)$ from the spectral analysis of
-\cite{bestenlehner:11} (\citealt{rubio-diez:17}). This value can be
-used to constrain the age of the star and its past evolution.
-
-% The star is very similar from the spectral point of view to R136a3, located inside the star cluster
-% R136 \citep{crowther:10}  for which \citet{crowther:16} report a
-% current mass estimate of $180^{+30}_{-30}\Msun$. The R136 cluster hosts
-% at least two more very massive WN5h stars, R136a1 and R136a2, whose
-% estimated current masses are even higher. However, VFTS682 stands
-% out by its isolation at a projected offset of 29\,pc from the center
-% of the cluster \citep{bestenlehner:11}.
-
-Further worth noticing is the variability of the
-star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
-lower than found by \citet{evans:11}.
-\citet{bestenlehner:11} discuss the Optical Gravitational Lensing
-Experiment (OGLE-III) light curves \citep{udalski:08} and show a
-variability in the V-band at $\sim$10\% level on a timescale of years.
-They commented that this is unusual for Wolf-Rayet stars and more reminiscent
-of Luminous Blue Variable (LBV) stars. % \citep[e.g.,][]{langer:94}.
-
-\cite{bressert:12} note that VFTS682 shows a significant radial
-velocity offset compared to the nebular lines from the gas surrounding
-it, suggesting peculiar motion in the line of sight direction.
-\citet{bestenlehner:11} quote a radial velocity measured from the
-N{\footnotesize V} $\lambda4944$ of   $300\pm10\kms$. The average radial velocity of the 30 Doradus region is 
-$270\pm10\kms$.
-Whether this measurement truly reflects a peculiar radial velocity is debated.
-The determination of radial velocities for emission-lines stars is difficult because of the presence of an optically thick, non-homogeneous, and variable wind obscuring the photosphere.
-The radial velocity errors quoted above are only statistical, and do not quantify the (possibly larger)
-systematic errors.  \todo{Possibly mention and cite paper in prep by Paco and student on Xshooter data?  Possibly mention (or even use) on their measurement for the RV?}
-
-For simplicity, throughout this study, we assume the same
-distance of $50$\,kpc to the star \citep[][]{pietrzynski:13}, and to
-the 30 Doradus region as a whole. We do not consider the error on
-the distance determination ($\lesssim2\%$) when converting proper motions into
-physical velocities.
-%We implicitly assume that the use of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not introduce significant errors.
-
-\subsection{HST astrometry for VFTS682}
-% observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
-% in October 3-8 2011 and the second epoch (GO-13359) in
-% October 6-11 2014, both using the F775W filter.
-% The second epoch was designed to match the pointing and
-% orientation of the first epoch as exactly as possible so
-% that the displacement measurements could be purely differential.
-
-For large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even exceed \emph{Gaia} DR2 data in quality.
-
-Recently, \citet{platais:18} presented HST astrometry and proper
-motions in the 30 Doradus region (GO-12499; P.I.: D.~J.~Lennon). The
-brightest stars (V$<$14) are missing from the catalogue because of
-saturation, but the high extinction towards VFTS682 makes is faint
-enough to be included. % in this catalogue.
-The  proper motions are
-listed in \Tabref{tab:vfts682}.
-
-The catalog from \citet{platais:18} lists the proper motion
-components of VFTS682 relative to the field, providing a
-measure of the motion of this star completely independent from
-\emph{Gaia}. They found proper
-motion components relative to the field
-$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
-  yr^{-1}}$ and
-$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
-  yr^{-1}}$, which added in quadrature give a relative proper motion of
-$\delta \mu^{HST} =0.20 \pm 0.15\,\mathrm{mas\ yr^{-1}}$.
-% \begin{equation}
-%   \label{eq:pm_around_HST}
-%   \delta \mu^{HST} = \sqrt{\left(\delta\mu_\mathrm{RA}^{HST}\right)^2+\left(\delta\mu_\mathrm{DEC}^{HST}\right)^2}
-%   = 0.20 \pm 0.15\,\mathrm{mas\
-%   yr^{-1}} \ \ .
-% \end{equation}
-
-These proper motion components can also be converted into projected
-velocities assuming a distance of 50\,kpc, obtaining $\delta
-v_\mathrm{RA}^{HST}=2\pm31\kms$ and $\delta
-v_\mathrm{RA}^{HST}=47\pm24\kms$% , corresponding to a two-dimensional
-% projected velocity of $47\pm24\kms$
-.
-Adding them in quadrature to the radial velocity
-measured by \citet{bestenlehner:11} we obtain a three-dimensional velocity of
-$v_\mathrm{pec}^{HST}=56 \pm 23  \kms$.
-% \begin{equation}
-%   \label{eq:speed_around_HST}
-%   v_\mathrm{pec}^{HST} = \sqrt{\left(\delta v_\mathrm{RA}^{HST}\right)^2
-%     +\left(\delta v_\mathrm{DEC}^{HST}\right)^2+\left(\delta
-%       v_\mathrm{rad}^{VFTS}\right)^2} = 56 \pm 23
-%   \kms \ .
-% \end{equation}
-
-\subsection{New data from \emph{Gaia} DR2  \label{data:gaia}}
-
-\begin{table}
-  \begin{center}
-    \caption{Astrometric parameters for VFTS682. }
-    \begin{tabular}{l|c|c}
-      \hline
-      \hline
-      Parameter & Value & Source\\
-      \hline
-      RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
-                        & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
-      DEC \hfill [degree] & -69.07 % $\pm$  0.05
-                        & \\[5pt]
-      \hline
-      $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \multirow{2}{*}{\emph{Gaia} DR2}\\[5pt]
-      $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
-      \hline
-      $\delta\mu_\mathrm{RA}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.01$\pm$0.13 & \multirow{2}{*}{\cite{platais:18}}\\[5pt]
-      $\delta\mu_\mathrm{DEC}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] &
-                                                                        0.2$\pm$0.1 &
-      \\[5pt]
-      \hline
-      $\delta v_\mathrm{rad}$  \hfill[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
-      \hline
-    \end{tabular}
-    \tablecomments
-    {The error on the RA and DEC positions, are of order
-      $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$ in \emph{Gaia}
-      DR2. %REPETITION: %  We adopt the peculiar
-      % radial velocity calculated using the  NV $\lambda4944$ line, which is
-      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
-    }
-  \end{center}
-  \label{tab:vfts682}
-\end{table}
-
-The second \emph{Gaia} Data Release (DR2) % , which
-% became available on 25 April 2018 and 
-provides positions,
-parallaxes, and proper motion components for more than a billion
-sources \citep{brown:18}. VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2
-catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Its
-G-band magnitude from \emph{Gaia}
-% reports a G-band magnitude of
-is 15.65. The star has a
-\texttt{visibility\_period} = 17, which counts how many observations have
-been used to reconstruct its astrometric solution
-\citep[][]{lindengren:18}, and the reported
-\texttt{astrometric\_excess\_noise} = 0. These values suggest that the \emph{Gaia}
-data for VFTS682 are reliable. % However, the effective temperature
-% reported in \emph{Gaia} DR2, based on the spectral fitting around the
-% CaII triplet, is one order of magnitude lower than what found by
-% \cite{bestenlehner:11}, and the best fit parallax of this star is
-% negative \citep[see, e.g.,][]{hogg:18}. We do not use the effective temperature of the star anywhere
-% in this study, and we attribute the unphysical value of the parallax
-% to the large distance to the LMC.
-% Our main findings do not rely on the
-% parallax nor the effective temperature values reported in the \emph{Gaia} DR2
-% catalog.
-
-While \emph{Gaia} DR2 data can be correlated, this is
-not an issue for VFTS682, and we treat the proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
-respectively) of the star as uncorrelated for simplicity.
-For the radial velocity of VFTS682 and of the 30 Doradus
-region as a whole, we instead use the VFTS data
-as quoted in \cite{bestenlehner:11}. \Tabref{tab:vfts682} lists the values adopted throughout
-this work for each of these quantities.
-
-We follow \citet{lennon:18} to define a local frame of reference and derive the peculiar velocity
-of VFTS682 with respect to the cluster R136. Their selection (see
-their Sec.~2.1) of 153
-bright ($G<17$) stars around R136 with reliable astrometric data from
-\emph{Gaia} DR2 yields $\langle\mu_\mathrm{RA}\rangle=1.74\pm0.01\,\mathrm{mas\
-  yr^{-1}}$ and $\langle\mu_\mathrm{DEC}\rangle=0.70\pm0.02\,\mathrm{mas\ yr^{-1}}$
-for the components of the mean motion of the region projected on the sky.
-
-% We define a local standard frame of reference to derive the peculiar velocity
-% of VFTS682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
-% We select all the stars in a target of 0.05 degrees around R136
-% (NGC2070) fulfilling the following criteria. First, we require a G-band
-% magnitude brighter than 17, corresponding roughly to the
-% completeness level of the VFTS survey \citep[here we implicitly assume
-% G$\sim$V,][]{evans:11}. Then we require \texttt{visibility\_period} $\geq$ 5,
-% \texttt{astrometric\_excess\_noise} $< 1$, and the errors on the proper
-% motion components to be smaller than 0.07\,$\mathrm{mas\
-% yr^{-1}}$. At the distance to the LMC, $0.1\mathrm{mas\
-% yr^{-1}}\simeq25\,\kms$ \citep[e.g.,][]{platais:18}. We
-% remove obvious foreground stars by requiring the absolute value of
-% the reported parallaxes smaller than $0.1\,\mathrm{mas\ yr^{-1}}$.
-% At the distance to the
-% LMC, $1\mathrm{mas\ yr^{-1}}\simeq250\,\kms$ \citep[e.g.,][]{lennon:18}, so the cut on the values
-% of the proper motions removes stars that would have projected
-% tangential velocities in excess of $\sim$500\,$\kms$, which are most
-% likely to be foreground stars. We checked that the additional
-% requirement of having parallaxes smaller than $2\,\mathrm{mas}$ does
-% not reduces further our sample.
-% This selection yields a sample of 66 stars.% ,
-
-% We calculate the averaged proper motion components for the whole
-% region using 
-% \begin{equation}
-%   \label{eq:mean}
-%   \langle \mu_i\rangle = \frac{\sum_\mathrm{stars}\frac{1}{\Delta
-%   \mu_i^2}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i^2}} \ \ , \
-%   \ \Delta \langle \mu_i\rangle = \frac{\sqrt{N}}{\sum_\mathrm{stars}
-%   \frac{1}{\Delta \mu_i^2}} \ \ ,
-% \end{equation}
-% where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
-% on the proper motion component reported by \emph{Gaia}. The sums run over
-% all the $N=66$ stars in our selected sample. We evaluate each proper motion
-% component separately.
-
-\section{The kinematics of VFTS682}
-\label{sec:results}
-
-\subsection{Is it a runaway star?}
-\label{sec:runaway}
-We first address the question of whether VFTS682 is a typical star
-from the kinematic point of view, or whether it is a runaway star with
-a significantly large peculiar velocity compared to its surrounding population. The former is what should
-be expected if it formed in the relative isolation that we observe today.% , in relative
-% isolation from other massive stars.
-
-% Using the stars selected as described in \Secref{data:gaia} (a
-% subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
-% $\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
-% $\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
-% yr^{-1}}$. These values are
-% in good agreement with what is found by \cite{lennon:18}. However, we
-% emphasize their sensitivity to the choice of stars adopted to define
-% the reference frame.
-Subtracting the mean proper motion components given by
-\citet{lennon:18} from the
-proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
-components of proper motion of the star relative to the surrounding region
-$\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\,\mathrm{mas\ yr^{-1}}$
-and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08
-\pm 0.09\,\mathrm{mas\ yr^{-1}}$. These components result in a
-two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\
-  yr^{-1}}$.
-
-% \begin{equation}
-%   \label{eq:pm_gaia_around}
-%   \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
-%   = 0.13\pm 0.09\,\mathrm{mas\
-%   yr^{-1}} \ \ .
-% \end{equation}
-
-% Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
-% of the stars used in \cite{lennon:18} to define the frame of reference
-% we also adopt here, cf.~the right panel in their Fig.~2. The
-% proper motions are decomposed in the radial and tangential direction
-% from R136, to highlight the likelihood of it being the origin of fast
-% moving stars. The plus sign marks VFTS682, which sits at the edge of
-% the proper motion distribution, but is not a clear outlier. Most
-% relevant is the direction of the proper motion of VFTS682, which we
-% discuss in \Secref{sec:r136_origin}.
-
-% \begin{figure}%[htbp]
-%   \centering
-%   \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
-%   \caption{Polar plot of the relative proper motion components for the stars
-%     defining our reference frame. VFTS682 is indicated by the plus sign, and
-%     other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
-%     0.1, 0.2, and 0.4$\,\mathrm{mas\ yr^{-1}}$. Positive angle
-%     indicate a tangential component pointing in the counterclockwise
-%     direction, with 0 degrees corresponding to proper motion pointing
-%     radially outward from R136. The proper motion for VFTS682
-%     corresponds to almost radial motion away from the cluster.
-%   }
-%   \label{fig:pm_polar}
-% \end{figure}
-
-The proper motion components can be converted into
-the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=24\pm19\,\kms$,
-$\delta v_\mathrm{DEC}^{Gaia}=20\pm23\,\kms$, assuming a distance of
-50\,kpc. These can be combined obtaining a projected two-dimensional
-velocity of $31\pm21\kms$. % (we do not account for the uncertainty in the distance
-% estimate when propagating errors).
-The radial velocity from
-\cite{bestenlehner:11} then gives the third component along
-the line of sight, % allowing us to calculate the three-dimensional
-% peculiar speed of the star:
-which added in quadrature to the transverse components results in a
-three-dimensional velocity of $44 \pm 21\kms$.
-% \begin{equation}
-%   \label{eq:speed_around_Gaia}
-%   v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
-%     +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
-%       v_\mathrm{rad}^{VFTS}\right)^2} = 44 \pm 21
-%   \kms \ .
-% \end{equation}
-
-Therefore, two completely independent measures of the proper motion of
-VFTS682 relative to the surrounding field, one from HST and one from
-\emph{Gaia} DR2, yield values of the peculiar three-dimensional
-velocity of VFTS682 % (\Eqref{eq:speed_around_HST} and
-% \Eqref{eq:speed_around_Gaia}, respectively)
-which would make it the most massive runaway star
-known to date. However, the large errors on both the proper motion measures
-require confirmation with future astrometric data.
-
-\subsection{Does it come from the R136 cluster?}
-\label{sec:r136_origin}
-
-Assuming that the best estimate of the peculiar velocity of VFTS682
-are reliable, we now address the question of its likely origin.
-
-Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
-of OB type and Wolf-Rayet stars included both in the VFTS survey and
-\emph{Gaia DR2} with reliable astrometric solutions. VFTS682 is not an
-extreme outlier compared to the two O-type runaways (re-)identified by
-\cite{lennon:18}. VFTS682 was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
-the dynamical ejection scenario. The green shade in
-\Figref{fig:mu_dist} shows the range of relative proper motions within
-the uncertainties on the apparent age of the star, assuming it was
-ejected from R136 very early in its life. The proper motion
-measurements from \emph{Gaia} DR2 are consistent with the hypothesis
-of dynamical ejection.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figures/dist_mu_region.pdf}
-  \caption{Distribution of OB-type and Wolf-Rayet stars in proper
-    motion relative to R136. VFTS682 is not an extreme outlier, but
-    its relative proper motion matches the expected value if it were indeed
-    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The dark blue histogram contains 317 
-    stars with error smaller than $0.1\,\mathrm{mas \
-      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
     lighter blue histogram contains 36 stars with errors smaller than $0.05\,\mathrm{mas \
@@
 
-\citet{lennon:18} carried out a study similar to ours on the fast
-moving O-type stars
-in the region, and found two massive runaway stars
-($\sim$$90\,M_\odot$) in the 30 Doradus region. One of them (VFTS 16)
-was previously known as a runaway star from its line of sight velocity
-\citep[][]{evans:10}. \citet{lennon:18} also concluded that VFTS16 is 
-the result of a dynamical ejection from the R136 cluster, while the
-origin of the other star (VFTS 72) is less clear given its direction
-of motion. The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
-VFTS682 (see \Secref{sec:r136_origin}) is smaller than the
-corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic
-age of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
-massive VFTS682 was ejected later than VFTS16 from the same cluster.
-
-The numerical simulations from \cite{oh:16} suggest that dynamical
-interaction eject the majority of the stars during or shortly after the cluster
-core-collapse. The large number of isolated massive stars around it
-suggest that R136 has already evolved past the
-time of maximum stellar density. This might have implications for the
-question of whether the cluster formed via a monolithic collapse, or
-as a (potentially ongoing) merger of several sub-structures \citep[e.g.,][]{sabbi:12}.
-
-\citet{oh:16} also showed that the mass and velocity distribution of the ejected star depends on the cluster initial conditions
-(whether it is segregated, its primordial binary fraction and initial period
-distribution of the binary population), therefore studies on
-the population of isolated massive stars in the surroundings of R136
-might shed light on its initial stellar population and dynamical
-state.
-
-VFTS682 is potentially the most massive runaway known to date, and its ejection
-from the cluster R136 likely implies that it is only the ``tip of the
-iceberg'' of possibly extremely massive runaways in the
-region. Studies of this population, enabled by recent HST and \emph{Gaia} observations will put constraints on the evolution
-of these extreme stars, together with the formation and evolution of
-the central cluster itself.
-
-% \bibliographystyle{aa}
-% \bibliography{bibliography/vfts682}
-\begin{thebibliography}{}
-\makeatletter
-\relax
-\def\mn@urlcharsother{\let\do\@makeother \do\$\do\&\do\#\do\^\do\_\do\%\do\~}
-\def\mn@doi{\begingroup\mn@urlcharsother \@ifnextchar [ {\mn@doi@}
-  {\mn@doi@[]}}
-\def\mn@doi@[#1]#2{\def\@tempa{#1}\ifx\@tempa\@empty \href
-  {http://dx.doi.org/#2} {doi:#2}\else \href {http://dx.doi.org/#2} {#1}\fi
-  \endgroup}
-\def\mn@eprint#1#2{\mn@eprint@#1:#2::\@nil}
-\def\mn@eprint@arXiv#1{\href {http://arxiv.org/abs/#1} {{\tt arXiv:#1}}}
-\def\mn@eprint@dblp#1{\href {http://dblp.uni-trier.de/rec/bibtex/#1.xml}
-  {dblp:#1}}
-\def\mn@eprint@#1:#2:#3:#4\@nil{\def\@tempa {#1}\def\@tempb {#2}\def\@tempc
-  {#3}\ifx \@tempc \@empty \let \@tempc \@tempb \let \@tempb \@tempa \fi \ifx
-  \@tempb \@empty \def\@tempb {arXiv}\fi \@ifundefined
-  {mn@eprint@\@tempb}{\@tempb:\@tempc}{\expandafter \expandafter \csname
-  mn@eprint@\@tempb\endcsname \expandafter{\@tempc}}}
-
-\bibitem[\protect\citeauthoryear{{Banerjee}, {Kroupa}  \& {Oh}}{{Banerjee}
-  et~al.}{2012}]{banerjee:12}
-{Banerjee} S.,  {Kroupa} P.,   {Oh} S.,  2012, \mn@doi [\apj]
-  {10.1088/0004-637X/746/1/15}, \href
-  {http://adsabs.harvard.edu/abs/2012ApJ...746...15B} {746, 15}
-
-\bibitem[\protect\citeauthoryear{{Bate}}{{Bate}}{2009}]{bate:09}
-{Bate} M.~R.,  2009, \mn@doi [\mnras] {10.1111/j.1365-2966.2008.14106.x}, \href
-  {http://adsabs.harvard.edu/abs/2009MNRAS.392..590B} {392, 590}
-
-\bibitem[\protect\citeauthoryear{{Bestenlehner} et~al.,}{{Bestenlehner}
-  et~al.}{2011}]{bestenlehner:11}
-{Bestenlehner} J.~M.,  et~al., 2011, \mn@doi [\aap]
-  {10.1051/0004-6361/201117043}, \href
-  {http://adsabs.harvard.edu/abs/2011A\%26A...530L..14B} {530, L14}
-
-\bibitem[\protect\citeauthoryear{{Blaauw}}{{Blaauw}}{1961}]{blaauw:61}
-{Blaauw} A.,  1961, \bain, \href
-  {http://adsabs.harvard.edu/abs/1961BAN....15..265B} {15, 265}
-
-\bibitem[\protect\citeauthoryear{{Bressert} et~al.,}{{Bressert}
-  et~al.}{2012}]{bressert:12}
-{Bressert} E.,  et~al., 2012, \mn@doi [\aap] {10.1051/0004-6361/201117247},
-  \href {http://adsabs.harvard.edu/abs/2012A\%26A...542A..49B} {542, A49}
-
-\bibitem[\protect\citeauthoryear{{Brott} et~al.,}{{Brott}
-  et~al.}{2011}]{brott:11}
-{Brott} I.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201016113}, \href
-  {http://adsabs.harvard.edu/abs/2011A\%26A...530A.115B} {530, A115}
-
-\bibitem[\protect\citeauthoryear{{Crowther}, {Schnurr}, {Hirschi}, {Yusof},
-  {Parker}, {Goodwin}  \& {Kassim}}{{Crowther} et~al.}{2010}]{crowther:10}
-{Crowther} P.~A.,  {Schnurr} O.,  {Hirschi} R.,  {Yusof} N.,  {Parker} R.~J.,
-  {Goodwin} S.~P.,   {Kassim} H.~A.,  2010, \mn@doi [\mnras]
-  {10.1111/j.1365-2966.2010.17167.x}, \href
-  {http://adsabs.harvard.edu/abs/2010MNRAS.408..731C} {408, 731}
-
-\bibitem[\protect\citeauthoryear{{Crowther} et~al.,}{{Crowther}
-  et~al.}{2016}]{crowther:16}
-{Crowther} P.~A.,  et~al., 2016, \mn@doi [\mnras] {10.1093/mnras/stw273}, \href
-  {http://adsabs.harvard.edu/abs/2016MNRAS.458..624C} {458, 624}
-
-\bibitem[\protect\citeauthoryear{{Davidson}, {Humphreys}  \& {Weis}}{{Davidson}
-  et~al.}{2016}]{davidson:16}
-{Davidson} K.,  {Humphreys} R.~M.,   {Weis} K.,  2016, arXiv:1608.02007, \href
-  {http://adsabs.harvard.edu/abs/2016arXiv160802007D} {}
-
-\bibitem[\protect\citeauthoryear{{de Koter}, {Heap}  \& {Hubeny}}{{de Koter}
-  et~al.}{1997}]{dekoter:97}
-{de Koter} A.,  {Heap} S.~R.,   {Hubeny} I.,  1997, \mn@doi [\apj]
-  {10.1086/303736}, \href {http://adsabs.harvard.edu/abs/1997ApJ...477..792D}
-  {477, 792}
-
-\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
-  et~al.}{2010}]{evans:10}
-{Evans} C.~J.,  et~al., 2010, \mn@doi [\apjl] {10.1088/2041-8205/715/2/L74},
-  \href {http://adsabs.harvard.edu/abs/2010ApJ...715L..74E} {715, L74}
-
-\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
-  et~al.}{2011}]{evans:11}
-{Evans} C.~J.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201116782},
-  \href {http://adsabs.harvard.edu/abs/2011A\%26A...530A.108E} {530, A108}
-
-\bibitem[\protect\citeauthoryear{{Fujii} \& {Portegies Zwart}}{{Fujii} \&
-  {Portegies Zwart}}{2011}]{fujii:11}
-{Fujii} M.~S.,  {Portegies Zwart} S.,  2011, \mn@doi [Science]
-  {10.1126/science.1211927}, \href
-  {http://adsabs.harvard.edu/abs/2011Sci...334.1380F} {334, 1380}
-
-\bibitem[\protect\citeauthoryear{{Gaia Collaboration} et~al.,}{{Gaia
-  Collaboration} et~al.}{2016}]{gaia:16}
-{Gaia Collaboration} et~al., 2016, \mn@doi [\aap]
-  {10.1051/0004-6361/201629272}, \href
-  {http://adsabs.harvard.edu/abs/2016A\%26A...595A...1G} {595, A1}
-
-\bibitem[\protect\citeauthoryear{{Gaia Collaboration}, {Brown}, {Vallenari},
-  {Prusti}, {de Bruijne}, {Babusiaux}  \& {Bailer-Jones}}{{Gaia Collaboration}
-  et~al.}{2018}]{brown:18}
-{Gaia Collaboration} {Brown} A.~G.~A.,  {Vallenari} A.,  {Prusti} T.,  {de
-  Bruijne} J.~H.~J.,  {Babusiaux} C.,   {Bailer-Jones} C.~A.~L.,  2018,
-  ArXiv:1804.09365, \href {http://adsabs.harvard.edu/abs/2018arXiv180409365G}
-  {}
-
-\bibitem[\protect\citeauthoryear{{Gruendl} \& {Chu}}{{Gruendl} \&
-  {Chu}}{2009}]{gruendl:09}
-{Gruendl} R.~A.,  {Chu} Y.-H.,  2009, \mn@doi [\apjs]
-  {10.1088/0067-0049/184/1/172}, \href
-  {http://adsabs.harvard.edu/abs/2009ApJS..184..172G} {184, 172}
-
-\bibitem[\protect\citeauthoryear{{Gvaramadze}, {Weidner}, {Kroupa}  \&
-  {Pflamm-Altenburg}}{{Gvaramadze} et~al.}{2012}]{gvaramadze:12}
-{Gvaramadze} V.~V.,  {Weidner} C.,  {Kroupa} P.,   {Pflamm-Altenburg} J.,
-  2012, \mn@doi [\mnras] {10.1111/j.1365-2966.2012.21452.x}, \href
-  {http://adsabs.harvard.edu/abs/2012MNRAS.424.3037G} {424, 3037}
-
-% \bibitem[\protect\citeauthoryear{{Hogg}}{{Hogg}}{2018}]{hogg:18}
-% {Hogg} D.~W.,  2018, ArXiv:1804.07766, \href
-%   {http://adsabs.harvard.edu/abs/2018arXiv180407766H} {}
-
-\bibitem[\protect\citeauthoryear{{Humphreys}, {Weis}, {Davidson}  \&
-  {Gordon}}{{Humphreys} et~al.}{2016}]{humphreys:16}
-{Humphreys} R.~M.,  {Weis} K.,  {Davidson} K.,   {Gordon} M.~S.,  2016, \mn@doi
-  [\apj] {10.3847/0004-637X/825/1/64}, \href
-  {http://adsabs.harvard.edu/abs/2016ApJ...825...64H} {825, 64}
-
-\bibitem[\protect\citeauthoryear{{K{\"o}hler} et~al.,}{{K{\"o}hler}
-  et~al.}{2015}]{kohler:15}
-{K{\"o}hler} K.,  et~al., 2015, \mn@doi [\aap] {10.1051/0004-6361/201424356},
-  \href {http://adsabs.harvard.edu/abs/2015A\%26A...573A..71K} {573, A71}
-
-\bibitem[\protect\citeauthoryear{{Kuiper}, {Yorke}  \& {Turner}}{{Kuiper}
-  et~al.}{2015}]{kuiper:15}
-{Kuiper} R.,  {Yorke} H.~W.,   {Turner} N.~J.,  2015, \mn@doi [\apj]
-  {10.1088/0004-637X/800/2/86}, \href
-  {http://adsabs.harvard.edu/abs/2015ApJ...800...86K} {800, 86}
-
-\bibitem[\protect\citeauthoryear{{Lada} \& {Lada}}{{Lada} \&
-  {Lada}}{2003}]{lada:03}
-{Lada} C.~J.,  {Lada} E.~A.,  2003, \mn@doi [\araa]
-  {10.1146/annurev.astro.41.011802.094844}, \href
-  {http://adsabs.harvard.edu/abs/2003ARA\%26A..41...57L} {41, 57}
-
-\bibitem[\protect\citeauthoryear{{Lamb}, {Oey}, {Segura-Cox}, {Graus},
-  {Kiminki}, {Golden-Marx}  \& {Parker}}{{Lamb} et~al.}{2016}]{lamb:16}
-{Lamb} J.~B.,  {Oey} M.~S.,  {Segura-Cox} D.~M.,  {Graus} A.~S.,  {Kiminki}
-  D.~C.,  {Golden-Marx} J.~B.,   {Parker} J.~W.,  2016, \mn@doi [\apj]
-  {10.3847/0004-637X/817/2/113}, \href
-  {http://adsabs.harvard.edu/abs/2016ApJ...817..113L} {817, 113}
-
-\bibitem[\protect\citeauthoryear{{Langer}, {Norman}, {de Koter}, {Vink},
-  {Cantiello}  \& {Yoon}}{{Langer} et~al.}{2007}]{langer:07}
-{Langer} N.,  {Norman} C.~A.,  {de Koter} A.,  {Vink} J.~S.,  {Cantiello} M.,
-  {Yoon} S.-C.,  2007, \mn@doi [\aap] {10.1051/0004-6361:20078482}, \href
-  {http://adsabs.harvard.edu/abs/2007A\%26A...475L..19L} {475, L19}
-
-\bibitem[\protect\citeauthoryear{{Lennon} et~al.,}{{Lennon}
-  et~al.}{2018}]{lennon:18}
-{Lennon} D.~J.,  et~al., 2018, ArXiv:1805.08277, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180508277L} {}
-
-\bibitem[\protect\citeauthoryear{{Lindegren} et~al.,}{{Lindegren}
-  et~al.}{2018}]{lindengren:18}
-{Lindegren} L.,  et~al., 2018, ArXiv:1804.09366, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180409366L} {}
-
-% \bibitem[\protect\citeauthoryear{{Maeder} \& {Meynet}}{{Maeder} \&
-%   {Meynet}}{2000}]{maeder:00}
-% {Maeder} A.,  {Meynet} G.,  2000, \aap, \href
-%   {http://adsabs.harvard.edu/abs/2000A\%26A...361..159M} {361, 159}
-
-\bibitem[\protect\citeauthoryear{{Oh} \& {Kroupa}}{{Oh} \&
-  {Kroupa}}{2016}]{oh:16}
-{Oh} S.,  {Kroupa} P.,  2016, \mn@doi [\aap] {10.1051/0004-6361/201628233},
-  \href {http://adsabs.harvard.edu/abs/2016A\%26A...590A.107O} {590, A107}
-
-\bibitem[\protect\citeauthoryear{{Parker}}{{Parker}}{1993}]{parker:93}
-{Parker} J.~W.,  1993, \mn@doi [\aj] {10.1086/116661}, \href
-  {http://adsabs.harvard.edu/abs/1993AJ....106..560P} {106, 560}
-
-\bibitem[\protect\citeauthoryear{{Pietrzy{\'n}ski} et~al.,}{{Pietrzy{\'n}ski}
-  et~al.}{2013}]{pietrzynski:13}
-{Pietrzy{\'n}ski} G.,  et~al., 2013, \mn@doi [\nat] {10.1038/nature11878},
-  \href {http://adsabs.harvard.edu/abs/2013Natur.495...76P} {495, 76}
-
-\bibitem[\protect\citeauthoryear{{Platais}, {van der Marel}, {Lennon},
-  {Anderson}, {Bellini}, {Sabbi}, {Sana}  \& {Bedin}}{{Platais}
-  et~al.}{2015}]{platais:15}
-{Platais} I.,  {van der Marel} R.~P.,  {Lennon} D.~J.,  {Anderson} J.,
-  {Bellini} A.,  {Sabbi} E.,  {Sana} H.,   {Bedin} L.~R.,  2015, \mn@doi [\aj]
-  {10.1088/0004-6256/150/3/89}, \href
-  {http://adsabs.harvard.edu/abs/2015AJ....150...89P} {150, 89}
-
-\bibitem[\protect\citeauthoryear{{Platais} et~al.,}{{Platais}
-  et~al.}{2018}]{platais:18}
-{Platais} I.,  et~al., 2018, ArXiv:1804.08678, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180408678P} {}
-
-\bibitem[\protect\citeauthoryear{{Poveda}, {Ruiz}  \& {Allen}}{{Poveda}
-  et~al.}{1967}]{poveda:67}
-{Poveda} A.,  {Ruiz} J.,   {Allen} C.,  1967, Boletin de los Observatorios
-  Tonantzintla y Tacubaya, \href
-  {http://adsabs.harvard.edu/abs/1967BOTT....4...86P} {4, 86}
-
-\bibitem[\protect\citeauthoryear{{Rakavy} \& {Shaviv}}{{Rakavy} \&
-  {Shaviv}}{1967}]{rakavy:67}
-{Rakavy} G.,  {Shaviv} G.,  1967, \mn@doi [\apj] {10.1086/149204}, \href
-  {http://adsabs.harvard.edu/abs/1967ApJ...148..803R} {148, 803}
-
-% \bibitem[\protect\citeauthoryear{{Ram{\'{\i}}rez-Agudelo}
-%   et~al.,}{{Ram{\'{\i}}rez-Agudelo} et~al.}{2015}]{ramirez-agudelo:15}
-% {Ram{\'{\i}}rez-Agudelo} O.~H.,  et~al., 2015, \mn@doi [\aap]
-%   {10.1051/0004-6361/201425424}, \href
-%   {http://adsabs.harvard.edu/abs/2015A\%26A...580A..92R} {580, A92}
-
-\bibitem[\protect\citeauthoryear{{Renzo} et~al.,}{{Renzo}
-  et~al.}{2018}]{renzo:18}
-{Renzo} M.,  et~al., 2018, ArXiv:1804.09164, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180409164R} {}
-
-% \bibitem[\protect\citeauthoryear{{Robitaille} \& {Bressert}}{{Robitaille} \&
-%   {Bressert}}{2012}]{robitaille:12}
-% {Robitaille} T.,  {Bressert} E.,  2012, {APLpy: Astronomical Plotting Library
-%   in Python}, Astrophysics Source Code Library (\mn@eprint {ascl} {1208.017})
-
-\bibitem[\protect\citeauthoryear{{Rosen}, {Krumholz}, {McKee}  \&
-  {Klein}}{{Rosen} et~al.}{2016}]{rosen:16}
-{Rosen} A.~L.,  {Krumholz} M.~R.,  {McKee} C.~F.,   {Klein} R.~I.,  2016,
-  \mn@doi [\mnras] {10.1093/mnras/stw2153}, \href
-  {http://adsabs.harvard.edu/abs/2016MNRAS.463.2553R} {463, 2553}
-
-\bibitem[\protect\citeauthoryear{{Rubio-D{\'{\i}}ez}, {Najarro},
-  {Garc{\'{\i}}a}  \& {Sundqvist}}{{Rubio-D{\'{\i}}ez}
-  et~al.}{2017}]{rubio-diez:17}
-{Rubio-D{\'{\i}}ez} M.~M.,  {Najarro} F.,  {Garc{\'{\i}}a} M.,   {Sundqvist}
-  J.~O.,  2017, in {Eldridge} J.~J.,  {Bray} J.~C.,  {McClelland} L.~A.~S.,
-  {Xiao} L.,  eds,  IAU Symposium Vol. 329, The Lives and Death-Throes of
-  Massive Stars. pp 131--135, \mn@doi{10.1017/S1743921317002447}
-
-\bibitem[\protect\citeauthoryear{{Sabbi} et~al.,}{{Sabbi}
-  et~al.}{2012}]{sabbi:12}
-{Sabbi} E.,  et~al., 2012, \mn@doi [\apjl] {10.1088/2041-8205/754/2/L37}, \href
-  {http://adsabs.harvard.edu/abs/2012ApJ...754L..37S} {754, L37}
-
-\bibitem[\protect\citeauthoryear{{Sana}, {Ram{\'{\i}}rez-Tannus}, {de Koter},
-  {Kaper}, {Tramper}  \& {Bik}}{{Sana} et~al.}{2017}]{sana:17}
-{Sana} H.,  {Ram{\'{\i}}rez-Tannus} M.~C.,  {de Koter} A.,  {Kaper} L.,
-  {Tramper} F.,   {Bik} A.,  2017, \mn@doi [\aap]
-  {10.1051/0004-6361/201630087}, \href
-  {http://adsabs.harvard.edu/abs/2017A\%26A...599L...9S} {599, L9}
-
-% \bibitem[\protect\citeauthoryear{{Schneider}, {Castro}, {Fossati}, {Langer}  \&
-%   {de Koter}}{{Schneider} et~al.}{2017}]{schneider:17}
-% {Schneider} F.~R.~N.,  {Castro} N.,  {Fossati} L.,  {Langer} N.,   {de Koter}
-%   A.,  2017, \mn@doi [\aap] {10.1051/0004-6361/201628409}, \href
-%   {http://adsabs.harvard.edu/abs/2017A\%26A...598A..60S} {598, A60}
-
-\bibitem[\protect\citeauthoryear{{Schneider} et~al.,}{{Schneider}
-  et~al.}{2018}]{schneider:18}
-{Schneider} F.~R.~N.,  et~al., 2018, \mn@doi [Science]
-  {10.1126/science.aan0106}, \href
-  {http://adsabs.harvard.edu/abs/2018Sci...359...69S} {359, 69}
-
-\bibitem[\protect\citeauthoryear{{Smith}}{{Smith}}{2016}]{smith:16}
-{Smith} N.,  2016, \mn@doi [\mnras] {10.1093/mnras/stw1533}, \href
-  {http://adsabs.harvard.edu/abs/2016MNRAS.461.3353S} {461, 3353}
-
-\bibitem[\protect\citeauthoryear{{Smith} \& {Tombleson}}{{Smith} \&
-  {Tombleson}}{2015}]{smith:15}
-{Smith} N.,  {Tombleson} R.,  2015, \mn@doi [\mnras] {10.1093/mnras/stu2430},
-  \href {http://adsabs.harvard.edu/abs/2015MNRAS.447..598S} {447, 598}
-
-\bibitem[\protect\citeauthoryear{{Udalski} et~al.,}{{Udalski}
-  et~al.}{2008}]{udalski:08}
-{Udalski} A.,  et~al., 2008, \actaa, \href
-  {http://adsabs.harvard.edu/abs/2008AcA....58..329U} {58, 329}
-
-\bibitem[\protect\citeauthoryear{{Ward} \& {Kruijssen}}{{Ward} \&
-  {Kruijssen}}{2018}]{ward:18}
-{Ward} J.~L.,  {Kruijssen} J.~M.~D.,  2018, \mn@doi [\mnras]
-  {10.1093/mnras/sty117}, \href
-  {http://adsabs.harvard.edu/abs/2018MNRAS.475.5659W} {475, 5659}
-
-\bibitem[\protect\citeauthoryear{{Zapartas} et~al.,}{{Zapartas}
-  et~al.}{2017}]{zapartas:17}
-{Zapartas} E.,  et~al., 2017, \mn@doi [\aap] {10.1051/0004-6361/201629685},
-  \href {http://adsabs.harvard.edu/abs/2017A\%26A...601A..29Z} {601, A29}
-
-\bibitem[\protect\citeauthoryear{{Zinnecker} \& {Yorke}}{{Zinnecker} \&
-  {Yorke}}{2007}]{zinnecker:07}
-{Zinnecker} H.,  {Yorke} H.~W.,  2007, \mn@doi [\araa]
-  {10.1146/annurev.astro.44.051905.092549}, \href
-  {http://adsabs.harvard.edu/abs/2007ARA\%26A..45..481Z} {45, 481}
-
-\bibitem[\protect\citeauthoryear{{Zwicky}}{{Zwicky}}{1957}]{zwicky:57}
-{Zwicky} F.,  1957, \zap, \href
-  {http://adsabs.harvard.edu/abs/1957ZA.....44...64Z} {44, 64}
-
-% \bibitem[\protect\citeauthoryear{{de Mink}, {Cantiello}, {Langer}, {Pols},
-%   {Brott}  \& {Yoon}}{{de Mink} et~al.}{2009}]{demink:09}
-% {de Mink} S.~E.,  {Cantiello} M.,  {Langer} N.,  {Pols} O.~R.,  {Brott} I.,
-%   {Yoon} S.-C.,  2009, \mn@doi [\aap] {10.1051/0004-6361/200811439}, \href
-%   {http://adsabs.harvard.edu/abs/2009A\%26A...497..243D} {497, 243}
-
-\makeatother
-\end{thebibliography}
-
-% \begin{acknowledgements}
-%   \small
-%   We are grateful to S.~Torres, M.~C.~Ramirez-Tannus,
-%   and C.~J.~Evans for help and discussions.
-SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC) (Grant agreement No.
-715063).
-VHB acknowledges support from the NRC-Canada Plaskett Fellowship.
-This work has made use of data from the ESA space mission \emph{Gaia} (\url{http://www.cosmos.esa.int/gaia}), processed by the \emph{Gaia} Data Processing and Analysis Consortium (DPAC, \url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}).
-Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \emph{Gaia} Multilateral Agreement.
-% The background image of \Figref{fig:main} is based on observations
-%   % made with ESO Telescopes at the La Silla Observatory under programme
-%   % ID 076.C-0888, processed and released by the ESO VOS/ADP group.
-%   % This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
-% \end{acknowledgements}
+\todo{maybe move around fig. 1
+\message{ !name(VFTS682_APJL.tex) !offset(844) }
 

### 2018-06-10 VFTS682_APJL.tex
table and moving figs

@@
 
+\begin{table}
+  \begin{center}
+    \caption{Observed and derived parameters for VFTS682. We also list
+    the average proper motion components of R136 derived by
+    \cite{lennon:18} for reference.}
+    \begin{tabular}{ll|c|c}
+      \hline
+      \hline
+      &Parameter & Value & Source\\
+
+       \hline
+       \multicolumn{4}{l}{\emph{Stellar Parameters VFTS 682}} \\
+     & present day $M$ \hfill $[M_\odot]$ & $137.8^{+27.5}_
+                                           {-15.9}$ &
+                                                     \multirow{3}{*}{\cite{schneider:18}}\\
+      &$M_\mathrm{ZAMS}$\hfill $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ &
+      \\
+      &age \hfill [Myr] & $1.0\pm0.2$ & \\[5pt]
+       \multicolumn{4}{l}{\emph{Average Proper motion in the region}} \\         &$\langle\mu_\mathrm{RA}\rangle$\hfill[$\mathrm{mas\ yr^{-1}}$] & $1.74\pm0.01$
+                        &  \multirow{2}{*}{\cite{lennon:18}}\\[5pt]
+      &$\langle\mu_\mathrm{DEC}\rangle$\hfill[$\mathrm{mas\ yr^{-1}}$]
+                & $0.70\pm0.02$ & \\
+  %    \multicolumn{4}{l}{\emph{Position}} \\
+
+    %  &RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
+   %                     & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
+    %  &DEC \hfill [degree] & -69.07 % $\pm$  0.05
+     %                   & \\[5pt]
+      \multicolumn{4}{l}{\emph{Relative Proper motion}} \\                 
+      &$\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \multirow{2}{*}{\emph{Gaia} DR2}\\[5pt]
+      &$\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
+
+      &$\delta\mu_\mathrm{RA}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.01$\pm$0.13 & \multirow{2}{*}{\cite{platais:18}}\\[5pt]
+      &$\delta\mu_\mathrm{DEC}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] &
+                                                                        0.2$\pm$0.1 &
+      \\[5pt]
+      \multicolumn{4}{l}{\emph{\SdM{ Motion in polar coordinates }}} \\     
+      &$\delta v_\mathrm{rad, R136}$  \hfill[$\kms$] & xxx & \cite{This Work}\\   
+       &$\theta_\mathrm{rad, R136}$  \hfill[$\kms$] & xxx & \cite{This Work}\\
+
+  %    &$\delta v_\mathrm{rad}$  \hfill[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
+
+      \hline
+
+    \end{tabular}
+    \tablecomments
+    {The error on the RA and DEC positions, are of order
+      $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$ in \emph{Gaia}
+      DR2. %REPETITION: %  We adopt the peculiar
+      % radial velocity calculated using the  NV $\lambda4944$ line, which is
+      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
+    }
+  \end{center}
+  \label{tab:vfts682}
+\end{table}
+
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
+  \includegraphics[width=0.315\textwidth]{figures/angle_dist}
+  \caption{Left panel: distribution of OB-type and Wolf-Rayet stars in proper
+    motion relative to R136. VFTS682 is not an outlier, but
+    its relative proper motion matches the expected value if it were indeed
+    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
+    assuming a distance of 50\,kpc. Right panel:  distribution of
+    angles between the relative proper motion direction and the radial
+    direction to the star. The error bars for VFTS682 are large, but
+    the best value is in agreement with the hypothesis of dynamical
+    ejection. In both
+    panels, the dark blue histograms contain 317 
+    stars with error smaller than $0.1\,\mathrm{mas \
+      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
+    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
+      yr^{-1}}$. \todo{ Todo: [1] insert correct errorbars in right pabel, [2] make linewidth, shading and figure size of both panels consistent [3] fix normalization to be consistent in both pabels}}
+  \label{fig:mu_dist}
+\end{figure*}
+
 \subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
@@
 \subsection{New data from \emph{Gaia} DR2  \label{data:gaia}}
-
-\begin{table}
-  \begin{center}
-    \caption{Observed and derived parameters for VFTS682. We also list
-    the average proper motion components of R136 derived by
-    \cite{lennon:18} for reference.}
-    \begin{tabular}{l|c|c}
-      \hline
-      \hline
-      Parameter & Value & Source\\
-      \hline
-      RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
-                        & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
-      DEC \hfill [degree] & -69.07 % $\pm$  0.05
-                        & \\[5pt]
-      \hline
-      $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \multirow{2}{*}{\emph{Gaia} DR2}\\[5pt]
-      $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
-      \hline
-      $\delta\mu_\mathrm{RA}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.01$\pm$0.13 & \multirow{2}{*}{\cite{platais:18}}\\[5pt]
-      $\delta\mu_\mathrm{DEC}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] &
-                                                                        0.2$\pm$0.1 &
-      \\[5pt]
-      \hline
-      $\delta v_\mathrm{rad}$  \hfill[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
-      \hline
-      present day $M$ \hfill $[M_\odot]$ & $137.8^{+27.5}_
-                                           {-15.9}$ &
-                                                     \multirow{3}{*}{\cite{schneider:18}}\\
-      $M_\mathrm{ZAMS}$\hfill $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ &
-      \\
-      age \hfill [Myr] & $1.0\pm0.2$ & \\
-      \hline
-      $\langle\mu_\mathrm{RA}\rangle$\hfill[$\mathrm{mas\ yr^{-1}}$] & $1.74\pm0.01$
-                        &  \multirow{2}{*}{\cite{lennon:18}}\\[5pt]
-      $\langle\mu_\mathrm{DEC}\rangle$\hfill[$\mathrm{mas\ yr^{-1}}$]
-                & $0.70\pm0.02$ & \\
-      \hline
-
-    \end{tabular}
-    \tablecomments
-    {The error on the RA and DEC positions, are of order
-      $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$ in \emph{Gaia}
-      DR2. %REPETITION: %  We adopt the peculiar
-      % radial velocity calculated using the  NV $\lambda4944$ line, which is
-      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
-    }
-  \end{center}
-  \label{tab:vfts682}
-\end{table}
 
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figures/dist_mu_region.pdf}\\
-  \includegraphics[width=0.5\textwidth]{figures/angle_dist}
-  \caption{Top panel: distribution of OB-type and Wolf-Rayet stars in proper
-    motion relative to R136. VFTS682 is not an outlier, but
-    its relative proper motion matches the expected value if it were indeed
-    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
-    assuming a distance of 50\,kpc. Bottom panel: \todo{redo with
-      errorbars} distribution of
-    angles between the relative proper motion direction and the radial
-    direction to the star. The error bars for VFTS682 are large, but
-    the best value is in agreement with the hypothesis of dynamical
-    ejection. In both
-    panels, the dark blue histograms contain 317 
-    stars with error smaller than $0.1\,\mathrm{mas \
-      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
-    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$. }
-  \label{fig:mu_dist}
-\end{figure}
-
 The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
@@
 
-\todo{maybe move around fig. 1}
 \citet{lennon:18} carried out a study similar to ours on the fast

### 2018-06-10 VFTS682_APJL.tex
Selma sunday - mostly section 2 and tables

@@
 %\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
-\usepackage[%draft, 
-pagebackref,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
-\renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
+\usepackage[%draft, pagebackref,
+colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
+%\renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
 \usepackage{aas_macros}
@@
 \newcommand{\Lsun}{{\,\mathrm{L}_\odot}}
-\newcommand{\masyr}{{\,\mathrm{mas}\,\mathrm{yr}^{-1}}}
+\newcommand{\Myr}{{\,\mathrm{Myr}}}
+\newcommand{\masyr}{\,\mathrm{mas}\,\mathrm{yr}^{-1}}
 
@@
 % runaway star known to date}
-\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
-
-\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
+%\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
+\title{Gaia's constraints on the extremely massive  ($\sim$$150\,M_\odot$) runaway star candidate VFTS682}
+
+\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
+  R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{4}$, 
+  J.~Bestenlehner$^{5}$, \SdM{C.~J.~Evans$^{X}$},
   V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{1,7,8}$,  A.~de~Koter$^{1}$,
-  N.~Langer$^{9}$, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
-\affil{$^{1}${Astronomical Institute Anton Pannekoek, University of
+  N.~Langer$^{9}$,  \SdM{F. Najarro$^{X}$}, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
+\affil{ $^{*}$ Corresponding author.  A list of the affiliations can be found at the end of this paper.}
+
+ \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+\begin{abstract}
+
+ How very massive stars form is still among the most intriguing open  questions in stellar astrophysics.
+VFTS 682 is among the most  extreme massive stars known, with an inferred initial mass of about  $150\,M_\odot$.
+It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.
+The absence of other massive stars in its immediate vicinity led to two intriguing hypotheses posed in earlier work: either (a) it formed in relative isolation through a new mode of star formation or (b) it was ejected dynamically from the central cluster.
+
+Aiming to shine light on this debate, we investigate the kinematics of  VFTS 682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space Telescope} photometry.
+We derive a projected velocity relative of to the cluster of $\sim$$30\pm20\kms$.
+The direction and magnitude of the proper is consistent with what is expected if VFTS 682 was ejected as a slow runaway from the central cluster, given its distance and the inferred age.
+However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.
+
+If future data confirms the runaway nature, this would make the VFTS682 the most massive runaway star known to date.
+While we cannot prove this solidly from the current data alone, we do consider this hypothesis as the most plausible because of a variety of circumstantial clues.
+The central cluster is known to harbor several other stars more massive than $150\,M_\odot$ similar in spectral type.
+The very massive O star VFTS 16, which has recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions by numerical simulations.
+\end{abstract}
+
+\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS 682}
+\maketitle{}
+
+\section{Introduction}
+\label{sec:intro}
+
+How massive stars form is one of the major longstanding questions in astrophysics
+\citep[e.g.,][]{zinnecker:07}. Improving our understanding of massive star formation, and its
+possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies, but also for understanding the transients that mark their death and the compact remnants they leave behind.
+%
+Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
+% \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
+evolve fast, typically reside in dense groups and remain enshrouded in
+their parent cloud during the entirety of their formation
+process. Important progress has been made on the theoretical side,
+\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally very expensive and therefore remain challenging.
+
+In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to reside in the innermost cores.
+In this picture, field stars are primarily the result of the dissolution of dense groups.
+However, a significant population of massive stars exists in relative isolation, far from dense clusters or OB associations and their origin remains matter of debate \citep{gvaramadze:12, lamb:16,ward:18}.
+One hypothesis to explain the population of relatively isolated massive stars is that they formed in the field.
+The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
+Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.
+
+% A contribution to the debate on whether or not massive
+% stars form in relative isolation was presented by
+% \cite{bestenlehner:11} and \cite{bressert:12}, who discussed the case of the very massive star
+% VFTS682.
+
+One of the most extreme examples that has been considered in this debate is the very massive star VFTS682  \citep[][]{bestenlehner:11, bressert:12}.
+This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}.
+It is a hydrogen-rich Wolf-Rayet star of spectral type WNh5.
+Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $\sim$$140^{+30}_{-16}\,M_\odot$ corresponding to an initial mass of $\sim$$150^{+30}_{-17}\,M_\odot$
+\citep{schneider:18}. 
+%
+%$137.8^{+27.5}_{-15.9}\,M_\odot$,
+%
+This makes VFTS682 one of the most massive stars known and one of the most extreme objects in the region.
+%
+From the spectral point of view, it is reminiscent of the very
+massive stars % , a.k.a. ``monster stars'',
+ in the core of the R136 cluster \citep{dekoter:97,crowther:10, crowther:16}. 
+ %
+ In particular, a remarkable similarity exist between the
+spectrum of VFTS682 and R136a3 \citep{rubio-diez:17} for which \citet{crowther:16} report a
+current mass estimate of $180^{+30}_{-30}\Msun$. R136 hosts
+at least two more very massive WN5h stars, R136a1 and R136a2, whose
+estimated current masses are even higher.
+
+%at a projected offset of 119.4 arcseconds, corresponding to
+
+VFTS682 stands out by its relative isolation at a projected distance of 119.4 arcseconds, corresponding to 
+$\sim$$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
+considered two possible explanation for the offset: either
+the star formed in situ as an isolated massive star, % , but remark that this
+                                % poses a challenge for starformation
+                                % theories (They don't)
+or it was ejected from  R136. N-body simulations % of dynamical interaction in star clusters
+indicate that the ejection of very massive stars like VFTS682 is expected \citep[e.g.][]{fujii:11, banerjee:12}.
+This is supported by the recent findings of other massive runaway stars in the region based on proper motion studies.
+\citet{platais:15,platais:18} analyze multi-epoch \emph{Hubble Space Telescope} (HST) photometry and identify 10 stars, which appear to be ejected from R136.
+\citet{lennon:18} investigate the kinematics of  isolated O-type stars in the region using the second \emph{Gaia} data release \cite[DR2,][]{gaia:16,brown:18} and show that the proper motion, postion and direction of the $\sim100\Msun$ star VFTS16 is consistent with a runaway origin from R136.
+
+%Testing the origin of VFTS682 is not only important for constraining isolated star formation; it should also help to improve understanding of early star-cluster dynamics.
+
+In this paper we present an analysis of the new kinematical
+constraints for VFTS682 provided by \emph{Gaia} DR2 and constraints
+from HST proper data  by \citet{platais:18}.   We discuss the
+implications for the hypothesis that VFTS682 is a slow runaway star
+ejected from R136.
+% which was not part of the sample of \citet{lennon:18} because of its  spectral type.
+
+%%% cut for length reasons: all is said below
+% Our results indicate that VFTS682 is a runaway star, although with
+% low statistical significance. The
+% direction and magnitude of the velocity vector are consistent with
+% dynamical ejection from R136. The age inferred from the spectral
+% analysis \citep[from][]{schneider:18} is consistent with the travel
+% time we calculate for this star. % Therefore, the hypothesis that VFTS
+% % 682 is formed in relative isolation is rejected.
+% If our results are confirmed by future astrometric data releases, VFTS
+% 682, with an inferred present day mass well above a hundred solar
+% masses, is the most massive runaway star known to date.
+
+\section{Observations}
+\label{sec:sample}
+
+\begin{table}
+  \begin{center}
+    \caption{Selected stellar parameters for VFTS682. }
+    \begin{tabular}{llc|c|c}
+      \hline
+      \hline
+      &Parameter & Units & Value & Ref.\\
+
+       \hline
+       \multicolumn{5}{l}{\emph{Selected stellar parameters VFTS 682}} \\
+     & present day mass  & $[M_\odot]$ & $137.8^{+27.5}_
+                                           {-15.9}$ &
+                                                    \\
+      & initial mass& $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ &
+      \\
+      &age & [Myr] & $1.0\pm0.2$ & \\
+      &mass loss rate &  & $\SdM{xxx}$ & \\
+      &helium abundance &  & $\SdM{xxx}$ & \\
+      \hline
+
+    \end{tabular}
+    \tablecomments
+    { 
+    }
+  \end{center}
+  \label{tab:star}
+\end{table}
+
+\begin{table}
+  \begin{center}
+    \caption{Kinematics of VFTS682. }
+    \begin{tabular}{llc|c|c}
+      \hline
+      \hline
+      &Parameter & Units & Value & Ref.\\
+
+       \hline
+           \multicolumn{5}{l}{\emph{Gaia's proper motion for VFTS 682 and average of the region}} \\
+          &$\mu_\mathrm{RA}$&[$\masyr$] & $1.843\pm 0.070$
+                     & (1) \\        
+               &$\mu_\mathrm{DEC}$&[$\masyr$] & $0.786\pm 0.080$
+                     &  (1) \\        
+                 & Correlation  ($\mu_\mathrm{RA}$, $\mu_\mathrm{DEC}$) &[$\masyr$] & $0.023$
+                        & (1)  \\         
+     %  \multicolumn{5}{l}{\emph{Gaia's average proper motion in the region}} \\         
+       &$\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$&[$\masyr$] & $1.74\pm0.01$
+                        & (2) \\
+      &$\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$&[$\masyr$]
+                & $0.70\pm0.02$ &  (2)\\
+\hline
+
+  %    \multicolumn{5}{l}{\emph{Position}} \\
+
+    %  &RA &[degree] &  \phantom{-}84.73 % $\pm$  0.03
+   %                     & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
+    %  &DEC & [degree] & -69.07 % $\pm$  0.05
+     %                   & \\[5pt]
+      \multicolumn{5}{l}{\emph{Gaia's relative proper motion for VFTS 682 }} \\                 
+      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & 0.10$\pm$0.07 & (1,4) \\
+      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & 0.08$\pm$0.09 &  (1,4) \\
+      &$\delta\mu_\mathrm{}$  &[$\masyr$] & \SdM{xxx}$\pm$\SdM{xxx} &  (1,4) \\
+                 &$v_\mathrm{2d}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
+ \hline     
+          \multicolumn{5}{l}{\emph{HST's independent determination of the relative proper motion for VFTS 682}} \\  
+      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & 0.01$\pm$0.13 & (3) \\
+      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & 0.2$\pm$0.1 &  (3) \\
+       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & 0.20$\pm$0.15 &  (3) \\
+                  &$v_\mathrm{2d, HST}$  &[$\kms$] & \SdM{$47\pm24$}\footnote{\SdM{This can't be right, the error bar is suddenly smaller than in the mas units!}} & (3)\\  
+\hline
+      \multicolumn{5}{l}{\emph{ Direction and magnitude w.r.t the vector pointing away from R136}} \\     
+       &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
+        &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
+ %          &$v_\mathrm{2d}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\
+
+\hline
+  %    &$\delta v_\mathrm{rad}$  &[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
+
+      \hline
+
+    \end{tabular}
+    \tablecomments
+    { Proper motions are based on Gaia data unless indicated otherwise in subscript. The angle $\theta$ is defined such that $\theta=0$ corresponds to the direction pointing away from R136.
+    $^1$\cite{Gaia-Collaboration+2018}, 
+    $^2$\cite{lennon:18}, 
+    $^3$\cite{platais:18},  
+    $^4${\color{blue}This work}
+    %The error on the RA and DEC positions, are of order $\sim$$10^{-2}\masyr$ in \emph{Gaia} DR2. %REPETITION: %  We adopt the peculiar
+      % radial velocity calculated using the  NV $\lambda4944$ line, which is
+      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
+    }
+  \end{center}
+  \label{tab:vfts682}
+\end{table}
+
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
+  \includegraphics[width=0.315\textwidth]{figures/angle_dist}
+  \caption{Left panel: distribution of OB-type and Wolf-Rayet stars in proper
+    motion relative to R136. VFTS682 is not an outlier, but
+    its relative proper motion matches the expected value if it were indeed
+    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
+    assuming a distance of 50\,kpc. Right panel:  distribution of
+    angles between the relative proper motion direction and the radial
+    direction to the star. The error bars for VFTS682 are large, but
+    the best value is in agreement with the hypothesis of dynamical
+    ejection. In both
+    panels, the dark blue histograms contain 317 
+    stars with error smaller than $0.1\,\mathrm{mas \
+      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
+    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
+      yr^{-1}}$. \todo{ Todo: [1] insert correct errorbars in right pabel, [2] make linewidth, shading and figure size of both panels consistent [3] fix normalization to be consistent in both pabels}}
+  \label{fig:mu_dist}
+\end{figure*}
+
+%\subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
+
+The star WNh5 star VFTS682, located at right ascension (RA)
+05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
+%
+%of spectral type WNh5 using multi-epoch spectra covering $\lambda$4000--7000 from the
+%VFTS survey. % The spectra available covered $\lambda$4000--7000 and
+% were taken at multiple epochs.
+%
+\citet{bestenlehner:11}  analyzed the spectra to infer the stellar parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of the brightest stars in the region.
+The star is unlikely to have a close companion, because of the absence of clear radial velocity variations \citet{bestenlehner:11}.
+Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see Tab~ \label{tab:vfts682} for an overview of the parameters.
+
+Estimates of the radial velocity are complicated by the variable, optically thick winds that is typical for emission line stars.
+\citet{bestenlehner:11} estimate a mass loss rate of $10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the possible effect of clumping.
+The V-band light curve of VFTS682  shows variations at a $\sim$10\% level on a timescale of years, which is unusual for Wolf-Rayet stars and more typical for Luminous Blue Variable (LBV) stars \citep{udalski:08, bestenlehner:11}.
+It also shows mid-infrared excess \citep{gruendl:09}.
+We therefore caution for over-interpreting the estimates for the radial velocities, but we will summarize the estimates that have been published.
+\citet{bestenlehner:11}  estimate a radial velocity (RV) of  $300\pm10\kms$ using the  N{\footnotesize V} $\lambda4944$ line, which is offset from the average radial velocity of the region, which is  $270\pm10\kms$.
+This is consistent with, but no proof of, a runaway nature.
+\cite{bressert:12} note an offset compared between the RV of the star and the nebular lines from the gas filaments in its vicinity.
+This is consistent with, but also no proof of, the expectation if the star was not formed in situ.
+Newer measurements based on new Xshooter data give different estimates for the RV ({\color{blue} Rubio-D{\' i}ez et al.
+in prep}).
+We will therefore refrain from using the RV measurements in this work.
+
+%of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
+
+%% Radial Velocity:
+
+The distance to 30 Doradus is $50$\,kpc  \citep[][]{pietrzynski:13}.
+The error on the distance determination is small ($\lesssim2\%$) and the possible difference between VFTS682 and the average of the region is probably even smaller (likely $\ll 1\%$).
+These uncertainties are negligible compared to the errors in the proper motion discussed below.
+
+%\citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude lower than found by \citet{evans:11}.  - Citing OGLE is enough
+
+%\SdM{\citet{bestenlehner:11} further report a wind mass loss rate of $10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, without correcting for the possible effect of clumping.
+The helium surface  abundance is  $Y=0.45\, (0.49)$ \cite{bestenlehner:11} (\citealt{rubio-diez:17}).
+IN TABLE ONLY TO SAFE SPACE}
+
+\subsection{ \emph{Gaia} astrometry for VFTS682\label{data:gaia}}
+
+%\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}
+
+VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band \citep{Gaia-Collaboration+2016a, Gaia-Collaboration+2018}.
+The number of visibility periods, i.e.
+groups of observations separated from other groups by a gap of at least 4 days, used in the astrometric solution is 17 for this star.
+The reported astrometric excess noise is zero.
+\citep[][]{lindengren:18}.
+These values suggest that the \emph{Gaia} DR2  data for VFTS682 are reliable.
+Generally the Gaia errors are correlated, but the correlation coefficient is small for this star, see Tab.~\ref{tab:vfts682}.
+We can therefore treat the proper motion components as uncorrelated.
+
+To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.
+They select  bright ($G<17$) stars within 0.05 degrees of R136 with and exclude sources with proper motion errors greater than $0.01\masyr$ in both co-ordinates (see their  Sect.~2.1 for further discussion).
+
+With these we compute the relative proper motion $\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.
+We also derive the angle $\theta$ between the direction of motion and the vector connecting the center of R136 with the current position of VFTS 682,  i.e.
+such that $\theta = 0$ is the angle expected for a star ejected from R136.
+We further provide the 2D velocity projected onto the radial vector pointing away from R136 $v_\mathrm{2d, \parallel}$  and the total projected 2d velocity $v_\mathrm{2d}$.
+All kinematic quantities are provided in Tab.~\ref{tab:vfts682}.
+
+\subsection{HST astrometry for VFTS682}
+% observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
+% in October 3-8 2011 and the second epoch (GO-13359) in
+% October 6-11 2014, both using the F775W filter.
+% The second epoch was designed to match the pointing and
+% orientation of the first epoch as exactly as possible so
+% that the displacement measurements could be purely differential.
+
+%For large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even exceed \emph{Gaia} DR2 data in quality. \SdM{Mathieu: do you have a ref for this?}
+
+The 30 doradus region was target of an a two-epoch photometric campaign with \emph{HST} providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{Sabbi+2013}).
+Visually, the star is clearly isolated and does not appear to have any close neighbor.
+This by itself gives further confidence in ability of \emph{Gaia} to measure the proper motion reliably without being affected by crowding.
+
+ \citet{platais:18} recently analyzed the HST data for candidate runaway stars.
+The brightest stars (V$<$14) are missing from the catalogue because of saturation effects, but the high extinction towards VFTS682 makes it faint enough to be included.
+ Our star did not make the more conservative selection criteria for runaway stars adopted in that study.
+However, the measurements are useful here, since they provides a measurement of the proper motion of VFTS 682 that is completely independent from the \emph{Gaia} measurement.
+
+We list their data in \Tabref{tab:vfts682}.
+The errorbars are larger than we obtain from \emph{Gaia}, but the values are consistent.
+\SdM{Actually they are not consistent with values found in one of Danny's emails.
+What is going on here?}
+
+%ID     /HST relative mu_x/HST relative mu_y/HST error_mu_x/HST error_my_y/  HST F775W mag/Gaia B-mag?/Gaia relative mu_x/  Gaia relative mu_y/Gaia error_pmRA/Gaia error_pmDec
+
+%682  -0.02   0.19 0.10 0.09 15.175 15.656   0.133   0.166 0.070   0.080
+
+%The catalog from \citet{platais:18} lists the proper motion components of VFTS682 relative to the field, providing a
+%measure of the motion of this star completely independent from
+%\emph{Gaia}. They found proper
+%motion components relative to the field
+%$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
+%  yr^{-1}}$ and
+%$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
+%  yr^{-1}}$, which added in quadrature give a relative proper motion of
+%$\delta \mu^{HST} =0.20 \pm 0.15\masyr$.
+% \begin{equation}
+%   \label{eq:pm_around_HST}
+%   \delta \mu^{HST} = \sqrt{\left(\delta\mu_\mathrm{RA}^{HST}\right)^2+\left(\delta\mu_\mathrm{DEC}^{HST}\right)^2}
+%   = 0.20 \pm 0.15\,\mathrm{mas\
+%   yr^{-1}} \ \ .
+% \end{equation}
+
+%These proper motion components can also be converted into projected
+%velocities assuming a distance of 50\,kpc, obtaining $\delta
+%v_\mathrm{RA}^{HST}=2\pm31\kms$ and $\delta
+%v_\mathrm{RA}^{HST}=47\pm24\kms$% , corresponding to a two-dimensional
+%% projected velocity of $47\pm24\kms$
+%.
+%Adding them in quadrature to the radial velocity
+%measured by \citet{bestenlehner:11} we obtain a three-dimensional velocity of
+%$v_\mathrm{pec}^{HST}=56 \pm 23  \kms$.
+% \begin{equation}
+%   \label{eq:speed_around_HST}
+%   v_\mathrm{pec}^{HST} = \sqrt{\left(\delta v_\mathrm{RA}^{HST}\right)^2
+%     +\left(\delta v_\mathrm{DEC}^{HST}\right)^2+\left(\delta
+%       v_\mathrm{rad}^{VFTS}\right)^2} = 56 \pm 23
+%   \kms \ .
+% \end{equation}
+
+\section{The kinematics of VFTS682}
+\label{sec:results}
+
+\subsection{Is it a runaway star?}
+\label{sec:runaway}
+We first address the question of whether VFTS682 is a typical star
+from the kinematic point of view, or whether it is a runaway star with
+a significantly large peculiar velocity compared to its surrounding population. The former is what should
+be expected if it formed in the relative isolation that we observe today.% , in relative
+% isolation from other massive stars.
+
+% Using the stars selected as described in \Secref{data:gaia} (a
+% subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
+% $\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\masyr$ and
+% $\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
+% yr^{-1}}$. These values are
+% in good agreement with what is found by \cite{lennon:18}. However, we
+% emphasize their sensitivity to the choice of stars adopted to define
+% the reference frame.
+Subtracting the mean proper motion components given by
+\citet{lennon:18} from the
+proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
+components of proper motion of the star relative to the surrounding region
+$\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\masyr$
+and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08
+\pm 0.09\masyr$. These components result in a
+two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\
+  yr^{-1}}$.
+
+% \begin{equation}
+%   \label{eq:pm_gaia_around}
+%   \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
+%   = 0.13\pm 0.09\,\mathrm{mas\
+%   yr^{-1}} \ \ .
+% \end{equation}
+
+% Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
+% of the stars used in \cite{lennon:18} to define the frame of reference
+% we also adopt here, cf.~the right panel in their Fig.~2. The
+% proper motions are decomposed in the radial and tangential direction
+% from R136, to highlight the likelihood of it being the origin of fast
+% moving stars. The plus sign marks VFTS682, which sits at the edge of
+% the proper motion distribution, but is not a clear outlier. Most
+% relevant is the direction of the proper motion of VFTS682, which we
+% discuss in \Secref{sec:r136_origin}.
+
+% \begin{figure}%[htbp]
+%   \centering
+%   \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
+%   \caption{Polar plot of the relative proper motion components for the stars
+%     defining our reference frame. VFTS682 is indicated by the plus sign, and
+%     other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
+%     0.1, 0.2, and 0.4$\masyr$. Positive angle
+%     indicate a tangential component pointing in the counterclockwise
+%     direction, with 0 degrees corresponding to proper motion pointing
+%     radially outward from R136. The proper motion for VFTS682
+%     corresponds to almost radial motion away from the cluster.
+%   }
+%   \label{fig:pm_polar}
+% \end{figure}
+
+The proper motion components can be converted into
+the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=24\pm19\,\kms$,
+$\delta v_\mathrm{DEC}^{Gaia}=20\pm23\,\kms$, assuming a distance of
+50\,kpc. These can be combined obtaining a projected two-dimensional
+velocity of $31\pm21\kms$. % (we do not account for the uncertainty in the distance
+% estimate when propagating errors).
+The radial velocity from
+\cite{bestenlehner:11} then gives the third component along
+the line of sight, % allowing us to calculate the three-dimensional
+% peculiar speed of the star:
+which added in quadrature to the transverse components results in a
+three-dimensional velocity of $44 \pm 21\kms$.
+% \begin{equation}
+%   \label{eq:speed_around_Gaia}
+%   v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
+%     +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
+%       v_\mathrm{rad}^{VFTS}\right)^2} = 44 \pm 21
+%   \kms \ .
+% \end{equation}
+
+Therefore, two completely independent measures of the proper motion of
+VFTS682 relative to the surrounding field, one from HST and one from
+\emph{Gaia} DR2, yield values of the peculiar three-dimensional
+velocity of VFTS682 % (\Eqref{eq:speed_around_HST} and
+% \Eqref{eq:speed_around_Gaia}, respectively)
+which would make it the most massive runaway star
+known to date. However, the large errors on both the proper motion measures
+require confirmation with future astrometric data.
+
+\subsection{Does it come from the R136 cluster?}
+\label{sec:r136_origin}
+
+% Assuming that the best estimate of the peculiar velocity of VFTS682
+% are reliable, we now address the question of its likely origin.
+
+Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
+of OB-type and Wolf-Rayet stars included both in the VFTS survey and
+\emph{Gaia DR2} with reliable astrometric solutions. The bottom panel
+shows the distribution in angles $\theta$ between the relative proper motion
+direction and the direction from the core of R136 to the star. We
+emphasize that our subset of stars with reliable proper motion
+measurements is biased towards the fast moving objects, which results in a
+distribution of angles mildly peaked at small angles.
+
+VFTS682 is not an outlier in relative proper motion%  compared
+% to the two O-type runaways (re-)identified by \cite{lennon:18} (see
+% also below)
+. However, the star was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
+the dynamical ejection scenario. The green shade in
+\Figref{fig:mu_dist} shows the range of relative proper motions
+required to reach the present day location within
+the uncertainties on the apparent age of the star, assuming it was
+ejected from R136 very early in its life.
+
+Because of the large
+error bars, the angle between the relative proper motion is not very
+constraining, but it is suggestive that the best value is close to zero.
+Therefore, the relative proper motion from \emph{Gaia} DR2 are consistent with the hypothesis
+of dynamical ejection.
+
+The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
+VFTS682 from \emph{Gaia} DR2, and the
+yellow arrows illustrate the uncertainty. % show the possible
+% range of directions within the uncertainties in the measured relative proper
+% motion.
+These arrows cross at the present-day location of VFTS682 and
+are prolonged in the direction opposite to the motion to illustrate
+the possible range of origins. % The yellow arrow shows the direction of
+% the relative proper motion from HST (see \Eqref{eq:pm_around_HST}),
+% but we omit to show the possible range of directions from HST data, since it is
+% insufficiently constrained to be useful.
+
+% Although the large uncertainties on the relative proper motion
+% components results in a wide range of possible directions,we argue that the most likely origin of the star is R136.
+
+The kinematic age of this star, assuming it originates from R136, is
+
+\begin{equation}
+  \label{eq:kin_age}
+  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu^\mathrm{Gaia}} \simeq
+  \frac{119.4\,\mathrm{arcsec}}{0.13\masyr} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
+\end{equation}
+where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
+the core of the cluster \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
+As in the rest of this study, we neglect for
+simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
+The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
+ejection from the cluster, given is apparent age
+of $1.0\pm 0.2$\,Myr \citep{schneider:18}. % , which
+% corroborates the idea that the star is the result of a dynamical
+% ejection very soon in the cluster evolution.
+
+% the thing below is not relevant if it is a merger, better cut out
+% for now, since we are too long anyway
+
+%Note that the
+% present-day surface helium abundance
+% \citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
+% limit on the age of the star of $\sim$0.9\,Myr, corresponding to the time needed to
+% synthesize this amount of helium in the models from \cite{kohler:15}.
+
+\begin{figure}%[tbp]
+  \centering
+  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
+  \caption{The red solid arrow indicates the proper motion of VFTS682
+    relative to the region from \emph{Gaia} DR2, starting from the present day position of
+    the star. The yellow arrows indicate the possible
+    directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
+    backwards (dashed) to illustrate the uncertainty on the origin of the
+    star. The length of the prolongations is proportional to the relative proper motion
+    times the age of VFTS682 \citep[$1.0\pm0.2$\,Myr,][]{schneider:18}.
+    % \todo{load color picture on background? Fig with HST data
+    % available, but confusing...}
+  }
+
+  \label{fig:main}
+\end{figure}
+
+\section{Discussion}
+\label{sec:discussion}
+
+Based on our results, we tentatively claim that VFTS682 is the most massive
+runaway known to date, with a peculiar three-dimensional speed of
+$44\pm21\,\kms$. Due to the large error bars, this result will need
+to be revisited with future astrometric data. % to be revisited as updated
+% astrometric parameters from future \emph{Gaia} data releases become
+% available.
+If confirmed, it means that isolated star formation is
+\emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
+$0.9\pm0.6$\,Myr ago. Because of the exceptionally large mass
+of this star, this raises the question of which stars must populate
+the core of the cluster.
+
+Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
+massive star among those interacting \cite[e.g.,][]{banerjee:12}. This means that, just
+based on the kinematic properties of VFTS682, we would expect several
+stars with initial masses larger than $\sim$$150\,M_\odot$ in the
+cluster R136.
+This is consistent with the detection
+of extremely massive stars in the core of the
+cluster. % The projected rotational equatorial
+% velocity\footnote{However, the determination of the rotational
+%   velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
+%   considered with caution.} of VFTS682
+% reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
+% line with the average rotation rate of massive stars in the region
+% \citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
+% experienced rotationally induced chemically homogeneous evolution
+% \citep[][]{maeder:00,demink:09}, and (ii) it has not
+% accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
+% data of the VFTS survey rule out the presence of a close companion at the
+% present day. Moreover,
+
+The spectral type of VFTS682
+\citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
+three most massive stars detected in the core of the cluster%  by
+% \cite{dekoter:97,crowther:10,crowther:16}
+, with an astonishing similarity in particular with
+the spectrum of R136a3. Therefore, the isolation of
+VFTS682 makes it an ideal target to constrain the stellar physics of
+stars with masses well above $\sim$$100\,M_\odot$ while avoiding
+crowding issues. %  : its isolation makes
+% it an easier target for observations compared to the similar stars
+% present in the crowded core of R136.
+
+\citet{banerjee:12} used N-body simulations of fully segregated
+clusters with all massive stars in binaries to suggest that VFTS682
+was ejected from R136. They
+demonstrated that the cluster potential does not significantly change
+the velocity of the star after the ejection. In their
+model, they relied on (dynamically driven) stellar mergers to explain the high masses of
+VFTS682 and the massive members of R136.%  While it is not possible to
+% robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
+% like fast rotation, which however could have already slowed down
+% because of the wind angular momentum losses during and after the
+% merger.
+
+To eject such a massive object, the cluster is
+expected to have produced
+a large number of massive runaways. Indeed, several %relatively
+isolated massive stars are observed in the region, some with known
+large radial velocities and/or proper motion. % (see
+% \Figref{fig:pm_polar} and Sana et al., in prep.).
+A comprehensive study of the kinematic
+properties of all the massive stars surrounding R136 might shed light
+on whether some can be unequivocally identified as merger products. It
+is also possible that the star or binary that caused the ejection of
+VFTS682 might have been ejected in the opposite direction, and is also
+isolated at present day. If the ejection was caused by an interaction
+with a binary, however, it is likely that the binary scattered in the
+opposite direction will experience further dynamical interactions on
+its way, modifying its trajectory and making it difficult to find.  %absence of proof would not undermine our conclusions
+
+The similarities between VFTS682 and the WNh5 stars in the core of
+R136 are also in agreement with the ``bully binary'' model of
+\cite{fujii:11}. Based on their numerical results, they suggested that
+early in the evolution of a cluster, dynamical interactions form an extremely
+massive binary, which then tightens its orbit by ejecting other stars passing
+by. Interpreting our results for VFTS682 through the lens of their simulations
+suggests the presence of a close binary with total mass
+$M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. Such bully
+binary could be R145 according to \cite{fujii:11}, and it might be an
+ideal observational candidate for a dynamically formed progenitor system of
+a binary black-hole, provided that stars this massive can avoid a
+pair-instability supernova \cite[e.g.,][]{rakavy:67} at LMC
+metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS682 could be either a
+pair-instability supernova without compact remnant formation, or
+possibly direct collapse to a black hole above the $2^\mathrm{nd}$
+mass gap. The amount of mass loss of these stars will determine their final core
+mass and thus their final fate.
+
+The kinematic age of VFTS682 puts an
+upper limit to the timescale to form the ``bully binary'' in
+R136. The cluster must have been at the very beginning of its
+evolution, given the age estimate of $\lesssim 2$\,Myr
+\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS682. If the
+cluster is indeed younger than the shortest stellar lifetime
+\citep[$\sim$3\,Myr, e.g.,][]{brott:11, zapartas:17}, then the alternative
+explanation for ejection of VFTS682 from the disruption of a binary
+by a core-collapse event is excluded since the region is too young for stars
+to have experienced core-collapse already.
+
+The variability of VFTS682, reminiscent of LBV stars, suggests
+that VFTS682 (and therefore its analogs in the core of R136) might
+experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
+debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
+claim that LBV stars are typically isolated form O-type stars. The fact that VFTS682 is a dynamically
+ejected runaway which might evolve into an LBV star suggests that
+N-body interactions also play a role in explaining the apparent
+isolation of at least some LBV stars.
+
+\citet{lennon:18} carried out a study similar to ours on the fast
+moving O-type stars
+in the region, and found two massive runaway stars
+($\sim$$90\,M_\odot$) in the 30 Doradus region. One of them (VFTS 16)
+was previously known as a runaway star from its line of sight velocity
+\citep[][]{evans:10}. \citet{lennon:18} also concluded that VFTS16 is 
+the result of a dynamical ejection from the R136 cluster, while the
+origin of the other star (VFTS 72) is less clear given its direction
+of motion. The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
+VFTS682 (see \Secref{sec:r136_origin}) is smaller than the
+corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic
+age of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
+massive VFTS682 was ejected later than VFTS16 from the same cluster.
+
+The numerical simulations from \cite{oh:16} suggest that dynamical
+interaction eject the majority of the stars during or shortly after the cluster
+core-collapse. The large number of isolated massive stars around it
+suggest that R136 has already evolved past the
+time of maximum stellar density. This might have implications for the
+question of whether the cluster formed via a monolithic collapse, or
+as a (potentially ongoing) merger of several sub-structures \citep[e.g.,][]{sabbi:12}.
+
+\citet{oh:16} also showed that the mass and velocity distribution of the ejected star depends on the cluster initial conditions
+(whether it is segregated, its primordial binary fraction and initial period
+distribution of the binary population), therefore studies on
+the population of isolated massive stars in the surroundings of R136
+might shed light on its initial stellar population and dynamical
+state.
+
+VFTS682 is potentially the most massive runaway known to date, and its ejection
+from the cluster R136 likely implies that it is only the ``tip of the
+iceberg'' of possibly extremely massive runaways in the
+region. Studies of this population, enabled by recent HST and \emph{Gaia} observations will put constraints on the evolution
+of these extreme stars, together with the formation and evolution of
+the central cluster itself.
+
+ \bibliographystyle{apj}
+ \bibliography{bibliography/vfts682,my_bib}
+
+%\begin{thebibliography}{}
+%\makeatletter
+%\relax
+%\def\mn@urlcharsother{\let\do\@makeother \do\$\do\&\do\#\do\^\do\_\do\%\do\~}
+%\def\mn@doi{\begingroup\mn@urlcharsother \@ifnextchar [ {\mn@doi@}
+%  {\mn@doi@[]}}
+%\def\mn@doi@[#1]#2{\def\@tempa{#1}\ifx\@tempa\@empty \href
+%  {http://dx.doi.org/#2} {doi:#2}\else \href {http://dx.doi.org/#2} {#1}\fi
+%  \endgroup}
+%\def\mn@eprint#1#2{\mn@eprint@#1:#2::\@nil}
+%\def\mn@eprint@arXiv#1{\href {http://arxiv.org/abs/#1} {{\tt arXiv:#1}}}
+%\def\mn@eprint@dblp#1{\href {http://dblp.uni-trier.de/rec/bibtex/#1.xml}
+%  {dblp:#1}}
+%\def\mn@eprint@#1:#2:#3:#4\@nil{\def\@tempa {#1}\def\@tempb {#2}\def\@tempc
+%  {#3}\ifx \@tempc \@empty \let \@tempc \@tempb \let \@tempb \@tempa \fi \ifx
+%  \@tempb \@empty \def\@tempb {arXiv}\fi \@ifundefined
+%  {mn@eprint@\@tempb}{\@tempb:\@tempc}{\expandafter \expandafter \csname
+%  mn@eprint@\@tempb\endcsname \expandafter{\@tempc}}}
+%
+%\bibitem[\protect\citeauthoryear{{Banerjee}, {Kroupa}  \& {Oh}}{{Banerjee}
+%  et~al.}{2012}]{banerjee:12}
+%{Banerjee} S.,  {Kroupa} P.,   {Oh} S.,  2012, \mn@doi [\apj]
+%  {10.1088/0004-637X/746/1/15}, \href
+%  {http://adsabs.harvard.edu/abs/2012ApJ...746...15B} {746, 15}
+%
+%\bibitem[\protect\citeauthoryear{{Bate}}{{Bate}}{2009}]{bate:09}
+%{Bate} M.~R.,  2009, \mn@doi [\mnras] {10.1111/j.1365-2966.2008.14106.x}, \href
+%  {http://adsabs.harvard.edu/abs/2009MNRAS.392..590B} {392, 590}
+%
+%\bibitem[\protect\citeauthoryear{{Bestenlehner} et~al.,}{{Bestenlehner}
+%  et~al.}{2011}]{bestenlehner:11}
+%{Bestenlehner} J.~M.,  et~al., 2011, \mn@doi [\aap]
+%  {10.1051/0004-6361/201117043}, \href
+%  {http://adsabs.harvard.edu/abs/2011A\%26A...530L..14B} {530, L14}
+%
+%\bibitem[\protect\citeauthoryear{{Blaauw}}{{Blaauw}}{1961}]{blaauw:61}
+%{Blaauw} A.,  1961, \bain, \href
+%  {http://adsabs.harvard.edu/abs/1961BAN....15..265B} {15, 265}
+%
+%\bibitem[\protect\citeauthoryear{{Bressert} et~al.,}{{Bressert}
+%  et~al.}{2012}]{bressert:12}
+%{Bressert} E.,  et~al., 2012, \mn@doi [\aap] {10.1051/0004-6361/201117247},
+%  \href {http://adsabs.harvard.edu/abs/2012A\%26A...542A..49B} {542, A49}
+%
+%\bibitem[\protect\citeauthoryear{{Brott} et~al.,}{{Brott}
+%  et~al.}{2011}]{brott:11}
+%{Brott} I.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201016113}, \href
+%  {http://adsabs.harvard.edu/abs/2011A\%26A...530A.115B} {530, A115}
+%
+%\bibitem[\protect\citeauthoryear{{Crowther}, {Schnurr}, {Hirschi}, {Yusof},
+%  {Parker}, {Goodwin}  \& {Kassim}}{{Crowther} et~al.}{2010}]{crowther:10}
+%{Crowther} P.~A.,  {Schnurr} O.,  {Hirschi} R.,  {Yusof} N.,  {Parker} R.~J.,
+%  {Goodwin} S.~P.,   {Kassim} H.~A.,  2010, \mn@doi [\mnras]
+%  {10.1111/j.1365-2966.2010.17167.x}, \href
+%  {http://adsabs.harvard.edu/abs/2010MNRAS.408..731C} {408, 731}
+%
+%\bibitem[\protect\citeauthoryear{{Crowther} et~al.,}{{Crowther}
+%  et~al.}{2016}]{crowther:16}
+%{Crowther} P.~A.,  et~al., 2016, \mn@doi [\mnras] {10.1093/mnras/stw273}, \href
+%  {http://adsabs.harvard.edu/abs/2016MNRAS.458..624C} {458, 624}
+%
+%\bibitem[\protect\citeauthoryear{{Davidson}, {Humphreys}  \& {Weis}}{{Davidson}
+%  et~al.}{2016}]{davidson:16}
+%{Davidson} K.,  {Humphreys} R.~M.,   {Weis} K.,  2016, arXiv:1608.02007, \href
+%  {http://adsabs.harvard.edu/abs/2016arXiv160802007D} {}
+%
+%\bibitem[\protect\citeauthoryear{{de Koter}, {Heap}  \& {Hubeny}}{{de Koter}
+%  et~al.}{1997}]{dekoter:97}
+%{de Koter} A.,  {Heap} S.~R.,   {Hubeny} I.,  1997, \mn@doi [\apj]
+%  {10.1086/303736}, \href {http://adsabs.harvard.edu/abs/1997ApJ...477..792D}
+%  {477, 792}
+%
+%  
+%\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
+%  et~al.}{2010}]{evans:10}
+%{Evans} C.~J.,  et~al., 2010, \mn@doi [\apjl] {10.1088/2041-8205/715/2/L74},
+%  \href {http://adsabs.harvard.edu/abs/2010ApJ...715L..74E} {715, L74}
+%
+%\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
+%  et~al.}{2011}]{evans:11}
+%{Evans} C.~J.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201116782},
+%  \href {http://adsabs.harvard.edu/abs/2011A\%26A...530A.108E} {530, A108}
+%
+%\bibitem[\protect\citeauthoryear{{Fujii} \& {Portegies Zwart}}{{Fujii} \&
+%  {Portegies Zwart}}{2011}]{fujii:11}
+%{Fujii} M.~S.,  {Portegies Zwart} S.,  2011, \mn@doi [Science]
+%  {10.1126/science.1211927}, \href
+%  {http://adsabs.harvard.edu/abs/2011Sci...334.1380F} {334, 1380}
+%
+%\bibitem[\protect\citeauthoryear{{Gaia Collaboration} et~al.,}{{Gaia
+%  Collaboration} et~al.}{2016}]{gaia:16}
+%{Gaia Collaboration} et~al., 2016, \mn@doi [\aap]
+%  {10.1051/0004-6361/201629272}, \href
+%  {http://adsabs.harvard.edu/abs/2016A\%26A...595A...1G} {595, A1}
+%
+%\bibitem[\protect\citeauthoryear{{Gaia Collaboration}, {Brown}, {Vallenari},
+%  {Prusti}, {de Bruijne}, {Babusiaux}  \& {Bailer-Jones}}{{Gaia Collaboration}
+%  et~al.}{2018}]{brown:18}
+%{Gaia Collaboration} {Brown} A.~G.~A.,  {Vallenari} A.,  {Prusti} T.,  {de
+%  Bruijne} J.~H.~J.,  {Babusiaux} C.,   {Bailer-Jones} C.~A.~L.,  2018,
+%  ArXiv:1804.09365, \href {http://adsabs.harvard.edu/abs/2018arXiv180409365G}
+%  {}
+%
+%\bibitem[\protect\citeauthoryear{{Gruendl} \& {Chu}}{{Gruendl} \&
+%  {Chu}}{2009}]{gruendl:09}
+%{Gruendl} R.~A.,  {Chu} Y.-H.,  2009, \mn@doi [\apjs]
+%  {10.1088/0067-0049/184/1/172}, \href
+%  {http://adsabs.harvard.edu/abs/2009ApJS..184..172G} {184, 172}
+%
+%\bibitem[\protect\citeauthoryear{{Gvaramadze}, {Weidner}, {Kroupa}  \&
+%  {Pflamm-Altenburg}}{{Gvaramadze} et~al.}{2012}]{gvaramadze:12}
+%{Gvaramadze} V.~V.,  {Weidner} C.,  {Kroupa} P.,   {Pflamm-Altenburg} J.,
+%  2012, \mn@doi [\mnras] {10.1111/j.1365-2966.2012.21452.x}, \href
+%  {http://adsabs.harvard.edu/abs/2012MNRAS.424.3037G} {424, 3037}
+%
+%% \bibitem[\protect\citeauthoryear{{Hogg}}{{Hogg}}{2018}]{hogg:18}
+%% {Hogg} D.~W.,  2018, ArXiv:1804.07766, \href
+%%   {http://adsabs.harvard.edu/abs/2018arXiv180407766H} {}
+%
+%\bibitem[\protect\citeauthoryear{{Humphreys}, {Weis}, {Davidson}  \&
+%  {Gordon}}{{Humphreys} et~al.}{2016}]{humphreys:16}
+%{Humphreys} R.~M.,  {Weis} K.,  {Davidson} K.,   {Gordon} M.~S.,  2016, \mn@doi
+%  [\apj] {10.3847/0004-637X/825/1/64}, \href
+%  {http://adsabs.harvard.edu/abs/2016ApJ...825...64H} {825, 64}
+%
+%\bibitem[\protect\citeauthoryear{{K{\"o}hler} et~al.,}{{K{\"o}hler}
+%  et~al.}{2015}]{kohler:15}
+%{K{\"o}hler} K.,  et~al., 2015, \mn@doi [\aap] {10.1051/0004-6361/201424356},
+%  \href {http://adsabs.harvard.edu/abs/2015A\%26A...573A..71K} {573, A71}
+%
+%\bibitem[\protect\citeauthoryear{{Kuiper}, {Yorke}  \& {Turner}}{{Kuiper}
+%  et~al.}{2015}]{kuiper:15}
+%{Kuiper} R.,  {Yorke} H.~W.,   {Turner} N.~J.,  2015, \mn@doi [\apj]
+%  {10.1088/0004-637X/800/2/86}, \href
+%  {http://adsabs.harvard.edu/abs/2015ApJ...800...86K} {800, 86}
+%
+%\bibitem[\protect\citeauthoryear{{Lada} \& {Lada}}{{Lada} \&
+%  {Lada}}{2003}]{lada:03}
+%{Lada} C.~J.,  {Lada} E.~A.,  2003, \mn@doi [\araa]
+%  {10.1146/annurev.astro.41.011802.094844}, \href
+%  {http://adsabs.harvard.edu/abs/2003ARA\%26A..41...57L} {41, 57}
+%
+%\bibitem[\protect\citeauthoryear{{Lamb}, {Oey}, {Segura-Cox}, {Graus},
+%  {Kiminki}, {Golden-Marx}  \& {Parker}}{{Lamb} et~al.}{2016}]{lamb:16}
+%{Lamb} J.~B.,  {Oey} M.~S.,  {Segura-Cox} D.~M.,  {Graus} A.~S.,  {Kiminki}
+%  D.~C.,  {Golden-Marx} J.~B.,   {Parker} J.~W.,  2016, \mn@doi [\apj]
+%  {10.3847/0004-637X/817/2/113}, \href
+%  {http://adsabs.harvard.edu/abs/2016ApJ...817..113L} {817, 113}
+%
+%\bibitem[\protect\citeauthoryear{{Langer}, {Norman}, {de Koter}, {Vink},
+%  {Cantiello}  \& {Yoon}}{{Langer} et~al.}{2007}]{langer:07}
+%{Langer} N.,  {Norman} C.~A.,  {de Koter} A.,  {Vink} J.~S.,  {Cantiello} M.,
+%  {Yoon} S.-C.,  2007, \mn@doi [\aap] {10.1051/0004-6361:20078482}, \href
+%  {http://adsabs.harvard.edu/abs/2007A\%26A...475L..19L} {475, L19}
+%
+%\bibitem[\protect\citeauthoryear{{Lennon} et~al.,}{{Lennon}
+%  et~al.}{2018}]{lennon:18}
+%{Lennon} D.~J.,  et~al., 2018, ArXiv:1805.08277, \href
+%  {http://adsabs.harvard.edu/abs/2018arXiv180508277L} {}
+%
+%\bibitem[\protect\citeauthoryear{{Lindegren} et~al.,}{{Lindegren}
+%  et~al.}{2018}]{lindengren:18}
+%{Lindegren} L.,  et~al., 2018, ArXiv:1804.09366, \href
+%  {http://adsabs.harvard.edu/abs/2018arXiv180409366L} {}
+%
+%% \bibitem[\protect\citeauthoryear{{Maeder} \& {Meynet}}{{Maeder} \&
+%%   {Meynet}}{2000}]{maeder:00}
+%% {Maeder} A.,  {Meynet} G.,  2000, \aap, \href
+%%   {http://adsabs.harvard.edu/abs/2000A\%26A...361..159M} {361, 159}
+%
+%\bibitem[\protect\citeauthoryear{{Oh} \& {Kroupa}}{{Oh} \&
+%  {Kroupa}}{2016}]{oh:16}
+%{Oh} S.,  {Kroupa} P.,  2016, \mn@doi [\aap] {10.1051/0004-6361/201628233},
+%  \href {http://adsabs.harvard.edu/abs/2016A\%26A...590A.107O} {590, A107}
+%
+%\bibitem[\protect\citeauthoryear{{Parker}}{{Parker}}{1993}]{parker:93}
+%{Parker} J.~W.,  1993, \mn@doi [\aj] {10.1086/116661}, \href
+%  {http://adsabs.harvard.edu/abs/1993AJ....106..560P} {106, 560}
+%
+%\bibitem[\protect\citeauthoryear{{Pietrzy{\'n}ski} et~al.,}{{Pietrzy{\'n}ski}
+%  et~al.}{2013}]{pietrzynski:13}
+%{Pietrzy{\'n}ski} G.,  et~al., 2013, \mn@doi [\nat] {10.1038/nature11878},
+%  \href {http://adsabs.harvard.edu/abs/2013Natur.495...76P} {495, 76}
+%
+%\bibitem[\protect\citeauthoryear{{Platais}, {van der Marel}, {Lennon},
+%  {Anderson}, {Bellini}, {Sabbi}, {Sana}  \& {Bedin}}{{Platais}
+%  et~al.}{2015}]{platais:15}
+%{Platais} I.,  {van der Marel} R.~P.,  {Lennon} D.~J.,  {Anderson} J.,
+%  {Bellini} A.,  {Sabbi} E.,  {Sana} H.,   {Bedin} L.~R.,  2015, \mn@doi [\aj]
+%  {10.1088/0004-6256/150/3/89}, \href
+%  {http://adsabs.harvard.edu/abs/2015AJ....150...89P} {150, 89}
+%
+%\bibitem[\protect\citeauthoryear{{Platais} et~al.,}{{Platais}
+%  et~al.}{2018}]{platais:18}
+%{Platais} I.,  et~al., 2018, ArXiv:1804.08678, \href
+%  {http://adsabs.harvard.edu/abs/2018arXiv180408678P} {}
+%
+%\bibitem[\protect\citeauthoryear{{Poveda}, {Ruiz}  \& {Allen}}{{Poveda}
+%  et~al.}{1967}]{poveda:67}
+%{Poveda} A.,  {Ruiz} J.,   {Allen} C.,  1967, Boletin de los Observatorios
+%  Tonantzintla y Tacubaya, \href
+%  {http://adsabs.harvard.edu/abs/1967BOTT....4...86P} {4, 86}
+%
+%\bibitem[\protect\citeauthoryear{{Rakavy} \& {Shaviv}}{{Rakavy} \&
+%  {Shaviv}}{1967}]{rakavy:67}
+%{Rakavy} G.,  {Shaviv} G.,  1967, \mn@doi [\apj] {10.1086/149204}, \href
+%  {http://adsabs.harvard.edu/abs/1967ApJ...148..803R} {148, 803}
+%
+%% \bibitem[\protect\citeauthoryear{{Ram{\'{\i}}rez-Agudelo}
+%%   et~al.,}{{Ram{\'{\i}}rez-Agudelo} et~al.}{2015}]{ramirez-agudelo:15}
+%% {Ram{\'{\i}}rez-Agudelo} O.~H.,  et~al., 2015, \mn@doi [\aap]
+%%   {10.1051/0004-6361/201425424}, \href
+%%   {http://adsabs.harvard.edu/abs/2015A\%26A...580A..92R} {580, A92}
+%
+%\bibitem[\protect\citeauthoryear{{Renzo} et~al.,}{{Renzo}
+%  et~al.}{2018}]{renzo:18}
+%{Renzo} M.,  et~al., 2018, ArXiv:1804.09164, \href
+%  {http://adsabs.harvard.edu/abs/2018arXiv180409164R} {}
+%
+%% \bibitem[\protect\citeauthoryear{{Robitaille} \& {Bressert}}{{Robitaille} \&
+%%   {Bressert}}{2012}]{robitaille:12}
+%% {Robitaille} T.,  {Bressert} E.,  2012, {APLpy: Astronomical Plotting Library
+%%   in Python}, Astrophysics Source Code Library (\mn@eprint {ascl} {1208.017})
+%
+%\bibitem[\protect\citeauthoryear{{Rosen}, {Krumholz}, {McKee}  \&
+%  {Klein}}{{Rosen} et~al.}{2016}]{rosen:16}
+%{Rosen} A.~L.,  {Krumholz} M.~R.,  {McKee} C.~F.,   {Klein} R.~I.,  2016,
+%  \mn@doi [\mnras] {10.1093/mnras/stw2153}, \href
+%  {http://adsabs.harvard.edu/abs/2016MNRAS.463.2553R} {463, 2553}
+%
+%\bibitem[\protect\citeauthoryear{{Rubio-D{\'{\i}}ez}, {Najarro},
+%  {Garc{\'{\i}}a}  \& {Sundqvist}}{{Rubio-D{\'{\i}}ez}
+%  et~al.}{2017}]{rubio-diez:17}
+%{Rubio-D{\'{\i}}ez} M.~M.,  {Najarro} F.,  {Garc{\'{\i}}a} M.,   {Sundqvist}
+%  J.~O.,  2017, in {Eldridge} J.~J.,  {Bray} J.~C.,  {McClelland} L.~A.~S.,
+%  {Xiao} L.,  eds,  IAU Symposium Vol. 329, The Lives and Death-Throes of
+%  Massive Stars. pp 131--135, \mn@doi{10.1017/S1743921317002447}
+%
+%\bibitem[\protect\citeauthoryear{{Sabbi} et~al.,}{{Sabbi}
+%  et~al.}{2012}]{sabbi:12}
+%{Sabbi} E.,  et~al., 2012, \mn@doi [\apjl] {10.1088/2041-8205/754/2/L37}, \href
+%  {http://adsabs.harvard.edu/abs/2012ApJ...754L..37S} {754, L37}
+%
+%\bibitem[\protect\citeauthoryear{{Sana}, {Ram{\'{\i}}rez-Tannus}, {de Koter},
+%  {Kaper}, {Tramper}  \& {Bik}}{{Sana} et~al.}{2017}]{sana:17}
+%{Sana} H.,  {Ram{\'{\i}}rez-Tannus} M.~C.,  {de Koter} A.,  {Kaper} L.,
+%  {Tramper} F.,   {Bik} A.,  2017, \mn@doi [\aap]
+%  {10.1051/0004-6361/201630087}, \href
+%  {http://adsabs.harvard.edu/abs/2017A\%26A...599L...9S} {599, L9}
+%
+%% \bibitem[\protect\citeauthoryear{{Schneider}, {Castro}, {Fossati}, {Langer}  \&
+%%   {de Koter}}{{Schneider} et~al.}{2017}]{schneider:17}
+%% {Schneider} F.~R.~N.,  {Castro} N.,  {Fossati} L.,  {Langer} N.,   {de Koter}
+%%   A.,  2017, \mn@doi [\aap] {10.1051/0004-6361/201628409}, \href
+%%   {http://adsabs.harvard.edu/abs/2017A\%26A...598A..60S} {598, A60}
+%
+%\bibitem[\protect\citeauthoryear{{Schneider} et~al.,}{{Schneider}
+%  et~al.}{2018}]{schneider:18}
+%{Schneider} F.~R.~N.,  et~al., 2018, \mn@doi [Science]
+%  {10.1126/science.aan0106}, \href
+%  {http://adsabs.harvard.edu/abs/2018Sci...359...69S} {359, 69}
+%
+%\bibitem[\protect\citeauthoryear{{Smith}}{{Smith}}{2016}]{smith:16}
+%{Smith} N.,  2016, \mn@doi [\mnras] {10.1093/mnras/stw1533}, \href
+%  {http://adsabs.harvard.edu/abs/2016MNRAS.461.3353S} {461, 3353}
+%
+%\bibitem[\protect\citeauthoryear{{Smith} \& {Tombleson}}{{Smith} \&
+%  {Tombleson}}{2015}]{smith:15}
+%{Smith} N.,  {Tombleson} R.,  2015, \mn@doi [\mnras] {10.1093/mnras/stu2430},
+%  \href {http://adsabs.harvard.edu/abs/2015MNRAS.447..598S} {447, 598}
+%
+%\bibitem[\protect\citeauthoryear{{Udalski} et~al.,}{{Udalski}
+%  et~al.}{2008}]{udalski:08}
+%{Udalski} A.,  et~al., 2008, \actaa, \href
+%  {http://adsabs.harvard.edu/abs/2008AcA....58..329U} {58, 329}
+%
+%\bibitem[\protect\citeauthoryear{{Ward} \& {Kruijssen}}{{Ward} \&
+%  {Kruijssen}}{2018}]{ward:18}
+%{Ward} J.~L.,  {Kruijssen} J.~M.~D.,  2018, \mn@doi [\mnras]
+%  {10.1093/mnras/sty117}, \href
+%  {http://adsabs.harvard.edu/abs/2018MNRAS.475.5659W} {475, 5659}
+%
+%\bibitem[\protect\citeauthoryear{{Zapartas} et~al.,}{{Zapartas}
+%  et~al.}{2017}]{zapartas:17}
+%{Zapartas} E.,  et~al., 2017, \mn@doi [\aap] {10.1051/0004-6361/201629685},
+%  \href {http://adsabs.harvard.edu/abs/2017A\%26A...601A..29Z} {601, A29}
+%
+%\bibitem[\protect\citeauthoryear{{Zinnecker} \& {Yorke}}{{Zinnecker} \&
+%  {Yorke}}{2007}]{zinnecker:07}
+%{Zinnecker} H.,  {Yorke} H.~W.,  2007, \mn@doi [\araa]
+%  {10.1146/annurev.astro.44.051905.092549}, \href
+%  {http://adsabs.harvard.edu/abs/2007ARA\%26A..45..481Z} {45, 481}
+%
+%\bibitem[\protect\citeauthoryear{{Zwicky}}{{Zwicky}}{1957}]{zwicky:57}
+%{Zwicky} F.,  1957, \zap, \href
+%  {http://adsabs.harvard.edu/abs/1957ZA.....44...64Z} {44, 64}
+%
+%
+%% \bibitem[\protect\citeauthoryear{{de Mink}, {Cantiello}, {Langer}, {Pols},
+%%   {Brott}  \& {Yoon}}{{de Mink} et~al.}{2009}]{demink:09}
+%% {de Mink} S.~E.,  {Cantiello} M.,  {Langer} N.,  {Pols} O.~R.,  {Brott} I.,
+%%   {Yoon} S.-C.,  2009, \mn@doi [\aap] {10.1051/0004-6361/200811439}, \href
+%%   {http://adsabs.harvard.edu/abs/2009A\%26A...497..243D} {497, 243}
+
+%\makeatother
+%\end{thebibliography}
+
+\section*{Affiliations}
+\noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
     Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
@@
   $^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
-}
-
- \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
-\date{}
-\begin{abstract}
-
- How very massive stars form is still among the most intriguing open
- questions in stellar astrophysics.  VFTS 682 is among the most
- extreme massive stars known, with an inferred initial mass of about
- $150\,M_\odot$.
-It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.
-The absence of other massive stars in its immediate vicinity lead to two intriguing possible hypotheses posed in earlier work.
-Either it formed in relative isolation through a new mode of star formation or it was ejected dynamically from the central cluster.
-
-Aiming to shine light on this debate, we investigate the kinematics of  VFTS 682 as obtained by \emph{Gaia} and a \emph{Hubble Space Telescope} proper motion campaign.
-We derive a projected velocity relative of to the cluster of $\sim$$30\pm20\kms$.
-The direction and magnitude of the proper motion are consistent with ejection from the central cluster considering its inferred age.
-However, the error bars and spread in proper motions derived for other sources in the region are substantial, which makes us cautious to derive hard conclusions.
-
-If future data confirms the runaway nature, this would make the VFTS682 the most massive runaway star known to date. 
-While we cannot prove this solidly from the current data, we do
-consider this hypothesis plausible. The central cluster is known to
-harbor several other stars more massive than $150\,M_\odot$ similar in
-spectral type.  Moreover, the current record holders are O type stars
-including VFTS 16 which provides direct evidence that R136 is indeed
-capable of ejecting some of its most massive members. 
-\end{abstract}
-
-\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS 682}
-\maketitle{}
-
-\section{Introduction}
-\label{sec:intro}
-
-How massive stars form is one of the major longstanding questions in astrophysics
-\citep[e.g.,][]{zinnecker:07}. Improving our understanding of massive star formation, and its
-possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies, but also for understanding the transients that mark their death and the compact remnants they leave behind.
-%
-Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
-% \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
-evolve fast, typically reside in dense groups and remain enshrouded in
-their parent cloud during the entirety of their formation
-process. Important progress has been made on the theoretical side,
-\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally very expensive and therefore remain challenging.
-
-In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to reside in the innermost cores.
-In this picture, field stars are primarily the result of the dissolution of dense groups.
-However, a significant population of massive stars exists in relative isolation, far from dense clusters or OB associations and their origin remains matter of debate \citep{gvaramadze:12, lamb:16,ward:18}.
-One hypothesis to explain the population of relatively isolated massive stars is that they formed in the field.
-The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
-Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.
-
-% A contribution to the debate on whether or not massive
-% stars form in relative isolation was presented by
-% \cite{bestenlehner:11} and \cite{bressert:12}, who discussed the case of the very massive star
-% VFTS682.
-
-One of the most extreme examples that has been considered in this debate is the very massive star VFTS682  \citep[][]{bestenlehner:11, bressert:12}.
-This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}.
-It is a hydrogen-rich Wolf-Rayet star of spectral type WNh5.
-Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $\sim$$140^{+30}_{-16}\,M_\odot$ corresponding to an initial mass of $\sim$$150^{+30}_{-17}\,M_\odot$
-\citep{schneider:18}. 
-%
-%$137.8^{+27.5}_{-15.9}\,M_\odot$,
-%
-This makes VFTS682 one of the most massive stars known and one of the most extreme objects in the region.
-%
-From the spectral point of view, it is reminiscent of the very
-massive stars % , a.k.a. ``monster stars'',
- in the core of the R136 cluster \citep{dekoter:97,crowther:10, crowther:16}. 
- %
- In particular, a remarkable similarity exist between the
-spectrum of VFTS682 and R136a3 for which \citet{crowther:16} report a
-current mass estimate of $180^{+30}_{-30}\Msun$. R136 hosts
-at least two more very massive WN5h stars, R136a1 and R136a2, whose
-estimated current masses are even higher.
-
-%at a projected offset of 119.4 arcseconds, corresponding to
-
-VFTS682 stands out by its relative isolation at a projected distance of 119.4 arcseconds, corresponding to 
-$\sim$$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
-considered two possible explanation for the offset: either
-the star formed in situ as an isolated massive star, % , but remark that this
-                                % poses a challenge for starformation
-                                % theories (They don't)
-or it was ejected from  R136. N-body simulations % of dynamical interaction in star clusters
-indicate that the ejection of very massive stars like VFTS682 is expected \citep[e.g.][]{fujii:11, banerjee:12}.
-This is supported by the recent findings of other massive runaway stars in the region based on proper motion studies.
-\citet{platais:15,platais:18} analyze multi-epoch \emph{Hubble Space Telescope} (HST) photometry and identify 10 stars, which appear to be ejected from R136.
-\citet{lennon:18} investigate the kinematics of  isolated O-type stars in the region using the second \emph{Gaia} data release \cite[DR2,][]{gaia:16,brown:18} and show that the proper motion, postion and direction of the $\sim100\Msun$ star VFTS16 is consistent with a runaway origin from R136.
-
-%Testing the origin of VFTS682 is not only important for constraining isolated star formation; it should also help to improve understanding of early star-cluster dynamics.
-
-In this paper we present an analysis of the new kinematical
-constraints for VFTS682 provided by \emph{Gaia} DR2 and constraints
-from HST proper data  by \citet{platais:18}.   We discuss the
-implications for the hypothesis that VFTS682 is a slow runaway star
-ejected from R136.
-% which was not part of the sample of \citet{lennon:18} because of its  spectral type.
-
-%%% cut for length reasons: all is said below
-% Our results indicate that VFTS682 is a runaway star, although with
-% low statistical significance. The
-% direction and magnitude of the velocity vector are consistent with
-% dynamical ejection from R136. The age inferred from the spectral
-% analysis \citep[from][]{schneider:18} is consistent with the travel
-% time we calculate for this star. % Therefore, the hypothesis that VFTS
-% % 682 is formed in relative isolation is rejected.
-% If our results are confirmed by future astrometric data releases, VFTS
-% 682, with an inferred present day mass well above a hundred solar
-% masses, is the most massive runaway star known to date.
-
-\section{Observations}
-\label{sec:sample}
-
-\begin{table}
-  \begin{center}
-    \caption{Observed and derived parameters for VFTS682. We also list
-    the average proper motion components of R136 derived by
-    \cite{lennon:18} for reference.}
-    \begin{tabular}{ll|c|c}
-      \hline
-      \hline
-      &Parameter & Value & Source\\
-
-       \hline
-       \multicolumn{4}{l}{\emph{Stellar Parameters VFTS 682}} \\
-     & present day $M$ \hfill $[M_\odot]$ & $137.8^{+27.5}_
-                                           {-15.9}$ &
-                                                     \multirow{3}{*}{\cite{schneider:18}}\\
-      &$M_\mathrm{ZAMS}$\hfill $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ &
-      \\
-      &age \hfill [Myr] & $1.0\pm0.2$ & \\[5pt]
-       \multicolumn{4}{l}{\emph{Average Proper motion in the region}} \\         &$\langle\mu_\mathrm{RA}\rangle$\hfill[$\mathrm{mas\ yr^{-1}}$] & $1.74\pm0.01$
-                        &  \multirow{2}{*}{\cite{lennon:18}}\\[5pt]
-      &$\langle\mu_\mathrm{DEC}\rangle$\hfill[$\mathrm{mas\ yr^{-1}}$]
-                & $0.70\pm0.02$ & \\
-  %    \multicolumn{4}{l}{\emph{Position}} \\
-
-    %  &RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
-   %                     & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
-    %  &DEC \hfill [degree] & -69.07 % $\pm$  0.05
-     %                   & \\[5pt]
-      \multicolumn{4}{l}{\emph{Relative Proper motion}} \\                 
-      &$\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \multirow{2}{*}{\emph{Gaia} DR2}\\[5pt]
-      &$\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
-
-      &$\delta\mu_\mathrm{RA}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.01$\pm$0.13 & \multirow{2}{*}{\cite{platais:18}}\\[5pt]
-      &$\delta\mu_\mathrm{DEC}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] &
-                                                                        0.2$\pm$0.1 &
-      \\[5pt]
-      \multicolumn{4}{l}{\emph{\SdM{ Motion in polar coordinates }}} \\     
-      &$\delta v_\mathrm{rad, R136}$  \hfill[$\kms$] & xxx & \cite{This Work}\\   
-       &$\theta_\mathrm{rad, R136}$  \hfill[$\kms$] & xxx & \cite{This Work}\\
-
-  %    &$\delta v_\mathrm{rad}$  \hfill[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
-
-      \hline
-
-    \end{tabular}
-    \tablecomments
-    {The error on the RA and DEC positions, are of order
-      $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$ in \emph{Gaia}
-      DR2. %REPETITION: %  We adopt the peculiar
-      % radial velocity calculated using the  NV $\lambda4944$ line, which is
-      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
-    }
-  \end{center}
-  \label{tab:vfts682}
-\end{table}
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
-  \includegraphics[width=0.315\textwidth]{figures/angle_dist}
-  \caption{Left panel: distribution of OB-type and Wolf-Rayet stars in proper
-    motion relative to R136. VFTS682 is not an outlier, but
-    its relative proper motion matches the expected value if it were indeed
-    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
-    assuming a distance of 50\,kpc. Right panel:  distribution of
-    angles between the relative proper motion direction and the radial
-    direction to the star. The error bars for VFTS682 are large, but
-    the best value is in agreement with the hypothesis of dynamical
-    ejection. In both
-    panels, the dark blue histograms contain 317 
-    stars with error smaller than $0.1\,\mathrm{mas \
-      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
-    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$. \todo{ Todo: [1] insert correct errorbars in right pabel, [2] make linewidth, shading and figure size of both panels consistent [3] fix normalization to be consistent in both pabels}}
-  \label{fig:mu_dist}
-\end{figure*}
-
-\subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
-
-The star VFTS682, located at right ascension (RA)
-05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000 \citep[][% see also
-% \Tabref{tab:vfts682} for the coordinates from \emph{Gaia} DR2
-]{evans:11}
-was originally classified as a young stellar object based on its
-mid-infrared excess \citep{gruendl:09}. \citet{evans:11} reclassified the
-object as Wolf-Rayet star of spectral type WNh5 using multi-epoch
-spectra covering $\lambda$4000--7000 from the
-VFTS survey. % The spectra available covered $\lambda$4000--7000 and
-% were taken at multiple epochs.
-
-Using the same dataset,
-\citet{bestenlehner:11} excluded the presence of a close companion
-with high confidence from the absence of radial velocity variations.
-\citet{bestenlehner:11} % further analyzed the spectra % using the non-LTE
-% % model atmosphere code CMFGEN \citep{hillier:98}
-also derived stellar
-parameters and surface abundances. They inferred an extinction
-of $A_V=4.45\pm0.12$, leading to the high luminosity
-$\log_{10}(L/L_\odot) =  6.5\pm0.2$. \citet{schneider:18} estimated
-a present-day mass for VFTS682 of $137.8^{+27.5}_ {-15.9}\Msun$, an
-apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass
-of $150.0^{+28.7}_{-17.4}\Msun$.% , using the Bayesian analysis tool BONNSAI
-% \citep{schneider:17}, based on the evolutionary models from
-% \citet{brott:11} and \cite{kohler:15}.
-% These values place VFTS682 around the
-% ``canonical upper limit'' of $\sim$$150\Msun$ by \citet{figer:05}.
-
-\citet{bestenlehner:11} reported for VFTS682 a wind mass loss rate of
-$10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, % inferred from
-% spectral analysis assuming a smooth outflow, i.e.,
-not accounting for
-clumping. The present-day surface helium mass
-fraction is $Y=0.45\, (0.49)$ from the spectral analysis of
-\cite{bestenlehner:11} (\citealt{rubio-diez:17}). This value can be
-used to constrain the age of the star and its past evolution.
-
-% The star is very similar from the spectral point of view to R136a3, located inside the star cluster
-% R136 \citep{crowther:10}  for which \citet{crowther:16} report a
-% current mass estimate of $180^{+30}_{-30}\Msun$. The R136 cluster hosts
-% at least two more very massive WN5h stars, R136a1 and R136a2, whose
-% estimated current masses are even higher. However, VFTS682 stands
-% out by its isolation at a projected offset of 29\,pc from the center
-% of the cluster \citep{bestenlehner:11}.
-
-Further worth noticing is the variability of the
-star. \citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude
-lower than found by \citet{evans:11}.
-\citet{bestenlehner:11} discuss the Optical Gravitational Lensing
-Experiment (OGLE-III) light curves \citep{udalski:08} and show a
-variability in the V-band at $\sim$10\% level on a timescale of years.
-They commented that this is unusual for Wolf-Rayet stars and more reminiscent
-of Luminous Blue Variable (LBV) stars. % \citep[e.g.,][]{langer:94}.
-
-\cite{bressert:12} note that VFTS682 shows a significant radial
-velocity offset compared to the nebular lines from the gas surrounding
-it, suggesting peculiar motion in the line of sight direction.
-\citet{bestenlehner:11} quote a radial velocity measured from the
-N{\footnotesize V} $\lambda4944$ of   $300\pm10\kms$. The average radial velocity of the 30 Doradus region is 
-$270\pm10\kms$.
-Whether this measurement truly reflects a peculiar radial velocity is debated.
-The determination of radial velocities for emission-lines stars is difficult because of the presence of an optically thick, non-homogeneous, and variable wind obscuring the photosphere.
-The radial velocity errors quoted above are only statistical, and do not quantify the (possibly larger)
-systematic errors.  \todo{Possibly mention and cite paper in prep by Paco and student on Xshooter data?  Possibly mention (or even use) on their measurement for the RV?}
-
-For simplicity, throughout this study, we assume the same
-distance of $50$\,kpc to the star \citep[][]{pietrzynski:13}, and to
-the 30 Doradus region as a whole. We do not consider the error on
-the distance determination ($\lesssim2\%$) when converting proper motions into
-physical velocities.
-%We implicitly assume that the use of a slightly different reference frame for $\delta v_\mathrm{rad}$ does not introduce significant errors.
-
-\subsection{HST astrometry for VFTS682}
-% observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
-% in October 3-8 2011 and the second epoch (GO-13359) in
-% October 6-11 2014, both using the F775W filter.
-% The second epoch was designed to match the pointing and
-% orientation of the first epoch as exactly as possible so
-% that the displacement measurements could be purely differential.
-
-For large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even exceed \emph{Gaia} DR2 data in quality.
-
-Recently, \citet{platais:18} presented HST astrometry and proper
-motions in the 30 Doradus region (GO-12499; P.I.: D.~J.~Lennon). The
-brightest stars (V$<$14) are missing from the catalogue because of
-saturation, but the high extinction towards VFTS682 makes is faint
-enough to be included. % in this catalogue.
-The  proper motions are
-listed in \Tabref{tab:vfts682}.
-
-The catalog from \citet{platais:18} lists the proper motion
-components of VFTS682 relative to the field, providing a
-measure of the motion of this star completely independent from
-\emph{Gaia}. They found proper
-motion components relative to the field
-$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
-  yr^{-1}}$ and
-$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
-  yr^{-1}}$, which added in quadrature give a relative proper motion of
-$\delta \mu^{HST} =0.20 \pm 0.15\,\mathrm{mas\ yr^{-1}}$.
-% \begin{equation}
-%   \label{eq:pm_around_HST}
-%   \delta \mu^{HST} = \sqrt{\left(\delta\mu_\mathrm{RA}^{HST}\right)^2+\left(\delta\mu_\mathrm{DEC}^{HST}\right)^2}
-%   = 0.20 \pm 0.15\,\mathrm{mas\
-%   yr^{-1}} \ \ .
-% \end{equation}
-
-These proper motion components can also be converted into projected
-velocities assuming a distance of 50\,kpc, obtaining $\delta
-v_\mathrm{RA}^{HST}=2\pm31\kms$ and $\delta
-v_\mathrm{RA}^{HST}=47\pm24\kms$% , corresponding to a two-dimensional
-% projected velocity of $47\pm24\kms$
-.
-Adding them in quadrature to the radial velocity
-measured by \citet{bestenlehner:11} we obtain a three-dimensional velocity of
-$v_\mathrm{pec}^{HST}=56 \pm 23  \kms$.
-% \begin{equation}
-%   \label{eq:speed_around_HST}
-%   v_\mathrm{pec}^{HST} = \sqrt{\left(\delta v_\mathrm{RA}^{HST}\right)^2
-%     +\left(\delta v_\mathrm{DEC}^{HST}\right)^2+\left(\delta
-%       v_\mathrm{rad}^{VFTS}\right)^2} = 56 \pm 23
-%   \kms \ .
-% \end{equation}
-
-\subsection{New data from \emph{Gaia} DR2  \label{data:gaia}}
-
-% \begin{table}
-%   \begin{center}
-%     \caption{Astrometric parameters for VFTS682. }
-%     \begin{tabular}{l|c|c}
-%       \hline
-%       \hline
-%       Parameter & Value & Source\\
-%       \hline
-%       RA \hfill[degree] &  \phantom{-}84.73 % $\pm$  0.03
-%                         & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
-%       DEC \hfill [degree] & -69.07 % $\pm$  0.05
-%                         & \\[5pt]
-%       \hline
-%       $\delta\mu_\mathrm{RA}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.10$\pm$0.07 & \multirow{2}{*}{\emph{Gaia} DR2}\\[5pt]
-%       $\delta\mu_\mathrm{DEC}^{Gaia}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.08$\pm$0.09 & \\[5pt]
-%       \hline
-%       $\delta\mu_\mathrm{RA}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] & 0.01$\pm$0.13 & \multirow{2}{*}{\cite{platais:18}}\\[5pt]
-%       $\delta\mu_\mathrm{DEC}^{HST}$  \hfill[$\mathrm{mas\ yr^{-1}}$] &
-%                                                                         0.2$\pm$0.1 &
-%       \\[5pt]
-%       \hline
-%       $\delta v_\mathrm{rad}$  \hfill[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
-%       \hline
-%     \end{tabular}
-%     \tablecomments
-%     {The error on the RA and DEC positions, are of order
-%       $\sim$$10^{-2}\,\mathrm{mas\ yr^{-1}}$ in \emph{Gaia}
-%       DR2. %REPETITION: %  We adopt the peculiar
-%       % radial velocity calculated using the  NV $\lambda4944$ line, which is
-%       % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
-%     }
-%   \end{center}
-%   \label{tab:vfts682}
-% \end{table}
-
-The second \emph{Gaia} Data Release (DR2) % , which
-% became available on 25 April 2018 and 
-provides positions,
-parallaxes, and proper motion components for more than a billion
-sources \citep{brown:18}. VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2
-catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}. Its
-G-band magnitude from \emph{Gaia}
-% reports a G-band magnitude of
-is 15.65. The star has a
-\texttt{visibility\_period} = 17, which counts how many observations have
-been used to reconstruct its astrometric solution
-\citep[][]{lindengren:18}, and the reported
-\texttt{astrometric\_excess\_noise} = 0. These values suggest that the \emph{Gaia}
-data for VFTS682 are reliable. % However, the effective temperature
-% reported in \emph{Gaia} DR2, based on the spectral fitting around the
-% CaII triplet, is one order of magnitude lower than what found by
-% \cite{bestenlehner:11}, and the best fit parallax of this star is
-% negative \citep[see, e.g.,][]{hogg:18}. We do not use the effective temperature of the star anywhere
-% in this study, and we attribute the unphysical value of the parallax
-% to the large distance to the LMC.
-% Our main findings do not rely on the
-% parallax nor the effective temperature values reported in the \emph{Gaia} DR2
-% catalog.
-
-While \emph{Gaia} DR2 data can be correlated, this is
-not an issue for VFTS682, and we treat the proper motion components ($\mu_\mathrm{RA}$, and $\mu_\mathrm{DEC}$,
-respectively) of the star as uncorrelated for simplicity.
-For the radial velocity of VFTS682 and of the 30 Doradus
-region as a whole, we instead use the VFTS data
-as quoted in \cite{bestenlehner:11}. \Tabref{tab:vfts682} lists the values adopted throughout
-this work for each of these quantities.
-
-We follow \citet{lennon:18} to define a local frame of reference and derive the peculiar velocity
-of VFTS682 with respect to the cluster R136. Their selection (see
-their Sec.~2.1) of 153
-bright ($G<17$) stars around R136 with reliable astrometric data from
-\emph{Gaia} DR2 yields $\langle\mu_\mathrm{RA}\rangle=1.74\pm0.01\,\mathrm{mas\
-  yr^{-1}}$ and $\langle\mu_\mathrm{DEC}\rangle=0.70\pm0.02\,\mathrm{mas\ yr^{-1}}$
-for the components of the mean motion of the region projected on the sky.
-
-% We define a local standard frame of reference to derive the peculiar velocity
-% of VFTS682 with respect to the cluster R136 by selecting from the \emph{Gaia} DR2 catalog following closely the approach of \cite{vandermarel:02,lennon:18}.
-% We select all the stars in a target of 0.05 degrees around R136
-% (NGC2070) fulfilling the following criteria. First, we require a G-band
-% magnitude brighter than 17, corresponding roughly to the
-% completeness level of the VFTS survey \citep[here we implicitly assume
-% G$\sim$V,][]{evans:11}. Then we require \texttt{visibility\_period} $\geq$ 5,
-% \texttt{astrometric\_excess\_noise} $< 1$, and the errors on the proper
-% motion components to be smaller than 0.07\,$\mathrm{mas\
-% yr^{-1}}$. At the distance to the LMC, $0.1\mathrm{mas\
-% yr^{-1}}\simeq25\,\kms$ \citep[e.g.,][]{platais:18}. We
-% remove obvious foreground stars by requiring the absolute value of
-% the reported parallaxes smaller than $0.1\,\mathrm{mas\ yr^{-1}}$.
-% At the distance to the
-% LMC, $1\mathrm{mas\ yr^{-1}}\simeq250\,\kms$ \citep[e.g.,][]{lennon:18}, so the cut on the values
-% of the proper motions removes stars that would have projected
-% tangential velocities in excess of $\sim$500\,$\kms$, which are most
-% likely to be foreground stars. We checked that the additional
-% requirement of having parallaxes smaller than $2\,\mathrm{mas}$ does
-% not reduces further our sample.
-% This selection yields a sample of 66 stars.% ,
-
-% We calculate the averaged proper motion components for the whole
-% region using 
-% \begin{equation}
-%   \label{eq:mean}
-%   \langle \mu_i\rangle = \frac{\sum_\mathrm{stars}\frac{1}{\Delta
-%   \mu_i^2}\mu_i}{\sum_\mathrm{stars} \frac{1}{\Delta \mu_i^2}} \ \ , \
-%   \ \Delta \langle \mu_i\rangle = \frac{\sqrt{N}}{\sum_\mathrm{stars}
-%   \frac{1}{\Delta \mu_i^2}} \ \ ,
-% \end{equation}
-% where $i = \mathrm{RA}, \mathrm{DEC}$, and $\Delta \mu_i$ is the error
-% on the proper motion component reported by \emph{Gaia}. The sums run over
-% all the $N=66$ stars in our selected sample. We evaluate each proper motion
-% component separately.
-
-\section{The kinematics of VFTS682}
-\label{sec:results}
-
-\subsection{Is it a runaway star?}
-\label{sec:runaway}
-We first address the question of whether VFTS682 is a typical star
-from the kinematic point of view, or whether it is a runaway star with
-a significantly large peculiar velocity compared to its surrounding population. The former is what should
-be expected if it formed in the relative isolation that we observe today.% , in relative
-% isolation from other massive stars.
-
-% Using the stars selected as described in \Secref{data:gaia} (a
-% subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
-% $\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
-% $\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
-% yr^{-1}}$. These values are
-% in good agreement with what is found by \cite{lennon:18}. However, we
-% emphasize their sensitivity to the choice of stars adopted to define
-% the reference frame.
-Subtracting the mean proper motion components given by
-\citet{lennon:18} from the
-proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
-components of proper motion of the star relative to the surrounding region
-$\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\,\mathrm{mas\ yr^{-1}}$
-and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08
-\pm 0.09\,\mathrm{mas\ yr^{-1}}$. These components result in a
-two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\
-  yr^{-1}}$.
-
-% \begin{equation}
-%   \label{eq:pm_gaia_around}
-%   \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
-%   = 0.13\pm 0.09\,\mathrm{mas\
-%   yr^{-1}} \ \ .
-% \end{equation}
-
-% Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
-% of the stars used in \cite{lennon:18} to define the frame of reference
-% we also adopt here, cf.~the right panel in their Fig.~2. The
-% proper motions are decomposed in the radial and tangential direction
-% from R136, to highlight the likelihood of it being the origin of fast
-% moving stars. The plus sign marks VFTS682, which sits at the edge of
-% the proper motion distribution, but is not a clear outlier. Most
-% relevant is the direction of the proper motion of VFTS682, which we
-% discuss in \Secref{sec:r136_origin}.
-
-% \begin{figure}%[htbp]
-%   \centering
-%   \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
-%   \caption{Polar plot of the relative proper motion components for the stars
-%     defining our reference frame. VFTS682 is indicated by the plus sign, and
-%     other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
-%     0.1, 0.2, and 0.4$\,\mathrm{mas\ yr^{-1}}$. Positive angle
-%     indicate a tangential component pointing in the counterclockwise
-%     direction, with 0 degrees corresponding to proper motion pointing
-%     radially outward from R136. The proper motion for VFTS682
-%     corresponds to almost radial motion away from the cluster.
-%   }
-%   \label{fig:pm_polar}
-% \end{figure}
-
-The proper motion components can be converted into
-the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=24\pm19\,\kms$,
-$\delta v_\mathrm{DEC}^{Gaia}=20\pm23\,\kms$, assuming a distance of
-50\,kpc. These can be combined obtaining a projected two-dimensional
-velocity of $31\pm21\kms$. % (we do not account for the uncertainty in the distance
-% estimate when propagating errors).
-The radial velocity from
-\cite{bestenlehner:11} then gives the third component along
-the line of sight, % allowing us to calculate the three-dimensional
-% peculiar speed of the star:
-which added in quadrature to the transverse components results in a
-three-dimensional velocity of $44 \pm 21\kms$.
-% \begin{equation}
-%   \label{eq:speed_around_Gaia}
-%   v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
-%     +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
-%       v_\mathrm{rad}^{VFTS}\right)^2} = 44 \pm 21
-%   \kms \ .
-% \end{equation}
-
-Therefore, two completely independent measures of the proper motion of
-VFTS682 relative to the surrounding field, one from HST and one from
-\emph{Gaia} DR2, yield values of the peculiar three-dimensional
-velocity of VFTS682 % (\Eqref{eq:speed_around_HST} and
-% \Eqref{eq:speed_around_Gaia}, respectively)
-which would make it the most massive runaway star
-known to date. However, the large errors on both the proper motion measures
-require confirmation with future astrometric data.
-
-\subsection{Does it come from the R136 cluster?}
-\label{sec:r136_origin}
-
-% Assuming that the best estimate of the peculiar velocity of VFTS682
-% are reliable, we now address the question of its likely origin.
-
-Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
-of OB-type and Wolf-Rayet stars included both in the VFTS survey and
-\emph{Gaia DR2} with reliable astrometric solutions. The bottom panel
-shows the distribution in angles $\theta$ between the relative proper motion
-direction and the direction from the core of R136 to the star. We
-emphasize that our subset of stars with reliable proper motion
-measurements is biased towards the fast moving objects, which results in a
-distribution of angles mildly peaked at small angles.
-
-VFTS682 is not an outlier in relative proper motion%  compared
-% to the two O-type runaways (re-)identified by \cite{lennon:18} (see
-% also below)
-. However, the star was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
-the dynamical ejection scenario. The green shade in
-\Figref{fig:mu_dist} shows the range of relative proper motions
-required to reach the present day location within
-the uncertainties on the apparent age of the star, assuming it was
-ejected from R136 very early in its life.
-
-Because of the large
-error bars, the angle between the relative proper motion is not very
-constraining, but it is suggestive that the best value is close to zero.
-Therefore, the relative proper motion from \emph{Gaia} DR2 are consistent with the hypothesis
-of dynamical ejection.
-
-The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
-VFTS682 from \emph{Gaia} DR2, and the
-yellow arrows illustrate the uncertainty. % show the possible
-% range of directions within the uncertainties in the measured relative proper
-% motion.
-These arrows cross at the present-day location of VFTS682 and
-are prolonged in the direction opposite to the motion to illustrate
-the possible range of origins. % The yellow arrow shows the direction of
-% the relative proper motion from HST (see \Eqref{eq:pm_around_HST}),
-% but we omit to show the possible range of directions from HST data, since it is
-% insufficiently constrained to be useful.
-
-% Although the large uncertainties on the relative proper motion
-% components results in a wide range of possible directions,we argue that the most likely origin of the star is R136.
-
-The kinematic age of this star, assuming it originates from R136, is
-
-\begin{equation}
-  \label{eq:kin_age}
-  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu^\mathrm{Gaia}} \simeq
-  \frac{119.4\,\mathrm{arcsec}}{0.13\,\mathrm{mas\ yr^{-1}}} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
-\end{equation}
-where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
-the core of the cluster \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
-As in the rest of this study, we neglect for
-simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
-The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
-ejection from the cluster, given is apparent age
-of $1.0\pm 0.2$\,Myr \citep{schneider:18}. % , which
-% corroborates the idea that the star is the result of a dynamical
-% ejection very soon in the cluster evolution.
-
-% the thing below is not relevant if it is a merger, better cut out
-% for now, since we are too long anyway
-
-%Note that the
-% present-day surface helium abundance
-% \citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
-% limit on the age of the star of $\sim$0.9\,Myr, corresponding to the time needed to
-% synthesize this amount of helium in the models from \cite{kohler:15}.
-
-\begin{figure}%[tbp]
-  \centering
-  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
-  \caption{The red solid arrow indicates the proper motion of VFTS682
-    relative to the region from \emph{Gaia} DR2, starting from the present day position of
-    the star. The yellow arrows indicate the possible
-    directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
-    backwards (dashed) to illustrate the uncertainty on the origin of the
-    star. The length of the prolongations is proportional to the relative proper motion
-    times the age of VFTS682 \citep[$1.0\pm0.2$\,Myr,][]{schneider:18}.
-    % \todo{load color picture on background? Fig with HST data
-    % available, but confusing...}
-  }
-
-  \label{fig:main}
-\end{figure}
-
-\section{Discussion}
-\label{sec:discussion}
-
-Based on our results, we tentatively claim that VFTS682 is the most massive
-runaway known to date, with a peculiar three-dimensional speed of
-$44\pm21\,\kms$. Due to the large error bars, this result will need
-to be revisited with future astrometric data. % to be revisited as updated
-% astrometric parameters from future \emph{Gaia} data releases become
-% available.
-If confirmed, it means that isolated star formation is
-\emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
-$0.9\pm0.6$\,Myr ago. Because of the exceptionally large mass
-of this star, this raises the question of which stars must populate
-the core of the cluster.
-
-Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
-massive star among those interacting \cite[e.g.,][]{banerjee:12}. This means that, just
-based on the kinematic properties of VFTS682, we would expect several
-stars with initial masses larger than $\sim$$150\,M_\odot$ in the
-cluster R136.
-This is consistent with the detection
-of extremely massive stars in the core of the
-cluster. % The projected rotational equatorial
-% velocity\footnote{However, the determination of the rotational
-%   velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
-%   considered with caution.} of VFTS682
-% reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
-% line with the average rotation rate of massive stars in the region
-% \citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
-% experienced rotationally induced chemically homogeneous evolution
-% \citep[][]{maeder:00,demink:09}, and (ii) it has not
-% accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
-% data of the VFTS survey rule out the presence of a close companion at the
-% present day. Moreover,
-
-The spectral type of VFTS682
-\citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
-three most massive stars detected in the core of the cluster%  by
-% \cite{dekoter:97,crowther:10,crowther:16}
-, with an astonishing similarity in particular with
-the spectrum of R136a3. Therefore, the isolation of
-VFTS682 makes it an ideal target to constrain the stellar physics of
-stars with masses well above $\sim$$100\,M_\odot$ while avoiding
-crowding issues. %  : its isolation makes
-% it an easier target for observations compared to the similar stars
-% present in the crowded core of R136.
-
-\citet{banerjee:12} used N-body simulations of fully segregated
-clusters with all massive stars in binaries to suggest that VFTS682
-was ejected from R136. They
-demonstrated that the cluster potential does not significantly change
-the velocity of the star after the ejection. In their
-model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS682 and the massive members of R136.%  While it is not possible to
-% robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
-% like fast rotation, which however could have already slowed down
-% because of the wind angular momentum losses during and after the
-% merger.
-
-To eject such a massive object, the cluster is
-expected to have produced
-a large number of massive runaways. Indeed, several %relatively
-isolated massive stars are observed in the region, some with known
-large radial velocities and/or proper motion. % (see
-% \Figref{fig:pm_polar} and Sana et al., in prep.).
-A comprehensive study of the kinematic
-properties of all the massive stars surrounding R136 might shed light
-on whether some can be unequivocally identified as merger products. It
-is also possible that the star or binary that caused the ejection of
-VFTS682 might have been ejected in the opposite direction, and is also
-isolated at present day. If the ejection was caused by an interaction
-with a binary, however, it is likely that the binary scattered in the
-opposite direction will experience further dynamical interactions on
-its way, modifying its trajectory and making it difficult to find.  %absence of proof would not undermine our conclusions
-
-The similarities between VFTS682 and the WNh5 stars in the core of
-R136 are also in agreement with the ``bully binary'' model of
-\cite{fujii:11}. Based on their numerical results, they suggested that
-early in the evolution of a cluster, dynamical interactions form an extremely
-massive binary, which then tightens its orbit by ejecting other stars passing
-by. Interpreting our results for VFTS682 through the lens of their simulations
-suggests the presence of a close binary with total mass
-$M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. Such bully
-binary could be R145 according to \cite{fujii:11}, and it might be an
-ideal observational candidate for a dynamically formed progenitor system of
-a binary black-hole, provided that stars this massive can avoid a
-pair-instability supernova \cite[e.g.,][]{rakavy:67} at LMC
-metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS682 could be either a
-pair-instability supernova without compact remnant formation, or
-possibly direct collapse to a black hole above the $2^\mathrm{nd}$
-mass gap. The amount of mass loss of these stars will determine their final core
-mass and thus their final fate.
-
-The kinematic age of VFTS682 puts an
-upper limit to the timescale to form the ``bully binary'' in
-R136. The cluster must have been at the very beginning of its
-evolution, given the age estimate of $\lesssim 2$\,Myr
-\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS682. If the
-cluster is indeed younger than the shortest stellar lifetime
-\citep[$\sim$3\,Myr, e.g.,][]{brott:11, zapartas:17}, then the alternative
-explanation for ejection of VFTS682 from the disruption of a binary
-by a core-collapse event is excluded since the region is too young for stars
-to have experienced core-collapse already.
-
-The variability of VFTS682, reminiscent of LBV stars, suggests
-that VFTS682 (and therefore its analogs in the core of R136) might
-experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
-debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
-claim that LBV stars are typically isolated form O-type stars. The fact that VFTS682 is a dynamically
-ejected runaway which might evolve into an LBV star suggests that
-N-body interactions also play a role in explaining the apparent
-isolation of at least some LBV stars.
-
-\citet{lennon:18} carried out a study similar to ours on the fast
-moving O-type stars
-in the region, and found two massive runaway stars
-($\sim$$90\,M_\odot$) in the 30 Doradus region. One of them (VFTS 16)
-was previously known as a runaway star from its line of sight velocity
-\citep[][]{evans:10}. \citet{lennon:18} also concluded that VFTS16 is 
-the result of a dynamical ejection from the R136 cluster, while the
-origin of the other star (VFTS 72) is less clear given its direction
-of motion. The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
-VFTS682 (see \Secref{sec:r136_origin}) is smaller than the
-corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic
-age of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
-massive VFTS682 was ejected later than VFTS16 from the same cluster.
-
-The numerical simulations from \cite{oh:16} suggest that dynamical
-interaction eject the majority of the stars during or shortly after the cluster
-core-collapse. The large number of isolated massive stars around it
-suggest that R136 has already evolved past the
-time of maximum stellar density. This might have implications for the
-question of whether the cluster formed via a monolithic collapse, or
-as a (potentially ongoing) merger of several sub-structures \citep[e.g.,][]{sabbi:12}.
-
-\citet{oh:16} also showed that the mass and velocity distribution of the ejected star depends on the cluster initial conditions
-(whether it is segregated, its primordial binary fraction and initial period
-distribution of the binary population), therefore studies on
-the population of isolated massive stars in the surroundings of R136
-might shed light on its initial stellar population and dynamical
-state.
-
-VFTS682 is potentially the most massive runaway known to date, and its ejection
-from the cluster R136 likely implies that it is only the ``tip of the
-iceberg'' of possibly extremely massive runaways in the
-region. Studies of this population, enabled by recent HST and \emph{Gaia} observations will put constraints on the evolution
-of these extreme stars, together with the formation and evolution of
-the central cluster itself.
-
-% \bibliographystyle{aa}
-% \bibliography{bibliography/vfts682}
-\begin{thebibliography}{}
-\makeatletter
-\relax
-\def\mn@urlcharsother{\let\do\@makeother \do\$\do\&\do\#\do\^\do\_\do\%\do\~}
-\def\mn@doi{\begingroup\mn@urlcharsother \@ifnextchar [ {\mn@doi@}
-  {\mn@doi@[]}}
-\def\mn@doi@[#1]#2{\def\@tempa{#1}\ifx\@tempa\@empty \href
-  {http://dx.doi.org/#2} {doi:#2}\else \href {http://dx.doi.org/#2} {#1}\fi
-  \endgroup}
-\def\mn@eprint#1#2{\mn@eprint@#1:#2::\@nil}
-\def\mn@eprint@arXiv#1{\href {http://arxiv.org/abs/#1} {{\tt arXiv:#1}}}
-\def\mn@eprint@dblp#1{\href {http://dblp.uni-trier.de/rec/bibtex/#1.xml}
-  {dblp:#1}}
-\def\mn@eprint@#1:#2:#3:#4\@nil{\def\@tempa {#1}\def\@tempb {#2}\def\@tempc
-  {#3}\ifx \@tempc \@empty \let \@tempc \@tempb \let \@tempb \@tempa \fi \ifx
-  \@tempb \@empty \def\@tempb {arXiv}\fi \@ifundefined
-  {mn@eprint@\@tempb}{\@tempb:\@tempc}{\expandafter \expandafter \csname
-  mn@eprint@\@tempb\endcsname \expandafter{\@tempc}}}
-
-\bibitem[\protect\citeauthoryear{{Banerjee}, {Kroupa}  \& {Oh}}{{Banerjee}
-  et~al.}{2012}]{banerjee:12}
-{Banerjee} S.,  {Kroupa} P.,   {Oh} S.,  2012, \mn@doi [\apj]
-  {10.1088/0004-637X/746/1/15}, \href
-  {http://adsabs.harvard.edu/abs/2012ApJ...746...15B} {746, 15}
-
-\bibitem[\protect\citeauthoryear{{Bate}}{{Bate}}{2009}]{bate:09}
-{Bate} M.~R.,  2009, \mn@doi [\mnras] {10.1111/j.1365-2966.2008.14106.x}, \href
-  {http://adsabs.harvard.edu/abs/2009MNRAS.392..590B} {392, 590}
-
-\bibitem[\protect\citeauthoryear{{Bestenlehner} et~al.,}{{Bestenlehner}
-  et~al.}{2011}]{bestenlehner:11}
-{Bestenlehner} J.~M.,  et~al., 2011, \mn@doi [\aap]
-  {10.1051/0004-6361/201117043}, \href
-  {http://adsabs.harvard.edu/abs/2011A\%26A...530L..14B} {530, L14}
-
-\bibitem[\protect\citeauthoryear{{Blaauw}}{{Blaauw}}{1961}]{blaauw:61}
-{Blaauw} A.,  1961, \bain, \href
-  {http://adsabs.harvard.edu/abs/1961BAN....15..265B} {15, 265}
-
-\bibitem[\protect\citeauthoryear{{Bressert} et~al.,}{{Bressert}
-  et~al.}{2012}]{bressert:12}
-{Bressert} E.,  et~al., 2012, \mn@doi [\aap] {10.1051/0004-6361/201117247},
-  \href {http://adsabs.harvard.edu/abs/2012A\%26A...542A..49B} {542, A49}
-
-\bibitem[\protect\citeauthoryear{{Brott} et~al.,}{{Brott}
-  et~al.}{2011}]{brott:11}
-{Brott} I.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201016113}, \href
-  {http://adsabs.harvard.edu/abs/2011A\%26A...530A.115B} {530, A115}
-
-\bibitem[\protect\citeauthoryear{{Crowther}, {Schnurr}, {Hirschi}, {Yusof},
-  {Parker}, {Goodwin}  \& {Kassim}}{{Crowther} et~al.}{2010}]{crowther:10}
-{Crowther} P.~A.,  {Schnurr} O.,  {Hirschi} R.,  {Yusof} N.,  {Parker} R.~J.,
-  {Goodwin} S.~P.,   {Kassim} H.~A.,  2010, \mn@doi [\mnras]
-  {10.1111/j.1365-2966.2010.17167.x}, \href
-  {http://adsabs.harvard.edu/abs/2010MNRAS.408..731C} {408, 731}
-
-\bibitem[\protect\citeauthoryear{{Crowther} et~al.,}{{Crowther}
-  et~al.}{2016}]{crowther:16}
-{Crowther} P.~A.,  et~al., 2016, \mn@doi [\mnras] {10.1093/mnras/stw273}, \href
-  {http://adsabs.harvard.edu/abs/2016MNRAS.458..624C} {458, 624}
-
-\bibitem[\protect\citeauthoryear{{Davidson}, {Humphreys}  \& {Weis}}{{Davidson}
-  et~al.}{2016}]{davidson:16}
-{Davidson} K.,  {Humphreys} R.~M.,   {Weis} K.,  2016, arXiv:1608.02007, \href
-  {http://adsabs.harvard.edu/abs/2016arXiv160802007D} {}
-
-\bibitem[\protect\citeauthoryear{{de Koter}, {Heap}  \& {Hubeny}}{{de Koter}
-  et~al.}{1997}]{dekoter:97}
-{de Koter} A.,  {Heap} S.~R.,   {Hubeny} I.,  1997, \mn@doi [\apj]
-  {10.1086/303736}, \href {http://adsabs.harvard.edu/abs/1997ApJ...477..792D}
-  {477, 792}
-
-\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
-  et~al.}{2010}]{evans:10}
-{Evans} C.~J.,  et~al., 2010, \mn@doi [\apjl] {10.1088/2041-8205/715/2/L74},
-  \href {http://adsabs.harvard.edu/abs/2010ApJ...715L..74E} {715, L74}
-
-\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
-  et~al.}{2011}]{evans:11}
-{Evans} C.~J.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201116782},
-  \href {http://adsabs.harvard.edu/abs/2011A\%26A...530A.108E} {530, A108}
-
-\bibitem[\protect\citeauthoryear{{Fujii} \& {Portegies Zwart}}{{Fujii} \&
-  {Portegies Zwart}}{2011}]{fujii:11}
-{Fujii} M.~S.,  {Portegies Zwart} S.,  2011, \mn@doi [Science]
-  {10.1126/science.1211927}, \href
-  {http://adsabs.harvard.edu/abs/2011Sci...334.1380F} {334, 1380}
-
-\bibitem[\protect\citeauthoryear{{Gaia Collaboration} et~al.,}{{Gaia
-  Collaboration} et~al.}{2016}]{gaia:16}
-{Gaia Collaboration} et~al., 2016, \mn@doi [\aap]
-  {10.1051/0004-6361/201629272}, \href
-  {http://adsabs.harvard.edu/abs/2016A\%26A...595A...1G} {595, A1}
-
-\bibitem[\protect\citeauthoryear{{Gaia Collaboration}, {Brown}, {Vallenari},
-  {Prusti}, {de Bruijne}, {Babusiaux}  \& {Bailer-Jones}}{{Gaia Collaboration}
-  et~al.}{2018}]{brown:18}
-{Gaia Collaboration} {Brown} A.~G.~A.,  {Vallenari} A.,  {Prusti} T.,  {de
-  Bruijne} J.~H.~J.,  {Babusiaux} C.,   {Bailer-Jones} C.~A.~L.,  2018,
-  ArXiv:1804.09365, \href {http://adsabs.harvard.edu/abs/2018arXiv180409365G}
-  {}
-
-\bibitem[\protect\citeauthoryear{{Gruendl} \& {Chu}}{{Gruendl} \&
-  {Chu}}{2009}]{gruendl:09}
-{Gruendl} R.~A.,  {Chu} Y.-H.,  2009, \mn@doi [\apjs]
-  {10.1088/0067-0049/184/1/172}, \href
-  {http://adsabs.harvard.edu/abs/2009ApJS..184..172G} {184, 172}
-
-\bibitem[\protect\citeauthoryear{{Gvaramadze}, {Weidner}, {Kroupa}  \&
-  {Pflamm-Altenburg}}{{Gvaramadze} et~al.}{2012}]{gvaramadze:12}
-{Gvaramadze} V.~V.,  {Weidner} C.,  {Kroupa} P.,   {Pflamm-Altenburg} J.,
-  2012, \mn@doi [\mnras] {10.1111/j.1365-2966.2012.21452.x}, \href
-  {http://adsabs.harvard.edu/abs/2012MNRAS.424.3037G} {424, 3037}
-
-% \bibitem[\protect\citeauthoryear{{Hogg}}{{Hogg}}{2018}]{hogg:18}
-% {Hogg} D.~W.,  2018, ArXiv:1804.07766, \href
-%   {http://adsabs.harvard.edu/abs/2018arXiv180407766H} {}
-
-\bibitem[\protect\citeauthoryear{{Humphreys}, {Weis}, {Davidson}  \&
-  {Gordon}}{{Humphreys} et~al.}{2016}]{humphreys:16}
-{Humphreys} R.~M.,  {Weis} K.,  {Davidson} K.,   {Gordon} M.~S.,  2016, \mn@doi
-  [\apj] {10.3847/0004-637X/825/1/64}, \href
-  {http://adsabs.harvard.edu/abs/2016ApJ...825...64H} {825, 64}
-
-\bibitem[\protect\citeauthoryear{{K{\"o}hler} et~al.,}{{K{\"o}hler}
-  et~al.}{2015}]{kohler:15}
-{K{\"o}hler} K.,  et~al., 2015, \mn@doi [\aap] {10.1051/0004-6361/201424356},
-  \href {http://adsabs.harvard.edu/abs/2015A\%26A...573A..71K} {573, A71}
-
-\bibitem[\protect\citeauthoryear{{Kuiper}, {Yorke}  \& {Turner}}{{Kuiper}
-  et~al.}{2015}]{kuiper:15}
-{Kuiper} R.,  {Yorke} H.~W.,   {Turner} N.~J.,  2015, \mn@doi [\apj]
-  {10.1088/0004-637X/800/2/86}, \href
-  {http://adsabs.harvard.edu/abs/2015ApJ...800...86K} {800, 86}
-
-\bibitem[\protect\citeauthoryear{{Lada} \& {Lada}}{{Lada} \&
-  {Lada}}{2003}]{lada:03}
-{Lada} C.~J.,  {Lada} E.~A.,  2003, \mn@doi [\araa]
-  {10.1146/annurev.astro.41.011802.094844}, \href
-  {http://adsabs.harvard.edu/abs/2003ARA\%26A..41...57L} {41, 57}
-
-\bibitem[\protect\citeauthoryear{{Lamb}, {Oey}, {Segura-Cox}, {Graus},
-  {Kiminki}, {Golden-Marx}  \& {Parker}}{{Lamb} et~al.}{2016}]{lamb:16}
-{Lamb} J.~B.,  {Oey} M.~S.,  {Segura-Cox} D.~M.,  {Graus} A.~S.,  {Kiminki}
-  D.~C.,  {Golden-Marx} J.~B.,   {Parker} J.~W.,  2016, \mn@doi [\apj]
-  {10.3847/0004-637X/817/2/113}, \href
-  {http://adsabs.harvard.edu/abs/2016ApJ...817..113L} {817, 113}
-
-\bibitem[\protect\citeauthoryear{{Langer}, {Norman}, {de Koter}, {Vink},
-  {Cantiello}  \& {Yoon}}{{Langer} et~al.}{2007}]{langer:07}
-{Langer} N.,  {Norman} C.~A.,  {de Koter} A.,  {Vink} J.~S.,  {Cantiello} M.,
-  {Yoon} S.-C.,  2007, \mn@doi [\aap] {10.1051/0004-6361:20078482}, \href
-  {http://adsabs.harvard.edu/abs/2007A\%26A...475L..19L} {475, L19}
-
-\bibitem[\protect\citeauthoryear{{Lennon} et~al.,}{{Lennon}
-  et~al.}{2018}]{lennon:18}
-{Lennon} D.~J.,  et~al., 2018, ArXiv:1805.08277, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180508277L} {}
-
-\bibitem[\protect\citeauthoryear{{Lindegren} et~al.,}{{Lindegren}
-  et~al.}{2018}]{lindengren:18}
-{Lindegren} L.,  et~al., 2018, ArXiv:1804.09366, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180409366L} {}
-
-% \bibitem[\protect\citeauthoryear{{Maeder} \& {Meynet}}{{Maeder} \&
-%   {Meynet}}{2000}]{maeder:00}
-% {Maeder} A.,  {Meynet} G.,  2000, \aap, \href
-%   {http://adsabs.harvard.edu/abs/2000A\%26A...361..159M} {361, 159}
-
-\bibitem[\protect\citeauthoryear{{Oh} \& {Kroupa}}{{Oh} \&
-  {Kroupa}}{2016}]{oh:16}
-{Oh} S.,  {Kroupa} P.,  2016, \mn@doi [\aap] {10.1051/0004-6361/201628233},
-  \href {http://adsabs.harvard.edu/abs/2016A\%26A...590A.107O} {590, A107}
-
-\bibitem[\protect\citeauthoryear{{Parker}}{{Parker}}{1993}]{parker:93}
-{Parker} J.~W.,  1993, \mn@doi [\aj] {10.1086/116661}, \href
-  {http://adsabs.harvard.edu/abs/1993AJ....106..560P} {106, 560}
-
-\bibitem[\protect\citeauthoryear{{Pietrzy{\'n}ski} et~al.,}{{Pietrzy{\'n}ski}
-  et~al.}{2013}]{pietrzynski:13}
-{Pietrzy{\'n}ski} G.,  et~al., 2013, \mn@doi [\nat] {10.1038/nature11878},
-  \href {http://adsabs.harvard.edu/abs/2013Natur.495...76P} {495, 76}
-
-\bibitem[\protect\citeauthoryear{{Platais}, {van der Marel}, {Lennon},
-  {Anderson}, {Bellini}, {Sabbi}, {Sana}  \& {Bedin}}{{Platais}
-  et~al.}{2015}]{platais:15}
-{Platais} I.,  {van der Marel} R.~P.,  {Lennon} D.~J.,  {Anderson} J.,
-  {Bellini} A.,  {Sabbi} E.,  {Sana} H.,   {Bedin} L.~R.,  2015, \mn@doi [\aj]
-  {10.1088/0004-6256/150/3/89}, \href
-  {http://adsabs.harvard.edu/abs/2015AJ....150...89P} {150, 89}
-
-\bibitem[\protect\citeauthoryear{{Platais} et~al.,}{{Platais}
-  et~al.}{2018}]{platais:18}
-{Platais} I.,  et~al., 2018, ArXiv:1804.08678, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180408678P} {}
-
-\bibitem[\protect\citeauthoryear{{Poveda}, {Ruiz}  \& {Allen}}{{Poveda}
-  et~al.}{1967}]{poveda:67}
-{Poveda} A.,  {Ruiz} J.,   {Allen} C.,  1967, Boletin de los Observatorios
-  Tonantzintla y Tacubaya, \href
-  {http://adsabs.harvard.edu/abs/1967BOTT....4...86P} {4, 86}
-
-\bibitem[\protect\citeauthoryear{{Rakavy} \& {Shaviv}}{{Rakavy} \&
-  {Shaviv}}{1967}]{rakavy:67}
-{Rakavy} G.,  {Shaviv} G.,  1967, \mn@doi [\apj] {10.1086/149204}, \href
-  {http://adsabs.harvard.edu/abs/1967ApJ...148..803R} {148, 803}
-
-% \bibitem[\protect\citeauthoryear{{Ram{\'{\i}}rez-Agudelo}
-%   et~al.,}{{Ram{\'{\i}}rez-Agudelo} et~al.}{2015}]{ramirez-agudelo:15}
-% {Ram{\'{\i}}rez-Agudelo} O.~H.,  et~al., 2015, \mn@doi [\aap]
-%   {10.1051/0004-6361/201425424}, \href
-%   {http://adsabs.harvard.edu/abs/2015A\%26A...580A..92R} {580, A92}
-
-\bibitem[\protect\citeauthoryear{{Renzo} et~al.,}{{Renzo}
-  et~al.}{2018}]{renzo:18}
-{Renzo} M.,  et~al., 2018, ArXiv:1804.09164, \href
-  {http://adsabs.harvard.edu/abs/2018arXiv180409164R} {}
-
-% \bibitem[\protect\citeauthoryear{{Robitaille} \& {Bressert}}{{Robitaille} \&
-%   {Bressert}}{2012}]{robitaille:12}
-% {Robitaille} T.,  {Bressert} E.,  2012, {APLpy: Astronomical Plotting Library
-%   in Python}, Astrophysics Source Code Library (\mn@eprint {ascl} {1208.017})
-
-\bibitem[\protect\citeauthoryear{{Rosen}, {Krumholz}, {McKee}  \&
-  {Klein}}{{Rosen} et~al.}{2016}]{rosen:16}
-{Rosen} A.~L.,  {Krumholz} M.~R.,  {McKee} C.~F.,   {Klein} R.~I.,  2016,
-  \mn@doi [\mnras] {10.1093/mnras/stw2153}, \href
-  {http://adsabs.harvard.edu/abs/2016MNRAS.463.2553R} {463, 2553}
-
-\bibitem[\protect\citeauthoryear{{Rubio-D{\'{\i}}ez}, {Najarro},
-  {Garc{\'{\i}}a}  \& {Sundqvist}}{{Rubio-D{\'{\i}}ez}
-  et~al.}{2017}]{rubio-diez:17}
-{Rubio-D{\'{\i}}ez} M.~M.,  {Najarro} F.,  {Garc{\'{\i}}a} M.,   {Sundqvist}
-  J.~O.,  2017, in {Eldridge} J.~J.,  {Bray} J.~C.,  {McClelland} L.~A.~S.,
-  {Xiao} L.,  eds,  IAU Symposium Vol. 329, The Lives and Death-Throes of
-  Massive Stars. pp 131--135, \mn@doi{10.1017/S1743921317002447}
-
-\bibitem[\protect\citeauthoryear{{Sabbi} et~al.,}{{Sabbi}
-  et~al.}{2012}]{sabbi:12}
-{Sabbi} E.,  et~al., 2012, \mn@doi [\apjl] {10.1088/2041-8205/754/2/L37}, \href
-  {http://adsabs.harvard.edu/abs/2012ApJ...754L..37S} {754, L37}
-
-\bibitem[\protect\citeauthoryear{{Sana}, {Ram{\'{\i}}rez-Tannus}, {de Koter},
-  {Kaper}, {Tramper}  \& {Bik}}{{Sana} et~al.}{2017}]{sana:17}
-{Sana} H.,  {Ram{\'{\i}}rez-Tannus} M.~C.,  {de Koter} A.,  {Kaper} L.,
-  {Tramper} F.,   {Bik} A.,  2017, \mn@doi [\aap]
-  {10.1051/0004-6361/201630087}, \href
-  {http://adsabs.harvard.edu/abs/2017A\%26A...599L...9S} {599, L9}
-
-% \bibitem[\protect\citeauthoryear{{Schneider}, {Castro}, {Fossati}, {Langer}  \&
-%   {de Koter}}{{Schneider} et~al.}{2017}]{schneider:17}
-% {Schneider} F.~R.~N.,  {Castro} N.,  {Fossati} L.,  {Langer} N.,   {de Koter}
-%   A.,  2017, \mn@doi [\aap] {10.1051/0004-6361/201628409}, \href
-%   {http://adsabs.harvard.edu/abs/2017A\%26A...598A..60S} {598, A60}
-
-\bibitem[\protect\citeauthoryear{{Schneider} et~al.,}{{Schneider}
-  et~al.}{2018}]{schneider:18}
-{Schneider} F.~R.~N.,  et~al., 2018, \mn@doi [Science]
-  {10.1126/science.aan0106}, \href
-  {http://adsabs.harvard.edu/abs/2018Sci...359...69S} {359, 69}
-
-\bibitem[\protect\citeauthoryear{{Smith}}{{Smith}}{2016}]{smith:16}
-{Smith} N.,  2016, \mn@doi [\mnras] {10.1093/mnras/stw1533}, \href
-  {http://adsabs.harvard.edu/abs/2016MNRAS.461.3353S} {461, 3353}
-
-\bibitem[\protect\citeauthoryear{{Smith} \& {Tombleson}}{{Smith} \&
-  {Tombleson}}{2015}]{smith:15}
-{Smith} N.,  {Tombleson} R.,  2015, \mn@doi [\mnras] {10.1093/mnras/stu2430},
-  \href {http://adsabs.harvard.edu/abs/2015MNRAS.447..598S} {447, 598}
-
-\bibitem[\protect\citeauthoryear{{Udalski} et~al.,}{{Udalski}
-  et~al.}{2008}]{udalski:08}
-{Udalski} A.,  et~al., 2008, \actaa, \href
-  {http://adsabs.harvard.edu/abs/2008AcA....58..329U} {58, 329}
-
-\bibitem[\protect\citeauthoryear{{Ward} \& {Kruijssen}}{{Ward} \&
-  {Kruijssen}}{2018}]{ward:18}
-{Ward} J.~L.,  {Kruijssen} J.~M.~D.,  2018, \mn@doi [\mnras]
-  {10.1093/mnras/sty117}, \href
-  {http://adsabs.harvard.edu/abs/2018MNRAS.475.5659W} {475, 5659}
-
-\bibitem[\protect\citeauthoryear{{Zapartas} et~al.,}{{Zapartas}
-  et~al.}{2017}]{zapartas:17}
-{Zapartas} E.,  et~al., 2017, \mn@doi [\aap] {10.1051/0004-6361/201629685},
-  \href {http://adsabs.harvard.edu/abs/2017A\%26A...601A..29Z} {601, A29}
-
-\bibitem[\protect\citeauthoryear{{Zinnecker} \& {Yorke}}{{Zinnecker} \&
-  {Yorke}}{2007}]{zinnecker:07}
-{Zinnecker} H.,  {Yorke} H.~W.,  2007, \mn@doi [\araa]
-  {10.1146/annurev.astro.44.051905.092549}, \href
-  {http://adsabs.harvard.edu/abs/2007ARA\%26A..45..481Z} {45, 481}
-
-\bibitem[\protect\citeauthoryear{{Zwicky}}{{Zwicky}}{1957}]{zwicky:57}
-{Zwicky} F.,  1957, \zap, \href
-  {http://adsabs.harvard.edu/abs/1957ZA.....44...64Z} {44, 64}
-
-% \bibitem[\protect\citeauthoryear{{de Mink}, {Cantiello}, {Langer}, {Pols},
-%   {Brott}  \& {Yoon}}{{de Mink} et~al.}{2009}]{demink:09}
-% {de Mink} S.~E.,  {Cantiello} M.,  {Langer} N.,  {Pols} O.~R.,  {Brott} I.,
-%   {Yoon} S.-C.,  2009, \mn@doi [\aap] {10.1051/0004-6361/200811439}, \href
-%   {http://adsabs.harvard.edu/abs/2009A\%26A...497..243D} {497, 243}
-
-\makeatother
-\end{thebibliography}
-
-% \begin{acknowledgements}
-%   \small
-%   We are grateful to S.~Torres, M.~C.~Ramirez-Tannus,
-%   and C.~J.~Evans for help and discussions.
-SdM has received funding under the European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC) (Grant agreement No.
-715063).
-VHB acknowledges support from the NRC-Canada Plaskett Fellowship.
-This work has made use of data from the ESA space mission \emph{Gaia} (\url{http://www.cosmos.esa.int/gaia}), processed by the \emph{Gaia} Data Processing and Analysis Consortium (DPAC, \url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}).
-Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \emph{Gaia} Multilateral Agreement.
+ \SdM{$^{x}$ {Affiliation: Chris Evans}\\
+  $^{x}$ {Affiliation: Paco Najorro}\\}
+
+ \begin{acknowledgements}
+   \small
+   We are grateful to S.\,Torres, M.\,C.\,Ramirez-Tannus,
+   and C.\,J.\,Evans for help and discussions, European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC), Grant agreement No.
+715063 [SdM], the NRC-Canada Plaskett Fellowship [VHB].
+
+This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
+Funding for the DPAC has been provided by national institutions, in particular the institutions
+participating in the {\it Gaia} Multilateral Agreement.
+
  % The background image of \Figref{fig:main} is based on observations
-%   % made with ESO Telescopes at the La Silla Observatory under programme
-%   % ID 076.C-0888, processed and released by the ESO VOS/ADP group.
-%   % This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
-% \end{acknowledgements}
+   % made with ESO Telescopes at the La Silla Observatory under programme
+   % ID 076.C-0888, processed and released by the ESO VOS/ADP group.
+   % This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
+ \end{acknowledgements}
 
 \end{document}
-
 %%% Local Variables:

### 2018-06-10 VFTS682_APJL.tex
table-inconsistencies removed

@@
       \multicolumn{5}{l}{\emph{Gaia's relative proper motion for VFTS 682 }} \\                 
-      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & 0.10$\pm$0.07 & (1,4) \\
-      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & 0.08$\pm$0.09 &  (1,4) \\
-      &$\delta\mu_\mathrm{}$  &[$\masyr$] & \SdM{xxx}$\pm$\SdM{xxx} &  (1,4) \\
-                 &$v_\mathrm{2d}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
+      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & 0.103$\pm$0.080 & (1,4) \\
+      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & 0.086$\pm$0.100 &  (1,4) \\
+      &$\delta\mu_\mathrm{}$  &[$\masyr$] & \SdM{0.13}$\pm$\SdM{0.09} &  (1,4) \\
+                 &$v_\mathrm{2d}$  &[$\kms$] & \SdM{32}$\pm$\SdM{21} & (1,4)\\  
  \hline     
@@
       &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & 0.01$\pm$0.13 & (3) \\
-      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & 0.2$\pm$0.1 &  (3) \\
-       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & 0.20$\pm$0.15 &  (3) \\
-                  &$v_\mathrm{2d, HST}$  &[$\kms$] & \SdM{$47\pm24$}\footnote{\SdM{This can't be right, the error bar is suddenly smaller than in the mas units!}} & (3)\\  
+      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & 0.20$\pm$0.10 &  (3) \\
+       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & 0.20$\pm$0.10 &  (3) \\
+                  &$v_\mathrm{2d, HST}$  &[$\kms$] & \SdM{$47\pm24$} & (3)\\  
 \hline
@@
 
-We list their data in \Tabref{tab:vfts682}.
-The errorbars are larger than we obtain from \emph{Gaia}, but the values are consistent.
-\SdM{Actually they are not consistent with values found in one of Danny's emails.
-What is going on here?}
+We list their data in \Tabref{tab:vfts682}.  The errorbars are comparable with those we obtain from \emph{Gaia}.  However the direction and angle are different.  
+\SdM{Can we please get theta Gaia and theta HST in the table.}
 

### 2018-06-10 VFTS682_APJL.tex
theta

@@
                  &$v_\mathrm{2d}$  &[$\kms$] & \SdM{32}$\pm$\SdM{21} & (1,4)\\  
+                     &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
+                         &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
+ %
+
  \hline     
@@
                   &$v_\mathrm{2d, HST}$  &[$\kms$] & \SdM{$47\pm24$} & (3)\\  
-\hline
-      \multicolumn{5}{l}{\emph{ Direction and magnitude w.r.t the vector pointing away from R136}} \\     
                       &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
                           &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
+ %
+
+%\hline
+     % \multicolumn{5}{l}{\emph{ Direction and magnitude w.r.t the vector pointing away from R136}} \\     
+ %      &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
+     %   &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
  %          &$v_\mathrm{2d}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\

### 2018-06-10 VFTS682_APJL.tex
merging

@@
 
-VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band \citep{Gaia-Collaboration+2016a, Gaia-Collaboration+2018}.
+VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band \citep{gaia:16,brown:18}.
 The number of visibility periods, i.e.

### 2018-06-10 VFTS682_APJL.tex
proof read sec.2 and filled tab 1

@@
 % \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
-evolve fast, typically reside in dense groups and remain enshrouded in
+evolve fast, typically reside in dense groups, and remain enshrouded in
 their parent cloud during the entirety of their formation
@@
 or it was ejected from  R136. N-body simulations % of dynamical interaction in star clusters
-indicate that the ejection of very massive stars like VFTS682 is expected \citep[e.g.][]{fujii:11, banerjee:12}.
-This is supported by the recent findings of other massive runaway stars in the region based on proper motion studies.
-\citet{platais:15,platais:18} analyze multi-epoch \emph{Hubble Space Telescope} (HST) photometry and identify 10 stars, which appear to be ejected from R136.
-\citet{lennon:18} investigate the kinematics of  isolated O-type stars in the region using the second \emph{Gaia} data release \cite[DR2,][]{gaia:16,brown:18} and show that the proper motion, postion and direction of the $\sim100\Msun$ star VFTS16 is consistent with a runaway origin from R136.
+indicate that the ejection of very massive stars like VFTS682 is
+expected \citep[e.g.][]{fujii:11, banerjee:12}.  This is supported by
+the recent findings of other massive runaway stars in the region based
+on proper motion studies.
+
+\citet{platais:15,platais:18} analyzed multi-epoch \emph{Hubble Space
+  Telescope} (HST) photometry and identified 10 stars likely ejected
+from R136.   \citet{lennon:18} investigate the kinematics of  isolated
+O-type stars in the region using the second \emph{Gaia} data release
+\cite[DR2,][]{gaia:16,brown:18} and show that the proper motion,
+postion and direction of the $\sim100\Msun$ star VFTS16 is consistent
+with a runaway origin from R136. They also found a less clear case for VFTS72.
 
@@
        \hline
-       \multicolumn{5}{l}{\emph{Selected stellar parameters VFTS 682}} \\
+       \multicolumn{5}{l}{\emph{Selected stellar parameters VFTS 682}}
+      \\
+      \hline
      & present day mass  & $[M_\odot]$ & $137.8^{+27.5}_
-                                           {-15.9}$ &
+                                           {-15.9}$ & (1)
                                                     \\
-      & initial mass& $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ &
+      & initial mass& $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ & (1)
       \\
-      &age & [Myr] & $1.0\pm0.2$ & \\
-      &mass loss rate &  & $\SdM{xxx}$ & \\
-      &helium abundance &  & $\SdM{xxx}$ & \\
+      &age & [Myr] & $1.0\pm0.2$ & (1) \\
+      &mass loss rate & [$M_\odot \ \mathrm{yr}^{-1}$] & $10^{-4.1\pm0.2}$ & (2)\\
+      &\multirow{2}{*}{surface helium abundance} &  & 0.45 & (2)\\
+      & &  & 0.49 & (3)\\
       \hline
@@
     { 
+      (1)~\cite{schneider:18}
+      (2)~\cite{bestenlehner:11}
+      (3)~\cite{rubio-diez:17}
     }
@@
        \hline
-           \multicolumn{5}{l}{\emph{Gaia's proper motion for VFTS 682 and average of the region}} \\
+           \multicolumn{5}{l}{\emph{Gaia's proper motion for VFTS 682
+      and average of the region}} \\
+      \hline
           &$\mu_\mathrm{RA}$&[$\masyr$] & $1.843\pm 0.070$
@@
      %                   & \\[5pt]
-      \multicolumn{5}{l}{\emph{Gaia's relative proper motion for VFTS 682 }} \\                 
+      \multicolumn{5}{l}{\emph{Gaia's relative proper motion for VFTS
+      682 }} \\
+      \hline
       &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & 0.103$\pm$0.080 & (1,4) \\
@@
  \hline     
-          \multicolumn{5}{l}{\emph{HST's independent determination of the relative proper motion for VFTS 682}} \\  
+          \multicolumn{5}{l}{\emph{HST's independent determination of
+      the relative proper motion for VFTS 682}} \\
+            \hline
       &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & 0.01$\pm$0.13 & (3) \\
@@
     { Proper motions are based on Gaia data unless indicated otherwise in subscript. The angle $\theta$ is defined such that $\theta=0$ corresponds to the direction pointing away from R136.
-    $^1$\cite{Gaia-Collaboration+2018}, 
-    $^2$\cite{lennon:18}, 
-    $^3$\cite{platais:18},  
-    $^4${\color{blue}This work}
+    (1) \cite{brown:18}, 
+    (2) \cite{lennon:18}, 
+    (3) \cite{platais:18},  
+    (4) {\color{blue}This work}
     %The error on the RA and DEC positions, are of order $\sim$$10^{-2}\masyr$ in \emph{Gaia} DR2. %REPETITION: %  We adopt the peculiar
@@
 
-The star WNh5 star VFTS682, located at right ascension (RA)
+The WNh5 star VFTS682, located at right ascension (RA)
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
@@
 The star is unlikely to have a close companion, because of the absence of clear radial velocity variations \citet{bestenlehner:11}.
-Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see Tab~ \label{tab:vfts682} for an overview of the parameters.
+Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see \Tabref{tab:star} for an overview of the parameters.
 

### 2018-06-10 VFTS682_APJL.tex
fixed bibliography

@@
 
-VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band \citep{gaia:16,brown:18}.
-The number of visibility periods, i.e.
-groups of observations separated from other groups by a gap of at least 4 days, used in the astrometric solution is 17 for this star.
-The reported astrometric excess noise is zero.
-\citep[][]{lindengren:18}.
-These values suggest that the \emph{Gaia} DR2  data for VFTS682 are reliable.
-Generally the Gaia errors are correlated, but the correlation coefficient is small for this star, see Tab.~\ref{tab:vfts682}.
-We can therefore treat the proper motion components as uncorrelated.
+VFTS682 is identified with the source id 4657685637907503744 in the
+\emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band
+\citep{gaia:16,brown:18}.   The number of visibility periods,
+i.e. groups of observations separated from other groups by a gap of at
+least 4 days, used in the astrometric solution is 17 for this
+star. The reported astrometric excess noise is zero.  These values
+suggest that the \emph{Gaia} DR2  data for VFTS682 are
+reliable. Generally the Gaia errors are correlated  \citep[][]{lindengren:18}, but the correlation coefficients are small for this star% , see Tab.~\ref{tab:vfts682}
+. We can therefore treat the proper motion components as uncorrelated.
 
 To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.
-They select  bright ($G<17$) stars within 0.05 degrees of R136 with and exclude sources with proper motion errors greater than $0.01\masyr$ in both co-ordinates (see their  Sect.~2.1 for further discussion).
-
-With these we compute the relative proper motion $\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.
-We also derive the angle $\theta$ between the direction of motion and the vector connecting the center of R136 with the current position of VFTS 682,  i.e.
-such that $\theta = 0$ is the angle expected for a star ejected from R136.
-We further provide the 2D velocity projected onto the radial vector pointing away from R136 $v_\mathrm{2d, \parallel}$  and the total projected 2d velocity $v_\mathrm{2d}$.
-All kinematic quantities are provided in Tab.~\ref{tab:vfts682}.
+They select  bright ($G<17$) stars within 0.05 degrees of R136 with and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1 for further discussion).
+
+With these we compute the relative proper motion
+$\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.  We also derive
+the angle $\theta$ between the direction of motion and the vector
+connecting the center of R136 with the current position of VFTS 682,
+i.e. such that $\theta = 0$ is the angle corresponding to perfectly
+radial motion away from R136.    We further provide the 2D velocity
+projected onto the radial vector pointing away from R136
+$v_\mathrm{2d, \parallel}$  and the total projected 2d velocity
+$v_\mathrm{2d}$.   All kinematic quantities are provided in
+Tab.~\ref{tab:vfts682}. %%what does the projection on the proper
+                        %%motion in the radial direction tell us?
+                        %%Seems like a useless number to me...
 
@@
 
-The 30 doradus region was target of an a two-epoch photometric campaign with \emph{HST} providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{Sabbi+2013}).
+The 30 doradus region was target of an a two-epoch photometric campaign with \emph{HST} providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
 Visually, the star is clearly isolated and does not appear to have any close neighbor.

### 2018-06-10 VFTS682_APJL.tex
added angle, stopping now

@@
                  &$v_\mathrm{2d}$  &[$\kms$] & \SdM{32}$\pm$\SdM{21} & (1,4)\\  
-                     &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
+                     &$\theta$  &[degrees] &  14.2$\pm$\SdM{xxx}  & (1,4)\\  
                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
@@
 suggest that the \emph{Gaia} DR2  data for VFTS682 are
-reliable. Generally the Gaia errors are correlated  \citep[][]{lindengren:18}, but the correlation coefficients are small for this star% , see Tab.~\ref{tab:vfts682}
+reliable. Generally, the Gaia errors are correlated
+\citep[][]{lindengren:18}, but for this star the correlation coefficient is small, see Tab.~\ref{tab:vfts682}
 . We can therefore treat the proper motion components as uncorrelated.

### 2018-06-11 VFTS682_APJL.tex
selma minor things

@@
       &Parameter & Units & Value & Ref.\\
+      \hline
+      \multicolumn{5}{l}{\emph{Coordinates}} \\
+      \hline
+         &$\mathrm{RA}_\mathrm{VFTS682}$&[degrees] & 84.7314 %$ \pm 0.036\,\mathrm{mas} $ 
+                     & (1) \\        
+               &$\mathrm{DEC}_\mathrm{VFTS682}$&[degrees] & -69.0741%$ \pm 0.048\,\mathrm{mas} $
+                     & (1)  \\
+
+                        &$\mathrm{RA}_\mathrm{R136}$&[degrees] & 84.6767
+                     &  SIMBAD  \\        
+               &$     \mathrm{DEC}_\mathrm{R136}$&[degrees] &  -69.1009
+                     &  SIMBAD \\       
+        &$      \delta\mathrm{RA}$  &[degrees] & 0.0547                      
+        &  \SdM{your value? }
+  \\        
+               &$     \delta\mathrm{DEC}$  &[degrees] &0.0268 
+                     &  \SdM{your value ? } \\
 
@@
                      &  (1) \\        
-                 & Correlation  ($\mu_\mathrm{RA}$, $\mu_\mathrm{DEC}$) &[$\masyr$] & $0.023$
+                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &[$\masyr$] & $0.023$
                         & (1)  \\         
@@
 
-    %  &RA &[degree] &  \phantom{-}84.73 % $\pm$  0.03
-   %                     & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
-    %  &DEC & [degree] & -69.07 % $\pm$  0.05
-     %                   & \\[5pt]
       \multicolumn{5}{l}{\emph{Gaia's relative proper motion for VFTS
@@
       \hline
-      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & 0.103$\pm$0.080 & (1,4) \\
-      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & 0.086$\pm$0.100 &  (1,4) \\
-      &$\delta\mu_\mathrm{}$  &[$\masyr$] & \SdM{0.13}$\pm$\SdM{0.09} &  (1,4) \\
-                 &$v_\mathrm{2d}$  &[$\kms$] & \SdM{32}$\pm$\SdM{21} & (1,4)\\  
-                     &$\theta$  &[degrees] &  14.2$\pm$\SdM{xxx}  & (1,4)\\  
-                         &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
- %
-
- \hline     
-          \multicolumn{5}{l}{\emph{HST's independent determination of
-      the relative proper motion for VFTS 682}} \\
-            \hline
-      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & 0.01$\pm$0.13 & (3) \\
-      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & 0.20$\pm$0.10 &  (3) \\
-       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & 0.20$\pm$0.10 &  (3) \\
-                  &$v_\mathrm{2d, HST}$  &[$\kms$] & \SdM{$47\pm24$} & (3)\\  
-                      &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
-                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
+      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & $0.103\pm0.080$ & (1,4) \\
+      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & $0.086\pm0.100$ &  (1,4) \\
+      &$\delta\mu_\mathrm{}$  &[$\masyr$] & $0.13\pm0.09$ &  (1,4) \\
+                 &$v_\mathrm{2d}$  &[$\kms$] & \SdM{$31\pm 21$} (not $32\pm21$!) & (1,4)\\  
+                     &$\theta$  &[degrees] &  \SdM{$14.2_{-xx}^{+xx}$}  & (1,4)\\  
+                         &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,4)\\  
+ %
+
+ \hline     
+          \multicolumn{5}{l}{\emph{HST's relative proper motion for VFTS 682}} \\
+            \hline
+      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & $0.01\pm0.13$ & (3) \\
+      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (3) \\
+       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (3) \\
+                  &$v_\mathrm{2d, HST}$  &[$\kms$] & $47\pm24$ & (3)\\  
+                      &$\theta^{*}$  &[degrees] &   \SdM{$xx_{-xx}^{+xx}$}   & (1,4)\\  
+                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,4)\\  
  %
@@
     \tablecomments
-    { Proper motions are based on Gaia data unless indicated otherwise in subscript. The angle $\theta$ is defined such that $\theta=0$ corresponds to the direction pointing away from R136.
+    { Proper motions are based on Gaia data unless indicated otherwise in subscript.
+The conversion factor assuming a distance of 50\,kpc is such that $1\masyr$ corresponds to $237\kms$.
+The correlation coefficient is indicated as $\rho$.
+The angle $\theta$ is defined such that $\theta=0$ for motions pointing away radially from R136.
     (1) \cite{brown:18}, 
     (2) \cite{lennon:18}, 
-    (3) \cite{platais:18},  
-    (4) {\color{blue}This work}
+    (3) \cite{platais:18} and
+    (4) {\color{blue}This work}.
     %The error on the RA and DEC positions, are of order $\sim$$10^{-2}\masyr$ in \emph{Gaia} DR2. %REPETITION: %  We adopt the peculiar
@@
   \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
-  \includegraphics[width=0.315\textwidth]{figures/angle_dist}
+  \includegraphics[width=0.35\textwidth]{figures/angle_dist}
   \caption{Left panel: distribution of OB-type and Wolf-Rayet stars in proper
@@
     lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$. \todo{ Todo: [1] insert correct errorbars in right pabel, [2] make linewidth, shading and figure size of both panels consistent [3] fix normalization to be consistent in both pabels}}
+      yr^{-1}}$.
+\SdM{ Todo: [1] insert correct errorbars in right pabel, [2] make (a) linewidth, (b) line width of the hashed shading and (c)  figure height consistent.
+(You could try labeling the top x axis in the right panel in degrees.)  [3] fix normalization to be consistent in both pabels, [4] unit in axis label in righthand plot is not correct.
+Degrees but radian.
+[4] Legen: the rel error limits that we adopted are missing a unit.
+Milia arrsecond?
+}}
   \label{fig:mu_dist}
@@
 
-The distance to 30 Doradus is $50$\,kpc  \citep[][]{pietrzynski:13}.
-The error on the distance determination is small ($\lesssim2\%$) and the possible difference between VFTS682 and the average of the region is probably even smaller (likely $\ll 1\%$).
+For the distance we adopt the distance to the LMC of $50$\,kpc.
+The error on the distance determination is small ($\lesssim2\%$, \citealt{pietrzynski:13}) and any possible offset between  R136 or VFTS682  and the distance we adopted for the LMC is probably much smaller (likely $\ll 1\%$, e.g.
+\citealt{Luks+1992}).
 These uncertainties are negligible compared to the errors in the proper motion discussed below.
@@
 i.e. groups of observations separated from other groups by a gap of at
-least 4 days, used in the astrometric solution is 17 for this
+least four days, used in the astrometric solution is seventeen for this
 star. The reported astrometric excess noise is zero.  These values
 suggest that the \emph{Gaia} DR2  data for VFTS682 are
-reliable. Generally, the Gaia errors are correlated
-\citep[][]{lindengren:18}, but for this star the correlation coefficient is small, see Tab.~\ref{tab:vfts682}
-. We can therefore treat the proper motion components as uncorrelated.
+reliable.
+Gaia errors are correlated \citep[][]{lindengren:18}, but for this star the correlation coefficient is very small, $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$  = 0.023, where zero is for completely un-correlated data and one is for fully correlated data.
+We can thus safely treat the proper motion components as uncorrelated.
 
@@
 
-We list their data in \Tabref{tab:vfts682}.  The errorbars are comparable with those we obtain from \emph{Gaia}.  However the direction and angle are different.  
-\SdM{Can we please get theta Gaia and theta HST in the table.}
+We list their data in \Tabref{tab:vfts682}.
+
+%The errorbars are comparable with those we obtain from \emph{Gaia}.  However the direction and angle are different.
 
@@
 
+\SdM{Since the errors for HST and Gaia are so similar and yield to different directions, I think we should probably restructure this paragraph.
+I liked how you break down the story in two separate questions, but given Fig 1 I think it may not work well.
+The first question should have adressed it in context of the motions of the other stars.
+And it has a very typical motion.
+I think that instead, you should simply describe Fig 1 step by step.
+I think that will be stronger.
+}
+
 \subsection{Is it a runaway star?}
 \label{sec:runaway}
-We first address the question of whether VFTS682 is a typical star
-from the kinematic point of view, or whether it is a runaway star with
-a significantly large peculiar velocity compared to its surrounding population. The former is what should
-be expected if it formed in the relative isolation that we observe today.% , in relative
-% isolation from other massive stars.
-
-% Using the stars selected as described in \Secref{data:gaia} (a
-% subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
-% $\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\masyr$ and
-% $\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
-% yr^{-1}}$. These values are
-% in good agreement with what is found by \cite{lennon:18}. However, we
-% emphasize their sensitivity to the choice of stars adopted to define
-% the reference frame.
-Subtracting the mean proper motion components given by
-\citet{lennon:18} from the
-proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
-components of proper motion of the star relative to the surrounding region
-$\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\masyr$
-and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08
-\pm 0.09\masyr$. These components result in a
-two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\
-  yr^{-1}}$.
+We first address the question whether VFTS682 is a typical star
+from the kinematic point of view (which is what would be expected if it formed in relative isolation), or whether has a significant peculiar velocity compared to its surrounding population (which is what we would expect if the star has a runaway origin).
+
+After subtracting the mean proper motion of the field, we obtain the relative two-dimensional relative proper motion.
+We obtain  $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$   and $\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$.
+
+Both indicate a non-zero peculiar 2D motion that would make it a slow runaway star, but in both cases the meaurements are also consistent with zero within 2 sigma.
+
+Figure~\ref{fig:mu_dist} shows the distribution of relative proper motions for other VFTS O, B and WR stars subselecting either those with relative proper motion errors smaller than $0.1\masyr$ (yielding 317 stars) or a more restrictive cut including only those with proper motion errors smaller than $0.05\masyr$ (yielding 36) stars.
+We see that the derived Gaia proper motion is very typical for stars in the field.
+
+\SdM{XXX We need to rethink }
+
+%Therefore, two completely independent measures of the proper motion of VFTS682 relative to the surrounding field, one from HST and one from \emph{Gaia} DR2, yield values of the peculiar three-dimensional velocity of VFTS682 % (\Eqref{eq:speed_around_HST} and
+% \Eqref{eq:speed_around_Gaia}, respectively)
+%which would make it the most massive runaway star known to date. However, the large errors on both the proper motion measures require confirmation with future astrometric data.
+
+.
+
+%Subtracting the mean proper motion components given by \citet{lennon:18} from the proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the components of proper motion of the star relative to the surrounding region $\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\masyr$ and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08 \pm 0.09\masyr$.
+These components result in a two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\ yr^{-1}}$.
 
@@
 
-% \begin{figure}%[htbp]
-%   \centering
-%   \includegraphics[width=0.5\textwidth]{figures/figure_polar_682-1.pdf}
-%   \caption{Polar plot of the relative proper motion components for the stars
-%     defining our reference frame. VFTS682 is indicated by the plus sign, and
-%     other notable outliers \citep[see][]{lennon:18} are also labeled. Concentric dashed-lines denote relative proper motions of
-%     0.1, 0.2, and 0.4$\masyr$. Positive angle
-%     indicate a tangential component pointing in the counterclockwise
-%     direction, with 0 degrees corresponding to proper motion pointing
-%     radially outward from R136. The proper motion for VFTS682
-%     corresponds to almost radial motion away from the cluster.
-%   }
-%   \label{fig:pm_polar}
-% \end{figure}
-
-The proper motion components can be converted into
-the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=24\pm19\,\kms$,
-$\delta v_\mathrm{DEC}^{Gaia}=20\pm23\,\kms$, assuming a distance of
-50\,kpc. These can be combined obtaining a projected two-dimensional
-velocity of $31\pm21\kms$. % (we do not account for the uncertainty in the distance
-% estimate when propagating errors).
-The radial velocity from
-\cite{bestenlehner:11} then gives the third component along
-the line of sight, % allowing us to calculate the three-dimensional
-% peculiar speed of the star:
-which added in quadrature to the transverse components results in a
-three-dimensional velocity of $44 \pm 21\kms$.
+%The proper motion components can be converted into the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=24\pm19\,\kms$, $\delta v_\mathrm{DEC}^{Gaia}=20\pm23\,\kms$, assuming a distance of 50\,kpc.
+These can be combined obtaining a projected two-dimensional velocity of $31\pm21\kms$.
+
+%The radial velocity from \cite{bestenlehner:11} then gives the third component along the line of sight,  which added in quadrature to the transverse components results in a three-dimensional velocity of $44 \pm 21\kms$.
 % \begin{equation}
@@
 % \end{equation}
-
-Therefore, two completely independent measures of the proper motion of
-VFTS682 relative to the surrounding field, one from HST and one from
-\emph{Gaia} DR2, yield values of the peculiar three-dimensional
-velocity of VFTS682 % (\Eqref{eq:speed_around_HST} and
-% \Eqref{eq:speed_around_Gaia}, respectively)
-which would make it the most massive runaway star
-known to date. However, the large errors on both the proper motion measures
-require confirmation with future astrometric data.
 
@@
   \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
-  \caption{The red solid arrow indicates the proper motion of VFTS682
+  \caption{\SdM{I don't see a reason why we would not show the HST data, except for wanting to hide it maybe.
+Let's sit together and find a good way to plot both.
+(Hope you still have the script to show the HST direction.
+)}  The red solid arrow indicates the proper motion of VFTS682.
     relative to the region from \emph{Gaia} DR2, starting from the present day position of

### 2018-06-11 VFTS682_APJL.tex
fixed typo in correlation coefficient in tab.2
reverted to non-normalized plots in fig.1, it is so much clearer for me what is happening

@@
                      &  (1) \\        
-                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &[$\masyr$] & $0.023$
+                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &[$\masyr$] & $0.026$
                         & (1)  \\         
@@
   \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
-  \includegraphics[width=0.35\textwidth]{figures/angle_dist}
+  \includegraphics[width=0.35\textwidth]{figures/angle}
   \caption{Left panel: distribution of OB-type and Wolf-Rayet stars in proper
@@
 reliable.
-Gaia errors are correlated \citep[][]{lindengren:18}, but for this star the correlation coefficient is very small, $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$  = 0.023, where zero is for completely un-correlated data and one is for fully correlated data.
+Gaia errors are correlated \citep[][]{lindengren:18}, but for this star the correlation coefficient is very small, $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$  = 0.026, where zero is for completely un-correlated data and one is for fully correlated data.
 We can thus safely treat the proper motion components as uncorrelated.
@@
 To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.
-They select  bright ($G<17$) stars within 0.05 degrees of R136 with and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1 for further discussion).
+They select  bright ($G<17$) stars within 0.05 degrees of R136 and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1 for further discussion).
 
@@
 
-The 30 doradus region was target of an a two-epoch photometric campaign with \emph{HST} providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
+The 30 doradus region was target of an a two-epoch photometric campaign with HST providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
 Visually, the star is clearly isolated and does not appear to have any close neighbor.
@@
 
- \citet{platais:18} recently analyzed the HST data for candidate runaway stars.
-The brightest stars (V$<$14) are missing from the catalogue because of saturation effects, but the high extinction towards VFTS682 makes it faint enough to be included.
- Our star did not make the more conservative selection criteria for runaway stars adopted in that study.
-However, the measurements are useful here, since they provides a measurement of the proper motion of VFTS 682 that is completely independent from the \emph{Gaia} measurement.
+ \citet{platais:18} recently analyzed the HST data for candidate
+ OB-type runaway stars. The brightest stars (V$<$14) are missing from
+ the catalogue because of saturation effects, but the high extinction
+ towards VFTS682 makes it faint enough to be included.
+ This star was not included in their study because of its WNh5
+ spectral type and because it would not pass the more conservative
+ selection criteria they adopted.
+ However, the measurements are useful here, since they provides an
+ estimate of the proper motion of VFTS 682 that is completely independent from the \emph{Gaia} measurement.
 
@@
 (Hope you still have the script to show the HST direction.
-)}  The red solid arrow indicates the proper motion of VFTS682.
+)}  The red solid arrow indicates the proper motion of VFTS682
     relative to the region from \emph{Gaia} DR2, starting from the present day position of

### 2018-06-11 VFTS682_APJL.tex
improved plot fig. 1

@@
 
-\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS 682}
+\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS682}
 \maketitle{}
@@
       &$\delta\mu_\mathrm{}$  &[$\masyr$] & $0.13\pm0.09$ &  (1,4) \\
-                 &$v_\mathrm{2d}$  &[$\kms$] & \SdM{$31\pm 21$} (not $32\pm21$!) & (1,4)\\  
+                 &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,4)\\  
                      &$\theta$  &[degrees] &  \SdM{$14.2_{-xx}^{+xx}$}  & (1,4)\\  
@@
     lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$.
-\SdM{ Todo: [1] insert correct errorbars in right pabel, [2] make (a) linewidth, (b) line width of the hashed shading and (c)  figure height consistent.
-(You could try labeling the top x axis in the right panel in degrees.)  [3] fix normalization to be consistent in both pabels, [4] unit in axis label in righthand plot is not correct.
-Degrees but radian.
-[4] Legen: the rel error limits that we adopted are missing a unit.
-Milia arrsecond?
-}}
+      yr^{-1}}$. \SdM{ Todo: [1] insert correct errorbars in right pabel}}
   \label{fig:mu_dist}

### 2018-06-11 VFTS682_APJL.tex
new fig and angle data: TO BE CHECKED

@@
                  &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,4)\\  
-                     &$\theta$  &[degrees] &  \SdM{$14.2_{-xx}^{+xx}$}  & (1,4)\\  
+                     &$\theta$  &[degrees] &  $14.2_{-43}^{+44}$  & (1,4)\\  
                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,4)\\  
@@
                   &$v_\mathrm{2d, HST}$  &[$\kms$] & $47\pm24$ & (3)\\  
-                      &$\theta^{*}$  &[degrees] &   \SdM{$xx_{-xx}^{+xx}$}   & (1,4)\\  
+                      &$\theta^{*}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,4)\\  
                           &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,4)\\  
@@
     angles between the relative proper motion direction and the radial
-    direction to the star. The error bars for VFTS682 are large, but
+    direction to the star. The error bars from \emph{Gaia} for VFTS682 are large, but
     the best value is in agreement with the hypothesis of dynamical
@@
     lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$. \SdM{ Todo: [1] insert correct errorbars in right pabel}}
+      yr^{-1}}$.}
   \label{fig:mu_dist}

### 2018-06-12 VFTS682_APJL.tex
new fig and angle data: TO BE CHECKED
text update: rewrote sec. 3, description of fig. 2 still drafty. Need the figure to improve, and I'm not convinced about the direction we are trying for it with Eva

@@
   R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{4}$, 
-  J.~Bestenlehner$^{5}$, \SdM{C.~J.~Evans$^{X}$},
-  V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{1,7,8}$,  A.~de~Koter$^{1}$,
-  N.~Langer$^{9}$,  \SdM{F. Najarro$^{X}$}, H.~Sana$^{10}$, F.~R.~Schneider$^{11}$, J.~S.~Vink$^{12}$}
-\affil{ $^{*}$ Corresponding author.  A list of the affiliations can be found at the end of this paper.}
-
- \thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+  J.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
+  V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{1,8,9}$,  A.~de~Koter$^{1}$,
+  N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~Schneider$^{13}$, J.~S.~Vink$^{14}$}
+\affil{ $^{*}$ Corresponding author, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}.  A list of the affiliations can be found at the end of this paper.}
+
 \begin{abstract}
@@
       &mass loss rate & [$M_\odot \ \mathrm{yr}^{-1}$] & $10^{-4.1\pm0.2}$ & (2)\\
-      &\multirow{2}{*}{surface helium abundance} &  & 0.45 & (2)\\
+      &\multirow{2}{*}{surface helium mass fraction} &  & 0.45 & (2)\\
       & &  & 0.49 & (3)\\
@@
       \hline
-         &$\mathrm{RA}_\mathrm{VFTS682}$&[degrees] & 84.7314 %$ \pm 0.036\,\mathrm{mas} $ 
+         &$\mathrm{RA}_\mathrm{VFTS682}$&[degrees] & \phantom{-}84.7314 %$ \pm 0.036\,\mathrm{mas} $ 
                      & (1) \\        
@@
 
-                        &$\mathrm{RA}_\mathrm{R136}$&[degrees] & 84.6767
+                        &$\mathrm{RA}_\mathrm{R136}$&[degrees] & \phantom{-}84.6767
                      &  SIMBAD  \\        
@@
                      &  SIMBAD \\       
-        &$      \delta\mathrm{RA}$  &[degrees] & 0.0547                      
-        &  \SdM{your value? }
+        &$      \delta\mathrm{RA}$  &[degrees] & \phantom{-0}0.0547                      
+        &  (2, 5)
   \\        
-               &$     \delta\mathrm{DEC}$  &[degrees] &0.0268 
-                     &  \SdM{your value ? } \\
-
-       \hline
-           \multicolumn{5}{l}{\emph{Gaia's proper motion for VFTS 682
+               &$     \delta\mathrm{DEC}$  &[degrees] & \phantom{-0}0.0268 
+                     &  (2, 5) \\
+
+       \hline
+           \multicolumn{5}{l}{\emph{Gaia proper motion for VFTS682
       and average of the region}} \\
@@
        &$\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$&[$\masyr$] & $1.74\pm0.01$
-                        & (2) \\
+                        & (3) \\
       &$\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$&[$\masyr$]
-                & $0.70\pm0.02$ &  (2)\\
+                & $0.70\pm0.02$ &  (3)\\
 \hline
@@
 
-      \multicolumn{5}{l}{\emph{Gaia's relative proper motion for VFTS
-      682 }} \\
-      \hline
-      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & $0.103\pm0.080$ & (1,4) \\
-      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & $0.086\pm0.100$ &  (1,4) \\
-      &$\delta\mu_\mathrm{}$  &[$\masyr$] & $0.13\pm0.09$ &  (1,4) \\
-                 &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,4)\\  
-                     &$\theta$  &[degrees] &  $14.2_{-43}^{+44}$  & (1,4)\\  
-                         &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,4)\\  
- %
-
- \hline     
-          \multicolumn{5}{l}{\emph{HST's relative proper motion for VFTS 682}} \\
-            \hline
-      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & $0.01\pm0.13$ & (3) \\
-      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (3) \\
-       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (3) \\
-                  &$v_\mathrm{2d, HST}$  &[$\kms$] & $47\pm24$ & (3)\\  
-                      &$\theta^{*}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,4)\\  
-                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,4)\\  
+      \multicolumn{5}{l}{\emph{Gaia relative proper motion for VFTS682 }} \\
+      \hline
+      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & $0.103\pm0.080$ & (1,5) \\
+      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & $0.086\pm0.100$ &  (1,5) \\
+      &$\delta\mu_{Gaia}$  &[$\masyr$] & $0.13\pm0.09$ &  (1,5) \\
+      &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
+      &$\theta$  &[degrees] &  $14.2_{-43}^{+44}$  & (1,5)\\  
+      &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
+ %
+
+ \hline     
+      \multicolumn{5}{l}{\emph{HST relative proper motion for VFTS682}} \\
+            \hline
+      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & $0.01\pm0.13$ & (4) \\
+      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (4) \\
+       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (4) \\
+                  &$v_\mathrm{2d, HST}$  &[$\kms$] & $47\pm24$ & (4)\\  
+                      &$\theta^{*}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,5)\\  
+                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
  %
@@
       (1) \cite{brown:18},
-    (2) \cite{lennon:18}, 
-    (3) \cite{platais:18} and
-    (4) {\color{blue}This work}.
+      (2)~\cite{bestenlehner:11},
+    (3) \cite{lennon:18}, 
+    (4) \cite{platais:18} and
+    (5) {\color{blue}this study}.
     %The error on the RA and DEC positions, are of order $\sim$$10^{-2}\masyr$ in \emph{Gaia} DR2. %REPETITION: %  We adopt the peculiar
@@
 \end{table}
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
-  \includegraphics[width=0.35\textwidth]{figures/angle}
-  \caption{Left panel: distribution of OB-type and Wolf-Rayet stars in proper
-    motion relative to R136. VFTS682 is not an outlier, but
-    its relative proper motion matches the expected value if it were indeed
-    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
-    assuming a distance of 50\,kpc. Right panel:  distribution of
-    angles between the relative proper motion direction and the radial
-    direction to the star. The error bars from \emph{Gaia} for VFTS682 are large, but
-    the best value is in agreement with the hypothesis of dynamical
-    ejection. In both
-    panels, the dark blue histograms contain 317 
-    stars with error smaller than $0.1\,\mathrm{mas \
-      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
-    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$.}
-  \label{fig:mu_dist}
-\end{figure*}
 
@@
 
-%\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
+  \includegraphics[width=0.35\textwidth]{figures/angle}
+  \caption{\emph{Left panel}: distribution of OB-type and Wolf-Rayet stars in proper
+    motion relative to R136. VFTS682 is not an outlier, but
+    its relative proper motion matches the expected value if it were indeed
+    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
+    assuming a distance of 50\,kpc. \emph{Right panel}:  distribution of
+    angles between the relative proper motion direction and the radial
+    direction from the center of the cluster to the star. The error bars from \emph{Gaia} for VFTS682 are large, but
+    the best value is in agreement with the hypothesis of dynamical
+    ejection. In both
+    panels, the dark blue histograms contain 317 
+    stars with error smaller than $0.1\,\mathrm{mas \
+      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
+    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
+      yr^{-1}}$.\todo{possibly: cartoon to illustrate angle definition, or add to
+    fig.2 if not too messy}}
+  \label{fig:dist}
+\end{figure*}
 
 VFTS682 is identified with the source id 4657685637907503744 in the
-\emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band
+\emph{Gaia} DR2 catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}
+  as a 15.65 mag star in the G band
 \citep{gaia:16,brown:18}.   The number of visibility periods,
@@
 
-\SdM{Since the errors for HST and Gaia are so similar and yield to different directions, I think we should probably restructure this paragraph.
-I liked how you break down the story in two separate questions, but given Fig 1 I think it may not work well.
-The first question should have adressed it in context of the motions of the other stars.
-And it has a very typical motion.
-I think that instead, you should simply describe Fig 1 step by step.
-I think that will be stronger.
-}
-
-\subsection{Is it a runaway star?}
-\label{sec:runaway}
-We first address the question whether VFTS682 is a typical star
-from the kinematic point of view (which is what would be expected if it formed in relative isolation), or whether has a significant peculiar velocity compared to its surrounding population (which is what we would expect if the star has a runaway origin).
-
-After subtracting the mean proper motion of the field, we obtain the relative two-dimensional relative proper motion.
+Figure~\ref{fig:dist} illustrates the kinematic properties of VFTS OB-type
+and Wolf-Rayet stars in the 30 Doradus region with \emph{Gaia} DR2 errors on the proper motion
+components of less than $0.1\,\mathrm{mas\ yr^{-1}}$ (dark blue lines,
+including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light shaded
+blue). We highlight VFTS682 together with the two O-type
+runaways \mbox{(re-)identified} by \cite{lennon:18} (VFTS16 and VFTS72,
+respectively). We emphasize that selecting stars with small error bars
+implicitly biases the sample toward fast moving objects.
+
+The left panel shows the distribution in two-dimensional proper motion relative
+to R136, and shows that VFTS682 is not a peculiar star in terms of
+its motion relative to R136. Subtracting the mean motion of the field,
+we obtain $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$ and
+$\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$. Both indicate
+a non-zero peculiar projected motion that would make it a slow runaway
+star as expected by \cite{bestenlehner:11} if the star was ejected
+from the cluster,  but in both cases the meaurements are also consistent with zero within 2$\sigma$. 
+More importantly, both the proper motion
+measurement from \emph{Gaia} and HST fall in the range expected to
+reach its present day location % at $\sim$29\,pc from the cluster core
+if the
+star was ejected very early in its evolution. This range is
+highlighted by the green vertical band, and its width is determined by
+the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star}).
+
+The right panel shows the distributions in flight direction for these
+stars: $\theta$ is the angle between the radial direction from the
+center of R136 to each star and its proper motion relative to
+R136 (see also \Figref{fig:main}). Dynamical ejections from the
+cluster should %are expected to
+produce
+close to radial ejections, i.e.~$\theta\simeq0$. On this panel, the green band highlights
+the range of angles corresponding to an opening angle around the
+radial direction of 45\,degrees. The direction of motion of VFTS682 is
+poorly constrained because of the large errors on the proper motion
+components both in HST and \emph{Gaia} data, but
+it is consistent with dynamical ejection from R136.
+
+Figure~\ref{fig:main} shows the projected motion of VFTS682 relative to R136 on the
+sky. \todo{fix colors} The dashed circle shows the location of the cluster, and the red
+(cyan) arrow shows the relative proper motion from \emph{Gaia} DR2
+(HST). The lighter arrows show the corresponding opening angle within
+the large uncertainties. The length of each arrow is proportional to
+the speed of the star, and dashed prolungations show the possible
+region of origin assuming an age of $1.0\pm0.2$\,Myr
+\citep[][]{schneider:18}. This figure illustrates that R136 is clearly
+the most likely origin of the star, although the large error bars
+prevent a robust identification.
+
+Assuming VFTS682 originates from R136, we can calculate its kinematic
+age simply as:
+\begin{equation}
+  \label{eq:kin_age}
+  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu_{Gaia}} \simeq
+  \frac{119.4\,\mathrm{arcsec}}{0.13\masyr} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
+\end{equation}
+where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
+the core of the cluster \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
+As in the rest of this study, we neglect for
+simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
+The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
+ejection from the cluster, given is apparent age.
+% of $1.0\pm 0.2$\,Myr 
+% \citep{schneider:18}.
+
+Therefore, although the motion of VFTS682 is not an outlier compared
+to other massive stars, it is possible to tentatively claim that it is
+a slow runaway with velocity compatible with dynamical ejection from
+R136 early in the evolution of both the star and the cluster.
+
+% \subsection{Is it a runaway star?}
+% \label{sec:runaway}
+% We first address the question whether VFTS682 is a typical star
+% from the kinematic point of view (which is what would be expected if it formed in relative isolation), or whether has a significant peculiar velocity compared to its surrounding population (which is what we would expect if the star has a runaway origin).
+
+% After subtracting the mean proper motion of the field, we obtain the relative two-dimensional relative proper motion.
 We obtain  $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$   and $\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$.
 
-Both indicate a non-zero peculiar 2D motion that would make it a slow runaway star, but in both cases the meaurements are also consistent with zero within 2 sigma.
-
-Figure~\ref{fig:mu_dist} shows the distribution of relative proper motions for other VFTS O, B and WR stars subselecting either those with relative proper motion errors smaller than $0.1\masyr$ (yielding 317 stars) or a more restrictive cut including only those with proper motion errors smaller than $0.05\masyr$ (yielding 36) stars.
+% Both indicate a non-zero peculiar 2D motion that would make it a slow runaway star, but in both cases the meaurements are also consistent with zero within 2 sigma.
+
+% Figure~\ref{fig:mu_dist} shows the distribution of relative proper motions for other VFTS O, B and WR stars subselecting either those with relative proper motion errors smaller than $0.1\masyr$ (yielding 317 stars) or a more restrictive cut including only those with proper motion errors smaller than $0.05\masyr$ (yielding 36) stars.
 We see that the derived Gaia proper motion is very typical for stars in the field.
 
-\SdM{XXX We need to rethink }
+% \SdM{XXX We need to rethink }
 
@@
 %which would make it the most massive runaway star known to date. However, the large errors on both the proper motion measures require confirmation with future astrometric data.
-
-.
 
@@
 
-\subsection{Does it come from the R136 cluster?}
-\label{sec:r136_origin}
+% \subsection{Does it come from the R136 cluster?}
+% \label{sec:r136_origin}
 
@@
 
-Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
-of OB-type and Wolf-Rayet stars included both in the VFTS survey and
-\emph{Gaia DR2} with reliable astrometric solutions. The bottom panel
-shows the distribution in angles $\theta$ between the relative proper motion
-direction and the direction from the core of R136 to the star. We
-emphasize that our subset of stars with reliable proper motion
-measurements is biased towards the fast moving objects, which results in a
-distribution of angles mildly peaked at small angles.
-
-VFTS682 is not an outlier in relative proper motion%  compared
-% to the two O-type runaways (re-)identified by \cite{lennon:18} (see
-% also below)
-. However, the star was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
-the dynamical ejection scenario. The green shade in
-\Figref{fig:mu_dist} shows the range of relative proper motions
-required to reach the present day location within
-the uncertainties on the apparent age of the star, assuming it was
-ejected from R136 very early in its life.
-
-Because of the large
-error bars, the angle between the relative proper motion is not very
-constraining, but it is suggestive that the best value is close to zero.
-Therefore, the relative proper motion from \emph{Gaia} DR2 are consistent with the hypothesis
-of dynamical ejection.
-
-The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
-VFTS682 from \emph{Gaia} DR2, and the
-yellow arrows illustrate the uncertainty. % show the possible
-% range of directions within the uncertainties in the measured relative proper
-% motion.
-These arrows cross at the present-day location of VFTS682 and
-are prolonged in the direction opposite to the motion to illustrate
-the possible range of origins. % The yellow arrow shows the direction of
+% Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
+% of OB-type and Wolf-Rayet stars included both in the VFTS survey and
+% \emph{Gaia DR2} with reliable astrometric solutions. The bottom panel
+% shows the distribution in angles $\theta$ between the relative proper motion
+% direction and the direction from the core of R136 to the star. We
+% emphasize that our subset of stars with reliable proper motion
+% measurements is biased towards the fast moving objects, which results in a
+% distribution of angles mildly peaked at small angles.
+
+% VFTS682 is not an outlier in relative proper motion%  compared
+% % to the two O-type runaways (re-)identified by \cite{lennon:18} (see
+% % also below)
+% . However, the star was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
+% the dynamical ejection scenario. The green shade in
+% \Figref{fig:mu_dist} shows the range of relative proper motions
+% required to reach the present day location within
+% the uncertainties on the apparent age of the star, assuming it was
+% ejected from R136 very early in its life.
+
+% Because of the large
+% error bars, the angle between the relative proper motion is not very
+% constraining, but it is suggestive that the best value is close to zero.
+% Therefore, the relative proper motion from \emph{Gaia} DR2 are consistent with the hypothesis
+% of dynamical ejection.
+
+% \todo{discuss and find a place for Fig.2}
+% The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
+% VFTS682 from \emph{Gaia} DR2, and the
+% yellow arrows illustrate the uncertainty. % show the possible
+% % range of directions within the uncertainties in the measured relative proper
+% % motion.
+% These arrows cross at the present-day location of VFTS682 and
+% are prolonged in the direction opposite to the motion to illustrate
+% the possible range of origins.
+% The yellow arrow shows the direction of
 % the relative proper motion from HST (see \Eqref{eq:pm_around_HST}),
@@
 
-The kinematic age of this star, assuming it originates from R136, is
-
-\begin{equation}
-  \label{eq:kin_age}
-  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu^\mathrm{Gaia}} \simeq
-  \frac{119.4\,\mathrm{arcsec}}{0.13\masyr} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
-\end{equation}
-where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
-the core of the cluster \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
-As in the rest of this study, we neglect for
-simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
-The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
-ejection from the cluster, given is apparent age
-of $1.0\pm 0.2$\,Myr \citep{schneider:18}. % , which
+ % , which
 % corroborates the idea that the star is the result of a dynamical
@@
 Based on our results, we tentatively claim that VFTS682 is the most massive
-runaway known to date, with a peculiar three-dimensional speed of
-$44\pm21\,\kms$. Due to the large error bars, this result will need
+(slow) runaway known to date, with a peculiar three-dimensional speed of
+$44\pm21\,\kms$ \todo{make 2d? use LOS velocity from Marimar?}. Due to the large error bars, this result will need
 to be revisited with future astrometric data. % to be revisited as updated
@@
 debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
-claim that LBV stars are typically isolated form O-type stars. The fact that VFTS682 is a dynamically
+claim that LBV stars are typically isolated from O-type stars. The fact that VFTS682 is a dynamically
 ejected runaway which might evolve into an LBV star suggests that
@@
 moving O-type stars
-in the region, and found two massive runaway stars
-($\sim$$90\,M_\odot$) in the 30 Doradus region. One of them (VFTS 16)
+in the region, and found two very massive runaway stars in the 30 Doradus region. One of them (VFTS 16)
 was previously known as a runaway star from its line of sight velocity
@@
 origin of the other star (VFTS 72) is less clear given its direction
-of motion. The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
-VFTS682 (see \Secref{sec:r136_origin}) is smaller than the
+of motion (cf.~\Figref{fig:dist}). The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
+VFTS682 is smaller than the
 corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic
@@
     Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
-  $^{6}$ {National Research Council, Herzberg Astronomy \&
+  $^{6}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}\\
+  $^{7}$ {National Research Council, Herzberg Astronomy \&
     Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
     Canada}\\
-  $^{7}$ {School of Astronomy \& Space Science, University of the Chinese
+  $^{8}$ {School of Astronomy \& Space Science, University of the Chinese
     Academy of Sciences, Beijing 100012, China}\\
-  $^{8}$ {National Astronomical Observatories, Chinese Academy of
+  $^{9}$ {National Astronomical Observatories, Chinese Academy of
     Sciences, Beijing 100012, China}\\
-  $^{9}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
+  $^{10}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
     Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
-  $^{10}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
-  $^{11}$ {Department of Physics, University of Oxford, Keble Road,
+  $^{11}$ {Centro de Astrobiología, CSIC-INTA, Carretera de Torrejón a Ajalvir km-4, E-28850 Torrejón de Ardoz, Madrid, Spain}\\
+  $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
+  $^{13}$ {Department of Physics, University of Oxford, Keble Road,
     Oxford OX1 3RH, UK} \\
-  $^{12}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
- \SdM{$^{x}$ {Affiliation: Chris Evans}\\
-  $^{x}$ {Affiliation: Paco Najorro}\\}
+  $^{14}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
+}
 
@@
    \small
-   We are grateful to S.\,Torres, M.\,C.\,Ramirez-Tannus,
-   and C.\,J.\,Evans for help and discussions, European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC), Grant agreement No.
+   We are grateful to S.\,Torres and M.\,C.\,Ramirez-Tannus
+   % and C.\,J.\,Evans
+   for help and discussions, European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC), Grant agreement No.
 715063 [SdM], the NRC-Canada Plaskett Fellowship [VHB].

### 2018-06-13 VFTS682_APJL.tex
minor

@@
   \end{center}
-  \label{tab:star}
+  \label{tab:star_param}
 \end{table}
@@
 The star is unlikely to have a close companion, because of the absence of clear radial velocity variations \citet{bestenlehner:11}.
-Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see \Tabref{tab:star} for an overview of the parameters.
+Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see \Tabref{tab:star_param} for an overview of the parameters.
 
@@
       yr^{-1}}$.\todo{possibly: cartoon to illustrate angle definition, or add to
-    fig.2 if not too messy}}
+    fig.2 if not too messy, maybe also show HST measurements on this
+    plot for 682?}}
   \label{fig:dist}
@@
 highlighted by the green vertical band, and its width is determined by
-the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star}).
+the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star_param}).
 
@@
 (cyan) arrow shows the relative proper motion from \emph{Gaia} DR2
-(HST). The lighter arrows show the corresponding opening angle within
-the large uncertainties. The length of each arrow is proportional to
+(HST). The lighter arrows show the uncertainty on the direction of
+motion due to the large uncertainties. The length of each arrow is proportional to
 the speed of the star, and dashed prolungations show the possible
@@
 
-Assuming VFTS682 originates from R136, we can calculate its kinematic
+Assuming VFTS682 indeed originates from R136, we can calculate its kinematic
 age simply as:
@@
 Based on our results, we tentatively claim that VFTS682 is the most massive
-(slow) runaway known to date, with a peculiar three-dimensional speed of
-$44\pm21\,\kms$ \todo{make 2d? use LOS velocity from Marimar?}. Due to the large error bars, this result will need
+(slow) runaway known to date, with a peculiar two-dimensional
+projected velocity with respect to R136 of
+$\sim38\pm22\kms$ (averaging the \emph{Gaia} DR2 and HST
+results\todo{check numbers}). Due to the large error bars, this result will need
 to be revisited with future astrometric data. % to be revisited as updated
@@
   $^{14}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
-}
 

### 2018-06-20 VFTS682_APJL.tex
minor
significant minors

@@
       &mass loss rate & [$M_\odot \ \mathrm{yr}^{-1}$] & $10^{-4.1\pm0.2}$ & (2)\\
-      &\multirow{2}{*}{surface helium mass fraction} &  & 0.45 & (2)\\
-      & &  & 0.49 & (3)\\
+      % &\multirow{2}{*}{surface helium mass fraction} &  & 0.45 & (2)\\
+      % & &  & 0.49 & (3)\\ %% we don't use this: if we believe it's a
+      % merger it's not relevant, if we don't we still don't use it
       \hline
@@
       (2)~\cite{bestenlehner:11}
-      (3)~\cite{rubio-diez:17}
+      % (3)~\cite{rubio-diez:17}
     }
@@
       &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
-      &$\theta$  &[degrees] &  $14.2_{-43}^{+44}$  & (1,5)\\  
-      &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
+      &$\theta$  &[degrees] &  $14_{-43}^{+44}$  & (1,5)\\  
+      % &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
  %
@@
                       &$\theta^{*}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,5)\\  
-                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
+                          % &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
  %
@@
 
-Estimates of the radial velocity are complicated by the variable, optically thick winds that is typical for emission line stars.
-\citet{bestenlehner:11} estimate a mass loss rate of $10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the possible effect of clumping.
-The V-band light curve of VFTS682  shows variations at a $\sim$10\% level on a timescale of years, which is unusual for Wolf-Rayet stars and more typical for Luminous Blue Variable (LBV) stars \citep{udalski:08, bestenlehner:11}.
-It also shows mid-infrared excess \citep{gruendl:09}.
+Estimates of the radial velocity are complicated by the variable,
+optically thick winds that is typical for emission line
+stars. \citet{bestenlehner:11} estimate a mass loss rate of
+$10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the
+possible effect of clumping.  The V-band light curve of VFTS682  shows
+variations at a $\sim$10\% level on a timescale of years, which is
+unusual for Wolf-Rayet stars and more typical for Luminous Blue
+Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
+shows mid-infrared excess \citep{gruendl:09}.
 We therefore caution for over-interpreting the estimates for the radial velocities, but we will summarize the estimates that have been published.
@@
 %of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
+VFTS682 is relatively isolated in \emph{Spitzer} images, except for the extended
+source S10 interpreted as star formation triggered by the wind of
+VFTS682 itself by \cite{walborn:13}. In X-rays
+\todo{questions: does 682 show excess X-rays (wind collisions?)}
 
@@
     fig.2 if not too messy, maybe also show HST measurements on this
-    plot for 682?}}
+    plot for 682? Redo right panel taking care of projection of the
+    radial directions!It's misleading!}}
   \label{fig:dist}
@@
 i.e. such that $\theta = 0$ is the angle corresponding to perfectly
-radial motion away from R136.    We further provide the 2D velocity
-projected onto the radial vector pointing away from R136
-$v_\mathrm{2d, \parallel}$  and the total projected 2d velocity
+radial motion away from R136.    We further provide % the 2D velocity
+% projected onto the radial vector pointing away from R136
+% $v_\mathrm{2d, \parallel}$  and
+the %total
+projected 2d velocity
 $v_\mathrm{2d}$.   All kinematic quantities are provided in
@@
 The right panel shows the distributions in flight direction for these
-stars: $\theta$ is the angle between the radial direction from the
+stars: $\theta$ is the angle between the projected radial direction from the
 center of R136 to each star and its proper motion relative to
@@
 components both in HST and \emph{Gaia} data, but
-it is consistent with dynamical ejection from R136.
+it is consistent with dynamical ejection from R136. We note that
+VFTS72 has a small radial velocity, while VFTS16 (and possibly VRTS682) has
+large peculiar radial velocity and therefore accurate distances are
+needed to contrain the flight direction in three dimensions.
 
@@
 projected velocity with respect to R136 of
-$\sim38\pm22\kms$ (averaging the \emph{Gaia} DR2 and HST
-results\todo{check numbers}). Due to the large error bars, this result will need
+$\sim38\kms$ (averaging the \emph{Gaia} DR2 and HST
+results) but compatible with zero within $2\sigma$. Due to the large error bars, this result will need
 to be revisited with future astrometric data. % to be revisited as updated

### 2018-06-20 VFTS682_APJL.tex
more caveats and some smoothing

@@
 
-The right panel shows the distributions in flight direction for these
+The right panel shows the distributions in projected flight direction for these
 stars: $\theta$ is the angle between the projected radial direction from the
-center of R136 to each star and its proper motion relative to
-R136 (see also \Figref{fig:main}). Dynamical ejections from the
+center of R136 to each star and its projected relative proper motion (see also \Figref{fig:main}). Dynamical ejections from the
 cluster should %are expected to
 produce
-close to radial ejections, i.e.~$\theta\simeq0$. On this panel, the green band highlights
+close to radial ejections, i.e.~$\theta\simeq0$ \todo{this is only
+  true for small RV}. On this panel, the green band highlights
 the range of angles corresponding to an opening angle around the
-radial direction of 45\,degrees. The direction of motion of VFTS682 is
+projected radial direction of 45\,degrees. The direction of motion of VFTS682 is
 poorly constrained because of the large errors on the proper motion

### 2018-06-20 VFTS682_APJL.tex
added mention of X-rays

@@
 %
-\citet{bestenlehner:11}  analyzed the spectra to infer the stellar parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of the brightest stars in the region.
-The star is unlikely to have a close companion, because of the absence of clear radial velocity variations \citet{bestenlehner:11}.
-Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see \Tabref{tab:star_param} for an overview of the parameters.
+\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
+parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
+luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
+the brightest stars in the region. The star is unlikely to have a
+close companion, because of the absence of clear radial velocity
+variations \citet{bestenlehner:11}. Bayesian fits of the stellar
+parameters against evolutionary tracks \citep{brott:11, kohler:15}
+using the Bonsaii code \citep{schneider:17, schneider:18} provide
+estimates for the age, present mass and initial mass, indicating that
+this star among the most massive stars known, see
+\Tabref{tab:star_param} for an overview of the parameters.
+
+\cite{townsley:06} did not find
+an X-ray point source at the position of VFTS682 in their
+\emph{Chandra} survey of the region, which suggests the absence of
+colliding winds that would be expected in the presence of companions
+even for extreme mass ratios, given the extremely large mass of VFTS682.
 
@@
 source S10 interpreted as star formation triggered by the wind of
-VFTS682 itself by \cite{walborn:13}. In X-rays
-\todo{questions: does 682 show excess X-rays (wind collisions?)}
+VFTS682 itself by \cite{walborn:13}.
 
@@
    \small
-   We are grateful to S.\,Torres and M.\,C.\,Ramirez-Tannus
+   We are grateful to M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
    % and C.\,J.\,Evans

### 2018-06-23 VFTS682_APJL.tex
selma's commits

--- 2018-06-20 VFTS682_APJL.tex
@@
 %\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
-\title{Gaia's constraints on the extremely massive  ($\sim$$150\,M_\odot$) runaway star candidate VFTS682}
-
-\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
-  R.~P.~van~der~Marel$^{3,4}$, I.~Platais$^{4}$, 
+\title{Gaia's constraints on the very massive  (initially $\sim$$150\,M_\odot$) runaway star candidate VFTS682}
+
+\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, \SdM{I.~Platais$^{3}$,}
+  R.~P.~van~der~Marel$^{3,4}$,  \SdM{E.~L.~Laplace$^{1}$,}
   J.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
@@
   N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~Schneider$^{13}$, J.~S.~Vink$^{14}$}
-\affil{ $^{*}$ Corresponding author, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}.  A list of the affiliations can be found at the end of this paper.}
+\affil{ $^{*}$ Corresponding author: \href{mailto:M.Renzo@uva.nl}{M.Renzo@UvA.nl}.  A list of the affiliations can be found at the end of this paper.}
 
@@
       &mass loss rate & [$M_\odot \ \mathrm{yr}^{-1}$] & $10^{-4.1\pm0.2}$ & (2)\\
-      % &\multirow{2}{*}{surface helium mass fraction} &  & 0.45 & (2)\\
-      % & &  & 0.49 & (3)\\ %% we don't use this: if we believe it's a
-      % merger it's not relevant, if we don't we still don't use it
+      &\multirow{2}{*}{surface helium mass fraction} &  & 0.45 & (2)\\
+      & &  & 0.49 & (3)\\
       \hline
@@
       (2)~\cite{bestenlehner:11}
-      % (3)~\cite{rubio-diez:17}
+      (3)~\cite{rubio-diez:17}
     }
   \end{center}
-  \label{tab:star_param}
+  \label{tab:star}
 \end{table}
@@
       \hline
-      \multicolumn{5}{l}{\emph{Coordinates}} \\
+      \multicolumn{5}{l}{\emph{Absolute and relative position}} \\
       \hline
@@
                      &  (2, 5) \\  
-
-       \hline
-           \multicolumn{5}{l}{\emph{Gaia proper motion for VFTS682
-      and average of the region}} \\
+                       &$  \mathrm{d_\mathrm{2d}}$  &[pc] & 29
+                     &  (Besthenlehner) \\
+
+                     \hline
+           \multicolumn{5}{l}{\emph{Gaia absolute proper motion for VFTS682
+      and the region}} \\
       \hline
@@
       &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
-      &$\theta$  &[degrees] &  $14_{-43}^{+44}$  & (1,5)\\  
+      &$\theta$  &[degrees] &  $14.2_{-43}^{+44}$  & (1,5)\\  
 %      &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
@@
                   %        &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
- %
-
-%\hline
-     % \multicolumn{5}{l}{\emph{ Direction and magnitude w.r.t the vector pointing away from R136}} \\     
- %      &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
-     %   &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
- %          &$v_\mathrm{2d}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\
+    \hline
+       \multicolumn{5}{l}{\emph{Expected proper motion if ejected from R136 at zero age}} \\
+      \hline
+      &$v_\mathrm{2d}$  &[$\kms$] & \SdM{$xx\pm xx$} & (Bestenlehner) \\  
+      &$\theta$  &[degrees] &  $0$  & \\ 
+       \hline
 
@@
 The star is unlikely to have a close companion, because of the absence of clear radial velocity variations \citet{bestenlehner:11}.
-Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see \Tabref{tab:star_param} for an overview of the parameters.
-
-Estimates of the radial velocity are complicated by the variable,
-optically thick winds that is typical for emission line
-stars. \citet{bestenlehner:11} estimate a mass loss rate of
-$10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the
-possible effect of clumping.  The V-band light curve of VFTS682  shows
-variations at a $\sim$10\% level on a timescale of years, which is
-unusual for Wolf-Rayet stars and more typical for Luminous Blue
-Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
-shows mid-infrared excess \citep{gruendl:09}.
+Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see \Tabref{tab:star} for an overview of the parameters.
+
+Estimates of the radial velocity are complicated by the variable, optically thick winds that is typical for emission line stars.
+\citet{bestenlehner:11} estimate a mass loss rate of $10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the possible effect of clumping.
+The V-band light curve of VFTS682  shows variations at a $\sim$10\% level on a timescale of years, which is unusual for Wolf-Rayet stars and more typical for Luminous Blue Variable (LBV) stars \citep{udalski:08, bestenlehner:11}.
+It also shows mid-infrared excess \citep{gruendl:09}.
 We therefore caution for over-interpreting the estimates for the radial velocities, but we will summarize the estimates that have been published.
@@
 This is consistent with, but also no proof of, the expectation if the star was not formed in situ.
-Newer measurements based on new Xshooter data give different estimates for the RV ({\color{blue} Rubio-D{\' i}ez et al.
+
+Newer measurements based on a fit to new Xshooter data covering the full spectrum from 3200 Angstr{\"o}m to 2.45 microns lead to an estimate of $260 \pm 15 \kms$ ({\color{blue} Rubio-D{\' i}ez et al.
 in prep}).
-We will therefore refrain from using the RV measurements in this work.
+This estimate, which we consider to be the most reliable fo those quoted here, indicates that the relative motion along the line-of-sight is small or negligible.
+This in consistent with both hypotheses.
 
 %of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
-VFTS682 is relatively isolated in \emph{Spitzer} images, except for the extended
-source S10 interpreted as star formation triggered by the wind of
-VFTS682 itself by \cite{walborn:13}. In X-rays
-\todo{questions: does 682 show excess X-rays (wind collisions?)}
 
@@
       yr^{-1}}$.\todo{possibly: cartoon to illustrate angle definition, or add to
-    fig.2 if not too messy, maybe also show HST measurements on this
-    plot for 682? Redo right panel taking care of projection of the
-    radial directions!It's misleading!}}
+    fig.2 if not too messy}}
   \label{fig:dist}
@@
 
+Gaia provides absolute proper motions.
 To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.
@@
 With these we compute the relative proper motion
-$\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.  We also derive
+$\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.   We also compute the total projected 2d velocity
+$v_\mathrm{2d}$ and
 the angle $\theta$ between the direction of motion and the vector
-connecting the center of R136 with the current position of VFTS 682,
-i.e. such that $\theta = 0$ is the angle corresponding to perfectly
-radial motion away from R136.    We further provide % the 2D velocity
-% projected onto the radial vector pointing away from R136
-% $v_\mathrm{2d, \parallel}$  and
-the %total
-projected 2d velocity
-$v_\mathrm{2d}$.   All kinematic quantities are provided in
-Tab.~\ref{tab:vfts682}. %%what does the projection on the proper
-                        %%motion in the radial direction tell us?
-                        %%Seems like a useless number to me...
+connecting the center of R136 with the current position of VFTS 682.
+This means that $\theta = 0$ is the angle that is expected for a star that is moving away from R136 with a perfect radial motion.
+All kinematic quantities are provided in
+Tab.~\ref{tab:vfts682}.
 
@@
 
- \citet{platais:18} recently analyzed the HST data for candidate
- OB-type runaway stars. The brightest stars (V$<$14) are missing from
- the catalogue because of saturation effects, but the high extinction
- towards VFTS682 makes it faint enough to be included.
- This star was not included in their study because of its WNh5
- spectral type and because it would not pass the more conservative
- selection criteria they adopted.
- However, the measurements are useful here, since they provides an
- estimate of the proper motion of VFTS 682 that is completely independent from the \emph{Gaia} measurement.
-
-We list their data in \Tabref{tab:vfts682}.
+ \citet{platais:15, platais:18} analyzed the HST data to determine the relative proper motion and identify candidate runaway stars.
+The brightest stars (V$<$14) are saturated in the data set and have been excluded from the analysis.
+The very massive star VFTS682 should be one of the brightest stars in the region, the high extinction makes it faint enough such that saturation effects are not an issue.
+VFTS682 is relatively slow and is therefore not discussed in their paper.
+However, their measurements are useful here, since they provides an  estimate of the proper motion of VFTS 682 that is independent from the \emph{Gaia} measurement.
+
+The HST provides proper motions that are relative to the bulk motion of the majority of the stars in the field of view.
+The full 30 Dor field is covered by different pointings and there is some systematic distortion.
+However, even for stars far from 30 Dor the effect is small, no more than $0.05\masyr$ across the whole 30 Dor field.
+The effect is much smaller for stars close to the center of field, such as VFTS 682 \citet{platais:18}.
+We can therefore use the relative proper motion as a good estimate for the proper motion relative to R136.
+
+ We list the relative proper motion in \Tabref{tab:vfts682}.
 
@@
 
+At first glance the magnitude of the relative proper motions derived by Gaia,  $32\pm 21\kms$, and by HST, $47 \pm 24\kms$, appear consistent.  In principle they are consistent with the velocity
+
 Figure~\ref{fig:dist} illustrates the kinematic properties of VFTS OB-type
@@
 highlighted by the green vertical band, and its width is determined by
-the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star_param}).
+the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star}).
 
 The right panel shows the distributions in flight direction for these
-stars: $\theta$ is the angle between the projected radial direction from the
+stars: $\theta$ is the angle between the radial direction from the
 center of R136 to each star and its proper motion relative to
@@
 components both in HST and \emph{Gaia} data, but
-it is consistent with dynamical ejection from R136. We note that
-VFTS72 has a small radial velocity, while VFTS16 (and possibly VRTS682) has
-large peculiar radial velocity and therefore accurate distances are
-needed to contrain the flight direction in three dimensions.
+it is consistent with dynamical ejection from R136.
 
@@
 (cyan) arrow shows the relative proper motion from \emph{Gaia} DR2
-(HST). The lighter arrows show the uncertainty on the direction of
-motion due to the large uncertainties. The length of each arrow is proportional to
+(HST). The lighter arrows show the corresponding opening angle within
+the large uncertainties. The length of each arrow is proportional to
 the speed of the star, and dashed prolungations show the possible
@@
 
-Assuming VFTS682 indeed originates from R136, we can calculate its kinematic
+Assuming VFTS682 originates from R136, we can calculate its kinematic
 age simply as:
@@
 Based on our results, we tentatively claim that VFTS682 is the most massive
-(slow) runaway known to date, with a peculiar two-dimensional
-projected velocity with respect to R136 of
-$\sim38\kms$ (averaging the \emph{Gaia} DR2 and HST
-results) but compatible with zero within $2\sigma$. Due to the large error bars, this result will need
+(slow) runaway known to date, with a peculiar three-dimensional speed of
+$44\pm21\,\kms$ \todo{make 2d? use LOS velocity from Marimar?}. Due to the large error bars, this result will need
 to be revisited with future astrometric data. % to be revisited as updated
@@
 
-%\begin{thebibliography}{}
-%\makeatletter
-%\relax
-%\def\mn@urlcharsother{\let\do\@makeother \do\$\do\&\do\#\do\^\do\_\do\%\do\~}
-%\def\mn@doi{\begingroup\mn@urlcharsother \@ifnextchar [ {\mn@doi@}
-%  {\mn@doi@[]}}
-%\def\mn@doi@[#1]#2{\def\@tempa{#1}\ifx\@tempa\@empty \href
-%  {http://dx.doi.org/#2} {doi:#2}\else \href {http://dx.doi.org/#2} {#1}\fi
-%  \endgroup}
-%\def\mn@eprint#1#2{\mn@eprint@#1:#2::\@nil}
-%\def\mn@eprint@arXiv#1{\href {http://arxiv.org/abs/#1} {{\tt arXiv:#1}}}
-%\def\mn@eprint@dblp#1{\href {http://dblp.uni-trier.de/rec/bibtex/#1.xml}
-%  {dblp:#1}}
-%\def\mn@eprint@#1:#2:#3:#4\@nil{\def\@tempa {#1}\def\@tempb {#2}\def\@tempc
-%  {#3}\ifx \@tempc \@empty \let \@tempc \@tempb \let \@tempb \@tempa \fi \ifx
-%  \@tempb \@empty \def\@tempb {arXiv}\fi \@ifundefined
-%  {mn@eprint@\@tempb}{\@tempb:\@tempc}{\expandafter \expandafter \csname
-%  mn@eprint@\@tempb\endcsname \expandafter{\@tempc}}}
-%
-%\bibitem[\protect\citeauthoryear{{Banerjee}, {Kroupa}  \& {Oh}}{{Banerjee}
-%  et~al.}{2012}]{banerjee:12}
-%{Banerjee} S.,  {Kroupa} P.,   {Oh} S.,  2012, \mn@doi [\apj]
-%  {10.1088/0004-637X/746/1/15}, \href
-%  {http://adsabs.harvard.edu/abs/2012ApJ...746...15B} {746, 15}
-%
-%\bibitem[\protect\citeauthoryear{{Bate}}{{Bate}}{2009}]{bate:09}
-%{Bate} M.~R.,  2009, \mn@doi [\mnras] {10.1111/j.1365-2966.2008.14106.x}, \href
-%  {http://adsabs.harvard.edu/abs/2009MNRAS.392..590B} {392, 590}
-%
-%\bibitem[\protect\citeauthoryear{{Bestenlehner} et~al.,}{{Bestenlehner}
-%  et~al.}{2011}]{bestenlehner:11}
-%{Bestenlehner} J.~M.,  et~al., 2011, \mn@doi [\aap]
-%  {10.1051/0004-6361/201117043}, \href
-%  {http://adsabs.harvard.edu/abs/2011A\%26A...530L..14B} {530, L14}
-%
-%\bibitem[\protect\citeauthoryear{{Blaauw}}{{Blaauw}}{1961}]{blaauw:61}
-%{Blaauw} A.,  1961, \bain, \href
-%  {http://adsabs.harvard.edu/abs/1961BAN....15..265B} {15, 265}
-%
-%\bibitem[\protect\citeauthoryear{{Bressert} et~al.,}{{Bressert}
-%  et~al.}{2012}]{bressert:12}
-%{Bressert} E.,  et~al., 2012, \mn@doi [\aap] {10.1051/0004-6361/201117247},
-%  \href {http://adsabs.harvard.edu/abs/2012A\%26A...542A..49B} {542, A49}
-%
-%\bibitem[\protect\citeauthoryear{{Brott} et~al.,}{{Brott}
-%  et~al.}{2011}]{brott:11}
-%{Brott} I.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201016113}, \href
-%  {http://adsabs.harvard.edu/abs/2011A\%26A...530A.115B} {530, A115}
-%
-%\bibitem[\protect\citeauthoryear{{Crowther}, {Schnurr}, {Hirschi}, {Yusof},
-%  {Parker}, {Goodwin}  \& {Kassim}}{{Crowther} et~al.}{2010}]{crowther:10}
-%{Crowther} P.~A.,  {Schnurr} O.,  {Hirschi} R.,  {Yusof} N.,  {Parker} R.~J.,
-%  {Goodwin} S.~P.,   {Kassim} H.~A.,  2010, \mn@doi [\mnras]
-%  {10.1111/j.1365-2966.2010.17167.x}, \href
-%  {http://adsabs.harvard.edu/abs/2010MNRAS.408..731C} {408, 731}
-%
-%\bibitem[\protect\citeauthoryear{{Crowther} et~al.,}{{Crowther}
-%  et~al.}{2016}]{crowther:16}
-%{Crowther} P.~A.,  et~al., 2016, \mn@doi [\mnras] {10.1093/mnras/stw273}, \href
-%  {http://adsabs.harvard.edu/abs/2016MNRAS.458..624C} {458, 624}
-%
-%\bibitem[\protect\citeauthoryear{{Davidson}, {Humphreys}  \& {Weis}}{{Davidson}
-%  et~al.}{2016}]{davidson:16}
-%{Davidson} K.,  {Humphreys} R.~M.,   {Weis} K.,  2016, arXiv:1608.02007, \href
-%  {http://adsabs.harvard.edu/abs/2016arXiv160802007D} {}
-%
-%\bibitem[\protect\citeauthoryear{{de Koter}, {Heap}  \& {Hubeny}}{{de Koter}
-%  et~al.}{1997}]{dekoter:97}
-%{de Koter} A.,  {Heap} S.~R.,   {Hubeny} I.,  1997, \mn@doi [\apj]
-%  {10.1086/303736}, \href {http://adsabs.harvard.edu/abs/1997ApJ...477..792D}
-%  {477, 792}
-%
-%  
-%\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
-%  et~al.}{2010}]{evans:10}
-%{Evans} C.~J.,  et~al., 2010, \mn@doi [\apjl] {10.1088/2041-8205/715/2/L74},
-%  \href {http://adsabs.harvard.edu/abs/2010ApJ...715L..74E} {715, L74}
-%
-%\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
-%  et~al.}{2011}]{evans:11}
-%{Evans} C.~J.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201116782},
-%  \href {http://adsabs.harvard.edu/abs/2011A\%26A...530A.108E} {530, A108}
-%
-%\bibitem[\protect\citeauthoryear{{Fujii} \& {Portegies Zwart}}{{Fujii} \&
-%  {Portegies Zwart}}{2011}]{fujii:11}
-%{Fujii} M.~S.,  {Portegies Zwart} S.,  2011, \mn@doi [Science]
-%  {10.1126/science.1211927}, \href
-%  {http://adsabs.harvard.edu/abs/2011Sci...334.1380F} {334, 1380}
-%
-%\bibitem[\protect\citeauthoryear{{Gaia Collaboration} et~al.,}{{Gaia
-%  Collaboration} et~al.}{2016}]{gaia:16}
-%{Gaia Collaboration} et~al., 2016, \mn@doi [\aap]
-%  {10.1051/0004-6361/201629272}, \href
-%  {http://adsabs.harvard.edu/abs/2016A\%26A...595A...1G} {595, A1}
-%
-%\bibitem[\protect\citeauthoryear{{Gaia Collaboration}, {Brown}, {Vallenari},
-%  {Prusti}, {de Bruijne}, {Babusiaux}  \& {Bailer-Jones}}{{Gaia Collaboration}
-%  et~al.}{2018}]{brown:18}
-%{Gaia Collaboration} {Brown} A.~G.~A.,  {Vallenari} A.,  {Prusti} T.,  {de
-%  Bruijne} J.~H.~J.,  {Babusiaux} C.,   {Bailer-Jones} C.~A.~L.,  2018,
-%  ArXiv:1804.09365, \href {http://adsabs.harvard.edu/abs/2018arXiv180409365G}
-%  {}
-%
-%\bibitem[\protect\citeauthoryear{{Gruendl} \& {Chu}}{{Gruendl} \&
-%  {Chu}}{2009}]{gruendl:09}
-%{Gruendl} R.~A.,  {Chu} Y.-H.,  2009, \mn@doi [\apjs]
-%  {10.1088/0067-0049/184/1/172}, \href
-%  {http://adsabs.harvard.edu/abs/2009ApJS..184..172G} {184, 172}
-%
-%\bibitem[\protect\citeauthoryear{{Gvaramadze}, {Weidner}, {Kroupa}  \&
-%  {Pflamm-Altenburg}}{{Gvaramadze} et~al.}{2012}]{gvaramadze:12}
-%{Gvaramadze} V.~V.,  {Weidner} C.,  {Kroupa} P.,   {Pflamm-Altenburg} J.,
-%  2012, \mn@doi [\mnras] {10.1111/j.1365-2966.2012.21452.x}, \href
-%  {http://adsabs.harvard.edu/abs/2012MNRAS.424.3037G} {424, 3037}
-%
-%% \bibitem[\protect\citeauthoryear{{Hogg}}{{Hogg}}{2018}]{hogg:18}
-%% {Hogg} D.~W.,  2018, ArXiv:1804.07766, \href
-%%   {http://adsabs.harvard.edu/abs/2018arXiv180407766H} {}
-%
-%\bibitem[\protect\citeauthoryear{{Humphreys}, {Weis}, {Davidson}  \&
-%  {Gordon}}{{Humphreys} et~al.}{2016}]{humphreys:16}
-%{Humphreys} R.~M.,  {Weis} K.,  {Davidson} K.,   {Gordon} M.~S.,  2016, \mn@doi
-%  [\apj] {10.3847/0004-637X/825/1/64}, \href
-%  {http://adsabs.harvard.edu/abs/2016ApJ...825...64H} {825, 64}
-%
-%\bibitem[\protect\citeauthoryear{{K{\"o}hler} et~al.,}{{K{\"o}hler}
-%  et~al.}{2015}]{kohler:15}
-%{K{\"o}hler} K.,  et~al., 2015, \mn@doi [\aap] {10.1051/0004-6361/201424356},
-%  \href {http://adsabs.harvard.edu/abs/2015A\%26A...573A..71K} {573, A71}
-%
-%\bibitem[\protect\citeauthoryear{{Kuiper}, {Yorke}  \& {Turner}}{{Kuiper}
-%  et~al.}{2015}]{kuiper:15}
-%{Kuiper} R.,  {Yorke} H.~W.,   {Turner} N.~J.,  2015, \mn@doi [\apj]
-%  {10.1088/0004-637X/800/2/86}, \href
-%  {http://adsabs.harvard.edu/abs/2015ApJ...800...86K} {800, 86}
-%
-%\bibitem[\protect\citeauthoryear{{Lada} \& {Lada}}{{Lada} \&
-%  {Lada}}{2003}]{lada:03}
-%{Lada} C.~J.,  {Lada} E.~A.,  2003, \mn@doi [\araa]
-%  {10.1146/annurev.astro.41.011802.094844}, \href
-%  {http://adsabs.harvard.edu/abs/2003ARA\%26A..41...57L} {41, 57}
-%
-%\bibitem[\protect\citeauthoryear{{Lamb}, {Oey}, {Segura-Cox}, {Graus},
-%  {Kiminki}, {Golden-Marx}  \& {Parker}}{{Lamb} et~al.}{2016}]{lamb:16}
-%{Lamb} J.~B.,  {Oey} M.~S.,  {Segura-Cox} D.~M.,  {Graus} A.~S.,  {Kiminki}
-%  D.~C.,  {Golden-Marx} J.~B.,   {Parker} J.~W.,  2016, \mn@doi [\apj]
-%  {10.3847/0004-637X/817/2/113}, \href
-%  {http://adsabs.harvard.edu/abs/2016ApJ...817..113L} {817, 113}
-%
-%\bibitem[\protect\citeauthoryear{{Langer}, {Norman}, {de Koter}, {Vink},
-%  {Cantiello}  \& {Yoon}}{{Langer} et~al.}{2007}]{langer:07}
-%{Langer} N.,  {Norman} C.~A.,  {de Koter} A.,  {Vink} J.~S.,  {Cantiello} M.,
-%  {Yoon} S.-C.,  2007, \mn@doi [\aap] {10.1051/0004-6361:20078482}, \href
-%  {http://adsabs.harvard.edu/abs/2007A\%26A...475L..19L} {475, L19}
-%
-%\bibitem[\protect\citeauthoryear{{Lennon} et~al.,}{{Lennon}
-%  et~al.}{2018}]{lennon:18}
-%{Lennon} D.~J.,  et~al., 2018, ArXiv:1805.08277, \href
-%  {http://adsabs.harvard.edu/abs/2018arXiv180508277L} {}
-%
-%\bibitem[\protect\citeauthoryear{{Lindegren} et~al.,}{{Lindegren}
-%  et~al.}{2018}]{lindengren:18}
-%{Lindegren} L.,  et~al., 2018, ArXiv:1804.09366, \href
-%  {http://adsabs.harvard.edu/abs/2018arXiv180409366L} {}
-%
-%% \bibitem[\protect\citeauthoryear{{Maeder} \& {Meynet}}{{Maeder} \&
-%%   {Meynet}}{2000}]{maeder:00}
-%% {Maeder} A.,  {Meynet} G.,  2000, \aap, \href
-%%   {http://adsabs.harvard.edu/abs/2000A\%26A...361..159M} {361, 159}
-%
-%\bibitem[\protect\citeauthoryear{{Oh} \& {Kroupa}}{{Oh} \&
-%  {Kroupa}}{2016}]{oh:16}
-%{Oh} S.,  {Kroupa} P.,  2016, \mn@doi [\aap] {10.1051/0004-6361/201628233},
-%  \href {http://adsabs.harvard.edu/abs/2016A\%26A...590A.107O} {590, A107}
-%
-%\bibitem[\protect\citeauthoryear{{Parker}}{{Parker}}{1993}]{parker:93}
-%{Parker} J.~W.,  1993, \mn@doi [\aj] {10.1086/116661}, \href
-%  {http://adsabs.harvard.edu/abs/1993AJ....106..560P} {106, 560}
-%
-%\bibitem[\protect\citeauthoryear{{Pietrzy{\'n}ski} et~al.,}{{Pietrzy{\'n}ski}
-%  et~al.}{2013}]{pietrzynski:13}
-%{Pietrzy{\'n}ski} G.,  et~al., 2013, \mn@doi [\nat] {10.1038/nature11878},
-%  \href {http://adsabs.harvard.edu/abs/2013Natur.495...76P} {495, 76}
-%
-%\bibitem[\protect\citeauthoryear{{Platais}, {van der Marel}, {Lennon},
-%  {Anderson}, {Bellini}, {Sabbi}, {Sana}  \& {Bedin}}{{Platais}
-%  et~al.}{2015}]{platais:15}
-%{Platais} I.,  {van der Marel} R.~P.,  {Lennon} D.~J.,  {Anderson} J.,
-%  {Bellini} A.,  {Sabbi} E.,  {Sana} H.,   {Bedin} L.~R.,  2015, \mn@doi [\aj]
-%  {10.1088/0004-6256/150/3/89}, \href
-%  {http://adsabs.harvard.edu/abs/2015AJ....150...89P} {150, 89}
-%
-%\bibitem[\protect\citeauthoryear{{Platais} et~al.,}{{Platais}
-%  et~al.}{2018}]{platais:18}
-%{Platais} I.,  et~al., 2018, ArXiv:1804.08678, \href
-%  {http://adsabs.harvard.edu/abs/2018arXiv180408678P} {}
-%
-%\bibitem[\protect\citeauthoryear{{Poveda}, {Ruiz}  \& {Allen}}{{Poveda}
-%  et~al.}{1967}]{poveda:67}
-%{Poveda} A.,  {Ruiz} J.,   {Allen} C.,  1967, Boletin de los Observatorios
-%  Tonantzintla y Tacubaya, \href
-%  {http://adsabs.harvard.edu/abs/1967BOTT....4...86P} {4, 86}
-%
-%\bibitem[\protect\citeauthoryear{{Rakavy} \& {Shaviv}}{{Rakavy} \&
-%  {Shaviv}}{1967}]{rakavy:67}
-%{Rakavy} G.,  {Shaviv} G.,  1967, \mn@doi [\apj] {10.1086/149204}, \href
-%  {http://adsabs.harvard.edu/abs/1967ApJ...148..803R} {148, 803}
-%
-%% \bibitem[\protect\citeauthoryear{{Ram{\'{\i}}rez-Agudelo}
-%%   et~al.,}{{Ram{\'{\i}}rez-Agudelo} et~al.}{2015}]{ramirez-agudelo:15}
-%% {Ram{\'{\i}}rez-Agudelo} O.~H.,  et~al., 2015, \mn@doi [\aap]
-%%   {10.1051/0004-6361/201425424}, \href
-%%   {http://adsabs.harvard.edu/abs/2015A\%26A...580A..92R} {580, A92}
-%
-%\bibitem[\protect\citeauthoryear{{Renzo} et~al.,}{{Renzo}
-%  et~al.}{2018}]{renzo:18}
-%{Renzo} M.,  et~al., 2018, ArXiv:1804.09164, \href
-%  {http://adsabs.harvard.edu/abs/2018arXiv180409164R} {}
-%
-%% \bibitem[\protect\citeauthoryear{{Robitaille} \& {Bressert}}{{Robitaille} \&
-%%   {Bressert}}{2012}]{robitaille:12}
-%% {Robitaille} T.,  {Bressert} E.,  2012, {APLpy: Astronomical Plotting Library
-%%   in Python}, Astrophysics Source Code Library (\mn@eprint {ascl} {1208.017})
-%
-%\bibitem[\protect\citeauthoryear{{Rosen}, {Krumholz}, {McKee}  \&
-%  {Klein}}{{Rosen} et~al.}{2016}]{rosen:16}
-%{Rosen} A.~L.,  {Krumholz} M.~R.,  {McKee} C.~F.,   {Klein} R.~I.,  2016,
-%  \mn@doi [\mnras] {10.1093/mnras/stw2153}, \href
-%  {http://adsabs.harvard.edu/abs/2016MNRAS.463.2553R} {463, 2553}
-%
-%\bibitem[\protect\citeauthoryear{{Rubio-D{\'{\i}}ez}, {Najarro},
-%  {Garc{\'{\i}}a}  \& {Sundqvist}}{{Rubio-D{\'{\i}}ez}
-%  et~al.}{2017}]{rubio-diez:17}
-%{Rubio-D{\'{\i}}ez} M.~M.,  {Najarro} F.,  {Garc{\'{\i}}a} M.,   {Sundqvist}
-%  J.~O.,  2017, in {Eldridge} J.~J.,  {Bray} J.~C.,  {McClelland} L.~A.~S.,
-%  {Xiao} L.,  eds,  IAU Symposium Vol. 329, The Lives and Death-Throes of
-%  Massive Stars. pp 131--135, \mn@doi{10.1017/S1743921317002447}
-%
-%\bibitem[\protect\citeauthoryear{{Sabbi} et~al.,}{{Sabbi}
-%  et~al.}{2012}]{sabbi:12}
-%{Sabbi} E.,  et~al., 2012, \mn@doi [\apjl] {10.1088/2041-8205/754/2/L37}, \href
-%  {http://adsabs.harvard.edu/abs/2012ApJ...754L..37S} {754, L37}
-%
-%\bibitem[\protect\citeauthoryear{{Sana}, {Ram{\'{\i}}rez-Tannus}, {de Koter},
-%  {Kaper}, {Tramper}  \& {Bik}}{{Sana} et~al.}{2017}]{sana:17}
-%{Sana} H.,  {Ram{\'{\i}}rez-Tannus} M.~C.,  {de Koter} A.,  {Kaper} L.,
-%  {Tramper} F.,   {Bik} A.,  2017, \mn@doi [\aap]
-%  {10.1051/0004-6361/201630087}, \href
-%  {http://adsabs.harvard.edu/abs/2017A\%26A...599L...9S} {599, L9}
-%
-%% \bibitem[\protect\citeauthoryear{{Schneider}, {Castro}, {Fossati}, {Langer}  \&
-%%   {de Koter}}{{Schneider} et~al.}{2017}]{schneider:17}
-%% {Schneider} F.~R.~N.,  {Castro} N.,  {Fossati} L.,  {Langer} N.,   {de Koter}
-%%   A.,  2017, \mn@doi [\aap] {10.1051/0004-6361/201628409}, \href
-%%   {http://adsabs.harvard.edu/abs/2017A\%26A...598A..60S} {598, A60}
-%
-%\bibitem[\protect\citeauthoryear{{Schneider} et~al.,}{{Schneider}
-%  et~al.}{2018}]{schneider:18}
-%{Schneider} F.~R.~N.,  et~al., 2018, \mn@doi [Science]
-%  {10.1126/science.aan0106}, \href
-%  {http://adsabs.harvard.edu/abs/2018Sci...359...69S} {359, 69}
-%
-%\bibitem[\protect\citeauthoryear{{Smith}}{{Smith}}{2016}]{smith:16}
-%{Smith} N.,  2016, \mn@doi [\mnras] {10.1093/mnras/stw1533}, \href
-%  {http://adsabs.harvard.edu/abs/2016MNRAS.461.3353S} {461, 3353}
-%
-%\bibitem[\protect\citeauthoryear{{Smith} \& {Tombleson}}{{Smith} \&
-%  {Tombleson}}{2015}]{smith:15}
-%{Smith} N.,  {Tombleson} R.,  2015, \mn@doi [\mnras] {10.1093/mnras/stu2430},
-%  \href {http://adsabs.harvard.edu/abs/2015MNRAS.447..598S} {447, 598}
-%
-%\bibitem[\protect\citeauthoryear{{Udalski} et~al.,}{{Udalski}
-%  et~al.}{2008}]{udalski:08}
-%{Udalski} A.,  et~al., 2008, \actaa, \href
-%  {http://adsabs.harvard.edu/abs/2008AcA....58..329U} {58, 329}
-%
-%\bibitem[\protect\citeauthoryear{{Ward} \& {Kruijssen}}{{Ward} \&
-%  {Kruijssen}}{2018}]{ward:18}
-%{Ward} J.~L.,  {Kruijssen} J.~M.~D.,  2018, \mn@doi [\mnras]
-%  {10.1093/mnras/sty117}, \href
-%  {http://adsabs.harvard.edu/abs/2018MNRAS.475.5659W} {475, 5659}
-%
-%\bibitem[\protect\citeauthoryear{{Zapartas} et~al.,}{{Zapartas}
-%  et~al.}{2017}]{zapartas:17}
-%{Zapartas} E.,  et~al., 2017, \mn@doi [\aap] {10.1051/0004-6361/201629685},
-%  \href {http://adsabs.harvard.edu/abs/2017A\%26A...601A..29Z} {601, A29}
-%
-%\bibitem[\protect\citeauthoryear{{Zinnecker} \& {Yorke}}{{Zinnecker} \&
-%  {Yorke}}{2007}]{zinnecker:07}
-%{Zinnecker} H.,  {Yorke} H.~W.,  2007, \mn@doi [\araa]
-%  {10.1146/annurev.astro.44.051905.092549}, \href
-%  {http://adsabs.harvard.edu/abs/2007ARA\%26A..45..481Z} {45, 481}
-%
-%\bibitem[\protect\citeauthoryear{{Zwicky}}{{Zwicky}}{1957}]{zwicky:57}
-%{Zwicky} F.,  1957, \zap, \href
-%  {http://adsabs.harvard.edu/abs/1957ZA.....44...64Z} {44, 64}
-%
-%
-%% \bibitem[\protect\citeauthoryear{{de Mink}, {Cantiello}, {Langer}, {Pols},
-%%   {Brott}  \& {Yoon}}{{de Mink} et~al.}{2009}]{demink:09}
-%% {de Mink} S.~E.,  {Cantiello} M.,  {Langer} N.,  {Pols} O.~R.,  {Brott} I.,
-%%   {Yoon} S.-C.,  2009, \mn@doi [\aap] {10.1051/0004-6361/200811439}, \href
-%%   {http://adsabs.harvard.edu/abs/2009A\%26A...497..243D} {497, 243}
-
-%\makeatother
-%\end{thebibliography}
-
 \section*{Affiliations}
@@
     E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
-  $^{3}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+ $^{3}$ {
+ %Center for Astrophysical Sciences, 
+ Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
+  $^{4}$ {Space Telescope Science Institute, 3700 San Martin Drive,
     Baltimore, MD 21218, USA}\\
-  $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
   $^{5}${Department of Physics and Astronomy, Hicks Building,

### 2018-06-23 VFTS682_APJL.tex

@@
       (2)~\cite{bestenlehner:11}
-      (3)~\cite{rubio-diez:17}
+      % (3)~\cite{rubio-diez:17}
     }
   \end{center}
-  \label{tab:star}
+  \label{tab:star_param}
 \end{table}
@@
       &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
-      &$\theta$  &[degrees] &  $14.2_{-43}^{+44}$  & (1,5)\\  
+      &$\theta$  &[degrees] &  $14_{-43}^{+44}$  & (1,5)\\  
       % &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
@@
 %
-\citet{bestenlehner:11}  analyzed the spectra to infer the stellar parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of the brightest stars in the region.
-The star is unlikely to have a close companion, because of the absence of clear radial velocity variations \citet{bestenlehner:11}.
-Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see \Tabref{tab:star} for an overview of the parameters.
 
-Estimates of the radial velocity are complicated by the variable, optically thick winds that is typical for emission line stars.
-\citet{bestenlehner:11} estimate a mass loss rate of $10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the possible effect of clumping.
-The V-band light curve of VFTS682  shows variations at a $\sim$10\% level on a timescale of years, which is unusual for Wolf-Rayet stars and more typical for Luminous Blue Variable (LBV) stars \citep{udalski:08, bestenlehner:11}.
-It also shows mid-infrared excess \citep{gruendl:09}.
+\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
+parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
+luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
+the brightest stars in the region. The star is unlikely to have a
+close companion, because of the absence of clear radial velocity
+variations \citet{bestenlehner:11}. Bayesian fits of the stellar
+parameters against evolutionary tracks \citep{brott:11, kohler:15}
+using the Bonsaii code \citep{schneider:17, schneider:18} provide
+estimates for the age, present mass and initial mass, indicating that
+this star among the most massive stars known, see
+\Tabref{tab:star_param} for an overview of the parameters.
+
+\cite{townsley:06} did not find
+an X-ray point source at the position of VFTS682 in their
+\emph{Chandra} survey of the region, which suggests the absence of
+colliding winds that would be expected in the presence of companions
+even for extreme mass ratios, given the extremely large mass of VFTS682.
+
+Estimates of the radial velocity are complicated by the variable,
+optically thick winds that is typical for emission line
+stars. \citet{bestenlehner:11} estimate a mass loss rate of
+$10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the
+possible effect of clumping.  The V-band light curve of VFTS682  shows
+variations at a $\sim$10\% level on a timescale of years, which is
+unusual for Wolf-Rayet stars and more typical for Luminous Blue
+Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
+shows mid-infrared excess \citep{gruendl:09}.
 We therefore caution for over-interpreting the estimates for the radial velocities, but we will summarize the estimates that have been published.
@@
 This is consistent with, but also no proof of, the expectation if the star was not formed in situ.
-
-Newer measurements based on a fit to new Xshooter data covering the full spectrum from 3200 Angstr{\"o}m to 2.45 microns lead to an estimate of $260 \pm 15 \kms$ ({\color{blue} Rubio-D{\' i}ez et al.
+Newer measurements based on new Xshooter data give different estimates for the RV ({\color{blue} Rubio-D{\' i}ez et al.
 in prep}).
-This estimate, which we consider to be the most reliable fo those quoted here, indicates that the relative motion along the line-of-sight is small or negligible.
-This in consistent with both hypotheses.
+We will therefore refrain from using the RV measurements in this work.
 
 %of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
+VFTS682 is relatively isolated in \emph{Spitzer} images, except for the extended
+source S10 interpreted as star formation triggered by the wind of
+VFTS682 itself by \cite{walborn:13}.
 
@@
       yr^{-1}}$.\todo{possibly: cartoon to illustrate angle definition, or add to
-    fig.2 if not too messy}}
+    fig.2 if not too messy, maybe also show HST measurements on this
+    plot for 682? Redo right panel taking care of projection of the
+    radial directions!It's misleading!}}
   \label{fig:dist}
@@
 the angle $\theta$ between the direction of motion and the vector
-connecting the center of R136 with the current position of VFTS 682.
-This means that $\theta = 0$ is the angle that is expected for a star that is moving away from R136 with a perfect radial motion.
-All kinematic quantities are provided in
+connecting the center of R136 with the current position of VFTS 682,
+i.e. such that $\theta = 0$ is the angle corresponding to perfectly
+radial motion away from R136.    We further provide the projected 2d velocity
+$v_\mathrm{2d}$.   All kinematic quantities are provided in
 Tab.~\ref{tab:vfts682}.
@@
 highlighted by the green vertical band, and its width is determined by
-the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star}).
+the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star_param}).
 
-The right panel shows the distributions in flight direction for these
-stars: $\theta$ is the angle between the radial direction from the
-center of R136 to each star and its proper motion relative to
-R136 (see also \Figref{fig:main}). Dynamical ejections from the
+The right panel shows the distributions in projected flight direction for these
+stars: $\theta$ is the angle between the projected radial direction from the
+center of R136 to each star and its projected relative proper motion (see also \Figref{fig:main}). Dynamical ejections from the
 cluster should %are expected to
 produce
-close to radial ejections, i.e.~$\theta\simeq0$. On this panel, the green band highlights
+close to radial ejections, i.e.~$\theta\simeq0$ \todo{this is only
+  true for small RV}. On this panel, the green band highlights
 the range of angles corresponding to an opening angle around the
-radial direction of 45\,degrees. The direction of motion of VFTS682 is
+projected radial direction of 45\,degrees. The direction of motion of VFTS682 is
 poorly constrained because of the large errors on the proper motion
 components both in HST and \emph{Gaia} data, but
-it is consistent with dynamical ejection from R136.
+it is consistent with dynamical ejection from R136. We note that
+VFTS72 has a small radial velocity, while VFTS16 (and possibly VRTS682) has
+large peculiar radial velocity and therefore accurate distances are
+needed to contrain the flight direction in three dimensions.
 
@@
 Based on our results, we tentatively claim that VFTS682 is the most massive
-(slow) runaway known to date, with a peculiar three-dimensional speed of
-$44\pm21\,\kms$ \todo{make 2d? use LOS velocity from Marimar?}. Due to the large error bars, this result will need
+(slow) runaway known to date, with a peculiar two-dimensional
+projected velocity with respect to R136 of
+$\sim38\kms$ (averaging the \emph{Gaia} DR2 and HST
+results) but compatible with zero within $2\sigma$. Due to the large error bars, this result will need
 to be revisited with future astrometric data. % to be revisited as updated

### 2018-06-25 VFTS682_APJL.tex
abstract editing

@@
 
-\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, \SdM{I.~Platais$^{3}$,}
-  R.~P.~van~der~Marel$^{3,4}$,  \SdM{E.~L.~Laplace$^{1}$,}
+\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
+  R.~P.~van~der~Marel$^{3,4}$, E.~L.~Laplace$^{1}$,
   J.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
@@
 
- How very massive stars form is still among the most intriguing open  questions in stellar astrophysics.
-VFTS 682 is among the most  extreme massive stars known, with an inferred initial mass of about  $150\,M_\odot$.
+ How very massive stars form is still an intriguing open question in stellar astrophysics.
+VFTS682 is among the most massive stars known, with an inferred initial mass of about  $150\,M_\odot$.
 It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.
@@
 
-Aiming to shine light on this debate, we investigate the kinematics of  VFTS 682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space Telescope} photometry.
-We derive a projected velocity relative of to the cluster of $\sim$$30\pm20\kms$.
-The direction and magnitude of the proper is consistent with what is expected if VFTS 682 was ejected as a slow runaway from the central cluster, given its distance and the inferred age.
+Aiming to shine light on this debate, we investigate the kinematics of
+VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space
+  Telescope} photometry. We derive a projected velocity relative of to
+the cluster of $\sim$$30\pm20\kms$.  The direction and magnitude of
+the proper motion are consistent with what is expected if VFTS682 was ejected as a slow runaway from the central cluster, given its distance and the inferred age.
 However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.
 
-If future data confirms the runaway nature, this would make the VFTS682 the most massive runaway star known to date.
-While we cannot prove this solidly from the current data alone, we do consider this hypothesis as the most plausible because of a variety of circumstantial clues.
-The central cluster is known to harbor several other stars more massive than $150\,M_\odot$ similar in spectral type.
-The very massive O star VFTS 16, which has recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions by numerical simulations.
+If future data confirm the runaway nature, this would make the VFTS682
+the most massive runaway star known to date.  While we cannot prove
+this solidly from the current data alone, we do consider this
+hypothesis as the most plausible because of a variety of
+circumstantial clues. The central cluster is known to harbor several
+other stars more massive than $150\,M_\odot$ similar in spectral type.
+The very massive O stars VFTS16 and VFTS72, which have recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
 \end{abstract}

### 2018-06-25 VFTS682_APJL.tex
sec. 2 and 3 edited

@@
 constraints for VFTS682 provided by \emph{Gaia} DR2 and constraints
-from HST proper data  by \citet{platais:18}.   We discuss the
+from HST proper motions  by \citet{platais:18}.   We discuss the
 implications for the hypothesis that VFTS682 is a slow runaway star
@@
 
+The WNh5 star VFTS682, located at right ascension (RA)
+05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
+\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
+parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
+luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
+the brightest stars in the region. The star is unlikely to have a
+close companion, because of the absence of clear radial velocity
+variations \citet{bestenlehner:11}. Bayesian fits of the stellar
+parameters against evolutionary tracks \citep{brott:11, kohler:15}
+using the BONNSAI code \citep{schneider:17, schneider:18} provide
+estimates for the age, present mass and initial mass, indicating that
+this star among the most massive stars known, see
+\Tabref{tab:star_param}. % for an overview of the parameters.
+
 \begin{table}
@@
                      &  SIMBAD \\       
-        &$      \delta\mathrm{RA}$  &[degrees] & \phantom{-0}0.0547                      
+        &$      \delta\mathrm{RA}$  &[marcsec] & \phantom{-0}0.0547                      
         &  (2, 5)
   \\        
-               &$     \delta\mathrm{DEC}$  &[degrees] & \phantom{-0}0.0268 
+               &$     \delta\mathrm{DEC}$  &[marcsec] & \phantom{-0}0.0268 
                      &  (2, 5) \\  
                        &$  \mathrm{d_\mathrm{2d}}$  &[pc] & 29
-                     &  (Besthenlehner) \\
+                     &  (2) \\
 
@@
       \hline
-      &$v_\mathrm{2d}$  &[$\kms$] & \SdM{$xx\pm xx$} & (Bestenlehner) \\  
-      &$\theta$  &[degrees] &  $0$  & \\ 
+      &$v_\mathrm{2d}$  &[$\kms$] & $29\pm 6$ & (2) \\  
+      &$\theta$  &[degrees] &  $\sim0$  & \\ 
        \hline
@@
 
-%\subsection{ Overview of VFTS682 from previous studies \label{data:vfts683}}
-
-The WNh5 star VFTS682, located at right ascension (RA)
-05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
-%
-%of spectral type WNh5 using multi-epoch spectra covering $\lambda$4000--7000 from the
-%VFTS survey. % The spectra available covered $\lambda$4000--7000 and
-% were taken at multiple epochs.
-%
-
-\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
-parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
-luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
-the brightest stars in the region. The star is unlikely to have a
-close companion, because of the absence of clear radial velocity
-variations \citet{bestenlehner:11}. Bayesian fits of the stellar
-parameters against evolutionary tracks \citep{brott:11, kohler:15}
-using the Bonsaii code \citep{schneider:17, schneider:18} provide
-estimates for the age, present mass and initial mass, indicating that
-this star among the most massive stars known, see
-\Tabref{tab:star_param} for an overview of the parameters.
-
 \cite{townsley:06} did not find
@@
 even for extreme mass ratios, given the extremely large mass of VFTS682.
+This star is also relatively isolated in \emph{Spitzer} images, except for the extended
+source S10 interpreted as star formation triggered by the wind of
+VFTS682 itself by \cite{walborn:13}.
 
 Estimates of the radial velocity are complicated by the variable,
-optically thick winds that is typical for emission line
+possibly inhomogeneous, optically thick wind  typical of emission line
 stars. \citet{bestenlehner:11} estimate a mass loss rate of
@@
 \cite{bressert:12} note an offset compared between the RV of the star and the nebular lines from the gas filaments in its vicinity.
-This is consistent with, but also no proof of, the expectation if the star was not formed in situ.
-Newer measurements based on new Xshooter data give different estimates for the RV ({\color{blue} Rubio-D{\' i}ez et al.
+This is consistent with the expectation if the star was not formed in situ.
+Newer measurements based on new X-shooter data give different estimates for the RV ({\color{blue} Rubio-D{\' i}ez et al.
 in prep}).
@@
 %of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
-VFTS682 is relatively isolated in \emph{Spitzer} images, except for the extended
-source S10 interpreted as star formation triggered by the wind of
-VFTS682 itself by \cite{walborn:13}.
 
@@
 
-For the distance we adopt the distance to the LMC of $50$\,kpc.
+We adopt as distance to the LMC the value of $50$\,kpc.
 The error on the distance determination is small ($\lesssim2\%$, \citealt{pietrzynski:13}) and any possible offset between  R136 or VFTS682  and the distance we adopted for the LMC is probably much smaller (likely $\ll 1\%$, e.g.
@@
 \citep{gaia:16,brown:18}.   The number of visibility periods,
-i.e. groups of observations separated from other groups by a gap of at
+i.e., groups of observations separated from other groups by a gap of at
 least four days, used in the astrometric solution is seventeen for this
@@
 suggest that the \emph{Gaia} DR2  data for VFTS682 are
-reliable.
+reliable. % We now include the correlation coefficient in the error
+          % calculation, so no need for this paragraph?
 Gaia errors are correlated \citep[][]{lindengren:18}, but for this star the correlation coefficient is very small, $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$  = 0.026, where zero is for completely un-correlated data and one is for fully correlated data.
@@
 
-The 30 doradus region was target of an a two-epoch photometric campaign with HST providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
+The 30 Doradus region was target of an a two-epoch photometric campaign with HST providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
 Visually, the star is clearly isolated and does not appear to have any close neighbor.
@@
 
- \citet{platais:15, platais:18} analyzed the HST data to determine the relative proper motion and identify candidate runaway stars.
-The brightest stars (V$<$14) are saturated in the data set and have been excluded from the analysis.
-The very massive star VFTS682 should be one of the brightest stars in the region, the high extinction makes it faint enough such that saturation effects are not an issue.
+\citet{platais:15, platais:18} analyzed the HST data to determine the
+relative proper motion and identify candidate runaway stars. The
+brightest stars (V$<$14) are saturated in the data set and have been
+excluded from the analysis.  VFTS682 should be one of the brightest
+stars in the region, however, the high extinction makes it faint enough such that saturation effects are not an issue.
 VFTS682 is relatively slow and is therefore not discussed in their paper.
-However, their measurements are useful here, since they provides an  estimate of the proper motion of VFTS 682 that is independent from the \emph{Gaia} measurement.
+However, their measurements are useful here, since they provides an  estimate of the proper motion of VFTS682 that is independent from the \emph{Gaia} measurement.
 
@@
 
-At first glance the magnitude of the relative proper motions derived by Gaia,  $32\pm 21\kms$, and by HST, $47 \pm 24\kms$, appear consistent.  In principle they are consistent with the velocity
+% At first glance the magnitude of the relative proper motions derived by Gaia,  $32\pm 21\kms$, and by HST, $47 \pm 24\kms$, appear consistent.  In principle they are consistent with the velocity
 
@@
 we obtain $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$ and
-$\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$. Both indicate
+$\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$. Both values
+are consistent with each other and indicate
 a non-zero peculiar projected motion that would make it a slow runaway
@@
 produce
-close to radial ejections, i.e.~$\theta\simeq0$ \todo{this is only
-  true for small RV}. On this panel, the green band highlights
-the range of angles corresponding to an opening angle around the
+close to radial ejections, i.e.~$\theta\simeq0$. On this panel, the green band highlights
+% the range of angles corresponding to
+an opening angle around the
 projected radial direction of 45\,degrees. The direction of motion of VFTS682 is
@@
 it is consistent with dynamical ejection from R136. We note that
-VFTS72 has a small radial velocity, while VFTS16 (and possibly VRTS682) has
+VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
 large peculiar radial velocity and therefore accurate distances are
@@
 The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
-ejection from the cluster, given is apparent age.
+ejection from the cluster, given the apparent age of the star.
 % of $1.0\pm 0.2$\,Myr 
@@
 Therefore, although the motion of VFTS682 is not an outlier compared
-to other massive stars, it is possible to tentatively claim that it is
+to other massive stars with small proper motion errors in \emph{Gaia} DR2, it is possible to tentatively claim that it is
 a slow runaway with velocity compatible with dynamical ejection from

### 2018-06-25 VFTS682_APJL.tex
sec. discussion edited

@@
     lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$.\todo{possibly: cartoon to illustrate angle definition, or add to
-    fig.2 if not too messy, maybe also show HST measurements on this
-    plot for 682? Redo right panel taking care of projection of the
-    radial directions!It's misleading!}}
+      yr^{-1}}$.}
   \label{fig:dist}
@@
   \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
-  \caption{\SdM{I don't see a reason why we would not show the HST data, except for wanting to hide it maybe.
-Let's sit together and find a good way to plot both.
-(Hope you still have the script to show the HST direction.
-)}  The red solid arrow indicates the proper motion of VFTS682
+  \caption{The red solid arrow indicates the proper motion of VFTS682 
     relative to the region from \emph{Gaia} DR2, starting from the present day position of

### 2018-06-25 VFTS682_APJL.tex
selma

--- 2018-06-25 VFTS682_APJL.tex
@@
 
-\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
-  R.~P.~van~der~Marel$^{3,4}$, E.~L.~Laplace$^{1}$,
+\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, \SdM{I.~Platais$^{3}$,}
+  R.~P.~van~der~Marel$^{3,4}$,  \SdM{E.~L.~Laplace$^{1}$,}
   J.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
@@
 
- How very massive stars form is still an intriguing open question in stellar astrophysics.
-VFTS682 is among the most massive stars known, with an inferred initial mass of about  $150\,M_\odot$.
+ How very massive stars form is still among the most intriguing open  questions in stellar astrophysics.
+VFTS 682 is among the most  extreme massive stars known, with an inferred initial mass of about  $150\,M_\odot$.
 It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.
@@
 
-Aiming to shine light on this debate, we investigate the kinematics of
-VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space
-  Telescope} photometry. We derive a projected velocity relative of to
-the cluster of $\sim$$30\pm20\kms$.  The direction and magnitude of
-the proper motion are consistent with what is expected if VFTS682 was ejected as a slow runaway from the central cluster, given its distance and the inferred age.
+Aiming to shine light on this debate, we investigate the kinematics of  VFTS 682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space Telescope} photometry.
+We derive a projected velocity relative of to the cluster of $\sim$$30\pm20\kms$.
+The direction and magnitude of the proper is consistent with what is expected if VFTS 682 was ejected as a slow runaway from the central cluster, given its distance and the inferred age.
 However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.
 
-If future data confirm the runaway nature, this would make the VFTS682
-the most massive runaway star known to date.  While we cannot prove
-this solidly from the current data alone, we do consider this
-hypothesis as the most plausible because of a variety of
-circumstantial clues. The central cluster is known to harbor several
-other stars more massive than $150\,M_\odot$ similar in spectral type.
-The very massive O stars VFTS16 and VFTS72, which have recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
+If future data confirms the runaway nature, this would make the VFTS682 the most massive runaway star known to date.
+While we cannot prove this solidly from the current data alone, we do consider this hypothesis as the most plausible because of a variety of circumstantial clues.
+The central cluster is known to harbor several other stars more massive than $150\,M_\odot$ similar in spectral type.
+The very massive O star VFTS 16, which has recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions by numerical simulations.
 \end{abstract}
@@
 The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
-Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, renzo:18}.
+Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the \SdM{ companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, eldridge:11, renzo:18}}.
 
@@
 (DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}.
+
 %
@@
 luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
-the brightest stars in the region. The star is unlikely to have a
-close companion, because of the absence of clear radial velocity
-variations \citet{bestenlehner:11}. Bayesian fits of the stellar
-parameters against evolutionary tracks \citep{brott:11, kohler:15}
+the brightest stars in the region. 
+%
+\SdM{The star does not show clear radial velocity variations, which makes the presence of a close companion unlikely
+ \citet{bestenlehner:11}.  There are also no indications for the presence of a companion based on observations with the Chandra X-Ray Observatory. 
+The detection of a point source in X-rays could have pointed at the presence of a compact companion or a colliding wind binary \citep{townsley:06}. }
+% 
+%\emph{Chandra} survey of the region, which suggests the absence of colliding winds that would be expected in the presence of companions even for extreme mass ratios, given the extremely large mass of VFTS682.
+- Selma: that depends on the separation.
+I don't know what the predictions are for very unequaly winds and the LMC is far away.
+I shortened it a bit and moved this up to merge it where we were discussing the presence of a companion already, to use it in the flow of the argument, instead of disrupting the flow.
+%
+ Bayesian fits of the stellar parameters derived from the spectra against single star evolutionary tracks \citep{brott:11, kohler:15}
 using the Bonsaii code \citep{schneider:17, schneider:18} provide
@@
 \Tabref{tab:star_param} for an overview of the parameters.
-
-\cite{townsley:06} did not find
-an X-ray point source at the position of VFTS682 in their
-\emph{Chandra} survey of the region, which suggests the absence of
-colliding winds that would be expected in the presence of companions
-even for extreme mass ratios, given the extremely large mass of VFTS682.
 
@@
 
+\SdM{VFTS682 also appears to be relatively isolated in (near) infrared.
+There nearest bright (near) infrared sources detected by \emph{Spitzer}  \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e.
+about 2.4 pc.
+\cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but we chance alignment cannot be excluded.
+}
+
+% interpreted as star formation triggered by the wind of  VFTS682 itself by \cite{walborn:13}.
+- SdM: too strong lets be more careful.
+This is a controversial topic.
+Walborn formulates more careful than you summarize him.
+
 %of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
-VFTS682 is relatively isolated in \emph{Spitzer} images, except for the extended
-source S10 interpreted as star formation triggered by the wind of
-VFTS682 itself by \cite{walborn:13}.
 

### 2018-06-25 VFTS682_APJL.tex
needs merger

--- 2018-06-25 VFTS682_APJL.tex
@@
 
-\cite{townsley:06} did not find
+The WNh5 star VFTS682, located at right ascension (RA)
+05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}.
+
+\cite{townsley:06} and \textcolor{blue}{Townsley et al., in prep.} did not find
 an X-ray point source at the position of VFTS682 in their
@@
 VFTS682 itself by \cite{walborn:13}.
+
+\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
+parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
+luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
+the brightest stars in the region. 
+%
+The star does not show clear radial velocity variations, which makes the presence of a close companion unlikely
+ \citet{bestenlehner:11}. 
+% 
+%\emph{Chandra} survey of the region, which suggests the absence of colliding winds that would be expected in the presence of companions even for extreme mass ratios, given the extremely large mass of VFTS682.
+- Selma: that depends on the separation.
+I don't know what the predictions are for very unequaly winds and the LMC is far away.
+I shortened it a bit and moved this up to merge it where we were discussing the presence of a companion already, to use it in the flow of the argument, instead of disrupting the flow.
+%
+ Bayesian fits of the stellar parameters derived from the spectra against single star evolutionary tracks \citep{brott:11, kohler:15}
+using the Bonsaii code \citep{schneider:17, schneider:18} provide
+estimates for the age, present mass and initial mass, indicating that
+this star among the most massive stars known, see
+\Tabref{tab:star_param} for an overview of the parameters.
+
+% Observations with the Chandra X-Ray Observatory do not reveal an X-ray point source at the position of VFTS682 \cite{townsley:06}.
+Such a detection would be expected if the star hosted an accreting compact object or if the strong winds of VFTS682 are colliding with those of a stellar companion.
+This star is also relatively isolated in \emph{Spitzer} images \citep{meixner:06}, except for the extended source S10 interpreted as star formation triggered by the wind of VFTS682 itself by \cite{walborn:13}.
+% >>>>>>> df97f49fb6e094f5c7a3bd15402c437202858eaa
 
@@
 We will therefore refrain from using the RV measurements in this work.
+
+\SdM{VFTS682 also appears to be relatively isolated in (near) infrared.
+There nearest bright (near) infrared sources detected by \emph{Spitzer}  \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e.
+about 2.4 pc.
+\cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but we chance alignment cannot be excluded.
+}
+
+% interpreted as star formation triggered by the wind of  VFTS682 itself by \cite{walborn:13}.
+- SdM: too strong lets be more careful.
+This is a controversial topic.
+Walborn formulates more careful than you summarize him.
 
@@
 participating in the {\it Gaia} Multilateral Agreement. 
-
- % The background image of \Figref{fig:main} is based on observations
-   % made with ESO Telescopes at the La Silla Observatory under programme
-   % ID 076.C-0888, processed and released by the ESO VOS/ADP group.
-   % This research made use of APLpy, an open-source plotting package for Python \citep[][]{robitaille:12}.
 \end{acknowledgements}

### 2018-06-29 VFTS682_APJL.tex
removed Eva's middle initial for consistency with her previous papers

@@
 \author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
-  R.~P.~van~der~Marel$^{3,4}$, E.~L.~Laplace$^{1}$,
+  R.~P.~van~der~Marel$^{3,4}$, E.~Laplace$^{1}$,
   J.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,

### 2018-07-01 VFTS682_APJL.tex
acknowledgements
update to fig caption and text

@@
 Figure~\ref{fig:main} shows the projected motion of VFTS682 relative to R136 on the
-sky. \todo{fix colors} The dashed circle shows the location of the cluster, and the red
-(cyan) arrow shows the relative proper motion from \emph{Gaia} DR2
-(HST). The lighter arrows show the uncertainty on the direction of
-motion due to the large uncertainties. The length of each arrow is proportional to
-the speed of the star, and dashed prolungations show the possible
-region of origin assuming an age of $1.0\pm0.2$\,Myr
-\citep[][]{schneider:18}. This figure illustrates that R136 is clearly
-the most likely origin of the star, although the large error bars
-prevent a robust identification.
+sky. We also show VFTS16 and VFTS72 for completeness
+\citep[see][]{lennon:18}. The thick cyan arrows are proportional to
+the relative proper motion from \emph{Gaia} DR2, and the yellow thick
+line illustrates the relative proper motion from HST. The error cone
+on the direction of motion is illustrated by the corresponding
+prolongation in the direction opposite to the motion, and we also show
+the most likely origin of the stars accounting for their apparent age.
+This figure illustrates that R136 is the most likely origin of these stars, although the large error bars
+prevent a robust identification for VFTS682, and there is some tension
+between the apparent age and the present day distance from the cluster
+core for VFTS16 and VFTS72 \citep[][]{lennon:18}.
 
@@
   \centering
-  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
-  \caption{The red solid arrow indicates the proper motion of VFTS682 
-    relative to the region from \emph{Gaia} DR2, starting from the present day position of
-    the star. The yellow arrows indicate the possible
-    directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
-    backwards (dashed) to illustrate the uncertainty on the origin of the
-    star. The length of the prolongations is proportional to the relative proper motion
-    times the age of VFTS682 \citep[$1.0\pm0.2$\,Myr,][]{schneider:18}.
-    % \todo{load color picture on background? Fig with HST data
-    % available, but confusing...}
+  \includegraphics[width=0.48\textwidth]{./figures/Figure2_paper_vfts682_hst}  
+  \caption{The thick cyan arrows indicate proper motions relative R136 (blue
+    circle) from \emph{Gaia} DR2, multiplied by 1\,Myr. The prolongations in the opposite
+    direction are proportional to the apparent ages of the stars, and
+    the thin cyan lines illustrate the error cone. The yellow arrow
+    indicates the HST proper motion of VFTS682, with the relative
+    error cone and prolongation proportional to its apparent age.
     }
@@
    \small
-   We are grateful to M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
+   We are grateful to J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
    % and C.\,J.\,Evans

### 2018-07-05 VFTS682_APJL.tex
updated sec.3 and caption/legend of fig.1

@@
 The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
-Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the \SdM{ companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, eldridge:11, renzo:18}}.
+Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, eldridge:11, renzo:18}.
 
@@
 
-% Observations with the Chandra X-Ray Observatory do not reveal an X-ray point source at the position of VFTS682 \cite{townsley:06}.
-Such a detection would be expected if the star hosted an accreting compact object or if the strong winds of VFTS682 are colliding with those of a stellar companion.
-This star is also relatively isolated in \emph{Spitzer} images \citep{meixner:06}, except for the extended source S10 interpreted as star formation triggered by the wind of VFTS682 itself by \cite{walborn:13}.
-% >>>>>>> df97f49fb6e094f5c7a3bd15402c437202858eaa
-
 Estimates of the radial velocity are complicated by the variable,
@@
 
-\SdM{VFTS682 also appears to be relatively isolated in (near) infrared.
-There nearest bright (near) infrared sources detected by \emph{Spitzer}  \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e.
+VFTS682 also appears to be relatively isolated in (near) infrared.
+There nearest bright (near-)infrared sources detected by \emph{Spitzer} \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e.
 about 2.4 pc.
 \cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but we chance alignment cannot be excluded.
-}
-
-% interpreted as star formation triggered by the wind of  VFTS682 itself by \cite{walborn:13}.
-- SdM: too strong lets be more careful.
-This is a controversial topic.
-Walborn formulates more careful than you summarize him.
-
-%of $137.8^{+27.5}_ {-15.9}\Msun$, an apparent age of $1.0^{+0.2}_{-0.2}$\,Myr and an inferred initial mass of $150.0^{+28.7}_{-17.4}\Msun$.  Further parameters of are summarized in Table~\ref{X}.
 
@@
 %\citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude lower than found by \citet{evans:11}.  - Citing OGLE is enough
-
-%\SdM{\citet{bestenlehner:11} further report a wind mass loss rate of $10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, without correcting for the possible effect of clumping.
-The helium surface  abundance is  $Y=0.45\, (0.49)$ \cite{bestenlehner:11} (\citealt{rubio-diez:17}).
-IN TABLE ONLY TO SAFE SPACE}
 
@@
     the best value is in agreement with the hypothesis of dynamical
-    ejection. In both
+    ejection. The peak at $\theta\simeq\pi/2$ is due to stars
+    belonging to NGC2060. In both
     panels, the dark blue histograms contain 317 
@@
     lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$.}
+      yr^{-1}}$. }
   \label{fig:dist}
@@
 suggest that the \emph{Gaia} DR2  data for VFTS682 are
-reliable. % We now include the correlation coefficient in the error
-          % calculation, so no need for this paragraph?
-Gaia errors are correlated \citep[][]{lindengren:18}, but for this star the correlation coefficient is very small, $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$  = 0.026, where zero is for completely un-correlated data and one is for fully correlated data.
-We can thus safely treat the proper motion components as uncorrelated.
+reliable.
 
@@
 \subsection{HST astrometry for VFTS682}
-% observations; the first epoch (GO-12499; P.I.: D.~J.~Lennon)
-% in October 3-8 2011 and the second epoch (GO-13359) in
-% October 6-11 2014, both using the F775W filter.
-% The second epoch was designed to match the pointing and
-% orientation of the first epoch as exactly as possible so
-% that the displacement measurements could be purely differential.
-
-%For large distances ($\gtrsim50$\,kpc), HST astrometry can compete or even exceed \emph{Gaia} DR2 data in quality. \SdM{Mathieu: do you have a ref for this?}
 
@@
 
-%The errorbars are comparable with those we obtain from \emph{Gaia}.  However the direction and angle are different.
-
-%ID     /HST relative mu_x/HST relative mu_y/HST error_mu_x/HST error_my_y/  HST F775W mag/Gaia B-mag?/Gaia relative mu_x/  Gaia relative mu_y/Gaia error_pmRA/Gaia error_pmDec
-
-%682  -0.02   0.19 0.10 0.09 15.175 15.656   0.133   0.166 0.070   0.080
-
-%The catalog from \citet{platais:18} lists the proper motion components of VFTS682 relative to the field, providing a
-%measure of the motion of this star completely independent from
-%\emph{Gaia}. They found proper
-%motion components relative to the field
-%$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
-%  yr^{-1}}$ and
-%$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
-%  yr^{-1}}$, which added in quadrature give a relative proper motion of
-%$\delta \mu^{HST} =0.20 \pm 0.15\masyr$.
-% \begin{equation}
-%   \label{eq:pm_around_HST}
-%   \delta \mu^{HST} = \sqrt{\left(\delta\mu_\mathrm{RA}^{HST}\right)^2+\left(\delta\mu_\mathrm{DEC}^{HST}\right)^2}
-%   = 0.20 \pm 0.15\,\mathrm{mas\
-%   yr^{-1}} \ \ .
-% \end{equation}
-
-%These proper motion components can also be converted into projected
-%velocities assuming a distance of 50\,kpc, obtaining $\delta
-%v_\mathrm{RA}^{HST}=2\pm31\kms$ and $\delta
-%v_\mathrm{RA}^{HST}=47\pm24\kms$% , corresponding to a two-dimensional
-%% projected velocity of $47\pm24\kms$
-%.
-%Adding them in quadrature to the radial velocity
-%measured by \citet{bestenlehner:11} we obtain a three-dimensional velocity of
-%$v_\mathrm{pec}^{HST}=56 \pm 23  \kms$.
-% \begin{equation}
-%   \label{eq:speed_around_HST}
-%   v_\mathrm{pec}^{HST} = \sqrt{\left(\delta v_\mathrm{RA}^{HST}\right)^2
-%     +\left(\delta v_\mathrm{DEC}^{HST}\right)^2+\left(\delta
-%       v_\mathrm{rad}^{VFTS}\right)^2} = 56 \pm 23
-%   \kms \ .
-% \end{equation}
-
 \section{The kinematics of VFTS682}
 \label{sec:results}
-
-% At first glance the magnitude of the relative proper motions derived by Gaia,  $32\pm 21\kms$, and by HST, $47 \pm 24\kms$, appear consistent.  In principle they are consistent with the velocity
 
@@
 including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light shaded
-blue). We highlight VFTS682 together with the two O-type
+blue). The green vertical bands show the expectations for VFTS682 if
+it were ejected from R136. We highlight the star together with the two O-type
 runaways \mbox{(re-)identified} by \cite{lennon:18} (VFTS16 and VFTS72,
-respectively). We emphasize that selecting stars with small error bars
-implicitly biases the sample toward fast moving objects.
+respectively).
 
 The left panel shows the distribution in two-dimensional proper motion relative
-to R136, and shows that VFTS682 is not a peculiar star in terms of
-its motion relative to R136. Subtracting the mean motion of the field,
-we obtain $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$ and
+to R136. Although VFTS682 is not peculiar, its relative proper motion
+falls in the range expected for it to reach its present day isolation
+within its apparent age (green vertical band), hinting at it being a ``slow-runaway'' as
+suggested by \cite{bestenlehner:11}. Subtracting
+the mean motion of R136, we obtain relative proper motions (projected
+velocities) of $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$ and
 $\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$. Both values
-are consistent with each other and indicate
-a non-zero peculiar projected motion that would make it a slow runaway
-star as expected by \cite{bestenlehner:11} if the star was ejected
-from the cluster,  but in both cases the meaurements are also consistent with zero within 2$\sigma$. 
-More importantly, both the proper motion
-measurement from \emph{Gaia} and HST fall in the range expected to
-reach its present day location % at $\sim$29\,pc from the cluster core
-if the
-star was ejected very early in its evolution. This range is
-highlighted by the green vertical band, and its width is determined by
-the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star_param}).
+are consistent with each other, but also with no relative motion
+within $2\sigma$.
+
+% The left panel shows the distribution in two-dimensional proper motion relative
+% to R136, and shows that, although VFTS682 is not a peculiar star in terms of
+% its motion relative to R136. Subtracting the mean motion of the field,
+% we obtain $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$ and
+% $\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$. Both values
+% are consistent with each other and indicate
+% a non-zero peculiar projected motion that would make it a slow runaway
+% star as expected by \cite{bestenlehner:11} if the star was ejected
+% from the cluster,  but in both cases the meaurements are also consistent with zero within 2$\sigma$. 
+% More importantly, both the proper motion
+% measurement from \emph{Gaia} and HST fall in the range expected to
+% reach its present day location % at $\sim$29\,pc from the cluster core
+% if the
+% star was ejected very early in its evolution. This range is
+% highlighted by the green vertical band, and its width is determined by
+% the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star_param}).
 
@@
 an opening angle around the
-projected radial direction of 45\,degrees. The direction of motion of VFTS682 is
-poorly constrained because of the large errors on the proper motion
-components both in HST and \emph{Gaia} data, but
-it is consistent with dynamical ejection from R136. We note that
+projected radial direction of 45\,degrees. The direction of motion of VFTS682 % is
+% poorly constrained because of the large errors on the proper motion
+% components both in HST and \emph{Gaia} data, but
+% it
+is consistent with dynamical ejection from R136. We note that
 VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
-large peculiar radial velocity and therefore accurate distances are
+large peculiar radial velocity and therefore accurate distances along
+the line of sight are
 needed to contrain the flight direction in three dimensions.
@@
     indicates the HST proper motion of VFTS682, with the relative
-    error cone and prolongation proportional to its apparent age.
-    }
+    error cone and prolongation proportional to its apparent
+    age.}
 

### 2018-07-05 VFTS682_APJL.tex
change of colors in fig2

@@
 %\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
-\title{Gaia's constraints on the very massive  (initially $\sim$$150\,M_\odot$) runaway star candidate VFTS682}
+\title{Gaia's constraints on the very massive $\sim$$150\,M_\odot$ runaway star candidate VFTS682}
 
@@
 sky. We also show VFTS16 and VFTS72 for completeness
-\citep[see][]{lennon:18}. The thick cyan arrows are proportional to
-the relative proper motion from \emph{Gaia} DR2, and the yellow thick
+\citep[see][]{lennon:18}. The thick yellow arrows are proportional to
+the relative proper motion from \emph{Gaia} DR2, and the thick green
 line illustrates the relative proper motion from HST. The error cone
@@
   \includegraphics[width=0.48\textwidth]{./figures/Figure2_paper_vfts682_hst}  
-  \caption{The thick cyan arrows indicate proper motions relative R136 (blue
-    circle) from \emph{Gaia} DR2, multiplied by 1\,Myr. The prolongations in the opposite
+  \caption{The thick yellow arrows indicate proper motions relative R136 (blue
+    circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The prolongations in the opposite
     direction are proportional to the apparent ages of the stars, and
-    the thin cyan lines illustrate the error cone. The yellow arrow
+    the thin yellow lines illustrate the error cone on the potential
+    origin. The green thick arrow
     indicates the HST proper motion of VFTS682, with the relative

### 2018-07-09 scripts/calc_angle_pm.py
fixed typo in table

@@ -1,0 +1,121 @@
+
+# coding: utf-8
+
+# # This notebooks is intended to calculate the PM angles, being originally written in IDL by D. Lennon
+
+
+import time
+import numpy as np
+import h5py as hp
+import sympy as sp
+import scipy.stats as st
+import glob
+import matplotlib.pyplot as plt
+import psutil
+import matplotlib.cm as cm
+import math as math
+
+from pyspark.sql.functions import floor
+
+#calculate opening angle of proper motion
+
+
+#star='VFTS16'
+#relpmra = -0.336
+#relpmdec = -0.038
+#relpm = np.array([-0.336,-0.038])
+
+#pmra_error = 0.04593
+#pmdec_error = 0.04536
+#pmra_pmdec_corr = 0.10909
+
+# star='VFTS72'
+
+# relpmra = -0.372
+# relpmdec = 0.125
+
+# pmra_error = 0.05038
+# pmdec_error = 0.06077
+# pmra_pmdec_corr = 0.1414
+
+star='VFTS682'
+relpmra =  0.102976611081
+pmra_error = 0.0797415003151
+relpmdec =  0.0855134899438
+pmdec_error = 0.0999780525252
+pmra_pmdec_corr = 0.026013786
+
+# components of the correlation matrix
+c33=pmra_error * pmra_error
+c44=pmdec_error * pmdec_error
+c34=pmra_error * pmdec_error * pmra_pmdec_corr
+
+#calculate semi-major and -minor axes of error ellipse
+a = math.sqrt(0.5*(c33+c44) + 0.5*math.sqrt((c44-c33)**2 +4.*c34*c34))
+b = math.sqrt(0.5*(c33+c44) - 0.5*math.sqrt((c44-c33)**2 +4.*c34*c34))
+theta2 = math.atan2(2*c34,c44-c33)/2.
+#review atan IDL definition with 2 arguments
+#theta2=np.arctan2(c44-c33,2*c34)/2.
+
+#;define plot range
+#range=0.5
+
+#;plot original proper motion and error ellipse
+
+#plot,[0,relpmra],[0,relpmdec],xrange=[-range,range],yrange=[-range,range],title=star
+
[…]

### 2018-07-09 scripts/calc_angle_pm.py
fixed tables

@@ -20,30 +20,32 @@ from pyspark.sql.functions import floor
 #calculate opening angle of proper motion
 
 
-#star='VFTS16'
-#relpmra = -0.336
-#relpmdec = -0.038
-#relpm = np.array([-0.336,-0.038])
+# star='VFTS16'
+# relpmra = -0.336
+# relpmdec = -0.038
+# relpm = np.array([-0.336,-0.038])
 
-#pmra_error = 0.04593
-#pmdec_error = 0.04536
-#pmra_pmdec_corr = 0.10909
+# pmra_error = 0.04593
+# pmdec_error = 0.04536
+# pmra_pmdec_corr = 0.10909
 
-# star='VFTS72'
+star='VFTS72'
 
-# relpmra = -0.372
-# relpmdec = 0.125
+relpmra = -0.372
+relpmdec = 0.125
 
-# pmra_error = 0.05038
-# pmdec_error = 0.06077
-# pmra_pmdec_corr = 0.1414
+pmra_error = 0.05038
+pmdec_error = 0.06077
+pmra_pmdec_corr = 0.1414
 
-star='VFTS682'
-relpmra =  0.102976611081
-pmra_error = 0.0797415003151
-relpmdec =  0.0855134899438
-pmdec_error = 0.0999780525252
-pmra_pmdec_corr = 0.026013786
+# star='VFTS682'
+# relpmra =  0.102976611081
+# pmra_error = 0.0797415003151
+
+# relpmdec =  0.0855134899438
+# pmdec_error = 0.0999780525252
+
+# pmra_pmdec_corr = 0.022639342
 
 # components of the correlation matrix
 c33=pmra_error * pmra_error
@@ -114,7 +116,7 @@ m3 = y/x
 alpha1=math.atan((m1 - m3)/(1.+m1*m3))
 alpha2=math.atan((m2 - m3)/(1.+m2*m3))
 
-
+print star
 print(alpha1,alpha2)
 print(180.*alpha1/math.pi,180.*alpha2/math.pi)
 

### 2018-07-10 VFTS682_APJL.tex
fixed typo in table
fixed tables
updated figures

@@
     {
+      The quoted errors are statistical, and do not include systematic
+      effects in the modeling.\\
       (1)~\cite{schneider:18}
@@
       \hline
-         &$\mathrm{RA}_\mathrm{VFTS682}$&[degrees] & \phantom{-}84.7314 %$ \pm 0.036\,\mathrm{mas} $ 
+         &$\mathrm{RA}_\mathrm{VFTS682}$&[degrees] & \phantom{-0}84.7314 %$ \pm 0.036\,\mathrm{mas} $ 
                      & (1) \\        
-               &$\mathrm{DEC}_\mathrm{VFTS682}$&[degrees] & -69.0741%$ \pm 0.048\,\mathrm{mas} $
+               &$\mathrm{DEC}_\mathrm{VFTS682}$&[degrees] & \phantom{0}-69.0741%$ \pm 0.048\,\mathrm{mas} $
                      & (1)  \\
 
-                        &$\mathrm{RA}_\mathrm{R136}$&[degrees] & \phantom{-}84.6767
+                        &$\mathrm{RA}_\mathrm{R136}$&[degrees] & \phantom{-0}84.6767
                      &  SIMBAD  \\        
-               &$     \mathrm{DEC}_\mathrm{R136}$&[degrees] &  -69.1009
+               &$     \mathrm{DEC}_\mathrm{R136}$&[degrees] &  \phantom{0}-69.1009
                      &  SIMBAD \\       
@@
                      &  (2, 5) \\  
-                       &$  \mathrm{d_\mathrm{2d}}$  &[pc] & 29
+                       &$  d_\parallel$  & [arcsec]
+                         & 119.4
                                  &  (2) \\
+      &$d_\parallel$ & [pc] & 29 & (2) \\
 
@@
                      &  (1) \\        
-                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &[$\masyr$] & $0.026$
+                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &[$\masyr$] & $0.0226$
                         & (1)  \\         
@@
       &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
-      &$\theta$  &[degrees] &  $14_{-43}^{+44}$  & (1,5)\\  
+      &$\theta$  &[degrees] &  $14_{-31}^{+36}$  & (1,5)\\  
       % &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
@@
   \centering
-  \includegraphics[width=0.48\textwidth]{./figures/Figure2_paper_vfts682_hst}  
+  \includegraphics[width=0.48\textwidth]{./figures/Figure2_paper_vfts682_hst_corr_nor136err}  
   \caption{The thick yellow arrows indicate proper motions relative R136 (blue

### 2018-07-10 scripts/calc_angle_pm.py

@@ -29,23 +29,22 @@ from pyspark.sql.functions import floor
 # pmdec_error = 0.04536
 # pmra_pmdec_corr = 0.10909
 
-star='VFTS72'
+# star='VFTS72'
+# relpmra = -0.372
+# relpmdec = 0.125
 
-relpmra = -0.372
-relpmdec = 0.125
+# pmra_error = 0.05038
+# pmdec_error = 0.06077
+# pmra_pmdec_corr = 0.1414
 
-pmra_error = 0.05038
-pmdec_error = 0.06077
-pmra_pmdec_corr = 0.1414
+star='VFTS682'
+relpmra =  0.102976611081
+pmra_error = 0.06974150031513375
 
-# star='VFTS682'
-# relpmra =  0.102976611081
-# pmra_error = 0.0797415003151
+relpmdec =  0.0855134899438
+pmdec_error = 0.0797415003151
 
-# relpmdec =  0.0855134899438
-# pmdec_error = 0.0999780525252
-
-# pmra_pmdec_corr = 0.022639342
+pmra_pmdec_corr = 0.022639342
 
 # components of the correlation matrix
 c33=pmra_error * pmra_error

### 2018-07-10 VFTS682_APJL.tex
updated text

@@
                          & 119.4
-                                 &  (2) \\
-      &$d_\parallel$ & [pc] & 29 & (2) \\
+                                 &  SIMBAD \\
+      &$L_\parallel$ & [pc] & 29 & (2) \\
 
@@
     \hline
-       \multicolumn{5}{l}{\emph{Expected proper motion if ejected from R136 at zero age}} \\
+       \multicolumn{5}{l}{\emph{Expected proper motion if ejected from
+      R136 at age zero}} \\
       \hline
@@
     \tablecomments
-    { Proper motions are based on Gaia data unless indicated otherwise in subscript.
-The conversion factor assuming a distance of 50\,kpc is such that $1\masyr$ corresponds to $237\kms$.
-The correlation coefficient is indicated as $\rho$.
-The angle $\theta$ is defined such that $\theta=0$ for motions pointing away radially from R136.
+    { Proper motions are based on Gaia data unless indicated otherwise
+      in subscript.    The conversion factor assuming a distance of
+      50\,kpc is such that $1\masyr$ corresponds to $237\kms$.  The
+      correlation coefficient is indicated as $\rho$. The angle
+      $\theta$ is defined such that $\theta=0$ for motions pointing
+      away radially from R136, we neglect the errors on
+      $\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$ and $\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$ to determine the uncertainty
+      on $\theta$.
       (1) \cite{brown:18},
@@
     direction are proportional to the apparent ages of the stars, and
-    the thin yellow lines illustrate the error cone on the potential
+    the thin lines illustrate the error cone on the potential
     origin. The green thick arrow

### 2018-07-11 VFTS682_APJL.tex
minor and typos

@@
 components of less than $0.1\,\mathrm{mas\ yr^{-1}}$ (dark blue lines,
-including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light shaded
+including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light hatched
 blue). The green vertical bands show the expectations for VFTS682 if
-it were ejected from R136. We highlight the star together with the two O-type
+it were ejected from R136. We highlight this star together with the two O-type
 runaways \mbox{(re-)identified} by \cite{lennon:18} (VFTS16 and VFTS72,
@@
 falls in the range expected for it to reach its present day isolation
-within its apparent age (green vertical band), hinting at it being a ``slow-runaway'' as
+within its apparent age (green vertical band). This suggests it is indeed a ``slow-runaway'' as
 suggested by \cite{bestenlehner:11}. Subtracting
@@
 where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
-the core of the cluster \citep[corresponding to $\sim$29\,pc at LMC distance,][]{bestenlehner:11}.
-As in the rest of this study, we neglect for
-simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
+the core of the cluster \citep[corresponding to $\sim$29\,pc at a
+distance of 
+50\,kpc,][]{bestenlehner:11}.
+% As in the rest of this study, we neglect for
+% simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
 The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
@@
 pair-instability supernova \cite[e.g.,][]{rakavy:67} at LMC
-metallicity \citep[see also][]{langer:07}. Similarly, the final fate of VFTS682 could be either a
+metallicity \citep[see also][]{langer:07}.
+
+Similarly, the final fate of VFTS682 could be either a
 pair-instability supernova without compact remnant formation, or

### 2018-07-13 VFTS682_APJL.tex
typos fixing, and re-arranging of sec.2 where a merger went wrong. Some text cut.

@@
   N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~Schneider$^{13}$, J.~S.~Vink$^{14}$}
-\affil{ $^{*}$ Corresponding author: \href{mailto:M.Renzo@uva.nl}{M.Renzo@UvA.nl}.  A list of the affiliations can be found at the end of this paper.}
+\affil{ $^{*}$ Corresponding author: \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}.  A list of the affiliations can be found at the end of this paper.}
 
@@
 
- How very massive stars form is still an intriguing open question in stellar astrophysics.
-VFTS682 is among the most massive stars known, with an inferred initial mass of about  $150\,M_\odot$.
-It is located in 30 Doradus at a projected distance of 29\,pc from the central cluster R136.
-The absence of other massive stars in its immediate vicinity led to two intriguing hypotheses posed in earlier work: either (a) it formed in relative isolation through a new mode of star formation or (b) it was ejected dynamically from the central cluster.
+How very massive stars form is still an intriguing open question in
+stellar astrophysics.  VFTS682 is among the most massive stars known,
+with an inferred initial mass of $\sim$$150\,M_\odot$. It is located
+in 30 Doradus at a projected distance of 29\,pc from the central
+cluster R136. % Its apparent isolation
+The absence of other massive stars in its immediate vicinity
+led to two intriguing hypotheses: % posed in earlier work:
+either (a) it formed in relative isolation % through a new mode of star formation
+or (b) it was ejected dynamically from the central cluster.
 
-Aiming to shine light on this debate, we investigate the kinematics of
+% Aiming to shine light on this debate,
+In this context, 
+we investigate the kinematics of
 VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space
@@
 circumstantial clues. The central cluster is known to harbor several
-other stars more massive than $150\,M_\odot$ similar in spectral type.
+other stars more massive than $150\,M_\odot$ of similar in spectral type.
 The very massive O stars VFTS16 and VFTS72, which have recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
@@
 %
-Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
+Obtaining clues from observations is challenging, because massive stars are intrinsically rare, 
 % \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
 evolve fast, typically reside in dense groups, and remain enshrouded in
-their parent cloud during the entirety of their formation
+their parent cloud during the entirety of the formation
 process. Important progress has been made on the theoretical side,
-\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally very expensive and therefore remain challenging.
+\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally expensive and remain challenging.
 
@@
 The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
-Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{zwicky:57, blaauw:61, eldridge:11, renzo:18}.
+Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{blaauw:61, renzo:18}.
 
@@
 It is a hydrogen-rich Wolf-Rayet star of spectral type WNh5.
-Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $\sim$$140^{+30}_{-16}\,M_\odot$ corresponding to an initial mass of $\sim$$150^{+30}_{-17}\,M_\odot$
+Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $\sim$$137.8^{+27.5}_{-15.9}\,M_\odot$ corresponding to an initial mass of $\sim$$150.0^{+28.7}_{-17.4}\,M_\odot$
 \citep{schneider:18}. 
@@
 spectrum of VFTS682 and R136a3 \citep{rubio-diez:17} for which \citet{crowther:16} report a
-current mass estimate of $180^{+30}_{-30}\Msun$. R136 hosts
-at least two more very massive WN5h stars, R136a1 and R136a2, whose
-estimated current masses are even higher.
+current mass estimate of $180^{+30}_{-30}\Msun$. % R136 hosts
+% at least two more very massive WN5h stars, R136a1 and R136a2, whose
+% estimated current masses are even higher.
 
@@
 \cite[DR2,][]{gaia:16,brown:18} and show that the proper motion,
-postion and direction of the $\sim100\Msun$ star VFTS16 is consistent
+postion and direction of the $\sim$$100\Msun$ star VFTS16 is consistent
 with a runaway origin from R136. They also found a less clear case for VFTS72.
@@
       \hline
-
+      \label{tab:star_param}
     \end{tabular}
@@
   \end{center}
-  \label{tab:star_param}
 \end{table}
@@
 using the BONNSAI code \citep{schneider:17, schneider:18} provide
-estimates for the age, present mass and initial mass, indicating that
-this star among the most massive stars known, see
+estimates for the age, present mass and initial mass, % indicating that
+% this star among the most massive stars known,
+% see
 \Tabref{tab:star_param}. % for an overview of the parameters.
@@
       &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
-      &$\theta$  &[degrees] &  $14_{-31}^{+36}$  & (1,5)\\  
+      &$\theta_{Gaia}$  &[degrees] &  $14_{-31}^{+36}$  & (1,5)\\  
       % &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
@@
                   &$v_\mathrm{2d, HST}$  &[$\kms$] & $47\pm24$ & (4)\\  
-                      &$\theta^{*}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,5)\\  
+                      &$\theta_\mathrm{HST}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,5)\\  
                   %        &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
@@
       $\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$ and $\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$ to determine the uncertainty
-      on $\theta$.
+      on $\theta_{Gaia}$.
       (1) \cite{brown:18},
@@
 
-The WNh5 star VFTS682, located at right ascension (RA)
-05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}.
-
 \cite{townsley:06} and \textcolor{blue}{Townsley et al., in prep.} did not find
@@
 colliding winds that would be expected in the presence of companions
-even for extreme mass ratios, given the extremely large mass of VFTS682.
-This star is also relatively isolated in \emph{Spitzer} images, except for the extended
-source S10 interpreted as star formation triggered by the wind of
-VFTS682 itself by \cite{walborn:13}.
+even for extreme mass ratios, given the extremely large mass of
+VFTS682.
+This star is also relatively isolated in (near) infrared.
+There nearest bright (near-)infrared sources detected by \emph{Spitzer} \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e.
+about 2.4 pc.
+\cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but we chance alignment cannot be excluded.
 
-\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
-parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
-luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
-the brightest stars in the region. 
-%
-The star does not show clear radial velocity variations, which makes the presence of a close companion unlikely
- \citet{bestenlehner:11}. 
-% 
-%\emph{Chandra} survey of the region, which suggests the absence of colliding winds that would be expected in the presence of companions even for extreme mass ratios, given the extremely large mass of VFTS682.
-- Selma: that depends on the separation.
-I don't know what the predictions are for very unequaly winds and the LMC is far away.
-I shortened it a bit and moved this up to merge it where we were discussing the presence of a companion already, to use it in the flow of the argument, instead of disrupting the flow.
-%
- Bayesian fits of the stellar parameters derived from the spectra against single star evolutionary tracks \citep{brott:11, kohler:15}
-using the Bonsaii code \citep{schneider:17, schneider:18} provide
-estimates for the age, present mass and initial mass, indicating that
-this star among the most massive stars known, see
-\Tabref{tab:star_param} for an overview of the parameters.
+The V-band light curve of VFTS682  shows
+variations at a $\sim$10\% level on a timescale of years, which is
+unusual for Wolf-Rayet stars and more typical for Luminous Blue
+Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
+shows mid-infrared excess \citep{gruendl:09}.
 
@@
 $10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the
-possible effect of clumping.  The V-band light curve of VFTS682  shows
-variations at a $\sim$10\% level on a timescale of years, which is
-unusual for Wolf-Rayet stars and more typical for Luminous Blue
-Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
-shows mid-infrared excess \citep{gruendl:09}.
-We therefore caution for over-interpreting the estimates for the radial velocities, but we will summarize the estimates that have been published.
-\citet{bestenlehner:11}  estimate a radial velocity (RV) of  $300\pm10\kms$ using the  N{\footnotesize V} $\lambda4944$ line, which is offset from the average radial velocity of the region, which is  $270\pm10\kms$.
+possible effect of clumping.  We therefore caution
+for over-interpreting the estimates for the radial velocities. % , but we will summarize the estimates that have been published. 
+\citet{bestenlehner:11}  estimate a radial velocity (RV) of
+$300\pm10\kms$ using the  N{\footnotesize V} $\lambda4944$ line, which
+is offset from the average radial velocity of the region of $270\pm10\kms$.
 This is consistent with, but no proof of, a runaway nature.
@@
 
-VFTS682 also appears to be relatively isolated in (near) infrared.
-There nearest bright (near-)infrared sources detected by \emph{Spitzer} \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e.
-about 2.4 pc.
-\cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but we chance alignment cannot be excluded.
-
-%% Radial Velocity:
-
 We adopt as distance to the LMC the value of $50$\,kpc.
-The error on the distance determination is small ($\lesssim2\%$, \citealt{pietrzynski:13}) and any possible offset between  R136 or VFTS682  and the distance we adopted for the LMC is probably much smaller (likely $\ll 1\%$, e.g.
-\citealt{Luks+1992}).
+The error on the distance determination is small ($\lesssim2\%$, \citealt{pietrzynski:13}) and any possible offset between  R136 or VFTS682  and the distance we adopted for the LMC is probably much smaller (likely $\ll 1\%$, e.g., \citealt{Luks+1992}).
 These uncertainties are negligible compared to the errors in the proper motion discussed below.
@@
 
+\vspace*{10pt}% to move new sec. to newpage
+
 \subsection{ \emph{Gaia} astrometry for VFTS682\label{data:gaia}}
 
-\begin{figure*}[htbp]
+\begin{figure*}[tbp]
   \centering
@@
 
-Gaia provides absolute proper motions.
-To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.
-They select  bright ($G<17$) stars within 0.05 degrees of R136 and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1 for further discussion).
+Gaia provides absolute proper motions.  To determine the relative
+proper motion with respect to R136, we follow  \citet{lennon:18} to
+define the motion of the local frame of reference using the average
+proper motion of nearby stars with reliable astrometric data.  They
+selected bright ($G<17$) stars within 0.05 degrees of R136 and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1).
 
@@
 i.e. such that $\theta = 0$ is the angle corresponding to perfectly
-radial motion away from R136.    We further provide the projected 2d velocity
-$v_\mathrm{2d}$.   All kinematic quantities are provided in
-Tab.~\ref{tab:vfts682}.
+radial motion away from R136. All kinematic quantities are provided in
+\Tabref{tab:vfts682}.
 
@@
 The 30 Doradus region was target of an a two-epoch photometric campaign with HST providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
-Visually, the star is clearly isolated and does not appear to have any close neighbor.
+% Visually, the star is clearly isolated and does not appear to have any close neighbor.
 This by itself gives further confidence in ability of \emph{Gaia} to measure the proper motion reliably without being affected by crowding.
@@
 brightest stars (V$<$14) are saturated in the data set and have been
-excluded from the analysis.  VFTS682 should be one of the brightest
-stars in the region, however, the high extinction makes it faint enough such that saturation effects are not an issue.
+excluded from the analysis. The high extinction around VFTS682
+prevents this issue. % should be one of the brightest
+% stars in the region, however, the high extinction makes it faint enough such that saturation effects are not an issue.
 VFTS682 is relatively slow and is therefore not discussed in their paper.
@@
 
-The HST provides proper motions that are relative to the bulk motion of the majority of the stars in the field of view.
-The full 30 Dor field is covered by different pointings and there is some systematic distortion.
-However, even for stars far from 30 Dor the effect is small, no more than $0.05\masyr$ across the whole 30 Dor field.
-The effect is much smaller for stars close to the center of field, such as VFTS 682 \citet{platais:18}.
-We can therefore use the relative proper motion as a good estimate for the proper motion relative to R136.
+The HST provides proper motions that are relative to the bulk motion
+of the majority of the stars in the field of view. The full 30 Dor
+field is covered by different pointings and there is some systematic
+distortion.  However, even for stars far from 30 Dor the effect is
+small, no more than $0.05\masyr$ across the whole 30 Dor field. The
+effect is much smaller for stars close to the center of field, such as
+VFTS 682 \citep{platais:18}.  We can therefore use the relative proper
+motion (\Tabref{tab:vfts682}) as a good estimate for the proper motion relative to R136.
 
-We list the relative proper motion in \Tabref{tab:vfts682}.
+% We list the relative proper motion in \Tabref{tab:vfts682}.
 
@@
 blue). The green vertical bands show the expectations for VFTS682 if
-it were ejected from R136. We highlight this star together with the two O-type
-runaways \mbox{(re-)identified} by \cite{lennon:18} (VFTS16 and VFTS72,
-respectively).
+it were ejected from R136 early in its lifetime. We highlight this star together with the two O-type
+runaways, VFTS16 and VFTS72, \mbox{(re-)identified} by \cite{lennon:18}.
 
-The left panel shows the distribution in two-dimensional proper motion relative
-to R136. Although VFTS682 is not peculiar, its relative proper motion
+The left panel shows the distribution in proper motion relative
+to R136 from \emph{Gaia} DR2, selecting stars with small error bars only. Although VFTS682 is not peculiar, its relative proper motion
 falls in the range expected for it to reach its present day isolation
@@
 model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS682 and the massive members of R136.%  While it is not possible to
+VFTS682 and the massive members of R136. %  While it is not possible to
 % robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
@@
 % merger.
-
 To eject such a massive object, the cluster is
@@
 early in the evolution of a cluster, dynamical interactions form an extremely
-massive binary, which then tightens its orbit by ejecting other stars passing
-by. Interpreting our results for VFTS682 through the lens of their simulations
+massive binary, which then tightens its orbit by ejecting other
+stars. Interpreting the kinematics of VFTS682 through the lens of their simulations
 suggests the presence of a close binary with total mass

### 2018-07-22 VFTS682_APJL.tex
added ignore comments folder
included Steve Shore's comments

@@
 VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space
-  Telescope} photometry. We derive a projected velocity relative of to
-the cluster of $\sim$$30\pm20\kms$.  The direction and magnitude of
-the proper motion are consistent with what is expected if VFTS682 was ejected as a slow runaway from the central cluster, given its distance and the inferred age.
+  Telescope} photometry. We derive a projected velocity relative to
+the cluster of $\sim$$30\pm20\kms$ at one $\sigma$ level.  The direction and magnitude of
+the proper motion are consistent with VFTS682 being a slow runaway
+ejected from the central cluster, given its distance and the inferred age.
 However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.
@@
 luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
-the brightest stars in the region. The star is unlikely to have a
-close companion, because of the absence of clear radial velocity
-variations \citet{bestenlehner:11}. Bayesian fits of the stellar
+the brightest stars in the region. The absence of radial velocity
+variations suggests that the star is unlikely to have close binary
+companions \citet{bestenlehner:11}, unless the orbital inclination is
+very high. Bayesian fits of the stellar
 parameters against evolutionary tracks \citep{brott:11, kohler:15}
@@
 
-\cite{townsley:06} and \textcolor{blue}{Townsley et al., in prep.} did not find
+\cite{townsley:06} did not find
 an X-ray point source at the position of VFTS682 in their
@@
 
-We adopt as distance to the LMC the value of $50$\,kpc.
-The error on the distance determination is small ($\lesssim2\%$, \citealt{pietrzynski:13}) and any possible offset between  R136 or VFTS682  and the distance we adopted for the LMC is probably much smaller (likely $\ll 1\%$, e.g., \citealt{Luks+1992}).
+We adopt as distance to the LMC the value of $50$\,kpc. The error on
+the distance determination is small ($\lesssim2\%$,
+\citealt{pietrzynski:13}) and any possible offset in the radial direction between R136 or VFTS682 and the distance we adopted for the LMC is probably much smaller (likely $\ll 1\%$, e.g., \citealt{Luks+1992}).
 These uncertainties are negligible compared to the errors in the proper motion discussed below.
@@
 \vspace*{10pt}% to move new sec. to newpage
-
-\subsection{ \emph{Gaia} astrometry for VFTS682\label{data:gaia}}
-
 \begin{figure*}[tbp]
@@
 \end{figure*}
+
+\subsection{ \emph{Gaia} astrometry for VFTS682\label{data:gaia}}
 
@@
 % simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
-The kinematic age $\tau_\mathrm{kin}$ is compatible with a very early
+The kinematic age $\tau_\mathrm{kin}$ is consistent with a very early
 ejection from the cluster, given the apparent age of the star.
@@
 R136 early in the evolution of both the star and the cluster.
-
-% \subsection{Is it a runaway star?}
-% \label{sec:runaway}
-% We first address the question whether VFTS682 is a typical star
-% from the kinematic point of view (which is what would be expected if it formed in relative isolation), or whether has a significant peculiar velocity compared to its surrounding population (which is what we would expect if the star has a runaway origin).
-
-% After subtracting the mean proper motion of the field, we obtain the relative two-dimensional relative proper motion.
-We obtain  $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$   and $\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$.
-
-% Both indicate a non-zero peculiar 2D motion that would make it a slow runaway star, but in both cases the meaurements are also consistent with zero within 2 sigma.
-
-% Figure~\ref{fig:mu_dist} shows the distribution of relative proper motions for other VFTS O, B and WR stars subselecting either those with relative proper motion errors smaller than $0.1\masyr$ (yielding 317 stars) or a more restrictive cut including only those with proper motion errors smaller than $0.05\masyr$ (yielding 36) stars.
-We see that the derived Gaia proper motion is very typical for stars in the field.
-
-% \SdM{XXX We need to rethink }
-
-%Therefore, two completely independent measures of the proper motion of VFTS682 relative to the surrounding field, one from HST and one from \emph{Gaia} DR2, yield values of the peculiar three-dimensional velocity of VFTS682 % (\Eqref{eq:speed_around_HST} and
-% \Eqref{eq:speed_around_Gaia}, respectively)
-%which would make it the most massive runaway star known to date. However, the large errors on both the proper motion measures require confirmation with future astrometric data.
-
-%Subtracting the mean proper motion components given by \citet{lennon:18} from the proper motion of VFTS682 (see \Tabref{tab:vfts682}), we obtain the components of proper motion of the star relative to the surrounding region $\delta\mu_\mathrm{RA}^{Gaia} = 0.10 \pm 0.07\masyr$ and $\delta\mu_\mathrm{DEC}^{Gaia} = 0.08 \pm 0.09\masyr$.
-These components result in a two-dimensional relative proper motion of $\delta \mu^{Gaia}=0.13\pm 0.09\,\mathrm{mas\ yr^{-1}}$.
-
-% \begin{equation}
-%   \label{eq:pm_gaia_around}
-%   \delta \mu^{Gaia} = \sqrt{\left(\delta\mu_\mathrm{RA}^{Gaia}\right)^2+\left(\delta\mu_\mathrm{DEC}^{Gaia}\right)^2}
-%   = 0.13\pm 0.09\,\mathrm{mas\
-%   yr^{-1}} \ \ .
-% \end{equation}
-
-% Figure \Figref{fig:pm_polar} shows the distribution in relative proper motion
-% of the stars used in \cite{lennon:18} to define the frame of reference
-% we also adopt here, cf.~the right panel in their Fig.~2. The
-% proper motions are decomposed in the radial and tangential direction
-% from R136, to highlight the likelihood of it being the origin of fast
-% moving stars. The plus sign marks VFTS682, which sits at the edge of
-% the proper motion distribution, but is not a clear outlier. Most
-% relevant is the direction of the proper motion of VFTS682, which we
-% discuss in \Secref{sec:r136_origin}.
-
-%The proper motion components can be converted into the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=24\pm19\,\kms$, $\delta v_\mathrm{DEC}^{Gaia}=20\pm23\,\kms$, assuming a distance of 50\,kpc.
-These can be combined obtaining a projected two-dimensional velocity of $31\pm21\kms$.
-
-%The radial velocity from \cite{bestenlehner:11} then gives the third component along the line of sight,  which added in quadrature to the transverse components results in a three-dimensional velocity of $44 \pm 21\kms$.
-% \begin{equation}
-%   \label{eq:speed_around_Gaia}
-%   v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
-%     +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
-%       v_\mathrm{rad}^{VFTS}\right)^2} = 44 \pm 21
-%   \kms \ .
-% \end{equation}
-
-% \subsection{Does it come from the R136 cluster?}
-% \label{sec:r136_origin}
-
-% Assuming that the best estimate of the peculiar velocity of VFTS682
-% are reliable, we now address the question of its likely origin.
-
-% Figure~\ref{fig:mu_dist} shows the distribution in proper motion relative to R136
-% of OB-type and Wolf-Rayet stars included both in the VFTS survey and
-% \emph{Gaia DR2} with reliable astrometric solutions. The bottom panel
-% shows the distribution in angles $\theta$ between the relative proper motion
-% direction and the direction from the core of R136 to the star. We
-% emphasize that our subset of stars with reliable proper motion
-% measurements is biased towards the fast moving objects, which results in a
-% distribution of angles mildly peaked at small angles.
-
-% VFTS682 is not an outlier in relative proper motion%  compared
-% % to the two O-type runaways (re-)identified by \cite{lennon:18} (see
-% % also below)
-% . However, the star was expected to be a ``slow runaway'' by \citet{bestenlehner:11} in
-% the dynamical ejection scenario. The green shade in
-% \Figref{fig:mu_dist} shows the range of relative proper motions
-% required to reach the present day location within
-% the uncertainties on the apparent age of the star, assuming it was
-% ejected from R136 very early in its life.
-
-% Because of the large
-% error bars, the angle between the relative proper motion is not very
-% constraining, but it is suggestive that the best value is close to zero.
-% Therefore, the relative proper motion from \emph{Gaia} DR2 are consistent with the hypothesis
-% of dynamical ejection.
-
-% \todo{discuss and find a place for Fig.2}
-% The red arrow in \Figref{fig:main} shows the direction of relative proper motion of
-% VFTS682 from \emph{Gaia} DR2, and the
-% yellow arrows illustrate the uncertainty. % show the possible
-% % range of directions within the uncertainties in the measured relative proper
-% % motion.
-% These arrows cross at the present-day location of VFTS682 and
-% are prolonged in the direction opposite to the motion to illustrate
-% the possible range of origins.
-% The yellow arrow shows the direction of
-% the relative proper motion from HST (see \Eqref{eq:pm_around_HST}),
-% but we omit to show the possible range of directions from HST data, since it is
-% insufficiently constrained to be useful.
-
-% Although the large uncertainties on the relative proper motion
-% components results in a wide range of possible directions,we argue that the most likely origin of the star is R136.
-
- % , which
-% corroborates the idea that the star is the result of a dynamical
-% ejection very soon in the cluster evolution.
-
-% the thing below is not relevant if it is a merger, better cut out
-% for now, since we are too long anyway
-
-%Note that the
-% present-day surface helium abundance
-% \citep[$Y\simeq0.5$,][]{bestenlehner:11, rubio-diez:17} puts a lower
-% limit on the age of the star of $\sim$0.9\,Myr, corresponding to the time needed to
-% synthesize this amount of helium in the models from \cite{kohler:15}.
 
@@
 
-\citet{lennon:18} carried out a study similar to ours on the fast
-moving O-type stars
-in the region, and found two very massive runaway stars in the 30 Doradus region. One of them (VFTS 16)
+In \citet{lennon:18} we carried out a similar study on the fast
+moving O-type stars in the region, and found two very massive runaway stars in the 30 Doradus region. One of them (VFTS 16)
 was previously known as a runaway star from its line of sight velocity
@@
 the result of a dynamical ejection from the R136 cluster, while the
-origin of the other star (VFTS 72) is less clear given its direction
+origin of the other star (VFTS72) is less clear given its direction
 of motion (cf.~\Figref{fig:dist}). The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
 VFTS682 is smaller than the
-corresponding value for VFTS16: \cite{lennon:18} inferred a kinematic
-age of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
+corresponding value for VFTS16 of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
 massive VFTS682 was ejected later than VFTS16 from the same cluster.

### 2018-07-22 VFTS682_APJL.tex
reordering

@@
 
-Based on our results, we tentatively claim that VFTS682 is the most massive
+Based on our results, we suggest that VFTS682 is the most massive
 (slow) runaway known to date, with a peculiar two-dimensional
@@
 % present in the crowded core of R136.
+
+The variability of VFTS682, reminiscent of LBV stars, suggests
+that VFTS682 (and therefore its analogs in the core of R136) might
+experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
+debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
+claim that LBV stars are typically isolated from O-type stars. The fact that VFTS682 is a dynamically
+ejected runaway which might evolve into an LBV star suggests that
+N-body interactions also play a role in explaining the apparent
+isolation of at least some LBV stars.
 
@@
 to have experienced core-collapse already.
-
-The variability of VFTS682, reminiscent of LBV stars, suggests
-that VFTS682 (and therefore its analogs in the core of R136) might
-experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
-debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
-claim that LBV stars are typically isolated from O-type stars. The fact that VFTS682 is a dynamically
-ejected runaway which might evolve into an LBV star suggests that
-N-body interactions also play a role in explaining the apparent
-isolation of at least some LBV stars.
 

### 2018-07-26 VFTS682_APJL.tex
added ref. to drew+18
started working on Selma's comments

@@
 
-Figure~\ref{fig:dist} illustrates the kinematic properties of VFTS OB-type
-and Wolf-Rayet stars in the 30 Doradus region with \emph{Gaia} DR2 errors on the proper motion
+The black points in~\Figref{fig:dist} show the proper motion 
+(left panel) and the projected flight direction (right panel) relative
+to R136 of
+VFTS682. $\theta$ is the angle between the projected radial direction from the
+center of R136 to each star and its projected relative proper motion
+(see also \Figref{fig:main}). Dynamical ejections from the
+cluster should %are expected to
+produce
+close to radial ejections, i.e.~$\theta\simeq0$. The green vertical bands highlight the expectations for these
+two quantities if VFTS682 were ejected from R136 early in its
+lifetime. Although the error bars are substantial and VFTS682 is not
+an outlier, the agreement
+suggests that the star is indeed a ``slow-runaway'' as suggested by
+\cite{bestenlehner:11}. For comparison, we also show the relative
+proper motion of VFTS16 and VFTS72 (gray points), and the distribution
+in relative proper motion and flight direction for all the VFTS O/B-type and Wolf-Rayet stars with
+\emph{Gaia} DR2 errors on the proper motion
 components of less than $0.1\,\mathrm{mas\ yr^{-1}}$ (dark blue lines,
 including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light hatched
-blue). The green vertical bands show the expectations for VFTS682 if
-it were ejected from R136 early in its lifetime. We highlight this star together with the two O-type
-runaways, VFTS16 and VFTS72, \mbox{(re-)identified} by \cite{lennon:18}.
+blue).
 
-The left panel shows the distribution in proper motion relative
-to R136 from \emph{Gaia} DR2, selecting stars with small error bars only. Although VFTS682 is not peculiar, its relative proper motion
-falls in the range expected for it to reach its present day isolation
-within its apparent age (green vertical band). This suggests it is indeed a ``slow-runaway'' as
-suggested by \cite{bestenlehner:11}. Subtracting
+% Figure~\ref{fig:dist} illustrates the kinematic properties of VFTS OB-type
+% and Wolf-Rayet stars in the 30 Doradus region with \emph{Gaia} DR2 errors on the proper motion
+% components of less than $0.1\,\mathrm{mas\ yr^{-1}}$ (dark blue lines,
+% including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light hatched
+% blue). The green vertical bands show the expectations for VFTS682 if
+% it were ejected from R136 early in its lifetime. We highlight this star together with the two O-type
+% runaways, VFTS16 and VFTS72, \mbox{(re-)identified} by \cite{lennon:18}.
+
+% The left panel shows the distribution in proper motion relative
+% to R136 from \emph{Gaia} DR2, selecting stars with small error bars only.
+% Although VFTS682 is not peculiar, its relative proper motion
+% falls in the range expected for it to reach its present day isolation
+% within its apparent age (green vertical band). This suggests it is indeed a ``slow-runaway'' as
+% suggested by \cite{bestenlehner:11}.
+Subtracting
 the mean motion of R136, we obtain relative proper motions (projected
@@
 
-The right panel shows the distributions in projected flight direction for these
-stars: $\theta$ is the angle between the projected radial direction from the
-center of R136 to each star and its projected relative proper motion (see also \Figref{fig:main}). Dynamical ejections from the
-cluster should %are expected to
-produce
-close to radial ejections, i.e.~$\theta\simeq0$. On this panel, the green band highlights
-% the range of angles corresponding to
-an opening angle around the
-projected radial direction of 45\,degrees. The direction of motion of VFTS682 % is
-% poorly constrained because of the large errors on the proper motion
-% components both in HST and \emph{Gaia} data, but
-% it
-is consistent with dynamical ejection from R136. We note that
+% The right panel shows the distributions in projected flight direction for these
+% stars: $\theta$ is the angle between the projected radial direction from the
+% center of R136 to each star and its projected relative proper motion (see also \Figref{fig:main}). Dynamical ejections from the
+% cluster should %are expected to
+% produce
+% close to radial ejections, i.e.~$\theta\simeq0$. On this panel, the green band highlights
+% % the range of angles corresponding to
+% an opening angle around the
+% projected radial direction of 45\,degrees. The direction of motion of VFTS682 % is
+% % poorly constrained because of the large errors on the proper motion
+% % components both in HST and \emph{Gaia} data, but
+% % it
+% is consistent with dynamical ejection from R136.
+We note that
 VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
@@
 
-Therefore, although the motion of VFTS682 is not an outlier compared
-to other massive stars with small proper motion errors in \emph{Gaia} DR2, it is possible to tentatively claim that it is
+% Therefore, although the motion of VFTS682 is not an outlier compared
+% to other massive stars with small proper motion errors in \emph{Gaia} DR2,
+% it is possible to tentatively claim that
+\todo{from here}
+it is
 a slow runaway with velocity compatible with dynamical ejection from
@@
 
-\begin{figure}%[tbp]
+\begin{figure*}%[tbp]
   \centering
-  \includegraphics[width=0.48\textwidth]{./figures/Figure2_paper_vfts682_hst_corr_nor136err}  
+  \includegraphics[width=\textwidth]{./figures/Figure2_paper_vfts682_hst_corr_nor136err}  
   \caption{The thick yellow arrows indicate proper motions relative R136 (blue
@@
     error cone and prolongation proportional to its apparent
-    age.}
+    age. The two blue circles indicate the regions of radii 0.01 and 0.05
+    degrees around the core of R136.}
 
   \label{fig:main}
-\end{figure}
+\end{figure*}
 
@@
 corresponding value for VFTS16 of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
-massive VFTS682 was ejected later than VFTS16 from the same cluster.
+massive VFTS682 was ejected later than VFTS16 from the same
+cluster. We note that also Galactic clusters can eject massive stars,
+and \emph{Gaia} data have been used to study the population
+surrounding Westerlund 2 \citep[][]{drew:18}.
 

### 2018-08-21 VFTS682_APJL.tex
addresses Jorik's comments

@@
 %\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
-\title{Gaia's constraints on the very massive $\sim$$150\,M_\odot$ runaway star candidate VFTS682}
+\title{Space astrometry of the very massive $\sim$$150\,M_\odot$ runaway star candidate VFTS682}
 
@@
  In particular, a remarkable similarity exist between the
-spectrum of VFTS682 and R136a3 \citep{rubio-diez:17} for which \citet{crowther:16} report a
-current mass estimate of $180^{+30}_{-30}\Msun$. % R136 hosts
+spectrum of VFTS682 and R136a3 \citep{rubio-diez:17}.%  for which \citet{crowther:16} report a
+% current mass estimate of $180^{+30}_{-30}\Msun$.
+% R136 hosts
 % at least two more very massive WN5h stars, R136a1 and R136a2, whose
@@
 postion and direction of the $\sim$$100\Msun$ star VFTS16 is consistent
-with a runaway origin from R136. They also found a less clear case for VFTS72.
+with a runaway origin from R136. They also found a less clear case for
+VFTS72, and in both cases some tension between the kinematic age of
+these stars and their apparent age remains.
 
@@
 about 2.4 pc.
-\cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but we chance alignment cannot be excluded.
+\cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but chance alignment cannot be excluded.
 
@@
 $300\pm10\kms$ using the  N{\footnotesize V} $\lambda4944$ line, which
-is offset from the average radial velocity of the region of $270\pm10\kms$.
-This is consistent with, but no proof of, a runaway nature.
-\cite{bressert:12} note an offset compared between the RV of the star and the nebular lines from the gas filaments in its vicinity.
-This is consistent with the expectation if the star was not formed in situ.
-Newer measurements based on new X-shooter data give different estimates for the RV ({\color{blue} Rubio-D{\' i}ez et al.
-in prep}).
+is offset from the average radial velocity of the region of
+$270\pm10\kms$. This is consistent with, but no proof of, a runaway
+nature.  \cite{bressert:12} note an offset compared between the RV of
+the star and the nebular lines from the gas filaments in its
+vicinity. This is consistent with the expectation if the star was not
+formed in situ. Newer measurements based on new X-shooter data give different estimates for the RV ({\color{blue} Rubio-D{\' i}ez et al. in prep}).
 We will therefore refrain from using the RV measurements in this work.
@@
 VFTS682 is identified with the source id 4657685637907503744 in the
-\emph{Gaia} DR2 catalog\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}
+\emph{Gaia} DR2 catalog %\footnote{\url{https://vizier.u-strasbg.fr/viz-bin/VizieR-3?-source=I/345/gaia2}}
   as a 15.65 mag star in the G band
@@
 
-% Therefore, although the motion of VFTS682 is not an outlier compared
-% to other massive stars with small proper motion errors in \emph{Gaia} DR2,
-% it is possible to tentatively claim that
-\todo{from here}
-it is
-a slow runaway with velocity compatible with dynamical ejection from
-R136 early in the evolution of both the star and the cluster.
+In summary, both \emph{Gaia} and HST relative proper motions are consistent with the dynamical ejection of
+VFTS682 from the cluster, although we cannot confidently rule out the
+counter hypothesis.
 
@@
 
-Based on our results, we suggest that VFTS682 is the most massive
-(slow) runaway known to date, with a peculiar two-dimensional
+Based on our results, we consider likely that VFTS682 is the most massive
+(slow) runaway known to date, with a two-dimensional
 projected velocity with respect to R136 of
 $\sim38\kms$ (averaging the \emph{Gaia} DR2 and HST
-results) but compatible with zero within $2\sigma$. Due to the large error bars, this result will need
+results), but compatible with zero within $2\sigma$. Due to the large error bars, this result will need
 to be revisited with future astrometric data. % to be revisited as updated
@@
 \emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
-$0.9\pm0.6$\,Myr ago. Because of the exceptionally large mass
-of this star, this raises the question of which stars must populate
-the core of the cluster.
+$0.9\pm0.6$\,Myr ago. If the
+cluster age \citep[$\lesssim2$\,Myr,][]{crowther:10, sabbi:12} is
+indeed smaller than the shortest stellar lifetime
+\citep[$\sim$3\,Myr,][]{brott:11,zapartas:17}, the ejection of VFTS682
+from the disruption of a binary is excluded. %  since the region is too young for stars
+% to have experienced core-collapse already. 
+The kinematic age we infer is smaller than the kinematic age of
+$\sim$1.5\,Myr for VFTS16 found in \cite{lennon:18}, which indicates
+that VFTS682 was ejected later than VFTS16, at a time similar to VFTS72.
 
-Dynamical ejections due to N-body interactions typically (although, not necessarily) eject the least
-massive star among those interacting \cite[e.g.,][]{banerjee:12}. This means that, just
+If the star were ejected dynamically, its exceptionally large mass raises the question of which stars must populate
+the core of the cluster. N-body dynamics typically ejects the least
+massive star among those interacting \cite[e.g.,][]{banerjee:12}. Just
 based on the kinematic properties of VFTS682, we would expect several
 stars with initial masses larger than $\sim$$150\,M_\odot$ in the
-cluster R136.
-This is consistent with the detection
+cluster R136, which is consistent with the detection
 of extremely massive stars in the core of the
-cluster. % The projected rotational equatorial
-% velocity\footnote{However, the determination of the rotational
-%   velocity for stars showing lines in emission is complicated by the optically thick wind screening the surface of the star, and should be
-%   considered with caution.} of VFTS682
-% reported by \cite{schneider:18} is $v\sin(i)<200\,\kms$, which is in
-% line with the average rotation rate of massive stars in the region
-% \citep[][]{ramirez-agudelo:15}. This suggests that VFTS682 (i) has not
-% experienced rotationally induced chemically homogeneous evolution
-% \citep[][]{maeder:00,demink:09}, and (ii) it has not
-% accreted mass from or merged with a binary companion, nor will it, since the multi-epoch
-% data of the VFTS survey rule out the presence of a close companion at the
-% present day. Moreover,
+cluster.
 
@@
 \citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
-three most massive stars detected in the core of the cluster%  by
+three most massive stars detected in the core of the cluster, %  by
 % \cite{dekoter:97,crowther:10,crowther:16}
-, with an astonishing similarity in particular with
-the spectrum of R136a3. Therefore, the isolation of
+ with an remarkable similarity in particular with
+the spectrum of R136a3. The isolation of
 VFTS682 makes it an ideal target to constrain the stellar physics of
 stars with masses well above $\sim$$100\,M_\odot$ while avoiding
-crowding issues. %  : its isolation makes
-% it an easier target for observations compared to the similar stars
-% present in the crowded core of R136.
+crowding issues.
 
-The variability of VFTS682, reminiscent of LBV stars, suggests
-that VFTS682 (and therefore its analogs in the core of R136) might
-experience enhanced mass loss episodes in LBV eruptions. \citet{smith:15} made the highly
-debated\footnote{See, e.g., \cite{humphreys:16, davidson:16, smith:16}.}
-claim that LBV stars are typically isolated from O-type stars. The fact that VFTS682 is a dynamically
-ejected runaway which might evolve into an LBV star suggests that
-N-body interactions also play a role in explaining the apparent
-isolation of at least some LBV stars.
-
-\citet{banerjee:12} used N-body simulations of fully segregated
-clusters with all massive stars in binaries to suggest that VFTS682
-was ejected from R136. They
-demonstrated that the cluster potential does not significantly change
-the velocity of the star after the ejection. In their
-model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS682 and the massive members of R136. %  While it is not possible to
-% robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
-% like fast rotation, which however could have already slowed down
-% because of the wind angular momentum losses during and after the
-% merger.
-To eject such a massive object, the cluster is
-expected to have produced
-a large number of massive runaways. Indeed, several %relatively
-isolated massive stars are observed in the region, some with known
-large radial velocities and/or proper motion. % (see
-% \Figref{fig:pm_polar} and Sana et al., in prep.).
-A comprehensive study of the kinematic
-properties of all the massive stars surrounding R136 might shed light
-on whether some can be unequivocally identified as merger products. It
-is also possible that the star or binary that caused the ejection of
-VFTS682 might have been ejected in the opposite direction, and is also
-isolated at present day. If the ejection was caused by an interaction
-with a binary, however, it is likely that the binary scattered in the
-opposite direction will experience further dynamical interactions on
-its way, modifying its trajectory and making it difficult to find.  %absence of proof would not undermine our conclusions
-
-The similarities between VFTS682 and the WNh5 stars in the core of
-R136 are also in agreement with the ``bully binary'' model of
-\cite{fujii:11}. Based on their numerical results, they suggested that
+The simulations of \cite{fujii:11} suggest that
 early in the evolution of a cluster, dynamical interactions form an extremely
 massive binary, which then tightens its orbit by ejecting other
-stars. Interpreting the kinematics of VFTS682 through the lens of their simulations
+stars. The spectral similarities between VFTS682 and stars in the core
+of R136 are in agreement with this ``bully binary'' model. Interpreting the kinematics of VFTS682 through the lens of their simulations
 suggests the presence of a close binary with total mass
-$M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. Such bully
-binary could be R145 according to \cite{fujii:11}, and it might be an
+$M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. The kinematic age of VFTS682 puts an
+upper limit to the timescale to form the ``bully binary'' in
+R136. Such binary could be R145 according to \cite{fujii:11}, and it might be an
 ideal observational candidate for a dynamically formed progenitor system of
@@
 
-The kinematic age of VFTS682 puts an
-upper limit to the timescale to form the ``bully binary'' in
-R136. The cluster must have been at the very beginning of its
-evolution, given the age estimate of $\lesssim 2$\,Myr
-\citep[][]{crowther:10,sabbi:12} and the kinematic age of VFTS682. If the
-cluster is indeed younger than the shortest stellar lifetime
-\citep[$\sim$3\,Myr, e.g.,][]{brott:11, zapartas:17}, then the alternative
-explanation for ejection of VFTS682 from the disruption of a binary
-by a core-collapse event is excluded since the region is too young for stars
-to have experienced core-collapse already.
-
-In \citet{lennon:18} we carried out a similar study on the fast
-moving O-type stars in the region, and found two very massive runaway stars in the 30 Doradus region. One of them (VFTS 16)
-was previously known as a runaway star from its line of sight velocity
-\citep[][]{evans:10}. \citet{lennon:18} also concluded that VFTS16 is 
-the result of a dynamical ejection from the R136 cluster, while the
-origin of the other star (VFTS72) is less clear given its direction
-of motion (cf.~\Figref{fig:dist}). The value of $\tau_\mathrm{kin}\simeq0.9$\,Myr we find for
-VFTS682 is smaller than the
-corresponding value for VFTS16 of $\sim$1.5\,Myr, possibly in tension with the apparent age of that star. This means that the more
-massive VFTS682 was ejected later than VFTS16 from the same
-cluster. We note that also Galactic clusters can eject massive stars,
-and \emph{Gaia} data have been used to study the population
-surrounding Westerlund 2 \citep[][]{drew:18}.
-
-The numerical simulations from \cite{oh:16} suggest that dynamical
-interaction eject the majority of the stars during or shortly after the cluster
-core-collapse. The large number of isolated massive stars around it
-suggest that R136 has already evolved past the
-time of maximum stellar density. This might have implications for the
-question of whether the cluster formed via a monolithic collapse, or
+Also \citet{banerjee:12} suggested, based on their N-body simulations,
+% N-body simulations % of fully segregated
+% clusters with all massive stars in binaries
+that VFTS682 was ejected from R136. In their
+model, they relied on (dynamically driven) stellar mergers to explain the high masses of
+VFTS682 and the massive members of R136. They
+demonstrated that the cluster potential does not significantly change
+the velocity of the star after the ejection.
+% While it is not possible to
+% robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
+% like fast rotation, which however could have already slowed down
+% because of the wind angular momentum losses during and after the
+% merger.
+To eject such a massive object, the cluster is
+expected to have produced
+a large number of massive runaways. Indeed, several %relatively
+isolated massive stars are observed in the region \citep[][]{evans:10,lennon:18}, some with known
+large radial velocities and/or proper motion. % (see
+% \Figref{fig:pm_polar} and Sana et al., in prep.).
+A comprehensive study of the kinematic
+properties of all the massive stars surrounding R136 could shed light
+on whether some can be unequivocally identified as merger products,
+but also on the initial conditions for the cluster dynamics
+\citep[e.g.,][]{oh:16}, and whether it formed via a monolithic collapse, or
 as a (potentially ongoing) merger of several sub-structures \citep[e.g.,][]{sabbi:12}.
-
-\citet{oh:16} also showed that the mass and velocity distribution of the ejected star depends on the cluster initial conditions
-(whether it is segregated, its primordial binary fraction and initial period
-distribution of the binary population), therefore studies on
-the population of isolated massive stars in the surroundings of R136
-might shed light on its initial stellar population and dynamical
-state.
 

### 2018-08-21 VFTS682_APJL.tex
addresses Joachim's comments

@@
   R.~P.~van~der~Marel$^{3,4}$, E.~Laplace$^{1}$,
-  J.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
+  J.~M.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
   V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{1,8,9}$,  A.~de~Koter$^{1}$,
@@
 \cite{townsley:06} did not find
-an X-ray point source at the position of VFTS682 in their
+a bright X-ray point source at the position of VFTS682 in their
 \emph{Chandra} survey of the region, which suggests the absence of

### 2018-08-21 VFTS682_APJL.tex
addresses Chris's comments

@@
 is offset from the average radial velocity of the region of
-$270\pm10\kms$. This is consistent with, but no proof of, a runaway
-nature.  \cite{bressert:12} note an offset compared between the RV of
+$270\pm10\kms$. This was suggested as indicating a runaway nature, but
+it is no proof of it, and more recent measurements based on X-shooter
+data give different estimates for the RV ({\color{blue} Rubio-D{\'
+    i}ez et al. in prep}). \cite{bressert:12} note an offset compared between the RV of
 the star and the nebular lines from the gas filaments in its
 vicinity. This is consistent with the expectation if the star was not
-formed in situ. Newer measurements based on new X-shooter data give different estimates for the RV ({\color{blue} Rubio-D{\' i}ez et al. in prep}).
-We will therefore refrain from using the RV measurements in this work.
+formed in situ. We will refrain from using the RV measurements in this
+work, and focus on the velocities on the plane of the sky.
 

### 2018-08-21 VFTS682_APJL.tex
addresses Chris's comments 2

@@
 %\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
-\title{Space astrometry of the very massive $\sim$$150\,M_\odot$ runaway star candidate VFTS682}
+\title{Space astrometry of the very massive $\sim$$150\,M_\odot$
+  candidate runaway star VFTS682}
 
@@
 VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space
-  Telescope} photometry. We derive a projected velocity relative to
+  Telescope} astrometry. We derive a projected velocity relative to
 the cluster of $\sim$$30\pm20\kms$ at one $\sigma$ level.  The direction and magnitude of
@@
 
-In has been proposed that most, if not all, stars form in clusters \citep{lada:03}, where massive stars are thought to reside in the innermost cores.
-In this picture, field stars are primarily the result of the dissolution of dense groups.
-However, a significant population of massive stars exists in relative isolation, far from dense clusters or OB associations and their origin remains matter of debate \citep{gvaramadze:12, lamb:16,ward:18}.
-One hypothesis to explain the population of relatively isolated massive stars is that they formed in the field.
+In has been proposed that most, if not all, stars form in clusters
+\citep{lada:03}, where massive stars are thought to reside in the
+innermost cores. In this picture, field stars are primarily the result
+of the dissolution of dense groups.  However, a significant population
+of massive stars exists in relative isolation, far from dense clusters
+or OB associations and their origin remains matter of debate
+\citep{gvaramadze:12, lamb:16,ward:18}.   One hypothesis to explain
+the population of relatively isolated massive stars is that they
+formed in the field \citep[e.g.,][]{parker:07}.
 The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
@@
 a bright X-ray point source at the position of VFTS682 in their
-\emph{Chandra} survey of the region, which suggests the absence of
+\emph{Chandra} survey of the region. The X-ray luminosity of VFTS682
+is significantly lower than known massive binaries in the region, which suggests the absence of
 colliding winds that would be expected in the presence of companions

### 2018-08-21 VFTS682_APJL.tex
addresses Danny's comments 2

@@
 other stars more massive than $150\,M_\odot$ of similar in spectral type.
-The very massive O stars VFTS16 and VFTS72, which have recently been identified as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
+The very massive O stars VFTS16 and (possibly) VFTS72, which have recently been proposed as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
 \end{abstract}
@@
 In has been proposed that most, if not all, stars form in clusters
-\citep{lada:03}, where massive stars are thought to reside in the
+\citep{lada:03}, where massive stars reside in the
 innermost cores. In this picture, field stars are primarily the result
@@
 the population of relatively isolated massive stars is that they
-formed in the field \citep[e.g.,][]{parker:07}.
-The alternative hypothesis is that these massive stars were ejected from the clusters in which they formed.
+formed in the field \citep[e.g.,][]{parker:07}. Another, not
+necessarily mutually exclusive,
+hypothesis is that these massive stars were ejected from the clusters
+in which they formed.
 Such ejections may result from dynamical interactions \citep[e.g.,][]{poveda:67} or from the disruption of binary systems at the death of the companion  star \citep[e.g.,][]{blaauw:61, renzo:18}.
@@
 \citep{schneider:18}. 
-%
-%$137.8^{+27.5}_{-15.9}\,M_\odot$,
-%
 This makes VFTS682 one of the most massive stars known and one of the most extreme objects in the region.
-%
 From the spectral point of view, it is reminiscent of the very
-massive stars % , a.k.a. ``monster stars'',
- in the core of the R136 cluster \citep{dekoter:97,crowther:10, crowther:16}. 
- %
+massive stars in the core of the R136 cluster \citep{dekoter:97,crowther:10, crowther:16}. 
 In particular, a remarkable similarity exist between the
@@
 % current mass estimate of $180^{+30}_{-30}\Msun$.
-% R136 hosts
-% at least two more very massive WN5h stars, R136a1 and R136a2, whose
-% estimated current masses are even higher.
-
-%at a projected offset of 119.4 arcseconds, corresponding to
 
@@
 considered two possible explanation for the offset: either
-the star formed in situ as an isolated massive star, % , but remark that this
-                                % poses a challenge for starformation
-                                % theories (They don't)
-or it was ejected from  R136. N-body simulations % of dynamical interaction in star clusters
+the star formed in situ as an isolated massive star, or it was ejected from  R136. N-body simulations 
 indicate that the ejection of very massive stars like VFTS682 is
-expected \citep[e.g.][]{fujii:11, banerjee:12}.  This is supported by
-the recent findings of other massive runaway stars in the region based
-on proper motion studies.
+expected \citep[e.g.][]{fujii:11, banerjee:12}. The capability of
+young cluster of ejecting a large number of (very) massive stars is supported by
+the recent findings of proper motion studies \citep[][]{lennon:18, drew:18}.
 
@@
   \begin{center}
-    \caption{Selected stellar parameters for VFTS682. }
+    \caption{Stellar parameters of VFTS682. }
     \begin{tabular}{llc|c|c}
@@
        \hline
-       \multicolumn{5}{l}{\emph{Selected stellar parameters VFTS 682}}
-      \\
-      \hline
+      %  \multicolumn{5}{l}{\emph{Selected stellar parameters VFTS 682}}
+      % \\
+      % \hline
      & present day mass  & $[M_\odot]$ & $137.8^{+27.5}_
@@
     {
-      The quoted errors are statistical, and do not include systematic
+      The quoted uncertainties are statistical, and do not include systematic
       effects in the modeling.\\
@@
 luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
-the brightest stars in the region. The absence of radial velocity
+the most luminous stars in the region. The absence of radial velocity
 variations suggests that the star is unlikely to have close binary
@@
 Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
-shows mid-infrared excess \citep{gruendl:09}.
+shows a mid-infrared excess \citep{gruendl:09}.
 
@@
   \label{fig:dist}
+\end{figure*}
+
+\begin{figure*}%[tbp]
+  \centering
+  \includegraphics[width=\textwidth]{./figures/Figure2_paper_vfts682_hst_corr_nor136err}  
+  \caption{The thick yellow arrows indicate proper motions relative R136 (blue
+    circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The prolongations in the opposite
+    direction are proportional to the apparent ages of the stars, and
+    the thin lines illustrate the error cone on the potential
+    origin. The green thick arrow
+    indicates the HST proper motion of VFTS682, with the relative
+    error cone and prolongation proportional to its apparent
+    age. The two blue circles indicate the regions of radii 0.01 and 0.05
+    degrees around the core of R136.}
+
+  \label{fig:main}
 \end{figure*}
@@
 
-\begin{figure*}%[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{./figures/Figure2_paper_vfts682_hst_corr_nor136err}  
-  \caption{The thick yellow arrows indicate proper motions relative R136 (blue
-    circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The prolongations in the opposite
-    direction are proportional to the apparent ages of the stars, and
-    the thin lines illustrate the error cone on the potential
-    origin. The green thick arrow
-    indicates the HST proper motion of VFTS682, with the relative
-    error cone and prolongation proportional to its apparent
-    age. The two blue circles indicate the regions of radii 0.01 and 0.05
-    degrees around the core of R136.}
-
-  \label{fig:main}
-\end{figure*}
-
 \section{Discussion}
@@
 Also \citet{banerjee:12} suggested, based on their N-body simulations,
-% N-body simulations % of fully segregated
-% clusters with all massive stars in binaries
 that VFTS682 was ejected from R136. In their

### 2018-08-21 VFTS682_APJL.tex
addressed minors from Roeland

@@
 led to two intriguing hypotheses: % posed in earlier work:
-either (a) it formed in relative isolation % through a new mode of star formation
-or (b) it was ejected dynamically from the central cluster.
+either it formed in relative isolation % through a new mode of star formation
+or it was ejected dynamically from the central cluster.
 
@@
   Telescope} astrometry. We derive a projected velocity relative to
-the cluster of $\sim$$30\pm20\kms$ at one $\sigma$ level.  The direction and magnitude of
+the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval).  The direction and magnitude of
 the proper motion are consistent with VFTS682 being a slow runaway
-ejected from the central cluster, given its distance and the inferred age.
+ejected from the central cluster, given its distance and the inferred
+age. The \todo{rephrase positive, see Fabian and Roeland's comments}
+
 However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.
 
-If future data confirm the runaway nature, this would make the VFTS682
-the most massive runaway star known to date.  While we cannot prove
-this solidly from the current data alone, we do consider this
-hypothesis as the most plausible because of a variety of
+If future data confirm the runaway nature, this would make VFTS682
+the most massive runaway star known to date. This hypothesis is
+further supported by a variety of
 circumstantial clues. The central cluster is known to harbor several
 other stars more massive than $150\,M_\odot$ of similar in spectral type.
-The very massive O stars VFTS16 and (possibly) VFTS72, which have recently been proposed as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
+The very massive O stars VFTS16 (and possibly VFTS72), which has recently been proposed as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
 \end{abstract}
@@
 of massive stars exists in relative isolation, far from dense clusters
-or OB associations and their origin remains matter of debate
+or OB associations and their origin remains a matter of debate
 \citep{gvaramadze:12, lamb:16,ward:18}. One hypothesis to explain
@@
 $\sim$$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
-considered two possible explanation for the offset: either
+considered two possible explanations for the offset: either
 the star formed in situ as an isolated massive star, or it was ejected from  R136. N-body simulations 
@@
 variations suggests that the star is unlikely to have close binary
-companions \citet{bestenlehner:11}, unless the orbital inclination is
+companions \citep[][]{bestenlehner:11}, unless the orbital inclination is
 very high. Bayesian fits of the stellar
@@
                      &  SIMBAD \\       
-        &$      \delta\mathrm{RA}$  &[marcsec] & \phantom{-0}0.0547                      
+        &$      \delta\mathrm{RA}$  &[mas] & \phantom{-0}0.0547                      
         &  (2, 5)
   \\        
-               &$     \delta\mathrm{DEC}$  &[marcsec] & \phantom{-0}0.0268 
+               &$     \delta\mathrm{DEC}$  &[mas] & \phantom{-0}0.0268 
                      &  (2, 5) \\  
@@
                      &  (1) \\        
-                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &[$\masyr$] & $0.0226$
+                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &  & $0.0226$
                         & (1)  \\         
@@
       &$\delta\mu_{Gaia}$  &[$\masyr$] & $0.13\pm0.09$ &  (1,5) \\
-      &$v_\mathrm{2d}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
+      &$v_\mathrm{2D}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
       &$\theta_{Gaia}$  &[degrees] &  $14_{-31}^{+36}$  & (1,5)\\  
-      % &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
+      % &$v_\mathrm{2D, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
  %
@@
        &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (4) \\
-                  &$v_\mathrm{2d, HST}$  &[$\kms$] & $47\pm24$ & (4)\\  
+                  &$v_\mathrm{2D, HST}$  &[$\kms$] & $47\pm24$ & (4)\\  
                       &$\theta_\mathrm{HST}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,5)\\  
-                  %        &$v_\mathrm{2d, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
+                  %        &$v_\mathrm{2D, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
     \hline
@@
       \hline
-      &$v_\mathrm{2d}$  &[$\kms$] & $29\pm 6$ & (2) \\  
+      &$v_\mathrm{2D}$  &[$\kms$] & $29\pm 6$ & (2) \\  
       &$\theta$  &[degrees] &  $\sim0$  & \\ 
@@
 VFTS682.
-This star is also relatively isolated in (near) infrared.
+This star is also relatively isolated in the (near) infrared.
 There nearest bright (near-)infrared sources detected by \emph{Spitzer} \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e.
@@
   \caption{\emph{Left panel}: distribution of OB-type and Wolf-Rayet stars in proper
-    motion relative to R136. VFTS682 is not an outlier, but
-    its relative proper motion matches the expected value if it were indeed
-    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
+    motion relative to R136 from \emph{Gaia} DR2. VFTS682 is not an outlier, but
+    its relative proper motion matches the expected value (green
+    vertical band) if it were indeed
+    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to $\kms$
     assuming a distance of 50\,kpc. \emph{Right panel}:  distribution of
@@
     the best value is in agreement with the hypothesis of dynamical
-    ejection. The peak at $\theta\simeq\pi/2$ is due to stars
+    ejection (green vertical band). The peak at $\theta\simeq\pi/2$ is due to stars
     belonging to NGC2060. In both
@@
   \includegraphics[width=\textwidth]{./figures/Figure2_paper_vfts682_hst_corr_nor136err}  
-  \caption{The thick yellow arrows indicate proper motions relative R136 (blue
+  \caption{The thick yellow arrows indicate proper motions relative to
+    R136 (blue
     circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The prolongations in the opposite
@@
 i.e., groups of observations separated from other groups by a gap of at
-least four days, used in the astrometric solution is seventeen for this
-star. The reported astrometric excess noise is zero.  These values
+least four days, used in the astrometric solution is seventeen. The reported astrometric excess noise is zero.  These values
 suggest that the \emph{Gaia} DR2  data for VFTS682 are
@@
 proper motion of nearby stars with reliable astrometric data.  They
-selected bright ($G<17$) stars within 0.05 degrees of R136 and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1).
-
-With these we compute the relative proper motion
-$\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.   We also compute the total projected 2d velocity
-$v_\mathrm{2d}$ and
-the angle $\theta$ between the direction of motion and the vector
+selected bright ($G<17$) stars within 0.05 degrees of R136 and exclude
+sources with proper motion errors greater than $0.01\masyr$ in both
+coordinates (see their  Sect.~2.1).  Using this definition of the
+local frame, we compute the relative proper motion
+$\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.   We also compute the total projected 2D velocity
+$v_\mathrm{2D}$ and the angle $\theta$ between the direction of motion and the vector
 connecting the center of R136 with the current position of VFTS 682,
@@
 
-The 30 Doradus region was target of an a two-epoch photometric campaign with HST providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
+The 30 Doradus region was targeted by a two-epoch photometric campaign with HST providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
 % Visually, the star is clearly isolated and does not appear to have any close neighbor.
@@
 \citet{platais:15, platais:18} analyzed the HST data to determine the
-relative proper motion and identify candidate runaway stars. The
+relative proper motions and identify candidate runaway stars. The
 brightest stars (V$<$14) are saturated in the data set and have been
@@
 VFTS682 is relatively slow and is therefore not discussed in their paper.
-However, their measurements are useful here, since they provides an  estimate of the proper motion of VFTS682 that is independent from the \emph{Gaia} measurement.
+However, their measurements are useful here, since they provide an  estimate of the proper motion of VFTS682 that is independent from the \emph{Gaia} measurement.
 
-The HST provides proper motions that are relative to the bulk motion
+The HST study provides proper motions that are relative to the bulk motion
 of the majority of the stars in the field of view. The full 30 Dor
@@
 to R136 of
-VFTS682. $\theta$ is the angle between the projected radial direction from the
+VFTS682. For comparison, we also show the relative
+proper motion of VFTS16 and VFTS72 (gray points), and the distribution
+in relative proper motion and flight direction for all the VFTS O/B-type and Wolf-Rayet stars with
+\emph{Gaia} DR2 errors on the proper motion
+components of less than $0.1\,\mathrm{mas\ yr^{-1}}$ (dark blue lines,
+including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light hatched
+blue). $\theta$ is the angle between the projected radial direction from the
 center of R136 to each star and its projected relative proper motion
@@
 lifetime. Although the error bars are substantial and VFTS682 is not
-an outlier, the agreement
+an outlier compared to other OB-type and Wolf-Rayet stars, the agreement
 suggests that the star is indeed a ``slow-runaway'' as suggested by
-\cite{bestenlehner:11}. For comparison, we also show the relative
-proper motion of VFTS16 and VFTS72 (gray points), and the distribution
-in relative proper motion and flight direction for all the VFTS O/B-type and Wolf-Rayet stars with
-\emph{Gaia} DR2 errors on the proper motion
-components of less than $0.1\,\mathrm{mas\ yr^{-1}}$ (dark blue lines,
-including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light hatched
-blue).
+\cite{bestenlehner:11}.
 
@@
 The kinematic age $\tau_\mathrm{kin}$ is consistent with a very early
-ejection from the cluster, given the apparent age of the star.
+ejection from the cluster (see \Tabref{tab:star_param}).
 % of $1.0\pm 0.2$\,Myr 
@@
 % \cite{dekoter:97,crowther:10,crowther:16}
- with an remarkable similarity in particular with
+ with a remarkable similarity in particular with
 the spectrum of R136a3. The isolation of
@@
 upper limit to the timescale to form the ``bully binary'' in
-R136. Such binary could be R145 according to \cite{fujii:11}, and it might be an
+R136. Such a binary could be R145 according to \cite{fujii:11}, and it might be an
 ideal observational candidate for a dynamically formed progenitor system of
@@
    % and C.\,J.\,Evans
-   for help and discussions, European Unions Horizon 2020 research and innovation programme from the European Research Council (ERC), Grant agreement No.
-715063 [SdM], the NRC-Canada Plaskett Fellowship [VHB].
-
+   for help and discussions. We also thank the HSTPROMO
+(High-resolution Space Telescope PROper MOtion) Collaboration for
+sharing and discussing ideas relevant to this work. SdM acknowledges
+the European Unions Horizon 2020 research and innovation programme
+from the European Research Council (ERC), Grant agreement No. 715063
+and VHB the NRC-Canada Plaskett Fellowship.
 This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).

### 2018-08-21 VFTS682_APJL.tex
addressed minors from Fabian -- only major thing about chance results from HST and Gaia left

@@
 In has been proposed that most, if not all, stars form in clusters
-\citep{lada:03}, where massive stars reside in the
-innermost cores. In this picture, field stars are primarily the result
+\citep{lada:03}. In this picture, field stars are primarily the result
 of the dissolution of dense groups.  However, a significant population
@@
 postion and direction of the $\sim$$100\Msun$ star VFTS16 is consistent
-with a runaway origin from R136. They also found a less clear case for
+with a runaway origin from R136. They found a less clear case for
 VFTS72, and in both cases some tension between the kinematic age of
@@
       &age & [Myr] & $1.0\pm0.2$ & (1) \\
-      &mass loss rate & [$M_\odot \ \mathrm{yr}^{-1}$] & $10^{-4.1\pm0.2}$ & (2)\\
+      & mass loss rate & $\log_{10}(\dot{M}/[M_\odot \ \mathrm{yr}^{-1}])$ & $-4.1\pm0.2$ & (2)\\
       % &\multirow{2}{*}{surface helium mass fraction} &  & 0.45 & (2)\\
@@
 parameters against evolutionary tracks \citep{brott:11, kohler:15}
-using the BONNSAI code \citep{schneider:17, schneider:18} provide
+using the BONNSAI code \citep{schneider:14,schneider:17} provide
 estimates for the age, present mass and initial mass, % indicating that
@@
 is significantly lower than known massive binaries in the region, which suggests the absence of
-colliding winds that would be expected in the presence of companions
+colliding winds. These would be expected in the presence of companions
 even for extreme mass ratios, given the extremely large mass of
-VFTS682.
-This star is also relatively isolated in the (near) infrared.
-There nearest bright (near-)infrared sources detected by \emph{Spitzer} \citep{meixner:06} and resolved in the \emph{VISTA} Magellanic Clouds Survey \citep{cioni:11} are located at a distance of about 10 arcsecond, i.e.
-about 2.4 pc.
-\cite{walborn:13} speculate that these nearby young stars may represent a case of starformation triggered by the the wind of VFTS682, but chance alignment cannot be excluded.
+VFTS682. This star is also relatively isolated in the (near)
+infrared. The nearest bright (near-)infrared sources detected by
+\emph{Spitzer} \citep{meixner:06} and resolved in the \emph{VISTA}
+Magellanic Clouds Survey \citep{cioni:11} are located at a distance of
+about 10 arcsecond, i.e. about 2.4 pc. \cite{walborn:13} speculate
+that these nearby young stars may represent a case of star formation triggered by the the wind of VFTS682.
 
@@
 stars. \citet{bestenlehner:11} estimate a mass loss rate of
-$10^{-4.1\pm0.2}\,M_\odot \ \mathrm{yr}^{-1}$, not accounting for the
+$\log_{10}(\dot{M}/[M_\odot \ \mathrm{yr}^{-1}])=-4.1\pm0.2$, not accounting for the
 possible effect of clumping.  We therefore caution
-for over-interpreting the estimates for the radial velocities. % , but we will summarize the estimates that have been published. 
+for over-interpreting the radial velocities estimates. % , but we will summarize the estimates that have been published. 
 \citet{bestenlehner:11}  estimate a radial velocity (RV) of
@@
 $270\pm10\kms$. This was suggested as indicating a runaway nature, but
-it is no proof of it, and more recent measurements based on X-shooter
-data give different estimates for the RV ({\color{blue} Rubio-D{\'
-    i}ez et al. in prep}). \cite{bressert:12} note an offset compared between the RV of
+it is no proof of it. % , and more recent measurements based on X-shooter
+% data give different estimates for the RV ({\color{blue} Rubio-D{\'
+%     i}ez et al. in prep}).
+\cite{bressert:12} note an offset compared between the RV of
 the star and the nebular lines from the gas filaments in its
@@
 
-We adopt as distance to the LMC the value of $50$\,kpc. The error on
+We adopt a distance to the LMC of $50$\,kpc. The error on
 the distance determination is small ($\lesssim2\%$,
-\citealt{pietrzynski:13}) and any possible offset in the radial direction between R136 or VFTS682 and the distance we adopted for the LMC is probably much smaller (likely $\ll 1\%$, e.g., \citealt{Luks+1992}).
-These uncertainties are negligible compared to the errors in the proper motion discussed below.
+\citealt{pietrzynski:13}) and any possible offset in the radial
+direction between R136 or VFTS682 and the distance we adopted for the
+LMC is probably much smaller ($\sim$$0.5\%$, e.g., \citealt{Luks+1992}). These uncertainties are negligible compared to the errors in the proper motion discussed below.
 
@@
 % stars in the region, however, the high extinction makes it faint enough such that saturation effects are not an issue.
-VFTS682 is relatively slow and is therefore not discussed in their paper.
-However, their measurements are useful here, since they provide an  estimate of the proper motion of VFTS682 that is independent from the \emph{Gaia} measurement.
+VFTS682 is relatively slow and is therefore not discussed in
+\citealt{platais:18}.  However, their measurements provide useful
+complementary estimate of the proper motion of VFTS682 independent from the \emph{Gaia} data.
 
@@
 Assuming VFTS682 indeed originates from R136, we can calculate its kinematic
-age simply as:
+age as:
 \begin{equation}
@@
 where $d_\parallel = 119.4\,\mathrm{arcsec}$ is the angular distance from VFTS682 to
-the core of the cluster \citep[corresponding to $\sim$29\,pc at a
-distance of 
-50\,kpc,][]{bestenlehner:11}.
+the core of the cluster \citep[][]{bestenlehner:11}. % \citep[corresponding to $\sim$29\,pc at a
+% distance of 
+% 50\,kpc,][]{bestenlehner:11}.
 % As in the rest of this study, we neglect for
@@
 % available.
+
 If confirmed, it means that isolated star formation is
@@
 indeed smaller than the shortest stellar lifetime
-\citep[$\sim$3\,Myr,][]{brott:11,zapartas:17}, the ejection of VFTS682
+\citep[$\sim$3\,Myr,][]{brott:11,kohler:15, zapartas:17}, the ejection of VFTS682
 from the disruption of a binary is excluded. %  since the region is too young for stars
@@
 $\sim$1.5\,Myr for VFTS16 found in \cite{lennon:18}, which indicates
-that VFTS682 was ejected later than VFTS16, at a time similar to VFTS72.
+that VFTS682 was ejected later than VFTS16, potentially at a time similar to VFTS72.
 
-If the star were ejected dynamically, its exceptionally large mass raises the question of which stars must populate
+If the star were ejected dynamically, its isolation makes it an ideal target to constrain the stellar physics of
+stars with masses well above $\sim$$100\,M_\odot$ while avoiding
+crowding issues. Moreover, 
+ its exceptionally large mass raises the question of which stars must populate
 the core of the cluster. N-body dynamics typically ejects the least
@@
 cluster.
-
-The spectral type of VFTS682
-\citep[WNh5,][]{bestenlehner:11} is the same as R136a1-a3, i.e.~the
-three most massive stars detected in the core of the cluster, %  by
-% \cite{dekoter:97,crowther:10,crowther:16}
- with a remarkable similarity in particular with
-the spectrum of R136a3. The isolation of
-VFTS682 makes it an ideal target to constrain the stellar physics of
-stars with masses well above $\sim$$100\,M_\odot$ while avoiding
-crowding issues.
 
@@
 pair-instability supernova without compact remnant formation, or
-possibly direct collapse to a black hole above the $2^\mathrm{nd}$
+possibly direct collapse to a black hole above the pair-instability
 mass gap. The amount of mass loss of these stars will determine their final core

### 2018-08-22 MNRASformat/VFTS682_MNRASL.tex
MNRAS Letter ready for submission

--- 2018-08-21 VFTS682_APJL.tex
@@
-\documentclass[apjl,twocolumn]{emulateapj}
-\usepackage{graphicx}
-\usepackage[varg]{txfonts}
-%\usepackage[pagebackref,breaklinks,colorlinks,citecolor=blue]{hyperref}        
-\usepackage[%draft, pagebackref,
-colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue]{hyperref}
-%\renewcommand*{\backref}[1]{[#1]}  % gives neat back references in the pdf
-\usepackage{aas_macros}
-%\usepackage{amsmath} % clashes with emulateapjj for some reaons
+\documentclass[a4paper,fleqn,usenatbib]{mnras}
+\usepackage{newtxtext,newtxmath}
+\usepackage[T1]{fontenc}
+\usepackage{ae,aecompl}
+
+%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
+
+% Only include extra packages if you really need them. Common packages are:
+\usepackage{graphicx}	% Including figure files
+\usepackage{amsmath}	% Advanced maths commands
+\usepackage{amssymb}	% Extra maths symbols
+\usepackage[%draft, 
+% pagebackref,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue
+]{hyperref}
 \usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{comment}
+\usepackage{url}
 \usepackage{multirow}
-%\usepackage{chngpage}
-%\usepackage{lscape}
-\usepackage{url}
-\newcommand{\todo}[1]{{\large $\blacksquare$~\textbf{\color{red}[#1]}}~$\blacksquare$}
-\newcommand{\udef}{\stackrel{\mathrm{def}}{=}}
-% for tree diagram
-%\usepackage{forest}
-%\usepackage{tikz-qtree}
-%\usetikzlibrary{shadows,trees}
 
@@
 \newcommand{\SdM}[1]{{{\color{brown}{#1}}}}
-%\renewcommand{\SdM}[1]{{{{#1}}}}
-
 \newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}
@@
 
+\title[Space astrometry of the very massive $\sim$$150\,M_\odot$ candidate runaway star VFTS682]{Space astrometry of the very massive $\sim$$150\,M_\odot$  candidate runaway star VFTS682}
+
+\author[Renzo et al.]{M.~Renzo$^{1}$\thanks{Corresponding author:
+    M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
+  R.~P.~van~der~Marel$^{3,4}$,
+  \newauthor{E.~Laplace$^{1}$, J.~M.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
+    V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{1,8,9}$,
+  }
+\newauthor{A.~de~Koter$^{1}$,
+      N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~Schneider$^{13}$, J.~S.~Vink$^{14}$}}
+
+% These dates will be filled out by the publisher
+\date{Accepted XXX. Received YYY; in original form ZZZ \\ Affiliations
+can be found at the end of this manuscript}
+
+% Enter the current year, for the copyright statements etc.
+\pubyear{2018}
+
 \begin{document}
-
-% \title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
-%                                 % ejected %% to fit in one line
-% runaway star known to date}
-%\title{Gaia's constraints on the  $\sim$$150\,M_\odot$  runaway star candidate VFTS682}
-\title{Space astrometry of the very massive $\sim$$150\,M_\odot$
-  candidate runaway star VFTS682}
-
-\author{M.~Renzo$^{*,1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
-  R.~P.~van~der~Marel$^{3,4}$, E.~Laplace$^{1}$,
-  J.~M.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
-  V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{1,8,9}$,  A.~de~Koter$^{1}$,
-  N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~Schneider$^{13}$, J.~S.~Vink$^{14}$}
-\affil{ $^{*}$ Corresponding author: \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}.  A list of the affiliations can be found at the end of this paper.}
+\label{firsxtpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\maketitle
 
@@
 
-How very massive stars form is still an intriguing open question in
+How very massive stars form is an intriguing open question in
 stellar astrophysics.  VFTS682 is among the most massive stars known,
-with an inferred initial mass of $\sim$$150\,M_\odot$. It is located
+with an inferred initial mass of $\gtrsim$$150\,M_\odot$. It is located
 in 30 Doradus at a projected distance of 29\,pc from the central
@@
 or it was ejected dynamically from the central cluster. 
-
-% Aiming to shine light on this debate,
 In this context, 
@@
   Telescope} astrometry. We derive a projected velocity relative to
-the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval).  The direction and magnitude of
-the proper motion are consistent with VFTS682 being a slow runaway
+the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval). Although
+the error bars are substantial, two completely independent measures
+suggest that VFTS682 might be a slow runaway
 ejected from the central cluster, given its distance and the inferred
-age. The \todo{rephrase positive, see Fabian and Roeland's comments}
-
-However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.
-
-If future data confirm the runaway nature, this would make VFTS682
-the most massive runaway star known to date. This hypothesis is
-further supported by a variety of
+age. This hypothesis is further supported by a variety of
 circumstantial clues. The central cluster is known to harbor several
-other stars more massive than $150\,M_\odot$ of similar in spectral type.
-The very massive O stars VFTS16 (and possibly VFTS72), which has recently been proposed as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
+other stars more massive than $150\,M_\odot$ of similar in spectral
+type. The very massive O stars VFTS16 (and possibly VFTS72), which has
+recently been proposed as a dynamically ejected runaway, provides
+direct evidence that R136 is indeed capable of ejecting some of its
+most massive members, consistent with the predictions of numerical
+simulations. If future data confirm the runaway nature, this would make VFTS682
+the most massive runaway star known to date. 
 \end{abstract}
 
-\keywords{stars: kinematics, stars: runaways, stars: individual: VFTS682}
-\maketitle{}
+\begin{keywords}
+  stars: kinematics, stars: runaways, stars: individual: VFTS682
+\end{keywords}
 
@@
 the population of relatively isolated massive stars is that they
-formed in the field \citep[e.g.,][]{parker:07}. Another, not
-necessarily mutually exclusive,
+formed in the field \citep[e.g.,][]{parker:07}. Another
 hypothesis is that these massive stars were ejected from the clusters
@@
 It is a hydrogen-rich Wolf-Rayet star of spectral type WNh5.
-Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $\sim$$137.8^{+27.5}_{-15.9}\,M_\odot$ corresponding to an initial mass of $\sim$$150.0^{+28.7}_{-17.4}\,M_\odot$
+Spectral analysis and comparison with evolutionary models lead to an inferred present-day mass of $137.8^{+27.5}_{-15.9}\,M_\odot$ corresponding to an initial mass of $150.0^{+28.7}_{-17.4}\,M_\odot$
 \citep{schneider:18}. 
@@
 VFTS682 stands out by its relative isolation at a projected distance of 119.4 arcseconds, corresponding to 
-$\sim$$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
+$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
 considered two possible explanations for the offset: either
@@
 young cluster of ejecting a large number of (very) massive stars is supported by
-the recent findings of proper motion studies \citep[][]{lennon:18, drew:18}.
+the recent findings of proper motion studies \citep[e.g.,][]{lennon:18, drew:18}.
 
@@
   Telescope} (HST) photometry and identified 10 stars likely ejected
-from R136. \citet{lennon:18} investigate the kinematics of  isolated
+from R136. \citet{lennon:18} investigated the kinematics of  isolated
 O-type stars in the region using the second \emph{Gaia} data release
-\cite[DR2,][]{gaia:16,brown:18} and show that the proper motion,
+\cite[DR2,][]{gaia:16,brown:18} and showed that the proper motion,
 postion and direction of the $\sim$$100\Msun$ star VFTS16 is consistent
@@
 
-In this paper we present an analysis of the new kinematical
+In this paper we present an analysis of the new kinematic
 constraints for VFTS682 provided by \emph{Gaia} DR2 and constraints
 from HST proper motions  by \citet{platais:18}.   We discuss the
-implications for the hypothesis that VFTS682 is a slow runaway star
+implications of the hypothesis that VFTS682 is a slow runaway star
 ejected from R136.
@@
     \end{tabular}
-    \tablecomments
-    {
+    % \tablecomments
+    % {
       The quoted uncertainties are statistical, and do not include systematic
@@
       % (3)~\cite{rubio-diez:17}
-    }
+    % }
   \end{center}
@@
 05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} (J2000), was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
 \citet{bestenlehner:11}  analyzed the spectra to infer the stellar
@@
 luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
-the most luminous stars in the region. The absence of radial velocity
+the most luminous stars in the region. The absence of periodic radial
+velocity (RV)
 variations suggests that the star is unlikely to have close binary
@@
                      &  SIMBAD \\       
-        &$      \delta\mathrm{RA}$  &[mas] & \phantom{-0}0.0547                      
+        &$      \delta\mathrm{RA}$  &[mas] & \phantom{-00}0.0547                      
         &  (2, 5)
   \\        
-               &$     \delta\mathrm{DEC}$  &[mas] & \phantom{-0}0.0268 
+               &$     \delta\mathrm{DEC}$  &[mas] & \phantom{-00}0.0268 
                      &  (2, 5) \\  
@@
       \hline
-      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & $0.103\pm0.080$ & (1,5) \\
+      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & $0.10\pm0.08$ & (1,5) \\
       &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & $0.086\pm0.100$ &  (1,5) \\
@@
       \hline
+      \multicolumn{5}{l}{\emph{Weighted average relative proper motion
+      for VFTS682}}\\
+      \hline
+      &$\delta\mu_\mathrm{RA, avg}$  &[$\masyr$] & $0.08\pm0.07$ & (5) \\
+      &$\delta\mu_\mathrm{DEC, avg}$  &[$\masyr$] & $0.14\pm0.07$ &  (5) \\
+       &$\delta\mu_\mathrm{avg}$  &[$\masyr$] & $0.16\pm0.07$ &  (5) \\
+      &$v_\mathrm{2D, avg}$  &[$\kms$] & $38\pm17$ & (5)\\
+      % &$\theta_\mathrm{avg}$  &[degrees] &  \todo{TBD} $_{-}^{+}$   & (5)\\  
+      \hline
        \multicolumn{5}{l}{\emph{Expected proper motion if ejected from
@@
 
-\hline
-  %    &$\delta v_\mathrm{rad}$  &[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
+% \hline
+%   %    &$\delta v_\mathrm{rad}$  &[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
 
-      \hline
+%       \hline
 
     \end{tabular}
-    \tablecomments
-    { Proper motions are based on Gaia data unless indicated otherwise
-      in subscript.    The conversion factor assuming a distance of
-      50\,kpc is such that $1\masyr$ corresponds to $237\kms$.  The
+    % \tablecomments
+    % {
+
+    Assuming a distance of 50\,kpc, the conversion from proper motion
+    to projected velocities is such that $1\masyr$ corresponds to $237\kms$.  The
       correlation coefficient is indicated as $\rho$. The angle
@@
       % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
-    }
+    % }
   \end{center}
@@
 unusual for Wolf-Rayet stars and more typical for Luminous Blue
-Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
+Variable (LBV) stars \citep{parker:93, udalski:08, bestenlehner:11}. It also
 shows a mid-infrared excess \citep{gruendl:09}.
@@
 for over-interpreting the radial velocities estimates. % , but we will summarize the estimates that have been published. 
-\citet{bestenlehner:11}  estimate a radial velocity (RV) of
+\citet{bestenlehner:11}  estimate a RV of
 $300\pm10\kms$ using the  N{\footnotesize V} $\lambda4944$ line, which
@@
 direction between R136 or VFTS682 and the distance we adopted for the
-LMC is probably much smaller ($\sim$$0.5\%$, e.g., \citealt{Luks+1992}). These uncertainties are negligible compared to the errors in the proper motion discussed below.
+LMC is probably much smaller ($\sim$$0.5\%$, e.g., \citealt{luks:92}). These uncertainties are negligible compared to the errors in the proper motion discussed below.
 
@@
 $v_\mathrm{2D}$ and the angle $\theta$ between the direction of motion and the vector
-connecting the center of R136 with the current position of VFTS 682,
+connecting the center of R136 with the current position of VFTS682,
 i.e. such that $\theta = 0$ is the angle corresponding to perfectly
@@
 This by itself gives further confidence in ability of \emph{Gaia} to measure the proper motion reliably without being affected by crowding.
-
 \citet{platais:15, platais:18} analyzed the HST data to determine the
@@
 to R136 of
-VFTS682. For comparison, we also show the relative
+VFTS682 from \emph{Gaia} DR2. $\theta$ is the angle between the projected radial direction from the
+center of R136 to each star and its projected relative proper motion
+(see also \Figref{fig:main}). Dynamical ejections from the
+cluster should %are expected to
+produce close to radial ejections, i.e.~$\theta\simeq0$. The green
+vertical bands highlight the expectations for these two quantities if VFTS682 were ejected from R136 early in its
+lifetime. For comparison, we also show the relative
 proper motion of VFTS16 and VFTS72 (gray points), and the distribution
-in relative proper motion and flight direction for all the VFTS O/B-type and Wolf-Rayet stars with
+in relative proper motion and flight direction for all the VFTS OB-type and Wolf-Rayet stars with
 \emph{Gaia} DR2 errors on the proper motion
@@
 including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light hatched
-blue). $\theta$ is the angle between the projected radial direction from the
-center of R136 to each star and its projected relative proper motion
-(see also \Figref{fig:main}). Dynamical ejections from the
-cluster should %are expected to
-produce
-close to radial ejections, i.e.~$\theta\simeq0$. The green vertical bands highlight the expectations for these
-two quantities if VFTS682 were ejected from R136 early in its
-lifetime. Although the error bars are substantial and VFTS682 is not
+blue). Although the error bars are substantial and VFTS682 is not
 an outlier compared to other OB-type and Wolf-Rayet stars, the agreement
@@
 
-% Figure~\ref{fig:dist} illustrates the kinematic properties of VFTS OB-type
-% and Wolf-Rayet stars in the 30 Doradus region with \emph{Gaia} DR2 errors on the proper motion
-% components of less than $0.1\,\mathrm{mas\ yr^{-1}}$ (dark blue lines,
-% including VFTS682), and less than $0.05\,\mathrm{mas\ yr^{-1}}$ (light hatched
-% blue). The green vertical bands show the expectations for VFTS682 if
-% it were ejected from R136 early in its lifetime. We highlight this star together with the two O-type
-% runaways, VFTS16 and VFTS72, \mbox{(re-)identified} by \cite{lennon:18}.
-
-% The left panel shows the distribution in proper motion relative
-% to R136 from \emph{Gaia} DR2, selecting stars with small error bars only.
-% Although VFTS682 is not peculiar, its relative proper motion
-% falls in the range expected for it to reach its present day isolation
-% within its apparent age (green vertical band). This suggests it is indeed a ``slow-runaway'' as
-% suggested by \cite{bestenlehner:11}.
 Subtracting
@@
 are consistent with each other, but also with no relative motion
-within $2\sigma$.
-
-% The left panel shows the distribution in two-dimensional proper motion relative
-% to R136, and shows that, although VFTS682 is not a peculiar star in terms of
-% its motion relative to R136. Subtracting the mean motion of the field,
-% we obtain $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$ and
-% $\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$. Both values
-% are consistent with each other and indicate
-% a non-zero peculiar projected motion that would make it a slow runaway
-% star as expected by \cite{bestenlehner:11} if the star was ejected
-% from the cluster,  but in both cases the meaurements are also consistent with zero within 2$\sigma$. 
-% More importantly, both the proper motion
-% measurement from \emph{Gaia} and HST fall in the range expected to
-% reach its present day location % at $\sim$29\,pc from the cluster core
-% if the
-% star was ejected very early in its evolution. This range is
-% highlighted by the green vertical band, and its width is determined by
-% the uncertainty in the present-day age of the star (cf.~\Tabref{tab:star_param}).
-
-% The right panel shows the distributions in projected flight direction for these
-% stars: $\theta$ is the angle between the projected radial direction from the
-% center of R136 to each star and its projected relative proper motion (see also \Figref{fig:main}). Dynamical ejections from the
-% cluster should %are expected to
-% produce
-% close to radial ejections, i.e.~$\theta\simeq0$. On this panel, the green band highlights
-% % the range of angles corresponding to
-% an opening angle around the
-% projected radial direction of 45\,degrees. The direction of motion of VFTS682 % is
-% % poorly constrained because of the large errors on the proper motion
-% % components both in HST and \emph{Gaia} data, but
-% % it
-% is consistent with dynamical ejection from R136.
-We note that
-VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
-large peculiar radial velocity and therefore accurate distances along
-the line of sight are
-needed to contrain the flight direction in three dimensions.
+within $2\sigma$. The weighted average of these two independent
+measurements is $\delta \mu_\mathrm{avg}=0.16\pm0.07\masyr
+(38\pm17\kms)$.
 
@@
 between the apparent age and the present day distance from the cluster
-core for VFTS16 and VFTS72 \citep[][]{lennon:18}.
+core for VFTS16 and VFTS72 \citep[][]{lennon:18}. We note that
+VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
+large peculiar radial velocity and therefore accurate distances along
+the line of sight are
+needed to contrain the flight direction in three dimensions.
 
@@
   \label{eq:kin_age}
-  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu_{Gaia}} \simeq
-  \frac{119.4\,\mathrm{arcsec}}{0.13\masyr} \simeq 0.9\pm\,0.6\, \mathrm{Myr} \ \ ,
+  \tau_\mathrm{kin} = \frac{d_\parallel}{\delta\mu_\mathrm{avg}} \simeq
+  \frac{119.4\,\mathrm{arcsec}}{0.16\masyr} \simeq 0.7\pm\,0.3\, \mathrm{Myr} \ \ ,
 \end{equation}
@@
 % simplicity the error on the distance estimates, because it is negligible compared to other uncertainties.
-The kinematic age $\tau_\mathrm{kin}$ is consistent with a very early
+The kinematic age $\tau_\mathrm{kin}$ is consistent with an early
 ejection from the cluster (see \Tabref{tab:star_param}).
@@
 Based on our results, we consider likely that VFTS682 is the most massive
-(slow) runaway known to date, with a two-dimensional
+runaway known to date, with a two-dimensional
 projected velocity with respect to R136 of
-$\sim38\kms$ (averaging the \emph{Gaia} DR2 and HST
-results), but compatible with zero within $2\sigma$. Due to the large error bars, this result will need
+$38\pm17\kms$ (averaging the \emph{Gaia} DR2 and HST
+results). Due to the large error bars, this result will need
 to be revisited with future astrometric data. % to be revisited as updated
@@
 \emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
-$0.9\pm0.6$\,Myr ago. If the
+$0.7\pm0.3$\,Myr ago. If the
 cluster age \citep[$\lesssim2$\,Myr,][]{crowther:10, sabbi:12} is
@@
 $\sim$1.5\,Myr for VFTS16 found in \cite{lennon:18}, which indicates
-that VFTS682 was ejected later than VFTS16, potentially at a time similar to VFTS72.
+that VFTS682 was ejected later than VFTS16, and potentially later than
+VFTS72 too.
 
@@
 
-The simulations of \cite{fujii:11} suggest that
+The N-body simulations of \citet{banerjee:12} suggest that VFTS682 was ejected from R136. In their
+model, they relied on (dynamically driven) stellar mergers to explain the high masses of
+VFTS682 and the massive members of R136. They
+demonstrated that the cluster potential does not significantly change
+the velocity of the star after the ejection.
+% While it is not possible to
+% robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
+% like fast rotation, which however could have already slowed down
+% because of the wind angular momentum losses during and after the
+% merger.
+To eject such a massive object, the cluster is
+expected to have produced a large number of massive runaways. Indeed, several %relatively
+isolated massive stars are observed in the region \citep[][]{evans:10,lennon:18}, some with known
+large radial velocities and/or proper motion. 
+A comprehensive study of the kinematic
+properties of all the massive stars surrounding R136 could shed light
+on whether some can be unequivocally identified as merger products,
+but also on the initial conditions for the cluster dynamics
+\citep[e.g.,][]{oh:16}, and whether it formed via a monolithic collapse, or
+as a (potentially ongoing) merger of several sub-structures \citep[e.g.,][]{sabbi:12}.
+
+Also \cite{fujii:11} suggest that
 early in the evolution of a cluster, dynamical interactions form an extremely
@@
 upper limit to the timescale to form the ``bully binary'' in
-R136. Such a binary could be R145 according to \cite{fujii:11}, and it might be an
-ideal observational candidate for a dynamically formed progenitor system of
+R136. If identified, such a binary % could be R145 according to \cite{fujii:11}, and it
+might be an ideal observational candidate for a dynamically formed progenitor system of
 a binary black-hole, provided that stars this massive can avoid a
@@
 
-Also \citet{banerjee:12} suggested, based on their N-body simulations,
-that VFTS682 was ejected from R136. In their
-model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS682 and the massive members of R136. They
-demonstrated that the cluster potential does not significantly change
-the velocity of the star after the ejection.
-% While it is not possible to
-% robustly exclude that VFTS682 is itself a merger, its spectrum does not show any of the obvious signatures
-% like fast rotation, which however could have already slowed down
-% because of the wind angular momentum losses during and after the
-% merger.
-To eject such a massive object, the cluster is
-expected to have produced
-a large number of massive runaways. Indeed, several %relatively
-isolated massive stars are observed in the region \citep[][]{evans:10,lennon:18}, some with known
-large radial velocities and/or proper motion. % (see
-% \Figref{fig:pm_polar} and Sana et al., in prep.).
-A comprehensive study of the kinematic
-properties of all the massive stars surrounding R136 could shed light
-on whether some can be unequivocally identified as merger products,
-but also on the initial conditions for the cluster dynamics
-\citep[e.g.,][]{oh:16}, and whether it formed via a monolithic collapse, or
-as a (potentially ongoing) merger of several sub-structures \citep[e.g.,][]{sabbi:12}.
-
 VFTS682 is potentially the most massive runaway known to date, and its ejection
@@
 
- \bibliographystyle{apj}
- \bibliography{bibliography/vfts682,my_bib}
+\bibliographystyle{mnras}
+\bibliography{bibliography/vfts682}
+
+% Don't change these lines
+\bsp	% typesetting comment
+\label{lastpage}
+
+ % \bibliographystyle{apj}
+ % \bibliography{bibliography/vfts682,my_bib}
 
@@
     Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
-  $^{11}$ {Centro de Astrobiología, CSIC-INTA, Carretera de Torrejón a Ajalvir km-4, E-28850 Torrejón de Ardoz, Madrid, Spain}\\
+  $^{11}$ {Centro de Astrobiología, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}\\
   $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
   $^{13}$ {Department of Physics, University of Oxford, Keble Road,
-    Oxford OX1 3RH, UK} \\
-  $^{14}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
+    Oxford OX13RH, UK} \\
+  $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}\\
 
- \begin{acknowledgements}
-   \small
+ % \begin{acknowledgements}
+{   \small
    We are grateful to J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
@@
 participating in the {\it Gaia} Multilateral Agreement. 
-\end{acknowledgements}
+% \end{acknowledgements}
+}
 

### 2018-08-22 VFTS682_APJL.tex
MNRAS Letter ready for submission

@@
   Telescope} astrometry. We derive a projected velocity relative to
-the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval).  The direction and magnitude of
-the proper motion are consistent with VFTS682 being a slow runaway
+the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval). Although
+the error bars are substantial, two completely independent measures
+suggest that VFTS682 might be a slow runaway
 ejected from the central cluster, given its distance and the inferred
-age. The \todo{rephrase positive, see Fabian and Roeland's comments}
-
-However, the error bars are substantial and the proper motion measurements alone cannot rule out the counter hypothesis with high confidence.
-
-If future data confirm the runaway nature, this would make VFTS682
-the most massive runaway star known to date. This hypothesis is
-further supported by a variety of
+age. This hypothesis is further supported by a variety of
 circumstantial clues. The central cluster is known to harbor several
-other stars more massive than $150\,M_\odot$ of similar in spectral type.
-The very massive O stars VFTS16 (and possibly VFTS72), which has recently been proposed as a dynamically ejected runaway, provides direct evidence that R136 is indeed capable of ejecting some of its most massive members, consistent with the predictions of numerical simulations.
+other stars more massive than $150\,M_\odot$ of similar in spectral
+type. The very massive O stars VFTS16 (and possibly VFTS72), which has
+recently been proposed as a dynamically ejected runaway, provides
+direct evidence that R136 is indeed capable of ejecting some of its
+most massive members, consistent with the predictions of numerical
+simulations. If future data confirm the runaway nature, this would make VFTS682
+the most massive runaway star known to date. 
 \end{abstract}
@@
                   %        &$v_\mathrm{2D, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
+      \hline
+      \multicolumn{5}{l}{\emph{Weighted average relative proper motion
+      for VFTS682}}\\
+      \hline
+      &$\delta\mu_\mathrm{RA, avg}$  &[$\masyr$] & $0.08\pm0.07$ & (5) \\
+      &$\delta\mu_\mathrm{DEC, avg}$  &[$\masyr$] & $0.14\pm0.07$ &  (5) \\
+       &$\delta\mu_\mathrm{avg}$  &[$\masyr$] & $0.16\pm0.07$ &  (5) \\
+      &$v_\mathrm{2D, avg}$  &[$\kms$] & $38\pm17$ & (5)\\
+      % &$\theta_\mathrm{avg}$  &[degrees] &  \todo{TBD} $_{-}^{+}$   & (5)\\  
       \hline
@@
 projected velocity with respect to R136 of
-$\sim38\kms$ (averaging the \emph{Gaia} DR2 and HST
-results), but compatible with zero within $2\sigma$. Due to the large error bars, this result will need
+$38\pm17\kms$ (averaging the \emph{Gaia} DR2 and HST
+results). Due to the large error bars, this result will need
 to be revisited with future astrometric data. % to be revisited as updated

### 2018-08-22 MNRASformat/VFTS682_MNRASL.tex
MNRAS Letter ready for submission

@@
 \newauthor{A.~de~Koter$^{1}$,
-      N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~Schneider$^{13}$, J.~S.~Vink$^{14}$}}
+      N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~N.~Schneider$^{13}$, J.~S.~Vink$^{14}$}}
 
@@
 circumstantial clues. The central cluster is known to harbor several
-other stars more massive than $150\,M_\odot$ of similar in spectral
+other stars more massive than $150\,M_\odot$ of similar spectral
 type. The very massive O stars VFTS16 (and possibly VFTS72), which has
@@
 \emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
-$0.7\pm0.3$\,Myr ago. If the
+$0.7\pm0.3$\,Myr ago, which is compatible with the evolutionary age of
+the star. If the
 cluster age \citep[$\lesssim2$\,Myr,][]{crowther:10, sabbi:12} is

### 2018-08-22 MNRASformat/VFTS682_MNRASL.tex
ready to submit to MNRASL

@@
 in 30 Doradus at a projected distance of 29\,pc from the central
-cluster R136. % Its apparent isolation
-The absence of other massive stars in its immediate vicinity
-led to two intriguing hypotheses: % posed in earlier work:
-either it formed in relative isolation % through a new mode of star formation
-or it was ejected dynamically from the central cluster. 
-In this context, 
-we investigate the kinematics of
-VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space
-  Telescope} astrometry. We derive a projected velocity relative to
+cluster R136. 
+Its apparent isolation led to two intriguing hypotheses: 
+either it formed in relative isolation 
+or it was ejected dynamically from the cluster. 
+We investigate the kinematics of
+VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space Telescope} astrometry. We derive a projected velocity relative to
 the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval). Although
-the error bars are substantial, two completely independent measures
-suggest that VFTS682 might be a slow runaway
-ejected from the central cluster, given its distance and the inferred
-age. This hypothesis is further supported by a variety of
-circumstantial clues. The central cluster is known to harbor several
+the error bars are substantial, two independent measures
+suggest that VFTS682 might be a runaway
+ejected from the central cluster. This hypothesis is further supported by a variety of
+circumstantial clues. The central cluster is known to harbor 
 other stars more massive than $150\,M_\odot$ of similar spectral
-type. The very massive O stars VFTS16 (and possibly VFTS72), which has
-recently been proposed as a dynamically ejected runaway, provides
-direct evidence that R136 is indeed capable of ejecting some of its
+type. The very massive O stars VFTS16 (and possibly VFTS72), 
+recently 
+proposed as a dynamically ejected runaway, provides
+direct evidence that R136 is
+capable of ejecting some of its
 most massive members, consistent with the predictions of numerical
@@
 {   \small
-   We are grateful to J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
+   We are grateful to P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
    % and C.\,J.\,Evans

### 2018-08-23 VFTS682_APJL.tex
Chris' last round

@@
 core for VFTS16 and VFTS72 \citep[][]{lennon:18}. We note that
-VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
-large peculiar radial velocity and therefore accurate distances along
-the line of sight are
-needed to contrain the flight direction in three dimensions.
+% VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
+% large peculiar radial velocity and therefore
+accurate
+distances along the line of sight are needed to contrain the flight direction in three dimensions.
 
@@
 
-Based on our results, we consider likely that VFTS682 is the most massive
-(slow) runaway known to date, with a two-dimensional
+Based on our results, we consider it plausible that VFTS682 is the most massive
+runaway known to date, with a two-dimensional
 projected velocity with respect to R136 of
@@
 
-VFTS682 is potentially the most massive runaway known to date, and its ejection
+VFTS682 is potentially the most massive runaway known to date, and its
+possible ejection
 from the cluster R136 likely implies that it is only the ``tip of the
-iceberg'' of possibly extremely massive runaways in the
+iceberg'' of massive runaways in the
 region. Studies of this population, enabled by recent HST and \emph{Gaia} observations will put constraints on the evolution

### 2018-08-23 MNRASformat/VFTS682_MNRASL.tex
imant's comments

@@
 
-One of the most extreme examples that has been considered in this debate is the very massive star VFTS682  \citep[][]{bestenlehner:11, bressert:12}.
-This star is located in the field of the 30 Doradus region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}.
+One of the most extreme examples that has been considered in this
+debate is the very massive star VFTS682  \citep[][]{bestenlehner:11,
+  bressert:12}. This star is located in the field of the 30 Doradus
+(30Dor) region in the Large Magellanic Cloud (LMC) and was studied as part of the multi-epoch spectroscopic VLT-FLAMES Tarantula Survey \citep[VFTS,][]{evans:11}.
 It is a hydrogen-rich Wolf-Rayet star of spectral type WNh5.
@@
 
-The WNh5 star VFTS682, located at right ascension (RA)
-05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
-(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} (J2000), was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
-\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
-parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
-luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
-the most luminous stars in the region. The absence of periodic radial
-velocity (RV)
-variations suggests that the star is unlikely to have close binary
-companions \citep[][]{bestenlehner:11}, unless the orbital inclination is
-very high. Bayesian fits of the stellar
-parameters against evolutionary tracks \citep{brott:11, kohler:15}
-using the BONNSAI code \citep{schneider:14,schneider:17} provide
-estimates for the age, present mass and initial mass, % indicating that
-% this star among the most massive stars known,
-% see
-\Tabref{tab:star_param}. % for an overview of the parameters.
-
 \begin{table}
@@
       \hline
-         &$\mathrm{RA}_\mathrm{VFTS682}$&[degrees] & \phantom{-0}84.7314 %$ \pm 0.036\,\mathrm{mas} $ 
+         &$\mathrm{RA}_\mathrm{VFTS682}$&[degrees] & \phantom{-0}84.73136339876477 
                      & (1) \\        
-               &$\mathrm{DEC}_\mathrm{VFTS682}$&[degrees] & \phantom{0}-69.0741%$ \pm 0.048\,\mathrm{mas} $
+               &$\mathrm{DEC}_\mathrm{VFTS682}$&[degrees] &
+                                                            \phantom{0}-69.07411071794998
                      & (1)  \\
 
-                        &$\mathrm{RA}_\mathrm{R136}$&[degrees] & \phantom{-0}84.6767
-                     &  SIMBAD  \\        
-               &$     \mathrm{DEC}_\mathrm{R136}$&[degrees] &  \phantom{0}-69.1009
-                     &  SIMBAD \\       
+                        &$\mathrm{RA}_\mathrm{R136}$&[degrees] & \phantom{00}84.6750
+                     &  (2) \\        
+               &$     \mathrm{DEC}_\mathrm{R136}$&[degrees] &  \phantom{0}-69.1006
+                     &  (2) \\       
         &$      \delta\mathrm{RA}$  &[mas] & \phantom{-00}0.0547                      
-        &  (2, 5)
+        &  (3, 5)
   \\        
                &$     \delta\mathrm{DEC}$  &[mas] & \phantom{-00}0.0268 
-                     &  (2, 5) \\  
+                     &  (3, 5) \\  
                        &$  d_\parallel$  & [arcsec]
                          & 119.4
-                                 &  SIMBAD \\
-      &$L_\parallel$ & [pc] & 29 & (2) \\
+                                 &  (3) \\
+      &$L_\parallel$ & [pc] & 29 & (3) \\
 
@@
       \hline
-          &$\mu_\mathrm{RA}$&[$\masyr$] & $1.843\pm 0.070$
+          &$\mu_\mathrm{RA}$&[$\masyr$] & $1.84\pm 0.07$
                      & (1) \\        
-               &$\mu_\mathrm{DEC}$&[$\masyr$] & $0.786\pm 0.080$
+               &$\mu_\mathrm{DEC}$&[$\masyr$] & $0.79\pm 0.08$
                      &  (1) \\        
@@
        &$\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$&[$\masyr$] & $1.74\pm0.01$
-                        & (3) \\
+                        & (4) \\
       &$\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$&[$\masyr$]
-                & $0.70\pm0.02$ &  (3)\\
+                & $0.70\pm0.02$ &  (4)\\
 \hline
@@
       \hline
-      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & $0.10\pm0.08$ & (1,5) \\
-      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & $0.086\pm0.100$ &  (1,5) \\
-      &$\delta\mu_{Gaia}$  &[$\masyr$] & $0.13\pm0.09$ &  (1,5) \\
-      &$v_\mathrm{2D}$  &[$\kms$] & $32\pm21$ & (1,5)\\  
-      &$\theta_{Gaia}$  &[degrees] &  $14_{-31}^{+36}$  & (1,5)\\  
+      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & $0.10\pm0.08$ & (1,6) \\
+      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & $0.08\pm0.10$ &  (1,6) \\
+      &$\delta\mu_{Gaia}$  &[$\masyr$] & $0.13\pm0.09$ &  (1,6) \\
+      &$v_\mathrm{2D}$  &[$\kms$] & $32\pm21$ & (1,6)\\  
+      &$\theta_{Gaia}$  &[degrees] &  $14_{-31}^{+36}$  & (1,6)\\  
       % &$v_\mathrm{2D, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
@@
             \hline
-      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & $0.01\pm0.13$ & (4) \\
-      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (4) \\
-       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & $0.20\pm0.10$ &  (4) \\
-                  &$v_\mathrm{2D, HST}$  &[$\kms$] & $47\pm24$ & (4)\\  
-                      &$\theta_\mathrm{HST}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,5)\\  
+      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & $0.02\pm0.10$ & (5) \\
+      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & $0.19\pm0.09$ &  (5) \\
+       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & $0.19\pm0.09$ &  (5) \\
+                  &$v_\mathrm{2D, HST}$  &[$\kms$] & $45\pm21$ & (5)\\  
+                      &$\theta_\mathrm{HST}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,6)\\  
                   %        &$v_\mathrm{2D, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
@@
       \hline
-      &$\delta\mu_\mathrm{RA, avg}$  &[$\masyr$] & $0.08\pm0.07$ & (5) \\
-      &$\delta\mu_\mathrm{DEC, avg}$  &[$\masyr$] & $0.14\pm0.07$ &  (5) \\
-       &$\delta\mu_\mathrm{avg}$  &[$\masyr$] & $0.16\pm0.07$ &  (5) \\
-      &$v_\mathrm{2D, avg}$  &[$\kms$] & $38\pm17$ & (5)\\
+      &$\delta\mu_\mathrm{RA, avg}$  &[$\masyr$] & $0.08\pm0.07$ & (6) \\
+      &$\delta\mu_\mathrm{DEC, avg}$  &[$\masyr$] & $0.14\pm0.07$ &  (6) \\
+       &$\delta\mu_\mathrm{avg}$  &[$\masyr$] & $0.16\pm0.07$ &  (6) \\
+      &$v_\mathrm{2D, avg}$  &[$\kms$] & $38\pm17$ & (6)\\
       % &$\theta_\mathrm{avg}$  &[degrees] &  \todo{TBD} $_{-}^{+}$   & (5)\\  
@@
       \hline
-      &$v_\mathrm{2D}$  &[$\kms$] & $29\pm 6$ & (2) \\  
+      &$v_\mathrm{2D}$  &[$\kms$] & $29\pm 6$ & (3) \\  
       &$\theta$  &[degrees] &  $\sim0$  & \\ 
@@
       on $\theta_{Gaia}$.
-      (1) \cite{brown:18},
-      (2)~\cite{bestenlehner:11},
-      (3) \cite{lennon:18}, 
-      (4) \cite{platais:18} and
-      (5) {\color{blue}this study}.
+      (1)~\cite{brown:18},
+      (2)~\cite{henault-brunet:12},
+      (3)~\cite{bestenlehner:11},
+      (4)~\cite{lennon:18}, 
+      (5)~\cite{platais:18} and
+      (6)~{\color{blue}this study}.
     %The error on the RA and DEC positions, are of order $\sim$$10^{-2}\masyr$ in \emph{Gaia} DR2. %REPETITION: %  We adopt the peculiar
@@
 \end{table}
+
+The WNh5 star VFTS682, located at right ascension (RA)
+05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
+(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} (J2000), was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
+\citet{bestenlehner:11}  analyzed the spectra to infer the stellar
+parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
+luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
+the most luminous stars in the region. The absence of periodic radial
+velocity (RV)
+variations suggests that the star is unlikely to have close binary
+companions \citep[][]{bestenlehner:11}, unless the orbital inclination is
+very high. Bayesian fits of the stellar
+parameters against evolutionary tracks \citep{brott:11, kohler:15}
+using the BONNSAI code \citep{schneider:14,schneider:17} provide
+estimates for the age, present mass and initial mass, % indicating that
+% this star among the most massive stars known,
+% see
+\Tabref{tab:star_param}. % for an overview of the parameters.
 
@@
 selected bright ($G<17$) stars within 0.05 degrees of R136 and exclude
-sources with proper motion errors greater than $0.01\masyr$ in both
+sources with proper motion errors greater than $0.1\masyr$ in both
 coordinates (see their  Sect.~2.1).  Using this definition of the
@@
 
-\subsection{HST astrometry for VFTS682}
+\subsection{HST (WFC3/UVIS) astrometry for VFTS682}
 
-The 30 Doradus region was targeted by a two-epoch photometric campaign with HST providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}).
-% Visually, the star is clearly isolated and does not appear to have any close neighbor.
-This by itself gives further confidence in ability of \emph{Gaia} to measure the proper motion reliably without being affected by crowding.
+The 30 Dor region was targeted by a two-epoch photometric campaign
+with HST providing observations in the F775W filter in October 2011
+and October 2014 (GO-12499; P.I.: D.~J.~Lennon). 
 \citet{platais:15, platais:18} analyzed the HST data to determine the
@@
 excluded from the analysis. The high extinction around VFTS682
-prevents this issue. % should be one of the brightest
-% stars in the region, however, the high extinction makes it faint enough such that saturation effects are not an issue.
-VFTS682 is relatively slow and is therefore not discussed in
-\citealt{platais:18}.  However, their measurements provide useful
+makes it redder and fainter \citep[$V=16.08$,
+$B-V=0.58$,][]{evans:11}, hence it has reasonably accurate HST astrometry
+with the WFC3/UVIS camera. This star did not pass a full set of stringent
+conditions to be considered as a candidate runaway \citep[][]{platais:18}.
+In retrospect, VFTS682 may have been included in the list of likely
+OB runaway stars, and it is identified
+with the ID source 330375 in their catalog. 
+Therefore, their measurements provide useful
 complementary estimate of the proper motion of VFTS682 independent from the \emph{Gaia} data.
@@
 velocities) of $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$ and
-$\delta \mu_{HST}=0.20\pm 0.10\masyr (47 \pm 24\kms )$. Both values
+$\delta \mu_{HST}=0.19\pm 0.09\masyr (45 \pm 21/\kms )$. Both values
 are consistent with each other, but also with no relative motion

### 2018-08-24 MNRASformat/VFTS682_MNRASL.tex
new pm position angle
to be submitted after Stephen's comments

@@
   \newauthor{E.~Laplace$^{1}$, J.~M.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
-    V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{1,8,9}$,
+    V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{8,9,1}$,
   }
@@
 the error bars are substantial, two independent measures
-suggest that VFTS682 might be a runaway
+suggest that VFTS682 is a runaway
 ejected from the central cluster. This hypothesis is further supported by a variety of
@@
 other stars more massive than $150\,M_\odot$ of similar spectral
-type. The very massive O stars VFTS16 (and possibly VFTS72), 
-recently 
-proposed as a dynamically ejected runaway, provides
-direct evidence that R136 is
-capable of ejecting some of its
+type. Recent analysis of the astrometry for the very massive O star VFTS16 (and possibly VFTS72), provides
+direct evidence that R136 can eject some of its
 most massive members, consistent with the predictions of numerical
@@
 In has been proposed that most, if not all, stars form in clusters
-\citep{lada:03}. In this picture, field stars are primarily the result
+\citep[and references therein]{lada:03}. In this picture, field stars are primarily the result
 of the dissolution of dense groups.  However, a significant population
@@
 In particular, a remarkable similarity exist between the
-spectrum of VFTS682 and R136a3 \citep{rubio-diez:17}.%  for which \citet{crowther:16} report a
+spectra of VFTS682 and R136a3 \citep{rubio-diez:17}.%  for which \citet{crowther:16} report a
 % current mass estimate of $180^{+30}_{-30}\Msun$.
@@
 expected \citep[e.g.][]{fujii:11, banerjee:12}. The capability of
-young cluster of ejecting a large number of (very) massive stars is supported by
+young clusters of ejecting a large number of (very) massive stars is supported by
 the recent findings of proper motion studies \citep[e.g.,][]{lennon:18, drew:18}.
@@
                   &$v_\mathrm{2D, HST}$  &[$\kms$] & $45\pm21$ & (5)\\  
-                      &$\theta_\mathrm{HST}$  &[degrees] &   $-35_{-51}^{+24}$   & (1,6)\\  
+                      &$\theta_\mathrm{HST}$  &[degrees] &   $-30_{-51}^{+24}$   & (1,6)\\  
                   %        &$v_\mathrm{2D, \parallel}$  &[$\kms$] & $\SdM{xxx}\pm\SdM{xxx}$ & (1,5)\\  
@@
       $\theta$ is defined such that $\theta=0$ for motions pointing
-      away radially from R136, we neglect the errors on
+      away radially from R136, we neglect the error bars on
       $\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$ and $\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$ to determine the uncertainty
@@
 % see
-\Tabref{tab:star_param}. % for an overview of the parameters.
+as given in \Tabref{tab:star_param}. % for an overview of the parameters.
 
-\cite{townsley:06} did not find
-a bright X-ray point source at the position of VFTS682 in their
-\emph{Chandra} survey of the region. The X-ray luminosity of VFTS682
+VFTS682 is not a bright X-ray point source. It was not detected in the
+\emph{Chandra} survey of \cite{townsley:06}, and shows a few counts in
+the deeper survey of \cite{townsley:14}.
+% \cite{townsley:06} did not find
+% a bright X-ray point source at the position of VFTS682 in their
+% \emph{Chandra} survey of the region.
+The X-ray luminosity of VFTS682
 is significantly lower than known massive binaries in the region, which suggests the absence of
@@
 even for extreme mass ratios, given the extremely large mass of
-VFTS682. This star is also relatively isolated in the (near)
+VFTS682.
+
+This star is also relatively isolated in the (near)
 infrared. The nearest bright (near-)infrared sources detected by
@@
 possible effect of clumping.  We therefore caution
-for over-interpreting the radial velocities estimates. % , but we will summarize the estimates that have been published. 
+against over-interpreting the radial velocities estimates. % , but we will summarize the estimates that have been published. 
 \citet{bestenlehner:11}  estimate a RV of
@@
     panels, the dark blue histograms contain 317 
-    stars with error smaller than $0.1\,\mathrm{mas \
+    stars with error bars smaller than $0.1\,\mathrm{mas \
       yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
-    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
+    lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
       yr^{-1}}$. }
@@
   \centering
-  \includegraphics[width=\textwidth]{./figures/Figure2_paper_vfts682_hst_corr_nor136err}  
-  \caption{The thick yellow arrows indicate proper motions relative to
-    R136 (blue
-    circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The prolongations in the opposite
-    direction are proportional to the apparent ages of the stars, and
-    the thin lines illustrate the error cone on the potential
-    origin. The green thick arrow
-    indicates the HST proper motion of VFTS682, with the relative
-    error cone and prolongation proportional to its apparent
-    age. The two blue circles indicate the regions of radii 0.01 and 0.05
+  \includegraphics[width=\textwidth]{./figures/fig2}  
+  % \caption{The thick yellow arrows indicate proper motions relative to
+  %   R136 (blue
+  %   circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The extensions in the opposite
+  %   direction are proportional to the apparent ages of the stars, and
+  %   the thin lines illustrate the error cone on the potential
+  %   origin. The green thick arrow
+  %   indicates the HST proper motion of VFTS682, with the relative
+  %   error cone and extensions proportional to its apparent
+  %   age. The two blue circles indicate the regions of radii 0.01 and 0.05
+  %   degrees around the core of R136.}
+  \caption{The thick purple arrow shows proper motion
+    relative to R136 for VFTS682 averaging the \emph{Gaia} DR2 and HST
+  astrometry, multiplied by 0.4\,Myr. The extension in the opposite
+  direction is proportional to the apparent age of the star, and the
+  thin lines illustrate the error cone on the potential origin. The
+  yellow arrows show the \emph{Gaia} DR2 results alone, and the green
+  arrow the HST results alone. The two blue circles indicate the regions of radii 0.01 and 0.05
    degrees around the core of R136.}  
-
   \label{fig:main}
@@
 selected bright ($G<17$) stars within 0.05 degrees of R136 and exclude
-sources with proper motion errors greater than $0.1\masyr$ in both
+sources with proper motion error bars greater than $0.1\masyr$ in both
 coordinates (see their  Sect.~2.1).  Using this definition of the
@@
 an outlier compared to other OB-type and Wolf-Rayet stars, the agreement
-suggests that the star is indeed a ``slow-runaway'' as suggested by
+suggests that the star is indeed a ``slow runaway'' as suggested by
 \cite{bestenlehner:11}.
@@
 $\delta \mu_{HST}=0.19\pm 0.09\masyr (45 \pm 21/\kms )$. Both values
-are consistent with each other, but also with no relative motion
-within $2\sigma$. The weighted average of these two independent
+are consistent with each other, but also with no motion relative to R136
+within $2\sigma$. The average (weighted with $1/\sigma^2$) of these two independent
 measurements is $\delta \mu_\mathrm{avg}=0.16\pm0.07\masyr
@@
 Figure~\ref{fig:main} shows the projected motion of VFTS682 relative to R136 on the
-sky. We also show VFTS16 and VFTS72 for completeness
+sky. We also show VFTS16 and VFTS72 %for completeness
 \citep[see][]{lennon:18}. The thick yellow arrows are proportional to
@@
 on the direction of motion is illustrated by the corresponding
-prolongation in the direction opposite to the motion, and we also show
+extension in the direction opposite to the motion, and we also show
 the most likely origin of the stars accounting for their apparent age.
@@
 \citep[$\sim$3\,Myr,][]{brott:11,kohler:15, zapartas:17}, the ejection of VFTS682
-from the disruption of a binary is excluded. %  since the region is too young for stars
+from the disruption of a binary by a supernova is excluded. %  since the region is too young for stars
 % to have experienced core-collapse already. 
@@
 If the star were ejected dynamically, its isolation makes it an ideal target to constrain the stellar physics of
-stars with masses well above $\sim$$100\,M_\odot$ while avoiding
+stars with masses well above $\sim$$100\,M_\odot$ in the inner
+cluster, while avoiding
 crowding issues. Moreover, 
@@
 iceberg'' of possibly extremely massive runaways in the
-region. Studies of this population, enabled by recent HST and \emph{Gaia} observations will put constraints on the evolution
+region. Studies of this population, enabled by recent and future HST and \emph{Gaia} observations will put constraints on the evolution
 of these extreme stars, together with the formation and evolution of
@@
     Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
-  $^{11}$ {Centro de Astrobiología, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}\\
+  $^{11}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}\\
   $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\

### 2018-08-24 MNRASformat/VFTS682_MNRASL.tex
re-submitting 5 pages - no acknowledgements/affiliations/smallfig

@@
 \begin{keywords}
-  stars: kinematics, stars: runaways, stars: individual: VFTS682
+  stars: astrometry, stars: kinematics and dynamics, stars: individual: VFTS682
 \end{keywords}
@@
 How massive stars form is one of the major longstanding questions in astrophysics
-\citep[e.g.,][]{zinnecker:07}. Improving our understanding of massive star formation, and its
-possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies, but also for understanding the transients that mark their death and the compact remnants they leave behind.
+\citep[e.g.,][]{zinnecker:07}. %  Improving our understanding of massive star formation, and its
+% possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies, but also for understanding the transients that mark their death and the compact remnants they leave behind.
 %
@@
 process. Important progress has been made on the theoretical side,
-\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally expensive and remain challenging.
+\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations % of this multi-scale and multi-physics problem
+% are computationally expensive and
+remain challenging.
 
@@
 % masses, is the most massive runaway star known to date. 
-
+\vspace*{-20pt}
 \section{Observations}
@@
 unusual for Wolf-Rayet stars and more typical for Luminous Blue
-Variable (LBV) stars \citep{parker:93, udalski:08, bestenlehner:11}. It also
+Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
 shows a mid-infrared excess \citep{gruendl:09}.
@@
     vertical band) if it were indeed
-    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to $\kms$
-    assuming a distance of 50\,kpc. \emph{Right panel}:  distribution of
+    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. %The top axis shows the conversion to $\kms$
+    % assuming a distance of 50\,kpc.
+    \emph{Right panel}:  distribution of
     angles between the relative proper motion direction and the radial
@@
 
-\begin{figure*}%[tbp]
+% \begin{figure*}%[tbp]
+%   \centering
+%   \includegraphics[width=\textwidth]{./figures/fig2}  
+%   % \caption{The thick yellow arrows indicate proper motions relative to
+%   %   R136 (blue
+%   %   circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The extensions in the opposite
+%   %   direction are proportional to the apparent ages of the stars, and
+%   %   the thin lines illustrate the error cone on the potential
+%   %   origin. The green thick arrow
+%   %   indicates the HST proper motion of VFTS682, with the relative
+%   %   error cone and extensions proportional to its apparent
+%   %   age. The two blue circles indicate the regions of radii 0.01 and 0.05
+%   %   degrees around the core of R136.}
+%   \caption{The thick purple arrow shows proper motion
+%     relative to R136 for VFTS682 averaging the \emph{Gaia} DR2 and HST
+%   astrometry, multiplied by 0.4\,Myr. The extension in the opposite
+%   direction is proportional to the apparent age of the star, and the
+%   thin lines illustrate the error cone on the potential origin. The
+%   yellow arrows show the \emph{Gaia} DR2 results alone, and the green
+%   arrow the HST results alone. The two blue circles indicate the regions of radii 0.01 and 0.05
+%    degrees around the core of R136.}  
+%   \label{fig:main}
+% \end{figure*}
+
+\begin{figure}%[tbp]
   \centering
-  \includegraphics[width=\textwidth]{./figures/fig2}  
+  \includegraphics[width=0.5\textwidth]{./figures/fig2}  
   % \caption{The thick yellow arrows indicate proper motions relative to
@@
   \label{fig:main}
-\end{figure*}
+\end{figure}
 
+\vspace*{-20pt}
 \subsection{ \emph{Gaia} astrometry for VFTS682\label{data:gaia}}
@@
 
+\vspace*{-20pt}
 \subsection{HST (WFC3/UVIS) astrometry for VFTS682}
@@
 % We list the relative proper motion in \Tabref{tab:vfts682}. 
-
+\vspace*{-20pt}
 \section{The kinematics of VFTS682}
@@
 counter hypothesis. 
-
+\vspace*{-20pt}
 \section{Discussion}
@@
 stars with initial masses larger than $\sim$$150\,M_\odot$ in the
-cluster R136, which is consistent with the detection
-of extremely massive stars in the core of the
-cluster.
+cluster R136, as it is observed.% which is consistent with the detection
+% of extremely massive stars in the core of the
+% cluster.
 
@@
 metallicity \citep[see also][]{langer:07}.
-
 Similarly, the final fate of VFTS682 could be either a
@@
 the central cluster itself.
-
+\vspace*{-20pt}
 \bibliographystyle{mnras}
@@
 % Don't change these lines
-\bsp	% typesetting comment
+%\bsp	% typesetting comment
 \label{lastpage}
@@
 
-\section*{Affiliations}
-\noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
-  $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
-    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
- $^{3}$ {
- %Center for Astrophysical Sciences, 
- Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-  $^{4}$ {Space Telescope Science Institute, 3700 San Martin Drive,
-    Baltimore, MD 21218, USA}\\
-  $^{5}${Department of Physics and Astronomy, Hicks Building,
-    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
-  $^{6}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}\\
-  $^{7}$ {National Research Council, Herzberg Astronomy \&
-    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
-    Canada}\\
-  $^{8}$ {School of Astronomy \& Space Science, University of the Chinese
-    Academy of Sciences, Beijing 100012, China}\\
-  $^{9}$ {National Astronomical Observatories, Chinese Academy of
-    Sciences, Beijing 100012, China}\\
-  $^{10}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
-    Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
-  $^{11}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}\\
-  $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
-  $^{13}$ {Department of Physics, University of Oxford, Keble Road,
-    Oxford OX13RH, UK} \\
-  $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}\\
+% \section*{Affiliations}
+% \noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
+%     Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
+%   $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
+%     E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
+%  $^{3}$ {
+%  %Center for Astrophysical Sciences, 
+%  Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
+%   $^{4}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+%     Baltimore, MD 21218, USA}\\
+%   $^{5}${Department of Physics and Astronomy, Hicks Building,
+%     Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
+%   $^{6}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}\\
+%   $^{7}$ {National Research Council, Herzberg Astronomy \&
+%     Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
+%     Canada}\\
+%   $^{8}$ {School of Astronomy \& Space Science, University of the Chinese
+%     Academy of Sciences, Beijing 100012, China}\\
+%   $^{9}$ {National Astronomical Observatories, Chinese Academy of
+%     Sciences, Beijing 100012, China}\\
+%   $^{10}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
+%     Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
+%   $^{11}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}\\
+%   $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
+%   $^{13}$ {Department of Physics, University of Oxford, Keble Road,
+%     Oxford OX13RH, UK} \\
+%   $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}\\
 
- % \begin{acknowledgements}
-{   \small
-   We are grateful to P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
-   % and C.\,J.\,Evans
-   for help and discussions. We also thank the HSTPROMO
-(High-resolution Space Telescope PROper MOtion) Collaboration for
-sharing and discussing ideas relevant to this work. SdM acknowledges
-the European Unions Horizon 2020 research and innovation programme
-from the European Research Council (ERC), Grant agreement No. 715063
-and VHB the NRC-Canada Plaskett Fellowship.
+%  % \begin{acknowledgements}
+% {   \small
+%    We are grateful to P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
+%    % and C.\,J.\,Evans
+%    for help and discussions. We also thank the HSTPROMO
+% (High-resolution Space Telescope PROper MOtion) Collaboration for
+% sharing and discussing ideas relevant to this work. SdM acknowledges
+% the European Unions Horizon 2020 research and innovation programme
+% from the European Research Council (ERC), Grant agreement No. 715063
+% and VHB the NRC-Canada Plaskett Fellowship.
 This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
 Funding for the DPAC has been provided by national institutions, in particular the institutions
-participating in the {\it Gaia} Multilateral Agreement. 
-% \end{acknowledgements}
-}
+% participating in the {\it Gaia} Multilateral Agreement. 
+% % \end{acknowledgements}
+% }
 

### 2018-08-28 MNRASformat/VFTS682_MNRASL.tex
re-added affiliations and minor re-wording to remove a page

@@
 
-\author[Renzo et al.]{M.~Renzo$^{1}$\thanks{Corresponding author:
-    M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
+\author[Renzo et al.]{M.~Renzo$^{1}$% \thanks{Corresponding author:
+    % M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
+  , S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
   R.~P.~van~der~Marel$^{3,4}$,
@@
 
-How very massive stars form is an intriguing open question in
-stellar astrophysics.  VFTS682 is among the most massive stars known,
+How very massive stars form is still an open question in
+astrophysics.  VFTS682 is among the most massive stars known,
 with an inferred initial mass of $\gtrsim$$150\,M_\odot$. It is located
@@
 other stars more massive than $150\,M_\odot$ of similar spectral
-type. Recent analysis of the astrometry for the very massive O star VFTS16 (and possibly VFTS72), provides
-direct evidence that R136 can eject some of its
-most massive members, consistent with the predictions of numerical
-simulations. If future data confirm the runaway nature, this would make VFTS682
+type and recent astrometric studies on VFTS16 (and possibly VFTS72)
+provide direct evidence that the cluster can eject some of its most
+massive members, in agreement with theoretical predictions.
+% . Recent analysis of the astrometry for the very massive O star VFTS16 (and possibly VFTS72), provides
+% direct evidence that R136 can eject some of its
+% most massive members, consistent with the predictions of numerical
+% simulations. 
+If future data confirm the runaway nature, this would make VFTS682
 the most massive runaway star known to date. 
@@
 \begin{keywords}
-  stars: astrometry, stars: kinematics and dynamics, stars: individual: VFTS682
+  stars: astrometry, kinematics and dynamics, individual: VFTS682
 \end{keywords}
@@
 
+\vspace*{-9pt}
 % A contribution to the debate on whether or not massive
@@
 % VFTS682.
-
 One of the most extreme examples that has been considered in this
@@
   \begin{center}
-    \caption{Stellar parameters of VFTS682. }
+    \caption{Stellar parameters of VFTS682.}
+    \vspace*{-15pt}
     \begin{tabular}{llc|c|c}
+      \hline\hline
+      &Parameter & Units & Value & Ref.\\     
        \hline
-      \hline
-      &Parameter & Units & Value & Ref.\\
-
-       \hline
-      %  \multicolumn{5}{l}{\emph{Selected stellar parameters VFTS 682}}
-      % \\
-      % \hline
      & present day mass  & $[M_\odot]$ & $137.8^{+27.5}_
@@
       \\
-      &age & [Myr] & $1.0\pm0.2$ & (1) \\
+      &age & $[\mathrm{Myr}]$ & $1.0\pm0.2$ & (1) \\
       & mass loss rate & $\log_{10}(\dot{M}/[M_\odot \ \mathrm{yr}^{-1}])$ & $-4.1\pm0.2$ & (2)\\
-      % &\multirow{2}{*}{surface helium mass fraction} &  & 0.45 & (2)\\
-      % & &  & 0.49 & (3)\\ %% we don't use this: if we believe it's a
-      % merger it's not relevant, if we don't we still don't use it
       \hline
@@
     \end{tabular}
-    % \tablecomments
-    % {
-      The quoted uncertainties are statistical, and do not include systematic
-      effects in the modeling.\\
+    \vspace*{-5pt}
+    {\tiny The quoted uncertainties are statistical, and do not include systematic
+      effects in the modeling.
       (1)~\cite{schneider:18}
-      (2)~\cite{bestenlehner:11}
-      % (3)~\cite{rubio-diez:17}
-    % }
+      (2)~\cite{bestenlehner:11}}
   \end{center}
@@
   \begin{center}
-    \caption{Kinematics of VFTS682. }
+    \caption{Kinematics of VFTS682.}
+    \vspace*{-9pt}
     \begin{tabular}{llc|c|c}
@@
        \hline
-
-% \hline
-%   %    &$\delta v_\mathrm{rad}$  &[$\kms$] & 30$\pm$20 & \cite{bestenlehner:11}\\
-
-%       \hline
-
     \end{tabular}
@@
 
-    Assuming a distance of 50\,kpc, the conversion from proper motion
-    to projected velocities is such that $1\masyr$ corresponds to $237\kms$.  The
-      correlation coefficient is indicated as $\rho$. The angle
-      $\theta$ is defined such that $\theta=0$ for motions pointing
-      away radially from R136, we neglect the error bars on
+    % Assuming a distance of 50\,kpc, the conversion from proper motion
+    % to projected velocities is such that $1\masyr$ corresponds to $237\kms$.
+    {\tiny The error on the RA and DEC positions, are of order
+      $\sim$$10^{-2}\masyr$ in \emph{Gaia}
+      DR2. $\rho(\mu_\mathrm{RA},\mu_\mathrm{DEC})$ is the
+      correlation coefficient. The position angle
+      $\theta$ is defined such that $\theta=0$ for radial motion
+      away from R136. We neglect the error bars on
       $\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$ and $\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$ to determine the uncertainty
@@
       (6)~{\color{blue}this study}.
-    %The error on the RA and DEC positions, are of order $\sim$$10^{-2}\masyr$ in \emph{Gaia} DR2. %REPETITION: %  We adopt the peculiar
-      % radial velocity calculated using the  NV $\lambda4944$ line, which is
-      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
-    % }
+    }
   \end{center}
@@
 
-%\citet{parker:93} reported B and V magnitudes roughly 0.5\,magnitude lower than found by \citet{evans:11}.  - Citing OGLE is enough
-
-\vspace*{10pt}% to move new sec. to newpage
-\begin{figure*}[tbp]
+\begin{figure}%[tbp]
   \centering
-  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
+  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}\\
+  \vspace*{-5pt}
   \includegraphics[width=0.35\textwidth]{figures/angle}
-  \caption{\emph{Left panel}: distribution of OB-type and Wolf-Rayet stars in proper
-    motion relative to R136 from \emph{Gaia} DR2. VFTS682 is not an outlier, but
-    its relative proper motion matches the expected value (green
-    vertical band) if it were indeed
-    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. %The top axis shows the conversion to $\kms$
-    % assuming a distance of 50\,kpc.
-    \emph{Right panel}:  distribution of
-    angles between the relative proper motion direction and the radial
-    direction from the center of the cluster to the star. The error bars from \emph{Gaia} for VFTS682 are large, but
-    the best value is in agreement with the hypothesis of dynamical
-    ejection (green vertical band). The peak at $\theta\simeq\pi/2$ is due to stars
-    belonging to NGC2060. In both
-    panels, the dark blue histograms contain 317 
+  % \caption{\emph{Top panel}: distribution of OB-type and Wolf-Rayet stars in proper
+  %   motion relative to R136 from \emph{Gaia} DR2. VFTS682 is not an outlier, but
+  %   its relative proper motion matches the expected value (green
+  %   vertical band) if it were indeed
+  %   ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. %The top axis shows the conversion to $\kms$
+  %   % assuming a distance of 50\,kpc.
+  %   \emph{Bottom panel}:  distribution of
+  %   angles between the relative proper motion direction and the radial
+  %   direction from the center of the cluster to the star. The error bars from \emph{Gaia} for VFTS682 are large, but
+  %   the best value is in agreement with the hypothesis of dynamical
+  %   ejection (green vertical band). The peak at $\theta\simeq\pi/2$ is due to stars
+  %   belonging to NGC2060. In both
+  %   panels, the dark blue histograms contain 317 
+  %   stars with error bars smaller than $0.1\,\mathrm{mas \
+  %     yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
+  %   lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
+  %   yr^{-1}}$. }
+  \caption{Distribution of OB-type and Wolf-Rayet stars in proper
+    motion relative to R136 (top panel) and proper motion angle
+    (bottom panel), from \emph{Gaia} DR2. Although VFTS682 is not an outlier, its relative
+    proper motion matches the value expected for an early dynamical ejection. In both
+    panels, the green vertical bands highlight these expectations, the dark blue histograms contain 317 
     stars with error bars smaller than $0.1\,\mathrm{mas \
@@
     lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
-      yr^{-1}}$. }
+    yr^{-1}}$. The peak at $\theta\simeq\pi/2$ in the bottom panel is due to stars
+  belonging to NGC2060. }
   \label{fig:dist}
-\end{figure*}
+\end{figure}
 
@@
 
- % \bibliographystyle{apj}
- % \bibliography{bibliography/vfts682,my_bib}
+\vspace*{-20pt}
+\section*{Affiliations}
+ \tiny
+\noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
+    Amsterdam, 1098 XH Amsterdam, The Netherlands} 
+  $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
+    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} 
+ $^{3}$ {
+ %Center for Astrophysical Sciences, 
+ Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}
+  $^{4}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+    Baltimore, MD 21218, USA}
+  $^{5}${Department of Physics and Astronomy, Hicks Building,
+    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
+  $^{6}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}
+  $^{7}$ {National Research Council, Herzberg Astronomy \&
+    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
+    Canada}
+  $^{8}$ {School of Astronomy \& Space Science, University of the Chinese
+    Academy of Sciences, Beijing 100012, China}
+  $^{9}$ {National Astronomical Observatories, Chinese Academy of
+    Sciences, Beijing 100012, China}
+  $^{10}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
+    Auf dem H\"ugel 71, 53121, Bonn, Germany}
+  $^{11}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}
+  $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}
+  $^{13}$ {Department of Physics, University of Oxford, Keble Road,
+    Oxford OX13RH, UK} 
+  $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}\\
 
-% \section*{Affiliations}
-% \noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
-%     Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
-%   $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
-%     E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
-%  $^{3}$ {
-%  %Center for Astrophysical Sciences, 
-%  Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
-%   $^{4}$ {Space Telescope Science Institute, 3700 San Martin Drive,
-%     Baltimore, MD 21218, USA}\\
-%   $^{5}${Department of Physics and Astronomy, Hicks Building,
-%     Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
-%   $^{6}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}\\
-%   $^{7}$ {National Research Council, Herzberg Astronomy \&
-%     Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
-%     Canada}\\
-%   $^{8}$ {School of Astronomy \& Space Science, University of the Chinese
-%     Academy of Sciences, Beijing 100012, China}\\
-%   $^{9}$ {National Astronomical Observatories, Chinese Academy of
-%     Sciences, Beijing 100012, China}\\
-%   $^{10}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
-%     Auf dem H\"ugel 71, 53121, Bonn, Germany}\\
-%   $^{11}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}\\
-%   $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}\\
-%   $^{13}$ {Department of Physics, University of Oxford, Keble Road,
-%     Oxford OX13RH, UK} \\
-%   $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}\\
-
-%  % \begin{acknowledgements}
-% {   \small
-%    We are grateful to P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore, and S.\,Torres
-%    % and C.\,J.\,Evans
-%    for help and discussions. We also thank the HSTPROMO
-% (High-resolution Space Telescope PROper MOtion) Collaboration for
-% sharing and discussing ideas relevant to this work. SdM acknowledges
-% the European Unions Horizon 2020 research and innovation programme
-% from the European Research Council (ERC), Grant agreement No. 715063
-% and VHB the NRC-Canada Plaskett Fellowship.
+  % \begin{acknowledgements}
+  {   \tiny
+    We thank P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore,
+    and S.\,Torres, and the HSTPROMO Collaboration
+    for help and discussions. SdM acknowledges
+    the European Unions Horizon 2020 research and innovation programme
+    from the European Research Council (ERC), Grant agreement No. 715063
+    and VHB the NRC-Canada Plaskett Fellowship.
 This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
-Funding for the DPAC has been provided by national institutions, in particular the institutions
-% participating in the {\it Gaia} Multilateral Agreement. 
-% % \end{acknowledgements}
-% }
+  }
+% \end{acknowledgements}
 

### 2018-08-28 MNRASformat/VFTS682_MNRASL.tex
proof-read

@@
 cluster R136. 
-Its apparent isolation led to two intriguing hypotheses: 
+Its apparent isolation led to two hypotheses: 
 either it formed in relative isolation 
@@
 from HST proper motions  by \citet{platais:18}.   We discuss the
-implications of the hypothesis that VFTS682 is a slow runaway star
+implications of the hypothesis that VFTS682 is a runaway star
 ejected from R136.
@@
     \end{tabular}
-    % \tablecomments
-    % {
-
-    % Assuming a distance of 50\,kpc, the conversion from proper motion
-    % to projected velocities is such that $1\masyr$ corresponds to $237\kms$.
     {\tiny The error on the RA and DEC positions, are of order
       $\sim$$10^{-2}\masyr$ in \emph{Gaia}
-      DR2. $\rho(\mu_\mathrm{RA},\mu_\mathrm{DEC})$ is the
+      DR2. Assuming a distance of 50\,kpc, the conversion from proper motion
+    to projected velocities is such that $1\masyr$ corresponds to $237\kms$.
+$\rho(\mu_\mathrm{RA},\mu_\mathrm{DEC})$ is the
       correlation coefficient. The position angle
@@
 colliding winds. These would be expected in the presence of companions
-even for extreme mass ratios, given the extremely large mass of
+even for extreme mass ratios, given the large mass of
 VFTS682.
 
-This star is also relatively isolated in the (near)
-infrared. The nearest bright (near-)infrared sources detected by
+This star is also relatively isolated in the \mbox{(near-)infrared}. The nearest bright (near-)infrared sources detected by
 \emph{Spitzer} \citep{meixner:06} and resolved in the \emph{VISTA}
@@
 possibly inhomogeneous, optically thick wind  typical of emission line
-stars. \citet{bestenlehner:11} estimate a mass loss rate of
+stars.  We therefore caution
+against over-interpreting the radial velocities estimates. \citet{bestenlehner:11} estimate a mass loss rate of
 $\log_{10}(\dot{M}/[M_\odot \ \mathrm{yr}^{-1}])=-4.1\pm0.2$, not accounting for the
-possible effect of clumping.  We therefore caution
-against over-interpreting the radial velocities estimates. % , but we will summarize the estimates that have been published. 
-\citet{bestenlehner:11}  estimate a RV of
+possible effect of clumping.  % , but we will summarize the estimates that have been published. 
+They estimate a RV of
 $300\pm10\kms$ using the  N{\footnotesize V} $\lambda4944$ line, which
@@
   \caption{Distribution of OB-type and Wolf-Rayet stars in proper
-    motion relative to R136 (top panel) and proper motion angle
+    motion relative to R136 (top panel) and proper motion position angle
     (bottom panel), from \emph{Gaia} DR2. Although VFTS682 is not an outlier, its relative
@@
   %   degrees around the core of R136.}
-  \caption{The thick purple arrow shows proper motion
-    relative to R136 for VFTS682 averaging the \emph{Gaia} DR2 and HST
+  \caption{The thick purple arrow shows the proper motion
+    relative to R136 for VFTS682 from averaging the \emph{Gaia} DR2 and HST
   astrometry, multiplied by 0.4\,Myr. The extension in the opposite
@@
 
-\vspace*{-20pt}
+\vspace*{-15pt}
 \subsection{ \emph{Gaia} astrometry for VFTS682\label{data:gaia}}
@@
 \citep{gaia:16,brown:18}.   The number of visibility periods,
-i.e., groups of observations separated from other groups by a gap of at
+i.e., groups of observations separated from each other by at
 least four days, used in the astrometric solution is seventeen. The reported astrometric excess noise is zero.  These values
@@
 
-Gaia provides absolute proper motions.  To determine the relative
-proper motion with respect to R136, we follow  \citet{lennon:18} to
+Gaia provides absolute proper motions.  To determine the
+proper motion relative to R136, we follow  \citet{lennon:18} to
 define the motion of the local frame of reference using the average
@@
 $\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.   We also compute the total projected 2D velocity
-$v_\mathrm{2D}$ and the angle $\theta$ between the direction of motion and the vector
-connecting the center of R136 with the current position of VFTS682,
-i.e. such that $\theta = 0$ is the angle corresponding to perfectly
-radial motion away from R136. All kinematic quantities are provided in
+($v_\mathrm{2D}$) and the angle $\theta$ between the direction of motion and the vector
+connecting the center of R136 with the current position of VFTS682.
+% i.e. such that $\theta = 0$ is the angle corresponding to perfectly
+% radial motion away from R136.
+All kinematic quantities are provided in
 \Tabref{tab:vfts682}.
 
-\vspace*{-20pt}
+\vspace*{-15pt}
 \subsection{HST (WFC3/UVIS) astrometry for VFTS682}
@@
 an outlier compared to other OB-type and Wolf-Rayet stars, the agreement
-suggests that the star is indeed a ``slow runaway'' as suggested by
+suggests that the star is indeed a runaway as suggested by
 \cite{bestenlehner:11}.
@@
 velocities) of $\delta \mu_{Gaia}=0.13\pm 0.09\masyr (32\pm 21\kms)$ and
-$\delta \mu_{HST}=0.19\pm 0.09\masyr (45 \pm 21/\kms )$. Both values
+$\delta \mu_{HST}=0.19\pm 0.09\masyr (45 \pm 21\kms )$. Both values
 are consistent with each other, but also with no motion relative to R136
@@
 
-Figure~\ref{fig:main} shows the projected motion of VFTS682 relative to R136 on the
+Figure~\ref{fig:main} shows the motion of VFTS682 relative to R136
+projected on the
 sky. We also show VFTS16 and VFTS72 %for completeness
 \citep[see][]{lennon:18}. The thick yellow arrows are proportional to
-the relative proper motion from \emph{Gaia} DR2, and the thick green
-line illustrates the relative proper motion from HST. The error cone
+the relative proper motion from \emph{Gaia} DR2, the thick green
+line illustrates the relative proper motion from HST, and the thick
+purple line shows the averaged result. The error cone
 on the direction of motion is illustrated by the corresponding

### 2018-09-03 MNRASformat/VFTS682_MNRASL.tex
upgraded colors for fig

@@
 We investigate the kinematics of
-VFTS682 as obtained by \emph{Gaia} and multi-epoch \emph{Hubble Space Telescope} astrometry. We derive a projected velocity relative to
+VFTS682 as obtained by \emph{Gaia} and \emph{Hubble Space Telescope} astrometry. We derive a projected velocity relative to
 the cluster of $38\pm17\kms$ ($1\sigma$ confidence interval). Although
@@
 massive members, in agreement with theoretical predictions.
-% . Recent analysis of the astrometry for the very massive O star VFTS16 (and possibly VFTS72), provides
-% direct evidence that R136 can eject some of its
-% most massive members, consistent with the predictions of numerical
-% simulations. 
 If future data confirm the runaway nature, this would make VFTS682
@@
   %   degrees around the core of R136.}
-  \caption{The thick purple arrow shows the proper motion
+  \caption{The thick red arrow shows the proper motion
     relative to R136 for VFTS682 from averaging the \emph{Gaia} DR2 and HST
@@
   thin lines illustrate the error cone on the potential origin. The
-  yellow arrows show the \emph{Gaia} DR2 results alone, and the green
+  yellow arrows show the \emph{Gaia} DR2 results alone, and the orange
   arrow the HST results alone. The two blue circles indicate the regions of radii 0.01 and 0.05

### 2018-09-20 MNRASformat/referee_report_answer.txt
comments from Eva, Chris, and referee +  report draft

I have read the paper entitled "Space astrometry of the very massive ~ 150 M_sun candidate runaway star VFTS682” by Renzo, Mathieu et al. The paper focusses on the astrometry of the widely-discussed, massive, Wolf-Rayet star VFTS682 in the 30 Dor region. Although this particular very massive star (VMS) has been on focus for several years now since its identification by the VFTS survey and classification as a candidate runaway VMS from the central young massive cluster R136 (Bestenlehner et al. 2011), the present work is a significant leap forward in this candidature as it takes into account, for the first time, the latest proper motion measurements of this star from Gaia and HST. Given that VFTS682 is widely discussed in several pressing contexts such as the production mechanisms of runaway VMSs, the possibility of isolated massive star formation, and the estimation of VMSs’ ages, I think this paper is worth publishing as a letter. Overall, I find the paper well presented and concise. Before the acceptance, however, I recommend the following improvements in the paper.

We thank the referee for their careful reading of our manuscript, and for the suggestions below. The changes in the text are highlighted in green bold font.

1. Sec. 1,  paragraph 4: in the sentence beginning with "N-body simulations indicate that…”, it would be clearer to add the phrase “through dynamical ejections” after the word “expected”.

For the sake of conciseness, we have added "N- body simulations indicate that the *dynamical* ejection of...", which we believe achieves the clarification suggested by the referee with less words.

2. It is not immediately clear to me how the widths of the green vertical bands in Fig. 1 are estimated. It is important to explain this in the text - these bands are important to demonstrate the likeliness of VFTS682 for being an early dynamical runaway from R136. Perhaps similar bands can also be indicated for VFTS16 and VFTS72 in the panels of Fig. 1, for comparing their cases with that of VFTS682.

The green vertical band in the top panel is determined by the error bars on the apparent age of the star: the expected proper motion is the projected distance between the cluster and the star (~29pc) divided by the apparent age of the star (1+/-0.2Myr) which gives the v2D reported in Tab. 2. For the bottom panel we have assumed a width of 45 degrees centered in theta=0. We have added a clarification in the main text of Sec. 3 and a pointer to it in the figure caption.

3. I find Fig. 2 somewhat crowded with lines, especially around VFTS682. For improving the visibility, I suggest that either the central value can be back-traced with dashed lines (say) or the error cones can be shown with dashed lines (preserving the corresponding colours).

We have marked the error cones with dot-dashed lines in the updated version. We have also changed the color of the HST arrows to increase the contrast with the background figure.

4. Fig. 1 is in the top-bottom mode, not left-right. This should be rectified in the first sentence of Sec. 3.

Thank you for spotting this, we have corrected the mistake

5. Sec. 4, paragraph 3: in regard to the statement "N-body dynamics typically ejects the least massive star among those interacting…”, it is perhaps useful to remind the reader that the dynamical ejection fraction, nevertheless, overall increases with increasing stellar mass, as demonstrated in, e.g., Banerjee S. et al., 2012, ApJ, 746, 15 (see their Fig. 4), since the most massive entities would be most centrally segregated and hence interact most efficiently.

We have added a clarifying remark before the citation.

6. Sec. 4, paragraph 4: "In their model, they relied on (dynamically driven) stellar mergers to explain the high masses of VFTS682 and the massive members of R136.” Note that the dynamical merger scenario is not instrumental in explaining members (either residing in or dynamically ejected from R136) up to initial stellar masses of 150 M_sun, due to the adopted IMF in the models of Banerjee et al. (2012) that extends up to 150 M_sun. It is necessary to explain any single star (in-cluster or ejected) beyond 150 M_sun, given these models: see Banerjee S. et al., 2012, MNRAS, 426, 1416 and Crowther P. A. et al., 2010, MNRAS, 408, 731. (In case > 150 M_sun stars would also derive directly from massive-star-formation process, stellar mergers would not be instrumental also for such “super-canonical” stars - for such stars, the dynamical ejection fraction approaches unity as in Banerjee, S. et al., 2012, ApJ, 746, 15 and Fujii, M. S., Portegies Zwart S., 2011, Science, 334, 1380.) Therefore, I suggest rephrasing the quoted sentence accordingly.

We agree with the referee on this point. Our intent in emphasizing the importance of mergers in the simulations from Banerjee et al. is to implicitly pose the question of whether these VMS are mergers or not. While the mergers are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave them enough VMS stars both in the cluster and ejected around it suggest that if they had allowed for the formation of even more massive stars, then the efficiency of dynamically driven mergers together with initially VMS might possibly result in an over-prediction of VMS in the region. We think that this is an interesting point were observations can put constraints on the physics of cluster dynamics and stellar mergers and would prefer to keep the sentence as it is.

7. Sec. 4, second to last paragraph: I recommend also citing the recent work Belczynski, K. et al., 2016, A&A, 594, A97 in the context of the (im)possibility of avoiding pair-instability SN, as invoked there. According to the above work, pair-instability SN can easily be avoided at the LMC metallicity for at least up to 150 M_sun ZAMS (the initially approx. 150 M_sun VFTS682 can easily be a non-merger product; see point 6).

The reference suggested is not a primary one (the work of Belczynski et al. is based on the detailed simulations of Woosley 2017, http://adsabs.harvard.edu/abs/2017ApJ...836..244W), we have added a citation to the primary source instead.

### 2018-09-20 MNRASformat/VFTS682_MNRASL.tex
comments from Eva, Chris, and referee +  report draft

@@
 
-In has been proposed that most, if not all, stars form in clusters
+It has been proposed that most, if not all, stars form in clusters
 \citep[and references therein]{lada:03}. In this picture, field stars are primarily the result
-of the dissolution of dense groups.  However, a significant population
+of the dissolution of dense groups.  However, a \newtext{small but} significant population
 of massive stars exists in relative isolation, far from dense clusters
@@
 massive stars in the core of the R136 cluster \citep{dekoter:97,crowther:10, crowther:16}. 
-In particular, a remarkable similarity exist between the
+In particular, a remarkable similarity exists between the
 spectra of VFTS682 and R136a3 \citep{rubio-diez:17}.%  for which \citet{crowther:16} report a
@@
 the star formed in situ as an isolated massive star, or it was ejected from  R136. N-body simulations 
-indicate that the ejection of very massive stars like VFTS682 is
-expected \citep[e.g.][]{fujii:11, banerjee:12}. The capability of
-young clusters of ejecting a large number of (very) massive stars is supported by
+indicate that the \newtext{dynamical} ejection of very massive stars like VFTS682 is
+expected \citep[e.g.][]{fujii:11, banerjee:12}. The capability of a
+young cluster to eject a large number of (very) massive stars is supported by
 the recent findings of proper motion studies \citep[e.g.,][]{lennon:18, drew:18}.
@@
 \citet{bestenlehner:11}  analyzed the spectra to infer the stellar
-parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a
+parameters and measured an extinction of $A_V=4.45\pm0.12$, implying a
 luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
@@
 stars.  We therefore caution
-against over-interpreting the radial velocities estimates. \citet{bestenlehner:11} estimate a mass loss rate of
+against over-interpreting the \newtext{existing} radial velocities estimates. \citet{bestenlehner:11} estimate a mass loss rate of
 $\log_{10}(\dot{M}/[M_\odot \ \mathrm{yr}^{-1}])=-4.1\pm0.2$, not accounting for the
@@
 vicinity. This is consistent with the expectation if the star was not
-formed in situ. We will refrain from using the RV measurements in this
+formed in situ. \newtext{Given these issues, we refrain} from using the RV measurements in this
 work, and focus on the velocities on the plane of the sky.
@@
     (bottom panel), from \emph{Gaia} DR2. Although VFTS682 is not an outlier, its relative
-    proper motion matches the value expected for an early dynamical ejection. In both
-    panels, the green vertical bands highlight these expectations, the dark blue histograms contain 317 
+    proper motion matches the value expected for an early dynamical
+    ejection \newtext{(see \Secref{sec:results})}. In both
+    panels, the dark blue histograms contain 317 
     stars with error bars smaller than $0.1\,\mathrm{mas \
-      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
+      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc and the
     lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
@@
   \includegraphics[width=0.5\textwidth]{./figures/fig2}  
-  % \caption{The thick yellow arrows indicate proper motions relative to
-  %   R136 (blue
-  %   circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The extensions in the opposite
-  %   direction are proportional to the apparent ages of the stars, and
-  %   the thin lines illustrate the error cone on the potential
-  %   origin. The green thick arrow
-  %   indicates the HST proper motion of VFTS682, with the relative
-  %   error cone and extensions proportional to its apparent
-  %   age. The two blue circles indicate the regions of radii 0.01 and 0.05
-  %   degrees around the core of R136.}
   \caption{The thick red arrow shows the proper motion
@@
 with the ID source 330375 in their catalog. 
-Therefore, their measurements provide useful
-complementary estimate of the proper motion of VFTS682 independent from the \emph{Gaia} data.
+Therefore, their measurements provide \newtext{a} useful
+complementary estimate of the proper motion of VFTS682 \newtext{which is} independent from the \emph{Gaia} data.
 
@@
 % We list the relative proper motion in \Tabref{tab:vfts682}. 
-\vspace*{-20pt}
+\vspace*{-30pt}
 \section{The kinematics of VFTS682}
@@
 The black points in~\Figref{fig:dist} show the proper motion 
-(left panel) and the projected flight direction (right panel) relative
+(\newtext{top} panel) and the projected flight direction (\newtext{bottom} panel) relative
 to R136 of
@@
 cluster should %are expected to
-produce close to radial ejections, i.e.~$\theta\simeq0$. The green
-vertical bands highlight the expectations for these two quantities if VFTS682 were ejected from R136 early in its
-lifetime. For comparison, we also show the relative
+produce close to radial ejections, i.e.~$\theta\simeq0$. \newtext{The green
+vertical bands highlight the expectations for these two quantities. %  if VFTS682 were ejected from R136 early in its
+% lifetime.
+The width in the top panel is determined
+by the error bars on the star's apparent age, and we assume a width of
+$45$ degrees in the bottom panel.} For comparison, we also show the relative
 proper motion of VFTS16 and VFTS72 (gray points), and the distribution
@@
 VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
-large peculiar radial velocity and therefore accurate distances along
+a large peculiar radial velocity and therefore accurate distances along
 the line of sight are
@@
 
-Based on our results, we consider likely that VFTS682 is the most massive
+Based on our results, we consider that VFTS682 is
+\newtext{potentially} the most massive
 runaway known to date, with a two-dimensional
@@
 the core of the cluster. N-body dynamics typically ejects the least
-massive star among those interacting \cite[e.g.,][]{banerjee:12}. Just
+massive star among those interacting \cite[\newtext{although the
+  dynamical ejection fraction increases with mass because of mass
+  segregation}, e.g.,][]{banerjee:12}. Just
 based on the kinematic properties of VFTS682, we would expect several
@@
 pair-instability supernova \cite[e.g.,][]{rakavy:67} at LMC
-metallicity \citep[see also][]{langer:07}.
+metallicity \citep[see also][]{langer:07, woosley:17}.
 Similarly, the final fate of VFTS682 could be either a
 pair-instability supernova without compact remnant formation, or
-possibly direct collapse to a black hole above the pair-instability
-mass gap. The amount of mass loss of these stars will determine their final core
+possibly direct collapse to a black hole. The amount of mass loss of these stars will determine their final core
 mass and thus their final fate.
@@
 from the cluster R136 likely implies that it is only the ``tip of the
-iceberg'' of possibly extremely massive runaways in the
+iceberg'' of massive runaways in the
 region. Studies of this population, enabled by recent and future HST and \emph{Gaia} observations will put constraints on the evolution
@@
 the central cluster itself.
-\vspace*{-20pt}
+\vspace*{-40pt}
 \bibliographystyle{mnras}

### 2018-09-20 MNRASformat/referee_report_answer.txt
dynamical mergers

@@
 is to implicitly pose the question of whether these VMS are mergers or not.
-While the mergers are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave them enough VMS stars both in the cluster and ejected around it suggest that if they had allowed for the formation of even more massive stars, then the efficiency of dynamically driven mergers together with initially VMS might possibly result in an over-prediction of VMS in the region.
+While the mergers are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave them enough VMS stars both in the cluster and ejected around it suggest that if they had allowed for the formation of even more massive stars, then the efficiency of dynamically driven mergers together with stars initially more massivre than 150Msun might possibly result in an over-prediction of VMS in the region.
 We think that this is an interesting point were observations can put constraints on the physics of cluster dynamics and stellar mergers and would prefer to keep the sentence as it is.
+We have slightly rephrased the paragraph.
 

### 2018-09-20 MNRASformat/VFTS682_MNRASL.tex
dynamical mergers

@@
 
-The N-body simulations of \citet{banerjee:12} suggest that VFTS682 was ejected from R136. In their
-model, they relied on (dynamically driven) stellar mergers to explain the high masses of
-VFTS682 and the massive members of R136. They
+The N-body simulations of \citet{banerjee:12} suggest that VFTS682 was ejected from R136. They
 demonstrated that the cluster potential does not significantly change
@@
 To eject such a massive object, the cluster is
-expected to have produced a large number of massive runaways. Indeed, several %relatively
+expected to have produced a large number of massive runaways,
+\newtext{and their simulation suggest a significant incidence of
+ (dynamically driven) stellar mergers both in the cluster and among
+ the stars ejected}. Indeed, several %relatively
 isolated massive stars are observed in the region \citep[][]{evans:10,lennon:18}, some with known

### 2018-09-24 MNRASformat/referee_report_answer.txt
dynamical mergers
minor on ref. report

@@
 3 and a pointer to it in the figure caption.
+For clarity, we do not put similar bands for VFTS16 and 72 (which would overlap with the one for vfts682 in the bottom panel), however, a similar information is contained in the prolongation backwards of their Gaia proper motions in fig.
+2.
 
@@
 
-Thank you for spotting this, we have corrected the mistake
+Thank you for spotting this, we have corrected the mistake.
 
@@
 While the mergers are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave them enough VMS stars both in the cluster and ejected around it suggest that if they had allowed for the formation of even more massive stars, then the efficiency of dynamically driven mergers together with stars initially more massivre than 150Msun might possibly result in an over-prediction of VMS in the region.
-We think that this is an interesting point were observations can put constraints on the physics of cluster dynamics and stellar mergers and would prefer to keep the sentence as it is.
-We have slightly rephrased the paragraph.
+We think that this is an interesting point were observations can put constraints on the physics of cluster dynamics and stellar mergers.
+We have slightly rephrased the paragraph, hopefully now the dynamical mergers do not seem instrumental anymore.
 

### 2018-09-27 MNRASformat/referee_report_answer.txt
report selma comments and vfts682 paper alex comments

@@
 
-We thank the referee for their careful reading of our manuscript, and for the suggestions below. The changes in the text are highlighted in green bold font.
+ANSWER: =========================================================================================
+
+We thank the referee for their kind comments and suggestions. We have adapted the manuscript, and all changes are highlighted in green bold font in the pdf.
+
+=========================================================================================
 
@@
 
-For the sake of conciseness, we have added "N- body simulations indicate that the *dynamical* ejection of...", which we believe achieves the clarification suggested by the referee with less words.
+ANSWER: =========================================================================================
+
+We agree that this would be more clear, and have modified the sentence accordingly adding "N- body simulations indicate that the *dynamical* ejection of..."
+
+=================================================================================================
 
@@
 
-The green vertical band in the top panel is determined by the error bars on the apparent age of the star: the expected proper motion is the projected distance between the cluster and the star (~29pc) divided by the apparent age of the star (1+/-0.2Myr) which gives the v2D reported in Tab.
+ANSWER: =========================================================================================
+
+Indeed we agree that we should have stated this more clearly. We have added a clarification in the main text of Sec. 3 and a pointer to it in the figure caption.
+
+To answer the referee's question here: the green vertical band in the top panel is determined by the error bars on the apparent age of the star: the expected proper motion is the projected distance between the cluster and the star (~29pc) divided by the apparent age of the star (1+/-0.2Myr) which gives the v2D reported in Tab.
 2.
 For the bottom panel we have assumed a width of 45 degrees centered in theta=0.
-We have added a clarification in the main text of Sec.
-3 and a pointer to it in the figure caption.
-For clarity, we do not put similar bands for VFTS16 and 72 (which would overlap with the one for vfts682 in the bottom panel), however, a similar information is contained in the prolongation backwards of their Gaia proper motions in fig.
-2.
+
+We have chosen not to include similar bands for VFTS16 and 72 since they would overlap with the one for vfts682 in the bottom panel and we think this would make the figure less clear.
+We note that this information is also contained in Fig.
+2 (in the prolongation backward in time of the observed proper motions and the proper motion error cone).
+
+=================================================================================================
 
@@
 
-We have marked the error cones with dot-dashed lines in the updated version. We have also changed the color of the HST arrows to increase the contrast with the background figure.
+ANSWER: =========================================================================================
+
+Indeed we agree that the figure readability can be improved.
+We have marked the error cones with dot-dashed lines in the updated version.
+We have also changed the color of the HST arrows to increase the contrast with the background figure.
+
+=================================================================================================
 
@@
 
+ANSWER: =========================================================================================
+
 Thank you for spotting this, we have corrected the mistake.
+
+=================================================================================================
 
@@
 
+ANSWER: =========================================================================================
+
+We agree with the referee's comment.
 We have added a clarifying remark before the citation.
+The text now reads: "N-body dynamics typically ejects the least massive star among those interacting (although the dynamical ejection fraction increases with mass because of mass segregation, e.g., Banerjee et al.
+2012)."
+
+=================================================================================================
 
@@
 
-We agree with the referee on this point.
-Our intent in emphasizing the importance of mergers in the simulations from Banerjee et al.
-is to implicitly pose the question of whether these VMS are mergers or not.
-While the mergers are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave them enough VMS stars both in the cluster and ejected around it suggest that if they had allowed for the formation of even more massive stars, then the efficiency of dynamically driven mergers together with stars initially more massivre than 150Msun might possibly result in an over-prediction of VMS in the region.
+ANSWER: =========================================================================================
+
+We agree with the referee on this point, and have modified the text to make the dynamical mergers less upfront.
+
+We would prefer to still mention the mergers to implicitly pose the question of whether these VMS are mergers or not (and we agree with the referee that this remains to be seen for VFTS682).
+While these are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave Banerjee et al.
+enough VMS stars both in and around the cluster suggests that if they had allowed for the formation of even more massive stars, then the efficiency of dynamically driven mergers together with stars initially more massivre than 150Msun might possibly result in an over-prediction of VMS in the region.
 We think that this is an interesting point were observations can put constraints on the physics of cluster dynamics and stellar mergers.
-We have slightly rephrased the paragraph, hopefully now the dynamical mergers do not seem instrumental anymore.
+
+=================================================================================================
 
@@
 
-The reference suggested is not a primary one (the work of Belczynski et al.
-is based on the detailed simulations of Woosley 2017, http://adsabs.harvard.edu/abs/2017ApJ...836..244W), we have added a citation to the primary source instead.
+ANSWER: =========================================================================================
+
+We agree with the referee that the metallicity dependence of the pair-instability fate is also addressed in Belczynski et al.
+2016 paper, although it is a population synthesis work where the pulsational pair instability is prescribed rather than calculated, and therefore the reference suggested is not a primary one.
+The work of Belczynski et al.
+is based on the detailed simulations of Woosley 2017, http://adsabs.harvard.edu/abs/2017ApJ...836..244W, and we chose to add a citation to the primary source instead.
+
+=================================================================================================

### 2018-09-27 MNRASformat/VFTS682_MNRASL.tex
report selma comments and vfts682 paper alex comments

@@
 other stars more massive than $150\,M_\odot$ of similar spectral
-type and recent astrometric studies on VFTS16 (and possibly VFTS72)
+type and recent astrometric studies on VFTS16 and VFTS72
 provide direct evidence that the cluster can eject some of its most
@@
       \hline\hline
-      &Parameter & Units & Value & Ref.\\     
+      &Parameter & Units & Value & Ref.\\[2pt]     
        \hline
@@
                                            {-15.9}$ & (1)
-                                                    \\
+                                                    \\[2pt]
       & initial mass& $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ & (1)
-      \\
-      &age & $[\mathrm{Myr}]$ & $1.0\pm0.2$ & (1) \\
-      & mass loss rate & $\log_{10}(\dot{M}/[M_\odot \ \mathrm{yr}^{-1}])$ & $-4.1\pm0.2$ & (2)\\
+      \\[2pt]
+      &age & $[\mathrm{Myr}]$ & $1.0\pm0.2$ & (1) \\[2pt]
+      & mass loss rate & $\log_{10}(\dot{M}/[M_\odot \ \mathrm{yr}^{-1}])$ & $-4.1\pm0.2$ & (2)\\[2pt]
       \hline
@@
       \hline
-      \multicolumn{5}{l}{\emph{Absolute and relative position}} \\
+      \multicolumn{5}{l}{\emph{Absolute position and position relative
+      to R136}} \\
       \hline
@@
 
-      \multicolumn{5}{l}{\emph{Gaia relative proper motion for VFTS682 }} \\
+      \multicolumn{5}{l}{\emph{Gaia DR2 proper motion of VFTS682 relative
+      to R136}}\\
       \hline
@@
  \hline     
-      \multicolumn{5}{l}{\emph{HST relative proper motion for VFTS682}} \\
+      \multicolumn{5}{l}{\emph{HST proper motion of VFTS682 relative
+      to R136}} \\
             \hline
@@
 \citet{bestenlehner:11}  analyzed the spectra to infer the stellar
-parameters and measured an extinction of $A_V=4.45\pm0.12$, implying a
+parameters and measured a \newtext{visual} extinction of $A_V=4.45\pm0.12$, implying a
 luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
@@
 velocity (RV)
-variations suggests that the star is unlikely to have close binary
+variations suggests that the star is unlikely to have close
 companions \citep[][]{bestenlehner:11}, unless the orbital inclination is
@@
 % this star among the most massive stars known,
-% see
-as given in \Tabref{tab:star_param}. % for an overview of the parameters.
+see
+% as given in 
+\Tabref{tab:star_param}. % for an overview of the parameters.
 
@@
 about 10 arcsecond, i.e. about 2.4 pc. \cite{walborn:13} speculate
-that these nearby young stars may represent a case of star formation triggered by the the wind of VFTS682.
+that these nearby young stars may represent a case of star formation triggered by the wind of VFTS682.
 
@@
 unusual for Wolf-Rayet stars and more typical for Luminous Blue
-Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. It also
+Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. \newtext{The source} also
 shows a mid-infrared excess \citep{gruendl:09}.
@@
 They estimate a RV of
-$300\pm10\kms$ using the  N{\footnotesize V} $\lambda4944$ line, which
+$300\pm10\kms$ using the  N{\footnotesize V}\,$\lambda4944$ line, which
 is offset from the average radial velocity of the region of
@@
 %     i}ez et al. in prep}).
-\cite{bressert:12} note an offset compared between the RV of
+\cite{bressert:12} note an offset between the RV of
 the star and the nebular lines from the gas filaments in its
-vicinity. This is consistent with the expectation if the star was not
+vicinity. This is \newtext{in line} with the expectation \newtext{that} the star was not
 formed in situ. \newtext{Given these issues, we refrain} from using the RV measurements in this
@@
 define the motion of the local frame of reference using the average
-proper motion of nearby stars with reliable astrometric data.  They
+proper motion of nearby stars with reliable astrometric data \newtext{(see \Tabref{tab:vfts682})}.  They
 selected bright ($G<17$) stars within 0.05 degrees of R136 and exclude
@@
 small, no more than $0.05\masyr$ across the whole 30 Dor field. The
-effect is much smaller for stars close to the center of field, such as
+effect is much smaller for stars close to the center of the field, such as
 VFTS 682 \citep{platais:18}.  We can therefore use the relative proper
@@
 to R136 of
-VFTS682 from \emph{Gaia} DR2. $\theta$ is the angle between the projected radial direction from the
-center of R136 to each star and its projected relative proper motion
-(see also \Figref{fig:main}). Dynamical ejections from the
+VFTS682 from \emph{Gaia} DR2. % $\theta$ is the angle between the projected radial direction from the
+% center of R136 to each star and its projected relative proper motion
+% (see also \Figref{fig:main}).
+Dynamical ejections from the
 cluster should %are expected to
@@
 extension in the direction opposite to the motion, and we also show
-the most likely origin of the stars accounting for their apparent age.
+the most likely origin of the stars accounting for their apparent age
+\newtext{($0.7\pm0.1$\,Myr and $0.4^{+0.8}_{-0.4}$\,Myr for VFTS16 and VFTS72,
+respectively \citealt{schneider:18})}.
 This figure illustrates that R136 is the most likely origin of these stars, although the large error bars
@@
 VFTS72 has a small radial velocity, while VFTS16 (and possibly VFTS682) has
-a large peculiar radial velocity and therefore accurate distances along
+a large peculiar radial velocity, and therefore accurate distances along
 the line of sight are
-needed to contrain the flight direction in three dimensions.
+needed to constrain the flight direction in three dimensions.
 
@@
 VFTS682 from the cluster, although we cannot confidently rule out the
-counter hypothesis. 
+hypothesis \newtext{of in situ formation}. 
 \vspace*{-20pt}
@@
 projected velocity with respect to R136 of
-$38\pm17\kms$ (averaging the \emph{Gaia} DR2 and HST
+$38\pm17\kms$ (\newtext{taking a weighted average of} \emph{Gaia} DR2 and HST
 results). Due to the large error bars, this result will need
@@
 \citep[$\sim$3\,Myr,][]{brott:11,kohler:15, zapartas:17}, the ejection of VFTS682
-from the disruption of a binary by a supernova is excluded. %  since the region is too young for stars
+from the disruption of a \newtext{massive} binary by a supernova is excluded. %  since the region is too young for stars
 % to have experienced core-collapse already. 
@@
 suggests the presence of a close binary with total mass
-$M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. The kinematic age of VFTS682 puts an
+$M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. The
+\newtext{difference between the cluster age and the} kinematic age of VFTS682 puts an
 upper limit to the timescale to form the ``bully binary'' in
-R136. If identified, such a binary % could be R145 according to \cite{fujii:11}, and it
+R136 of \newtext{$\sim$$1.3\,\mathrm{Myr}$}. If identified, such a binary % could be R145 according to \cite{fujii:11}, and it
 might be an ideal observational candidate for a dynamically formed progenitor system of
@@
 iceberg'' of massive runaways in the
-region. Studies of this population, enabled by recent and future HST and \emph{Gaia} observations will put constraints on the evolution
+region. Studies of this population, enabled by recent and future %HST
+% and \emph{Gaia}
+observations will put constraints on the evolution
 of these extreme stars, together with the formation and evolution of

### 2018-09-28 MNRASformat/referee_report_answer.txt
added affiliation for Danny and cut fig. caption

@@
 
-We agree that this would be more clear, and have modified the sentence accordingly adding "N- body simulations indicate that the *dynamical* ejection of..."
+We agree that this would be more clear, and have modified the sentence accordingly. The text now reads: "N- body simulations indicate that the *dynamical* ejection of..."
 
@@
 
-Indeed we agree that we should have stated this more clearly. We have added a clarification in the main text of Sec. 3 and a pointer to it in the figure caption.
+Indeed, we agree that we should have stated this more clearly. We have added a clarification in the main text of Sec. 3 and a pointer to it in the figure caption.
 
@@
 
-We have chosen not to include similar bands for VFTS16 and 72 since they would overlap with the one for vfts682 in the bottom panel and we think this would make the figure less clear.
+We have chosen not to include similar bands for VFTS16 and 72 since they would overlap with the one for VFTS682 in the bottom panel and we think this would make the figure less clear.
 We note that this information is also contained in Fig.
@@
 
-We would prefer to still mention the mergers to implicitly pose the question of whether these VMS are mergers or not (and we agree with the referee that this remains to be seen for VFTS682).
-While these are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave Banerjee et al.
+We would prefer to still mention the mergers, to implicitly pose the question of whether these VMS are mergers or not (and we agree with the referee that this remains to be seen for VFTS682).
+While mergers are not instrumental in explaining their results, the fact that using an IMF up to 150Msun only already gave Banerjee et al.
 enough VMS stars both in and around the cluster suggests that if they had allowed for the formation of even more massive stars, then the efficiency of dynamically driven mergers together with stars initially more massivre than 150Msun might possibly result in an over-prediction of VMS in the region.
@@
 
-We agree with the referee that the metallicity dependence of the pair-instability fate is also addressed in Belczynski et al.
-2016 paper, although it is a population synthesis work where the pulsational pair instability is prescribed rather than calculated, and therefore the reference suggested is not a primary one.
+We agree with the referee that the metallicity dependence of the pair-instability occurrence is also addressed in Belczynski et al.
+2016, although it is a population synthesis study where the pulsational pair instability is prescribed rather than calculated, and therefore the reference suggested is not a primary one.
 The work of Belczynski et al.

### 2018-09-28 MNRASformat/VFTS682_MNRASL.tex
added affiliation for Danny and cut fig. caption

@@
     % M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
-  , S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$, I.~Platais$^{3}$,
-  R.~P.~van~der~Marel$^{3,4}$,
-  \newauthor{E.~Laplace$^{1}$, J.~M.~Bestenlehner$^{5}$, C.~J.~Evans$^{6}$,
-    V.~H\'enault-Brunet$^{7}$,  S.~Justham$^{8,9,1}$,
+  , S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2,3}$, I.~Platais$^{4}$,
+  R.~P.~van~der~Marel$^{4,5}$,
+  \newauthor{E.~Laplace$^{1}$, J.~M.~Bestenlehner$^{6}$, C.~J.~Evans$^{7}$,
+    V.~H\'enault-Brunet$^{8}$,  S.~Justham$^{9,10,1}$,
   }
 \newauthor{A.~de~Koter$^{1}$,
-      N.~Langer$^{10}$, F. Najarro$^{11}$, H.~Sana$^{12}$, F.~R.~N.~Schneider$^{13}$, J.~S.~Vink$^{14}$}}
+      N.~Langer$^{11}$, F. Najarro$^{12}$, H.~Sana$^{13}$, F.~R.~N.~Schneider$^{14}$, J.~S.~Vink$^{15}$}}
 
@@
 \end{keywords}
-
+\vspace{-30pt}
 \section{Introduction}
@@
     {\tiny The error on the RA and DEC positions, are of order
-      $\sim$$10^{-2}\masyr$ in \emph{Gaia}
-      DR2. Assuming a distance of 50\,kpc, the conversion from proper motion
-    to projected velocities is such that $1\masyr$ corresponds to $237\kms$.
+      $\sim$$0.01\masyr$ in \emph{Gaia}
+      DR2. Assuming a distance of 50\,kpc, $1\masyr$ corresponds to $237\kms$.
 $\rho(\mu_\mathrm{RA},\mu_\mathrm{DEC})$ is the
@@
   %   yr^{-1}}$. }
+  \vspace{-5pt}
   \caption{Distribution of OB-type and Wolf-Rayet stars in proper
@@
   belonging to NGC2060. }
+\vspace{-10pt}
   \label{fig:dist}
@@
 
-% We list the relative proper motion in \Tabref{tab:vfts682}. 
-\vspace*{-30pt}
+\vspace{-30pt}
 \section{The kinematics of VFTS682}
@@
 
-If confirmed, it means that isolated star formation is
+If confirmed, % it means that
+isolated star formation is
 \emph{not} required to explain the isolation of VFTS682. Its proper motion suggests that it was ejected from the cluster R136
@@
 
-\vspace*{-20pt}
+\vspace*{-25pt}
 \section*{Affiliations}
@@
     E-28691 Villanueva de la Ca\~nada, Madrid, Spain} 
- $^{3}$ {
+ $^{3}$ {Instituto de Astrof\'isica de Canarias, E-38205 La Laguna, Tenerife, Spain}
  %Center for Astrophysical Sciences, 
- Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}
-  $^{4}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+$^4$ {Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}
+  $^{5}$ {Space Telescope Science Institute, 3700 San Martin Drive,
     Baltimore, MD 21218, USA}
-  $^{5}${Department of Physics and Astronomy, Hicks Building,
+  $^{6}${Department of Physics and Astronomy, Hicks Building,
     Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
-  $^{6}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}
-  $^{7}$ {National Research Council, Herzberg Astronomy \&
+  $^{7}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}
+  $^{8}$ {National Research Council, Herzberg Astronomy \&
     Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
     Canada}
-  $^{8}$ {School of Astronomy \& Space Science, University of the Chinese
+  $^{9}$ {School of Astronomy \& Space Science, University of the Chinese
     Academy of Sciences, Beijing 100012, China}
-  $^{9}$ {National Astronomical Observatories, Chinese Academy of
+  $^{10}$ {National Astronomical Observatories, Chinese Academy of
     Sciences, Beijing 100012, China}
-  $^{10}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
+  $^{11}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
     Auf dem H\"ugel 71, 53121, Bonn, Germany}
-  $^{11}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}
-  $^{12}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}
-  $^{13}$ {Department of Physics, University of Oxford, Keble Road,
+  $^{12}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}
+  $^{13}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}
+  $^{14}$ {Department of Physics, University of Oxford, Keble Road,
     Oxford OX13RH, UK} 
-  $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}\\
+  $^{15}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}
 
@@
     and VHB the NRC-Canada Plaskett Fellowship.
-This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
+This study has used data from the ESA mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
   }

### 2018-09-28 MNRASformat/VFTS682_MNRASL.tex
added vink book ref

@@
 \end{table}
-
 The WNh5 star VFTS682, located at right ascension (RA)
@@
     yr^{-1}}$. The peak at $\theta\simeq\pi/2$ in the bottom panel is due to stars
-  belonging to NGC2060. }
-\vspace{-10pt}
+  belonging to NGC2060.}
+\vspace{-5pt}
   \label{fig:dist}
@@
   thin lines illustrate the error cone on the potential origin. The
-  yellow arrows show the \emph{Gaia} DR2 results alone, and the orange
-  arrow the HST results alone. The two blue circles indicate the regions of radii 0.01 and 0.05
+  yellow (orange) arrows show the \emph{Gaia} DR2 (HST) results alone. The two blue circles indicate the regions of radii 0.01 and 0.05
   degrees around the core of R136.}
+\vspace{-5pt}
   \label{fig:main}
@@
 
-\vspace*{-15pt}
 \subsection{ \emph{Gaia} astrometry for VFTS682\label{data:gaia}}
@@
 sky. We also show VFTS16 and VFTS72 %for completeness
-\citep[see][]{lennon:18}. The thick yellow arrows are proportional to
-the relative proper motion from \emph{Gaia} DR2, the thick green
-line illustrates the relative proper motion from HST, and the thick
-purple line shows the averaged result. The error cone
+\citep[see][]{lennon:18}. The yellow arrows are proportional to
+the relative proper motion from \emph{Gaia} DR2, the orange
+line illustrates the relative proper motion from HST, and the 
+red arrow shows the averaged result. The error cone
 on the direction of motion is illustrated by the corresponding
@@
 upper limit to the timescale to form the ``bully binary'' in
-R136 of \newtext{$\sim$$1.3\,\mathrm{Myr}$}. If identified, such a binary % could be R145 according to \cite{fujii:11}, and it
-might be an ideal observational candidate for a dynamically formed progenitor system of
+R136 of \newtext{$\sim$$1.3\,\mathrm{Myr}$}. % If identified, 
+Such a binary % could be R145 according to \cite{fujii:11}, and it
+might be a % an ideal observational
+candidate for a dynamically formed progenitor system of
 a binary black-hole, provided that stars this massive can avoid a
@@
 Similarly, the final fate of VFTS682 could be either a
-pair-instability supernova without compact remnant formation, or
-possibly direct collapse to a black hole. The amount of mass loss of these stars will determine their final core
-mass and thus their final fate.
+pair-instability supernova without compact remnant formation, or collapse to a black hole. The amount of mass loss of these stars will determine their final core
+mass and thus their final fate \newtext{\cite[e.g.,][]{vink:15}}.
 
@@
 observations will put constraints on the evolution
-of these extreme stars, together with the formation and evolution of
+of these stars, together with the formation and evolution of
 the central cluster itself.
-\vspace*{-40pt}
+\vspace*{-30pt}
 \bibliographystyle{mnras}
@@
 \vspace*{-25pt}
-\section*{Affiliations}
+\section*{Affiliations and Acknowledgements}
  \tiny
@@
     and VHB the NRC-Canada Plaskett Fellowship.
-This study has used data from the ESA mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}).
+This study has used data from the ESA mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by % the {\it Gaia} Data Processing and Analysis Consortium (
+    DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}.
   }

### 2018-10-04 MNRASformat/VFTS682_MNRASL.tex
removed hugues

@@
 \newauthor{A.~de~Koter$^{1}$,
-      N.~Langer$^{11}$, F. Najarro$^{12}$, H.~Sana$^{13}$, F.~R.~N.~Schneider$^{14}$, J.~S.~Vink$^{15}$}}
+      N.~Langer$^{11}$, F. Najarro$^{12}$, F.~R.~N.~Schneider$^{13}$, J.~S.~Vink$^{14}$}}
 
@@
   $^{12}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}
-  $^{13}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200 D, B-3001 Leuven, Belgium}
-  $^{14}$ {Department of Physics, University of Oxford, Keble Road,
+  $^{13}$ {Department of Physics, University of Oxford, Keble Road,
     Oxford OX13RH, UK} 
-  $^{15}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}
+  $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}
 

### 2018-10-31 MNRASformat/VFTS682_MNRASL.tex
stuff

@@
 
-%Selma's comments
-\definecolor{Wildstrawberry}{rgb}{1.0, 0.26, 0.64}
-%\newcommand{\SdM}[1]{{{\color{Sepia}{#1}}}}
-\newcommand{\SdM}[1]{{{\color{brown}{#1}}}}
-\newcommand{\newtext}[1]{{\color{ForestGreen}\bf{#1}}}
-
 \renewcommand{\labelitemii}{$\bullet$}
@@
 
-\title[Space astrometry of the very massive $\sim$$150\,M_\odot$ candidate runaway star VFTS682]{Space astrometry of the very massive $\sim$$150\,M_\odot$  candidate runaway star VFTS682}
+\title[Space astrometry of the very massive
+$\sim$$150\,M_\odot$ candidate runaway star VFTS682]{\emph{Gaia} and
+  HST astrometry of the very massive $\sim$$150\,M_\odot$  candidate runaway star VFTS682}
 
@@
   }
-\newauthor{A.~de~Koter$^{1}$,
-      N.~Langer$^{11}$, F. Najarro$^{12}$, F.~R.~N.~Schneider$^{13}$, J.~S.~Vink$^{14}$}}
+\newauthor{A.~de~Koter$^{1,11}$,
+      N.~Langer$^{12}$, F. Najarro$^{13}$, F.~R.~N.~Schneider$^{14}$, J.~S.~Vink$^{15}$}}
 
 % These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ \\ Affiliations
+\date{Accepted 2018 October 12. Received 2018 October 4; in original
+  form 2018 August 28 \\ Affiliations
 can be found at the end of this manuscript}
@@
 \citep[and references therein]{lada:03}. In this picture, field stars are primarily the result
-of the dissolution of dense groups.  However, a \newtext{small but} significant population
+of the dissolution of dense groups.  However, a small but significant population
 of massive stars exists in relative isolation, far from dense clusters
@@
 the star formed in situ as an isolated massive star, or it was ejected from  R136. N-body simulations 
-indicate that the \newtext{dynamical} ejection of very massive stars like VFTS682 is
+indicate that the dynamical ejection of very massive stars like VFTS682 is
 expected \citep[e.g.][]{fujii:11, banerjee:12}. The capability of a
@@
 % masses, is the most massive runaway star known to date. 
-\vspace*{-20pt}
+%\vspace*{-20pt}
 \section{Observations}
@@
     \caption{Stellar parameters of VFTS682.}
-    \vspace*{-15pt}
+    % \vspace*{-15pt}
     \begin{tabular}{llc|c|c}
@@
     \caption{Kinematics of VFTS682.}
-    \vspace*{-9pt}
+    % \vspace*{-9pt}
     \begin{tabular}{llc|c|c}
@@
 \citet{bestenlehner:11}  analyzed the spectra to infer the stellar
-parameters and measured a \newtext{visual} extinction of $A_V=4.45\pm0.12$, implying a
+parameters and measured a visual extinction of $A_V=4.45\pm0.12$, implying a
 luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of
@@
 unusual for Wolf-Rayet stars and more typical for Luminous Blue
-Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. \newtext{The source} also
+Variable (LBV) stars \citep{udalski:08, bestenlehner:11}. The source also
 shows a mid-infrared excess \citep{gruendl:09}.
@@
 stars.  We therefore caution
-against over-interpreting the \newtext{existing} radial velocities estimates. \citet{bestenlehner:11} estimate a mass loss rate of
+against over-interpreting the existing radial velocities estimates. \citet{bestenlehner:11} estimate a mass loss rate of
 $\log_{10}(\dot{M}/[M_\odot \ \mathrm{yr}^{-1}])=-4.1\pm0.2$, not accounting for the
@@
 the star and the nebular lines from the gas filaments in its
-vicinity. This is \newtext{in line} with the expectation \newtext{that} the star was not
-formed in situ. \newtext{Given these issues, we refrain} from using the RV measurements in this
+vicinity. This is in line with the expectation that the star was not
+formed in situ. Given these issues, we refrain from using the RV measurements in this
 work, and focus on the velocities on the plane of the sky.
@@
   \includegraphics[width=0.35\textwidth]{figures/angle}
-  % \caption{\emph{Top panel}: distribution of OB-type and Wolf-Rayet stars in proper
-  %   motion relative to R136 from \emph{Gaia} DR2. VFTS682 is not an outlier, but
-  %   its relative proper motion matches the expected value (green
-  %   vertical band) if it were indeed
-  %   ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. %The top axis shows the conversion to $\kms$
-  %   % assuming a distance of 50\,kpc.
-  %   \emph{Bottom panel}:  distribution of
-  %   angles between the relative proper motion direction and the radial
-  %   direction from the center of the cluster to the star. The error bars from \emph{Gaia} for VFTS682 are large, but
-  %   the best value is in agreement with the hypothesis of dynamical
-  %   ejection (green vertical band). The peak at $\theta\simeq\pi/2$ is due to stars
-  %   belonging to NGC2060. In both
-  %   panels, the dark blue histograms contain 317 
-  %   stars with error bars smaller than $0.1\,\mathrm{mas \
-  %     yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
-  %   lighter blue histograms contain 36 stars with error bars smaller than $0.05\,\mathrm{mas \
-  %   yr^{-1}}$. }
   \vspace{-5pt}
@@
     proper motion matches the value expected for an early dynamical
-    ejection \newtext{(see \Secref{sec:results})}. In both
+    ejection (see \Secref{sec:results}). In both
     panels, the dark blue histograms contain 317 
@@
 
-% \begin{figure*}%[tbp]
-%   \centering
-%   \includegraphics[width=\textwidth]{./figures/fig2}  
-%   % \caption{The thick yellow arrows indicate proper motions relative to
-%   %   R136 (blue
-%   %   circle) from \emph{Gaia} DR2, multiplied by 0.4\,Myr. The extensions in the opposite
-%   %   direction are proportional to the apparent ages of the stars, and
-%   %   the thin lines illustrate the error cone on the potential
-%   %   origin. The green thick arrow
-%   %   indicates the HST proper motion of VFTS682, with the relative
-%   %   error cone and extensions proportional to its apparent
-%   %   age. The two blue circles indicate the regions of radii 0.01 and 0.05
-%   %   degrees around the core of R136.}
-%   \caption{The thick purple arrow shows proper motion
-%     relative to R136 for VFTS682 averaging the \emph{Gaia} DR2 and HST
-%   astrometry, multiplied by 0.4\,Myr. The extension in the opposite
-%   direction is proportional to the apparent age of the star, and the
-%   thin lines illustrate the error cone on the potential origin. The
-%   yellow arrows show the \emph{Gaia} DR2 results alone, and the green
-%   arrow the HST results alone. The two blue circles indicate the regions of radii 0.01 and 0.05
-%    degrees around the core of R136.}  
-%   \label{fig:main}
-% \end{figure*}
-
-\begin{figure}%[tbp]
+\begin{figure*}
   \centering
-  \includegraphics[width=0.5\textwidth]{./figures/fig2}  
+  \includegraphics[width=\textwidth]{./figures/fig2}  
   \caption{The thick red arrow shows the proper motion
@@
   \label{fig:main}
-\end{figure}
+\end{figure*}
 
@@
 define the motion of the local frame of reference using the average
-proper motion of nearby stars with reliable astrometric data \newtext{(see \Tabref{tab:vfts682})}.  They
+proper motion of nearby stars with reliable astrometric data (see \Tabref{tab:vfts682}).  They
 selected bright ($G<17$) stars within 0.05 degrees of R136 and exclude
@@
 with the ID source 330375 in their catalog. 
-Therefore, their measurements provide \newtext{a} useful
-complementary estimate of the proper motion of VFTS682 \newtext{which is} independent from the \emph{Gaia} data.
+Therefore, their measurements provide a useful
+complementary estimate of the proper motion of VFTS682 which is independent from the \emph{Gaia} data.
 
@@
 
-\vspace{-30pt}
+%\vspace{-30pt}
 \section{The kinematics of VFTS682}
@@
 The black points in~\Figref{fig:dist} show the proper motion 
-(\newtext{top} panel) and the projected flight direction (\newtext{bottom} panel) relative
+(top panel) and the projected flight direction (bottom panel) relative
 to R136 of
@@
 cluster should %are expected to
-produce close to radial ejections, i.e.~$\theta\simeq0$. \newtext{The green
+produce close to radial ejections, i.e.~$\theta\simeq0$. The green
 vertical bands highlight the expectations for these two quantities. %  if VFTS682 were ejected from R136 early in its
@@
 by the error bars on the star's apparent age, and we assume a width of
-$45$ degrees in the bottom panel.} For comparison, we also show the relative
+$45$ degrees in the bottom panel. For comparison, we also show the relative
 proper motion of VFTS16 and VFTS72 (gray points), and the distribution
@@
 the most likely origin of the stars accounting for their apparent age
-\newtext{($0.7\pm0.1$\,Myr and $0.4^{+0.8}_{-0.4}$\,Myr for VFTS16 and VFTS72,
-respectively \citealt{schneider:18})}.
+($0.7\pm0.1$\,Myr and $0.4^{+0.8}_{-0.4}$\,Myr for VFTS16 and VFTS72,
+respectively \citealt{schneider:18}).
 This figure illustrates that R136 is the most likely origin of these stars, although the large error bars
@@
 VFTS682 from the cluster, although we cannot confidently rule out the
-hypothesis \newtext{of in situ formation}. 
+hypothesis of in situ formation. 
 \vspace*{-20pt}
@@
 Based on our results, we consider that VFTS682 is
-\newtext{potentially} the most massive
+potentially the most massive
 runaway known to date, with a two-dimensional
 projected velocity with respect to R136 of
-$38\pm17\kms$ (\newtext{taking a weighted average of} \emph{Gaia} DR2 and HST
+$38\pm17\kms$ (taking a weighted average of \emph{Gaia} DR2 and HST
 results). Due to the large error bars, this result will need
@@
 \citep[$\sim$3\,Myr,][]{brott:11,kohler:15, zapartas:17}, the ejection of VFTS682
-from the disruption of a \newtext{massive} binary by a supernova is excluded. %  since the region is too young for stars
+from the disruption of a massive binary by a supernova is excluded. %  since the region is too young for stars
 % to have experienced core-collapse already. 
@@
 the core of the cluster. N-body dynamics typically ejects the least
-massive star among those interacting \cite[\newtext{although the
+massive star among those interacting \cite[although the
   dynamical ejection fraction increases with mass because of mass
-  segregation}, e.g.,][]{banerjee:12}. Just
+  segregation, e.g.,][]{banerjee:12}. Just
 based on the kinematic properties of VFTS682, we would expect several
@@
 expected to have produced a large number of massive runaways,
-\newtext{and their simulation suggest a significant incidence of
+and their simulation suggest a significant incidence of
  (dynamically driven) stellar mergers both in the cluster and among
- the stars ejected}. Indeed, several %relatively
+ the stars ejected. Indeed, several %relatively
 isolated massive stars are observed in the region \citep[][]{evans:10,lennon:18}, some with known
@@
 $M_1+M_2\gtrsim 300\,M_\odot$ in the core of the cluster. The
-\newtext{difference between the cluster age and the} kinematic age of VFTS682 puts an
+difference between the cluster age and the kinematic age of VFTS682 puts an
 upper limit to the timescale to form the ``bully binary'' in
-R136 of \newtext{$\sim$$1.3\,\mathrm{Myr}$}. % If identified, 
+R136 of $\sim$$1.3\,\mathrm{Myr}$. % If identified, 
 Such a binary % could be R145 according to \cite{fujii:11}, and it
@@
 pair-instability supernova without compact remnant formation, or collapse to a black hole. The amount of mass loss of these stars will determine their final core
-mass and thus their final fate \newtext{\cite[e.g.,][]{vink:15}}.
+mass and thus their final fate \cite[e.g.,][]{vink:15}.
 
@@
 of these stars, together with the formation and evolution of
-the central cluster itself.
-\vspace*{-30pt}
-\bibliographystyle{mnras}
-\bibliography{bibliography/vfts682}
-
-% Don't change these lines
-%\bsp	% typesetting comment
-\label{lastpage}
-
-\vspace*{-25pt}
-\section*{Affiliations and Acknowledgements}
- \tiny
-\noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
-    Amsterdam, 1098 XH Amsterdam, The Netherlands} 
-  $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
-    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} 
- $^{3}$ {Instituto de Astrof\'isica de Canarias, E-38205 La Laguna, Tenerife, Spain}
- %Center for Astrophysical Sciences, 
-$^4$ {Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}
-  $^{5}$ {Space Telescope Science Institute, 3700 San Martin Drive,
-    Baltimore, MD 21218, USA}
-  $^{6}${Department of Physics and Astronomy, Hicks Building,
-    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
-  $^{7}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}
-  $^{8}$ {National Research Council, Herzberg Astronomy \&
-    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
-    Canada}
-  $^{9}$ {School of Astronomy \& Space Science, University of the Chinese
-    Academy of Sciences, Beijing 100012, China}
-  $^{10}$ {National Astronomical Observatories, Chinese Academy of
-    Sciences, Beijing 100012, China}
-  $^{11}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
-    Auf dem H\"ugel 71, 53121, Bonn, Germany}
-  $^{12}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}
-  $^{13}$ {Department of Physics, University of Oxford, Keble Road,
-    Oxford OX13RH, UK} 
-  $^{14}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}
+the central cluster itself.\\
+% \vspace*{-30pt}
 
   % \begin{acknowledgements}
-  {   \tiny
-    We thank P.~Crowther, J.~Heyl, M.\,C.\,Ramirez-Tannus, S.~N.~Shore,
+  { \small
+    We thank P.~Crowther, J.~Heyl, M.\,C.\,Ram\'irez-Tannus, S.~N.~Shore,
     and S.\,Torres, and the HSTPROMO Collaboration
@@
 
+% \bibliographystyle{mnras}
+% \bibliography{bibliography/vfts682}
+\begin{thebibliography}{}
+\makeatletter
+\relax
+\def\mn@urlcharsother{\let\do\@makeother \do\$\do\&\do\#\do\^\do\_\do\%\do\~}
+\def\mn@doi{\begingroup\mn@urlcharsother \@ifnextchar [ {\mn@doi@}
+  {\mn@doi@[]}}
+\def\mn@doi@[#1]#2{\def\@tempa{#1}\ifx\@tempa\@empty \href
+  {http://dx.doi.org/#2} {doi:#2}\else \href {http://dx.doi.org/#2} {#1}\fi
+  \endgroup}
+\def\mn@eprint#1#2{\mn@eprint@#1:#2::\@nil}
+\def\mn@eprint@arXiv#1{\href {http://arxiv.org/abs/#1} {{\tt arXiv:#1}}}
+\def\mn@eprint@dblp#1{\href {http://dblp.uni-trier.de/rec/bibtex/#1.xml}
+  {dblp:#1}}
+\def\mn@eprint@#1:#2:#3:#4\@nil{\def\@tempa {#1}\def\@tempb {#2}\def\@tempc
+  {#3}\ifx \@tempc \@empty \let \@tempc \@tempb \let \@tempb \@tempa \fi \ifx
+  \@tempb \@empty \def\@tempb {arXiv}\fi \@ifundefined
+  {mn@eprint@\@tempb}{\@tempb:\@tempc}{\expandafter \expandafter \csname
+  mn@eprint@\@tempb\endcsname \expandafter{\@tempc}}}
+
+\bibitem[\protect\citeauthoryear{{Banerjee}, {Kroupa}  \& {Oh}}{{Banerjee}
+  et~al.}{2012}]{banerjee:12}
+{Banerjee} S.,  {Kroupa} P.,   {Oh} S.,  2012, \mn@doi [\apj]
+  {10.1088/0004-637X/746/1/15}, \href
+  {http://adsabs.harvard.edu/abs/2012ApJ...746...15B} {746, 15}
+
+\bibitem[\protect\citeauthoryear{{Bate}}{{Bate}}{2009}]{bate:09}
+{Bate} M.~R.,  2009, \mn@doi [\mnras] {10.1111/j.1365-2966.2008.14106.x}, \href
+  {http://adsabs.harvard.edu/abs/2009MNRAS.392..590B} {392, 590}
+
+\bibitem[\protect\citeauthoryear{{Bestenlehner} et~al.,}{{Bestenlehner}
+  et~al.}{2011}]{bestenlehner:11}
+{Bestenlehner} J.~M.,  et~al., 2011, \mn@doi [\aap]
+  {10.1051/0004-6361/201117043}, \href
+  {http://adsabs.harvard.edu/abs/2011A%26A...530L..14B} {530, L14}
+
+\bibitem[\protect\citeauthoryear{{Blaauw}}{{Blaauw}}{1961}]{blaauw:61}
+{Blaauw} A.,  1961, \bain, \href
+  {http://adsabs.harvard.edu/abs/1961BAN....15..265B} {15, 265}
+
+\bibitem[\protect\citeauthoryear{{Bressert} et~al.,}{{Bressert}
+  et~al.}{2012}]{bressert:12}
+{Bressert} E.,  et~al., 2012, \mn@doi [\aap] {10.1051/0004-6361/201117247},
+  \href {http://adsabs.harvard.edu/abs/2012A%26A...542A..49B} {542, A49}
+
+\bibitem[\protect\citeauthoryear{{Brott} et~al.,}{{Brott}
+  et~al.}{2011}]{brott:11}
+{Brott} I.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201016113}, \href
+  {http://adsabs.harvard.edu/abs/2011A%26A...530A.115B} {530, A115}
+
+\bibitem[\protect\citeauthoryear{{Cioni} et~al.,}{{Cioni}
+  et~al.}{2011}]{cioni:11}
+{Cioni} M.-R.~L.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201016137},
+  \href {http://adsabs.harvard.edu/abs/2011A%26A...527A.116C} {527, A116}
+
+\bibitem[\protect\citeauthoryear{{Crowther}, {Schnurr}, {Hirschi}, {Yusof},
+  {Parker}, {Goodwin}  \& {Kassim}}{{Crowther} et~al.}{2010}]{crowther:10}
+{Crowther} P.~A.,  {Schnurr} O.,  {Hirschi} R.,  {Yusof} N.,  {Parker} R.~J.,
+  {Goodwin} S.~P.,   {Kassim} H.~A.,  2010, \mn@doi [\mnras]
+  {10.1111/j.1365-2966.2010.17167.x}, \href
+  {http://adsabs.harvard.edu/abs/2010MNRAS.408..731C} {408, 731}
+
+\bibitem[\protect\citeauthoryear{{Crowther} et~al.,}{{Crowther}
+  et~al.}{2016}]{crowther:16}
+{Crowther} P.~A.,  et~al., 2016, \mn@doi [\mnras] {10.1093/mnras/stw273}, \href
+  {http://adsabs.harvard.edu/abs/2016MNRAS.458..624C} {458, 624}
+
+\bibitem[\protect\citeauthoryear{{Drew}, {Herrero}, {Mohr-Smith},
+  {Mongui{\'o}}, {Wright}, {Kupfer}  \& {Napiwotzki}}{{Drew}
+  et~al.}{2018}]{drew:18}
+{Drew} J.~E.,  {Herrero} A.,  {Mohr-Smith} M.,  {Mongui{\'o}} M.,  {Wright}
+  N.~J.,  {Kupfer} T.,   {Napiwotzki} R.,  2018, \mn@doi [\mnras]
+  {10.1093/mnras/sty1905}, \href
+  {http://adsabs.harvard.edu/abs/2018MNRAS.tmp.1812D} {}
+
+\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
+  et~al.}{2010}]{evans:10}
+{Evans} C.~J.,  et~al., 2010, \mn@doi [\apjl] {10.1088/2041-8205/715/2/L74},
+  \href {http://adsabs.harvard.edu/abs/2010ApJ...715L..74E} {715, L74}
+
+\bibitem[\protect\citeauthoryear{{Evans} et~al.,}{{Evans}
+  et~al.}{2011}]{evans:11}
+{Evans} C.~J.,  et~al., 2011, \mn@doi [\aap] {10.1051/0004-6361/201116782},
+  \href {http://adsabs.harvard.edu/abs/2011A%26A...530A.108E} {530, A108}
+
+\bibitem[\protect\citeauthoryear{{Fujii} \& {Portegies Zwart}}{{Fujii} \&
+  {Portegies Zwart}}{2011}]{fujii:11}
+{Fujii} M.~S.,  {Portegies Zwart} S.,  2011, \mn@doi [Science]
+  {10.1126/science.1211927}, \href
+  {http://adsabs.harvard.edu/abs/2011Sci...334.1380F} {334, 1380}
+
+\bibitem[\protect\citeauthoryear{{Gaia Collaboration} et~al.,}{{Gaia
+  Collaboration} et~al.}{2016}]{gaia:16}
+{Gaia Collaboration} et~al., 2016, \mn@doi [\aap]
+  {10.1051/0004-6361/201629272}, \href
+  {http://adsabs.harvard.edu/abs/2016A%26A...595A...1G} {595, A1}
+
+\bibitem[\protect\citeauthoryear{{Gaia Collaboration} et~al.,}{{Gaia
+  Collaboration} et~al.}{2018}]{brown:18}
+{Gaia Collaboration} et~al., 2018, \mn@doi [\aap]
+  {10.1051/0004-6361/201833051}, \href
+  {http://adsabs.harvard.edu/abs/2018A%26A...616A...1G} {616, A1}
+
+\bibitem[\protect\citeauthoryear{{Gruendl} \& {Chu}}{{Gruendl} \&
+  {Chu}}{2009}]{gruendl:09}
+{Gruendl} R.~A.,  {Chu} Y.-H.,  2009, \mn@doi [\apjs]
+  {10.1088/0067-0049/184/1/172}, \href
+  {http://adsabs.harvard.edu/abs/2009ApJS..184..172G} {184, 172}
+
+\bibitem[\protect\citeauthoryear{{Gvaramadze}, {Weidner}, {Kroupa}  \&
+  {Pflamm-Altenburg}}{{Gvaramadze} et~al.}{2012}]{gvaramadze:12}
+{Gvaramadze} V.~V.,  {Weidner} C.,  {Kroupa} P.,   {Pflamm-Altenburg} J.,
+  2012, \mn@doi [\mnras] {10.1111/j.1365-2966.2012.21452.x}, \href
+  {http://adsabs.harvard.edu/abs/2012MNRAS.424.3037G} {424, 3037}
+
+\bibitem[\protect\citeauthoryear{{H{\'e}nault-Brunet}
+  et~al.,}{{H{\'e}nault-Brunet} et~al.}{2012}]{henault-brunet:12}
+{H{\'e}nault-Brunet} V.,  et~al., 2012, \mn@doi [\aap]
+  {10.1051/0004-6361/201219472}, \href
+  {http://adsabs.harvard.edu/abs/2012A%26A...545L...1H} {545, L1}
+
+\bibitem[\protect\citeauthoryear{{K{\"o}hler} et~al.,}{{K{\"o}hler}
+  et~al.}{2015}]{kohler:15}
+{K{\"o}hler} K.,  et~al., 2015, \mn@doi [\aap] {10.1051/0004-6361/201424356},
+  \href {http://adsabs.harvard.edu/abs/2015A%26A...573A..71K} {573, A71}
+
+\bibitem[\protect\citeauthoryear{{Kuiper}, {Yorke}  \& {Turner}}{{Kuiper}
+  et~al.}{2015}]{kuiper:15}
+{Kuiper} R.,  {Yorke} H.~W.,   {Turner} N.~J.,  2015, \mn@doi [\apj]
+  {10.1088/0004-637X/800/2/86}, \href
+  {http://adsabs.harvard.edu/abs/2015ApJ...800...86K} {800, 86}
+
+\bibitem[\protect\citeauthoryear{{Lada} \& {Lada}}{{Lada} \&
+  {Lada}}{2003}]{lada:03}
+{Lada} C.~J.,  {Lada} E.~A.,  2003, \mn@doi [\araa]
+  {10.1146/annurev.astro.41.011802.094844}, \href
+  {http://adsabs.harvard.edu/abs/2003ARA%26A..41...57L} {41, 57}
+
+\bibitem[\protect\citeauthoryear{{Lamb}, {Oey}, {Segura-Cox}, {Graus},
+  {Kiminki}, {Golden-Marx}  \& {Parker}}{{Lamb} et~al.}{2016}]{lamb:16}
+{Lamb} J.~B.,  {Oey} M.~S.,  {Segura-Cox} D.~M.,  {Graus} A.~S.,  {Kiminki}
+  D.~C.,  {Golden-Marx} J.~B.,   {Parker} J.~W.,  2016, \mn@doi [\apj]
+  {10.3847/0004-637X/817/2/113}, \href
+  {http://adsabs.harvard.edu/abs/2016ApJ...817..113L} {817, 113}
+
+\bibitem[\protect\citeauthoryear{{Langer}, {Norman}, {de Koter}, {Vink},
+  {Cantiello}  \& {Yoon}}{{Langer} et~al.}{2007}]{langer:07}
+{Langer} N.,  {Norman} C.~A.,  {de Koter} A.,  {Vink} J.~S.,  {Cantiello} M.,
+  {Yoon} S.-C.,  2007, \mn@doi [\aap] {10.1051/0004-6361:20078482}, \href
+  {http://adsabs.harvard.edu/abs/2007A%26A...475L..19L} {475, L19}
+
+\bibitem[\protect\citeauthoryear{{Lennon} et~al.,}{{Lennon}
+  et~al.}{2018}]{lennon:18}
+{Lennon} D.~J.,  et~al., 2018, ArXiv:1805.08277, \href
+  {http://adsabs.harvard.edu/abs/2018arXiv180508277L} {}
+
+\bibitem[\protect\citeauthoryear{{Luks} \& {Rohlfs}}{{Luks} \&
+  {Rohlfs}}{1992}]{luks:92}
+{Luks} T.,  {Rohlfs} K.,  1992, \aap, \href
+  {http://adsabs.harvard.edu/abs/1992A%26A...263...41L} {263, 41}
+
+\bibitem[\protect\citeauthoryear{{Meixner} et~al.,}{{Meixner}
+  et~al.}{2006}]{meixner:06}
+{Meixner} M.,  et~al., 2006, \mn@doi [\aj] {10.1086/508185}, \href
+  {http://adsabs.harvard.edu/abs/2006AJ....132.2268M} {132, 2268}
+
+\bibitem[\protect\citeauthoryear{{Oh} \& {Kroupa}}{{Oh} \&
+  {Kroupa}}{2016}]{oh:16}
+{Oh} S.,  {Kroupa} P.,  2016, \mn@doi [\aap] {10.1051/0004-6361/201628233},
+  \href {http://adsabs.harvard.edu/abs/2016A%26A...590A.107O} {590, A107}
+
+\bibitem[\protect\citeauthoryear{{Parker} \& {Goodwin}}{{Parker} \&
+  {Goodwin}}{2007}]{parker:07}
+{Parker} R.~J.,  {Goodwin} S.~P.,  2007, \mn@doi [\mnras]
+  {10.1111/j.1365-2966.2007.12179.x}, \href
+  {http://adsabs.harvard.edu/abs/2007MNRAS.380.1271P} {380, 1271}
+
+\bibitem[\protect\citeauthoryear{{Pietrzy{\'n}ski} et~al.,}{{Pietrzy{\'n}ski}
+  et~al.}{2013}]{pietrzynski:13}
+{Pietrzy{\'n}ski} G.,  et~al., 2013, \mn@doi [\nat] {10.1038/nature11878},
+  \href {http://adsabs.harvard.edu/abs/2013Natur.495...76P} {495, 76}
+
+\bibitem[\protect\citeauthoryear{{Platais}, {van der Marel}, {Lennon},
+  {Anderson}, {Bellini}, {Sabbi}, {Sana}  \& {Bedin}}{{Platais}
+  et~al.}{2015}]{platais:15}
+{Platais} I.,  {van der Marel} R.~P.,  {Lennon} D.~J.,  {Anderson} J.,
+  {Bellini} A.,  {Sabbi} E.,  {Sana} H.,   {Bedin} L.~R.,  2015, \mn@doi [\aj]
+  {10.1088/0004-6256/150/3/89}, \href
+  {http://adsabs.harvard.edu/abs/2015AJ....150...89P} {150, 89}
+
+\bibitem[\protect\citeauthoryear{{Platais} et~al.,}{{Platais}
+  et~al.}{2018}]{platais:18}
+{Platais} I.,  et~al., 2018, \mn@doi [\aj] {10.3847/1538-3881/aad280}, \href
+  {http://adsabs.harvard.edu/abs/2018AJ....156...98P} {156, 98}
+
+\bibitem[\protect\citeauthoryear{{Poveda}, {Ruiz}  \& {Allen}}{{Poveda}
+  et~al.}{1967}]{poveda:67}
+{Poveda} A.,  {Ruiz} J.,   {Allen} C.,  1967, Boletin de los Observatorios
+  Tonantzintla y Tacubaya, \href
+  {http://adsabs.harvard.edu/abs/1967BOTT....4...86P} {4, 86}
+
+\bibitem[\protect\citeauthoryear{{Rakavy} \& {Shaviv}}{{Rakavy} \&
+  {Shaviv}}{1967}]{rakavy:67}
+{Rakavy} G.,  {Shaviv} G.,  1967, \mn@doi [\apj] {10.1086/149204}, \href
+  {http://adsabs.harvard.edu/abs/1967ApJ...148..803R} {148, 803}
+
+\bibitem[\protect\citeauthoryear{{Renzo} et~al.,}{{Renzo}
+  et~al.}{2018}]{renzo:18}
+{Renzo} M.,  et~al., 2018, ArXiv:1804.09164, \href
+  {http://adsabs.harvard.edu/abs/2018arXiv180409164R} {}
+
+\bibitem[\protect\citeauthoryear{{Rosen}, {Krumholz}, {McKee}  \&
+  {Klein}}{{Rosen} et~al.}{2016}]{rosen:16}
+{Rosen} A.~L.,  {Krumholz} M.~R.,  {McKee} C.~F.,   {Klein} R.~I.,  2016,
+  \mn@doi [\mnras] {10.1093/mnras/stw2153}, \href
+  {http://adsabs.harvard.edu/abs/2016MNRAS.463.2553R} {463, 2553}
+
+\bibitem[\protect\citeauthoryear{{Rubio-D{\'{\i}}ez}, {Najarro},
+  {Garc{\'{\i}}a}  \& {Sundqvist}}{{Rubio-D{\'{\i}}ez}
+  et~al.}{2017}]{rubio-diez:17}
+{Rubio-D{\'{\i}}ez} M.~M.,  {Najarro} F.,  {Garc{\'{\i}}a} M.,   {Sundqvist}
+  J.~O.,  2017, in {Eldridge} J.~J.,  {Bray} J.~C.,  {McClelland} L.~A.~S.,
+  {Xiao} L.,  eds,  IAU Symposium Vol. 329, The Lives and Death-Throes of
+  Massive Stars. pp 131--135, \mn@doi{10.1017/S1743921317002447}
+
+\bibitem[\protect\citeauthoryear{{Sabbi} et~al.,}{{Sabbi}
+  et~al.}{2012}]{sabbi:12}
+{Sabbi} E.,  et~al., 2012, \mn@doi [\apjl] {10.1088/2041-8205/754/2/L37}, \href
+  {http://adsabs.harvard.edu/abs/2012ApJ...754L..37S} {754, L37}
+
+\bibitem[\protect\citeauthoryear{{Schneider}, {Langer}, {de Koter}, {Brott},
+  {Izzard}  \& {Lau}}{{Schneider} et~al.}{2014}]{schneider:14}
+{Schneider} F.~R.~N.,  {Langer} N.,  {de Koter} A.,  {Brott} I.,  {Izzard}
+  R.~G.,   {Lau} H.~H.~B.,  2014, \mn@doi [\aap] {10.1051/0004-6361/201424286},
+  \href {http://adsabs.harvard.edu/abs/2014A%26A...570A..66S} {570, A66}
+
+\bibitem[\protect\citeauthoryear{{Schneider}, {Castro}, {Fossati}, {Langer}  \&
+  {de Koter}}{{Schneider} et~al.}{2017}]{schneider:17}
+{Schneider} F.~R.~N.,  {Castro} N.,  {Fossati} L.,  {Langer} N.,   {de Koter}
+  A.,  2017, \mn@doi [\aap] {10.1051/0004-6361/201628409}, \href
+  {http://adsabs.harvard.edu/abs/2017A%26A...598A..60S} {598, A60}
+
+\bibitem[\protect\citeauthoryear{{Schneider} et~al.,}{{Schneider}
+  et~al.}{2018}]{schneider:18}
+{Schneider} F.~R.~N.,  et~al., 2018, \mn@doi [Science]
+  {10.1126/science.aan0106}, \href
+  {http://adsabs.harvard.edu/abs/2018Sci...359...69S} {359, 69}
+
+\bibitem[\protect\citeauthoryear{{Townsley}, {Broos}, {Feigelson}, {Garmire}
+  \& {Getman}}{{Townsley} et~al.}{2006}]{townsley:06}
+{Townsley} L.~K.,  {Broos} P.~S.,  {Feigelson} E.~D.,  {Garmire} G.~P.,
+  {Getman} K.~V.,  2006, \mn@doi [\aj] {10.1086/500535}, \href
+  {http://adsabs.harvard.edu/abs/2006AJ....131.2164T} {131, 2164}
+
+\bibitem[\protect\citeauthoryear{{Townsley}, {Broos}, {Garmire}, {Bouwman},
+  {Povich}, {Feigelson}, {Getman}  \& {Kuhn}}{{Townsley}
+  et~al.}{2014}]{townsley:14}
+{Townsley} L.~K.,  {Broos} P.~S.,  {Garmire} G.~P.,  {Bouwman} J.,  {Povich}
+  M.~S.,  {Feigelson} E.~D.,  {Getman} K.~V.,   {Kuhn} M.~A.,  2014, \mn@doi
+  [\apjs] {10.1088/0067-0049/213/1/1}, \href
+  {http://adsabs.harvard.edu/abs/2014ApJS..213....1T} {213, 1}
+
+\bibitem[\protect\citeauthoryear{{Udalski} et~al.,}{{Udalski}
+  et~al.}{2008}]{udalski:08}
+{Udalski} A.,  et~al., 2008, \actaa, \href
+  {http://adsabs.harvard.edu/abs/2008AcA....58..329U} {58, 329}
+
+\bibitem[\protect\citeauthoryear{{Vink}}{{Vink}}{2015}]{vink:15}
+{Vink} J.~S.,  2015, {Mass-Loss Rates of Very Massive Stars}.
+Springer
+
+\bibitem[\protect\citeauthoryear{{Walborn}, {Barb{\'a}}  \&
+  {Sewi{\l}o}}{{Walborn} et~al.}{2013}]{walborn:13}
+{Walborn} N.~R.,  {Barb{\'a}} R.~H.,   {Sewi{\l}o} M.~M.,  2013, \mn@doi [\aj]
+  {10.1088/0004-6256/145/4/98}, \href
+  {http://adsabs.harvard.edu/abs/2013AJ....145...98W} {145, 98}
+
+\bibitem[\protect\citeauthoryear{{Ward} \& {Kruijssen}}{{Ward} \&
+  {Kruijssen}}{2018}]{ward:18}
+{Ward} J.~L.,  {Kruijssen} J.~M.~D.,  2018, \mn@doi [\mnras]
+  {10.1093/mnras/sty117}, \href
+  {http://adsabs.harvard.edu/abs/2018MNRAS.475.5659W} {475, 5659}
+
+\bibitem[\protect\citeauthoryear{{Woosley}}{{Woosley}}{2017}]{woosley:17}
+{Woosley} S.~E.,  2017, \mn@doi [\apj] {10.3847/1538-4357/836/2/244}, \href
+  {http://adsabs.harvard.edu/abs/2017ApJ...836..244W} {836, 244}
+
+\bibitem[\protect\citeauthoryear{{Zapartas} et~al.,}{{Zapartas}
+  et~al.}{2017}]{zapartas:17}
+{Zapartas} E.,  et~al., 2017, \mn@doi [\aap] {10.1051/0004-6361/201629685},
+  \href {http://adsabs.harvard.edu/abs/2017A%26A...601A..29Z} {601, A29}
+
+\bibitem[\protect\citeauthoryear{{Zinnecker} \& {Yorke}}{{Zinnecker} \&
+  {Yorke}}{2007}]{zinnecker:07}
+{Zinnecker} H.,  {Yorke} H.~W.,  2007, \mn@doi [\araa]
+  {10.1146/annurev.astro.44.051905.092549}, \href
+  {http://adsabs.harvard.edu/abs/2007ARA%26A..45..481Z} {45, 481}
+
+\bibitem[\protect\citeauthoryear{{de Koter}, {Heap}  \& {Hubeny}}{{de Koter}
+  et~al.}{1997}]{dekoter:97}
+{de Koter} A.,  {Heap} S.~R.,   {Hubeny} I.,  1997, \mn@doi [\apj]
+  {10.1086/303736}, \href {http://adsabs.harvard.edu/abs/1997ApJ...477..792D}
+  {477, 792}
+
+\makeatother
+\end{thebibliography}
+
+% Don't change these lines
+%\bsp	% typesetting comment
+\label{lastpage}
+
+% \vspace*{-25pt}
+\section*{Affiliations}
+\noindent $^{1}${Astronomical Institute Anton Pannekoek, University of
+  Amsterdam, 1098 XH Amsterdam, The Netherlands}
+
+\noindent $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
+    E-28691 Villanueva de la Ca\~nada, Madrid, Spain}
+
+  \noindent $^{3}$ {Instituto de Astrof\'isica de Canarias, E-38205 La Laguna, Tenerife, Spain}
+ %Center for Astrophysical Sciences,
+
+  \noindent $^4$ {Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}
+
+  \noindent $^{5}$ {Space Telescope Science Institute, 3700 San Martin Drive,
+    Baltimore, MD 21218, USA}
+
+  \noindent $^{6}${Department of Physics and Astronomy, Hicks Building,
+    Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}
+
+  \noindent $^{7}${UK Astronomy Technology Centre, Royal Observatory Edinburgh, Blackford Hill, Edinburgh, EH9 3HJ, UK}
+
+  \noindent $^{8}$ {National Research Council, Herzberg Astronomy \&
+    Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7,
+    Canada}
+
+  \noindent $^{9}$ {School of Astronomy \& Space Science, University of the Chinese
+    Academy of Sciences, Beijing 100012, China}
+
+  \noindent $^{10}$ {National Astronomical Observatories, Chinese Academy of
+    Sciences, Beijing 100012, China}
+
+  \noindent $^{11}$ {Institute of Astronomy, KU Leuven, Celestijnenlaan 200D, 3001 Leuven, Belgium}
+
+  \noindent $^{12}$ {Argelander-Instit\"ut f\"ur Astronomie, Universit\"at Bonn,
+    Auf dem H\"ugel 71, 53121, Bonn, Germany}
+
+  \noindent $^{13}$ {Centro de Astrobiolog\'ia, CSIC-INTA, Carretera de Torrej\'on a Ajalvir km-4, E-28850 Torrej\'on de Ardoz, Madrid, Spain}
+
+  \noindent $^{14}$ {Department of Physics, University of Oxford, Keble Road,
+    Oxford OX13RH, UK}
+
+  \noindent $^{15}$ {Armagh Observatory, College Hill, Armagh BT619DG, UK}
+
 \end{document}
